### 2015-05-26 tools/doc/wip.tex
Initial commit of spline and innerproduct codes, with Faddeeva package.

\documentclass{revtex4}
\usepackage{amsmath}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage[usenames]{color}
\usepackage{ulem}
\newcommand{\x}{{\mathrm{x}}}

\begin{document}
\title{Method for direct integration of complex functions with smooth amplitude and phase.}
\begin{abstract}
In gravitational wave data analysis studies we encounter integrals of complex functions for which we have smooth and often slowly varying representations for each phase and amplitude.  Such integrals can be computed in segments with polynomial representations of the phase and amplitude for each segment.  Expressions for each integral can be derived by a combination of analytic expressions, recursion relations, and rapidly convering power series.
\end{abstract}
\author{John Baker}
\maketitle

In gravitational wave data analysis studies, we frequently must integrate waveform inner products.  In the Fourier domain, the signals may have some very fine features related to the long duration of the signals while spreading across a broad band.  If uniformly sampled in frequency, as with an FFT a very large number of samples (up to $10^7$ or more for LISA) would be needed to accurately represent the signals, and to carry out colored noise signal inner products. It is now recognized that for many signals, with more optimal spacing of frequency samples, far fewer frequency evaluations are needed for accurate spline representations.  We present a way to compute signal inner products with comparably few frequency samples by analytically integrating the spline-represented inner product.

For example, we compute the signal-to-noise ratio (SNR) for a signal $\tilde s(f)$ by
\begin{eqnarray}
\rho^2&=&2\int_0^\infty\frac{\tilde s(f)\tilde s(f)^*}{S_n(f)}df\\
      &=&\sum_{k=0}^{N-1}\int_{f_k}^{f_{k+1}} A(f)e^{i p(f)}df.
\end{eqnarray}
Consistent with a spline approximation, in  each of the component integrals we can approximate $A(f)$ and $p(f)$ locally as polynomials (cubic in this particular case).  Then given $p(f)$ the  task is to compute integrals of the general form:
\begin{equation}
I_{n,m}[p](\epsilon)=\int_{0}^{\epsilon} x^nE_m[p](x)dx
\end{equation}
where we have defined
\begin{equation}
E_m[p](x)=e^{i\sum\limits_{k=1}^{m}p_kx^k}.
\end{equation}
The task is made managable by recursion relations in either $n$ or $m$.  For $n\geq m$, integration by parts yields, suppressing unnecessary $p$ and $\epsilon$ arguments:
\begin{equation}
I_{n,m}=\frac{n-m+1}{i m p_m}\left[I_{n-m,0}E_m-I_{n-m,m}\right]-\sum_{k=1}^{m-1}\frac{k}{m}\frac{p_k}{p_m}I_{n-m+k,m}
\end{equation}
which for fixed $m$ provides a convenient way to compute all $n$-order integrals given the first few.  And for the case $0<n=m-1$,
\begin{equation}
I_{n,m}=\frac{E_{m-1}}{i m p_m}-\sum_{k=1}^{m-1}\frac{k}{m}\frac{p_k}{p_m}I_{k-1,m}
\end{equation}
For those first few with $n<m-1$ (or if $p_m$ is computationally small)  we need another way to compute the integral.  We an compute cases up to $m<=2$ with one additional closed form solution,
\begin{equation}
I_{0,2}=-\frac{e^{\frac{i\pi}4}}2\sqrt{\frac{\pi}{p_2}}\left(E_2w\left(e^{\frac{i\pi}4}\left(\epsilon \sqrt{p_2}+z_0\right)\right)-w\left(e^{\frac{i\pi}4}\left(z_0\right)\right)\right),
\end{equation}
where we have written the result in terms of the Faddeeva function $w(z)=e^{-z^2}\left(1-{\mathrm{erf}}(-iz)\right)$.  We implement this using an open source numerical implementation\cite{FaddeevaPackage}, based on the algorthims in \cite{Poppe1990,Zaghloul2011}.

For cases with small $p_m$ or $m>2$, expanding $e^{ip_m x^m}$ yields a rapidly converging series, allowing an rapidly converging infinite series type recursion relation,
\begin{equation}
I_{n,m}=\sum_{k=0}^{\infty}\frac{(i p_m)^k}{k!}I_{n+km,m-1}
\end{equation}
expresssing the integral in terms of lower-$m$ integrals.
\end{document}

### 2015-05-26 tools/spline.c

@@ -1,0 +1,268 @@
+//Cubic spline with derivatives
+//Written by John G Baker at gsfc.nasa.gov
+#include <stdio.h>
+#include <stdlib.h>
+#define false 0
+#define true 1
+
+int spline_natural=false;
+int spline_iold=0;
+int spline_findix(double xp, const double xs[],int n);
+int verbose=0;
+
+//Solves tridiagonal for the matched second derivatives z of the 3rd order interpoling polynomials at the sample points (knots).
+//The condition that 1st derivatives are continuous at the knots becomes:
+//  dx[i-1] z[i-1] + 2(dx[i]+dx[i-1])z[i] + dx[i]z[i+1] = r[i]
+//with:
+//  dx[i] = x[i+1]-x[i]
+// r[i>0] = 6(dy[i]/D[i]-dy[i-1]/D[i])
+//  dy[i] = y[i+1]-y[i]
+//
+//Solution part 1:
+//  For 0<i<n-1 defines a tridiagonal matrix  with diagonals b[i],upperdiags c[i] and lower diags a[i]
+//  with numbering starting from i=1 in each case
+//  All rows are identical but first and last which depend on BCs
+//
+//  The iterative solution algorithm by LU decomposition is:
+//  (with shorthand: bb = b[i] - a[i-1]*v[i-1]) 
+//  v[1] = c[1]/b[1] 
+//  s[1] = r[1]/b[1] 
+//  v[i] = c[i]/bb
+//  s[i] = (r[i] - a[i-1]*s[i-1])/bb
+//
+//  In the bulk: 1<i<n-2 the abc values come directly from the eqn at top:
+//  a[i]=c[i]=x[i+1]-x[i]  
+//  b[i]=2*(x[i+1]-x[i-1])
+//  
+//  The boundary conditions for the first and last row come from some condition which
+//  eliminates z[0] and z[n-1] in the i=1 and i=n-2 versions of the continuity condition at top
+//  This yields the values for b[1],c[1],b[n-2],a[n-3] in the first and last rows.
+//  
[…]

### 2015-05-26 tools/spline.h

@@ -1,0 +1,12 @@
+//Cubic spline with derivatives
+//Written by John G Baker at gsfc.nasa.gov
+extern int spline_verbose;
+extern int spline_natural;
+extern int spline_iold;
+
+void spline_construct(const double xs[],const double ys[],double zs[],int n);
+void spline_intd3(double xp, const double xs[],const double ys[],const double zs[],int n,double *q, double *dq1, double *dq2, double *dq3);
+void spline_intd3_vec(const double xp[],int nxp, const double xs[],const double ys[],const double zs[],int n,double q[], double dq1[], double dq2[], double dq3[]);
+double spline_int(double xp, const double xs[],const double ys[],const double zs[],int n);
+
+        

### 2015-05-26 tools/wip.c

@@ -1,0 +1,269 @@
+//Written by John G Baker at gsfc.nasa.gov
+//Weighted Inner-Product
+#include "spline.h"
+#include "Faddeeva.h"
+#include <complex.h>
+#include <math.h>
+#include <stdlib.h>
+#include <stdio.h>
+
+//# This function realizes integration of S-weighted inner product
+//# of the difference between two complex functions represented by
+//# slowly evolving phase and amplitude.
+//# 
+//# We write the integral as:
+//#
+//# I = \int{df A(f)exp(i\psi(f))} = \int{df |s_1(f)* s_2(f)|^2/S(f)}
+//#
+//# For comparison we have a slow classic approach applying standard quadrature
+//# encoded in the function wip_sum.  We also try a novel approach encoded
+//# in wip_phase.
+
+
+
+//# Our novel approach involves locally approximating 
+//# A(f) and \psi(f) as quadratic polynomials
+//# We can then evaluate the integral explicitly over short local intervals
+
+int wip_relative=1;
+double wip_mindf=1e-8;
+double wip_errtol=1.0e-8;
+int wip_uselogf=0;
+int wip_count;
+
+// To avoid loss of precision at small z, cexpm1i(zi) implements cos(zi)-1 + i*sin(zi)
+// The real analogue of this function is part of the standard library, and the C99 standard
+// reserves the name, but it is not generally implemented
+inline double complex cexpm1i(double zi){
+  double rr;//rr=cos(x)-1=2*sin(x/2)**2
+  double ri=sin(zi);
+  double sinhalf=sin(zi/2.0);
+  rr=-2.0*sinhalf*sinhalf;
+  return rr+I*ri;
+};
+
+//interpolation integrand on a new f-grid
[…]

### 2015-05-26 tools/wip.h

@@ -1,0 +1,9 @@
+// WIP.H
+//Written by John G Baker at gsfc.nasa.gov
+//Weighted Inner-Product
+#ifndef WIP_H
+#define WIP_H
+
+double complex wip_phase (double *f1, int n1, double *f2, int n2, double *s1Ar, double *s1Ai, double  *s1p, double *s2Ar, double*s2Ai, double *s2p, double (*Snoise)(double), double scalefactor);
+
+#endif

### 2015-05-26 tools/wiptest.c

@@ -1,0 +1,210 @@
+//test of wip 
+//Written by John G Baker at gsfc.nasa.gov
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <complex.h>
+#include <time.h>
+#include "wip.h"
+
+double testfn_I_p(double x, double c2,double c3,double x0){
+  double xo=x-x0;
+  return xo*xo*(c2+xo*c3);
+};
+
+double testfn_I_A(double x, double sig,double xc){
+  return erf((x-xc)/sig);
+};
+
+double complex testfn_dI_A(double x,double c2,double c3,double sig,double x0,double xc){
+  const double twoortpi=2.0/sqrt(M_PI);
+  double xo=x-x0;
+  double dp=xo*(2.0*c2+3*xo*c3);
+  double xnorm=(x-xc)/sig;
+  //double expfac=exp(-xnorm*xnorm);
+  //double complex wfac=I*dp*Faddeeva_w(I*xnorm,1.0e-11)
+  //double complex dI=I*dp + ( twoortpi/sig - wfac ) * expfac 
+  double complex dI=twoortpi/sig*exp(-xnorm*xnorm) + I*dp*erf(xnorm);
+  return dI;
+};
+
+double Swhitenoise(double x){
+  return 1;
+};
+
+extern int wip_count;
+
+int main (){
+  //Define the test function
+  double c2,c3,x0,xc,a0,sig;
+  c2=-1.41;
+  //c3=+0.0199;
+  c3=+0.0203;
+  c2*=5;
+  c3*=5;
+  xc=3.47;
+  x0=-20;
+  a0=10.0;
+  sig=1.51e2;
+
+  //Defin the f grid
+  double fmax, fmin;
+  fmax=12.0;
+  fmin=1e-4;
+  int N0=20,N=N0;
+
+  //wip_uselogf=1
+
+  //#numerical test function check
+  if(1){
+    srand(time(NULL));
+    double x=rand()/((double) RAND_MAX);
+    double dx=0.01;
+    int twotolevel=1;
+    double eps=dx;
+    int level;
+    for(level=0;level<10;level++){	
+      eps=dx/twotolevel;
+      double complex Ival = a0*cexp(I*testfn_I_p(x, c2, c3, x0))*testfn_I_A(x, sig, xc);
[…]

### 2015-12-08 docs/paper1.tex
Added time-domain yAB and X,Y,Z. Uploaded first draft of paper.
LISAgeometry: added time-domain yAB functions and time-domain X,Y,Z TDI observables.
docs: added first draft of paper and reference list.

\documentclass[aps,showpacs,twocolumn,prd,superscriptaddress,nofootinbib]{revtex4}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{color}
\usepackage{enumerate}
\usepackage{ulem}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand\ud{{\mathrm{d}}}
\newcommand\uD{{\mathrm{D}}}
\newcommand\calO{{\mathcal{O}}}
\newcommand\bfx{\mathbf{x}}
\newcommand{\ov}[1]{\overline{#1}}
\newcommand{\ph}[1]{\phantom{#1}}
\newcommand{\cte}{\mathrm{cte}}
\newcommand{\nn}{\nonumber}
\newcommand{\hatk}{\hat{k}}
\newcommand{\Hz}{\,\mathrm{Hz}}
\newcommand{\sinc}{\,\mathrm{sinc}}
\newcommand{\Msol}{M_{\odot}}

\begin{document}

\title{Accelerated prospective parameter estimation of non-spinning compact binaries with higher modes for LIGO/VIRGO and LISA-type detectors}

\author{John G. Baker}
\affiliation{Gravitational Astrophysics Laboratory, NASA Goddard Space Flight Center, 8800 Greenbelt Rd., Greenbelt, MD 20771, USA}
\author{Alessandra Buonanno}
\affiliation{Department of Physics, University of Maryland, College Park, MD 20742, USA}
\affiliation{Max Planck Institute for Gravitational Physics (Albert Einstein Institute), Am M\"uhlenberg 1, Potsdam-Golm, 14476, Germany}
\author{Philip. B. Graff}
\affiliation{Department of Physics, University of Maryland, College Park, MD 20742, USA}
\affiliation{Gravitational Astrophysics Laboratory, NASA Goddard Space Flight Center, 8800 Greenbelt Rd., Greenbelt, MD 20771, USA}
\author{Sylvain Marsat}
\affiliation{Department of Physics, University of Maryland, College Park, MD 20742, USA}
\affiliation{Gravitational Astrophysics Laboratory, NASA Goddard Space Flight Center, 8800 Greenbelt Rd., Greenbelt, MD 20771, USA}
\affiliation{Max Planck Institute for Gravitational Physics (Albert Einstein Institute), Am M\"uhlenberg 1, Potsdam-Golm, 14476, Germany}

\date{\today}

\begin{abstract}

[Abstract]

\end{abstract}

\pacs{
04.25.D-, % numerical relativity
04.70.Bw, % classical black holes
04.80.Nn, % Gravitational wave detectors and experiments
95.30.Sf, % relativity and gravitation
95.55.Ym, % Gravitational radiation detectors
97.60.Lf  % black holes (astrophysics)
}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}

[Introduction]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Methodology}
\label{sec:methodology}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Reduced order models for fast waveform generation}
\label{subsec:rom}

[review recent advances in this field: surrogates, ROQ, P\"urrer's SVD method in some more details]

[present the mode-by-mode time and phase offset tracking, specific to HM]

[specific choices: logarithmic sampling, 1D spline interpolation, smoothing spline for projection coefficients]

[give summary plot of unfaithfulness]

[compare speedup with ordinary EOB]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Fourier-domain response of LISA-type detectors}
\label{subsec:fdresponse}

We use the notations of the Kr\'olak \& al paper. We denote the annual frequency of the motion by $f_{0}=1/\mathrm{year}$, and $\Omega = 2\pi f_{0}$. We set $c=1$ except for numerical applications.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection*{Definitions}

The primary observables for a LISA-type instrument are given by (taking $y_{21}$ as an example)
%
\be
	y_{21}(t) = \left( 1-\hatk \cdot n_{3}\right) \left[ \Psi_{3}(t+\hatk\cdot p_{2} - L) - \Psi_{3}(t+\hatk\cdot p_{1}) \right] \,,
\ee
%
with the definitions
%
\be
	\Psi_{A} = \frac{\Phi_{A}}{1-(\hatk\cdot n_{A})^{2}} \,, \quad \Phi_{A} = \frac{1}{2} n_{A}^{i} H_{ij} n_{A}^{j} \,.
\ee
%
Here we follow the approximation made in~\cite{Krolak+04}: all vectors related to the positions, $n_{A}$ and $p_{A}$, will be considered evaluated at the time $t$, neglecting their variation on the time scale of light propagation in the system, which is motivated by $\Omega R/c \ll 1$, $\Omega L/c\ll 1$. It seems also that this approximation is used anyway in the first place when deriving the above form for the frequency variation of the lasers [to check]. We will also idealize the geometry of the constellation by keeping the linear order in eccentricity and neglecting higher-order orbital perturbations. Another convenient definition is
%
\be
	\Phi_{A} = P_{A}^{+} h_{+} + P_{A}^{\times}h_{\times} \,,
\ee
%
where $h_{+}$, $h_{\times}$ are the two polarizations of the wave and the $P_{A}$'s are combinations of trigonometric functions depending on $\Omega t$.

\subsection*{Fourier transform of a delayed signal}

Let us consider, in general, the problem of determining the Fourier transform of a signal $h$ (whose Fourier transform $\tilde{h}$ is known), when applying a varying delay $d(t)$ and also multiplying by some function $F(t)$. Defining
%
\be
	h_{d}(t) = h(t+d(t)) \,, \quad s(t) = F(t)h_{d}(t) \,,
\ee
%
we have
%
\be
	h_{d}(t) = \int \ud f \, e^{-2i\pi f (t+d(t))}\tilde{h}(f) \,,
\ee
%
and
%
\begin{align}
	\tilde{s}(f) &= \int \ud t \, e^{2i\pi f t} F(t)  \int \ud f' \, e^{-2i\pi f' (t+d(t))}\tilde{h}(f') \nn\\
	&= \int \ud f' \, \tilde{h}(f-f') \int \ud t \, e^{2i\pi f' t} e^{-2i\pi (f-f') d(t)} F(t) \,.
\end{align}
%
We can rewrite the last equation as a kind of convolution with a frequency-dependent Kernel, according to
%
\begin{align}\label{eq:freqkernel}
	\tilde{s}(f) &= \int \ud f' \, \tilde{h}(f-f') G(f,f') \,, \nn\\
	G(f,f') &= \int \ud t \, e^{2i\pi f' t} e^{-2i\pi (f-f') d(t)} F(t) \,.
\end{align}
%
Since in our case both $d(t)$ and $F(t)$ are $2\pi/\Omega$-periodic, the latter Kernel will reduce to a discrete comb in frequency. If we define $g[f](t) = e^{-2i\pi f d(t)} F(t)$, a Fourier series decomposition gives
%
\begin{align}
	g[f](t) &= \sum\limits_{n\in \mathbb{Z}} c_{n}[g](f) e^{-i n\Omega t} \,, \nn\\
	c_{n}[g](f) &= \frac{\Omega}{2\pi}\int_{0}^{\frac{2\pi}{\Omega}} \ud t \, g[f](t) e^{i n \Omega t} \,,
\end{align}
%
which transforms~\eqref{eq:freqkernel} into
%
\begin{align}
	G(f,f') &= \sum\limits_{n\in \mathbb{Z}} c_{n}[g](f-f') \delta\left(f' - n f_{0}\right) \,, \nn\\
	\tilde{s}(f) &= \sum\limits_{n\in \mathbb{Z}} c_{n}[g]\left(f - n f_{0}\right) \tilde{h}\left(f - n f_{0}\right) \,.
\end{align}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection*{Application to $y_{AB}$ observables}

We now apply the above results to $y_{AB}$ observables, with the example of $y_{21}$. We will assume our starting point are the Fourier transforms of $h_{+}$ and $h_{\times}$. We get for either polarization
%
\begin{widetext}
\begin{align}
	g_{21}^{+,\times}[f](t) &= \frac{P_{3}^{+,\times}}{1+\hatk\cdot n_{3}} \left[ e^{-2i\pi f (\hatk\cdot p_{2} - L)} - e^{-2i\pi f \hatk\cdot p_{1}} \right] \,, \nn\\
	&= i\pi L f P_{3}^{+,\times} \sinc\left[ \pi L f(1+\hatk \cdot n_{3})\right]\exp\left[ i \pi f L \left(  1 - \hatk \cdot \frac{p_{1}+p_{2}}{L}\right) \right] \,,
\end{align}
\end{widetext}
%
where we used $p_{2}-p_{1} = -L n_{3}$ and the dependence in $t$ is in the vectors $n_{A}$, $p_{A}$. $(p_{2}+p_{1})/2 $ also has a simple expression in the LISA frame. Thus the complete result before any further approximation is
%
\begin{widetext}
\begin{subequations}
\begin{align}
	\tilde{y}_{21}^{+,\times}(f) &= \sum\limits_{n\in \mathbb{Z}} c_{n}[g_{21}^{+,\times}]\left(f - n f_{0}\right) \tilde{h}_{+,\times}\left(f - n f_{0}\right)\,, \\
	c_{n}[g_{21}^{+,\times}]\left(f\right) &= \frac{\Omega}{2\pi}\int_{0}^{\frac{2\pi}{\Omega}} \ud t \, e^{i n \Omega t} i\pi L f P_{3}^{+,\times} \sinc\left[ \pi L f(1+\hatk \cdot n_{3})\right]\exp\left[ i \pi f L \left(  1 - \hatk \cdot \frac{p_{1}+p_{2}}{L}\right) \right]  \,.
\end{align}
\end{subequations}
\end{widetext}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection*{Separating the orbital delay from the constellation delay}

Some numerical tests for the coefficients $c_{n}$ on the frequency band $10^{-5}$Hz-$1$Hz show that the extent in $n$ of the frequency comb rises a lot (up to several hundreds) at the high frequency end. This seems attributable to the presence of $\hatk\cdot r/c$ in the argument of te exponential and the large number of oscillations this causes. Indeed, this contribution is roughly $f R/c$ with $c/R\sim 2.10^{-3}\Hz$, whereas $c/L\sim 2.10^{-1}\Hz$, so that the oscillations due to $L/c$ terms are relatively mild, while those due to the $R/c$ term are much more rapid and yield extended harmonic content.

Luckily, it is possible to factor out this term, and to control it analytically. It corresponds to the part of the delay that corresponds to the GW propagating from the solar system barycenter ($\Omega$) to the moving center of the constellation ($O$), and has a simple expression involving only the sine and cosine of $\Omega t$. If we define
%
\be
	h_{+,\times}^{O} = h_{+,\times} (t + \hatk \cdot r)
\ee
%
as the signal as measured at the center of the constellation $O$, since $\hatk \cdot r = R \cos \beta \cos(\Omega t - \lambda)$ (choosing $\eta_{0} = 0$), applying the same calculation as above yields for this simplified retardation
%
\begin{align}
	\tilde{h}_{+,\times}^{O}\left(f\right) &= \sum\limits_{n\in \mathbb{Z}} c_{n}[g_{O}^{+,\times}]\left(f - n f_{0}\right) \tilde{h}_{+,\times}\left(f - n f_{0}\right)\,, \\
	c_{n}[g_{O}^{+,\times}]\left(f\right) &= \frac{\Omega}{2\pi}\int_{0}^{\frac{2\pi}{\Omega}} \ud t \, e^{i n \Omega t} \exp\left[ -2i \pi f R \cos \beta \cos (\Omega t - \lambda) \right] \nn\\ 
	&= e^{i n \lambda} \frac{1}{2\pi}\int_{0}^{2\pi} \ud u \, e^{i n u} \exp\left[ -2i \pi f R \cos \beta \cos u \right] \nn\\
	&= i^{n}e^{i n \lambda} J_{n}\left[ -2\pi f R \cos \beta \right] \,.
\end{align}
%
Thus we obtain, in two steps,
%
\begin{widetext}
\begin{subequations}
\begin{align}
	\tilde{h}_{+,\times}^{O}\left(f\right) &= \sum\limits_{n\in \mathbb{Z}} i^{n}e^{i n \lambda} J_{n}\left[ -2\pi \left(f - n f_{0}\right) R \cos \beta \right]\tilde{h}_{+,\times}\left(f - n f_{0}\right)\,, \\
	\tilde{y}_{21}^{+,\times}(f) &= \sum\limits_{n\in \mathbb{Z}} c_{n}[\gamma_{21}^{+,\times}]\left(f - n f_{0}\right) \tilde{h}_{+,\times}^{O}\left(f - n f_{0}\right)\,, \\
	c_{n}[\gamma_{21}^{+,\times}]\left(f\right) &= \frac{\Omega}{2\pi}\int_{0}^{\frac{2\pi}{\Omega}} \ud t \, e^{i n \Omega t} 2 i\pi L f f_{3}^{+,\times}\sinc\left[ \pi L f (1+\hatk \cdot n_{3})\right]\exp\left[ i \pi f L\left(1 - \hatk \cdot O_{2} \cdot \frac{p_{1}^{L}+p_{2}^{L}}{L} \right) \right] \,.
\end{align}
\end{subequations}
\end{widetext}
%
Now we find that the Bessel comb that relates $\tilde{h}^{O}$ to $\tilde{h}$ (which we can control analytically) contains all the spread in $n$, as the new $c_{n}$ coefficients (expressed with integrals) do not extend beyond $n=\pm 10$. These comb coefficients show a smooth variation with frequency (they vary much less than the Fourier-domain signal): in the worst case, one should be able to compute them numerically at a handful of frequencies and interpolate inbetween.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection*{Locally linear phase approximation (LLP)}

If we put actual numbers, we have $\Omega/2\pi \simeq 3.1\times10^{-8}\mathrm{Hz} \ll f$ over all the frequency band we consider. Considering that the comb is of limited extend in $n$, we might want to Taylor expand the relevant quantities accross the comb. Since the phase is rapidly varying, and since we have to consider a constant and a linear term in the phase as arbitrary, it is natural to Taylor expand the phase at the first order (that is, approximate the phase linearly locally), neglect the $n\Omega/2\pi$ in $(f-n\Omega/2\pi)$ appearing in the argument of the amplitude andother terms, and assess the quality of this approximation.

Preliminary tests show that this approximation should be valid, unsurprisingly, for all the mildly varying functions encountered: $\sinc$, $\exp$ and the linear term in the $c_{n}$'s, and the Bessel functions $J_{n}$. However, $\tilde{h}$ has a Fourier-domain phase which becomes very rapidly varying at low frequencies, in the deep inspiral, $\Psi\propto f^{-5/3}$; here it is not obvious that we can neglect the second derivative of the phase.

For a Newtonian inspiral, applying the SPA gives $\tilde{h}(f) = A_{N}(f)e^{-i\Psi_{N}(f)}$ with
%
\begin{subequations}
\begin{align}
	A_{N}(f) &= -\sqrt{\frac{5\pi}{24}} \frac{G^{2}m^{2}}{Dc^{5}} \nu^{1/2}v^{-7/2}\,,\\
	\Psi_{N}(f) &= 2\pi f t_{0} - \phi_{0} + \frac{3}{128\nu v^{5}} \,, 
\end{align}
\end{subequations}
%
where $v=(G m \pi f/c^{3})^{1/3}$. The magnitude of the first terms in a Taylor expansion in $f_{0}$ is (taking $10^{-5}\mathrm{Hz}$ as a reference starting frequency):
%
\begin{widetext}
\begin{subequations}
\begin{align}
	|\delta A/A| &= \left| \frac{1}{A_{N}}\frac{\ud A_{N}}{\ud f} f_0 \right| \simeq 3\times10^{-3} \left(\frac{f}{10^{-5}\Hz}\right)^{-1} \,,\\
	|\delta^{1}\Psi | &= \left| f_0\frac{\ud \Psi_{N}}{\ud f} \right| \simeq 10^{3}\left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-8/3} \,, \\
	|\delta^{2}\Psi | &= \left| \frac{1}{2} f_0^{2}\frac{\ud^{2} \Psi_{N}}{\ud f^{2}} \right| \simeq 5 \left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-11/3} \,, \\
	|\delta^{3}\Psi | &= \left| \frac{1}{6} f_0^{3}\frac{\ud^{3} \Psi_{N}}{\ud f^{3}} \right| \simeq 2\times 10^{-2} \left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-14/3} \,.
\end{align}
\end{subequations}
\end{widetext}
%
The approximation degrades for lower starting frequencies, and, at a fixed frequency, when the total mass decreases and/or the mass ratio increases. However, if we consider only merging binaries, that is to say binaries of which the merger is seen during the observation period of the mission, we have a lower limit for the starting frequency which can be larger than the detector's lowest frequency in band.

Namely, if we take a Newtonian estimate for the relation between the frequency and the time to coalescence, we have
%
\be
	\pi f(t) = \left[ \frac{256\nu}{5c^{5}} (Gm)^{5/3} \Delta t \right]^{-3/8} \,,
\ee
%
with $\Delta t$ the time to coalescence, and the starting frequency is then
%
\be
	f_{\rm start} = 10^{-5}\mathrm{Hz} \left( \frac{m}{8.5\times10^{6}\Msol} \right)^{-5/8} \left( \frac{\nu}{1/4} \right)^{-3/8} \left( \frac{\Delta t}{5 \mathrm{yr}} \right)^{-3/8} \,.
\ee
%
So we see that the ``worst'' total mass is $8.5\times10^6\Msol (\nu/(1/4))^{3/5}$, for 5 years of observation and $f_{\rm low} = 10^{-5}\mathrm{Hz}$ for the entry in band, which are conservative values; for higher masses, part of the signal is cut because it is out-of-band, while for lower masses, the signal is cut because of the limited time of observation before merger, and we have the scaling
%
\be
	|\delta^{2}\Psi |_{f_\mathrm{start}} \propto m^{5/8}\nu^{3/8}\Delta t^{11/8} \,,
\ee
%
which shows that the effect of the truncation of the signal for a finite observation time wins: the approximation becomes better for lower mass and higher mass ratios systems. The worse point in the parameter space should be, for these Newtonian estimates, at equal mass and $m=8.5\times 10^{6} \Msol$, with $|\delta^{2}\Psi |_{\mathrm{worst}} \simeq 0.15$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection*{Direct summation}

The Newtonian estimates above indicate that the locally linear phase approximation is going to be at least marginally valid for all sources of which we observe the inspiral, merger and ringdown. We adopt for the results above the generic notation
%
\begin{align}
	\tilde{s}(f) &=  \sum\limits_{n\in \mathbb{Z}} c_{n}[G]\left(f - n f_{0}\right) \tilde{h}\left(f - n f_{0}\right)\,, \nn\\
	c_{n}[G](f) &= \frac{1}{T} \int_{0}^{T}\ud t\, e^{in\Omega t} G[f](t)
\end{align}
%
with $T=2\pi/\Omega=1\mathrm{yr}$ and $G[f](t)$ a function of frequency and time.

At the lowest order in the approximation, with $|\delta A/A| \ll 1$ and $|\delta^{2} \Psi | \ll 1$, and also neglecting $f_{0}$ in the frequency-dependence of $G$, we have $\tilde{h}(f-n f_{0}) \simeq \tilde{h}(f) \exp\left[ i n f_{0} \ud \Psi/\ud f \right]$ and (after commuting the sum and integral):
%
\be
	\tilde{s}(f) \simeq \tilde{h}(f) \frac{1}{T} \int_{0}^{T}\ud t\, G[f](t) \sum\limits_{n} e^{i n \Omega t} \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \,.
\ee
%
Using the Dirac comb relation
%
\be
	\sum\limits_{n} e^{i n \Omega (t-t_{0})} = \frac{2\pi}{\Omega} \sum\limits_{n} \delta\left( t-t_{0} - \frac{n\Omega}{2\pi} \right) \,,
\ee
%
we have
%
\be
	\sum\limits_{n} e^{i n \Omega t} \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f }\right] = T \sum\limits_{n} \delta\left( t + \frac{1}{2\pi}\frac{\ud \Psi}{\ud f} - n T \right) \,,
\ee
%
and since only one of the Dirac deltas has support on $[0,T]$, we obtain simply
%
\be
	\tilde{s}(f) \simeq \tilde{h}(f) \times G[f]\left( -\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right)\,.
\ee
%
Now, recall that when the SPA applies (which it does not in the merger-ringdown phase of the signal), we have the relations
%
\begin{align}
	\Psi_{\rm SPA} &= \phi(t_{f}) - 2\pi f t_{f} + \mathrm{const} \,,\nn\\
	\frac{\ud \Psi_{\rm SPA}}{\ud f} &= -2\pi t_{f} \,,
\end{align}
%
thus providing a simple physical interpretation: at leading order, we can simply use the time-domain modulation and delays, as evaluated at a fixed time $t_{f}$, the time at which the corresponding frequency is dominantly emitted. Notice however that the above result is more general, as it extends to the merger-ringdown phase, and that the limits of validity of the two approximations differ: the SPA becomes increasingly valid in the deep inspiral, whereas the LLP breaks down in the limit of almost monochromatic sources but becomes increasingly valid at the end of the inspiral, when the signal evolves quickly compared to the orbital timescale.

Notice that these Newtonian estimates are very poor to cover the merger and ringdown; for instance, above the dominant ringdown frequency the Fourier-domain amplitude drops rapidly in a way that is not captured by the power law above.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection*{Summation of $\tilde{h}_{O}\left(f\right)$}

The expression obtained above for $\tilde{h}_{O}\left(f\right)$ can be summed, either by applying the above general formulas with the function $G[f](t) = \exp\left[ -2i \pi f R \cos \beta \cos (\Omega t - \lambda) \right]$, either directly by means of the Jacobi-Anger expansion:
%
\begin{widetext}
\begin{align}
	\tilde{h}_{O}\left(f\right) &\simeq \sum\limits_{n\in \mathbb{Z}} i^{n}e^{i n \lambda} J_{n}\left[ -2\pi f R \cos \beta \right] A(f) \exp\left[ -i \Psi \left(f-\frac{n\Omega}{2\pi}\right) \right] \nn \\
	&\simeq A(f) e^{-i \Psi \left(f\right)} \sum\limits_{n\in \mathbb{Z}} i^{n}e^{i n \lambda} J_{n}\left[ -2\pi f R \cos \beta \right] \exp\left[ i \frac{n\Omega}{2\pi} \frac{\ud \Psi}{\ud f} + \calO(\Omega^{2})  \right]
\end{align}
\end{widetext}
%
We can then use the Jacobi-Anger expansion (with the notation $z\equiv 2\pi R f \cos\beta$)
%
\be
	e^{i z \cos \theta}= \sum\limits_{n\in \mathbb{Z}} i^{n}J_{n}[z]e^{i n \theta} \,,
\ee
%
to obtain a simple phase correction
%
\begin{subequations}
\begin{align}
	\tilde{h}_{O}\left(f\right) &\simeq \tilde{h}(f) \exp\left[ -iz \cos\left( \lambda + \frac{\Omega}{2\pi} \frac{\ud \Psi}{\ud f} \right) \right]\,.
\end{align}
\end{subequations}
%
The corrections beyond the leading order can be derived directly by a kind of Feynman trick, introducing derivatives with respect to the parameter $\lambda$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Accelerated overlap computations for modeled waveforms}
\label{subsec:overlaps}

[present recursion relations - Fresnel integrals]

[discuss numerical robustness, power-law cut in the space of phase coefficients]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Summary \& performance}
\label{subsec:performance}

[give speedup for linear overlap and Fresnel overlap, for 22 only and HM]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Impact of noise realizations on parameter estimation for LIGO}
\label{sec:noiseligo}

[no-noise corresponds to a geometric average on the posteriors themselves across noise realizations; how does this translate in practice ?]

[Phil's plot: example of the distribution of the MAP across noise realizations, compared to the posterior itself]

[Phil's plot: example(s) of 100 cumulative distributions compared to the no-noise case]

[evolution of results when increasing SNR: 12, 15, 20]

[table giving the ``averaged'' broadening of the posterior and the ``averaged'' departure of the MAP from injection across noise realizations, for all SNR=12 simulations]

[discussion: validation of the no-noise analysis as a first proxy to do parameter estimation estimation]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Impact of merger/ringdown and higher modes on parameter estimation for LISA}
\label{sec:pelisa}

[connection to previous work: McWilliams\&al, Littenberg\&al]

[plot: example posterior with/without HM, EOBNRv2HMROM injection]

[plot: example posterior with/without merger/ringdown, EOBNRv2HMROM injection]

[plot: posterior with information contained in the inspiral only: inspiral-only injection and templates, vs IMR injection and template]

[discuss multimodal structure of the posterior (sky position notably), if it holds]

[improvement on ]

[sky localization errors across the sky with full bayesian inference on a few example cases: a sufficiently large number of inferences for this to be interesting ?]

[accumulation of information with time ? - future studies]

[compare to Fisher matrix analysis on some examples ?]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Summary and Conclusions}
\label{sec:conc}

[extension to longer waveforms, lower masses]

[use of other fast FD waveforms: SEOBNRv2ROM, PhenomP]

[impact of aligned spins ? of simple precession ?]

[impact of merger-ringdown in the aligned spins case and simple precession case ?]

[validation of existing Fischer matrix analysis, at least on some examples, with full Bayesian inference]

[compare different instrument designs, across astrophysical models (Sesana-Barausse\&al)]

[extend analysis to a full population ?]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{4.5mm}

\hspace{0.85in}
{\bf Acknowledgments}

\vspace{3.5mm}

[Acknowledgments]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Details on Fourier-domain reduced order models}
\label{app:rom}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Details on the Fourier-domain LISA response}
\label{app:fdresponse}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{TDI observables}
\label{appsubsec:TDI}

For the constellation delay and modulation, if we associate as above a function $G_{AB}$ to each basic Doppler shift observable $y_{AB}$ (as with the example of $y_{21}$ above), for each of the polarizations we obtain:
%
\begin{align}
	G_{21}^{+,\times} &= i\pi f L \left(n_{3}\cdot P_{+,\times}\cdot n_{3}\right) \sinc\left[ \pi f L \left( 1 + k\cdot n_{3} \right) \right] \exp\left[ i\pi f L \left( 1 - k\cdot (p_{1}+p_{2})/L  \right) \right] \nn\\
	G_{12}^{+,\times} &= i\pi f L \left(n_{3}\cdot P_{+,\times}\cdot n_{3}\right) \sinc\left[ \pi f L \left( 1 - k\cdot n_{3} \right) \right] \exp\left[ i\pi f L \left( 1 - k\cdot (p_{1}+p_{2})/L  \right) \right] \nn\\
	G_{32}^{+,\times} &= i\pi f L \left(n_{1}\cdot P_{+,\times}\cdot n_{1}\right) \sinc\left[ \pi f L \left( 1 + k\cdot n_{1} \right) \right] \exp\left[ i\pi f L \left( 1 - k\cdot (p_{2}+p_{3})/L  \right) \right] \nn\\
	G_{23}^{+,\times} &= i\pi f L \left(n_{1}\cdot P_{+,\times}\cdot n_{1}\right) \sinc\left[ \pi f L \left( 1 - k\cdot n_{1} \right) \right] \exp\left[ i\pi f L \left( 1 - k\cdot (p_{2}+p_{3})/L  \right) \right] \nn\\
	G_{13}^{+,\times} &= i\pi f L \left(n_{2}\cdot P_{+,\times}\cdot n_{2}\right) \sinc\left[ \pi f L \left( 1 + k\cdot n_{2} \right) \right] \exp\left[ i\pi f L \left( 1 - k\cdot (p_{3}+p_{1})/L  \right) \right] \nn\\
	G_{31}^{+,\times} &= i\pi f L \left(n_{2}\cdot P_{+,\times}\cdot n_{2}\right) \sinc\left[ \pi f L \left( 1 - k\cdot n_{2} \right) \right] \exp\left[ i\pi f L \left( 1 - k\cdot (p_{3}+p_{1})/L  \right) \right] \,,
\end{align}
%
where the $p_{A}$ vectors refer to the positions with respect to the center of the constellation $O$ (that is to say, they are given by $O_{2}\cdot p_{A}^{L}$).

The expression for the $X_{A}$ second-generation TDI observables are given by (c.f. Kr\'olak\&al)
%
\begin{widetext}
\begin{align}
	X_{1} &= \left[ \left( y_{31} + y_{13,2} \right) - \left( y_{21} + y_{12,3} \right) - \left( y_{31} + y_{13,2} \right)_{,33} + \left( y_{21} + y_{12,3} \right)_{,22} \right] \nn\\
	& - \left[ \left( y_{31} + y_{13,2} \right) - \left( y_{21} + y_{12,3} \right) - \left( y_{31} + y_{13,2} \right)_{,33} + \left( y_{21} + y_{12,3} \right)_{,22} \right]_{,2233} \,.
\end{align}
\end{widetext}
%
Within the rigid approximation, all the delays are considered equal to $L/c$ as far as the gravitational wave response is concerned. We are applying successive constant delays to the basic observables, and the translation in Fourier domain is simply
%
\begin{widetext}
\begin{align}
	\tilde{X}_{1}(f) &= -4 \sin\left( 2\pi f L \right) \sin\left( 4\pi f L \right) e^{6 i \pi f L} \left[ \tilde{y}_{31} - \tilde{y}_{21} + e^{2i\pi f L} \left( \tilde{y}_{13} - \tilde{y}_{12} \right) \right] \,, \nn\\
	\tilde{X}_{2}(f) &= -4 \sin\left( 2\pi f L \right) \sin\left( 4\pi f L \right) e^{6 i \pi f L} \left[ \tilde{y}_{12} - \tilde{y}_{32} + e^{2i\pi f L} \left( \tilde{y}_{21} - \tilde{y}_{23} \right) \right] \,, \nn\\
	\tilde{X}_{3}(f) &= -4 \sin\left( 2\pi f L \right) \sin\left( 4\pi f L \right) e^{6 i \pi f L}\left[ \tilde{y}_{23} - \tilde{y}_{13} + e^{2i\pi f L} \left( \tilde{y}_{32} - \tilde{y}_{31} \right) \right] \,.
\end{align}
\end{widetext}
%

For the Sagnac observables, we have similarly ((8) of Shaddock\&al)
%
\begin{align}
	\alpha_{1} &= \left[ \left( y_{31} + y_{23,2} + y_{12,12} \right) - \left( y_{21} + y_{32,3} + y_{13,13} \right) \right] \nn\\
	& - \left[ \left( y_{31} + y_{23,2} + y_{12,12} \right)_{,213} - \left( y_{21} + y_{32,3} + y_{13,13} \right)_{,312} \right]
\end{align}
%
and
%
\begin{widetext}
\begin{align}
	\tilde{\alpha}_{1}(f) &= - 2i \sin\left(3\pi f L\right) e^{3 i \pi f L} \left[ \tilde{y}_{31} - \tilde{y}_{21} + e^{2i\pi f L} \left( \tilde{y}_{23} - \tilde{y}_{32} \right) + e^{4i\pi f L} \left( \tilde{y}_{12} - \tilde{y}_{13} \right) \right] \,, \nn\\
	\tilde{\alpha}_{2}(f) &= - 2i \sin\left(3\pi f L\right) e^{3 i \pi f L} \left[ \tilde{y}_{12} - \tilde{y}_{32} + e^{2i\pi f L} \left( \tilde{y}_{31} - \tilde{y}_{13} \right) + e^{4i\pi f L} \left( \tilde{y}_{23} - \tilde{y}_{21} \right) \right] \,, \nn\\
	\tilde{\alpha}_{3}(f) &= - 2i \sin\left(3\pi f L\right) e^{3 i \pi f L} \left[ \tilde{y}_{23} - \tilde{y}_{13} + e^{2i\pi f L} \left( \tilde{y}_{12} - \tilde{y}_{21} \right) + e^{4i\pi f L} \left( \tilde{y}_{31} - \tilde{y}_{32} \right) \right] \,.
\end{align}
\end{widetext}
%
The optimal second-generation combinations $A$, $E$ and $T$ are given by ((57) of Kr\'olak\&al)
%
\begin{align}
	A &= \frac{1}{\sqrt{2}} \left( \alpha_{3} - \alpha_{1} \right) \nn\\
	E &= \frac{1}{\sqrt{6}} \left( \alpha_{1} - 2\alpha_{2} + \alpha_{3} \right) \nn\\
	T &= \frac{1}{\sqrt{3}} \left( \alpha_{1} + \alpha_{2} + \alpha_{3} \right) \,,
\end{align}
%
which gives
%
\begin{widetext}
\begin{align}
	\tilde{A}(f) &= - \frac{2i}{\sqrt{2}} \sin\left(3\pi f L\right) e^{3 i \pi f L} \left[ \left( \tilde{y}_{13} + \tilde{y}_{31} \right) \left( e^{4i\pi f L} - 1 \right) + \left( \tilde{y}_{21} + \tilde{y}_{23} \right) \left( 1 - e^{2i\pi f L} \right) \right. \nn\\
	& \qquad\qquad\qquad\qquad\qquad\qquad \left. + \left( \tilde{y}_{32} + \tilde{y}_{12} \right) \left( e^{2i\pi f L} - e^{4i\pi f L} \right) \right] \nn\\
	\tilde{E}(f) &= - \frac{2i}{\sqrt{6}} \sin\left(3\pi f L\right) e^{3 i \pi f L} \left[ \left( \tilde{y}_{23} - \tilde{y}_{21} \right) \left( 1 + e^{2i\pi f L} - 2e^{4i \pi f L} \right) + \left( \tilde{y}_{31} - \tilde{y}_{13} \right) \left( 1 - 2e^{2i\pi f L} + e^{4i \pi f L} \right) \right. \nn\\
	& \qquad\qquad\qquad\qquad\qquad\qquad + \left( \tilde{y}_{12} - \tilde{y}_{32} \right) \left( -2 + e^{2i\pi f L} + e^{4i \pi f L} \right) \left. \right] \nn\\
	\tilde{T}(f) &= - \frac{2i}{\sqrt{3}} \sin\left(3\pi f L\right) e^{3 i \pi f L} \left( \tilde{y}_{31} - \tilde{y}_{13} + \tilde{y}_{12} - \tilde{y}_{21} + \tilde{y}_{23} - \tilde{y}_{32} \right) \left( 1 + e^{2i\pi f L} + e^{4i \pi f L} \right) \,.
\end{align}
\end{widetext}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Beyond the leading-order Fourier-domain response}
\label{appsubsec:nloresponse}

We derive in the following analytical corrections to this leading order formula. In practice, we expect that the first and second correction terms, at the very most, will be needed.

\subsubsection*{Frequency dependence of the modulation/delay}

If we want to keep the full frequency dependence of the modulation/delay function $G$ in the sum, then we can expand
%
\be
	G[f-n f_{0}](t) = \sum\limits_{k\geq 0} \frac{1}{k!}\left( \frac{-n \Omega}{2\pi} \right)^{k} \frac{\partial^{k} G}{\partial f^{k}}[f](t) \,,
\ee
%
and we have
%
\begin{widetext}
\begin{align}
	\tilde{s}(f) &\simeq \tilde{h}(f) \frac{1}{T} \int_{0}^{T}\ud t\, \sum\limits_{n}\sum\limits_{k} e^{i n \Omega t} \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \frac{1}{k!}\left( \frac{-n \Omega}{2\pi} \right)^{k} \frac{\partial^{k} G}{\partial f^{k}}[f](t) \nn\\
	&= \tilde{h}(f) \sum\limits_{n}\sum\limits_{k} \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \frac{1}{k!} \frac{1}{(2i\pi)^{k}} \frac{1}{T} \int_{0}^{T}\ud t\, (-i n \Omega)^{k}  e^{i n \Omega t}  \frac{\partial^{k} G}{\partial f^{k}}[f](t) \nn\\
	&= \tilde{h}(f) \sum\limits_{n}\sum\limits_{k} \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \frac{1}{k!} \frac{1}{(2i\pi)^{k}} \frac{1}{T} \int_{0}^{T}\ud t\, e^{i n \Omega t}  \frac{\partial^{2k} G}{\partial f^{k}\partial t^{k}}[f](t) \nn\\
	&= \tilde{h}(f) \sum\limits_{k\geq 0} \frac{1}{k!} \frac{1}{(2i\pi)^{k}}  \frac{\partial^{2k} G}{\partial f^{k}\partial t^{k}}[f]\left(-\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) \,,
\end{align}
\end{widetext}
%
where we performed $k$ integrations by parts (using the periodicity of $G$ for any frequency) and summed over $n$ in the same way as in the leading order calculation.

\subsubsection*{Quadratic term in the phase}

If we keep the first correction in the phase beyond the LLP, the quadratic term $\delta^{2}\Psi$, but ignore all the subsequent terms, we can Taylor-expand the exponential according to
%
\begin{widetext}
\begin{align}
	\tilde{h}(f-n f_{0}) &\simeq \tilde{h}(f)  \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \exp\left[ -\frac{1}{2} i (n f_{0})^{2} \frac{\ud^{2}\Psi}{\ud f^{2}} \right] \nn\\
	&= \tilde{h}(f)  \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \sum\limits_{k\geq 0} \frac{1}{k!}\left( \frac{-i}{2} \right)^{k} \left( \frac{n\Omega}{2\pi} \right)^{2k} \left( \frac{\ud^{2}\Psi}{\ud f^{2}}  \right)^{k} \,,
\end{align}
\end{widetext}
% 
and we have
%
\begin{widetext}
\begin{align}
	\tilde{s}(f) &\simeq \tilde{h}(f) \frac{1}{T} \int_{0}^{T}\ud t\, \sum\limits_{n}\sum\limits_{k} e^{i n \Omega t} \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \frac{1}{k!} \left( \frac{-i}{2} \right)^{k} \left( \frac{n\Omega}{2\pi} \right)^{2k} \left( \frac{\ud^{2}\Psi}{\ud f^{2}}  \right)^{k} G[f](t) \nn\\
	&= \tilde{h}(f) \sum\limits_{n}\sum\limits_{k} \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \frac{1}{k!} \left( \frac{i}{2} \right)^{k} \frac{1}{(2\pi)^{2k}} \left( \frac{\ud^{2}\Psi}{\ud f^{2}}  \right)^{k} \frac{1}{T} \int_{0}^{T}\ud t\, (i n \Omega)^{2 k}  e^{i n \Omega t} G[f](t) \nn\\
	&= \tilde{h}(f) \sum\limits_{k\geq 0} \frac{1}{2^k k!} \left( \frac{i}{4\pi^{2}}\frac{\ud^{2}\Psi}{\ud f^{2}}  \right)^{k} \frac{\partial^{k} G}{\partial t^{k}}[f]\left(-\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) \,.
\end{align}
\end{widetext}
%

\subsubsection*{Linear term in the amplitude}

If we keep the first correction in the amplitude, we have
%
\be
	\tilde{h}(f-n f_{0}) \simeq \tilde{h}(f)  \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \left( 1 - \frac{n \Omega}{2\pi} \frac{1}{A}\frac{\ud A}{\ud f} \right) \,,
\ee
%
and we obtain similarly
%
\be
	\tilde{s}(f) \simeq \tilde{h}(f) \left[ G[f]\left(-\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) + \frac{1}{2 i \pi} \frac{1}{A} \frac{\partial G}{\partial t}[f]\left(-\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) \right]\,.
\ee
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\bibliography{ListeRef.bib}
\bibliography{/Users/marsat/Documents/publications/bibliographie/ListeRef.bib}

\end{document}

### 2015-12-08 LISAsim/LISAgeometry.h

@@ -63,6 +63,53 @@ int EvaluateGABmode(
   const double complex Yfactorplus,        /* Spin-weighted spherical harmonic factor for plus */
   const double complex Yfactorcross);      /* Spin-weighted spherical harmonic factor for cross */
 
+/* Functions for the response in time domain */
+double y12TD(
+  gsl_spline* splinehp,                    /* Input spline for TD hplus */
+  gsl_spline* splinehc,                    /* Input spline for TD hcross */
+  gsl_interp_accel* accelhp,               /* Accelerator for hp spline */
+  gsl_interp_accel* accelhc,               /* Accelerator for hc spline */
+  const double t);                         /* Time */
+double y21TD(
+  gsl_spline* splinehp,                    /* Input spline for TD hplus */
+  gsl_spline* splinehc,                    /* Input spline for TD hcross */
+  gsl_interp_accel* accelhp,               /* Accelerator for hp spline */
+  gsl_interp_accel* accelhc,               /* Accelerator for hc spline */
+  const double t);                         /* Time */
+double y23TD(
+  gsl_spline* splinehp,                    /* Input spline for TD hplus */
+  gsl_spline* splinehc,                    /* Input spline for TD hcross */
+  gsl_interp_accel* accelhp,               /* Accelerator for hp spline */
+  gsl_interp_accel* accelhc,               /* Accelerator for hc spline */
+  const double t);                         /* Time */
+double y32TD(
+  gsl_spline* splinehp,                    /* Input spline for TD hplus */
+  gsl_spline* splinehc,                    /* Input spline for TD hcross */
+  gsl_interp_accel* accelhp,               /* Accelerator for hp spline */
+  gsl_interp_accel* accelhc,               /* Accelerator for hc spline */
+  const double t);                         /* Time */
+double y31TD(
+  gsl_spline* splinehp,                    /* Input spline for TD hplus */
+  gsl_spline* splinehc,                    /* Input spline for TD hcross */
+  gsl_interp_accel* accelhp,               /* Accelerator for hp spline */
+  gsl_interp_accel* accelhc,               /* Accelerator for hc spline */
+  const double t);                         /* Time */
+double y13TD(
+  gsl_spline* splinehp,                    /* Input spline for TD hplus */
+  gsl_spline* splinehc,                    /* Input spline for TD hcross */
+  gsl_interp_accel* accelhp,               /* Accelerator for hp spline */
+  gsl_interp_accel* accelhc,               /* Accelerator for hc spline */
[…]

### 2016-04-28 LISAinference/LISAinference_common.c
This commit implements an option for extending the waveforms to low frequency using TaylorF2 for the 22 mode and power-law extrapolation otherwise.  Use the flag --mfmatch=1e-5 for instance to turn on the matching.  On a couple example cases near 1e-6 Msun the results show good matching and only tiny (~1e-3) liklihood differences when the last ~1 order of mag of freq range is extended in this way.

@@ -23,8 +23,9 @@ void addendum(int argc, char *argv[],LISARunParams *runParams, int *ndim, int *n
   /* Parse commandline to read parameters of injection - copy the number of modes demanded for the injection */
   parse_args_LISA(argc, argv, injectedparams, globalparams, priorParams, runParams);
   injectedparams->nbmode = globalparams->nbmodeinj;
-  if(myid == 0) print_parameters_to_file_LISA(injectedparams, globalparams, priorParams, runParams);
 
+  int notLISAlike=strstr(argv[0],"LISAlike")==0;
+  if(myid == 0 && notLISAlike) print_parameters_to_file_LISA(injectedparams, globalparams, priorParams, runParams);
   /* Initialize the data structure for the injection */
   LISAInjectionCAmpPhase* injectedsignalCAmpPhase = NULL;
   LISAInjectionReIm* injectedsignalReIm = NULL;
@@ -65,7 +66,7 @@ void addendum(int argc, char *argv[],LISARunParams *runParams, int *ndim, int *n
       priorParams->dist_max *= SNR123 / priorParams->snr_target;
       if (myid == 0) printf("Distance prior (dist_min, dist_max) = (%g, %g) Mpc\n", priorParams->dist_min, priorParams->dist_max);
     }
-    if (myid == 0) print_rescaleddist_to_file_LISA(injectedparams, globalparams, priorParams, runParams);
+    if (myid == 0 && notLISAlike) print_rescaleddist_to_file_LISA(injectedparams, globalparams, priorParams, runParams);
     if(globalparams->tagint==0) {
       LISAGenerateInjectionCAmpPhase(injectedparams, injectedsignalCAmpPhase);
       SNR123 = sqrt(injectedsignalCAmpPhase->TDI123ss);

### 2016-04-28 LISAinference/LISAutils.c

@@ -136,6 +136,7 @@ Arguments are as follows:\n\
  --fRef                Reference frequency where phiRef is set (Hz, default=0, interpreted as Mf=0.14)\n\
  --deltatobs           Observation time (years, default=2)\n\
  --fmin                Minimal frequency (Hz, default=0) - when set to 0, use the first frequency covered by the noise data of the detector\n\
+ --mfmatch             Minimum matching frequency (1/mtot); if low, will use limit of ROM support., if <=0, then no extension.\n\
  --nbmodeinj           Number of modes of radiation to use for the injection (1-5, default=5)\n\
  --nbmodetemp          Number of modes of radiation to use for the templates (1-5, default=5)\n\
  --tagint              Tag choosing the integrator: 0 for Fresnel (default), 1 for linear integration\n\
@@ -209,6 +210,7 @@ Syntax: --PARAM-min\n\
     globalparams->fRef = 0.;
     globalparams->deltatobs = 2.;
     globalparams->fmin = 0.;
+    globalparams->mfmatch = -1;
     globalparams->nbmodeinj = 5;
     globalparams->nbmodetemp = 5;
     globalparams->tagint = 0;
@@ -300,6 +302,8 @@ Syntax: --PARAM-min\n\
             globalparams->deltatobs = atof(argv[++i]);
         } else if (strcmp(argv[i], "--fmin") == 0) {
             globalparams->fmin = atof(argv[++i]);
+        } else if (strcmp(argv[i], "--mfmatch") == 0) {
+            globalparams->mfmatch = atof(argv[++i]);
         } else if (strcmp(argv[i], "--nbmodeinj") == 0) {
             globalparams->nbmodeinj = atoi(argv[++i]);
         } else if (strcmp(argv[i], "--nbmodetemp") == 0) {
@@ -463,6 +467,7 @@ int print_parameters_to_file_LISA(
   fprintf(f, "fRef:         %.16e\n", globalparams->fRef);
   fprintf(f, "deltatobs:    %.16e\n", globalparams->deltatobs);
   fprintf(f, "fmin:         %.16e\n", globalparams->fmin);
+  fprintf(f, "mfmatch:      %.16e\n", globalparams->mfmatch);
   fprintf(f, "nbmodeinj:    %d\n", globalparams->nbmodeinj);
[…]

### 2016-04-28 LISAinference/LISAutils.h

@@ -47,6 +47,7 @@ typedef struct tagLISAGlobalParams {
   double fRef;               /* reference frequency (Hz, default 0 which is interpreted as Mf=0.14) */
   double deltatobs;          /* max duration of observation (years, default 2) - the start of the signals might be cut in time instead of cut in frequency */
   double fmin;               /* Minimal frequency (Hz) - when set to 0 (default), use the first frequency covered by the noise data of the detector */
+  double mfmatch;            /* Minimum matching frequency (1/mtot); if less than ROM model supports then will extrapolate ROM, if <=0, then no extension. */
   int nbmodeinj;             /* number of modes to include in the injection (starting with 22) - defaults to 5 (all modes) */
   int nbmodetemp;            /* number of modes to include in the templates (starting with 22) - defaults to 5 (all modes) */
   int tagint;                /* Tag choosing the integrator: 0 for wip (default), 1 for linear integration */
@@ -198,6 +199,35 @@ int LISAGenerateInjectionReIm(
   int nbpts,                                  /* Input: number of frequency samples */
   int tagsampling,                            /* Input: tag for using linear (0) or logarithmic (1) sampling */ 
   struct tagLISAInjectionReIm* signal);       /* Output: structure for the generated signal */
+
+/*Wrapper for waveform generation with possibly a combination of EOBNRv2HMROM and TaylorF2*/
+/* Note: GenerateWaveform accepts masses and distances in SI units, whereas LISA params is in solar masses and Mpc */
+int GenerateWaveform(
+  struct tagListmodesCAmpPhaseFrequencySeries **listhlm,  /* Output: list of modes in Frequency-domain amplitude and phase form */
+  int nbmode,                                    /* Number of modes to generate (starting with the 22) */
+  double f_match,                                /* Minimum frequency using EOBNRv2HMROM */
+  double f_min,                                  /* Minimum frequency required */
+  double deltatRef,                              /* Time shift so that the peak of the 22 mode occurs at deltatRef */
+  double phiRef,                                 /* Phase at reference frequency */
+  double fRef,                                   /* Reference frequency (Hz); 0 defaults to fLow */
+  double m1SI,                                   /* Mass of companion 1 (kg) */
+  double m2SI,                                   /* Mass of companion 2 (kg) */
+  double distance                               /* Distance of source (m) */
+		     );
+
+/* Non-spinning merger TaylorF2 waveform, copied and condensed from LAL */
+void TaylorF2nonspin(
[…]

### 2016-04-28 tools/constants.h

@@ -49,6 +49,7 @@ extern "C" {
 /*************** Mathematical constants ******************/
 
 #define PI 3.1415926535897932384626433832795029
+#define GAMMA 0.577215664901532860606512090082402431
 #define sqrt2 1.4142135623730950488
 #define sqrt3 1.7320508075688772935
 #define sqrt6 2.4494897427831780982

### 2016-04-28 docs/paper_method.tex

--- 2015-12-08 docs/paper1.tex
@@
-\documentclass[aps,showpacs,twocolumn,prd,superscriptaddress,nofootinbib]{revtex4}
+\documentclass[aps,showpacs,%twocolumn,
+prd,superscriptaddress,nofootinbib]{revtex4}
 
@@
 \newcommand\calO{{\mathcal{O}}}
+\newcommand\calF{{\mathcal{F}}}
 \newcommand\bfx{\mathbf{x}}
@@
 \newcommand{\Msol}{M_{\odot}}
+\newcommand{\tf}{\tilde{t}_{f}}
+\newcommand{\Tf}{\tilde{T}_{f}}
 
@@
 
-\title{Accelerated prospective parameter estimation of non-spinning compact binaries with higher modes for LIGO/VIRGO and LISA-type detectors}
+\title{Modulation and delays in Fourier domain: precessing binaries and LISA-type detector response}
 
@@
 \label{sec:methodology}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Fourier transform of a modulated and delayed signal}
+\label{subsec:FTgeneral}
+
+Let us consider, in general, the problem of determining the Fourier transform of a signal $h(t)$ to which we apply a varying delay $d(t)$ and a multiplication by a modulation function $F(t)$.
+In the case of precessing binaries, the delays will ba absent, and the modulation functions will be the time-dependent Wigner coefficients applied to rotate the waveform from a precessing frame, where it verifies a restored mode hierarchy, to an inertial frame where the observations take place.
+In the case of a LISA-type detector, the signal $h(t)$ will simply be the waveform in a fixed heliocentric frame, and the delays will come from the motion of each detector in the wave front, while the modulation will represent the time-variation in the detector orientation.
+Both can be considered as periodic functions with a period of a year, which allows for an additional rewriting in terms of Fourier series instead of Fourier transforms, as will be developed in Sec.~\ref{}.
+
+The convention we will be using for Fourier transform will be
+%
+\be\label{eq:defFT}
+	\tilde{h}(f) =  \int \ud t \, e^{+2i\pi f t} h(t) \,.
+\ee
+%
+Notice that this convention uses the opposite sign with respect to the one used in general.
+It was chosen to ensure that, with the conventions of~\cite{}, modes $h_{\ell m}$ with $m>0$ will have support for positive frequencies.
+One can revert to the more usual convention by taking $f\rightarrow -f$.
+
+Defining
+%
+\be
+	h_{d}(t) = h(t+d(t)) \,, \quad s(t) = F(t)h_{d}(t) \,,
+\ee
+%
+we have
+%
+\be
+	h_{d}(t) = \int \ud f \, e^{-2i\pi f (t+d(t))}\tilde{h}(f) \,,
+\ee
+%
+and
+%
+\begin{align}
+	\tilde{s}(f) &= \int \ud t \, e^{2i\pi f t} F(t)  \int \ud f' \, e^{-2i\pi f' (t+d(t))}\tilde{h}(f') \nn\\
+	&= \int \ud f' \, \tilde{h}(f-f') \int \ud t \, e^{2i\pi f' t} e^{-2i\pi (f-f') d(t)} F(t) \,.
+\end{align}
+%
+We can rewrite the last equation as a kind of convolution with a frequency-dependent Kernel, according to
+%
+\be\label{eq:FDkernel}
+	\tilde{s}(f) = \int \ud f' \, \tilde{h}(f-f') \tilde{g}(f-f',f') \,,
+\ee
+%
+where we introduced the frequency-dependent function of time $g(f,t)$, and its Fourier transform in the auxiliary frequency $f'$, denoted by $\tilde{g}(f,f')$, as
+%
+\begin{subequations}
+\begin{align}\label{eq:defg}
+	g(f,t) &= e^{-2i\pi f d(t)} F(t) \,, \\
+	\tilde{g}(f,f') &= \int \ud t \, e^{2i\pi f' t} e^{-2i\pi f d(t)} F(t) \,.
+\end{align}
+\end{subequations}
+%
+In the case of precessing binaries, where the delays $d(t)$ will be absent, the frequency-dependence of $g$ will vanish and this function will simply reduce to the modulation $F(t)$.
+
+In our approach, two assumptions will be crucial to the approximation:
+\begin{itemize}
+	\item the time scales entering the signal $h(t)$ are well separated from the ones entering the modulation and the delay, so that $\tilde{g}(f-f',f')$ has support limited to $[-f_{\rm max},f_{\rm max}]$, with $f_{\rm max}$ small compared the characteristic scale for the variation of the Fourier-domain signal;
+	\item the original signal has a smooth amplitude and phase in the Fourier-domain, slowly varying over the scale $f_{\rm max}$, so that a local treatment in $f'$ of the convolution in~\eqref{eq:FDkernel} will be possible.
+\end{itemize}
+Although well verified for the case of LISA-type detectors, the first assumption can be a limitation of the approximation in the case of preccesing binaries, and we will carefully explore its validity in Sec.~\ref{}.
+
+Notice that these approximations are not exactly the same as the ones usually made to justify the stationary phase approximation (thereafter SPA) for the treatment of the Fourier transform of the inspiral part of the signal.
+Indeed, we will exploit the smoothness of the Fourier-domain $\tilde{h}(f)$ to be able to cover the merger-ringdown part of the signal $h$, where the SPA itself breaks down.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Leading order: the locally linear phase approximation}
+\label{subsec:LLP}
+
+Under the assumptions above, it is possible to see the convolution in the $f'$ variable in~\eqref{eq:FDkernel} as being a local operation, allowing us to expand the signal $\tilde{h}(f-f')$ around $f$.
+
+To this end, we will separate the signal in a Fourier-domain amplitude and phase, according to
+%
+\be
+	\tilde{h}(f) = A(f) e^{-i\Psi(f)} \,,
+\ee
+%
+
+A first leading-order approximation, which could be called the locally linear phase (LLP) approximation, is to discard the $f'$ dependence in the first argument of $\tilde{g}(f-f', f')$, to keep only the leading order in the Fourier-domain amplitude, and to expand the Fourier-domain phase to the first order, according to
+%
+\be
+	\tilde{h}_{\rm LLP}(f-f') \simeq A(f) \exp\left[ -i\left( \Psi(f) - f' \frac{\ud \Psi}{\ud f} \right) \right] \,,
+\ee
+%
+Here and in the rest of the paper, derivatives with respect to $f$ will always be evaluated at $f$.
+
+Plugging this relation in~\eqref{eq:FDkernel}, we obtain
+%
+\begin{align}
+	\tilde{s}_{\rm LLP}(f) &= \tilde{h}(f) \int \ud f' \, \exp\left[ i f' \frac{\ud \Psi}{\ud f} \right] \tilde{g}(f,f') \nn\\
+	&= \tilde{h}(f) g\left( f, -\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) \,,
+\end{align}
+%
+which motivates the introduction of the effective time
+%
+\be\label{eq:deftf}
+	\tf \equiv -\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \,,
+\ee
+%
+where the tilde here does not indicate  a Fourier transform operation but simply that this quantity is an effective time defined from Fourier-domain waveform.
+This definition is a generalization of the time-to-frequency correspondence at the heart of the SPA, as will be shown in Sec.~\ref{}, but it is worth noting the it is defined entirely from the Fourier-domain waveform.
+In particular, $\tf$ will in general not be monotonously increasing with frequency in the high-frequency part of the waveform, corresponding to the ringdown.
+
+Using this notation, we have simply
+%
+\be\label{eq:LLP}
+	\tilde{s}_{\rm LLP}(f) = \tilde{h}(f) g(f, \tf) \,.
+\ee
+%
+The interpretation of this approximation is straightforward: the signal is simply multiplied by the response function evaluated at the time $\tf$.
+Additionally, when delays are present, the frequency-dependence of the response is evaluated directly at $f$.
+For precessing binary, this corresponds to the approximation used for the PhenomP approximant~\cite{}, as we will discuss in Sec.~\ref{}.
+
+It is worth noting that a shift in time of the time-domain signal will, by construction, be appropriately propagated to~\eqref{eq:LLP}.
+If $h'(t) = h(t -  \Delta t)$ but the modulation and delays are left unchanged (which is relevant for space-based detectors, not for precessing binaries), $\Psi'(f) = \Psi(f) - 2\pi f \Delta t$, so that
+%
+\be\label{eq:LLP}
+	\tilde{s}'_{\rm LLP}(f) = \tilde{h}'(f) g(f, \tf + \Delta t) \,. 
+\ee
+%
+This observation also justifies that the first derivative of the Fourier-domain phase cannot, in general, be considered a small quantity, as it can take arbitrary values for arbitrary shifts in time of the signal.
+The leading order approximation corresponds to neglecting the curvature in $\Psi(f)$, hence the name Locally Linear Phase approximation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{General Taylor expansion in the Fourier domain}
+\label{subsec:TaylorFD}
+
+We now extend our approach beyond the LLP approximation.
+We will give a formal derivation of the  general formula for the higher-order corrections in~\eqref{eq:FDkernel}, and then specialize it to the special cases of interest.
+
+We will formally Taylor-expand $\tilde{h}(f-f')$ in $f'$ as follows: we write
+%
+\begin{subequations}
+\begin{align}
+	\Psi(f-f') &= \Psi(f) + 2\pi f' \tf + \sum\limits_{p\geq 2}^{N} \frac{(-1)^{p}}{p!} {f'}^{p} \frac{\ud^{p} \Psi}{\ud f^{p}} \,, \label{eq:expandPsi}\\
+	A(f-f') &= A(f) \sum\limits_{q\geq 0} \frac{(-1)^{q}}{q!} {f'}^{q} \frac{1}{A}\frac{\ud^{q} A}{\ud f^{q}} \,, \label{eq:expandA}\\
+	\tilde{g}(f-f', f') &= \sum\limits_{r\geq 0} \frac{(-1)^{r}}{r!} {f'}^{r} \frac{\partial^{r} }{\partial f^{r}}  \tilde{g}(f,f') \label{eq:expandg}
+\end{align}
+\end{subequations}
+%
+This expansion will only be usable in practice when we will be able to keep only a few terms in each of these expansions.
+We therefore limited the expansion in $\Psi$ to a finite order $N$, to keep the formula readable.
+Notice that we expand $\tilde{g}(f-f',f')$ in $f'$ only in its first argument.
+We can also interchange the $f$-derivatives of $g$ with its Fourier transform.
+
+If we further expand the exponentials to obtain a pure $f'$-expansion, by applying the derivative rule
+%
+\be
+	\int \ud f'\, {f'}^{n} \frac{\partial^{m} }{\partial f^{m}}  \tilde{g}(f,f') e^{-2i\pi f' \tf} = \frac{1}{(-2i\pi)^{n}} \left( \frac{\partial^{m} }{\partial f^{m}} \frac{\partial^{n} }{\partial t^{n}} g \right)(f,\tf) \,,
+\ee
+%
+we arrive at the formal expansion
+%
+\begin{align}
+	\tilde{s}(f) &= \tilde{h}(f) \sum\limits_{q\geq 0} \sum\limits_{r\geq 0} \sum\limits_{k_{2}\geq 0} \dots \sum\limits_{k_{N}\geq 0} \frac{(-1)^{q}}{q!} \frac{(-1)^{r}}{r!} \prod\limits_{p=2}^{N} \frac{1}{k_{p}!}\left( \frac{(-1)^{p}(-i)}{p!} \frac{\ud^{p}\Psi}{\ud f^{p}}\right)^{k_{p}} \frac{1}{A} \frac{\ud^{q} A}{\ud f ^{q}} \frac{1}{(-2i\pi)^m} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{m} }{\partial t^{m}} g \right)(f,\tf) \,, \nn\\
+	\text{where } & m = p+r+2k_{2}+\dots+N k_{N} \,.
+\end{align}
+%
+
+This rather cumbersome expression is easy to simplify for special cases. For instance, if we expand only the frequency-dependence of $g$ as in~\eqref{eq:expandg}, we obtain
+%
+\be
+	\tilde{s}(f) = \tilde{h}(f) \sum\limits_{r\geq 0} \frac{1}{(2i\pi)^{r}r!} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{r} }{\partial t^{r}} g \right)(f,\tf)
+\ee
+%
+which interestingly looks like a Taylor expansion of $g$, but with joint derivatives in frequency and time. Similarly, the expansion~\eqref{eq:expandA} gives
+%
+\be
+	\tilde{s}(f) = \tilde{h}(f) \sum\limits_{q\geq 0} \frac{1}{(2i\pi)^{q}q!} \frac{1}{A} \frac{\ud^{q} A}{\ud f ^{q}}  \left( \frac{\partial^{q} }{\partial t^{q}} g \right)(f,\tf)
+\ee
+%
+
+Of particular interest for us will be the case where we keep only the first term in~\eqref{eq:expandPsi}, and expand the resulting exponential.
+Indeed, this term will be  the dominant correction beyond leading order in the applications that we will consider in Secs.~\ref{} and~\ref{}, and this particular approximation will turn out to be closely related to the SPA approach of~\cite{}.
+The above approach yields
+%
+\be\label{eq:resultdffPsi}
+	\tilde{s}(f) = \tilde{h}(f) \sum\limits_{p\geq 0} \frac{1}{p!} \left( \frac{i}{8\pi^{2}}\frac{\ud^{2} \Psi}{\ud f^{2}} \right)^{p} \left( \frac{\partial^{2p} }{\partial t^{2p}} g \right)(f,\tf) \,,
+\ee
+%
+and it is natural to define a new timescale, as a function of frequency, as
+%
+\be\label{eq:defTf}
+	\Tf^{2} = \frac{1}{4\pi^{2}}\left| \frac{\ud^{2}\Psi}{\ud f^{2}} \right| \,.
+\ee
+%
+The interpretation of this timescale for the inspiralling part of the signal, when the SPA applies, will be discussed in the next section.
+Whe can use this notation to rewrite~\eqref{eq:resultdffPsi} as
+%
+\begin{align}\label{eq:resultdffPsiTf}
+	 \tilde{s}(f) &= \tilde{h}(f) \sum\limits_{p\geq 0} \frac{(-i\epsilon)^{p}}{2^{p}p!} \Tf^{2p} \left( \frac{\partial^{2p} }{\partial t^{2p}} g \right)(f,\tf) \,, \nn \\
+	 \text{with } & \epsilon = -\mathrm{sgn}(\ud^{2}\Psi/\ud f^{2} ) \,.
+\end{align}
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Relation to the stationary phase approximation}
+\label{subsec:linkSPA}
+
+Of particular interest is the relation between these results and the SPA.
+One first writes the time-domain signal in amplitude and phase form as $h(t) = a(t) e^{-2i\varphi(t)}$, where for gravitational wave signals $\varphi$ will correspond to the orbital phase, with the orbital frequency being $\omega = \dot{\varphi}$.
+The factor of 2 in the phase of the wave is appropriate for the dominant 22 harmonic of the signal.
+The approximation then applies to signals verifying the conditions
+%
+\be\label{eq:conditionSPA}
+	\left| \frac{\dot{a}/a}{\omega} \right| \ll 1\,, \quad \left|\frac{\dot{\omega}}{\omega^{2}} \right| \ll 1\,, \quad \left| \frac{(\dot{a}/a)^{2}}{\dot{\omega}} \right| \ll 1 \,.
+\ee
+%
+Since the integral~\eqref{eq:defFT} defining the Fourier transform is rapidly oscillatory unless the term $2\pi f t$ cancels the evolution of $-2\varphi(t)$, its support is well centered around the point of stationary phase.
+This defines the time of stationary phase implicitly by the relation
+%
+\be\label{eq:deftfSPA}
+	\omega(t_{f}) = \pi  f \,.
+\ee
+% 
+For a chirping signal of increasing phase, $\omega>0$ and $\dot{\omega}>0$, there is a unique point of stationary phase, located in the positive frequency range $f>0$.
+Using the conditions above, one formally expands the signal around $t_{f}$, to quadratic order in the time.
+The resulting integral yields
+%
+\begin{subequations}
+\begin{align}
+	\tilde{h}_{\rm SPA}(f) &= A_{\rm SPA}(f) e^{-i\Psi_{\rm SPA}(f)} \,, \\
+	A_{\rm SPA}(f) &= a(t_{f}) \sqrt{\frac{\pi}{\dot{\omega}(t_{f})}} \,, \\
+	\Psi_{\rm SPA}(f) &= 2\varphi(t_{f}) - 2\pi f t_{f} + \frac{\pi}{4} \,.
+\end{align}
+\end{subequations}
+%
+For non-precessing gravitational wave signals, the SPA is a very good approximation through the inspiralling phase.
+Ref.~\cite{} evaluated the first correction to this approximation and found that it can be considered as a term of the fifth post-Newtonian order.
+However, it is important to note that the approximation breaks down when reaching the merger-ringdown part of the signal, as the rapidly varying features of the merger break the locality of the time-to-frequency relation~\eqref{eq:deftfSPA}.
+
+Taking the first derivative of the defining relation~\eqref{eq:deftfSPA} gives
+%
+\be
+	t_{f} = -\frac{1}{2\pi} \frac{\ud \Psi_{\rm SPA}}{\ud f} \,,
+\ee
+%
+which means that our generalized definition~\eqref{eq:deftf} for the effective time $\tf$ coincides with $t_{f}$ where the SPA is applicable. Similarly, if we define
+%
+\be
+	T_{f} \equiv \frac{1}{\sqrt{2\dot{\omega}(t_{f})}} \,,
+\ee
+%
+we have
+%
+\be
+	T_{f}^{2} = -\frac{1}{4\pi^{2}}  \frac{\ud^{2} \Psi_{\rm SPA}}{\ud f^{2}} \,,
+\ee
+%
+where $\ud^{2}\Psi/\ud f^{2} < 0$ in the SPA with our sign conventions.
+Comparing with~\eqref{eq:defTf} above, we see that our generalized time scale $\Tf$ also reduces to $T_{f}$ when the SPA applies.
+However, its definition extends to the merger-ringdown part of the signal, and we have to include an absolute value to allow for a possible change of sign of $\ud^{2}\Psi/\ud f^{2}$ in that regime.
+
+We can now compare our result to the approach of Ref.~\cite{}, where the authors extend the SPA, in a formalism called the shifted asymptotic expansion (SUA).
+The derivation includes an expansion of the modulation in Bessel functions, a study of the displacement of the stationary phase point induced by the precession, and a resummation of the result in the form of time derivatives.
+The main result of~\cite{}, their equation~(34), reads exactly as~\eqref{eq:resultdffPsiTf} with the identifications $\tilde{H}_{corr}(f)\rightarrow \tilde{s}(f)/\tilde{h}(f)$, $T\rightarrow \Tf$ and $e^{-i\delta\phi} \rightarrow F$.
+Thus, we are able to identify the Fourier-domain approximation that the treatment of~\cite{} amounts to, namely the approximation~\eqref{eq:expandPsi} where in the Fourier-domain convolution~\eqref{} the phase of the signal is expanded to quadratic order and the amplitude is not expanded.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Magnitude of the corrections beyond the LLP approximation}
+\label{subsec:sizecorr}
+
+\subsubsection{Inspiral of a precessing binary}
+\label{subsubsec:sizecorrprecession}
+
+We investigate now the post-Newtonian order counting of the expansion~\eqref{eq:resultdffPsiTf} during the inspiralling phase of a precessing binary, with modulation $F$ as in~\eqref{}.
+Using the notation $\calO(n) = \calO(c^{-n})$, we have for quasi-cirular inspirals that $\dot{\omega} = \calO(5)$, corresponding to the radiation reaction effect at the 2.5PN order.
+The mode mixing will be induced by the rotation of the orbital angular momentum $\bm{L}$ around the total angular momentum $\bm{J}$, with the angular velocity for this precession $\bm{\Omega} = \calO(2)$.
+The relevant scaling for the terms in the series~\eqref{eq:resultdffPsiTf}, for the inspiral phase where the SPA applies, is therefore
+%
+\be
+	\Tf^{2p} \partial_{t}^{2p} F = \calO(-1) \,,
+\ee
+%
+meaning that the expansion becomes {\it less} well behaved for {\it lower} frequencies.
+The presence of a $1/p!$ gives a formal sense to the expansion if the successive derivatives of $F(t)$ are appropriately bounded, but one may have to keep a large number of terms in the sum.
+The approximation that consists in taking the first term in the series, the LLP approximation~\eqref{} which was in particular used for the PhenomP~\cite{} waveform model, gets {\it worse} the further away from merger the signal is.
+[This does not seem to have been clearly stated before, to check].
+We investigate further in Sec.~\ref{} the validity of the treatment of precession used for PhenomP.
+
+Notice that, for the amplitude derivatives in~\eqref{eq:expandA} and the higher order derivatives of the phase in~\eqref{eq:expandPsi}, the same PN scaling arguments give however the opposite, more usual behaviour that the terms beyond the first one appear as corrections becoming less and less important at lower frequencies.
+For instance, within the SPA we obtain
+%
+\begin{subequations}
+\begin{align}
+	\frac{\ud^{3} \Psi}{\ud f^{3}} \partial_{t}^{3} F &= \calO(6) \,, \\
+	\frac{1}{A}\frac{\ud A}{\ud f} \partial_{t} F &= \calO(2) \,.
+\end{align}
+\end{subequations}
+%
+
+\subsubsection{The case of a LISA-type detector [TO BE UPDATED]}
+\label{subsubsec:sizecorrLISA}
+
+If we put actual numbers, we have $\Omega/2\pi \simeq 3.1\times10^{-8}\mathrm{Hz} \ll f$ over all the frequency band we consider.
+Considering that the comb is of limited extend in $n$, we might want to Taylor expand the relevant quantities accross the comb.
+Since the phase is rapidly varying, and since we have to consider a constant and a linear term in the phase as arbitrary, it is natural to Taylor expand the phase at the first order (that is, approximate the phase linearly locally), neglect the $n\Omega/2\pi$ in $(f-n\Omega/2\pi)$ appearing in the argument of the amplitude and other terms, and assess the quality of this approximation.
+
+Preliminary tests show that this approximation should be valid, unsurprisingly, for all the mildly varying functions encountered: $\sinc$, $\exp$ and the linear term in the $c_{n}$'s, and the Bessel functions $J_{n}$.
+However, $\tilde{h}$ has a Fourier-domain phase which becomes very rapidly varying at low frequencies, in the deep inspiral, $\Psi\propto f^{-5/3}$; here it is not obvious that we can neglect the second derivative of the phase.
+
+For a Newtonian inspiral, applying the SPA gives $\tilde{h}(f) = A_{N}(f)e^{-i\Psi_{N}(f)}$ with
+%
+\begin{subequations}
+\begin{align}
+	A_{N}(f) &= -\sqrt{\frac{5\pi}{24}} \frac{G^{2}m^{2}}{Dc^{5}} \nu^{1/2}v^{-7/2}\,,\\
+	\Psi_{N}(f) &= 2\pi f t_{0} - \phi_{0} + \frac{3}{128\nu v^{5}} \,, 
+\end{align}
+\end{subequations}
+%
+where $v=(G m \pi f/c^{3})^{1/3}$. The magnitude of the first terms in a Taylor expansion in $f_{0}$ is (taking $10^{-5}\mathrm{Hz}$ as a reference starting frequency):
+%
+\begin{widetext}
+\begin{subequations}
+\begin{align}
+	|\delta A/A| &= \left| \frac{1}{A_{N}}\frac{\ud A_{N}}{\ud f} f_0 \right| \simeq 3\times10^{-3} \left(\frac{f}{10^{-5}\Hz}\right)^{-1} \,,\\
+	|\delta^{1}\Psi | &= \left| f_0\frac{\ud \Psi_{N}}{\ud f} \right| \simeq 10^{3}\left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-8/3} \,, \\
+	|\delta^{2}\Psi | &= \left| \frac{1}{2} f_0^{2}\frac{\ud^{2} \Psi_{N}}{\ud f^{2}} \right| \simeq 5 \left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-11/3} \,, \\
+	|\delta^{3}\Psi | &= \left| \frac{1}{6} f_0^{3}\frac{\ud^{3} \Psi_{N}}{\ud f^{3}} \right| \simeq 2\times 10^{-2} \left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-14/3} \,.
+\end{align}
+\end{subequations}
+\end{widetext}
+%
+The approximation degrades for lower starting frequencies, and, at a fixed frequency, when the total mass decreases and/or the mass ratio increases.
+However, if we consider only merging binaries, that is to say binaries of which the merger is seen during the observation period of the mission, we have a lower limit for the starting frequency which can be larger than the detector's lowest frequency in band.
+
+Namely, if we take a Newtonian estimate for the relation between the frequency and the time to coalescence, we have
+%
+\be
+	\pi f(t) = \left[ \frac{256\nu}{5c^{5}} (Gm)^{5/3} \Delta t \right]^{-3/8} \,,
+\ee
+%
+with $\Delta t$ the time to coalescence, and the starting frequency is then
+%
+\be
+	f_{\rm start} = 10^{-5}\mathrm{Hz} \left( \frac{m}{8.5\times10^{6}\Msol} \right)^{-5/8} \left( \frac{\nu}{1/4} \right)^{-3/8} \left( \frac{\Delta t}{5 \mathrm{yr}} \right)^{-3/8} \,.
+\ee
+%
+So we see that the ``worst'' total mass is $8.5\times10^6\Msol (\nu/(1/4))^{3/5}$, for 5 years of observation and $f_{\rm low} = 10^{-5}\mathrm{Hz}$ for the entry in band, which are conservative values; for higher masses, part of the signal is cut because it is out-of-band, while for lower masses, the signal is cut because of the limited time of observation before merger, and we have the scaling
+%
+\be
+	|\delta^{2}\Psi |_{f_\mathrm{start}} \propto m^{5/8}\nu^{3/8}\Delta t^{11/8} \,,
+\ee
+%
+which shows that the effect of the truncation of the signal for a finite observation time wins: the approximation becomes better for lower mass and higher mass ratios systems.
+The worse point in the parameter space should be, for these Newtonian estimates, at equal mass and $m=8.5\times 10^{6} \Msol$, with $|\delta^{2}\Psi |_{\mathrm{worst}} \simeq 0.15$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Exact quadratic term in the phase and Fresnel transform}
+\label{subsec:fresneltransform}
+
+Since the second-derivative corrective term in~\eqref{} is the most interesting, since it contains the most important deviation from the LLP, we will investigate it in more details in this section.
+For simplicity of notation, here we will keep to the case of precessing binaries, with only a modulation $F$ and no delays.
+
+To obtain~\eqref{eq:resultdffPsi}, we expanded the exponential. If we keep it as it is, however, we have
+%
+\begin{subequations}
+\begin{align}
+	\tilde{s}(f) &= \tilde{h}(f) \int\ud f'\, \tilde{F}(f') e^{-2i\pi f'\tf} \exp\left[ -\frac{i}{2} \frac{\ud^{2} \Psi}{\ud f^{2}} {f'}^{2} \right] \, \\
+	&= \tilde{h}(f) \int \ud t\, F(t) \int\ud f'\, e^{2i\pi f' (t-\tf)} \exp\left[ 2i\pi^{2} \epsilon{f'}^{2} \Tf^{2} \right] \,,
+\end{align}
+\end{subequations}
+%
+where we recall that $\epsilon = -\mathrm{sgn}(\ud ^{2} \Psi/\ud f^{2})$. The resulting Gaussian integral yields [check how $\epsilon$ propagates]
+%
+\begin{subequations}
+\begin{align}\label{eq:resultFresnel}
+	\tilde{s}(f) &= \tilde{h}(f) \calF_{\Tf}[F](\tf)^{*} \quad \text{if } \epsilon = 1 \,,\\
+	\tilde{s}(f) &= \tilde{h}(f) \calF_{\Tf}[F](\tf) \quad \text{if } \epsilon = -1 \,,
+\end{align}
+\end{subequations}
+%
+where we introduced the Fresnel transform~\cite{} of $F$, defined as follows:
+%
+\be\label{eq:defFresnel}
+	\calF_{\tau}[F](t_{0}) \equiv \frac{e^{-i\frac{\pi}{4}}}{\sqrt{2\pi} \tau} \int \ud t \, \exp\left[ \frac{i}{2} \left( \frac{t-t_{0}}{\tau} \right)^{2}\right] F(t) \,.
+\ee
+%
+The Fresnel transform applies to a large class of functions [be more precise] and intervenes in the theory of diffraction [give references and examples].
+Although it is interesting to be able to relate our results to a known transformation used in signal processing, whether this opens avenues for efficient computation methods remains to be investigated (for instance by performing a linear decomposition of the modulation $F$ in wavelets, or chirplets, which have analytically computable Fresnel transforms).
+
+As a verification, upon performing a Taylor expansion in time of $F(t)$ around $\tf$ in the above integral and integrating  term by term the result, one readily recovers our previous result~\eqref{eq:resultdffPsiTf}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Resummation of derivatives}
+\label{subsec:resummation}
+
+The authors of~\cite{} did not reformulate their result as in~\eqref{eq:resultFresnel}, but they proposed  a resummation scheme.
+Indeed, the result~\eqref{} looks like a symmetrixed Taylor expansion, except for the factors $i^{p}$ and $1/p!$ instead of $1/(2p)!$.
+After having truncated the sum at some finite $N$, one can write
+%
+\begin{align}\label{eq:resum}
+	\tilde{s}(f) &\simeq \tilde{h}(f) \sum\limits_{p = 0}^{N} \frac{(-i\epsilon\Tf^{2})^{p}}{2^{p}p!} \partial_{t}^{2p}F(\tf) \nn\\
+	&\simeq \tilde{h}(f) \sum\limits_{p= 0}^{N} \sum\limits_{k=0}^{N} a_{N,k}\frac{(k\Tf)^{2p}}{(2p)!}  \partial_{t}^{2p}F(\tf) \nn\\
+	&\simeq \tilde{h}(f) \frac{1}{2}\sum\limits_{k=0}^{N} a_{N,k} \left( F(\tf + k\Tf) + F(t-k\Tf) \right)\,,
+\end{align}
+%
+where we applied a truncation approximation at the first and last stage.
+The second equality holds provided that the complex coefficients $a_{N,k}$ are a solution of the $N+1$-dimensional linear system
+%
+\be
+	(-i\epsilon)^{p} (2p-1)!! = \sum\limits_{k=0}^{N} a_{N,k} k^{2p} \quad \text{for } p=0,\dots,N \,.
+\ee
+%
+The two solutions for the comb coefficients in the two cases $\epsilon = \pm 1$ are simply related by a complex conjugation.
+
+An immediate advantage of this reformulation is its improved numerical stability.
+Although one should expect an ideal signal the scaling $\partial_{t}^{p} F \sim \Omega^{p} F$ with $\Omega$ the precession frequency, it can be extremely hard to control numerical derivatives of $F$ (which in any realistic model will be only provided as sampled at discrete times) well enough to enforce this scaling.
+Here, however, one simply evaluates the original smooth function $F$ at shifted times.
+
+Note, interestingly, that, if the validity of the comb approach was generic, and if the coefficients $a_{N,k}$ were to converge for $N\rightarrow +\infty$ to a well-defined sequence admitting an appropriate fall-off with $k$, then one would have found a magical way of computing the Fresnel transform of a signal by simply evaluating it on a comb made of at most a few dozen equally spaced times.
+
+This seems to good to be true.
+In general, one could have made any ansatz for the series expansion to be matched on, and one therefore cannot expect in general nice properties to emerge from a resummation.
+However, the transformation~\eqref{} between the two series above has a special structure.
+Namely, written in matrix form, the linear system is a Vandermonde system.
+Defining the Vandermonde matrix $V(x_{0},\dots,x_{N})$ as $V_{ij} = (x_{j})^{i}$ (with the convention $0^{0} = 1$), setting $x_{j} = j^{2}$, the linear system is simply
+%
+\be
+	b_{p} = V_{p,k} a_{N,k} \,.
+\ee
+%
+Now, since the Vandermonde system solves the Lagrange polynomial interpolation problem, we know that .
+Namely, $({}^{t}V^{-1})_{ij}$ is the coefficient of $X^{i}$ in the polynomial $\prod_{k\neq j} (X-x_{k})/(x_{j} - x_{k})$.
+If we introduce the symmetric polynomials $\sigma$ such that
+%
+\be	
+	\sigma(m, \{x_{1}, \dots, x_{n}\}) = \sum\limits_{1\leq i_{1}<\dots<i_{m}\leq n} x_{i_{1}}\dots x_{i_{m}} \,,
+\ee
+%
+we obtain for the inverse of the matrix $V$
+%
+\be
+	V^{-1}_{ij} = \frac{1}{\prod_{k\neq i} (i^{2} - k^{2})} (-1)^{N-j} \sigma(N-j, \{0,\dots,N^{2}\}\backslash \{i^{2}\}) \,.
+\ee
+%
+This allows us to write an ``analytical'' expression for the $a_{N,k}$ for every finite order $N$:
+%
+\be
+	a_{N,k} = \sum\limits_{p=0}^{N} (-i\epsilon)^{p}(2p-1)!!
+\frac{(-1)^{N-p}}{\prod_{k\neq p} (p^{2}-k^{2})} \sum\limits_{\substack{ 0 \leq i_{1} < \dots < i_{N-p} \leq N \\ i_{1}, \dots, i_{N-p} \neq p}} i_{1}^{2}\dots i_{N-p}^{2}
+\ee
+%
+The combinatorial explosion of the number of terms in the symmetric polynomials, however, makes this expression impractical to evaluate for $N$ beyond a few dozen.
+It could allow, however, one to control the asymptotics and maybe prove the convergence or non-convergence of the sequence of $a_{N,k}$ for $N\rightarrow +\infty$.
+[wishful thinking]
+
+\section*{---BELOW IS OUTDATED MATERIAL---}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection*{Locally linear phase approximation (LLP)}
-
-If we put actual numbers, we have $\Omega/2\pi \simeq 3.1\times10^{-8}\mathrm{Hz} \ll f$ over all the frequency band we consider.
-Considering that the comb is of limited extend in $n$, we might want to Taylor expand the relevant quantities accross the comb.
-Since the phase is rapidly varying, and since we have to consider a constant and a linear term in the phase as arbitrary, it is natural to Taylor expand the phase at the first order (that is, approximate the phase linearly locally), neglect the $n\Omega/2\pi$ in $(f-n\Omega/2\pi)$ appearing in the argument of the amplitude andother terms, and assess the quality of this approximation.
-
-Preliminary tests show that this approximation should be valid, unsurprisingly, for all the mildly varying functions encountered: $\sinc$, $\exp$ and the linear term in the $c_{n}$'s, and the Bessel functions $J_{n}$.
-However, $\tilde{h}$ has a Fourier-domain phase which becomes very rapidly varying at low frequencies, in the deep inspiral, $\Psi\propto f^{-5/3}$; here it is not obvious that we can neglect the second derivative of the phase.
-
-For a Newtonian inspiral, applying the SPA gives $\tilde{h}(f) = A_{N}(f)e^{-i\Psi_{N}(f)}$ with
-%
-\begin{subequations}
-\begin{align}
-	A_{N}(f) &= -\sqrt{\frac{5\pi}{24}} \frac{G^{2}m^{2}}{Dc^{5}} \nu^{1/2}v^{-7/2}\,,\\
-	\Psi_{N}(f) &= 2\pi f t_{0} - \phi_{0} + \frac{3}{128\nu v^{5}} \,, 
-\end{align}
-\end{subequations}
-%
-where $v=(G m \pi f/c^{3})^{1/3}$. The magnitude of the first terms in a Taylor expansion in $f_{0}$ is (taking $10^{-5}\mathrm{Hz}$ as a reference starting frequency):
-%
-\begin{widetext}
-\begin{subequations}
-\begin{align}
-	|\delta A/A| &= \left| \frac{1}{A_{N}}\frac{\ud A_{N}}{\ud f} f_0 \right| \simeq 3\times10^{-3} \left(\frac{f}{10^{-5}\Hz}\right)^{-1} \,,\\
-	|\delta^{1}\Psi | &= \left| f_0\frac{\ud \Psi_{N}}{\ud f} \right| \simeq 10^{3}\left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-8/3} \,, \\
-	|\delta^{2}\Psi | &= \left| \frac{1}{2} f_0^{2}\frac{\ud^{2} \Psi_{N}}{\ud f^{2}} \right| \simeq 5 \left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-11/3} \,, \\
-	|\delta^{3}\Psi | &= \left| \frac{1}{6} f_0^{3}\frac{\ud^{3} \Psi_{N}}{\ud f^{3}} \right| \simeq 2\times 10^{-2} \left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-14/3} \,.
-\end{align}
-\end{subequations}
-\end{widetext}
-%
-The approximation degrades for lower starting frequencies, and, at a fixed frequency, when the total mass decreases and/or the mass ratio increases.
-However, if we consider only merging binaries, that is to say binaries of which the merger is seen during the observation period of the mission, we have a lower limit for the starting frequency which can be larger than the detector's lowest frequency in band.
-
-Namely, if we take a Newtonian estimate for the relation between the frequency and the time to coalescence, we have
-%
-\be
-	\pi f(t) = \left[ \frac{256\nu}{5c^{5}} (Gm)^{5/3} \Delta t \right]^{-3/8} \,,
-\ee
-%
-with $\Delta t$ the time to coalescence, and the starting frequency is then
-%
-\be
-	f_{\rm start} = 10^{-5}\mathrm{Hz} \left( \frac{m}{8.5\times10^{6}\Msol} \right)^{-5/8} \left( \frac{\nu}{1/4} \right)^{-3/8} \left( \frac{\Delta t}{5 \mathrm{yr}} \right)^{-3/8} \,.
-\ee
-%
-So we see that the ``worst'' total mass is $8.5\times10^6\Msol (\nu/(1/4))^{3/5}$, for 5 years of observation and $f_{\rm low} = 10^{-5}\mathrm{Hz}$ for the entry in band, which are conservative values; for higher masses, part of the signal is cut because it is out-of-band, while for lower masses, the signal is cut because of the limited time of observation before merger, and we have the scaling
-%
-\be
-	|\delta^{2}\Psi |_{f_\mathrm{start}} \propto m^{5/8}\nu^{3/8}\Delta t^{11/8} \,,
-\ee
-%
-which shows that the effect of the truncation of the signal for a finite observation time wins: the approximation becomes better for lower mass and higher mass ratios systems.
-The worse point in the parameter space should be, for these Newtonian estimates, at equal mass and $m=8.5\times 10^{6} \Msol$, with $|\delta^{2}\Psi |_{\mathrm{worst}} \simeq 0.15$.
 

### 2016-07-02 docs/paper_method.tex
Small changes to datura Makefile and to the paper.

@@
 \begin{align}\label{eq:resultFresnel}
-	\tilde{s}(f) &= \tilde{h}(f) \calF_{\Tf}[F](\tf)^{*} \quad \text{if } \epsilon = 1 \,,\\
-	\tilde{s}(f) &= \tilde{h}(f) \calF_{\Tf}[F](\tf) \quad \text{if } \epsilon = -1 \,,
+	\tilde{s}(f) &= \tilde{h}(f) \calF_{\Tf}[F](\tf) \quad \text{if } \epsilon = 1 \,,\\
+	\tilde{s}(f) &= \tilde{h}(f) \calF_{\Tf}[F](\tf)^{*} \quad \text{if } \epsilon = -1 \,,
 \end{align}
@@
 \be\label{eq:defFresnel}
-	\calF_{\tau}[F](t_{0}) \equiv \frac{e^{-i\frac{\pi}{4}}}{\sqrt{2\pi} \tau} \int \ud t \, \exp\left[ \frac{i}{2} \left( \frac{t-t_{0}}{\tau} \right)^{2}\right] F(t) \,.
+	\calF_{\tau}[F](t_{0}) \equiv \frac{e^{i\frac{\pi}{4}}}{\sqrt{2\pi} \tau} \int \ud t \, \exp\left[ - \frac{i}{2} \left( \frac{t-t_{0}}{\tau} \right)^{2}\right] F(t) \,.
 \ee
@@
 	&\simeq \tilde{h}(f) \sum\limits_{p= 0}^{N} \sum\limits_{k=0}^{N} a_{N,k}\frac{(k\Tf)^{2p}}{(2p)!}  \partial_{t}^{2p}F(\tf) \nn\\
-	&\simeq \tilde{h}(f) \frac{1}{2}\sum\limits_{k=0}^{N} a_{N,k} \left( F(\tf + k\Tf) + F(t-k\Tf) \right)\,,
+	&\simeq \tilde{h}(f) \frac{1}{2}\sum\limits_{k=0}^{N} a_{N,k} \left( F(\tf + k\Tf) + F(\tf-k\Tf) \right)\,,
 \end{align}
@@
 Namely, written in matrix form, the linear system is a Vandermonde system.
-Defining the Vandermonde matrix $V(x_{0},\dots,x_{N})$ as $V_{ij} = (x_{j})^{i}$ (with the convention $0^{0} = 1$), setting $x_{j} = j^{2}$, the linear system is simply
+Defining the Vandermonde matrix $V(x_{0},\dots,x_{N})$ as $V_{ij} = (x_{j})^{i}$ (with the convention $0^{0} = 1$), setting $x_{j} = j^{2}$, and $b_{p} \equiv (-i\epsilon)^{p}(2p-1)!!$, the linear system is simply
 %
@@
 	a_{N,k} = \sum\limits_{p=0}^{N} (-i\epsilon)^{p}(2p-1)!!
-\frac{(-1)^{N-p}}{\prod_{k\neq p} (p^{2}-k^{2})} \sum\limits_{\substack{ 0 \leq i_{1} < \dots < i_{N-p} \leq N \\ i_{1}, \dots, i_{N-p} \neq p}} i_{1}^{2}\dots i_{N-p}^{2}
+\frac{(-1)^{N-p}}{\prod_{j\neq k} (j^{2}-k^{2})} \sum\limits_{\substack{ 0 \leq i_{1} < \dots < i_{N-p} \leq N \\ i_{1}, \dots, i_{N-p} \neq k}} i_{1}^{2}\dots i_{N-p}^{2}
 \ee

### 2016-08-13 docs/paper_method.tex
Changes to accomodate response at high frequencies. Various other changes.
The response at thigh frequencies (for low masses) shows much more structure, including zero-crossings. We now do a resampling of the high-f part to ensure enough points to resolve this structure. The response now also includes the new first order correction to the R-delay.

Other changes:
- EvaluateGABmode now has a tag to include or not the R-delay phase term
- interface for PN extension changed: now tagextpn is a tag to indicate wether to do a PN extension at all, Mfmatch used to indicate wodn to which frequency to u

@@
 %
-\be
+\be\label{eq:resultdfA}
 	\tilde{s}(f) = \tilde{h}(f) \sum\limits_{q\geq 0} \frac{1}{(2i\pi)^{q}q!} \frac{1}{A} \frac{\ud^{q} A}{\ud f ^{q}}  \left( \frac{\partial^{q} }{\partial t^{q}} g \right)(f,\tf)
@@
 %
+
+\subsubsection{Delays in the LLP approximation for LISA-type detectors}
+\label{subsubsec:LISAorbitaldelay}
+
+In the case of a LISA-type detector response, it is interesting to consider the case where the .
+The following will be mostly relevant for the delay induced by the motion of the constellation around the sun.
+Indeed, the analysis of the magnitude of the corrections beyond the LLP shows that it is the only one where the figure of merit of the approximation can exceed $10^{-1}$, due to the large baseline of the delay.
+
+If one can neglect the signal related amplitude and phase corrections as given in~\eqref{eq:resultdfA} and~\eqref{eq:resultdffPsiTf}, but one keeps the frequency-dependence in the convolution kernel in~\eqref{}, the general form of the response reads
+\begin{equation}
+	\tilde{s}(f) = \tilde{h}(f) \int \ud t \, F(t) e^{-2i\pi f d(t)} \int \ud f' \, e^{2i\pi f' (t+d(t) - t_{f})} \,,
+\end{equation}
+where we recall that we defined $t_{f} \equiv -1/(2\pi)\ud \Psi/\ud f$. If we introduce a new time, defined implicitly by
+\be
+	\left. (t + d(t) - t_{f})\right|_{t=t_{f}^{d}} = 0 \,,
+\ee
+The above integral can be formally computed as
+\begin{align}
+	\tilde{s}(f) &= \tilde{h}(f) \int \ud t \, F(t) e^{-2i\pi f d(t)} \delta(t + d(t) - t_{f}) \nn \\
+	&= \tilde{h}(f) F(t_{f}^{d}) \frac{e^{-2i\pi f d(t_{f}^{d})}}{1+\dot{d}(t_{f}^{d})}
+\end{align}
+In the case of the delay associated with the motion of the constellation around the Sun, the delay takes here a simple form (restoring factors $c$),
+\be
+	d(t) = \frac{R}{c} \cos(\beta) \cos(\Omega_{0} t -\lambda) \,,
+\ee
+with $R/c \simeq 500 \mathrm{s}$ and $\Omega_{0}R/c \simeq []$.
+In the case of the delays between spacecrafts of the constellation, the delays have a more complicated time dependence and scale instead as $L/c \simeq [] \mathrm{s}$ and $\Omega_{0} L/c \simeq []$.
+
+Since the propagation delays are small compared to the characteristic timescale of the orbital motion $T_{0} = 1\mathrm{yr}$, we will be able to treat $d$ as a perturbation in the equations above~\eqref{}.
+By doing so, we obtain for the delayed time at first order
+\begin{align}
+	t_{f}^{d} &= t_{f} - d(t_{f}) ( 1 - \dot{d}(t_{f})) + \calO(\Omega_{0}d) \,, \nn\\
+	d(t_{f}^{d}) &= d(t_{f}) ( 1 - \dot{d}(t_{f})) + \calO(\Omega_{0}d) \,.
+\end{align}
+Now, since $\dot{d} \ll 1$, we can consider that the only relevant correction in~\eqref{} comes from the phase factor, where the factor $2\pi f d$ gives an additional magnification by a factor $3.10^{3}$.
+We thus arrive in this case at the approximate form
+\be
+	\tilde{s}(f) \simeq \tilde{h}(f) F(t_{f})\exp\left[ -2i\pi f d(t_{f}) (1-\dot{d}(t_{f})) \right] \,.
+\ee
 

### 2016-08-13 EOBNRv2HMROM/EOBNRv2HMROM.c

@@ -4,7 +4,7 @@
  * \brief C code for EOBNRv2HM reduced order model (non-spinning version).
  * See CQG 31 195010, 2014, arXiv:1402.4146 for details on the reduced order method.
  * See arXiv:1106.1021 for the EOBNRv2HM model.
- *  
+ *
  * Borrows from the SEOBNR ROM LAL code written by Michael Puerrer and John Veitch.
  *
  * Put the untared data in the directory designated by the environment variable ROM_DATA_PATH.
@@ -55,11 +55,11 @@
 /* nbmodemax = 5 has been defined in the header */
 const int listmode[nbmodemax][2] = { {2,2}, {2,1}, {3,3}, {4,4}, {5,5} };
 
-/* Maximal mass ratio covered by the model - q=12 (almost) for now */ 
+/* Maximal mass ratio covered by the model - q=12 (almost) for now */
 static const double q_max = 11.9894197212;
-/* Minimal geometric frequency covered by the model - f=8Hz for M=10Msol for now */ 
+/* Minimal geometric frequency covered by the model - f=8Hz for M=10Msol for now */
 static const double Mf_ROM_min = 0.0003940393857519091;
-/* Maximal geometric frequency covered by the model - Mf=0.285 for the 55 mode - used as default */ 
+/* Maximal geometric frequency covered by the model - Mf=0.285 for the 55 mode - used as default */
 static const double Mf_ROM_max = 0.285;
 /* Total mass (in units of solar mass) used to generate the ROM - used to restore the correct amplitude (global factor M) when coming back to physical units */
 static const double M_ROM = 10.;
@@ -196,7 +196,7 @@ void EOBNRHMROMdata_interp_Cleanup(EOBNRHMROMdata_interp *data_interp) {
   free(data_interp);
 }
 
-/* Read binary ROM data for frequency vectors, coefficients matrices, basis functions matrices, and shiftvectors in time and phase */ 
+/* Read binary ROM data for frequency vectors, coefficients matrices, basis functions matrices, and shiftvectors in time and phase */
 int Read_Data_Mode(const char dir[], const int mode[2], EOBNRHMROMdata *data) {
[…]

### 2016-08-13 EOBNRv2HMROM/EOBNRv2HMROM.h

@@ -4,7 +4,7 @@
  * \brief C header for EOBNRv2HM reduced order model (non-spinning version).
  * See CQG 31 195010, 2014, arXiv:1402.4146 for details on the reduced order method.
  * See arXiv:1106.1021 for the EOBNRv2HM model.
- *  
+ *
  * Borrows from the SEOBNR ROM LAL code written by Michael Puerrer and John Veitch.
  *
  * Put the untared data in the directory designated by the environment variable ROM_DATA_PATH.
@@ -107,9 +107,21 @@ int EOBNRv2HMROMCore(
   double distance);
 
 int SimEOBNRv2HMROM(
-  struct tagListmodesCAmpPhaseFrequencySeries **listhlm,  /* Output: list of modes in Frequency-domain amplitude and phase form */
+  ListmodesCAmpPhaseFrequencySeries **listhlm,  /* Output: list of modes in Frequency-domain amplitude and phase form */
   int nbmode,                                    /* Number of modes to generate (starting with the 22) */
   double tRef,                                   /* Time shift with respect to the 22-fit removed waveform (s) */
+  double phiRef,                                 /* Phase at reference frequency */
+  double fRef,                                   /* Reference frequency (Hz); 0 defaults to fLow */
+  double m1SI,                                   /* Mass of companion 1 (kg) */
+  double m2SI,                                   /* Mass of companion 2 (kg) */
+  double distance);                              /* Distance of source (m) */
+
+int SimEOBNRv2HMROMExtTF2(
+  ListmodesCAmpPhaseFrequencySeries **listhlm,   /* Output: list of modes in Frequency-domain amplitude and phase form */
+  int nbmode,                                    /* Number of modes to generate (starting with the 22) */
+  double Mf_match,                               /* Minimum frequency using EOBNRv2HMROM in inverse total mass units*/
+  double minf,                                   /* Minimum frequency required */
+  double deltatRef,                              /* Time shift so that the peak of the 22 mode occurs at deltatRef */
   double phiRef,                                 /* Phase at reference frequency */
   double fRef,                                   /* Reference frequency (Hz); 0 defaults to fLow */
   double m1SI,                                   /* Mass of companion 1 (kg) */

### 2016-08-13 EOBNRv2HMROM/GenerateWaveform.c

@@ -43,6 +43,9 @@ Arguments are as follows:\n\
 --------------------------------------------------\n\
  --nbmode              Number of modes of radiation to generate (1-5, default=5)\n\
  --minf                Minimal frequency (Hz, default=0) - when too low, use first frequency covered by the ROM\n\
+ --deltatobs           Observation duration (years, default=2)\n\
+ --tagextpn            Tag to allow PN extension of the waveform at low frequencies (default=1)\n\
+ --Mfmatch             When PN extension allowed, geometric matching frequency: will use ROM above this value. If <=0, use ROM down to the lowest covered frequency (default=0.)\n\
  --taggenwave          Tag choosing the wf format: hlm (default: downsampled modes in Amp/Phase form), hphcFD (hlm interpolated and summed), hphcTD (IFFT of hphcFD),\n\
  --binaryout           Tag for outputting the data in gsl binary form instead of text (default false)\n\
  --outdir              Output directory\n\
@@ -63,6 +66,9 @@ Arguments are as follows:\n\
     /* set default values for the generation params */
     params->nbmode = 5;
     params->minf = 0.;
+    params->deltatobs = 2.;
+    params->tagextpn = 1;
+    params->Mfmatch = 0.;
     params->taggenwave = hlm;
     params->binaryout = 0;
     strcpy(params->outdir, ".");
@@ -91,6 +97,12 @@ Arguments are as follows:\n\
           params->nbmode = atoi(argv[++i]);
         } else if (strcmp(argv[i], "--minf") == 0) {
           params->minf = atof(argv[++i]);
+        } else if (strcmp(argv[i], "--deltatobs") == 0) {
+          params->deltatobs = atof(argv[++i]);
+        } else if (strcmp(argv[i], "--tagextpn") == 0) {
+          params->tagextpn = atoi(argv[++i]);
+        } else if (strcmp(argv[i], "--Mfmatch") == 0) {
+          params->Mfmatch = atof(argv[++i]);
         } else if (strcmp(argv[i], "--taggenwave") == 0) {
           params->taggenwave = ParseGenWavetag(argv[++i]);
         } else if (strcmp(argv[i], "--binaryout") == 0) {
[…]

### 2016-08-13 EOBNRv2HMROM/GenerateWaveform.h

@@ -58,6 +58,9 @@ typedef struct tagGenWaveParams {
   double distance;           /* distance of source (Mpc, default 1e3) */
   double inclination;        /* inclination of source (rad, default pi/3) */
   double minf;               /* Minimal frequency (Hz) - when set to 0 (default), use the first frequency covered by the ROM */
+  double deltatobs;          /* Observation duration (years, default=2) */
+  int tagextpn;              /* Tag to allow PN extension of the waveform at low frequencies */
+  double Mfmatch;            /* When PN extension allowed, geometric matching frequency: will use ROM above this value. If <=0, use ROM down to the lowest covered frequency */
   int nbmode;                /* number of modes to generate (starting with 22) - defaults to 5 (all modes) */
   int taggenwave;            /* Tag selecting the desired output format */
   int binaryout;             /* Tag for outputting the data in gsl binary form instead of text (default 0) */

### 2016-08-13 LISAinference/LISAutils.c

@@ -134,9 +134,10 @@ Arguments are as follows:\n\
 ----- Global Waveform/Inner products Parameters -----------------\n\
 -----------------------------------------------------------------\n\
  --fRef                Reference frequency where phiRef is set (Hz, default=0, interpreted as Mf=0.14)\n\
- --deltatobs           Observation time (years, default=2)\n\
+ --deltatobs           Observation duration (years, default=2)\n\
  --minf                Minimal frequency (Hz, default=0) - when set to 0, use the first frequency covered by the noise data of the detector\n\
- --mfmatch             Minimum matching frequency (1/mtot); if low, will use limit of ROM support., if <=0, then no extension.\n\
+ --tagextpn            Tag to allow PN extension of the waveform at low frequencies (default=1)\n\
+ --Mfmatch             When PN extension allowed, geometric matching frequency: will use ROM above this value. If <=0, use ROM down to the lowest covered frequency (default=0.)\n\
  --nbmodeinj           Number of modes of radiation to use for the injection (1-5, default=5)\n\
  --nbmodetemp          Number of modes of radiation to use for the templates (1-5, default=5)\n\
  --tagint              Tag choosing the integrator: 0 for Fresnel (default), 1 for linear integration\n\
@@ -211,7 +212,8 @@ Syntax: --PARAM-min\n\
     globalparams->fRef = 0.;
     globalparams->deltatobs = 2.;
     globalparams->minf = 0.;
-    globalparams->mfmatch = -1;
+    globalparams->tagextpn = 1;
+    globalparams->Mfmatch = 0.;
     globalparams->nbmodeinj = 5;
     globalparams->nbmodetemp = 5;
     globalparams->tagint = 0;
@@ -304,8 +306,10 @@ Syntax: --PARAM-min\n\
             globalparams->deltatobs = atof(argv[++i]);
         } else if (strcmp(argv[i], "--minf") == 0) {
             globalparams->minf = atof(argv[++i]);
-        } else if (strcmp(argv[i], "--mfmatch") == 0) {
-            globalparams->mfmatch = atof(argv[++i]);
[…]

### 2016-08-13 LISAinference/LISAutils.h

@@ -47,7 +47,8 @@ typedef struct tagLISAGlobalParams {
   double fRef;               /* reference frequency (Hz, default 0 which is interpreted as Mf=0.14) */
   double deltatobs;          /* max duration of observation (years, default 2) - the start of the signals might be cut in time instead of cut in frequency */
   double minf;               /* Minimal frequency (Hz) - when set to 0 (default), use the first frequency covered by the noise data of the detector */
-  double mfmatch;            /* Minimum matching frequency (1/mtot); if less than ROM model supports then will extrapolate ROM, if <=0, then no extension. */
+  int tagextpn;              /* Tag to allow PN extension of the waveform at low frequencies */
+  double Mfmatch;            /* When PN extension allowed, geometric matching frequency: will use ROM above this value. If <=0, use ROM down to the lowest covered frequency */
   int nbmodeinj;             /* number of modes to include in the injection (starting with 22) - defaults to 5 (all modes) */
   int nbmodetemp;            /* number of modes to include in the templates (starting with 22) - defaults to 5 (all modes) */
   int tagint;                /* Tag choosing the integrator: 0 for wip (default), 1 for linear integration */
@@ -169,6 +170,9 @@ int print_rescaleddist_to_file_LISA(
   LISAGlobalParams* globalparams,
   LISAPrior* prior,
   LISARunParams* run);
+/* Function printing injection/signal parameters to stdout */
+void report_LISAParams(
+  LISAParams* params);
 
 /* Initialization and clean-up for LISASignal structures */
 void LISASignalCAmpPhase_Cleanup(LISASignalCAmpPhase* signal);
@@ -224,10 +228,10 @@ void TaylorF2nonspin(
 		const int size,                         /** number of freq samples */
 		const double m1_SI,                     /**< mass of companion 1 (kg) */
 		const double m2_SI,                     /**< mass of companion 2 (kg) */
-		const double distance,                  /** distance (m) */   
-		const double imatch                     /**< index at which to match phase; 
+		const double distance,                  /** distance (m) */
+		const double imatch                     /**< index at which to match phase;
 							   assumes arrays are preloaded at imatch and imatch+1
-							   with the required result */ 
+							   with the required result */
 		     );
 
 /* checks prior boundaires */

### 2016-08-13 LISAsim/GenerateTDIFD.c

@@ -47,6 +47,9 @@ Arguments are as follows:\n\
 --------------------------------------------------\n\
  --nbmode              Number of modes of radiation to generate (1-5, default=5)\n\
  --minf                Minimal frequency (Hz, default=0) - when too low, use first frequency covered by the ROM\n\
+ --deltatobs           Observation duration (years, default=2)\n\
+ --tagextpn            Tag to allow PN extension of the waveform at low frequencies (default=1)\n\
+ --Mfmatch             When PN extension allowed, geometric matching frequency: will use ROM above this value. If <=0, use ROM down to the lowest covered frequency (default=0.)\n\
  --deltaf;             When generating frequency series from the mode contributions, deltaf for the output (0 to set automatically at 1/2*1/(2T))\n\
  --tagtdi              Tag choosing the set of TDI variables to use (default TDIAETXYZ)\n\
  --taggenwave          Tag choosing the wf format: TDIhlm (default: downsampled mode contributions to TDI in Amp/Phase form), TDIFD (hlm interpolated and summed accross modes)\n\
@@ -78,6 +81,9 @@ Arguments are as follows:\n\
     /* Set default values for the generation params */
     params->nbmode = 5;
     params->minf = 0.;
+    params->deltatobs = 2.;
+    params->tagextpn = 1;
+    params->Mfmatch = 0.;
     params->deltaf = 0.;
     params->tagtdi = TDIXYZ;
     params->taggenwave = TDIhlm;
@@ -118,6 +124,12 @@ Arguments are as follows:\n\
             params->nbmode = atoi(argv[++i]);
         } else if (strcmp(argv[i], "--minf") == 0) {
             params->minf = atof(argv[++i]);
+        } else if (strcmp(argv[i], "--deltatobs") == 0) {
+            params->deltatobs = atof(argv[++i]);
+        }  else if (strcmp(argv[i], "--tagextpn") == 0) {
+            params->tagextpn = atoi(argv[++i]);
+        } else if (strcmp(argv[i], "--Mfmatch") == 0) {
+            params->Mfmatch = atof(argv[++i]);
         } else if (strcmp(argv[i], "--deltaf") == 0) {
             params->deltaf = atof(argv[++i]);
[…]

### 2016-08-13 LISAsim/GenerateTDIFD.h

@@ -62,6 +62,9 @@ typedef struct tagGenTDIFDparams {
 
   int nbmode;                /* number of modes to generate (starting with 22) - defaults to 5 (all modes) */
   double minf;               /* Minimal frequency (Hz) - when set to 0 (default), use the first frequency covered by the ROM */
+  double deltatobs;          /* Observation duration (years, default=2) */
+  int tagextpn;              /* Tag to allow PN extension of the waveform at low frequencies (default=1) */
+  double Mfmatch;            /* When PN extension allowed, geometric matching frequency: will use ROM above this value. If <=0, use ROM down to the lowest covered frequency (default=0.) */
   double deltaf;             /* When generating frequency series from the mode contributions, deltaf for the output (0 to set automatically at 1/2*1/(2T)) */
   int tagtdi;                /* Tag selecting the desired output format */
   int taggenwave;            /* Tag selecting the desired output format */

### 2016-08-13 LISAsim/GenerateTDITD.c

@@ -167,11 +167,19 @@ int main(int argc, char *argv[])
   gsl_spline_init(spline_hp, gsl_vector_const_ptr(hptd->times,0), gsl_vector_const_ptr(hptd->h,0), nbtimes);
   gsl_spline_init(spline_hc, gsl_vector_const_ptr(hctd->times,0), gsl_vector_const_ptr(hctd->h,0), nbtimes);
 
+  //
+  //printf("(%g, %g)\n", gsl_vector_get(times, 0), gsl_vector_get(times, 200));
+  //printf("(%g, %g)\n", gsl_vector_get(times, times->size-1), gsl_vector_get(times, times->size-1-200));
+
   /* Evaluate TDI */
   RealTimeSeries* TDI1 = NULL;
   RealTimeSeries* TDI2 = NULL;
   RealTimeSeries* TDI3 = NULL;
-  int nptmargin = 200; /* Here, hardcoded margin that we will set to 0 on both sides, to avoid problems with delays extending past the input values */
+  /* Here, hardcoded margin that we will set to 0 on both sides, to avoid problems with delays extending past the input values */
+  /* We take R/c as the maximum delay that can occur in the response, and adjust the margin accordingly */
+  double maxdelay = R_SI/C_SI;
+  double deltat = gsl_vector_get(times, 1) - gsl_vector_get(times, 0);
+  int nptmargin = 2 * (int)(maxdelay/deltat);
   GenerateTDITD3Chan(&TDI1, &TDI2, &TDI3, spline_hp, spline_hc, accel_hp, accel_hc, times, nptmargin, params->tagtdi);
 
   /* Output */

### 2016-08-13 LISAsim/LISAFDresponse.c

@@ -103,7 +103,7 @@ int LISASimFDResponse21(
       gsl_vector_set(amp_imag, j, cimag(camp));
       gsl_vector_set(besselphi, j, bphi);
     }
-    
+
     /* Second step of the processing: constellation delay/modulation */
     /* Initializing spline for the bessel phase */
     gsl_spline* spline_besselphi = gsl_spline_alloc(gsl_interp_cspline, len);
@@ -117,7 +117,7 @@ int LISASimFDResponse21(
       /**/
       gsl_vector_set(amp_real, j, creal(camp));
       gsl_vector_set(amp_imag, j, cimag(camp));
-    } 
+    }
 
     listelement = listelement->next;
 
@@ -188,7 +188,7 @@ int LISASimFDResponsey12(
     gsl_spline* spline_phi = gsl_spline_alloc(gsl_interp_cspline, len);
     gsl_interp_accel* accel_phi = gsl_interp_accel_alloc();
     gsl_spline_init(spline_phi, gsl_vector_const_ptr(freq, 0), gsl_vector_const_ptr(phase, 0), len);
-    
+
     /* Orbital delay, constellation delay/modulation */
     /* Initializing frequency series structure for this mode, for each of the TDI observables */
     CAmpPhaseFrequencySeries *modefreqseries = NULL;
@@ -204,7 +204,7 @@ int LISASimFDResponsey12(
       f = gsl_vector_get(freq, j);
       tf = (gsl_spline_eval_deriv(spline_phi, f, accel_phi))/(2*PI);
       //clock_t tbegGAB = clock();
-      EvaluateGABmode(&g12mode, &g21mode, &g23mode, &g32mode, &g31mode, &g13mode, f, tf, Yfactorplus, Yfactorcross);
+      EvaluateGABmode(&g12mode, &g21mode, &g23mode, &g32mode, &g31mode, &g13mode, f, tf, Yfactorplus, Yfactorcross, 1);
       //clock_t tendGAB = clock();
       //timingcumulativeGABmode += (double) (tendGAB-tbegGAB) /CLOCKS_PER_SEC;
       /**/
@@ -220,7 +220,7 @@ int LISASimFDResponsey12(
     /* Copying the vectors of frequencies and phases */
     gsl_vector_memcpy(freq_y, freq);
     gsl_vector_memcpy(phase_y, phase);
-    
+
     /* Append the modes to the ouput list-of-modes structures */
[…]

### 2016-08-13 LISAsim/LISAgeometry.c

@@ -93,7 +93,8 @@ static double sinarray[4];
 /* Function to convert string input TDI string to TDItag */
 TDItag ParseTDItag(char* string) {
   TDItag tag;
-  if(strcmp(string, "TDIXYZ")==0) tag = TDIXYZ;
+  if(strcmp(string, "y12")==0) tag = y12;
+  else if(strcmp(string, "TDIXYZ")==0) tag = TDIXYZ;
   else if(strcmp(string, "TDIalphabetagamma")==0) tag = TDIalphabetagamma;
   else if(strcmp(string, "TDIAETXYZ")==0) tag = TDIAETXYZ;
   else if(strcmp(string, "TDIAETalphabetagamma")==0) tag = TDIAETalphabetagamma;
@@ -517,7 +518,8 @@ int EvaluateGABmode(
   const double f,                          /* Frequency */
   const double t,                          /* Time */
   const double complex Yfactorplus,        /* Spin-weighted spherical harmonic factor for plus */
-  const double complex Yfactorcross)       /* Spin-weighted spherical harmonic factor for cross */
+  const double complex Yfactorcross,       /* Spin-weighted spherical harmonic factor for cross */
+  int tagdelayR)                           /* Tag: when 1, include the phase term of the R-delay */
 {
   /* Precompute array of sine/cosine */
   for(int j=0; j<4; j++) {
@@ -565,7 +567,10 @@ int EvaluateGABmode(
   double complex factorcexp12 = cexp(I*prefactor * (1.+kp1plusp2));
   double complex factorcexp23 = cexp(I*prefactor * (1.+kp2plusp3));
   double complex factorcexp31 = cexp(I*prefactor * (1.+kp3plusp1));
-  double complex factorcexpkR = cexp(I*prefactorR * kR);
+  /* The tag tagdelayR aloows choose to include or not the R-delay phase term (here leading order) */
+  double complex factorcexpkR;
+  if(tagdelayR) factorcexpkR = cexp(I*prefactorR * kR);
+  else factorcexpkR = 1.;
   /* Output result */
   *G12 = I*prefactor * factorcexpkR * factn3Pn3 * sinc( prefactor * (1.-kn3)) * factorcexp12;
[…]

### 2016-08-13 LISAsim/LISAgeometry.h

@@ -47,6 +47,7 @@ extern "C" {
 /**************** TDI variables **********************/
 
 typedef enum TDItag {
+  y12,
   TDIXYZ,
   TDIalphabetagamma,
   TDIAETXYZ,
@@ -90,7 +91,8 @@ int EvaluateGABmode(
   const double f,                          /* Frequency */
   const double t,                          /* Time */
   const double complex Yfactorplus,        /* Spin-weighted spherical harmonic factor for plus */
-  const double complex Yfactorcross);      /* Spin-weighted spherical harmonic factor for cross */
+  const double complex Yfactorcross,       /* Spin-weighted spherical harmonic factor for cross */
+  int tagdelayR);                          /* Tag: when 1, include the phase term of the R-delay */
 
 /* Functions evaluating the Fourier-domain factors (combinations of the GAB's) for TDI observables */
 /* NOTE: factors have been scaled out, in parallel of what is done for the noise function */

### 2016-08-13 tools/likelihood.c

@@ -48,15 +48,6 @@
 
 /********************************* Utilities ****************************************/
 
-/* Newtonian estimate of the relation Mf(deltat/M) (for the 22 mode) - gives the starting geometric frequency for a given mass ratio and a given geometric duration of the observations */
-double NewtonianfoftGeom(
-  const double q,                      /* Mass ratio m1/m2 */
-  const double t)                      /* Duration of the observations in geometric units (t/M) */
-{
-  double nu = q/(1.+q)/(1.+q);
-  return 1./PI * pow(256*nu/5. * t, -3./8);
-}
-
 /* Function to evaluate a Noise function  */
 void EvaluateNoise(
   gsl_vector* noisevalues,                         /* Output: vector of the noise values */

### 2016-08-13 tools/likelihood.h

@@ -48,11 +48,6 @@ extern "C" {
 
 /****** Prototypes: utilities *******/
 
-/* Newtonian estimate of the relation Mf(deltat/M) (for the 22 mode) - gives the starting geometric frequency for a given mass ratio and a given geometric duration of the observations */
-double NewtonianfoftGeom(
-  const double q,                      /* Mass ratio m1/m2 */
-  const double t);                     /* Duration of the observations in geometric units (t/M) */
-
 /* Function to evaluate a Noise function  */
 void EvaluateNoise(
   gsl_vector* noisevalues,                         /* Output: vector of the noise values */

### 2016-08-13 tools/waveform.c

@@ -17,6 +17,30 @@
 
 /* NOTE: uses the list of modes of EOBNRv2HMROM (listmode), to be extended when more waveform models are added */
 
+/***************** Function estimating frequency corresponding to a given time to coalescence ****************/
+
+/* Newtonian estimate of the relation Mf(deltat/M) (for the 22 mode) - gives the starting geometric frequency for a given mass ratio and a given geometric duration of the observations */
+double NewtonianfoftGeom(
+  const double q,                      /* Mass ratio m1/m2 */
+  const double t)                      /* Duration of the observations in geometric units (t/M) */
+{
+  if(t<=0.) return 0.;
+  double nu = q/(1.+q)/(1.+q);
+  return 1./PI * pow(256*nu/5. * t, -3./8);
+}
+
+/* Newtonian estimate of the relation f(deltat) (for the 22 mode) - gives the starting geometric frequency for a given mass ratio and a given geometric duration of the observations - output in Hz */
+double Newtonianfoft(
+  const double m1,                     /* Mass 1 (solar masses) */
+  const double m2,                     /* Mass 2 (solar masses) */
+  const double t)                      /* Duration of the observations in years */
+{
+  if(t<=0.) return 0.;
+  double mtot = m1 + m2;
+  double q = m1/m2;
+  return NewtonianfoftGeom(q, t*YRSID_SI / (mtot*MTSUN_SI)) / (mtot*MTSUN_SI);
+}
+
 /***************** Function estimating time to coalescence ****************/
 
 /* Functions reading from a list of modes the minimal and maximal frequencies */
@@ -367,6 +391,9 @@ int GenerateFDReImFrequencySeries(
     nbpts = (int) ceil(fHigh/deltaf);
   }
   else nbpts = nbpt;
+
+  //
+  printf("nbpt in GenerateFDReImFrequencySeries: %d\n", nbpts);
 
   /* Initialize frequency series */
   ReImFrequencySeries_Init(freqseries, nbpts);

### 2016-08-13 tools/waveform.h

@@ -47,6 +47,14 @@ extern "C" {
 
 /* NOTE: uses the list of modes of EOBNRv2HMROM (listmode), to be extended when more waveform models are added */
 
+/***************** Function estimating frequency corresponding to a given time to coalescence ****************/
+
+/* Newtonian estimate of the relation Mf(deltat/M) (for the 22 mode) - gives the starting geometric frequency for a given mass ratio and a given geometric duration of the observations */
+double NewtonianfoftGeom(const double q, const double t); /* t here is t/M */
+
+/* Newtonian estimate of the relation f(deltat) (for the 22 mode) - gives the starting frequency in Hz for a given mass ratio and a given geometric duration of the observations */
+double Newtonianfoft(const double m1, const double m2, const double t); /* t here is in years, m1-m2 in solar masses */
+
 /***************** Function estimating time to coalescence and min/max frequency ****************/
 
 /* Functions reading from a list of modes the minimal and maximal frequencies */

### 2016-11-10 docs/paper_method.tex
Various modifications.
- added support for maxf along with minf
- added support for TD amp/phase representation (incl. phase unwrapping tool) -- still work in progress
- added support for TD response, orbital delay and y12L -- still work in progress
- update of the paper

--- 2016-07-02 docs/paper_method.tex
@@
 \newcommand\calO{{\mathcal{O}}}
+\newcommand\calM{{\mathcal{M}}}
 \newcommand\calF{{\mathcal{F}}}
@@
 \newcommand{\Msol}{M_{\odot}}
-\newcommand{\tf}{\tilde{t}_{f}}
-\newcommand{\Tf}{\tilde{T}_{f}}
+\newcommand{\tf}{t_{f}}
+\newcommand{\Tf}{T_{f}}
+\newcommand{\tfd}{t_{f}^{d}}
 
@@
 
-\title{Modulation and delays in Fourier domain: precessing binaries and LISA-type detector response}
+\title{Modulations and delays in Fourier domain: precessing binaries and LISA-type detector response}
 
@@
 \author{Sylvain Marsat}
-\affiliation{Department of Physics, University of Maryland, College Park, MD 20742, USA}
-\affiliation{Gravitational Astrophysics Laboratory, NASA Goddard Space Flight Center, 8800 Greenbelt Rd., Greenbelt, MD 20771, USA}
 \affiliation{Max Planck Institute for Gravitational Physics (Albert Einstein Institute), Am M\"uhlenberg 1, Potsdam-Golm, 14476, Germany}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\section{Methodology}
-\label{sec:methodology}
+\section{Fourier-domain waveforms and separation of timescales}
+\label{sec:motivation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Waveforms in the Fourier domain and stationary phase approximation}
+\label{subsec:SPA}
+
+The convention we will be using for the Fourier transform of a signal $h(t)$ is
+\be\label{eq:defFT}
+	\tilde{h}(f) =  \int \ud t \, e^{+2i\pi f t} h(t) \,.
+\ee
+Notice that this sign convention is not the most frequently used in the literature.
+We chose it to ensure that, with the conventions of~\cite{BlanchetLiving}, spin-weighted spherical modes $h_{\ell m}$ with $m>0$ will have support mostly for positive frequencies.
+One can revert to the more usual convention by taking $f\rightarrow -f$.
+
+The gravitational waveform emitted in the direction $(\Theta, \Phi)$, with its two polarizations $h_{+},h_{\times}$, can be decomposed in spin-weighted spherical modes as~\cite{Thorne80}
+\be\label{eq:defmodes}
+	h_{+} - i h_{\times} = \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{+\ell} {}_{-2}Y_{\ell m}(\Theta,\Phi) h_{\ell m} \,,
+\ee
+where the ${}_{-2}Y_{\ell m}(\Theta,\Phi)$ are spin-weighted spherical harmonics.
+Since we will work in the Fourier domain, it will be useful to write the contributions of the individual modes to the Fourier transforms of the polarizations as
+\begin{subequations}\label{eq:hpcfrommodes}
+\begin{align}
+	\tilde{h}_{+}(f) &= \frac{1}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{+\ell} \left( {}_{-2}Y_{\ell m} \widetilde{h_{\ell m}}(f) + {}_{-2}Y_{\ell m}^{*} \widetilde{h_{\ell m}}(-f)^{*} \right) \,, \\
+	\tilde{h}_{\times}(f) &= \frac{i}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{+\ell} \left( {}_{-2}Y_{\ell m} \widetilde{h_{\ell m}}(f) - {}_{-2}Y_{\ell m}^{*} \widetilde{h_{\ell m}}(-f)^{*} \right) \,,
+\end{align}
+\end{subequations}
+where we used $\widetilde{h_{\ell m}^{*}}(f) = \widetilde{h_{\ell m}}(-f)^{*}$.
+In the following, we will denote generically by $h$ an individual mode of the waveform, dropping the mode indices $\ell m$, and we will often refer to the Fourier-domain amplitude $A$ and phase $\Psi$ of the mode, according to
+\be
+	\tilde{h}(f) \equiv A(f) e^{-i\Psi(f)} \,.
+\ee
+We recall that, with our sign convention, the effect of a positive shift in time $h_{\Delta t} (t) = h(t- \Delta t)$ is
+\be
+	\tilde{h}_{\Delta t}(f) = e^{2i\pi f \Delta t} \tilde{h}(f) \,.
+\ee
+If we also allow for a global shift in phase of the signal, which corresponds to a rotation of the unit vector pointing towards the observer in the frame of the gravitational waves source, a constant term can also be added (note that the effect of such a rotation gives different phase changes $e^{\pm i m \delta \phi}$ for the different modes of the radiation).
+We see that both a constant and a linear term in the Fourier-domain phase can be adjusted by changing extrinsic parameters.
+Thus, as is well known, apart from the relative phases of the modes for waveforms that include higher harmonics, the intrinsic information on the phasing is contained in the second derivative of the Fourier-domain phase $\Psi(f)$.
+
+We recall below an approximation widely used to model gravitational waves signals emitted by inspiralling binaries, the stationary phase approximation (thereafter SPA).
+It applies in general to chirping signals, and we refer the reader to~\ref{} for details.
+One first writes the time-domain signal in amplitude and phase form as $h(t) = a(t) e^{-2i\varphi(t)}$, where for gravitational wave signals $\varphi$ will correspond to the orbital phase, with the orbital frequency being $\omega = \dot{\varphi}$.
+In order to keep close to the notations used in the gravitational wave literature, we introduced a factor of 2 in the phase of the wave, which is appropriate for the dominant 22 harmonic of the signal.
+The approximation then applies to signals verifying the conditions
+%
+\be\label{eq:conditionSPA}
+	\left| \frac{\dot{a}/a}{\omega} \right| \ll 1\,, \quad \left|\frac{\dot{\omega}}{\omega^{2}} \right| \ll 1\,, \quad \left| \frac{(\dot{a}/a)^{2}}{\dot{\omega}} \right| \ll 1 \,.
+\ee
+%
+Since the integral~\eqref{eq:defFT} defining the Fourier transform is rapidly oscillatory unless the term $2\pi f t$ cancels the evolution of $-2\varphi(t)$, its support is well centered around the point of stationary phase.
+This determines the time-to-frequency correspondence in the stationary phase approximation, and leads to the definition of this time as an implicit function of frequency by the relation
+%
+\be\label{eq:deftfSPA}
+	\omega(t_{f}^{\rm SPA}) = \pi  f \,.
+\ee
+% 
+For a chirping signal of increasing phase, $\omega>0$ and $\dot{\omega}>0$, there is a unique point of stationary phase, located in the positive frequency range $f>0$.
+Using the conditions above, one can formally expand the signal around $t_{f}^{\rm SPA}$ to quadratic order in time, according to
+\be
+	\tilde{h}_{\rm SPA} (f) \simeq a(t_{f}^{\rm SPA}) \exp\left[2i\pi f t_{f}^{\rm SPA}-2i\varphi(t_{f}^{\rm SPA}) \right]  \int \ud t \, e^{-i \dot{\omega} (t_{f}^{\rm SPA}) (t-t_{f}^{\rm SPA})^{2}} \,.
+\ee
+Note that to be able to treat the amplitude as a constant in the integral above, we used the third condition in~\eqref{eq:conditionSPA}.
+The resulting complex Gaussian integral yields\footnote{Note that the expressions given here are valid for the $22$ mode of the waveform.}
+\begin{subequations}
+\begin{align}
+	\tilde{h}_{\rm SPA}(f) &= A_{\rm SPA}(f) e^{-i\Psi_{\rm SPA}(f)} \,, \\
+	A_{\rm SPA}(f) &= a(t_{f}^{\rm SPA}) \sqrt{\frac{\pi}{\dot{\omega}(t_{f}^{\rm SPA})}} \,, \\
+	\Psi_{\rm SPA}(f) &= 2\varphi(t_{f}^{\rm SPA}) - 2\pi f t_{f}^{\rm SPA} + \frac{\pi}{4} \,.
+\end{align}
+\end{subequations}
+Ref.~\cite{Droz+99} evaluated the first correction to this approximation, within the context of post-Newtonian signals, and found that it can be considered as a term of the fifth post-Newtonian order, beyond the accuracy level of our current best models~\cite{BlanchetLiving}.
+
+To understand the separation of timescales in the problem, it will be useful to have at hand the leading-order scaling laws for an inspiral (at the Newtonian order in the PN language).
+Although inaccurate for the purpose of waveform modelling, these Newtonian estimates will give a good qualitative description of the order of magnitude of the relevant timescales in the inspiral.
+For a binary with masses $m_{1}, m_{2}$, we define the total mass $m=m_{1}+m_{2}$ and the symmetric mass ratio $\nu = m_{1}m_{2}/m^{2}$.
+Introducing a time of coalescence $t_{c}$, the relations between the orbital frequency and phase and the time to coalescence $t_{c} - t$ are then given by
+\begin{subequations}
+\begin{align}
+	\omega(t) &= \left[ \frac{256\nu}{5c^{5}} (Gm)^{5/3} (t_{c}-t) \right]^{-3/8} \,, \\
+	\varphi(t) &= -\left[ \frac{c^{3}}{G m \nu^{3/5}} (t_{c}-t) \right]^{5/8} \,.
+\end{align}
+\end{subequations}
+For this Newtonian inspiral, applying the SPA gives for the $22$ mode $\tilde{h}(f) = A_{\rm N}(f)e^{-i\Psi_{\rm N}(f)}$ with
+%
+\begin{subequations}
+\begin{align}
+	A_{N}(f) &= \frac{G^{2}m^{2} \pi}{Dc^{5}} \sqrt{\frac{2\nu}{3}} v^{-7/2}\,,\\
+	\Psi_{\rm SPA}^{\rm N}(f) &= 2\pi f t_{0} - \phi_{0} + \frac{3}{128\nu v^{5}} \,, 
+\end{align}
+\end{subequations}
+%
+where $v=(G m \pi f/c^{3})^{1/3}$, $D$ is the luminosity distance to the observer, and where $t_{0}, \phi_{0}$ are constants.
+It is also customary to rewrite the above relations in terms of the chirp mass $\calM \equiv m\nu^{3/5}$, which is the only mass combination arising at the leading PN order.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Instrumental modulations and delays for LISA-type detectors}
+\label{subsec:modulationLISA}
+
+The response of a detector of the LISA type to an incident gravitational wave can be written in two different, equivalent forms, in terms of phase or frequency measurements.
+Here we will work with the second representation, which will prove more convenient for our purposes.
+Moreover, various notations and conventions have been used in the literature to label the spacecrafts and describe their orbits.
+We refer the reader to~\cite{Vallisneri04} for a comparative account on these various conventions.
+In this work, we will use the conventions of~\cite{Vallisneri04}, which were also used in the Mock LISA Data Challenges (MLDC)~\cite{}.
+
+We use a coordinate system centered on the solar system barycenter (SSB), and we denote the center of the constellation by $O$.
+We introduce the notation $\hat{k}$ for the propagation vector of the gravitational wave, which we denote by $h_{ij}^{\rm TT}(t)$ in transverse-traceless matrix form, as measured at the SSB (thus, at position $p$, $h_{ij}^{\rm TT}(t, p) = h_{ij}^{\rm TT}(t - \hat{k}.p)$).
+We denote by $p_{A}$ ($A=1,2,3$) the position of the spacecrafts and $n_{l}$ the unit vectors of the three links, with the convention that $n_{3}$ points from 1 to 2.
+
+Written in terms of the fractional laser frequency shifts between two spacecrafts, the elementary response of the detector reads~\cite{EW75, RCP04, Vallisneri04}
+\be\label{eq:yslr}
+	y_{slr} \equiv \frac{\nu_{r} - \nu_{s}}{\nu} = \frac{1}{2} \frac{n_{l}^{i}n_{l}^{j}}{1 - \hat{k}\cdot n_{l}} \left[ h_{ij}^{\rm TT}(t - \hat{k}\cdot p_{s}) - h_{ij}^{\rm TT}(t - \hat{k}\cdot p_{r}) \right]
+\ee
+As detailed below in Sec.~\ref{}, we will use a rigid instantaneous model, approximating the geometry of the constellation by a moving equilateral triangle (neglecting the flexing of the arms induced by corrections in the orbits), and evaluating all geometric factors in~\eqref{eq:yslr} at a single time $t$ (neglecting pointing-ahead corrections).
+
+The first- and second-generation TDI observables are then built as combinations of these basic building blocks, evaluated at delayed times.
+Since these delays will take a simple form in Fourier domain in our approximation, we will focus on the $y_{slr}$ observables.
+We will give in Sec.~\ref{subsec:modelLISA} more details on the response of LISA-type detectors, on the TDI observables, and on the approximations that enter the derivation of the basic response~\eqref{eq:yslr} above.
+
+The structure of~\eqref{eq:yslr} is such that one can build the full signal from individual contributions of the form
+\be
+	s(t) = F(t) h(t + d(t))
+\ee
+Here, $h(t)$ represents one of the individual modes building the full gravitational wave signal (see Eq.~\eqref{eq:defmodes}), $d(t)$ represents the time-varying delays of the form $-\hat{k}\cdot p_{A}(t)$, and $F(t)$ incorporates all the relevant geometric prefactors.
+
+The functions $F(t)$ and $d(t)$ vary on a timescale of one year.
+When neglecting small non-periodic orbital perturbations (such as those caused by the influence of the Earth and of the other planets on the orbits), both $F(t)$ and $d(t)$ are periodic.
+It will also be useful to separate the delays in two types of terms: the first, $d_{O} = -\hat{k}\cdot p_{O}$, relates the waveform at the SSB to the waveform at the center of the constellation $O$, whereas the second, $d_{c}$, represents the various delays between the spacecrafts of the constellation.
+For an orbit radius $R=1\,\mathrm{au}$ and an armlength $L=5.10^{6}\mathrm{km}$, $d_{O} \sim R/c \sim 500\mathrm{s}$ and $d_{c} \sim L/c \sim 17\mathrm{s}$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Precession modulation for spinning binaries}
+\label{subsec:modulationPrec}
+
+For spinning compact objects, the presence of the spins leads to the precession of the orbit~\cite{Apostolatos+94, Kidder95}, which can have large effects on the waveform.
+In particular, it breaks the planar symmetry of the gravitational wave emission, and leads to amplitude modulations that are especially important for edge-on observers.
+
+It was suggested by a number of authors~\cite{} that the effect of the precession can be modeled, to a good approximation, by a time-dependent rotation of a non-precessing waveform.
+In this approach, one introduces a precessing frame following the evolution of the plane of the orbit, and approximates the waveform in this precessing frame by using a non-precessing model.
+Different prescriptions have been proposed for the construction of the precessing frame itself~\cite{}.
+
+If $(\alpha, \beta, \gamma)$ are the Euler angles relating the precessing frame to the inertial frame in the $(z,y,z)$ convention, the modes in the inertial frame $h_{\ell m}^{\rm I}$ are then related to the modes in the precessing frame $h_{\ell m}^{\rm P}$ by~\cite{Goldberg+67}
+\be
+	h_{\ell m}^{\rm I} = \sum\limits_{m=-\ell}^{\ell} D^{\ell}_{mm'} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm P} \,,
+\ee
+where the $D^{\ell}_{mm'}$ are Wigner matrices~\cite{}. Notice that there is no mixing of the modes with different values of $\ell$.
+
+If we make the assumption that the precessing-frame waveform $h^{\rm P}$ is approximated by a non-precessing model that provides us with a smooth Fourier-domain amplitude and phase, then the problem reduces to computing the Fourier transform of the signal
+\be\label{eq:defmodulationprec}
+	s(t) = F(t) h(t) \,,
+\ee
+where the modulation function $F(t)$ is given by a Wigner matrix and depends on time through the Euler angles $(\alpha, \beta, \gamma)(t)$.
+We will discuss more in Sec.~\ref{subsec:modelPrec} the limitations of the assumptions used to perform the decomposition above.
+
+In Eq.~\eqref{eq:defmodulationprec} above, the modulation function $F$ evolves on the precessional timescale, which evolves throughout the inspiral.
+The separation between the precessional timescale and the orbital timescale will be crucial for our analysis.
+In the limit of low frequencies, we will see in Sec.~\ref{} below that, although these timescales become more and more separated, the decrease in the chirping rate gives raise to a corrective contribution that does not vanish in this limit.
+In the other limit, as the system gets close to merger, the separation of timescales becomes weaker, and we will explore in Sec.~\ref{} the application of our formalism to the merger and ringdown phase.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Fourier-domain approach to modulations and delays}
+\label{sec:formalism}
 
@@
 Let us consider, in general, the problem of determining the Fourier transform of a signal $h(t)$ to which we apply a varying delay $d(t)$ and a multiplication by a modulation function $F(t)$.
-In the case of precessing binaries, the delays will ba absent, and the modulation functions will be the time-dependent Wigner coefficients applied to rotate the waveform from a precessing frame, where it verifies a restored mode hierarchy, to an inertial frame where the observations take place.
+In the case of precessing binaries, the delays will be absent, and the modulation functions will be the time-dependent Wigner coefficients applied to rotate the waveform from a precessing frame, where it verifies a restored mode hierarchy, to an inertial frame where the observations take place.
 In the case of a LISA-type detector, the signal $h(t)$ will simply be the waveform in a fixed heliocentric frame, and the delays will come from the motion of each detector in the wave front, while the modulation will represent the time-variation in the detector orientation.
 Both can be considered as periodic functions with a period of a year, which allows for an additional rewriting in terms of Fourier series instead of Fourier transforms, as will be developed in Sec.~\ref{}.
-
-The convention we will be using for Fourier transform will be
-%
-\be\label{eq:defFT}
-	\tilde{h}(f) =  \int \ud t \, e^{+2i\pi f t} h(t) \,.
-\ee
-%
-Notice that this convention uses the opposite sign with respect to the one used in general.
-It was chosen to ensure that, with the conventions of~\cite{}, modes $h_{\ell m}$ with $m>0$ will have support for positive frequencies.
-One can revert to the more usual convention by taking $f\rightarrow -f$.
 
@@
 %
-\begin{subequations}
-\begin{align}\label{eq:defg}
+\begin{subequations}\label{eq:defg}
+\begin{align}
 	g(f,t) &= e^{-2i\pi f d(t)} F(t) \,, \\
@@
 Here and in the rest of the paper, derivatives with respect to $f$ will always be evaluated at $f$.
+Notice that, because of the freedom of adding a linear term to $\Psi(f)$ by simply shifting the signal in time, no assumption can be made on the smallness of the first derivative of the phase, and this is really a leading order approximation.
 
@@
 %
-which motivates the introduction of the effective time
+which motivates the introduction of the effective time, seen as a function of frequency,
 %
@@
 %
-where the tilde here does not indicate  a Fourier transform operation but simply that this quantity is an effective time defined from Fourier-domain waveform.
+where the tilde here does not indicate  a Fourier transform operation but simply that this quantity is an effective time defined from the Fourier-domain waveform.
 This definition is a generalization of the time-to-frequency correspondence at the heart of the SPA, as will be shown in Sec.~\ref{}, but it is worth noting the it is defined entirely from the Fourier-domain waveform.
@@
 If we further expand the exponentials to obtain a pure $f'$-expansion, by applying the derivative rule
-%
 \be
@@
 \ee
-%
 we arrive at the formal expansion
-%
 \begin{align}
-	\tilde{s}(f) &= \tilde{h}(f) \sum\limits_{q\geq 0} \sum\limits_{r\geq 0} \sum\limits_{k_{2}\geq 0} \dots \sum\limits_{k_{N}\geq 0} \frac{(-1)^{q}}{q!} \frac{(-1)^{r}}{r!} \prod\limits_{p=2}^{N} \frac{1}{k_{p}!}\left( \frac{(-1)^{p}(-i)}{p!} \frac{\ud^{p}\Psi}{\ud f^{p}}\right)^{k_{p}} \frac{1}{A} \frac{\ud^{q} A}{\ud f ^{q}} \frac{1}{(-2i\pi)^m} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{m} }{\partial t^{m}} g \right)(f,\tf) \,, \nn\\
-	\text{where } & m = p+r+2k_{2}+\dots+N k_{N} \,.
+	\tilde{s}(f) &= \tilde{h}(f) \sum\limits_{q\geq 0} \sum\limits_{r\geq 0} \sum\limits_{k_{2}\geq 0} \dots \sum\limits_{k_{N}\geq 0} \frac{(-1)^{q}}{q!} \frac{(-1)^{r}}{r!} \prod\limits_{p=2}^{N} \frac{1}{k_{p}!}\left( \frac{(-1)^{p}(-i)}{p!} \frac{\ud^{p}\Psi}{\ud f^{p}}\right)^{k_{p}} \frac{1}{A} \frac{\ud^{q} A}{\ud f ^{q}} \frac{1}{(-2i\pi)^m} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{m} }{\partial t^{m}} g \right)(f,\tf) \,,
 \end{align}
-%
+where $m = p+r+2k_{2}+\dots+N k_{N}$.
 
-This rather cumbersome expression is easy to simplify for special cases. For instance, if we expand only the frequency-dependence of $g$ as in~\eqref{eq:expandg}, we obtain
-%
-\be
+This rather cumbersome expression will not be used as such, but is easy to simplify for special cases.
+For instance, if we expand only the frequency-dependence of $g$ as in~\eqref{eq:expandg}, we obtain
+\be\label{eq:taylordelay}
 	\tilde{s}(f) = \tilde{h}(f) \sum\limits_{r\geq 0} \frac{1}{(2i\pi)^{r}r!} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{r} }{\partial t^{r}} g \right)(f,\tf)
 \ee
-%
 which interestingly looks like a Taylor expansion of $g$, but with joint derivatives in frequency and time. Similarly, the expansion~\eqref{eq:expandA} gives
-%
-\be
+\be\label{eq:resultdfA}
 	\tilde{s}(f) = \tilde{h}(f) \sum\limits_{q\geq 0} \frac{1}{(2i\pi)^{q}q!} \frac{1}{A} \frac{\ud^{q} A}{\ud f ^{q}}  \left( \frac{\partial^{q} }{\partial t^{q}} g \right)(f,\tf)
 \ee
-%
 
 Of particular interest for us will be the case where we keep only the first term in~\eqref{eq:expandPsi}, and expand the resulting exponential.
-Indeed, this term will be  the dominant correction beyond leading order in the applications that we will consider in Secs.~\ref{} and~\ref{}, and this particular approximation will turn out to be closely related to the SPA approach of~\cite{}.
+Indeed, this term will be  the dominant correction beyond leading order in the applications that we will consider in Secs.~\ref{} and~\ref{}, and this particular approximation will turn out to be closely related to the approach of~\cite{Klein+14}.
 The above approach yields
-%
 \be\label{eq:resultdffPsi}
-	\tilde{s}(f) = \tilde{h}(f) \sum\limits_{p\geq 0} \frac{1}{p!} \left( \frac{i}{8\pi^{2}}\frac{\ud^{2} \Psi}{\ud f^{2}} \right)^{p} \left( \frac{\partial^{2p} }{\partial t^{2p}} g \right)(f,\tf) \,,
+	\tilde{s}(f) = \tilde{h}(f) \sum\limits_{p\geq 0} \frac{1}{p!} \left( \frac{i}{8\pi^{2}}\frac{\ud^{2} \Psi}{\ud f^{2}} \right)^{p} \left( \frac{\partial^{2p} }{\partial t^{2p}} g \right)(f, \tf) \,,
 \ee
-%
 and it is natural to define a new timescale, as a function of frequency, as
-%
 \be\label{eq:defTf}
@@
 \ee
-%
 The interpretation of this timescale for the inspiralling part of the signal, when the SPA applies, will be discussed in the next section.
 Whe can use this notation to rewrite~\eqref{eq:resultdffPsi} as
-%
 \begin{align}\label{eq:resultdffPsiTf}
-	 \tilde{s}(f) &= \tilde{h}(f) \sum\limits_{p\geq 0} \frac{(-i\epsilon)^{p}}{2^{p}p!} \Tf^{2p} \left( \frac{\partial^{2p} }{\partial t^{2p}} g \right)(f,\tf) \,, \nn \\
-	 \text{with } & \epsilon = -\mathrm{sgn}(\ud^{2}\Psi/\ud f^{2} ) \,.
+	 \tilde{s}(f) &= \tilde{h}(f) \sum\limits_{p\geq 0} \frac{(-i\epsilon)^{p}}{2^{p}p!} \Tf^{2p} \left( \frac{\partial^{2p} }{\partial t^{2p}} g \right)(f, \tf) \,,
 \end{align}
-%
+with $\epsilon = -\mathrm{sgn}(\ud^{2}\Psi/\ud f^{2} )$.
+
+The expansion~\eqref{eq:resultdfA} also leads to the natural introduction of a set of other timescales related to the successive derivatives of te amplitude.
+In an analogous manner to the definition~\eqref{eq:defTf} of the timescale $T_{f}$, we can define
+\be
+	\left( T_{Af}^{(n)} \right)^{n} \equiv \frac{1}{(2 \pi)^{n}} \frac{1}{A(f)} \left| \frac{\ud^{n} A}{\ud f^{n}} \right| \,,
+\ee
+where we included an absolute value to accomodate the possible sign changes in the right-hand side. In practice, only the first few of these timescales will be relevant.
 
@@
 
-\subsection{Relation to the stationary phase approximation}
+\subsection{Signal-dependent timescales and relation to the stationary phase approximation}
 \label{subsec:linkSPA}
@@
 Of particular interest is the relation between these results and the SPA.
-One first writes the time-domain signal in amplitude and phase form as $h(t) = a(t) e^{-2i\varphi(t)}$, where for gravitational wave signals $\varphi$ will correspond to the orbital phase, with the orbital frequency being $\omega = \dot{\varphi}$.
-The factor of 2 in the phase of the wave is appropriate for the dominant 22 harmonic of the signal.
-The approximation then applies to signals verifying the conditions
-%
-\be\label{eq:conditionSPA}
-	\left| \frac{\dot{a}/a}{\omega} \right| \ll 1\,, \quad \left|\frac{\dot{\omega}}{\omega^{2}} \right| \ll 1\,, \quad \left| \frac{(\dot{a}/a)^{2}}{\dot{\omega}} \right| \ll 1 \,.
-\ee
-%
-Since the integral~\eqref{eq:defFT} defining the Fourier transform is rapidly oscillatory unless the term $2\pi f t$ cancels the evolution of $-2\varphi(t)$, its support is well centered around the point of stationary phase.
-This defines the time of stationary phase implicitly by the relation
-%
-\be\label{eq:deftfSPA}
-	\omega(t_{f}) = \pi  f \,.
-\ee
-% 
-For a chirping signal of increasing phase, $\omega>0$ and $\dot{\omega}>0$, there is a unique point of stationary phase, located in the positive frequency range $f>0$.
-Using the conditions above, one formally expands the signal around $t_{f}$, to quadratic order in the time.
-The resulting integral yields
-%
-\begin{subequations}
-\begin{align}
-	\tilde{h}_{\rm SPA}(f) &= A_{\rm SPA}(f) e^{-i\Psi_{\rm SPA}(f)} \,, \\
-	A_{\rm SPA}(f) &= a(t_{f}) \sqrt{\frac{\pi}{\dot{\omega}(t_{f})}} \,, \\
-	\Psi_{\rm SPA}(f) &= 2\varphi(t_{f}) - 2\pi f t_{f} + \frac{\pi}{4} \,.
-\end{align}
-\end{subequations}
-%
-For non-precessing gravitational wave signals, the SPA is a very good approximation through the inspiralling phase.
-Ref.~\cite{} evaluated the first correction to this approximation and found that it can be considered as a term of the fifth post-Newtonian order.
+For non-precessing gravitational wave signals, the SPA is a very good approximation through the inspiralling phase~\cite{Droz+99}.
 However, it is important to note that the approximation breaks down when reaching the merger-ringdown part of the signal, as the rapidly varying features of the merger break the locality of the time-to-frequency relation~\eqref{eq:deftfSPA}.
@@
 Taking the first derivative of the defining relation~\eqref{eq:deftfSPA} gives
-%
 \be
@@
 \ee
-%
-which means that our generalized definition~\eqref{eq:deftf} for the effective time $\tf$ coincides with $t_{f}$ where the SPA is applicable. Similarly, if we define
-%
+which means that our generalized definition~\eqref{eq:deftf} for the effective time $t_{f}$ coincides with $t_{f}^{\rm SPA}$ in the inspiral phase of the signal, where the SPA is applicable.
+Similarly, if we define
 \be
-	T_{f} \equiv \frac{1}{\sqrt{2\dot{\omega}(t_{f})}} \,,
+	T_{f}^{\rm SPA} \equiv \frac{1}{\sqrt{2\dot{\omega}(t_{f})}} \,,
 \ee
-%
 we have
-%
 \be
-	T_{f}^{2} = -\frac{1}{4\pi^{2}}  \frac{\ud^{2} \Psi_{\rm SPA}}{\ud f^{2}} \,,
+	\left(T_{f}^{\rm SPA} \right)^{2} = -\frac{1}{4\pi^{2}}  \frac{\ud^{2} \Psi_{\rm SPA}}{\ud f^{2}} \,,
 \ee
-%
 where $\ud^{2}\Psi/\ud f^{2} < 0$ in the SPA with our sign conventions.
-Comparing with~\eqref{eq:defTf} above, we see that our generalized time scale $\Tf$ also reduces to $T_{f}$ when the SPA applies.
+Comparing with~\eqref{eq:defTf} above, we see that our generalized time scale $\Tf$ also reduces to $T_{f}^{\rm SPA}$ when the SPA applies.
 However, its definition extends to the merger-ringdown part of the signal, and we have to include an absolute value to allow for a possible change of sign of $\ud^{2}\Psi/\ud f^{2}$ in that regime.
 
-We can now compare our result to the approach of Ref.~\cite{}, where the authors extend the SPA, in a formalism called the shifted asymptotic expansion (SUA).
+We can now compare our result to the approach of Ref.~\cite{Klein+14}, where the authors extend the SPA in a formalism called the shifted uniform asymptotic expansion (SUA).
 The derivation includes an expansion of the modulation in Bessel functions, a study of the displacement of the stationary phase point induced by the precession, and a resummation of the result in the form of time derivatives.
-The main result of~\cite{}, their equation~(34), reads exactly as~\eqref{eq:resultdffPsiTf} with the identifications $\tilde{H}_{corr}(f)\rightarrow \tilde{s}(f)/\tilde{h}(f)$, $T\rightarrow \Tf$ and $e^{-i\delta\phi} \rightarrow F$.
-Thus, we are able to identify the Fourier-domain approximation that the treatment of~\cite{} amounts to, namely the approximation~\eqref{eq:expandPsi} where in the Fourier-domain convolution~\eqref{} the phase of the signal is expanded to quadratic order and the amplitude is not expanded.
+The main result of~\cite{Klein+14}, their equation~(34), reads exactly like~\eqref{eq:resultdffPsiTf} with the identifications $\tilde{H}_{corr}(f)\rightarrow \tilde{s}(f)/\tilde{h}(f)$ for the transfer function, $T\rightarrow \Tf$ for the radiation-reaction timescale and $e^{-i\delta\phi} \rightarrow F$ for the modulation function (which is restricted in their framework to include only the phase of the modulation, its amplitude being treated jointly with the amplitude of the signal).
+Thus, we are able to make a direct correspondence between the treatment of~\cite{Klein+14} and our framework, which provides us with a simple and alternative derivation of their result.
+Namely, the result of~\cite{Klein+14} corresponds to the approximation~\eqref{eq:expandPsi}, where in the Fourier-domain convolution the phase of the signal is expanded to quadratic order and the amplitude is not expanded.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Quadratic phase corrections and delays}
+\label{sec:quadphasedelay}
 
@@
 
-\subsection{Magnitude of the corrections beyond the LLP approximation}
-\label{subsec:sizecorr}
+\subsection{Resummation of derivatives for quadratic term in the phase}
+\label{subsec:resumquadphase}
 
-\subsubsection{Inspiral of a precessing binary}
-\label{subsubsec:sizecorrprecession}
+In this Section, we focus on the second-derivative corrective term in~\eqref{}.
+We will restrict to the case of a pure modulation, $d=0$ and $g(f,t) = F(t)$.
+The authors of~\cite{Klein+14} proposed a resummation scheme.
+Indeed, the result~\eqref{} looks like a symmetrized Taylor expansion, except for the factors $i^{p}$ and $1/p!$ instead of $1/(2p)!$.
+After having truncated the sum at some finite $N$, one can write
+%
+\begin{align}\label{eq:resumquadphase}
+	\tilde{s}(f) &\simeq \tilde{h}(f) \sum\limits_{p = 0}^{N} \frac{(-i\epsilon\Tf^{2})^{p}}{2^{p}p!} \partial_{t}^{2p}F(\tf) \nn\\
+	&\simeq \tilde{h}(f) \sum\limits_{p= 0}^{N} \sum\limits_{k=0}^{N} a_{N,k}\frac{(k\Tf)^{2p}}{(2p)!}  \partial_{t}^{2p}F(\tf) \nn\\
+	&\simeq \tilde{h}(f) \frac{1}{2}\sum\limits_{k=0}^{N} a_{N,k} \left( F(\tf + k\Tf) + F(\tf-k\Tf) \right)\,,
+\end{align}
+%
+where we applied a truncation approximation at the first and last stage.
+The second equality holds provided that the complex coefficients $a_{N,k}$ are a solution of the $N+1$-dimensional linear system
+\be\label{eq:combsystem}
+	(-i\epsilon)^{p} (2p-1)!! = \sum\limits_{k=0}^{N} a_{N,k} k^{2p} \quad \text{for } p=0,\dots,N \,.
+\ee
+The two solutions for the comb coefficients in the two cases $\epsilon = \pm 1$ are simply related by a complex conjugation.
+
+An immediate advantage of this reformulation is its improved numerical stability.
+For the signal of precessing binaries, although one should expect an ideal scaling $\partial_{t}^{p} F \sim \Omega^{p} F$ with $\Omega$ the precession frequency, it can be extremely hard to control numerical derivatives of $F$ (which in any realistic model will be only provided as sampled at discrete times) well enough to enforce this scaling.
+Here, however, one simply evaluates the original smooth function $F$ at shifted times.
+
+Note that the choice of the samples $\tf \pm k\Tf$ to approximate the successive derivatives of the expansion by finite differences is by no means unique.
+In principle, one could have made any ansatz of samples for the series expansion to be matched on.
+One cannot expect all choices to work equally well when applied to noisy data, and the problems of the numerical size of the coefficients and of the overall numerical stability of the scheme must be investigated numerically.
+
+However, we remark that the transformation~\eqref{eq:resumquadphase} between the two series above has a special structure, and the resulting linear system can be written as a Vandermonde system.
+Defining the Vandermonde matrix $V(x_{0},\dots,x_{N})$ as $V_{ij} = (x_{j})^{i}$ (with the convention $0^{0} = 1$), setting $x_{j} = j^{2}$, and $b_{p} \equiv (-i\epsilon)^{p}(2p-1)!!$, the linear system~\eqref{eq:combsystem} becomes
+\be
+	b_{p} = \sum\limits_{k=0}^{N} V_{p,k} a_{N,k} \,.
+\ee
+Now, since the Vandermonde system solves the Lagrange polynomial interpolation problem, we know that $({}^{t}V^{-1})_{ij}$ is the coefficient of $X^{i}$ in the polynomial $\prod_{k\neq j} (X-x_{k})/(x_{j} - x_{k})$.
+If we introduce the symmetric polynomials $\sigma$ such that
+\be	
+	\sigma(m, \{x_{1}, \dots, x_{n}\}) = \sum\limits_{1\leq i_{1}<\dots<i_{m}\leq n} x_{i_{1}}\dots x_{i_{m}} \,,
+\ee
+we obtain for the inverse of the matrix $V$
+\be
+	V^{-1}_{ij} = \frac{1}{\prod_{k\neq i} (i^{2} - k^{2})} (-1)^{N-j} \sigma(N-j, \{0,\dots,N^{2}\}\backslash \{i^{2}\}) \,.
+\ee
+This allows us to write a closed-form expression for the $a_{N,k}$, for every finite order $N$:
+\be
+	a_{N,k} = \sum\limits_{p=0}^{N} (-i\epsilon)^{p}(2p-1)!!
+\frac{(-1)^{N-p}}{\prod_{j\neq k} (j^{2}-k^{2})} \sum\limits_{\substack{ 0 \leq i_{1} < \dots < i_{N-p} \leq N \\ i_{1}, \dots, i_{N-p} \neq k}} i_{1}^{2}\dots i_{N-p}^{2}
+\ee
+The combinatorial explosion of the number of terms in the symmetric polynomials, however, makes this expression impractical to evaluate for $N$ beyond a few dozen.
+We leave a more complete investigation of the asymptotics and numerical stability of this resummation approach for future work.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Quadratic term in the phase and Fresnel transform}
+\label{subsec:fresneltransform}
+
+We can give an alternative interpretation of~\eqref{eq:resultdffPsi} and of Sec.~\ref{subsec:resumquadphase} above by using an integral approach.
+For simplicity of notation, here we will again keep to the case of precessing binaries, with only a modulation $F$ and no delays, $d=0$.
+We will reintroduce the delays below in Sec.~\ref{subsec:delays}.
+To obtain~\eqref{eq:resultdffPsi}, we expanded fully the exponential as a Taylor series.
+If it is not expanded, however, we have
+\be\label{eq:integralquadphase}
+	\tilde{s}(f)	\simeq \tilde{h}(f) \int \ud t\, F(t) \int\ud f'\, e^{2i\pi f' (t-\tf)} \exp\left[ 2i\pi^{2} \epsilon{f'}^{2} \Tf^{2} \right] \,,
+\ee
+where we recall that $\epsilon = -\mathrm{sgn}(\ud ^{2} \Psi/\ud f^{2})$.
+The integral on $f'$ is a simple complex Gaussian integral, and results in a complex Gaussian integral over time where we recognize a Fresnel transform~\cite{} of the function $F$.
+We introduce for the Fresnel transform the notation
+\be\label{eq:defFresnel}
+	\calF_{\tau}[F](t_{0}) \equiv \frac{e^{i\frac{\pi}{4}}}{\sqrt{2\pi} \tau} \int \ud t \, \exp\left[ - \frac{i}{2} \left( \frac{t-t_{0}}{\tau} \right)^{2}\right] F(t) \,,
+\ee
+together with the additional notation
+\be\label{eq:Fresnelsign}
+	\calF^{\epsilon}_{\tau}[F](t_{0}) \equiv
+\begin{cases}
+	 \calF_{\tau}[F](t_{0}) &\text{ if } \epsilon=1 \\
+	 \calF_{\tau}[F^{*}](t_{0})^{*} &\text{ if } \epsilon=-1
+\end{cases}
+\ee
+to accomodate for the possible sign change represented by $\epsilon$. The integral~\eqref{eq:integralquadphase} then reads
+\be\label{eq:resultFresnel}
+	\tilde{s}(f) = \tilde{h}(f) \calF^{\epsilon}_{\Tf}[F](\tf) \,.
+\ee
+
+The Fresnel transform is an integral transform that shares some properties with the Fourier transform.
+It intervenes in particular in the theory of diffraction~\cite{}.
+Contrarily to the Fourier transform, however, it is localized in the sense that the part of the integral that is centered around $t_{0}$ contributes predominantly, due to the cancelling oscillations far from $t_{0}$.
+Note also that only the part of the function $F$ that is symmetric about $t_{0}$ contributes to the integral in~\eqref{eq:defFresnel}.
+
+The parameter $\tau$ of the transform plays the role of a focusing parameter, determining how local the transform is.
+In the limit $\tau\rightarrow 0$, fast oscillations away from the central value $t_{0}$ will cancel out, leading to the integral taking the value $F(t_{0})$.
+For large values of $\tau$, however, the integral~\eqref{eq:defFresnel} has an extended support.
+Transposed in terms of the timescale $\Tf$, since we have seen that $\Tf$ can be interpreted as the radiation reaction timescale in the SPA regime, this means that a faster-chirping signal ($\Tf$ small) will have a Fresnel transform that is more focused, whereas a slower-chirping signal ($\Tf$ large) will have a Fresnel transform that is more extended.
+How much the Fresnel transform is focused in time is to be put in relation with how fast the function $F(t)$ in the integrand is varying.
+In the case of signals from precessing binaries, the precession timescale evolves as the system gets closer to merger, as will be discussed in details in~\ref{subsec:sizecorrPrec} below.
+In the case of the response of a LISA-like detectors, the modulation evolves with a timescale of one year, as will be detailed in Sec.\ref{subsec:sizecorrLISA}.
+
+Although it is interesting to be able to relate our results to a known transformation used in signal processing, whether this opens avenues for more efficient computation methods remains to be investigated, for instance by performing a decomposition of the modulation $F$ in chirplets, which have analytically computable Fresnel transforms.
+Note that in the result~\eqref{eq:resultFresnel}, both the scale $\Tf$ and the central time $\tf$ are functions of the frequency $f$.
+
+As a verification, upon performing a formal Taylor expansion in time of $F(t)$ around $\tf$ in the above integral and integrating term by term the result\footnote{Note that the intermediate integrals produced by such a Taylor expansion are formally divergent.
+One can regularize them for instance by introducing a small imaginary part in $\tau$ that and sending it to $0$ at the end of the computation.}, one readily recovers our previous result~\eqref{eq:resultdffPsiTf}.
+Now, the resummation of~\eqref{eq:resumquadphase} can be rephrased as a quadrature rule for computing the Fresnel integral~\eqref{eq:defFresnel}, according to
+\be\label{eq:Fresnelstencil}
+	\calF_{\tau}^{\epsilon} [F](t_{0}) \simeq \frac{1}{2} \sum\limits_{k=0}^{N} a_{N,k}^{\epsilon} \left( F(t_{0} + k\tau) + F(t_{0} - k\tau) \right)\,,
+\ee
+with $a_{N,k}^{1} = a_{N,k}$ and $a_{N,k}^{-1} = a_{N,k}^{*}$.
+Formally, if one allows for polynomial integrands in~\eqref{eq:defFresnel} (which can only be given sense by introducing a regularization), Eq.~\eqref{eq:Fresnelstencil} amounts to the building a quadrature rule for the particular choice of nodes $t_{0} \pm k \tau$, which is exact (with a regularization) if $F$ is a symmetric polynomial of degree $\leq 2N$.
+Note however that this quadrature rule does not fall into the class of Gaussian quadratures, which provide $N$-point rules that give an exact representation of the integral for polynomials of degree $2N-1$.
+Due to the complex character of the kernel in~\eqref{eq:defFresnel}, the induced inner product is not positive definite and some key properties are lost, such as the construction of the optimal Gaussian nodes from the zeros of the orthonormalized polynomials.
+For more details on quadrature rules when applied to integrals with oscillatory kernels, we refer to~\cite{}.
+
+%[TO CORRECT AND REWRITE]
+%The rewriting of~\eqref{eq:resultdffPsiTf} by means of an integral transform leads by analogy to another way of formulating the connection to the results of Ref.~\cite{Klein+14}.
+If we consider the inspiralling phase of a precessing binary, the SPA applies to the signal $h(t) = a(t)e^{-2i\varphi(t)}$.
+Refering to~\eqref{} above, if we consider as before that $a(t)$ is slowly variable
+%\be
+%	\tilde{s}(f) \simeq \tilde{h}_{\rm SPA} (f) \simeq a(t_{f}^{\rm SPA}) \exp\left[2i\pi f t_{f}^{\rm SPA}-2i%\varphi(t_{f}^{\rm SPA}) \right]  \int \ud t \, F(t) e^{-i \dot{\omega} (t_{f}^{\rm SPA}) (t-t_{f}^{\rm SPA})^{2}} \,.
+%\ee
+%We see that, within the context of the SPA, the result~\eqref{} for the quadratic term in the phase, equivalent to the result of~\cite{Klein+14}, can be recovered by a time-domain approach.
+However, our formalism is somewhat more general, as it does not make use of the SPA for the underlying signal $\tilde{h}(f)$ and can in principle be used for the merger-ringdown part of the signal as well.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Treatment of the delays}
+\label{subsec:delays}
+
+In the case of a LISA-type detector response, the presence of the delay $d(t)$ is responsible for the frequency-dependence of the kernel $g$ introduced in~\eqref{eq:defg}.
+We will show below in Sec.~\ref{} that the higher-order correction in the delay can become non-negligible at high frequencies, as measured by the magnitude of the first correction in~\eqref{eq:taylordelay}.
+This is especially true for the orbital delay related to the motion of the whole detector constellation around the Sun, due to the large baseline for this delay.
+
+Here we give an alternative treatment of the delays which is an alternative to the mere Taylor expansion of~\eqref{eq:taylordelay}.
+It is useful to look first at the case where one neglects the signal-related amplitude and phase corrections as given in~\eqref{eq:resultdfA} and~\eqref{eq:resultdffPsiTf}, but keeps the frequency-dependence in the convolution kernel in~\eqref{eq:defg}.
+The response then depends on the signal only by the time-to-frequency correspondence and reads
+\begin{equation}
+	\tilde{s}(f) = \tilde{h}(f) \int \ud t \, F(t) e^{-2i\pi f d(t)} \int \ud f' \, e^{2i\pi f' (t+d(t) - t_{f})} \,,
+\end{equation}
+where we recall that we defined $t_{f} \equiv -1/(2\pi)\ud \Psi/\ud f$.
+This motivates the introduction of the delayed time function $t_{d}:t \mapsto t+d(t)$, and of the reciprocal function $t_{d}^{-1}$, and of a modified time-to-frequency correspondence $\tfd = t_{d}^{-1}(\tf)$, defined implicitly by
+\be
+	\left. (t + d(t) - t_{f})\right|_{t=t_{f}^{d}} = 0 \,.
+\ee
+The above integral can then be formally computed as
+\begin{align}\label{eq:delaycorrleading}
+	\tilde{s}(f) &= \tilde{h}(f) \int \ud t \, F(t) e^{-2i\pi f d(t)} \delta(t + d(t) - t_{f}) \nn \\
+	&= \tilde{h}(f) F(t_{f}^{d}) \frac{e^{-2i\pi f d(t_{f}^{d})}}{1+\dot{d}(t_{f}^{d})}
+\end{align}
+
+For the standard LISA configuration, we have for the orbital delays the scaling $d_{O}\sim R/c \sim 500s$, for the constellation delays the scaling $d_{c}\sim L/c \sim 17s$ (with an additional dependence on angular factors).
+Since the motion of the constellation is periodic with a period of one year and $\Omega_{0} = 2.10^{-7}\mathrm{rad}.s^{-1}$, we have $\dot{d}_{O} \sim \Omega_{0} R/c \sim 10^{-4}$ and $\dot{d}_{c} \sim \Omega_{0} L/c \sim 3.10^{-6}$.
+The smallness of the dimensionless quantity $\dot{d} \ll 1$ (and of its subsequent derivatives) will allow us to treat it perturbatively with a very good approximation, and shows also that the function $t_{d}$ is univalued and that there is no ambiguity in defining its reciprocal $t_{d}^{-1}$.
+
+By treating $\dot{d}$ as a perturbation and keeping only first-order terms, we obtain for the delayed time reciprocal function
+\begin{align}
+	t_{d}^{-1}(t) &\simeq t-d(t) (1-\dot{d}(t)) \,,\nn\\
+	d(t_{f}^{d}) &\simeq d(t_{f}) ( 1 - \dot{d}(t_{f})) \,.
+\end{align}
+Now, the most relevant correction in~\eqref{eq:delaycorrleading} comes from the phase factor at high frequencies, where the factors $2\pi f d_{O}$ and $2\pi f d_{c}$ give a magnification by a factor of, respectively, $3.10^{3}$ and $10^{2}$ at $1\Hz$.
+Ignoring the other corrections, we thus arrive at the approximate form
+\be
+	\tilde{s}(f) \simeq \tilde{h}(f) F(t_{f})\exp\left[ -2i\pi f d(t_{f}) (1-\dot{d}(t_{f})) \right] \,.
+\ee
+Notice that this leading-order correction in the treatment of the delays affects purely the phase of the signal.
+
+Next, we consider the case where the quadratic phase correction is kept as well, as in Sec.~\ref{subsec:fresneltransform}.
+Keeping as before only the first-order terms in $\dot{d}$ (and neglecting its higher derivatives), we can write
+\begin{align}
+	\tilde{s}(f) &\simeq \tilde{h}(f) \int \ud t \, F(t) e^{-2i\pi f d(t)} \int \ud f' \, \exp\left[ 2i\pi \epsilon \Tf^{2} f'^{2} + 2i\pi f' (t+d(t) - \tf) \right] \nn\\
+	&\simeq \tilde{h}(f) \frac{e^{i\epsilon\frac{\pi}{4}}}{\sqrt{2\pi}\Tf} \int \ud \tau \, \frac{F(\tau - d(\tau))}{1+\dot{d}(\tau)} e^{-2i\pi f d(\tau)(1-\dot{d}(\tau))}\exp\left[ -\frac{i\epsilon}{2} \frac{(\tau - \tf)^{2}}{\Tf^{2}} \right] \,,
+\end{align}
+where we used a change of variable $\tau = t_{d}(t)$. We see that the result can again be expressed as a Fresnel transform.
+
+When considering amplitude corrections as well, as in~\eqref{}, additional powers of $f'$ can be translated as time derivatives with respect to the variable $\tau$ after performing the change of variables.
+Thus, we obtain our most general result for an exact treatment of the quadratic term in the phase and for an improved treatment of the delays as:
+\be
+	\tilde{s}(f) = \tilde{h}(f) \sum\limits_{k=0}^{+\infty} \frac{1}{(2i\pi)^{k} k!} \frac{1}{A} \frac{\ud^{k}A}{\ud f^{k}}  \calF^{\epsilon}_{\Tf} \left[ \frac{\ud^{k}}{\ud \tau^{k}} \left( \frac{F(\tau - d(\tau))}{1+\dot{d}(\tau)} e^{-2i\pi f d(\tau)(1-\dot{d}(\tau))} \right) \right] (\tf) \,.
+\ee
+Each term in this series (in practice, only the first few will be relevant) can in turn be approximated by using the stencil formula~\eqref{eq:Fresnelstencil}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Application to the response of LISA-type detectors}
+\label{sec:LISA}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{The response model and TDI observables}
+\label{subsec:modelLISA}
+
+In this Section, we expand Sec.~\ref{subsec:} and detail more the model that we use for the response of a LISA-like detector, together with the assumptions used and their limitations.
+
+As argued in~\cite{Krolak+04}, since we focus on the gravitational-wave contribution to the basic observable, we can use a simplified model for the response, ignoring corrections that are crucial from the point of view of noise cancellations.
+Our aim is indeed to assess the accuracy of a direct Fourier-domain treatment of the response, thus focusing on the separation of timescales in the problem, keeping in mind that corrections can be added later without affecting the conclusions of the present analysis.
+
+The frequency-shift response for a single link~\eqref{eq:yslr} was derived in Ref.~\cite{EW75} (see also~\cite{} for a more modern version).
+Several assumptions enter the result as written in~\eqref{eq:yslr}: (i) effects of the order $v/c$ are neglected, including for instance the special relativistic Doppler effect created by the relative speeds of the spacecrafts on their orbits (ii) the propagation is assumed to take place in a flat spacetime, perturbed only by the gravitational wave; thus the gravitational redshift as well as the deflection of light created by the gravitational potential of the Sun is ignored (iii) all geometric factors are evaluated at a single time, whereas one should consider the beam as propagating from the position of the first spacecraft at the time of emission to the position of the second scapecraft at the time of reception, leading to a pointing-ahead effect.
+
+Additionally, we limit ourselves to a rigid model for the detector, namely we assume that the constellation remains in an equilateral configuration with fixed armlengths $L=5.10^{6}\mathrm{km}$.
+This simplified orbits neglect (iv) the effect of excentricity of the individual Keplerian orbits, which introduces corrections at the level $e^{2}$ (v) the effect of gravitational perturbations coming from other celestial bodies, such as the Earth, the quadrupole of the Sun, and the other planets.
+
+Although all the effects (i)-(v) are neglected here, keeping track of these corrections is crucial for the purpose of noise cancellations, and led to the development of new generations of TDI observables~\cite{}.
+
+We now turn to the transposition of~\eqref{eq:yslr} to the Fourier-domain.
+We will choose to write the response directly in terms of spin-weighted spherical modes $h_{\ell m}$, whose Fourier transforms are assumed to have a smooth amplitude and phase.
+Since the orbital delay due to the orbit around the Sun is common to all $y_{slr}$ observables, we will introduce the orbital-delayed waveform and rewrite the response as
+\begin{align}
+	h^{\rm TT}_{O} (t) &\equiv h^{\rm TT}(t-\hat{k}\cdot p_{O}) \,, \nn\\
+	y_{slr} &= \frac{1}{2} \frac{1}{1 - \hat{k}\cdot n_{l}} n_{l}\cdot \left[ h_{O}^{\rm TT}(t - \hat{k}\cdot p^{L}_{s}) - h_{O}^{\rm TT}(t - \hat{k}\cdot p^{L}_{r}) \right] \cdot n_{l}\,,
+\end{align}
+where we introduced the positions of the spacecrafts relative to the center of the constellation, $p^{L}_{A} \equiv p_{A} - p_{O}$.
+
+When using the rigid approximation for the constellation, a particular simplification occurs when writing the transfer function at leading order in the delays.
+First, we define the matrices $P^{+},P^{\times}$ such that
+\be
+	h^{\rm TT} = h_{+}P_{+} + h_{\times}P_{\times} \,,
+\ee
+in the sense of matrices.
+The spin-weighted spherical harmonics factors in~\eqref{eq:hpcfrommodes} are mere constants of time, so we can also define a geometric projection factor $P_{slr}^{\ell m}$ for each individual mode.
+Under the assumption that we can neglect the negative frequencies for the modes $m>0$ and the positive frequencies for $m<0$ (which is also used for precessing waveforms), we have for $f>0$ and $m>0$
+%\begin{subequations}
+\begin{align}
+	P_{slr}^{\ell m} &= \frac{1}{4} \frac{1}{1-\hat{k}\cdot n_{l}} {}_{-2}Y_{\ell m} n_{l}\cdot \left( P_{+} - i P_{\times} \right) \cdot n_{l} \text{ for } m>0 \,,
+\end{align}
+%\end{subequations}
+
+Note that, if the corrections~\eqref{} are included for the constellation delays, this transfer function will not have this simple form anymore as $\dot{d}$ will have a different velocity-dependent expression for the sending and receiving spacecrafts.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Estimates for the magnitude of higher-order corrections}
+\label{subsec:sizecorrLISA}
+
+If we put actual numbers, we have $\Omega/2\pi \simeq 3.1\times10^{-8}\mathrm{Hz} \ll f$ over all the frequency band we consider.
+Considering that the comb is of limited extend in $n$, we might want to Taylor expand the relevant quantities accross the comb.
+Since the phase is rapidly varying, and since we have to consider a constant and a linear term in the phase as arbitrary, it is natural to Taylor expand the phase at the first order (that is, approximate the phase linearly locally), neglect the $n\Omega/2\pi$ in $(f-n\Omega/2\pi)$ appearing in the argument of the amplitude and other terms, and assess the quality of this approximation.
+
+Preliminary tests show that this approximation should be valid, unsurprisingly, for all the mildly varying functions encountered: $\sinc$, $\exp$ and the linear term in the $c_{n}$'s, and the Bessel functions $J_{n}$.
+However, $\tilde{h}$ has a Fourier-domain phase which becomes very rapidly varying at low frequencies, in the deep inspiral, $\Psi\propto f^{-5/3}$; here it is not obvious that we can neglect the second derivative of the phase.
+
+The magnitude of the first terms in a Taylor expansion in $f_{0}$ is (taking $10^{-5}\mathrm{Hz}$ as a reference starting frequency):
+%
+\begin{widetext}
+\begin{subequations}
+\begin{align}
+	|\delta A/A| &= \left| \frac{1}{A_{N}}\frac{\ud A_{N}}{\ud f} f_0 \right| \simeq 3\times10^{-3} \left(\frac{f}{10^{-5}\Hz}\right)^{-1} \,,\\
+	|\delta^{1}\Psi | &= \left| f_0\frac{\ud \Psi_{N}}{\ud f} \right| \simeq 10^{3}\left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-8/3} \,, \\
+	|\delta^{2}\Psi | &= \left| \frac{1}{2} f_0^{2}\frac{\ud^{2} \Psi_{N}}{\ud f^{2}} \right| \simeq 5 \left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-11/3} \,, \\
+	|\delta^{3}\Psi | &= \left| \frac{1}{6} f_0^{3}\frac{\ud^{3} \Psi_{N}}{\ud f^{3}} \right| \simeq 2\times 10^{-2} \left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-14/3} \,.
+\end{align}
+\end{subequations}
+\end{widetext}
+%
+The approximation degrades for lower starting frequencies, and, at a fixed frequency, when the total mass decreases and/or the mass ratio increases.
+However, if we consider only merging binaries, that is to say binaries of which the merger is seen during the observation period of the mission, we have a lower limit for the starting frequency which can be larger than the detector's lowest frequency in band.
+
+From the Newtonian estimate~\eqref{}, the starting frequency is then
+%
+\be
+	f_{\rm start} = 10^{-5}\mathrm{Hz} \left( \frac{m}{8.5\times10^{6}\Msol} \right)^{-5/8} \left( \frac{\nu}{1/4} \right)^{-3/8} \left( \frac{\Delta t}{5 \mathrm{yr}} \right)^{-3/8} \,.
+\ee
+%
+So we see that the ``worst'' total mass is $8.5\times10^6\Msol (\nu/(1/4))^{3/5}$, for 5 years of observation and $f_{\rm low} = 10^{-5}\mathrm{Hz}$ for the entry in band, which are conservative values; for higher masses, part of the signal is cut because it is out-of-band, while for lower masses, the signal is cut because of the limited time of observation before merger, and we have the scaling
+%
+\be
+	|\delta^{2}\Psi |_{f_\mathrm{start}} \propto m^{5/8}\nu^{3/8}\Delta t^{11/8} \,,
+\ee
+%
+which shows that the effect of the truncation of the signal for a finite observation time wins: the approximation becomes better for lower mass and higher mass ratios systems.
+The worse point in the parameter space should be, for these Newtonian estimates, at equal mass and $m=8.5\times 10^{6} \Msol$, with $|\delta^{2}\Psi |_{\mathrm{worst}} \simeq 0.15$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Error control for the Fourier-domain response}
+\label{subsec:errorsLISA}
+
+[implementation]
+[examples of errors]
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Application to waveforms from precessing binaries}
+\label{sec:Precession}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Morphology of precessing waveforms}
+\label{subsec:precmodel}
+
+[explain inertial frame/precessing frame]
+[explain Gatech formula for post-merger precession]
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Simplified model of precessing IMR waveform}
+\label{subsec:precmodel}
+
+Here we will only consider the inertial-frame modes induced by the precessing-frame dominant 22 mode.
+
+[introduce phenomP]
+[explain model with smoothing for the angles with its two versions]
+[qualify how bad the model is with respect to full NR]
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Estimates for the magnitude of higher-order corrections}
+\label{subsec:sizecorrPrec}
 
@@
 
-\subsubsection{The case of a LISA-type detector [TO BE UPDATED]}
-\label{subsubsec:sizecorrLISA}
-
-If we put actual numbers, we have $\Omega/2\pi \simeq 3.1\times10^{-8}\mathrm{Hz} \ll f$ over all the frequency band we consider.
-Considering that the comb is of limited extend in $n$, we might want to Taylor expand the relevant quantities accross the comb.
-Since the phase is rapidly varying, and since we have to consider a constant and a linear term in the phase as arbitrary, it is natural to Taylor expand the phase at the first order (that is, approximate the phase linearly locally), neglect the $n\Omega/2\pi$ in $(f-n\Omega/2\pi)$ appearing in the argument of the amplitude and other terms, and assess the quality of this approximation.
-
-Preliminary tests show that this approximation should be valid, unsurprisingly, for all the mildly varying functions encountered: $\sinc$, $\exp$ and the linear term in the $c_{n}$'s, and the Bessel functions $J_{n}$.
-However, $\tilde{h}$ has a Fourier-domain phase which becomes very rapidly varying at low frequencies, in the deep inspiral, $\Psi\propto f^{-5/3}$; here it is not obvious that we can neglect the second derivative of the phase.
-
-For a Newtonian inspiral, applying the SPA gives $\tilde{h}(f) = A_{N}(f)e^{-i\Psi_{N}(f)}$ with
-%
-\begin{subequations}
-\begin{align}
-	A_{N}(f) &= -\sqrt{\frac{5\pi}{24}} \frac{G^{2}m^{2}}{Dc^{5}} \nu^{1/2}v^{-7/2}\,,\\
-	\Psi_{N}(f) &= 2\pi f t_{0} - \phi_{0} + \frac{3}{128\nu v^{5}} \,, 
-\end{align}
-\end{subequations}
-%
-where $v=(G m \pi f/c^{3})^{1/3}$. The magnitude of the first terms in a Taylor expansion in $f_{0}$ is (taking $10^{-5}\mathrm{Hz}$ as a reference starting frequency):
-%
-\begin{widetext}
-\begin{subequations}
-\begin{align}
-	|\delta A/A| &= \left| \frac{1}{A_{N}}\frac{\ud A_{N}}{\ud f} f_0 \right| \simeq 3\times10^{-3} \left(\frac{f}{10^{-5}\Hz}\right)^{-1} \,,\\
-	|\delta^{1}\Psi | &= \left| f_0\frac{\ud \Psi_{N}}{\ud f} \right| \simeq 10^{3}\left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-8/3} \,, \\
-	|\delta^{2}\Psi | &= \left| \frac{1}{2} f_0^{2}\frac{\ud^{2} \Psi_{N}}{\ud f^{2}} \right| \simeq 5 \left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-11/3} \,, \\
-	|\delta^{3}\Psi | &= \left| \frac{1}{6} f_0^{3}\frac{\ud^{3} \Psi_{N}}{\ud f^{3}} \right| \simeq 2\times 10^{-2} \left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-14/3} \,.
-\end{align}
-\end{subequations}
-\end{widetext}
-%
-The approximation degrades for lower starting frequencies, and, at a fixed frequency, when the total mass decreases and/or the mass ratio increases.
-However, if we consider only merging binaries, that is to say binaries of which the merger is seen during the observation period of the mission, we have a lower limit for the starting frequency which can be larger than the detector's lowest frequency in band.
-
-Namely, if we take a Newtonian estimate for the relation between the frequency and the time to coalescence, we have
-%
-\be
-	\pi f(t) = \left[ \frac{256\nu}{5c^{5}} (Gm)^{5/3} \Delta t \right]^{-3/8} \,,
-\ee
-%
-with $\Delta t$ the time to coalescence, and the starting frequency is then
-%
-\be
-	f_{\rm start} = 10^{-5}\mathrm{Hz} \left( \frac{m}{8.5\times10^{6}\Msol} \right)^{-5/8} \left( \frac{\nu}{1/4} \right)^{-3/8} \left( \frac{\Delta t}{5 \mathrm{yr}} \right)^{-3/8} \,.
-\ee
-%
-So we see that the ``worst'' total mass is $8.5\times10^6\Msol (\nu/(1/4))^{3/5}$, for 5 years of observation and $f_{\rm low} = 10^{-5}\mathrm{Hz}$ for the entry in band, which are conservative values; for higher masses, part of the signal is cut because it is out-of-band, while for lower masses, the signal is cut because of the limited time of observation before merger, and we have the scaling
-%
-\be
-	|\delta^{2}\Psi |_{f_\mathrm{start}} \propto m^{5/8}\nu^{3/8}\Delta t^{11/8} \,,
-\ee
-%
-which shows that the effect of the truncation of the signal for a finite observation time wins: the approximation becomes better for lower mass and higher mass ratios systems.
-The worse point in the parameter space should be, for these Newtonian estimates, at equal mass and $m=8.5\times 10^{6} \Msol$, with $|\delta^{2}\Psi |_{\mathrm{worst}} \simeq 0.15$.
+[add plots and discussion of real figures of merit as measured from the model]
 
@@
 
-\subsection{Exact quadratic term in the phase and Fresnel transform}
-\label{subsec:fresneltransform}
+\subsection{Direct convolution approach for the merger-ringdown phase}
+\label{subsec:directconvolution}
 
-Since the second-derivative corrective term in~\eqref{} is the most interesting, since it contains the most important deviation from the LLP, we will investigate it in more details in this section.
-For simplicity of notation, here we will keep to the case of precessing binaries, with only a modulation $F$ and no delays.
-
-To obtain~\eqref{eq:resultdffPsi}, we expanded the exponential. If we keep it as it is, however, we have
-%
-\begin{subequations}
-\begin{align}
-	\tilde{s}(f) &= \tilde{h}(f) \int\ud f'\, \tilde{F}(f') e^{-2i\pi f'\tf} \exp\left[ -\frac{i}{2} \frac{\ud^{2} \Psi}{\ud f^{2}} {f'}^{2} \right] \, \\
-	&= \tilde{h}(f) \int \ud t\, F(t) \int\ud f'\, e^{2i\pi f' (t-\tf)} \exp\left[ 2i\pi^{2} \epsilon{f'}^{2} \Tf^{2} \right] \,,
-\end{align}
-\end{subequations}
-%
-where we recall that $\epsilon = -\mathrm{sgn}(\ud ^{2} \Psi/\ud f^{2})$. The resulting Gaussian integral yields [check how $\epsilon$ propagates]
-%
-\begin{subequations}
-\begin{align}\label{eq:resultFresnel}
-	\tilde{s}(f) &= \tilde{h}(f) \calF_{\Tf}[F](\tf) \quad \text{if } \epsilon = 1 \,,\\
-	\tilde{s}(f) &= \tilde{h}(f) \calF_{\Tf}[F](\tf)^{*} \quad \text{if } \epsilon = -1 \,,
-\end{align}
-\end{subequations}
-%
-where we introduced the Fresnel transform~\cite{} of $F$, defined as follows:
-%
-\be\label{eq:defFresnel}
-	\calF_{\tau}[F](t_{0}) \equiv \frac{e^{i\frac{\pi}{4}}}{\sqrt{2\pi} \tau} \int \ud t \, \exp\left[ - \frac{i}{2} \left( \frac{t-t_{0}}{\tau} \right)^{2}\right] F(t) \,.
-\ee
-%
-The Fresnel transform applies to a large class of functions [be more precise] and intervenes in the theory of diffraction [give references and examples].
-Although it is interesting to be able to relate our results to a known transformation used in signal processing, whether this opens avenues for efficient computation methods remains to be investigated (for instance by performing a linear decomposition of the modulation $F$ in wavelets, or chirplets, which have analytically computable Fresnel transforms).
-
-As a verification, upon performing a Taylor expansion in time of $F(t)$ around $\tf$ in the above integral and integrating  term by term the result, one readily recovers our previous result~\eqref{eq:resultdffPsiTf}.
+[describe treatment of MR by direct convolution]
 
@@
 
-\subsection{Resummation of derivatives}
-\label{subsec:resummation}
+\subsection{Error control for the Fourier-domain modulation}
+\label{subsec:errorsPrec}
 
-The authors of~\cite{} did not reformulate their result as in~\eqref{eq:resultFresnel}, but they proposed  a resummation scheme.
-Indeed, the result~\eqref{} looks like a symmetrixed Taylor expansion, except for the factors $i^{p}$ and $1/p!$ instead of $1/(2p)!$.
-After having truncated the sum at some finite $N$, one can write
-%
-\begin{align}\label{eq:resum}
-	\tilde{s}(f) &\simeq \tilde{h}(f) \sum\limits_{p = 0}^{N} \frac{(-i\epsilon\Tf^{2})^{p}}{2^{p}p!} \partial_{t}^{2p}F(\tf) \nn\\
-	&\simeq \tilde{h}(f) \sum\limits_{p= 0}^{N} \sum\limits_{k=0}^{N} a_{N,k}\frac{(k\Tf)^{2p}}{(2p)!}  \partial_{t}^{2p}F(\tf) \nn\\
-	&\simeq \tilde{h}(f) \frac{1}{2}\sum\limits_{k=0}^{N} a_{N,k} \left( F(\tf + k\Tf) + F(\tf-k\Tf) \right)\,,
-\end{align}
-%
-where we applied a truncation approximation at the first and last stage.
-The second equality holds provided that the complex coefficients $a_{N,k}$ are a solution of the $N+1$-dimensional linear system
-%
-\be
-	(-i\epsilon)^{p} (2p-1)!! = \sum\limits_{k=0}^{N} a_{N,k} k^{2p} \quad \text{for } p=0,\dots,N \,.
-\ee
-%
-The two solutions for the comb coefficients in the two cases $\epsilon = \pm 1$ are simply related by a complex conjugation.
-
-An immediate advantage of this reformulation is its improved numerical stability.
-Although one should expect an ideal signal the scaling $\partial_{t}^{p} F \sim \Omega^{p} F$ with $\Omega$ the precession frequency, it can be extremely hard to control numerical derivatives of $F$ (which in any realistic model will be only provided as sampled at discrete times) well enough to enforce this scaling.
-Here, however, one simply evaluates the original smooth function $F$ at shifted times.
-
-Note, interestingly, that, if the validity of the comb approach was generic, and if the coefficients $a_{N,k}$ were to converge for $N\rightarrow +\infty$ to a well-defined sequence admitting an appropriate fall-off with $k$, then one would have found a magical way of computing the Fresnel transform of a signal by simply evaluating it on a comb made of at most a few dozen equally spaced times.
-
-This seems to good to be true.
-In general, one could have made any ansatz for the series expansion to be matched on, and one therefore cannot expect in general nice properties to emerge from a resummation.
-However, the transformation~\eqref{} between the two series above has a special structure.
-Namely, written in matrix form, the linear system is a Vandermonde system.
-Defining the Vandermonde matrix $V(x_{0},\dots,x_{N})$ as $V_{ij} = (x_{j})^{i}$ (with the convention $0^{0} = 1$), setting $x_{j} = j^{2}$, and $b_{p} \equiv (-i\epsilon)^{p}(2p-1)!!$, the linear system is simply
-%
-\be
-	b_{p} = V_{p,k} a_{N,k} \,.
-\ee
-%
-Now, since the Vandermonde system solves the Lagrange polynomial interpolation problem, we know that .
-Namely, $({}^{t}V^{-1})_{ij}$ is the coefficient of $X^{i}$ in the polynomial $\prod_{k\neq j} (X-x_{k})/(x_{j} - x_{k})$.
-If we introduce the symmetric polynomials $\sigma$ such that
-%
-\be	
-	\sigma(m, \{x_{1}, \dots, x_{n}\}) = \sum\limits_{1\leq i_{1}<\dots<i_{m}\leq n} x_{i_{1}}\dots x_{i_{m}} \,,
-\ee
-%
-we obtain for the inverse of the matrix $V$
-%
-\be
-	V^{-1}_{ij} = \frac{1}{\prod_{k\neq i} (i^{2} - k^{2})} (-1)^{N-j} \sigma(N-j, \{0,\dots,N^{2}\}\backslash \{i^{2}\}) \,.
-\ee
-%
-This allows us to write an ``analytical'' expression for the $a_{N,k}$ for every finite order $N$:
-%
-\be
-	a_{N,k} = \sum\limits_{p=0}^{N} (-i\epsilon)^{p}(2p-1)!!
-\frac{(-1)^{N-p}}{\prod_{j\neq k} (j^{2}-k^{2})} \sum\limits_{\substack{ 0 \leq i_{1} < \dots < i_{N-p} \leq N \\ i_{1}, \dots, i_{N-p} \neq k}} i_{1}^{2}\dots i_{N-p}^{2}
-\ee
-%
-The combinatorial explosion of the number of terms in the symmetric polynomials, however, makes this expression impractical to evaluate for $N$ beyond a few dozen.
-It could allow, however, one to control the asymptotics and maybe prove the convergence or non-convergence of the sequence of $a_{N,k}$ for $N\rightarrow +\infty$.
-[wishful thinking]
-
-\section*{---BELOW IS OUTDATED MATERIAL---}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Reduced order models for fast waveform generation}
-\label{subsec:rom}
-
-[review recent advances in this field: surrogates, ROQ, P\"urrer's SVD method in some more details]
-
-[present the mode-by-mode time and phase offset tracking, specific to HM]
-
-[specific choices: logarithmic sampling, 1D spline interpolation, smoothing spline for projection coefficients]
-
-[give summary plot of unfaithfulness]
-
-[compare speedup with ordinary EOB]
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Fourier-domain response of LISA-type detectors}
-\label{subsec:fdresponse}
-
-We use the notations of the Kr\'olak \& al paper. We denote the annual frequency of the motion by $f_{0}=1/\mathrm{year}$, and $\Omega = 2\pi f_{0}$. We set $c=1$ except for numerical applications.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection*{Definitions}
-
-The primary observables for a LISA-type instrument are given by (taking $y_{21}$ as an example)
-%
-\be
-	y_{21}(t) = \left( 1-\hatk \cdot n_{3}\right) \left[ \Psi_{3}(t+\hatk\cdot p_{2} - L) - \Psi_{3}(t+\hatk\cdot p_{1}) \right] \,,
-\ee
-%
-with the definitions
-%
-\be
-	\Psi_{A} = \frac{\Phi_{A}}{1-(\hatk\cdot n_{A})^{2}} \,, \quad \Phi_{A} = \frac{1}{2} n_{A}^{i} H_{ij} n_{A}^{j} \,.
-\ee
-%
-Here we follow the approximation made in~\cite{Krolak+04}: all vectors related to the positions, $n_{A}$ and $p_{A}$, will be considered evaluated at the time $t$, neglecting their variation on the time scale of light propagation in the system, which is motivated by $\Omega R/c \ll 1$, $\Omega L/c\ll 1$.
-It seems also that this approximation is used anyway in the first place when deriving the above form for the frequency variation of the lasers [to check].
-We will also idealize the geometry of the constellation by keeping the linear order in eccentricity and neglecting higher-order orbital perturbations.
-Another convenient definition is
-%
-\be
-	\Phi_{A} = P_{A}^{+} h_{+} + P_{A}^{\times}h_{\times} \,,
-\ee
-%
-where $h_{+}$, $h_{\times}$ are the two polarizations of the wave and the $P_{A}$'s are combinations of trigonometric functions depending on $\Omega t$.
-
-\subsection*{Fourier transform of a delayed signal}
-
-Let us consider, in general, the problem of determining the Fourier transform of a signal $h$ (whose Fourier transform $\tilde{h}$ is known), when applying a varying delay $d(t)$ and also multiplying by some function $F(t)$.
-Defining
-%
-\be
-	h_{d}(t) = h(t+d(t)) \,, \quad s(t) = F(t)h_{d}(t) \,,
-\ee
-%
-we have
-%
-\be
-	h_{d}(t) = \int \ud f \, e^{-2i\pi f (t+d(t))}\tilde{h}(f) \,,
-\ee
-%
-and
-%
-\begin{align}
-	\tilde{s}(f) &= \int \ud t \, e^{2i\pi f t} F(t)  \int \ud f' \, e^{-2i\pi f' (t+d(t))}\tilde{h}(f') \nn\\
-	&= \int \ud f' \, \tilde{h}(f-f') \int \ud t \, e^{2i\pi f' t} e^{-2i\pi (f-f') d(t)} F(t) \,.
-\end{align}
-%
-We can rewrite the last equation as a kind of convolution with a frequency-dependent Kernel, according to
-%
-\begin{align}\label{eq:freqkernel}
-	\tilde{s}(f) &= \int \ud f' \, \tilde{h}(f-f') G(f,f') \,, \nn\\
-	G(f,f') &= \int \ud t \, e^{2i\pi f' t} e^{-2i\pi (f-f') d(t)} F(t) \,.
-\end{align}
-%
-Since in our case both $d(t)$ and $F(t)$ are $2\pi/\Omega$-periodic, the latter Kernel will reduce to a discrete comb in frequency.
-If we define $g[f](t) = e^{-2i\pi f d(t)} F(t)$, a Fourier series decomposition gives
-%
-\begin{align}
-	g[f](t) &= \sum\limits_{n\in \mathbb{Z}} c_{n}[g](f) e^{-i n\Omega t} \,, \nn\\
-	c_{n}[g](f) &= \frac{\Omega}{2\pi}\int_{0}^{\frac{2\pi}{\Omega}} \ud t \, g[f](t) e^{i n \Omega t} \,,
-\end{align}
-%
-which transforms~\eqref{eq:freqkernel} into
-%
-\begin{align}
-	G(f,f') &= \sum\limits_{n\in \mathbb{Z}} c_{n}[g](f-f') \delta\left(f' - n f_{0}\right) \,, \nn\\
-	\tilde{s}(f) &= \sum\limits_{n\in \mathbb{Z}} c_{n}[g]\left(f - n f_{0}\right) \tilde{h}\left(f - n f_{0}\right) \,.
-\end{align}
-%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection*{Application to $y_{AB}$ observables}
-
-We now apply the above results to $y_{AB}$ observables, with the example of $y_{21}$.
-We will assume our starting point are the Fourier transforms of $h_{+}$ and $h_{\times}$.
-We get for either polarization
-%
-\begin{widetext}
-\begin{align}
-	g_{21}^{+,\times}[f](t) &= \frac{P_{3}^{+,\times}}{1+\hatk\cdot n_{3}} \left[ e^{-2i\pi f (\hatk\cdot p_{2} - L)} - e^{-2i\pi f \hatk\cdot p_{1}} \right] \,, \nn\\
-	&= i\pi L f P_{3}^{+,\times} \sinc\left[ \pi L f(1+\hatk \cdot n_{3})\right]\exp\left[ i \pi f L \left(  1 - \hatk \cdot \frac{p_{1}+p_{2}}{L}\right) \right] \,,
-\end{align}
-\end{widetext}
-%
-where we used $p_{2}-p_{1} = -L n_{3}$ and the dependence in $t$ is in the vectors $n_{A}$, $p_{A}$.
-$(p_{2}+p_{1})/2 $ also has a simple expression in the LISA frame.
-Thus the complete result before any further approximation is
-%
-\begin{widetext}
-\begin{subequations}
-\begin{align}
-	\tilde{y}_{21}^{+,\times}(f) &= \sum\limits_{n\in \mathbb{Z}} c_{n}[g_{21}^{+,\times}]\left(f - n f_{0}\right) \tilde{h}_{+,\times}\left(f - n f_{0}\right)\,, \\
-	c_{n}[g_{21}^{+,\times}]\left(f\right) &= \frac{\Omega}{2\pi}\int_{0}^{\frac{2\pi}{\Omega}} \ud t \, e^{i n \Omega t} i\pi L f P_{3}^{+,\times} \sinc\left[ \pi L f(1+\hatk \cdot n_{3})\right]\exp\left[ i \pi f L \left(  1 - \hatk \cdot \frac{p_{1}+p_{2}}{L}\right) \right]  \,.
-\end{align}
-\end{subequations}
-\end{widetext}
-%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection*{Separating the orbital delay from the constellation delay}
-
-Some numerical tests for the coefficients $c_{n}$ on the frequency band $10^{-5}$Hz-$1$Hz show that the extent in $n$ of the frequency comb rises a lot (up to several hundreds) at the high frequency end.
-This seems attributable to the presence of $\hatk\cdot r/c$ in the argument of te exponential and the large number of oscillations this causes.
-Indeed, this contribution is roughly $f R/c$ with $c/R\sim 2.10^{-3}\Hz$, whereas $c/L\sim 2.10^{-1}\Hz$, so that the oscillations due to $L/c$ terms are relatively mild, while those due to the $R/c$ term are much more rapid and yield extended harmonic content.
-
-Luckily, it is possible to factor out this term, and to control it analytically.
-It corresponds to the part of the delay that corresponds to the GW propagating from the solar system barycenter ($\Omega$) to the moving center of the constellation ($O$), and has a simple expression involving only the sine and cosine of $\Omega t$.
-If we define
-%
-\be
-	h_{+,\times}^{O} = h_{+,\times} (t + \hatk \cdot r)
-\ee
-%
-as the signal as measured at the center of the constellation $O$, since $\hatk \cdot r = R \cos \beta \cos(\Omega t - \lambda)$ (choosing $\eta_{0} = 0$), applying the same calculation as above yields for this simplified retardation
-%
-\begin{align}
-	\tilde{h}_{+,\times}^{O}\left(f\right) &= \sum\limits_{n\in \mathbb{Z}} c_{n}[g_{O}^{+,\times}]\left(f - n f_{0}\right) \tilde{h}_{+,\times}\left(f - n f_{0}\right)\,, \\
-	c_{n}[g_{O}^{+,\times}]\left(f\right) &= \frac{\Omega}{2\pi}\int_{0}^{\frac{2\pi}{\Omega}} \ud t \, e^{i n \Omega t} \exp\left[ -2i \pi f R \cos \beta \cos (\Omega t - \lambda) \right] \nn\\ 
-	&= e^{i n \lambda} \frac{1}{2\pi}\int_{0}^{2\pi} \ud u \, e^{i n u} \exp\left[ -2i \pi f R \cos \beta \cos u \right] \nn\\
-	&= i^{n}e^{i n \lambda} J_{n}\left[ -2\pi f R \cos \beta \right] \,.
-\end{align}
-%
-Thus we obtain, in two steps,
-%
-\begin{widetext}
-\begin{subequations}
-\begin{align}
-	\tilde{h}_{+,\times}^{O}\left(f\right) &= \sum\limits_{n\in \mathbb{Z}} i^{n}e^{i n \lambda} J_{n}\left[ -2\pi \left(f - n f_{0}\right) R \cos \beta \right]\tilde{h}_{+,\times}\left(f - n f_{0}\right)\,, \\
-	\tilde{y}_{21}^{+,\times}(f) &= \sum\limits_{n\in \mathbb{Z}} c_{n}[\gamma_{21}^{+,\times}]\left(f - n f_{0}\right) \tilde{h}_{+,\times}^{O}\left(f - n f_{0}\right)\,, \\
-	c_{n}[\gamma_{21}^{+,\times}]\left(f\right) &= \frac{\Omega}{2\pi}\int_{0}^{\frac{2\pi}{\Omega}} \ud t \, e^{i n \Omega t} 2 i\pi L f f_{3}^{+,\times}\sinc\left[ \pi L f (1+\hatk \cdot n_{3})\right]\exp\left[ i \pi f L\left(1 - \hatk \cdot O_{2} \cdot \frac{p_{1}^{L}+p_{2}^{L}}{L} \right) \right] \,.
-\end{align}
-\end{subequations}
-\end{widetext}
-%
-Now we find that the Bessel comb that relates $\tilde{h}^{O}$ to $\tilde{h}$ (which we can control analytically) contains all the spread in $n$, as the new $c_{n}$ coefficients (expressed with integrals) do not extend beyond $n=\pm 10$.
-These comb coefficients show a smooth variation with frequency (they vary much less than the Fourier-domain signal): in the worst case, one should be able to compute them numerically at a handful of frequencies and interpolate inbetween.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection*{Direct summation}
-
-The Newtonian estimates above indicate that the locally linear phase approximation is going to be at least marginally valid for all sources of which we observe the inspiral, merger and ringdown.
-We adopt for the results above the generic notation
-%
-\begin{align}
-	\tilde{s}(f) &=  \sum\limits_{n\in \mathbb{Z}} c_{n}[G]\left(f - n f_{0}\right) \tilde{h}\left(f - n f_{0}\right)\,, \nn\\
-	c_{n}[G](f) &= \frac{1}{T} \int_{0}^{T}\ud t\, e^{in\Omega t} G[f](t)
-\end{align}
-%
-with $T=2\pi/\Omega=1\mathrm{yr}$ and $G[f](t)$ a function of frequency and time.
-
-At the lowest order in the approximation, with $|\delta A/A| \ll 1$ and $|\delta^{2} \Psi | \ll 1$, and also neglecting $f_{0}$ in the frequency-dependence of $G$, we have $\tilde{h}(f-n f_{0}) \simeq \tilde{h}(f) \exp\left[ i n f_{0} \ud \Psi/\ud f \right]$ and (after commuting the sum and integral):
-%
-\be
-	\tilde{s}(f) \simeq \tilde{h}(f) \frac{1}{T} \int_{0}^{T}\ud t\, G[f](t) \sum\limits_{n} e^{i n \Omega t} \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \,.
-\ee
-%
-Using the Dirac comb relation
-%
-\be
-	\sum\limits_{n} e^{i n \Omega (t-t_{0})} = \frac{2\pi}{\Omega} \sum\limits_{n} \delta\left( t-t_{0} - \frac{n\Omega}{2\pi} \right) \,,
-\ee
-%
-we have
-%
-\be
-	\sum\limits_{n} e^{i n \Omega t} \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f }\right] = T \sum\limits_{n} \delta\left( t + \frac{1}{2\pi}\frac{\ud \Psi}{\ud f} - n T \right) \,,
-\ee
-%
-and since only one of the Dirac deltas has support on $[0,T]$, we obtain simply
-%
-\be
-	\tilde{s}(f) \simeq \tilde{h}(f) \times G[f]\left( -\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right)\,.
-\ee
-%
-Now, recall that when the SPA applies (which it does not in the merger-ringdown phase of the signal), we have the relations
-%
-\begin{align}
-	\Psi_{\rm SPA} &= \phi(t_{f}) - 2\pi f t_{f} + \mathrm{const} \,,\nn\\
-	\frac{\ud \Psi_{\rm SPA}}{\ud f} &= -2\pi t_{f} \,,
-\end{align}
-%
-thus providing a simple physical interpretation: at leading order, we can simply use the time-domain modulation and delays, as evaluated at a fixed time $t_{f}$, the time at which the corresponding frequency is dominantly emitted.
-Notice however that the above result is more general, as it extends to the merger-ringdown phase, and that the limits of validity of the two approximations differ: the SPA becomes increasingly valid in the deep inspiral, whereas the LLP breaks down in the limit of almost monochromatic sources but becomes increasingly valid at the end of the inspiral, when the signal evolves quickly compared to the orbital timescale.
-
-Notice that these Newtonian estimates are very poor to cover the merger and ringdown; for instance, above the dominant ringdown frequency the Fourier-domain amplitude drops rapidly in a way that is not captured by the power law above.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection*{Summation of $\tilde{h}_{O}\left(f\right)$}
-
-The expression obtained above for $\tilde{h}_{O}\left(f\right)$ can be summed, either by applying the above general formulas with the function $G[f](t) = \exp\left[ -2i \pi f R \cos \beta \cos (\Omega t - \lambda) \right]$, either directly by means of the Jacobi-Anger expansion:
-%
-\begin{widetext}
-\begin{align}
-	\tilde{h}_{O}\left(f\right) &\simeq \sum\limits_{n\in \mathbb{Z}} i^{n}e^{i n \lambda} J_{n}\left[ -2\pi f R \cos \beta \right] A(f) \exp\left[ -i \Psi \left(f-\frac{n\Omega}{2\pi}\right) \right] \nn \\
-	&\simeq A(f) e^{-i \Psi \left(f\right)} \sum\limits_{n\in \mathbb{Z}} i^{n}e^{i n \lambda} J_{n}\left[ -2\pi f R \cos \beta \right] \exp\left[ i \frac{n\Omega}{2\pi} \frac{\ud \Psi}{\ud f} + \calO(\Omega^{2})  \right]
-\end{align}
-\end{widetext}
-%
-We can then use the Jacobi-Anger expansion (with the notation $z\equiv 2\pi R f \cos\beta$)
-%
-\be
-	e^{i z \cos \theta}= \sum\limits_{n\in \mathbb{Z}} i^{n}J_{n}[z]e^{i n \theta} \,,
-\ee
-%
-to obtain a simple phase correction
-%
-\begin{subequations}
-\begin{align}
-	\tilde{h}_{O}\left(f\right) &\simeq \tilde{h}(f) \exp\left[ -iz \cos\left( \lambda + \frac{\Omega}{2\pi} \frac{\ud \Psi}{\ud f} \right) \right]\,.
-\end{align}
-\end{subequations}
-%
-The corrections beyond the leading order can be derived directly by a kind of Feynman trick, introducing derivatives with respect to the parameter $\lambda$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Accelerated overlap computations for modeled waveforms}
-\label{subsec:overlaps}
-
-[present recursion relations - Fresnel integrals]
-
-[discuss numerical robustness, power-law cut in the space of phase coefficients]
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Summary \& performance}
-\label{subsec:performance}
-
-[give speedup for linear overlap and Fresnel overlap, for 22 only and HM]
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Impact of noise realizations on parameter estimation for LIGO}
-\label{sec:noiseligo}
-
-[no-noise corresponds to a geometric average on the posteriors themselves across noise realizations; how does this translate in practice ?]
-
-[Phil's plot: example of the distribution of the MAP across noise realizations, compared to the posterior itself]
-
-[Phil's plot: example(s) of 100 cumulative distributions compared to the no-noise case]
-
-[evolution of results when increasing SNR: 12, 15, 20]
-
-[table giving the ``averaged'' broadening of the posterior and the ``averaged'' departure of the MAP from injection across noise realizations, for all SNR=12 simulations]
-
-[discussion: validation of the no-noise analysis as a first proxy to do parameter estimation estimation]
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Impact of merger/ringdown and higher modes on parameter estimation for LISA}
-\label{sec:pelisa}
-
-[connection to previous work: McWilliams\&al, Littenberg\&al]
-
-[plot: example posterior with/without HM, EOBNRv2HMROM injection]
-
-[plot: example posterior with/without merger/ringdown, EOBNRv2HMROM injection]
-
-[plot: posterior with information contained in the inspiral only: inspiral-only injection and templates, vs IMR injection and template]
-
-[discuss multimodal structure of the posterior (sky position notably), if it holds]
-
-[improvement on ]
-
-[sky localization errors across the sky with full bayesian inference on a few example cases: a sufficiently large number of inferences for this to be interesting ?]
-
-[accumulation of information with time ? - future studies]
-
-[compare to Fisher matrix analysis on some examples ?]
+[show summary of amplitude/phase errors on the example case]
+[show mismatch calculations]
 
@@
 
-[extension to longer waveforms, lower masses]
-
-[use of other fast FD waveforms: SEOBNRv2ROM, PhenomP]
-
-[impact of aligned spins ? of simple precession ?]
-
-[impact of merger-ringdown in the aligned spins case and simple precession case ?]
-
-[validation of existing Fischer matrix analysis, at least on some examples, with full Bayesian inference]
-
-[compare different instrument designs, across astrophysical models (Sesana-Barausse\&al)]
-
-[extend analysis to a full population ?]
+[]
 
@@
 
-\section{Details on Fourier-domain reduced order models}
-\label{app:rom}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Details on the Fourier-domain LISA response}
-\label{app:fdresponse}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{TDI observables}
-\label{appsubsec:TDI}
-
-For the constellation delay and modulation, if we associate as above a function $G_{AB}$ to each basic Doppler shift observable $y_{AB}$ (as with the example of $y_{21}$ above), for each of the polarizations we obtain:
-%
-\begin{align}
-	G_{21}^{+,\times} &= i\pi f L \left(n_{3}\cdot P_{+,\times}\cdot n_{3}\right) \sinc\left[ \pi f L \left( 1 + k\cdot n_{3} \right) \right] \exp\left[ i\pi f L \left( 1 - k\cdot (p_{1}+p_{2})/L  \right) \right] \nn\\
-	G_{12}^{+,\times} &= i\pi f L \left(n_{3}\cdot P_{+,\times}\cdot n_{3}\right) \sinc\left[ \pi f L \left( 1 - k\cdot n_{3} \right) \right] \exp\left[ i\pi f L \left( 1 - k\cdot (p_{1}+p_{2})/L  \right) \right] \nn\\
-	G_{32}^{+,\times} &= i\pi f L \left(n_{1}\cdot P_{+,\times}\cdot n_{1}\right) \sinc\left[ \pi f L \left( 1 + k\cdot n_{1} \right) \right] \exp\left[ i\pi f L \left( 1 - k\cdot (p_{2}+p_{3})/L  \right) \right] \nn\\
-	G_{23}^{+,\times} &= i\pi f L \left(n_{1}\cdot P_{+,\times}\cdot n_{1}\right) \sinc\left[ \pi f L \left( 1 - k\cdot n_{1} \right) \right] \exp\left[ i\pi f L \left( 1 - k\cdot (p_{2}+p_{3})/L  \right) \right] \nn\\
-	G_{13}^{+,\times} &= i\pi f L \left(n_{2}\cdot P_{+,\times}\cdot n_{2}\right) \sinc\left[ \pi f L \left( 1 + k\cdot n_{2} \right) \right] \exp\left[ i\pi f L \left( 1 - k\cdot (p_{3}+p_{1})/L  \right) \right] \nn\\
-	G_{31}^{+,\times} &= i\pi f L \left(n_{2}\cdot P_{+,\times}\cdot n_{2}\right) \sinc\left[ \pi f L \left( 1 - k\cdot n_{2} \right) \right] \exp\left[ i\pi f L \left( 1 - k\cdot (p_{3}+p_{1})/L  \right) \right] \,,
-\end{align}
-%
-where the $p_{A}$ vectors refer to the positions with respect to the center of the constellation $O$ (that is to say, they are given by $O_{2}\cdot p_{A}^{L}$).
-
-The expression for the $X_{A}$ second-generation TDI observables are given by (c.f. Kr\'olak\&al)
-%
-\begin{widetext}
-\begin{align}
-	X_{1} &= \left[ \left( y_{31} + y_{13,2} \right) - \left( y_{21} + y_{12,3} \right) - \left( y_{31} + y_{13,2} \right)_{,33} + \left( y_{21} + y_{12,3} \right)_{,22} \right] \nn\\
-	& - \left[ \left( y_{31} + y_{13,2} \right) - \left( y_{21} + y_{12,3} \right) - \left( y_{31} + y_{13,2} \right)_{,33} + \left( y_{21} + y_{12,3} \right)_{,22} \right]_{,2233} \,.
-\end{align}
-\end{widetext}
-%
-Within the rigid approximation, all the delays are considered equal to $L/c$ as far as the gravitational wave response is concerned.
-We are applying successive constant delays to the basic observables, and the translation in Fourier domain is simply
-%
-\begin{widetext}
-\begin{align}
-	\tilde{X}_{1}(f) &= -4 \sin\left( 2\pi f L \right) \sin\left( 4\pi f L \right) e^{6 i \pi f L} \left[ \tilde{y}_{31} - \tilde{y}_{21} + e^{2i\pi f L} \left( \tilde{y}_{13} - \tilde{y}_{12} \right) \right] \,, \nn\\
-	\tilde{X}_{2}(f) &= -4 \sin\left( 2\pi f L \right) \sin\left( 4\pi f L \right) e^{6 i \pi f L} \left[ \tilde{y}_{12} - \tilde{y}_{32} + e^{2i\pi f L} \left( \tilde{y}_{21} - \tilde{y}_{23} \right) \right] \,, \nn\\
-	\tilde{X}_{3}(f) &= -4 \sin\left( 2\pi f L \right) \sin\left( 4\pi f L \right) e^{6 i \pi f L}\left[ \tilde{y}_{23} - \tilde{y}_{13} + e^{2i\pi f L} \left( \tilde{y}_{32} - \tilde{y}_{31} \right) \right] \,.
-\end{align}
-\end{widetext}
-%
-
-For the Sagnac observables, we have similarly ((8) of Shaddock\&al)
-%
-\begin{align}
-	\alpha_{1} &= \left[ \left( y_{31} + y_{23,2} + y_{12,12} \right) - \left( y_{21} + y_{32,3} + y_{13,13} \right) \right] \nn\\
-	& - \left[ \left( y_{31} + y_{23,2} + y_{12,12} \right)_{,213} - \left( y_{21} + y_{32,3} + y_{13,13} \right)_{,312} \right]
-\end{align}
-%
-and
-%
-\begin{widetext}
-\begin{align}
-	\tilde{\alpha}_{1}(f) &= - 2i \sin\left(3\pi f L\right) e^{3 i \pi f L} \left[ \tilde{y}_{31} - \tilde{y}_{21} + e^{2i\pi f L} \left( \tilde{y}_{23} - \tilde{y}_{32} \right) + e^{4i\pi f L} \left( \tilde{y}_{12} - \tilde{y}_{13} \right) \right] \,, \nn\\
-	\tilde{\alpha}_{2}(f) &= - 2i \sin\left(3\pi f L\right) e^{3 i \pi f L} \left[ \tilde{y}_{12} - \tilde{y}_{32} + e^{2i\pi f L} \left( \tilde{y}_{31} - \tilde{y}_{13} \right) + e^{4i\pi f L} \left( \tilde{y}_{23} - \tilde{y}_{21} \right) \right] \,, \nn\\
-	\tilde{\alpha}_{3}(f) &= - 2i \sin\left(3\pi f L\right) e^{3 i \pi f L} \left[ \tilde{y}_{23} - \tilde{y}_{13} + e^{2i\pi f L} \left( \tilde{y}_{12} - \tilde{y}_{21} \right) + e^{4i\pi f L} \left( \tilde{y}_{31} - \tilde{y}_{32} \right) \right] \,.
-\end{align}
-\end{widetext}
-%
-The optimal second-generation combinations $A$, $E$ and $T$ are given by ((57) of Kr\'olak\&al)
-%
-\begin{align}
-	A &= \frac{1}{\sqrt{2}} \left( \alpha_{3} - \alpha_{1} \right) \nn\\
-	E &= \frac{1}{\sqrt{6}} \left( \alpha_{1} - 2\alpha_{2} + \alpha_{3} \right) \nn\\
-	T &= \frac{1}{\sqrt{3}} \left( \alpha_{1} + \alpha_{2} + \alpha_{3} \right) \,,
-\end{align}
-%
-which gives
-%
-\begin{widetext}
-\begin{align}
-	\tilde{A}(f) &= - \frac{2i}{\sqrt{2}} \sin\left(3\pi f L\right) e^{3 i \pi f L} \left[ \left( \tilde{y}_{13} + \tilde{y}_{31} \right) \left( e^{4i\pi f L} - 1 \right) + \left( \tilde{y}_{21} + \tilde{y}_{23} \right) \left( 1 - e^{2i\pi f L} \right) \right.
-\nn\\
-	& \qquad\qquad\qquad\qquad\qquad\qquad \left. + \left( \tilde{y}_{32} + \tilde{y}_{12} \right) \left( e^{2i\pi f L} - e^{4i\pi f L} \right) \right] \nn\\
-	\tilde{E}(f) &= - \frac{2i}{\sqrt{6}} \sin\left(3\pi f L\right) e^{3 i \pi f L} \left[ \left( \tilde{y}_{23} - \tilde{y}_{21} \right) \left( 1 + e^{2i\pi f L} - 2e^{4i \pi f L} \right) + \left( \tilde{y}_{31} - \tilde{y}_{13} \right) \left( 1 - 2e^{2i\pi f L} + e^{4i \pi f L} \right) \right.
-\nn\\
-	& \qquad\qquad\qquad\qquad\qquad\qquad + \left( \tilde{y}_{12} - \tilde{y}_{32} \right) \left( -2 + e^{2i\pi f L} + e^{4i \pi f L} \right) \left. \right] \nn\\
-	\tilde{T}(f) &= - \frac{2i}{\sqrt{3}} \sin\left(3\pi f L\right) e^{3 i \pi f L} \left( \tilde{y}_{31} - \tilde{y}_{13} + \tilde{y}_{12} - \tilde{y}_{21} + \tilde{y}_{23} - \tilde{y}_{32} \right) \left( 1 + e^{2i\pi f L} + e^{4i \pi f L} \right) \,.
-\end{align}
-\end{widetext}
-%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Beyond the leading-order Fourier-domain response}
-\label{appsubsec:nloresponse}
-
-We derive in the following analytical corrections to this leading order formula. In practice, we expect that the first and second correction terms, at the very most, will be needed.
-
-\subsubsection*{Frequency dependence of the modulation/delay}
-
-If we want to keep the full frequency dependence of the modulation/delay function $G$ in the sum, then we can expand
-%
-\be
-	G[f-n f_{0}](t) = \sum\limits_{k\geq 0} \frac{1}{k!}\left( \frac{-n \Omega}{2\pi} \right)^{k} \frac{\partial^{k} G}{\partial f^{k}}[f](t) \,,
-\ee
-%
-and we have
-%
-\begin{widetext}
-\begin{align}
-	\tilde{s}(f) &\simeq \tilde{h}(f) \frac{1}{T} \int_{0}^{T}\ud t\, \sum\limits_{n}\sum\limits_{k} e^{i n \Omega t} \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \frac{1}{k!}\left( \frac{-n \Omega}{2\pi} \right)^{k} \frac{\partial^{k} G}{\partial f^{k}}[f](t) \nn\\
-	&= \tilde{h}(f) \sum\limits_{n}\sum\limits_{k} \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \frac{1}{k!} \frac{1}{(2i\pi)^{k}} \frac{1}{T} \int_{0}^{T}\ud t\, (-i n \Omega)^{k}  e^{i n \Omega t}  \frac{\partial^{k} G}{\partial f^{k}}[f](t) \nn\\
-	&= \tilde{h}(f) \sum\limits_{n}\sum\limits_{k} \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \frac{1}{k!} \frac{1}{(2i\pi)^{k}} \frac{1}{T} \int_{0}^{T}\ud t\, e^{i n \Omega t}  \frac{\partial^{2k} G}{\partial f^{k}\partial t^{k}}[f](t) \nn\\
-	&= \tilde{h}(f) \sum\limits_{k\geq 0} \frac{1}{k!} \frac{1}{(2i\pi)^{k}}  \frac{\partial^{2k} G}{\partial f^{k}\partial t^{k}}[f]\left(-\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) \,,
-\end{align}
-\end{widetext}
-%
-where we performed $k$ integrations by parts (using the periodicity of $G$ for any frequency) and summed over $n$ in the same way as in the leading order calculation.
-
-\subsubsection*{Quadratic term in the phase}
-
-If we keep the first correction in the phase beyond the LLP, the quadratic term $\delta^{2}\Psi$, but ignore all the subsequent terms, we can Taylor-expand the exponential according to
-%
-\begin{widetext}
-\begin{align}
-	\tilde{h}(f-n f_{0}) &\simeq \tilde{h}(f)  \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \exp\left[ -\frac{1}{2} i (n f_{0})^{2} \frac{\ud^{2}\Psi}{\ud f^{2}} \right] \nn\\
-	&= \tilde{h}(f)  \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \sum\limits_{k\geq 0} \frac{1}{k!}\left( \frac{-i}{2} \right)^{k} \left( \frac{n\Omega}{2\pi} \right)^{2k} \left( \frac{\ud^{2}\Psi}{\ud f^{2}}  \right)^{k} \,,
-\end{align}
-\end{widetext}
-% 
-and we have
-%
-\begin{widetext}
-\begin{align}
-	\tilde{s}(f) &\simeq \tilde{h}(f) \frac{1}{T} \int_{0}^{T}\ud t\, \sum\limits_{n}\sum\limits_{k} e^{i n \Omega t} \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \frac{1}{k!} \left( \frac{-i}{2} \right)^{k} \left( \frac{n\Omega}{2\pi} \right)^{2k} \left( \frac{\ud^{2}\Psi}{\ud f^{2}}  \right)^{k} G[f](t) \nn\\
-	&= \tilde{h}(f) \sum\limits_{n}\sum\limits_{k} \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \frac{1}{k!} \left( \frac{i}{2} \right)^{k} \frac{1}{(2\pi)^{2k}} \left( \frac{\ud^{2}\Psi}{\ud f^{2}}  \right)^{k} \frac{1}{T} \int_{0}^{T}\ud t\, (i n \Omega)^{2 k}  e^{i n \Omega t} G[f](t) \nn\\
-	&= \tilde{h}(f) \sum\limits_{k\geq 0} \frac{1}{2^k k!} \left( \frac{i}{4\pi^{2}}\frac{\ud^{2}\Psi}{\ud f^{2}}  \right)^{k} \frac{\partial^{k} G}{\partial t^{k}}[f]\left(-\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) \,.
-\end{align}
-\end{widetext}
-%
-
-\subsubsection*{Linear term in the amplitude}
-
-If we keep the first correction in the amplitude, we have
-%
-\be
-	\tilde{h}(f-n f_{0}) \simeq \tilde{h}(f)  \exp\left[ i n f_{0} \frac{\ud \Psi}{\ud f} \right] \left( 1 - \frac{n \Omega}{2\pi} \frac{1}{A}\frac{\ud A}{\ud f} \right) \,,
-\ee
-%
-and we obtain similarly
-%
-\be
-	\tilde{s}(f) \simeq \tilde{h}(f) \left[ G[f]\left(-\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) + \frac{1}{2 i \pi} \frac{1}{A} \frac{\partial G}{\partial t}[f]\left(-\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) \right]\,.
-\ee
-%
+\section{Appendix}
+\label{app:}
 
@@
 %\bibliography{ListeRef.bib}
-\bibliography{/Users/marsat/Documents/publications/bibliographie/ListeRef.bib}
+\bibliography{references.bib}
 

### 2016-11-10 EOBNRv2HMROM/GenerateWaveform.c

@@ -6,6 +6,7 @@
 static GenWavetag ParseGenWavetag(char* string) {
   GenWavetag tag;
   if(strcmp(string, "hlm")==0) tag = hlm;
+  else if(strcmp(string, "h22TD")==0) tag = h22TD;
   else if(strcmp(string, "hphcFD")==0) tag = hphcFD;
   else if(strcmp(string, "hphcTD")==0) tag = hphcTD;
   else {
@@ -42,11 +43,14 @@ Arguments are as follows:\n\
 ----- Generation Parameters ----------------------\n\
 --------------------------------------------------\n\
  --nbmode              Number of modes of radiation to generate (1-5, default=5)\n\
- --minf                Minimal frequency (Hz, default=0) - when too low, use first frequency covered by the ROM\n\
+ --minf                Minimal frequency, ignore if 0 (Hz, default=0) - will use first frequency covered by the ROM if higher\n\
+ --maxf                Maximal frequency, ignore if 0 (Hz, default=0) - will use last frequency covered by the ROM if lower\n\
  --deltatobs           Observation duration (years, default=2)\n\
  --tagextpn            Tag to allow PN extension of the waveform at low frequencies (default=1)\n\
  --Mfmatch             When PN extension allowed, geometric matching frequency: will use ROM above this value. If <=0, use ROM down to the lowest covered frequency (default=0.)\n\
- --taggenwave          Tag choosing the wf format: hlm (default: downsampled modes in Amp/Phase form), hphcFD (hlm interpolated and summed), hphcTD (IFFT of hphcFD),\n\
+ --taggenwave          Tag choosing the wf format: hlm (default: downsampled modes, Amp/Phase form),  h22TD (IFFT of h22, Amp/Phase form - currently not supported for higher modes), hphcFD (hlm interpolated and summed, Re/Im form), hphcTD (IFFT of hphcFD, Re/Im form)\n\
+ --f1windowend         If generating h22TD/hphcTD, start frequency for windowing at the end - set to 0 to ignore and use 0.995*fHighROM, where fHighROM is the highest frequency covered by the ROM (Hz, default=0)\n\
[…]

### 2016-11-10 EOBNRv2HMROM/GenerateWaveform.h

@@ -43,9 +43,10 @@
 /********************************** Structures ******************************************/
 
 typedef enum GenWavetag {
-  hphcTD,
+  hlm,
+  h22TD,
   hphcFD,
-  hlm
+  hphcTD
 } GenWavetag;
 
 /* Parameters for the generation of a LISA waveform (in the form of a list of modes) */
@@ -57,12 +58,15 @@ typedef struct tagGenWaveParams {
   double m2;                 /* mass of companion 2 (solar masses, default 1e6) */
   double distance;           /* distance of source (Mpc, default 1e3) */
   double inclination;        /* inclination of source (rad, default pi/3) */
-  double minf;               /* Minimal frequency (Hz) - when set to 0 (default), use the first frequency covered by the ROM */
+  double minf;               /* Minimal frequency, ignore if 0 (Hz, default=0) - will use first frequency covered by the ROM if higher */
+  double maxf;               /* Maximal frequency, ignore if 0 (Hz, default=0) - will use last frequency covered by the ROM if lower */
   double deltatobs;          /* Observation duration (years, default=2) */
   int tagextpn;              /* Tag to allow PN extension of the waveform at low frequencies */
   double Mfmatch;            /* When PN extension allowed, geometric matching frequency: will use ROM above this value. If <=0, use ROM down to the lowest covered frequency */
   int nbmode;                /* number of modes to generate (starting with 22) - defaults to 5 (all modes) */
   int taggenwave;            /* Tag selecting the desired output format */
+  double f1windowend;        /* If generating hphcTD, start frequency for windowing at the end - set to 0 to ignore and use 0.995*fHighROM, where fHighROM is the highest frequency covered by the ROM (Hz, default=0) */
+  double f2windowend;        /* If generating hphcTD, stop frequency for windowing at the end - set to 0 to ignore and use fHighROM, where fHighROM is the highest frequency covered by the ROM (Hz, default=0) */
   int binaryout;             /* Tag for outputting the data in gsl binary form instead of text (default 0) */
   char outdir[256];          /* Path for the output directory */
   char outfile[256];         /* Path for the output file */

### 2016-11-10 LISAinference/LISAinference.c

@@ -328,7 +328,7 @@ int main(int argc, char *argv[])
 	    /* Phys live points */
 	    fprintf(fphyslivepoints, "    %.18E", injectedparams->m1);
 	    fprintf(fphyslivepoints, "    %.18E", injectedparams->m2);
-	    fprintf(fphyslivepoints, "    %.18E", 0.); /* For templates, tRef is defined relatively to the injected value */
+	    fprintf(fphyslivepoints, "    %.18E", injectedparams->tRef);
 	    fprintf(fphyslivepoints, "    %.18E", injectedparams->distance);
 	    fprintf(fphyslivepoints, "    %.18E", injectedparams->phiRef);
 	    fprintf(fphyslivepoints, "    %.18E", injectedparams->inclination);

### 2016-11-10 LISAinference/LISAutils.c

@@ -910,11 +910,13 @@ int LISAGenerateSignalReIm(
 
   /* Compute the Re/Im frequency series - takes into account the length of the observation with deltatobs */
   double fstartobs = Newtonianfoft(params->m1, params->m2, globalparams->deltatobs);
+  double fLow = fmax(__LISASimFD_Noise_fLow, globalparams->minf);
+  double fHigh = fmin(__LISASimFD_Noise_fHigh, globalparams->maxf);
   //TESTING
   //tbeg = clock();
-  ReImFrequencySeries_SumListmodesCAmpPhaseFrequencySeries(TDI1, listTDI1, freq, fstartobs);
-  ReImFrequencySeries_SumListmodesCAmpPhaseFrequencySeries(TDI2, listTDI2, freq, fstartobs);
-  ReImFrequencySeries_SumListmodesCAmpPhaseFrequencySeries(TDI3, listTDI3, freq, fstartobs);
+  ReImFrequencySeries_SumListmodesCAmpPhaseFrequencySeries(TDI1, listTDI1, freq, fLow, fHigh, fstartobs);
+  ReImFrequencySeries_SumListmodesCAmpPhaseFrequencySeries(TDI2, listTDI2, freq, fLow, fHigh, fstartobs);
+  ReImFrequencySeries_SumListmodesCAmpPhaseFrequencySeries(TDI3, listTDI3, freq, fLow, fHigh, fstartobs);
   //tend = clock();
   //printf("time ReIm: %g\n", (double) (tend-tbeg)/CLOCKS_PER_SEC);
   //
@@ -996,9 +998,9 @@ int LISAGenerateInjectionReIm(
   /* Compute the Re/Im frequency series */
   //TESTING
   //tbeg = clock();
-  ReImFrequencySeries_SumListmodesCAmpPhaseFrequencySeries(TDI1, listTDI1, freq, fstartobs);
-  ReImFrequencySeries_SumListmodesCAmpPhaseFrequencySeries(TDI2, listTDI2, freq, fstartobs);
-  ReImFrequencySeries_SumListmodesCAmpPhaseFrequencySeries(TDI3, listTDI3, freq, fstartobs);
+  ReImFrequencySeries_SumListmodesCAmpPhaseFrequencySeries(TDI1, listTDI1, freq, fLow, fHigh, fstartobs);
[…]

### 2016-11-10 LISAsim/GenerateTDIFD.c

@@ -359,8 +359,9 @@ int main(int argc, char *argv[])
     SetCoeffsG(params->lambda, params->beta, params->polarization);
 
     /* Starting frequency - takes into account the duration of observation deltatobs and the arg minf */
-    double fstartobs = Newtonianfoft(params->m1, params->m2, params->deltatobs);
-    double flow = fmax(params->minf, fstartobs);
+    double fstartobs = 0.;
+    if(!(params->deltatobs==0)) fstartobs = Newtonianfoft(params->m1, params->m2, params->deltatobs);
+    double fLow = fmax(params->minf, fstartobs);
 
     /* Generate Fourier-domain waveform as a list of hlm modes */
     /* Use TF2 extension, if required to, to arbitrarily low frequencies */
@@ -371,7 +372,7 @@ int main(int argc, char *argv[])
       SimEOBNRv2HMROM(&listROM, params->nbmode, params->tRef, params->phiRef, params->fRef, (params->m1)*MSUN_SI, (params->m2)*MSUN_SI, (params->distance)*1e6*PC_SI);
     } else {
       //printf("Extending signal waveform.  mfmatch=%g\n",globalparams->mfmatch);
-      SimEOBNRv2HMROMExtTF2(&listROM, params->nbmode, params->Mfmatch, flow, 0, params->tRef, params->phiRef, params->fRef, (params->m1)*MSUN_SI, (params->m2)*MSUN_SI, (params->distance)*1e6*PC_SI);
+      SimEOBNRv2HMROMExtTF2(&listROM, params->nbmode, params->Mfmatch, fLow, 0, params->tRef, params->phiRef, params->fRef, (params->m1)*MSUN_SI, (params->m2)*MSUN_SI, (params->distance)*1e6*PC_SI);
     }
 
     /* Process through the Fourier-domain response for TDI observables */
@@ -413,27 +414,23 @@ int main(int argc, char *argv[])
       double deltaf = 0.;
       if(params->deltaf==0.) {
         /* Determine deltaf so that N deltat = 1/deltaf > 2*tc where tc is the time to coalescence estimated from Psi22 */
-        /* Assumes the TD waveform ends around t=0 */
-        /* Note: the phase is untouched by the response processing, so the phase in the 22 TDI contribution is still Psi22 and the same for all channels */
-        double tc = EstimateInitialTime(listTDI1, flow);
-        deltaf = 0.5 * 1./(2*(-tc)); /* Extra factor of 1/2 corresponding to 0-padding in TD by factor of 2 */
-        //
[…]

### 2016-11-10 LISAsim/GenerateTDITD.c

@@ -60,13 +60,13 @@ Arguments are as follows:\n\
         } else if (strcmp(argv[i], "--polarization") == 0) {
             params->polarization = atof(argv[++i]);
         } else if (strcmp(argv[i], "--tagtdi") == 0) {
-	  params->tagtdi = ParseTDItag(argv[++i]);
+	          params->tagtdi = ParseTDItag(argv[++i]);
         } else if (strcmp(argv[i], "--nsamplesinfile") == 0) {
             params->nsamplesinfile = atoi(argv[++i]);
         } else if (strcmp(argv[i], "--binaryin") == 0) {
-          params->binaryin = 1;
+            params->binaryin = 1;
         } else if (strcmp(argv[i], "--binaryout") == 0) {
-          params->binaryout = 1;
+            params->binaryout = 1;
         } else if (strcmp(argv[i], "--indir") == 0) {
             strcpy(params->indir, argv[++i]);
         } else if (strcmp(argv[i], "--infile") == 0) {
@@ -114,7 +114,30 @@ static void Read_Wave_hphcTD( RealTimeSeries** hptd, RealTimeSeries** hctd, cons
   /* Clean up */
   gsl_matrix_free(inmatrix);
 }
-/* Output waveform in downsampled form, FD Amp/Pase, all hlm modes in a single file */
+/* Read waveform time series in Amp/Phase form for a single mode h22TD */
+/* NOTE: assumes a single mode amp/phase in the input file */
+static void Read_Wave_h22TD( RealTimeSeries** hptd, const char dir[], const char file[], const int nsamples, int binary)
+{
+  /* Initalize and read input */
+  gsl_matrix* inmatrix =  gsl_matrix_alloc(nsamples, 3);
+  if (!binary) Read_Text_Matrix(dir, file, inmatrix);
+  else Read_Matrix(dir, file, inmatrix);
+
+  /* Initialize structures */
+  AmpPhaseTimeSeries_Init(h22td, nsamples);
+
+  /* Set values */
+  gsl_vector_view timesview = gsl_matrix_column(inmatrix, 0);
+  gsl_vector_view ampview = gsl_matrix_column(inmatrix, 1);
+  gsl_vector_view phaseview = gsl_matrix_column(inmatrix, 2);
+  gsl_vector_memcpy((*h22td)->times, &timesview.vector);
[…]

### 2016-11-10 LISAsim/LISAFDresponse.c

@@ -204,7 +204,7 @@ int LISASimFDResponsey12(
       f = gsl_vector_get(freq, j);
       tf = (gsl_spline_eval_deriv(spline_phi, f, accel_phi))/(2*PI);
       //clock_t tbegGAB = clock();
-      EvaluateGABmode(&g12mode, &g21mode, &g23mode, &g32mode, &g31mode, &g13mode, f, tf, Yfactorplus, Yfactorcross, 1);
+      EvaluateGABmode(&g12mode, &g21mode, &g23mode, &g32mode, &g31mode, &g13mode, f, tf, Yfactorplus, Yfactorcross, 1); /* does include the R-delay term */
       //clock_t tendGAB = clock();
       //timingcumulativeGABmode += (double) (tendGAB-tbegGAB) /CLOCKS_PER_SEC;
       /**/

### 2016-11-10 LISAsim/LISAgeometry.c

@@ -93,7 +93,9 @@ static double sinarray[4];
 /* Function to convert string input TDI string to TDItag */
 TDItag ParseTDItag(char* string) {
   TDItag tag;
-  if(strcmp(string, "y12")==0) tag = y12;
+  if(strcmp(string, "delayO")==0) tag = delayO;
+  else if(strcmp(string, "y12L")==0) tag = y12L;
+  else if(strcmp(string, "y12")==0) tag = y12;
   else if(strcmp(string, "TDIXYZ")==0) tag = TDIXYZ;
   else if(strcmp(string, "TDIalphabetagamma")==0) tag = TDIalphabetagamma;
   else if(strcmp(string, "TDIAETXYZ")==0) tag = TDIAETXYZ;
@@ -877,6 +879,34 @@ int RestoreInPlaceScaledFactorTDI(
 
 /*********************** Time-domain response ************************/
 
+/* Processing single mode in amp/phase form through orbital time delay */
+double hOTDAmpPhase(
+  double* amp,                             /* Output: amplitude */
+  double* phase,                           /* Output: phase */
+  gsl_spline* splineamp,                   /* Input spline for TD mode amplitude */
+  gsl_spline* splinephase,                 /* Input spline for TD mode phase */
+  gsl_interp_accel* accelamp,              /* Accelerator for amp spline */
+  gsl_interp_accel* accelphase,            /* Accelerator for phase spline */
+  const double t)                          /* Time */
+{
+  /* Precompute array of sine/cosine */
+  for(int j=0; j<4; j++) {
+    cosarray[j] = cos((j+1) * Omega_SI * t);
+    sinarray[j] = sin((j+1) * Omega_SI * t);
+  }
+  /* Scalar product k.R */
+  double kR = coeffkRconst;
+  for(int j=0; j<2; j++) {
+    kR += cosarray[j] * coeffkRcos[j] + sinarray[j] * coeffkRsin[j];
+  }
+  /* Common factor and delay */
+  double delay = -(kR*R_SI)/C_SI;
+
+  /* Output result */
+  *amp = gsl_spline_eval(splineamp, t+delay, accelamp);
+  *phase = gsl_spline_eval(splinephase, t+delay, accelphase);
+}
+
 /* Functions evaluating yAB observables in time domain */
 double y12TD(
   gsl_spline* splinehp,                    /* Input spline for TD hplus */
[…]

### 2016-11-10 LISAsim/LISAgeometry.h

@@ -47,7 +47,9 @@ extern "C" {
 /**************** TDI variables **********************/
 
 typedef enum TDItag {
-  y12,
+  delayO,                  /* Orbital delay */
+  y12L,                    /* Constellation-only response y12L */
+  y12,                     /* Complete response y12 (includees orbital delay) */
   TDIXYZ,
   TDIalphabetagamma,
   TDIAETXYZ,
@@ -137,6 +139,18 @@ int RestoreInPlaceScaledFactorTDI(
   int nchannel);                                  /* TDI channel number */
 
 /* Functions for the response in time domain */
+
+/* Processing single mode in amp/phase form through orbital time delay */
+double hOTDAmpPhase(
+  double* amp,                             /* Output: amplitude */
+  double* phase,                           /* Output: phase */
+  gsl_spline* splineamp,                   /* Input spline for TD mode amplitude */
+  gsl_spline* splinephase,                 /* Input spline for TD mode phase */
+  gsl_interp_accel* accelamp,              /* Accelerator for amp spline */
+  gsl_interp_accel* accelphase,            /* Accelerator for phase spline */
+  const double t);                         /* Time */
+
+/* Basic yslr observables (including orbital delay) from hplus, hcross */
 double y12TD(
   gsl_spline* splinehp,                    /* Input spline for TD hplus */
   gsl_spline* splinehc,                    /* Input spline for TD hcross */
@@ -173,7 +187,8 @@ double y13TD(
   gsl_interp_accel* accelhp,               /* Accelerator for hp spline */
   gsl_interp_accel* accelhc,               /* Accelerator for hc spline */
   const double t);                         /* Time */
-int EvaluateTDIXYZTD(
+/* TDI observables (including orbital delay) from hplus, hcross */
+int EvaluateTDIXYZTDhphc(
   double* TDIX,                            /* Output: value of TDI observable X */
   double* TDIY,                            /* Output: value of TDI observable Y */
   double* TDIZ,                            /* Output: value of TDI observable Z */
@@ -182,7 +197,7 @@ int EvaluateTDIXYZTD(
   gsl_interp_accel* accelhp,               /* Accelerator for hp spline */
   gsl_interp_accel* accelhc,               /* Accelerator for hc spline */
   const double t);                         /* Time */
-int EvaluateTDIAETXYZTD(
+int EvaluateTDIAETXYZTDhphc(
   double* TDIA,                            /* Output: value of TDI observable X */
[…]

### 2016-11-10 LLVsim/GenerateLLVFD.c

@@ -317,19 +317,27 @@ int main(int argc, char *argv[])
 
     else {
 
-      /* Determine deltaf so that N deltat = 1/deltaf > 2*tc where tc is the time to coalescence estimated from Psi22 */
-      /* Assumes the TD waveform ends around t=0 */
-      /* Note: the phase differs for the three detectors, and the signal can be shifted in time at worst by ~40 ms - should be ok as we take some margin */
-      double tc = EstimateInitialTime(listLLV1, params->minf);
-      double deltaf = 0.5 * 1./(2*(-tc)); /* Extra factor of 1/2 corresponding to 0-padding in TD by factor of 2 */
+      double deltaf = 0.;
+      if(params->deltaf==0.) {
+        /* Determine deltaf so that N deltat = 1/deltaf > 2*tc where tc is the time to coalescence estimated from Psi22 */
+        /* NOTE: assumes the TD waveform ends around t=0 */
+        /* NOTE: estimate based on the 22 mode - fstartobs will be scaled from mode to mode to ensure the same deltatobs for all (except for the 21 mode, which will turn on after the others) */
+        /* NOTE: the phase differs for the three detectors, and the signal can be shifted in time at worst by ~40 ms - should be ok as we take some margin */
+        double tc = EstimateInitialTime(listROM, params->minf);
+        double deltaf = 0.5 * 1./(2*(-tc)); /* Extra factor of 1/2 corresponding to 0-padding in TD by factor of 2 */
+      }
+      else deltaf = params->deltaf;
+
+      /* Compute TDI FD */
+      /* NOTE: we do note use deltatobs and fstartobs */
 
       /* Generate FD frequency series by summing mode contributions */
       ReImFrequencySeries* LLV1FD = NULL;
       ReImFrequencySeries* LLV2FD = NULL;
       ReImFrequencySeries* LLV3FD = NULL;
-      GenerateFDReImFrequencySeries(&LLV1FD, listLLV1, params->minf, deltaf, 0, params->nbmode); /* Here determines the number of points from deltaf and max frequency in input list of modes */
-      GenerateFDReImFrequencySeries(&LLV2FD, listLLV2, params->minf, deltaf, 0, params->nbmode); /* Here determines the number of points from deltaf and max frequency in input list of modes */
[…]

### 2016-11-10 tools/fft.c

@@ -128,9 +128,9 @@ int FFTTimeSeries(ReImFrequencySeries** freqseries, RealTimeSeries* timeseries, 
 }
 
 /* IFFT of frequency series */
-/* Note: assumes frequency series is FT of real data */
+/* Note: assumes frequency series is FT of real data - produces real time series */
 /* Note: FFT uses flipped convention (i.e. h(f) = int e^(+2ipift)h(t)) */
-int IFFTFrequencySeries(RealTimeSeries** timeseries, ReImFrequencySeries* freqseries, double f1windowbeg, double f2windowbeg, double f1windowend, double f2windowend, int nzeropad)
+int IFFTFrequencySeriesReal(RealTimeSeries** timeseries, ReImFrequencySeries* freqseries, double f1windowbeg, double f2windowbeg, double f1windowend, double f2windowend, int nzeropad)
 {
   /* deltaf of frequency series */
   /* Warning: assumes linear sampling in frequency */
@@ -154,15 +154,17 @@ int IFFTFrequencySeries(RealTimeSeries** timeseries, ReImFrequencySeries* freqse
   double* hrealval = hrealvalues->data;
   double* himagval = himagvalues->data;
   for (int i=0; i<n; i++) {
-    hrealval[i] = WindowFunction(freq[i], f1windowbeg, f2windowend, deltafwindowbeg, deltafwindowend) * hreal[i];
-    himagval[i] = WindowFunction(freq[i], f1windowbeg, f2windowend, deltafwindowbeg, deltafwindowend) * himag[i];
+    double window = WindowFunction(freq[i], f1windowbeg, f2windowend, deltafwindowbeg, deltafwindowend);
+    hrealval[i] = window * hreal[i];
+    himagval[i] = window * himag[i];
   }
 
   /* Input as array of fftw_complex */
   int N = (int) hrealvalues->size;
   fftw_complex* in = (fftw_complex*) fftw_malloc(sizeof(fftw_complex) * N);
+  /* NOTE: Restoring the standard sign convention for the FT, used by FFTW - change in convention amounts to f->-f, equivalent to a conjugation only for FFT of a real series  */
   for(int i=0; i<n; i++) {
-    in[i] = hrealval[i] - I*himagval[i] ; /* Restoring the standard sign convention for the FT, used by FFTW */
[…]

### 2016-11-10 tools/fft.h

@@ -55,8 +55,20 @@ int FFTTimeSeries(
 /* IFFT of frequency series */
 /* Note: assumes frequency series is FT of real data */
 /* Note: FFT uses flipped convention (i.e. h(f) = int e^(+2ipift)h(t)) */
-int IFFTFrequencySeries(
+int IFFTFrequencySeriesReal(
   RealTimeSeries** timeseries,       /* Output: real time series*/
+  ReImFrequencySeries* freqseries,   /* Input: complex frequency series, assumed to be the FT of a real time series */
+  double f1windowbeg,                /* Start of window at the beginning */
+  double f2windowbeg,                /* End of window at the beginning */
+  double f1windowend,                /* Start of window at the end */
+  double f2windowend,                /* End of window at the end */
+  int nzeropad);                     /* For 0-padding: length will be (upper power of 2)*2^nzeropad */
+
+/* IFFT of frequency series */
+/* Note: assumes frequency series is FT of complex data - produces complex output */
+/* Note: FFT uses flipped convention (i.e. h(f) = int e^(+2ipift)h(t)) */
+int IFFTFrequencySeries(
+  ReImTimeSeries** timeseries,       /* Output: complex time series */
   ReImFrequencySeries* freqseries,   /* Input: complex frequency series, assumed to be the FT of a real time series */
   double f1windowbeg,                /* Start of window at the beginning */
   double f2windowbeg,                /* End of window at the beginning */

### 2016-11-10 tools/likelihood.c

@@ -269,23 +269,24 @@ double FDListmodesLogLinearOverlap(
   double minf1, maxf1, minf2, maxf2, minf, maxf;
   ListmodesCAmpPhaseFrequencySeries* listelement1mode22 = ListmodesCAmpPhaseFrequencySeries_GetMode(list1, 2, 2);
   ListmodesCAmpPhaseFrequencySeries* listelement2mode22 = ListmodesCAmpPhaseFrequencySeries_GetMode(list2, 2, 2);
-  minf1 = fmax(gsl_vector_get(listelement1mode22->freqseries->freq, 0), fstartobs1);
-  minf2 = fmax(gsl_vector_get(listelement2mode22->freqseries->freq, 0), fstartobs2);
-  maxf1 = gsl_vector_get(list1->freqseries->freq, (int) list1->freqseries->freq->size - 1);
-  maxf2 = gsl_vector_get(list2->freqseries->freq, (int) list2->freqseries->freq->size - 1);
-  ListmodesCAmpPhaseFrequencySeries* listelement1 = list1;
-  ListmodesCAmpPhaseFrequencySeries* listelement2 = list2;
-  while(listelement1) {
-    maxf1 = fmax(maxf1, gsl_vector_get(listelement1->freqseries->freq, (int) listelement1->freqseries->freq->size - 1));
-    listelement1 = listelement1->next;
+  minf1 = gsl_vector_get(listelement1mode22->freqseries->freq, 0);
+  minf2 = gsl_vector_get(listelement2mode22->freqseries->freq, 0);
+  maxf1 = ListmodesCAmpPhaseFrequencySeries_maxf(list1);
+  maxf2 = ListmodesCAmpPhaseFrequencySeries_maxf(list2);
+  /* Taking into account fLow, fHigh -- fstartobs, determined from 22 mode, is also used as a fLow */
+  if(!(fHigh==0.)) {
+    maxf1 = fmin(maxf1, fHigh);
+    maxf2 = fmin(maxf2, fHigh);
   }
-  while(listelement2) {
-    maxf2 = fmax(maxf2, gsl_vector_get(listelement2->freqseries->freq, (int) listelement2->freqseries->freq->size - 2));
-    listelement2 = listelement2->next;
+  if(!(fLow==0.)) {
+    minf1 = fmax(minf1, fLow);
+    minf2 = fmax(minf2, fLow);
   }
[…]

### 2016-11-10 tools/struct.c

@@ -287,6 +287,47 @@ void ReImFrequencySeries_Cleanup(ReImFrequencySeries *freqseries) {
   free(freqseries);
 }
 
+/******** Functions to initialize and clean up ReImTimeSeries structure ********/
+void ReImTimeSeries_Init(ReImTimeSeries **timeseries, const int n) {
+  if(!timeseries) exit(1);
+  /* Create storage for structures */
+  if(!*timeseries) *timeseries=malloc(sizeof(ReImTimeSeries));
+  else
+  {
+    ReImTimeSeries_Cleanup(*timeseries);
+  }
+  gsl_set_error_handler(&Err_Handler);
+  (*timeseries)->times = gsl_vector_alloc(n);
+  (*timeseries)->h_real = gsl_vector_alloc(n);
+  (*timeseries)->h_imag = gsl_vector_alloc(n);
+}
+void ReImTimeSeries_Cleanup(ReImTimeSeries *timeseries) {
+  if(timeseries->times) gsl_vector_free(timeseries->times);
+  if(timeseries->h_real) gsl_vector_free(timeseries->h_real);
+  if(timeseries->h_imag) gsl_vector_free(timeseries->h_imag);
+  free(timeseries);
+}
+/******** Functions to initialize and clean up AmpPhaseTimeSeries structure ********/
+void AmpPhaseTimeSeries_Init(AmpPhaseTimeSeries **timeseries, const int n) {
+  if(!timeseries) exit(1);
+  /* Create storage for structures */
+  if(!*timeseries) *timeseries=malloc(sizeof(AmpPhaseTimeSeries));
+  else
+  {
+    AmpPhaseTimeSeries_Cleanup(*timeseries);
+  }
+  gsl_set_error_handler(&Err_Handler);
+  (*timeseries)->times = gsl_vector_alloc(n);
+  (*timeseries)->h_amp = gsl_vector_alloc(n);
+  (*timeseries)->h_phase = gsl_vector_alloc(n);
+}
+void AmpPhaseTimeSeries_Cleanup(AmpPhaseTimeSeries *timeseries) {
+  if(timeseries->times) gsl_vector_free(timeseries->times);
+  if(timeseries->h_amp) gsl_vector_free(timeseries->h_amp);
+  if(timeseries->h_phase) gsl_vector_free(timeseries->h_phase);
+  free(timeseries);
+}
+
 /******** Functions to initialize and clean up RealTimeSeries structure ********/
[…]

### 2016-11-10 tools/struct.h

@@ -83,7 +83,26 @@ typedef struct tagReImFrequencySeries
   gsl_vector* h_imag;
 } ReImFrequencySeries;
 
+/* Complex frequency series in real/imaginary part representation (for one mode, or their sum) */
+/* NOTE: for now, exact duplicata of ReImFrequencySeries - differentiated for readability of the code */
+typedef struct tagReImTimeSeries
+{
+  gsl_vector* times;
+  gsl_vector* h_real;
+  gsl_vector* h_imag;
+} ReImTimeSeries;
+
+/* Complex frequency series in amplitude/phase representation (representing one mode) */
+/* NOTE: for now, exact duplicata of ReImFrequencySeries - differentiated for readability of the code */
+typedef struct tagAmpPhaseTimeSeries
+{
+  gsl_vector* times;
+  gsl_vector* h_amp;
+  gsl_vector* h_phase;
+} AmpPhaseTimeSeries;
+
 /* Real time series */
+/* NOTE: could change the h to something more general, like values - also used for TD 22 amplitude, for instance  */
 typedef struct tagRealTimeSeries
 {
   gsl_vector* times;
@@ -172,13 +191,21 @@ void CAmpPhaseSpline_Init(
 	 const int n );                         /* length of the frequency series setting the splines */
 void CAmpPhaseSpline_Cleanup(CAmpPhaseSpline* splines);
 void CAmpPhaseGSLSpline_Init(
-	 CAmpPhaseGSLSpline** splines,             /* double pointer for initialization */
+	 CAmpPhaseGSLSpline** splines,          /* double pointer for initialization */
 	 const int n );                         /* length of the frequency series setting the splines */
 void CAmpPhaseGSLSpline_Cleanup(CAmpPhaseGSLSpline* splines);
 void ReImFrequencySeries_Init(
 	 ReImFrequencySeries** freqseries,      /* double pointer for initialization */
 	 const int n );                         /* length of the frequency series */
 void ReImFrequencySeries_Cleanup(ReImFrequencySeries* freqseries);
+void ReImTimeSeries_Init(
+	 ReImTimeSeries** timeseries,           /* double pointer for initialization */
+	 const int n );                         /* length of the time series */
+void ReImTimeSeries_Cleanup(ReImTimeSeries* timeseries);
+void AmpPhaseTimeSeries_Init(
+	 AmpPhaseTimeSeries** timeseries,       /* double pointer for initialization */
[…]

### 2016-11-10 tools/waveform.c

@@ -65,7 +65,7 @@ double ListmodesCAmpPhaseFrequencySeries_minf(ListmodesCAmpPhaseFrequencySeries*
   CAmpPhaseFrequencySeries* hlm;
   while(listelementhlm) {
     hlm = listelementhlm->freqseries;
-    minf = fmin(minf, gsl_vector_get(hlm->freq, hlm->freq->size - 1));
+    minf = fmin(minf, gsl_vector_get(hlm->freq, 0));
     listelementhlm = listelementhlm->next;
   }
   return minf;
@@ -88,6 +88,8 @@ double EstimateInitialTime(ListmodesCAmpPhaseFrequencySeries* listhlm, double fL
 void ReImFrequencySeries_AddCAmpPhaseFrequencySeries(
   struct tagReImFrequencySeries* freqseriesReIm,              /* Output Re/Im frequency series */
   struct tagCAmpPhaseFrequencySeries* freqseriesCAmpPhase,    /* Input CAmp/Phase frequency series, to be interpolated and added to the output */
+  double fLow,                                                /* Minimal frequency - set to 0 to ignore */
+  double fHigh,                                               /* Maximal frequency - set to 0 to ignore */
   double fstartobsmode)                                       /* Starting frequency in case of limited duration of observations- assumed to have been scaled with the proper factor m/2 for this mode - set to 0 to ignore */
 {
   /* Frequency vectors, and sizes */
@@ -95,6 +97,10 @@ void ReImFrequencySeries_AddCAmpPhaseFrequencySeries(
   gsl_vector* freqout = freqseriesReIm->freq;
   int sizein = (int) freqin->size;
   int sizeout = (int) freqout->size;
+
+  //
+  printf("sizeout %d\n", sizeout);
+
   /* Input, Output vectors */
   gsl_vector* vecampreal = freqseriesCAmpPhase->amp_real;
   gsl_vector* vecampimag = freqseriesCAmpPhase->amp_imag;
@@ -115,18 +121,21 @@ void ReImFrequencySeries_AddCAmpPhaseFrequencySeries(
   gsl_spline_init(phase, gsl_vector_const_ptr(freqin,0), gsl_vector_const_ptr(vecphase,0), sizein);
 
   /* First and last index of the output frequency vector that are covered by the CAmpPhase data */
+  /* Takes into account fLow, fHigh and fstartobs */
   int jStart = 0;
   int jStop = sizeout - 1;
[…]

### 2016-11-10 tools/waveform.h

@@ -69,13 +69,17 @@ double EstimateInitialTime(ListmodesCAmpPhaseFrequencySeries* listhlm, double fL
 void ReImFrequencySeries_AddCAmpPhaseFrequencySeries(
   struct tagReImFrequencySeries* freqseriesReIm,              /* Output Re/Im frequency series */
   struct tagCAmpPhaseFrequencySeries* freqseriesCAmpPhase,    /* Input CAmp/Phase frequency series, to be interpolated and added to the output */
+  double fLow,                                                /* Minimal frequency - set to 0 to ignore */
+  double fHigh,                                               /* Maximal frequency - set to 0 to ignore */
   double fstartobsmode);                                      /* Starting frequency in case of limited duration of observations- assumed to have been scaled with the proper factor m/2 for this mode - set to 0 to ignore */
 /* Function evaluating a ReImFrequencySeries by interpolating wach mode of a ListmodesCAmpPhaseFrequencySeries and summing them, given a set of frequencies */
 void ReImFrequencySeries_SumListmodesCAmpPhaseFrequencySeries(
   struct tagReImFrequencySeries* freqseriesReIm,                    /* Output Re/Im frequency series - already initialized */
   struct tagListmodesCAmpPhaseFrequencySeries* listmodesCAmpPhase,  /* Input CAmp/Phase frequency series, to be interpolated */
   gsl_vector* freq,                                                 /* Input set of frequencies on which evaluating */
-  double fstartobs);                                                /* Starting frequency in case of limited duration of observations - set to 0 to ignore */
+  double fLow,                                                      /* Minimal frequency - set to 0 to ignore */
+  double fHigh,                                                     /* Maximal frequency - set to 0 to ignore */
+  double fstartobs);                                                /* For limited duration of observation, starting frequency for the 22 mode - set to 0 to ignore */
 /* Helper function to add a mode to hplus, hcross in Fourier domain
  * - copies the function XLALSimAddMode, which was done only for TD structures */
 int FDAddMode(
@@ -92,27 +96,45 @@ int GeneratehphcFDReImFrequencySeries(
   ReImFrequencySeries** hptilde,                 /* Output: frequency series for hplus */
   ReImFrequencySeries** hctilde,                 /* Output: frequency series for hcross */
   ListmodesCAmpPhaseFrequencySeries* listhlm,    /* Input: frequency series for the mode hlm  */
-  double fLow,                                   /* Minimal frequency */
+  double fLow,                                   /* Minimal frequency - set to 0 to ignore */
[…]

### 2016-11-15 docs/paper_method.tex
Edits to the paper, added several plots.

@@
 \usepackage{ulem}
+
+\usepackage{graphicx}
+\usepackage[caption=false]{subfig}
+%\usepackage{caption}
+%\usepackage{subcaption}
+%\captionsetup{compatibility=false}
 
@@
 \ee
+Since we intend to work in the Fourier domain, we will use PhenomD waveforms~\cite{Khan+15}.
+These phenomenological waveforms provide analytic fits for both the Fourier-domain amplitude and phase of the 22 mode, with fit coefficients that were calibrated on EOB/NR hybrids [describe smoothing across intervals].
+Fig.~\ref{fig:ampphase} shows an example of amplitude and phase in geometric units for an equal-mass binary.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/Af.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/Psif.pdf}
+  \caption{Fourier-domain amplitude and phase in geometric units for an equal-mass, non-spinning system.
+The vertical line represents $\omega_{22}^{\rm peak}/2\pi$, the frequency corresponding to the time-domain frequency of the waveform at the peak amplitude.
+The starting frequency $Mf_{min} = 2.6\times 10^{-4}$ is appropriate for a 2-years observation of a $M=10^{6} M_{\odot}$ system by LISA, or for an observation by LIGO/Virgo of a $M=5.4 M_{\odot}$ system entering the instrument band at $10\Hz$.}
+  \label{fig:ampphase}
+\end{figure}
+
 We recall that, with our sign convention, the effect of a positive shift in time $h_{\Delta t} (t) = h(t- \Delta t)$ is
@@
 where the tilde here does not indicate a Fourier transform operation but simply that this quantity is an effective time defined from the Fourier-domain waveform.
-This definition is a generalization of the time-to-frequency correspondence at the heart of the SPA, as will be shown in Sec.~\ref{}, but it is worth noting the it is defined entirely from the Fourier-domain waveform.
-In particular, $\tf$ will in general not be monotonously increasing with frequency in the high-frequency part of the waveform, corresponding to the ringdown.
+This definition is a generalization of the time-to-frequency correspondence at the heart of the SPA, as will be discussed in Sec.~\ref{subsec:linkSPA}, but it is worth noting that here it is defined entirely from the Fourier-domain waveform.
+Fig.~\ref{fig:tf} shows for the example of an equal-mass waveform.
+In particular, one should note that $\tf$ is not monotonously increasing with frequency anymore when reaching in the high-frequency part of the waveform, corresponding to the ringdown.
+This would forbid an unambiguous definition of an inverse fequency-of-time correspondence $f(t)$.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/tf.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/tfzoom.pdf}
+  \caption{Time-to-frequency correspondence, as defined directly from the phase of the Fourier-domain signal as in~\eqref{eq:deftf}, in geometric units for an equal-mass, non-spinning system.
+The right panel shows the merger and ringdown part of the signal in more details.
+The time axis is such that the time-domain amplitude peaks at $t=0$.
+The vertical line shows the instantaneous frequency at the peak, and the fact that the curve $t_{f}$ does not pass exactly by the crossing of the two lines reflects the fact that the link between time-domain frequency and Fourier-domain f is not accurate anymore in the merger part of the signal.
+Note that $t_{f}$ does not go beyond $20M$ after the peak, and is not monotonous.}
+  \label{fig:tf}
+\end{figure}
 
@@
 
-Of particular interest for us will be the case where we keep only the first term in~\eqref{eq:expandPsi}, and expand the resulting exponential.
-Indeed, this term will be  the dominant correction beyond leading order in the applications that we will consider in Secs.~\ref{} and~\ref{}, and this particular approximation will turn out to be closely related to the approach of~\cite{Klein+14}.
-The above approach yields
+Of particular interest for us will be the case where we keep only the first term in~\eqref{eq:expandPsi}, and expand the resulting exponential, which yields
 \be\label{eq:resultdffPsi}
-	\tilde{s}(f) = \tilde{h}(f) \sum\limits_{p\geq 0} \frac{1}{p!} \left( \frac{i}{8\pi^{2}}\frac{\ud^{2} \Psi}{\ud f^{2}} \right)^{p} \left( \frac{\partial^{2p} }{\partial t^{2p}} g \right)(f, \tf) \,,
+	\tilde{s}(f) = \tilde{h}(f) \sum\limits_{p\geq 0} \frac{1}{p!} \left( \frac{i}{8\pi^{2}}\frac{\ud^{2} \Psi}{\ud f^{2}} \right)^{p} \left( \frac{\partial^{2p} }{\partial t^{2p}} g \right)(f, \tf) \,.
 \ee
-and it is natural to define a new timescale, as a function of frequency, as
-\be\label{eq:defTf}
-	\Tf^{2} = \frac{1}{4\pi^{2}}\left| \frac{\ud^{2}\Psi}{\ud f^{2}} \right| \,.
-\ee
-The interpretation of this timescale for the inspiralling part of the signal, when the SPA applies, will be discussed in the next section.
-Whe can use this notation to rewrite~\eqref{eq:resultdffPsi} as
-\begin{align}\label{eq:resultdffPsiTf}
-	 \tilde{s}(f) &= \tilde{h}(f) \sum\limits_{p\geq 0} \frac{(-i\epsilon)^{p}}{2^{p}p!} \Tf^{2p} \left( \frac{\partial^{2p} }{\partial t^{2p}} g \right)(f, \tf) \,,
+Indeed, this correction will be the most significant correction beyond leading order in the applications that we will consider in Secs.~\ref{} and~\ref{}, and this particular approximation will turn out to be closely related to the approach of~\cite{Klein+14}.
+
+In the following, it will be convenient to introduce figures of merit of these Taylor-like expansions. Thus we define
+\begin{subequations}\label{eq:deffom}
+\begin{align}
+	\epsilon_{\Psi 2} &\equiv \frac{1}{4\pi^{2}} \left| \frac{\ud^{2} \Psi}{\ud f^{2}} \frac{1}{g} \frac{\partial^{2} g}{\partial t^{2}}(f, \tf) \right| \,, \\
+	\epsilon_{A 1} &\equiv \frac{1}{2\pi} \left| \frac{1}{A}\frac{\ud A}{\ud f} \frac{1}{g} \frac{\partial g}{\partial t}(f, \tf) \right| \,, \\
+	\epsilon_{A 2} &\equiv \frac{1}{4\pi^{2}} \left| \frac{1}{A}\frac{\ud^{2} A}{\ud f^{2}} \frac{1}{g} \frac{\partial^{2} g}{\partial t^{2}}(f, \tf) \right| \,, \\
+	\epsilon_{d} &\equiv \frac{1}{2\pi} \left| \frac{1}{g} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{r} }{\partial t^{r}} g \right)(f, \tf) \right| \,.
 \end{align}
-with $\epsilon = -\mathrm{sgn}(\ud^{2}\Psi/\ud f^{2} )$.
+\end{subequations}
 
-The expansion~\eqref{eq:resultdfA} also leads to the natural introduction of a set of other timescales related to the successive derivatives of te amplitude.
-In an analogous manner to the definition~\eqref{eq:defTf} of the timescale $T_{f}$, we can define
-\be
-	\left( T_{Af}^{(n)} \right)^{n} \equiv \frac{1}{(2 \pi)^{n}} \frac{1}{A(f)} \left| \frac{\ud^{n} A}{\ud f^{n}} \right| \,,
-\ee
-where we included an absolute value to accomodate the possible sign changes in the right-hand side. In practice, only the first few of these timescales will be relevant.
+We found in both the LISA case and in the precession case that the third and higher derivatives of both the amplitude and phase are negligible.
+Note that the figure of merit $\epsilon_{d}$, introduced to represent the effect of the frequency dependence in the convolution kernel~\eqref{eq:FDkernel}, is signal-independent and depends only on $g$.
 
@@
 
-Of particular interest is the relation between these results and the SPA.
+\begin{figure}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/TfTA.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/TfTAzoom.pdf}
+  \caption{Fourier-domain amplitude and phase timescales, as defined in~\eqref{eq:defTf} and~\eqref{eq:defTA}, in geometric units for an equal-mass, non-spinning system.
+The right panel shows the merger and ringdown part of the signal in more details, the time-domain frequency at merger being represented by the vertical line.
+The derivatives from which these timescales are built encounter zero-crossings in the high-frequency range.}
+  \label{fig:TfTA}
+\end{figure}
+
+Considering~\eqref{eq:eq:resultdffPsi} it is natural to define a new timescale, as a function of frequency, as
+\be\label{eq:defTf}
+	\Tf^{2} = \frac{1}{4\pi^{2}}\left| \frac{\ud^{2}\Psi}{\ud f^{2}} \right| \,.
+\ee
+The interpretation of this timescale for the inspiralling part of the signal, when the SPA applies, will be discussed below. Whe can use this notation to rewrite~\eqref{eq:resultdffPsi} as
+\begin{align}\label{eq:resultdffPsiTf}
+	 \tilde{s}(f) &= \tilde{h}(f) \sum\limits_{p\geq 0} \frac{(-i\epsilon)^{p}}{2^{p}p!} \Tf^{2p} \left( \frac{\partial^{2p} }{\partial t^{2p}} g \right)(f, \tf) \,,
+\end{align}
+with $\epsilon = -\mathrm{sgn}(\ud^{2}\Psi/\ud f^{2} )$.
+
+The expansion in amplitude corrections~\eqref{eq:resultdfA} also leads to the natural introduction of a set of other timescales related to the successive derivatives of te amplitude.
+In an analogous manner to the definition~\eqref{eq:defTf} of the timescale $T_{f}$, we can define
+\be\label{eq:defTA}
+	\left( T_{An} \right)^{n} \equiv \frac{1}{(2 \pi)^{n}} \frac{1}{A(f)} \left| \frac{\ud^{n} A}{\ud f^{n}} \right| \,,
+\ee
+where we included an absolute value to accomodate the possible sign changes in the right-hand side. In practice, only the first few of these timescales will be relevant.
+
+Of particular interest is the relation between out formal results and the SPA.
 For non-precessing gravitational wave signals, the SPA is a very good approximation through the inspiralling phase~\cite{Droz+99}.
@@
 Comparing with~\eqref{eq:defTf} above, we see that our generalized time scale $\Tf$ also reduces to $T_{f}^{\rm SPA}$ when the SPA applies.
-However, its definition extends to the merger-ringdown part of the signal, and we have to include an absolute value to allow for a possible change of sign of $\ud^{2}\Psi/\ud f^{2}$ in that regime.
+However, its definition extends to the merger-ringdown part of the signal, and we have to include an absolute value to allow for a possible change of sign of $\ud^{2}\Psi/\ud f^{2}$ in that regime, as shown in Fig.~\ref{fig:TfTA}.
 
@@
 We will choose to write the response directly in terms of spin-weighted spherical modes $h_{\ell m}$, whose Fourier transforms are assumed to have a smooth amplitude and phase.
-Since the orbital delay due to the orbit around the Sun is common to all $y_{slr}$ observables, we will introduce the orbital-delayed waveform and rewrite the response as
+Since the orbital delay due to the orbit around the Sun is common to all $y_{slr}$ observables, we will decompose the response in the orbital delay and the constellation-centered response as
+\begin{subequations}
 \begin{align}
-	h^{\rm TT}_{O} (t) &\equiv h^{\rm TT}(t-\hat{k}\cdot p_{O}) \,, \nn\\
-	y_{slr} &= \frac{1}{2} \frac{1}{1 - \hat{k}\cdot n_{l}} n_{l}\cdot \left[ h_{O}^{\rm TT}(t - \hat{k}\cdot p^{L}_{s}) - h_{O}^{\rm TT}(t - \hat{k}\cdot p^{L}_{r}) \right] \cdot n_{l}\,,
+	h^{\rm TT}_{O} (t) &\equiv h^{\rm TT}(t-\hat{k}\cdot p_{O}) \,, \label{eq:dfresponseO} \\
+	y_{slr} &= \frac{1}{2} \frac{1}{1 - \hat{k}\cdot n_{l}} n_{l}\cdot \left[ h_{O}^{\rm TT}(t - \hat{k}\cdot p^{L}_{s}) - h_{O}^{\rm TT}(t - \hat{k}\cdot p^{L}_{r}) \right] \cdot n_{l}\,, \label{eq:dfresponsec}
 \end{align}
+\end{subequations}
 where we introduced the positions of the spacecrafts relative to the center of the constellation, $p^{L}_{A} \equiv p_{A} - p_{O}$.
@@
 
-If we put actual numbers, we have $\Omega/2\pi \simeq 3.1\times10^{-8}\mathrm{Hz} \ll f$ over all the frequency band we consider.
-Considering that the comb is of limited extend in $n$, we might want to Taylor expand the relevant quantities accross the comb.
-Since the phase is rapidly varying, and since we have to consider a constant and a linear term in the phase as arbitrary, it is natural to Taylor expand the phase at the first order (that is, approximate the phase linearly locally), neglect the $n\Omega/2\pi$ in $(f-n\Omega/2\pi)$ appearing in the argument of the amplitude and other terms, and assess the quality of this approximation.
-
-Preliminary tests show that this approximation should be valid, unsurprisingly, for all the mildly varying functions encountered: $\sinc$, $\exp$ and the linear term in the $c_{n}$'s, and the Bessel functions $J_{n}$.
-However, $\tilde{h}$ has a Fourier-domain phase which becomes very rapidly varying at low frequencies, in the deep inspiral, $\Psi\propto f^{-5/3}$; here it is not obvious that we can neglect the second derivative of the phase.
+Due to the smallness of $\Omega/2\pi \simeq 3.1\times10^{-8}\mathrm{Hz} \ll f$ over all the frequency band we consider, one can expect that . In this Section we try to quantify to which extent.
 
@@
 \begin{align}
-	|\delta A/A| &= \left| \frac{1}{A_{N}}\frac{\ud A_{N}}{\ud f} f_0 \right| \simeq 3\times10^{-3} \left(\frac{f}{10^{-5}\Hz}\right)^{-1} \,,\\
-	|\delta^{1}\Psi | &= \left| f_0\frac{\ud \Psi_{N}}{\ud f} \right| \simeq 10^{3}\left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-8/3} \,, \\
-	|\delta^{2}\Psi | &= \left| \frac{1}{2} f_0^{2}\frac{\ud^{2} \Psi_{N}}{\ud f^{2}} \right| \simeq 5 \left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-11/3} \,, \\
-	|\delta^{3}\Psi | &= \left| \frac{1}{6} f_0^{3}\frac{\ud^{3} \Psi_{N}}{\ud f^{3}} \right| \simeq 2\times 10^{-2} \left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-14/3} \,.
+	\delta_{A1} &\equiv \left| \frac{1}{A_{N}}\frac{\ud A_{N}}{\ud f} f_0 \right| \simeq 3\times10^{-3} \left(\frac{f}{10^{-5}\Hz}\right)^{-1} \,,\\
+	\delta_{\Psi 2} &\equiv \left| \frac{1}{2} f_0^{2}\frac{\ud^{2} \Psi_{N}}{\ud f^{2}} \right| \simeq 5 \left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-11/3} \,, \\
+	\delta_{\Psi 3} &\equiv \left| \frac{1}{6} f_0^{3}\frac{\ud^{3} \Psi_{N}}{\ud f^{3}} \right| \simeq 2\times 10^{-2} \left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-14/3} \,.
 \end{align}
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/LISAfomdO.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/LISAfomdc.pdf}
+  %
+  \includegraphics[width=.48\linewidth]{plots/LISAfomPsi2.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/LISAfomA1.pdf}
+  \caption{Figures of merit of the approximation as defined in~\eqref{eq:deffom}, for an equal-mass, non-spinning system and for total masses of $10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$.
+The upper left panel corresponds to the orbital delay part of the response~\eqref{eq:dfresponseO}, and the upper right panel corresponds to the LISA-centered response~\eqref{eq:dfresponsec}.
+The central line and interval are the mean and $1\sigma$ standard deviation of the logarithm of $\epsilon$ over 400 random values for the position in the sky, inclination and polarization.}
+  \label{fig:fomLISA}
+\end{figure}
+
+Fig.~\ref{fig:fomLISA} shows the magnitude of the figures of merit $\epsilon$ for different total masses, for the orbital delay, for the rest of the response, for the quadratic-in-phase term and for the first derivative of the amplitude.
+Since individual signals can show significant variation depending on the angular parameters, we show the average over the sky position, inclination and polarization, and $\pm \sigma$ standard deviation.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\subsection{Morphology of precessing waveforms}
+\subsection{Simplified model of precessing IMR waveform}
 \label{subsec:precmodel}
@@
 [explain Gatech formula for post-merger precession]
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Simplified model of precessing IMR waveform}
-\label{subsec:precmodel}
 
@@
 [qualify how bad the model is with respect to full NR]
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/LISAfomdO.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/LISAfomdc.pdf}
+  \caption{Figures of merit of the approximation as defined in~\eqref{eq:deffom}, for an equal-mass, non-spinning system and for total masses of $10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$.
+The upper left panel corresponds to the orbital delay part of the response~\eqref{eq:dfresponseO}, and the upper right panel corresponds to the LISA-centered response~\eqref{eq:dfresponsec}.
+The central line and interval are the mean and $1\sigma$ standard deviation of the logarithm of $\epsilon$ over 400 random values for the position in the sky, inclination and polarization.}
+  \label{fig:TfTA}
+\end{figure}
+
+In the absence of delays, we have a simple modulation of the signal by a function $F(t)$, as given in~\eqref{eq:defmodulationprec} above.
+The convolution with a frequency-dependent kernel~\eqref{eq:FDkernel} becomes a simple convolution in the Fourier-domain,
+\be\label{eq:convolution}
+	\tilde{s} (f) = \int \ud f' \, \tilde{h}(f-f') \tilde{F} (f') \,.
+\ee
 
@@
 The relevant scaling for the terms in the series~\eqref{eq:resultdffPsiTf}, for the inspiral phase where the SPA applies, is therefore
-%
 \be
@@
 \ee
-%
 meaning that the expansion becomes {\it less} well behaved for {\it lower} frequencies.
@@
 For instance, within the SPA we obtain
-%
 \begin{subequations}
@@
 \end{subequations}
-%
 
@@
 
-[describe treatment of MR by direct convolution]
+As will be shown by Fig.~\ref{} below, although including higher-order amplitude corrections does capture some effects near merger, it is not sufficiently accurate and seems to be too sensitive to the details of the evolution of the modulation $F(t)$ in the merger region.
+Motivated by this shortcoming of the Taylor-like expansion approach, here we investigate an alternative way of handling the merger-ringdown part of the signal.
+
+Thanks to the separation of timescales between the precessional timescale and the orbital timescale, for suffciently high frequencies the convolution~\eqref{eq:convolution} will have a support that will be entirely comprised in the high-frequency part of the signal $\tilde{h}(f)$, which is rather featureless and slowly varying as a function of $f$.
+Taking advantage of this smoothness, we will adopt a trigonometric polynomial representation for $\tilde{h}(f)$.
+For frequencies high enough that the support of the convolution~\eqref{eq:convolution} does not extend beyond the range covered by this trigonometric representation, we will be able to write the result directly as a Discrete Fourier Transform (DFT) with a limited number of samples.
+
+We consider the high-frequency part of the signal above some frequency $f_{0}$, up to some maximal frequency $f_{\rm max}$ where the Fourier-domain amplitude of the signal has decayed to a negligible level.
+In our example of the equal-mass, non-spinning waveform, we take $[Mf_{0}, Mf_{\rm max}] = [0.05, 0.16]$.
+First, we eliminate a constant and a linear term in the phase by choosing another frequency central to the high-frequency range we want to represent, which we take to be $Mf_{p} = 0.1$.
+We can think of $f_{p}$ as representing roughly the frequency at merger, and of the associated time $t_{p}\equiv \tf(f_{p})$ as an estimate of the time of merger.
+Note that the method should only be weakly sensitive to the precise choice of $f_{p}$ and $t_{p}$.
+
+Additionaly, we taper the amplitude to arrive at a flat amplitude at $f_{0}$, at the price of introducing a deviation from the amplitude of the original signal.
+We localize this deviation to $f\leq f_{1}$, with $Mf_{1} = 0.06$ in our example, and in general the trigonometric representation will be accurate to the original signal only for $f\geq f_{1}$.
+We denote by $\mathcal{W}(f)$ this tapering, built in practice as the discrete integral of a cosine window function.
+To ensure good continuity properties, we perform an artificial symmetrization to a ficitious range $f\in [f_{0} - (f_{\rm max} - f_{0}), f_{0}]$, by imposing symmetric amplitudes and phases about $f_{0}$.
+Defining $\Delta f \equiv 2 (f_{\rm max} - f_{0})$, we write
+\be\label{eq:defhsym}
+	\tilde{h}_{\rm sym}(f) = 
+	\begin{cases} 
+	\mathcal{W}(f) \exp\left[- i \left( \Psi(f_{0}) - 2i\pi (f-f_{0}) t_{p} \right) \right] \tilde{h}(f) \,,  &\text{ for } f \in [f_{0}, f_{0} + \Delta f /2] \\
+	\tilde{h}_{\rm sym}(2 f_{0} - f) \,,  &\text{ for } f \in [f_{0} - \Delta f/2, f_{0}]
+	\end{cases}
+\ee
+Fig~\ref{fig:convolsymmetrized} shows both the original high-frequency part of the signal and its symmetrized version with a flattened amplitude.
+
+Next, we build a a trigonometric polynomial representation of $\tilde{h}_{\rm sym }$.
+This construction is intimately related to the DFT, but we recall it for completeness.
+For a periodic function $f(x)$ defined on $x\in [x_{0}, x_{0} + \Delta x]$, and represented by $N$ samples $x_{j} = \ov{x} + j \Delta x/N$ (with $N$ chosen to satisfy the Nyquist criterion), we can build a trigonometric interpolant $P(x)$ as
+\be
+	P(x) = \sum\limits_{k=-M}^{+M} c_{k} e^{2i\pi k \frac{x-x_{0}}{\Delta x}}
+\ee
+that will satisfy the system $P(x_{j}) = f(x_{j})$ for $j=0,\dots, N-1$.
+Here we set $M=N/2$ (assuming N is even).
+The coefficients $c_{k}$ are related to the coefficients of the inverse DFT of the series for $f$.
+If we set $\omega \equiv e^{2i\pi/N}$ and define
+\be
+	y_{k} = \frac{1}{N} \sum\limits_{j=0}^{N-1} f(x_{j}) \omega^{jk} \,,
+\ee
+which is the expression of the inverse DFT in our sign convention~\eqref{}, the coefficients $c_{k}$ are given by
+\begin{align}\label{eq:ckyk}
+	c_{k} &= y_{k} \text{ for } k=0,\dots, M-1 \,, \nn\\
+	c_{k} &= y_{k+N} \text{ for } k=-M+1,\dots, -1 \,, \nn\\
+	c_{M} &= c_{-M} = \frac{y_{M}}{2} \,,
+\end{align}
+where the condition $c_{M} = c_{-M}$ is an arbitrary condition enforced to match the number of degrees of freedom.
+
+Now, assuming that we wish to compute~\eqref{eq:convolution} for a frequency $f$ such that the extent of the convolution does not extend beyond $[f_{1}, f_{\rm max}]$, we can approximate
+\be\label{eq:hsymtrigo}
+	\tilde{h}_{\rm sym } (f) \simeq \sum\limits_{k=-M}^{+M} (-1)^{k} c_{k} e^{2i\pi k \frac{f-f_{0}}{\Delta f}} \,,
+\ee
+where the factor $(-1)^{k}$ comes from the fact that $f_{0}$ is here at the center of the interval.
+The coefficients $c_{k}$ are built following the rules~\eqref{eq:ckyk} from the inverse DFT coefficients
+\be
+	y_{k} = \frac{1}{N} \sum\limits_{j=0}^{N-1} \tilde{h}_{\rm sym}\left( f_{0} + \frac{2j - N}{N} \Delta f \right) \omega^{jk} \,.
+\ee
+When inserting this representation~\eqref{eq:defhsym} and~\eqref{eq:hsymtrigo} into~\eqref{eq:convolution}, we obtain
+\be
+	\tilde{s}(f) = e^{-i \Psi(f_{0})} e^{2i\pi (f-f_{0}) t_{p}} \sum\limits_{k=-M}^{+M} (-1)^{k} c_{k} e^{2i\pi k \frac{f-f_{0}}{\Delta f}} F(t_{p} + k\delta t) \,, 
+\ee
+where we defined the time sampling $\delta t \equiv 1/\Delta f$.
+We see that we are left with a DFT-like expression to compute from $N+1$ time samples of the modulation function $F$, centered around $t_{p}$.
+
+In terms of performance, the implementation of~\eqref{} and~\eqref{} above is expected to have a very reasonable cost.
+We obtained results with a good accuracy, as shown in Fig.~\ref{}, by including 32 samples (64 samples in the artificially symmetrized signal).
+Since we only need to evaluate , these operations effectively amount to two DFT operations on $\sim$128 samples, to cover .
+This is of the same order of magnitude as the number of (logarithmically-spaced) frequency samples used to represent the Fourier-domain waveform with its amplitude and phase in the first place, thus the cost is comparable to the Taylor-like approach presented in Sec.~\ref{sec:quadphasedelay}.
 

### 2016-11-15 EOBNRv2HMROM/GenerateWaveform.c

@@ -210,14 +210,22 @@ static void Write_Wave_hphcTD(const char dir[], const char file[], RealTimeSerie
   else Write_Matrix(dir, file, outmatrix);
 }
 /* Output waveform in frequency series form,Re/Im for hplus and hcross */
+    // Output Re/Im because difference between numpy and C unwrapping
+//static void Write_Wave_h22TD(const char dir[], const char file[], AmpPhaseTimeSeries* h22, int binary)
 static void Write_Wave_h22TD(const char dir[], const char file[], ReImTimeSeries* h22, int binary)
 {
+
   /* Initialize output */
   int nbtimes = h22->times->size;
   gsl_matrix* outmatrix = gsl_matrix_alloc(nbtimes, 3);
 
+  //
+//printf("nbtimes, h_amp, h_phase %d %d %d\n", nbtimes, (int) h22->h_amp->size, (int) h22->h_phase->size);
+
   /* Set output matrix */
   gsl_matrix_set_col(outmatrix, 0, h22->times);
+  //gsl_matrix_set_col(outmatrix, 1, h22->h_amp);
+  //gsl_matrix_set_col(outmatrix, 2, h22->h_phase);
   gsl_matrix_set_col(outmatrix, 1, h22->h_real);
   gsl_matrix_set_col(outmatrix, 2, h22->h_imag);
 
@@ -345,7 +353,13 @@ int main(int argc, char *argv[])
     printf("length h22TD %d\n", h22TD->times->size);
     //exit(0);
 
+    /* Convert to Amp/Phase representation */
+    // Output Re/Im because difference between numpy and C unwrapping
+    //AmpPhaseTimeSeries* h22TDAmpPhase = NULL;
+    //ReImTimeSeries_ToAmpPhase(&h22TDAmpPhase, h22TD);
+
     /* Output */
+    //Write_Wave_h22TD(params->outdir, params->outfile, h22TDAmpPhase, params->binaryout);
     Write_Wave_h22TD(params->outdir, params->outfile, h22TD, params->binaryout);
     exit(0);
   }

### 2016-11-15 LISAinference/ComputeLISASNR.c

@@ -234,9 +234,9 @@ int main(int argc, char *argv[])
       gsl_vector* times = TDI1->times;
       double twindowbeg = 0.05 * (gsl_vector_get(times, times->size - 1) - gsl_vector_get(times, 0)); /* Here hardcoded relative window lengths */
       double twindowend = 0.01 * (gsl_vector_get(times, times->size - 1) - gsl_vector_get(times, 0)); /* Here hardcoded relative window lengths */
-      FFTTimeSeries(&TDI1FFT, TDI1, twindowbeg, twindowend, 2); /* Here hardcoded 0-padding */
-      FFTTimeSeries(&TDI2FFT, TDI2, twindowbeg, twindowend, 2); /* Here hardcoded 0-padding */
-      FFTTimeSeries(&TDI3FFT, TDI3, twindowbeg, twindowend, 2); /* Here hardcoded 0-padding */
+      FFTRealTimeSeries(&TDI1FFT, TDI1, twindowbeg, twindowend, 2); /* Here hardcoded 0-padding */
+      FFTRealTimeSeries(&TDI2FFT, TDI2, twindowbeg, twindowend, 2); /* Here hardcoded 0-padding */
+      FFTRealTimeSeries(&TDI3FFT, TDI3, twindowbeg, twindowend, 2); /* Here hardcoded 0-padding */
 
       /* Restrict FFT on frequency interval of interest - no limitation put on the upper bound */
       ReImFrequencySeries* TDI1FFTrestr = NULL;

### 2016-11-15 LISAsim/GenerateTDIFD.c

@@ -7,6 +7,8 @@ static GenTDIFDtag ParseGenTDIFDtag(char* string) {
   GenTDIFDtag tag;
   if(strcmp(string, "TDIhlm")==0) tag = TDIhlm;
   else if(strcmp(string, "TDIFD")==0) tag = TDIFD;
+  else if(strcmp(string, "h22FFT")==0) tag = h22FFT;
+  else if(strcmp(string, "yslrFFT")==0) tag = yslrFFT;
   else {
     printf("Error in ParseGenTDIFDtag: string not recognized.\n");
     exit(1);
@@ -51,9 +53,11 @@ Arguments are as follows:\n\
  --deltatobs           Observation duration (years, default=2)\n\
  --tagextpn            Tag to allow PN extension of the waveform at low frequencies (default=1)\n\
  --Mfmatch             When PN extension allowed, geometric matching frequency: will use ROM above this value. If <=0, use ROM down to the lowest covered frequency (default=0.)\n\
- --deltaf;             When generating frequency series from the mode contributions, deltaf for the output (0 to set automatically at 1/2*1/(2T))\n\
+ --deltaf              When generating frequency series from the mode contributions, deltaf for the output (0 to set automatically at 1/2*1/(2T))\n\
+ --twindowbeg          When generating frequency series from file by FFT, twindowbeg (0 to set automatically at 0.05*duration)\n\
+ --twindowend          When generating frequency series from file by FFT, twindowend (0 to set automatically at 0.01*duration)\n\
  --tagtdi              Tag choosing the set of TDI variables to use (default TDIAETXYZ)\n\
- --taggenwave          Tag choosing the wf format: TDIhlm (default: downsampled mode contributions to TDI in Amp/Phase form), TDIFD (hlm interpolated and summed accross modes)\n\
+ --taggenwave          Tag choosing the wf format: TDIhlm (default: downsampled mode contributions to TDI in Amp/Phase form), TDIFD (hlm interpolated and summed accross modes), h22FFT and yslrFFT (used with fomrtditdfile -- loads time series and FFT)\n\
[…]

### 2016-11-15 LISAsim/GenerateTDIFD.h

@@ -45,7 +45,9 @@
 
 typedef enum GenTDIFDtag {
   TDIFD,
-  TDIhlm
+  TDIhlm,
+  h22FFT,
+  yslrFFT
 } GenTDIFDtag;
 
 /* Parameters for the generation of a LISA waveform (in the form of a list of modes) */
@@ -68,6 +70,8 @@ typedef struct tagGenTDIFDparams {
   int tagextpn;              /* Tag to allow PN extension of the waveform at low frequencies (default=1) */
   double Mfmatch;            /* When PN extension allowed, geometric matching frequency: will use ROM above this value. If <=0, use ROM down to the lowest covered frequency (default=0.) */
   double deltaf;             /* When generating frequency series from the mode contributions, deltaf for the output (0 to set automatically at 1/2*1/(2T)) */
+  double twindowbeg;         /* When generating frequency series from file by FFT, twindowbeg (0 to set automatically at 0.05*duration) */
+  double twindowend;         /* When generating frequency series from file by FFT, twindowend (0 to set automatically at 0.01*duration) */
   int tagtdi;                /* Tag selecting the desired output format */
   int taggenwave;            /* Tag selecting the desired output format */
   int restorescaledfactor;   /* If 1, restore the factors that were scaled out of TDI observables */

### 2016-11-15 LISAsim/GenerateTDITD.c

@@ -16,6 +16,8 @@ Arguments are as follows:\n\
 --------------------------------------------------\n\
 ----- Physical Parameters ------------------------\n\
 --------------------------------------------------\n\
+ --phiRef              Phase (radians, default=0) -- used when input is h22\n\
+ --inclination         Inclination (radians, default=PI/3) -- used when input is h22\n\
  --lambda              First angle for the position in the sky (radians, default=0)\n\
  --beta                Second angle for the position in the sky (radians, default=0)\n\
  --polarization        Polarization of source (radians, default=0)\n\
@@ -36,6 +38,8 @@ Arguments are as follows:\n\
     ssize_t i;
 
     /* Set default values for the physical params */
+    params->phiRef = 0;
+    params->inclination = PI/3;
     params->lambda = 0;
     params->beta = 0;
     params->polarization = 0;
@@ -53,6 +57,10 @@ Arguments are as follows:\n\
         if (strcmp(argv[i], "--help") == 0) {
             fprintf(stdout,"%s", help);
             exit(0);
+        } else if (strcmp(argv[i], "--phiRef") == 0) {
+            params->phiRef = atof(argv[++i]);
+        } else if (strcmp(argv[i], "--inclination") == 0) {
+            params->inclination = atof(argv[++i]);
         } else if (strcmp(argv[i], "--lambda") == 0) {
             params->lambda = atof(argv[++i]);
         } else if (strcmp(argv[i], "--beta") == 0) {
@@ -278,8 +286,7 @@ int main(int argc, char *argv[])
     else if(params->tagtdi==y12L) {
       /* Evaluate y12 signal - constellation response only, input assumed to be h22tdO already */
       RealTimeSeries* y12td = NULL;
-      //
-      //Generatey12LTD(&y12td, spline_amp, spline_phase, accel_amp, accel_phase, times, nptmargin);
+      Generatey12LTD(&y12td, spline_amp, spline_phase, accel_amp, accel_phase, times, params->inclination, params->phiRef, nptmargin);
 
       /* Output */
       Write_yslrTD(params->outdir, params->outfile, y12td, params->binaryout);

### 2016-11-15 LISAsim/LISAgeometry.c

@@ -32,6 +32,7 @@
 #include <gsl/gsl_complex.h>
 
 #include "constants.h"
+#include "waveform.h"
 #include "LISAgeometry.h"
 
 #include <time.h> /* for testing */
@@ -908,6 +909,55 @@ double hOTDAmpPhase(
 }
 
 /* Functions evaluating yAB observables in time domain */
+double y12LTDfromh22AmpPhase(
+  gsl_spline* splineamp,                   /* Input spline for h22 TD amp */
+  gsl_spline* splinephase,                 /* Input spline for h22 TD phase */
+  gsl_interp_accel* accelamp,              /* Accelerator for amp spline */
+  gsl_interp_accel* accelphase,            /* Accelerator for phase spline */
+  double complex Y22,                      /* Y22 factor needed to convert h22 to hplus, hcross */
+  const double t)                          /* Time */
+{
+  /* Precompute array of sine/cosine */
+  for(int j=0; j<4; j++) {
+    cosarray[j] = cos((j+1) * Omega_SI * t);
+    sinarray[j] = sin((j+1) * Omega_SI * t);
+  }
+  /* Scalar products with k */
+  double n3Pn3plus = coeffn3Hn3plusconst;
+  double n3Pn3cross = coeffn3Hn3crossconst;
+  for(int j=0; j<4; j++) {
+    n3Pn3plus += cosarray[j] * coeffn3Hn3pluscos[j] + sinarray[j] * coeffn3Hn3plussin[j];
+    n3Pn3cross += cosarray[j] * coeffn3Hn3crosscos[j] + sinarray[j] * coeffn3Hn3crosssin[j];
+  }
+  /* Scalar products with k */
+  double kn3 = coeffkn3const;
+  double kp1 = coeffkp1const;
+  double kp2 = coeffkp2const;
+  for(int j=0; j<2; j++) {
+    kn3 += cosarray[j] * coeffkn3cos[j] + sinarray[j] * coeffkn3sin[j];
+    kp1 += cosarray[j] * coeffkp1cos[j] + sinarray[j] * coeffkp1sin[j];
+    kp2 += cosarray[j] * coeffkp2cos[j] + sinarray[j] * coeffkp2sin[j];
+  }
+  /* Common factor and delay */
+  double factorp = (1./(1.-kn3)) * 0.5*n3Pn3plus;
[…]

### 2016-11-15 LISAsim/LISAgeometry.h

@@ -216,6 +216,17 @@ int Generateh22TDO(
   gsl_interp_accel* accelphase,            /* Accelerator for phase spline */
   gsl_vector* times,                       /* Vector of times to evaluate */
   int nbptmargin);                         /* Margin set to 0 on both side to avoid problems with delays out of the domain */
+/* Generate y12L from orbital-delayed h22 in amp/phase form */
+int Generatey12LTD(
+  RealTimeSeries** y12Ltd,                 /* Output: real time series for y12L */
+  gsl_spline* splineamp,                   /* Input spline for TD mode amplitude */
+  gsl_spline* splinephase,                 /* Input spline for TD mode phase */
+  gsl_interp_accel* accelamp,              /* Accelerator for amp spline */
+  gsl_interp_accel* accelphase,            /* Accelerator for phase spline */
+  gsl_vector* times,                       /* Vector of times to evaluate */
+  double Theta,                            /* Inclination */
+  double Phi,                              /* Phase */
+  int nbptmargin);                         /* Margin set to 0 on both side to avoid problems with delays out of the domain */
 
 
 /* Generate TDI observables (including orbital delay) for one mode contritbution from amp, phase */

### 2016-11-15 LLVsim/GenerateLLVFD.c

@@ -264,9 +264,9 @@ int main(int argc, char *argv[])
     gsl_vector* times = LLV1->times;
     double twindowbeg = 0.05 * (gsl_vector_get(times, times->size - 1) - gsl_vector_get(times, 0)); /* Here hardcoded relative window lengths */
     double twindowend = 0.01 * (gsl_vector_get(times, times->size - 1) - gsl_vector_get(times, 0)); /* Here hardcoded relative window lengths */
-    FFTTimeSeries(&LLV1FFT, LLV1, twindowbeg, twindowend, 2); /* Here hardcoded 0-padding */
-    FFTTimeSeries(&LLV2FFT, LLV2, twindowbeg, twindowend, 2); /* Here hardcoded 0-padding */
-    FFTTimeSeries(&LLV3FFT, LLV3, twindowbeg, twindowend, 2); /* Here hardcoded 0-padding */
+    FFTRealTimeSeries(&LLV1FFT, LLV1, twindowbeg, twindowend, 2); /* Here hardcoded 0-padding */
+    FFTRealTimeSeries(&LLV2FFT, LLV2, twindowbeg, twindowend, 2); /* Here hardcoded 0-padding */
+    FFTRealTimeSeries(&LLV3FFT, LLV3, twindowbeg, twindowend, 2); /* Here hardcoded 0-padding */
 
     /* Output */
     char *outfileLLV1 = malloc(256);

### 2016-11-15 tools/fft.c

@@ -54,10 +54,14 @@ double WindowFunctionRight(double x, double xi, double deltaxi)
   else return 1;
 }
 
-/* FFT of time series */
+/* FFT of real time series */
 /* Note: FFT uses flipped convention (i.e. h(f) = int e^(+2ipift)h(t)) */
-int FFTTimeSeries(ReImFrequencySeries** freqseries, RealTimeSeries* timeseries, double twindowbeg, double twindowend, int nzeropad)
+int FFTRealTimeSeries(ReImFrequencySeries** freqseries, RealTimeSeries* timeseries, double twindowbeg, double twindowend, int nzeropad)
 {
+  //
+  printf("aaa\n");
+  printf("%g, %g\n", twindowbeg, twindowend);
+
   /* deltat of time series */
   /* Warning: assumes linear sampling in time */
   double* times = timeseries->times->data;
@@ -70,9 +74,18 @@ int FFTTimeSeries(ReImFrequencySeries** freqseries, RealTimeSeries* timeseries, 
   gsl_vector* hvalues = gsl_vector_alloc(n + nzeros);
   gsl_vector_set_zero(hvalues);
 
+  //
+  printf("aaa\n");
+  printf("deltat %g\n", deltat);
+
   /* Compute input TD values, with windowing */
   int nbptswindowbeg = (int) ceil(twindowbeg/deltat) + 1;
   int nbptswindowend = (int) ceil(twindowend/deltat) + 1;
+
+
+    //
+    printf("aaa\n");
+    printf("%d, %d\n", nbptswindowbeg, nbptswindowend);
   double t1windowbeg = times[0];
   double t2windowbeg = times[nbptswindowbeg-1];
   double t1windowend = times[n-nbptswindowend];
@@ -81,6 +94,10 @@ int FFTTimeSeries(ReImFrequencySeries** freqseries, RealTimeSeries* timeseries, 
   double deltatwindowend = t2windowend - t1windowend;
   double* h = timeseries->h->data;
   double* hval = hvalues->data;
+
+  //
+  printf("aaa\n");
+
   for (int i=0; i<nbptswindowbeg; i++) {
     hval[i] = WindowFunctionRight(times[i], t1windowbeg, deltatwindowbeg) * h[i];
   }
@@ -91,6 +108,9 @@ int FFTTimeSeries(ReImFrequencySeries** freqseries, RealTimeSeries* timeseries, 
[…]

### 2016-11-15 tools/fft.h

@@ -43,11 +43,20 @@ double WindowFunction(double x, double xi, double xf, double deltaxi, double del
 double WindowFunctionLeft(double x, double xf, double deltaxf);
 double WindowFunctionRight(double x, double xi, double deltaxi);
 
-/* FFT of time series */
+/* FFT of real time series */
+/* Note: FFT uses flipped convention (i.e. h(f) = int e^(+2ipift)h(t)) */
+int FFTRealTimeSeries(
+  ReImFrequencySeries** freqseries,   /* Output: frequency series */
+  RealTimeSeries* timeseries,         /* Input: real time series */
+  double twindowbeg,                  /* Extent of the window at beginning (starts at the first point) */
+  double twindowend,                  /* Extent of the window at the end (end at the last point) */
+  int nzeropad);                      /* For 0-padding: length will be (upper power of 2)*2^nzeropad */
+
+/* FFT of Re/Im time series */
 /* Note: FFT uses flipped convention (i.e. h(f) = int e^(+2ipift)h(t)) */
 int FFTTimeSeries(
   ReImFrequencySeries** freqseries,   /* Output: frequency series */
-  RealTimeSeries* timeseries,         /* Input: real time series */
+  ReImTimeSeries* timeseries,         /* Input: Re/Im time series */
   double twindowbeg,                  /* Extent of the window at beginning (starts at the first point) */
   double twindowend,                  /* Extent of the window at the end (end at the last point) */
   int nzeropad);                      /* For 0-padding: length will be (upper power of 2)*2^nzeropad */

### 2016-11-15 tools/waveform.c

@@ -521,8 +521,9 @@ int RestrictFDReImFrequencySeries(
 }
 
 /* Function to unwrap the phase mod 2pi  - acts directly on a gsl_vector representing the phase */
+/* WARNING : found significant accumulating difference with the numpy unwrap function - to be understood */
 int UnwrapPhase(
-  gsl_vector** phaseout,   /* Output: unwrapped phase vector */
+  gsl_vector* phaseout,    /* Output: unwrapped phase vector, already allocated */
   gsl_vector*  phasein)    /* Input: phase vector */
 {
   size_t N = phasein->size;
@@ -530,10 +531,6 @@ int UnwrapPhase(
   double* diffcorr = (double*) malloc(sizeof(double) * N);
   double* cumsum = (double*) malloc(sizeof(double) * N);
   double* p = phasein->data;
-
-  /* Initialize output */
-  *phaseout = gsl_vector_alloc(N);
-  double* pout = (*phaseout)->data;
 
   /* Incremental phase variation - indexed from 1 to N-1 */
   diff[0] = 0;
@@ -549,6 +546,7 @@ int UnwrapPhase(
   /* One could preserve sign of variation if diff[i] = +pi (maps to -pi here) - here we simply assume that the phase jumps are strictly smaller than +-pi, which is the limiting case for the reconstruction to work  */
 
   /* Cumulative sum of diff */
+  double* pout = phaseout->data;
   pout[0] = p[0];
   for(int i=1; i<N; i++) {
     pout[i] = pout[i-1] + diff[i];
@@ -556,6 +554,39 @@ int UnwrapPhase(
 
   /* Cleanup */
   free(diff);
+}
+
+/* Function to convert a time series from Re/Im form to Amp/Phase form - unwrapping the phase */
+int ReImTimeSeries_ToAmpPhase(
+  AmpPhaseTimeSeries** timeseriesout,             /* Output: Amp/Phase time series */
+  ReImTimeSeries* timeseriesin)                   /* Input: Re/Im time series */
+{
+  /* Initialize output */
+  int npt = (int) timeseriesin->times->size;
+  AmpPhaseTimeSeries_Init(timeseriesout, npt);
+  gsl_vector_memcpy((*timeseriesout)->times, timeseriesin->times);
+
+  /* Compute amplitude and phase */
+  gsl_vector* amp = gsl_vector_alloc(npt);
[…]

### 2016-11-15 tools/waveform.h

@@ -133,8 +133,12 @@ int RestrictFDReImFrequencySeries(
   double fHigh);                                 /* Maximal frequency */
 /* Function to unwrap the phase mod 2pi  - acts directly on the gsl_vector representing the phase */
 int UnwrapPhase(
-  gsl_vector** phaseout,   /* Output: unwrapped phase vector */
+  gsl_vector*  phaseout,   /* Output: unwrapped phase vector - already allocated */
   gsl_vector*  phasein);   /* Input: phase vector */
+/* Function to convert a time series from Re/Im form to Amp/Phase form - unwrapping the phase */
+int ReImTimeSeries_ToAmpPhase(
+  AmpPhaseTimeSeries** timeseriesout,             /* Output: Amp/Phase time series */
+  ReImTimeSeries* timeseriesin);                  /* Input: Re/Im time series */
 
 /***************** Spin weighted spherical harmonics ****************/
 

### 2016-11-22 docs/paper_method.tex

@@
 \newcommand\calF{{\mathcal{F}}}
+\newcommand\calT{{\mathcal{T}}}
 \newcommand\bfx{\mathbf{x}}
@@
 
-\author{John G. Baker}
-\affiliation{Gravitational Astrophysics Laboratory, NASA Goddard Space Flight Center, 8800 Greenbelt Rd., Greenbelt, MD 20771, USA}
-\author{Alessandra Buonanno}
-\affiliation{Department of Physics, University of Maryland, College Park, MD 20742, USA}
-\affiliation{Max Planck Institute for Gravitational Physics (Albert Einstein Institute), Am M\"uhlenberg 1, Potsdam-Golm, 14476, Germany}
-\author{Philip. B. Graff}
-\affiliation{Department of Physics, University of Maryland, College Park, MD 20742, USA}
-\affiliation{Gravitational Astrophysics Laboratory, NASA Goddard Space Flight Center, 8800 Greenbelt Rd., Greenbelt, MD 20771, USA}
-\author{Sylvain Marsat}
-\affiliation{Max Planck Institute for Gravitational Physics (Albert Einstein Institute), Am M\"uhlenberg 1, Potsdam-Golm, 14476, Germany}
+\author{A. A. Author}
+\affiliation{U. U. University}
+
+%\author{John G. Baker}
+%\affiliation{Gravitational Astrophysics Laboratory, NASA Goddard Space Flight Center, 8800 Greenbelt Rd., Greenbelt, MD 20771, USA}
+%\author{Alessandra Buonanno}
+%\affiliation{Department of Physics, University of Maryland, College Park, MD 20742, USA}
+%\affiliation{Max Planck Institute for Gravitational Physics (Albert Einstein Institute), Am M\"uhlenberg 1, Potsdam-Golm, 14476, Germany}
+%\author{Philip. B. Graff}
+%\affiliation{Department of Physics, University of Maryland, College Park, MD 20742, USA}
+%\affiliation{Gravitational Astrophysics Laboratory, NASA Goddard Space Flight Center, 8800 Greenbelt Rd., Greenbelt, MD 20771, USA}
+%\author{Sylvain Marsat}
+%\affiliation{Max Planck Institute for Gravitational Physics (Albert Einstein Institute), Am M\"uhlenberg 1, Potsdam-Golm, 14476, Germany}
 
@@
 If $(\alpha, \beta, \gamma)$ are the Euler angles relating the precessing frame to the inertial frame in the $(z,y,z)$ convention, the modes in the inertial frame $h_{\ell m}^{\rm I}$ are then related to the modes in the precessing frame $h_{\ell m}^{\rm P}$ by~\cite{Goldberg+67}
-\be
-	h_{\ell m}^{\rm I} = \sum\limits_{m=-\ell}^{\ell} D^{\ell}_{mm'} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm P} \,,
+\be\label{eq:wignerrot}
+	h_{\ell m}^{\rm I} = \sum\limits_{m=-\ell}^{\ell} D^{\ell *}_{m'm} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm P} \,,
 \ee
@@
 
-\subsection{Leading order: the locally linear phase approximation}
+\subsection{Leading order: the local approximation}
 \label{subsec:LLP}
@@
 Under the assumptions above, it is possible to see the convolution in the $f'$ variable in~\eqref{eq:FDkernel} as being a local operation, allowing us to expand the signal $\tilde{h}(f-f')$ around $f$.
-
-To this end, we will separate the signal in a Fourier-domain amplitude and phase, according to
-%
-\be
-	\tilde{h}(f) = A(f) e^{-i\Psi(f)} \,,
-\ee
-%
-
-A first leading-order approximation, which could be called the locally linear phase (LLP) approximation, is to discard the $f'$ dependence in the first argument of $\tilde{g}(f-f', f')$, to keep only the leading order in the Fourier-domain amplitude, and to expand the Fourier-domain phase to the first order, according to
-%
-\be
-	\tilde{h}_{\rm LLP}(f-f') \simeq A(f) \exp\left[ -i\left( \Psi(f) - f' \frac{\ud \Psi}{\ud f} \right) \right] \,,
-\ee
-%
-Here and in the rest of the paper, derivatives with respect to $f$ will always be evaluated at $f$.
+A first leading-order approximation, which could be called the local approximation for reasons that will appear clear below, is to discard the $f'$ dependence in the first argument of $\tilde{g}(f-f', f')$, to keep only the leading order in the Fourier-domain amplitude, and to expand the Fourier-domain phase to the first order, according to
+\be
+	\tilde{h}(f-f') \simeq A(f) \exp\left[ -i\left( \Psi(f) - f' \frac{\ud \Psi}{\ud f} \right) \right] \,,
+\ee
+Here and in the rest of the paper, derivatives with respect to the frequency will always be evaluated at $f$.
 Notice that, because of the freedom of adding a linear term to $\Psi(f)$ by simply shifting the signal in time, no assumption can be made on the smallness of the first derivative of the phase, and this is really a leading order approximation.
@@
 Plugging this relation in~\eqref{eq:FDkernel}, we obtain
-%
-\begin{align}
-	\tilde{s}_{\rm LLP}(f) &= \tilde{h}(f) \int \ud f' \, \exp\left[ i f' \frac{\ud \Psi}{\ud f} \right] \tilde{g}(f,f') \nn\\
+\begin{align}
+	\tilde{s}(f) &= \tilde{h}(f) \int \ud f' \, \exp\left[ i f' \frac{\ud \Psi}{\ud f} \right] \tilde{g}(f,f') \nn\\
 	&= \tilde{h}(f) g\left( f, -\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) \,,
 \end{align}
-%
-which motivates the introduction of the effective time, seen as a function of frequency,
-%
+which motivates the introduction of an effective time, seen as a function of frequency,
 \be\label{eq:deftf}
@@
 \ee
-%
 where the tilde here does not indicate a Fourier transform operation but simply that this quantity is an effective time defined from the Fourier-domain waveform.
 This definition is a generalization of the time-to-frequency correspondence at the heart of the SPA, as will be discussed in Sec.~\ref{subsec:linkSPA}, but it is worth noting that here it is defined entirely from the Fourier-domain waveform.
-Fig.~\ref{fig:tf} shows for the example of an equal-mass waveform.
+Fig.~\ref{fig:tf} shows the time function $\tf$ for the example of an equal-mass waveform.
 In particular, one should note that $\tf$ is not monotonously increasing with frequency anymore when reaching in the high-frequency part of the waveform, corresponding to the ringdown.
@@
 %
-\be\label{eq:LLP}
-	\tilde{s}_{\rm LLP}(f) = \tilde{h}(f) g(f, \tf) \,.
+\be\label{eq:local}
+	\tilde{s}_{\rm local}(f) = \tilde{h}(f) g(f, \tf) \,.
 \ee
@@
 Additionally, when delays are present, the frequency-dependence of the response is evaluated directly at $f$.
-For precessing binary, this corresponds to the approximation used for the PhenomP approximant~\cite{}, as we will discuss in Sec.~\ref{}.
-
-It is worth noting that a shift in time of the time-domain signal will, by construction, be appropriately propagated to~\eqref{eq:LLP}.
-If $h'(t) = h(t -  \Delta t)$ but the modulation and delays are left unchanged (which is relevant for space-based detectors, not for precessing binaries), $\Psi'(f) = \Psi(f) - 2\pi f \Delta t$, so that
-%
-\be\label{eq:LLP}
-	\tilde{s}'_{\rm LLP}(f) = \tilde{h}'(f) g(f, \tf + \Delta t) \,. 
+For precessing binary, this corresponds to the approximation used for the PhenomP approximant~\cite{Hannam+13}, as will be discussed in Sec.~\ref{}.
+
+It is worth noting that a shift in time of the time-domain signal will, by construction, be appropriately propagated to~\eqref{eq:local}.
+If $h'(t) = h(t -  \Delta t)$ [but the modulation and delays are left unchanged (which is relevant for space-based detectors, not for precessing binaries)], $\Psi'(f) = \Psi(f) - 2\pi f \Delta t$, so that
+%
+\be\label{eq:localshifted}
+	\tilde{s}'_{\rm local}(f) = \tilde{h}'(f) g(f, \tf + \Delta t) \,. 
 \ee
@@
 This observation also justifies that the first derivative of the Fourier-domain phase cannot, in general, be considered a small quantity, as it can take arbitrary values for arbitrary shifts in time of the signal.
-The leading order approximation corresponds to neglecting the curvature in $\Psi(f)$, hence the name Locally Linear Phase approximation.
+The leading order approximation corresponds to neglecting the curvature in $\Psi(f)$, and yields a result~\eqref{eq:local} where the passage to the Fourier domain is done by simply using a time-to-frequency correspondence~\eqref{eq:deftf}, which motivates the name ``local approximation''.
 
@@
 %\end{subequations}
+the transfer function for~\eqref{eq:dfresponsec} reads in the local approximation
+\be
+	\calT_{slr}^{\ell m} = \frac{i \pi f L}{2} \sinc \left[ \pi f L\left(1-\hat{k}\cdot n_{l} \right) \right] \exp\left[ \pi f \left( L + \hat{k}\cdot \left( p_{1}^{L} + p_{2}^{L} \right) \right) \right] {}_{-2}Y_{\ell m} n_{l}\cdot \left( P_{+} - i P_{\times} \right) \cdot n_{l} \text{ for } m>0 \,,
+\ee
+where we used the fact that, in our rigid approximation for the trajectory of the constellation, $p^{L}_{r} - p^{L}_{s} =  L n_{l}$.
 
@@
 
-[implementation]
-[examples of errors]
+\begin{figure}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/tdifdAamp.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/tdifdAphase.pdf}
+  \caption{Fourier-domain amplitude and phase timescales, as defined in~\eqref{eq:defTf} and~\eqref{eq:defTA}, in geometric units for an equal-mass, non-spinning system.
+The right panel shows the merger and ringdown part of the signal in more details, the time-domain frequency at merger being represented by the vertical line.
+The derivatives from which these timescales are built encounter zero-crossings in the high-frequency range.}
+  \label{fig:tdifdA}
+\end{figure}
+
+Fig.~\ref{fig:tdifdA} shows the amplitude and phase errors in the Fourier-domain local response when compared to the FFT of the time-domain response~\eqref{eq:yslr}.
 
@@
 
+In the following, since our purpose here is not to provide a full-fledged model for precessing signals in the Fourier domain, we will not explore the full parameter space and we will limit our analysis to a simplified toy model for precession.
+Namely, we will approximate the Euler angles during the inspiral phase by using the PN expressions used in PhenomP~\cite{Hannam+13}, and we will adopt two alternatives asymptotic behaviours post-merger.
+In the first one (thereafter Case I), we simply send smoothly all three Euler angles to constants.
+In the second one (Case II), we send the opening angle of the precession cone $\beta$ to a constant, but we reproduce the behaviour describe letting the frame rotate around the direction of the final total angular momentum $\hat{\bm{J}}_{\rm final}$ with an angular velocity of $\Omega_{\rm frame} = \omega_{220} - \omega_{210}$.
+Fig.~\ref{fig:prectoymodel} shows the Euler angles evolution for Cases I and II.
+
 \begin{figure}
   \centering
-  \includegraphics[width=.48\linewidth]{plots/LISAfomdO.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/LISAfomdc.pdf}
-  \caption{Figures of merit of the approximation as defined in~\eqref{eq:deffom}, for an equal-mass, non-spinning system and for total masses of $10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$.
-The upper left panel corresponds to the orbital delay part of the response~\eqref{eq:dfresponseO}, and the upper right panel corresponds to the LISA-centered response~\eqref{eq:dfresponsec}.
-The central line and interval are the mean and $1\sigma$ standard deviation of the logarithm of $\epsilon$ over 400 random values for the position in the sky, inclination and polarization.}
-  \label{fig:TfTA}
+  \includegraphics[width=.48\linewidth]{plots/prectoymodelalphagamma.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/prectoymodelbeta.pdf}
+  \caption{Evolution of Euler angles $(\alpha, \beta, \gamma)$ for our two toy models.
+Dashed lines represent Case I and solid lines Case II.
+The vertical line indicates the time of merger.
+Dashed black lines indicate the enforced asymptotic behaviour: frame rotation with angular velocity $\omega_{220} -\omega_{210}$ in Case II, and asymptotically constant $\beta$.}
+  \label{fig:prectoymodel}
 \end{figure}
+
+These simplifications comes with obvious limitations.
+Notably, they produce modulation functions $F(t)$ that are very smooth by construction.
+However, these two toy models will allow us to explore the separation of timescales in the merger-ringdown part of the signal, and to highlight how the perturbative analysis can be challenged by the rapid rotation of the radiation frame after merger, even in this idealized model.
 
@@
 \ee
+
+It will be convenient to use the notation
+\be\label{eq:defFlmlmp}
+	F_{\ell m, \ell m'} \equiv D^{\ell *}_{m'm} (\alpha,\beta,\gamma)
+\ee
+which represents the modulation function corresponding to the contribution of each precessing-frame mode $h^{P}_{\ell m}$ mode to the inertial-frame mode $h^{I}_{\ell m'}$ (see Eq.~\eqref{eq:wignerrot}).
+One can also define Fourier-domain tranfer functions for each mode contribution, according to
+\be\label{eq:defTlmlmp}
+	\mathrm{FT} \left[F_{\ell m, \ell m'} h \right] (f) \equiv \calT_{\ell m, \ell m'} (f) \tilde{h}(f) \,,
+\ee
+where $\mathrm{FT}$ stands sfor the Fourier transform.
+Fig.~\ref{fig:prectransfer} gives the amplitude and phase of the transfer functions $\calT_{22, 2 m'}$ for the two cases I and II.
+As expected, the difference between the two toy models for the post-merger frame precession is localized to the high frequency range.
+One can see a clear hierarchy in the amplitude of the inertial-frame modes generated by the precessing-frame 22 mode.
+This hierarchy come from the Wigner geometric factors and is dependent on the opening angle of the precession cone, or second Euler angle $\beta$, which in turn is dependent on the magnitude of the misaligned spins in the system.
+Case II shows stronger amplitude features in the merger and ringdown.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/prectransferAcaseI.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/prectransferPsicaseI.pdf}
+  %
+  \includegraphics[width=.48\linewidth]{plots/prectransferAcaseII.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/prectransferPsicaseII.pdf}  
+  \caption{Amplitude and phase of the Fourier-domain transfer fonctions $\calT_{22,2m'}$ as defined in~\eqref{eq:defTlmlmp} for individual mode contributions.
+The left panels show the amplitude and the right panels the phase.}
+  \label{fig:prectransfer}
+\end{figure}
 
@@
 \begin{align}
-	\frac{\ud^{3} \Psi}{\ud f^{3}} \partial_{t}^{3} F &= \calO(6) \,, \\
-	\frac{1}{A}\frac{\ud A}{\ud f} \partial_{t} F &= \calO(2) \,.
+	\frac{\ud^{3} \Psi}{\ud f^{3}} \frac{1}{F} \partial_{t}^{3} F &= \calO(6) \,, \\
+	\frac{1}{A}\frac{\ud A}{\ud f} \frac{1}{F} \partial_{t} F &= \calO(2) \,.
 \end{align}
@@
 
-[add plots and discussion of real figures of merit as measured from the model]
+\begin{figure}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/precfom22caseI.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/precfom21caseI.pdf}
+  %
+  \includegraphics[width=.48\linewidth]{plots/precfom22caseII.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/precfom21caseII.pdf}
+  \caption{Figures of merit of the approximation as defined in~\eqref{eq:deffom} for the precession modulations.
+The precessing-frame waveform is taken to be the $22$ mode of an equal-mass, non-spinning system.
+The two upper panels show the case where the frame rotation is freezed at merger (Case I), while the two lower panels show the case where it rotates after merger (Case II) (see Sec.~\ref{subsec:precmodel} for their definitions).
+The left panels show the figures of merit for $F_{22, 22}$, and the right panels for $F_{22, 21}$.
+The vertical line represents the frequency at merger.
+The amplitude corrections become non-negligible after merger, and become large enough to challenge a perturbative treatment in Case II for $F_{22,21}$ (notice the change of scale for the lower right panel).}
+  \label{fig:TfTA}
+\end{figure}
 
@@
 We localize this deviation to $f\leq f_{1}$, with $Mf_{1} = 0.06$ in our example, and in general the trigonometric representation will be accurate to the original signal only for $f\geq f_{1}$.
-We denote by $\mathcal{W}(f)$ this tapering, built in practice as the discrete integral of a cosine window function.
-To ensure good continuity properties, we perform an artificial symmetrization to a ficitious range $f\in [f_{0} - (f_{\rm max} - f_{0}), f_{0}]$, by imposing symmetric amplitudes and phases about $f_{0}$.
+This tapering is built in practice as the discrete integral of a cosine window function.
+To ensure good continuity properties, we perform an artificial symmetrization to a fictitious range $f\in [f_{0} - (f_{\rm max} - f_{0}), f_{0}]$, by imposing symmetric amplitudes and phases about $f_{0}$.
 Defining $\Delta f \equiv 2 (f_{\rm max} - f_{0})$, we write
@@
 	\begin{cases} 
-	\mathcal{W}(f) \exp\left[- i \left( \Psi(f_{0}) - 2i\pi (f-f_{0}) t_{p} \right) \right] \tilde{h}(f) \,,  &\text{ for } f \in [f_{0}, f_{0} + \Delta f /2] \\
-	\tilde{h}_{\rm sym}(2 f_{0} - f) \,,  &\text{ for } f \in [f_{0} - \Delta f/2, f_{0}]
+		\exp\left[- i \left( \Psi(f_{0}) - 2i\pi (f-f_{0}) t_{p} \right) \right] \tilde{h}(f) \,,  &\text{ for } f \in [f_{0}, f_{0} + \Delta f /2] \\
+	\tilde{h}_{\rm sym}(2 f_{0} - f)^{*} \,,  &\text{ for } f \in [f_{0} - \Delta f/2, f_{0}]
 	\end{cases}
 \ee
-Fig~\ref{fig:convolsymmetrized} shows both the original high-frequency part of the signal and its symmetrized version with a flattened amplitude.
 
@@
 where the condition $c_{M} = c_{-M}$ is an arbitrary condition enforced to match the number of degrees of freedom.
+Fig~\ref{fig:convolsymmetrized} shows both the original high-frequency part of the signal and its reconstruction as a trigonometric polynomial.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/trigamp.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/trigphase.pdf}
+  \caption{Amplitude (left panel) and phase (right panel) of the high-frequency part of $\tilde{h}_{\rm sym}$, compared to its trigonometric polynomial representation.}
+  \label{fig:trig}
+\end{figure}
 
@@
 The coefficients $c_{k}$ are built following the rules~\eqref{eq:ckyk} from the inverse DFT coefficients
-\be
+\be\label{eq:ykDFT}
 	y_{k} = \frac{1}{N} \sum\limits_{j=0}^{N-1} \tilde{h}_{\rm sym}\left( f_{0} + \frac{2j - N}{N} \Delta f \right) \omega^{jk} \,.
@@
 When inserting this representation~\eqref{eq:defhsym} and~\eqref{eq:hsymtrigo} into~\eqref{eq:convolution}, we obtain
-\be
+\be\label{eq:resultdirectconvol}
 	\tilde{s}(f) = e^{-i \Psi(f_{0})} e^{2i\pi (f-f_{0}) t_{p}} \sum\limits_{k=-M}^{+M} (-1)^{k} c_{k} e^{2i\pi k \frac{f-f_{0}}{\Delta f}} F(t_{p} + k\delta t) \,, 
@@
 
-In terms of performance, the implementation of~\eqref{} and~\eqref{} above is expected to have a very reasonable cost.
-We obtained results with a good accuracy, as shown in Fig.~\ref{}, by including 32 samples (64 samples in the artificially symmetrized signal).
-Since we only need to evaluate , these operations effectively amount to two DFT operations on $\sim$128 samples, to cover .
-This is of the same order of magnitude as the number of (logarithmically-spaced) frequency samples used to represent the Fourier-domain waveform with its amplitude and phase in the first place, thus the cost is comparable to the Taylor-like approach presented in Sec.~\ref{sec:quadphasedelay}.
+In terms of performance, the implementation of~\eqref{eq:ykDFT} and~\eqref{eq:resultdirectconvol} above is expected to have a very reasonable cost.
+We obtained results with a good accuracy, as shown in Fig.~\ref{fig:convolsymmetrized}, by including $M=32$ samples ($N=64$ samples in total for the artificially symmetrized signal).
+The calculation described here is done only once for a given waveform and modulation function.
+The number of samples is of the same order of magnitude as the number of (logarithmically-spaced) frequency samples used to represent the Fourier-domain waveform with its amplitude and phase in the first place, thus the cost is comparable to the Taylor-like approach presented in Sec.~\ref{sec:quadphasedelay}.
 
@@
 
-[show summary of amplitude/phase errors on the example case]
-[show mismatch calculations]
+\begin{figure}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/precerrorA22caseI.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/precerrorPsi22caseI.pdf}
+  %
+  \includegraphics[width=.48\linewidth]{plots/precerrorA21caseI.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/precerrorPsi21caseI.pdf}  
+  \caption{Errors in the Fourier-domain transfer fonctions $\calT_{22,22}$ and $\calT_{22,21}$ as defined in~\eqref{eq:defTlmlmp}, for Case I and for various approximations.
+The black curve is the exact result provided by the FFT.
+The blue curve gives the leading-order, local approximation.
+The red curve includes quadratic corrections in phase and in amplitude.
+The yellow curve shows the result of the direct conolution approach of Sec.~\ref{subsec:directconvolution}.
+The left panels show the amplitude and the right panels the phase difference with the FFT.
+Note that the vertical scales are not uniform.}
+  \label{fig:precerrorcaseI}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/precerrorA22caseII.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/precerrorPsi22caseII.pdf}
+  %
+  \includegraphics[width=.48\linewidth]{plots/precerrorA21caseII.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/precerrorPsi21caseII.pdf}  
+  \caption{Errors in the Fourier-domain transfer fonctions $\calT_{22,22}$ and $\calT_{22,21}$.
+Conventions are the same as in Fig.~\ref{fig:precerrorcaseI} but for Case II, with a more rapid post-merger frame precession.
+Note that the vertical scales are not uniform.}
+  \label{fig:precerrorcaseII}
+\end{figure}
+
+Figs.~\ref{fig:precerrorcaseI} and~\ref{fig:precerrorcaseII} show the amplitude and the phase error of the various reconstructions of the Fourier-domain transfer functions $\calT_{22,22}$ and $\calT_{22,21}$.
+Alongside with the exact result provided by the FFT of the time-domain precession modulation, we show the result obtained with the leading-order approximation (the local approximation) and with.
+We also show the result for the direct convolution approach presented in Sec.~\ref{subsec:directconvolution}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/precmmsimple.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/precmmasympt.pdf}
+  \caption{Mismatches computed between the exact FFT and the various approximations respresented in Figs.~\ref{fig:precerrorcaseI} and~\ref{fig:precerrorcaseII}.}
+  \label{fig:precmm}
+\end{figure}
 

### 2016-12-02 docs/paper_method.tex
Updates to the paper.

@@
 
-The gravitational waveform emitted in the direction $(\Theta, \Phi)$, with its two polarizations $h_{+},h_{\times}$, can be decomposed in spin-weighted spherical modes as~\cite{Thorne80}
+A useful representation of the gravitational waveform is given by its decomposition in spin-weighted spherical.
+The waveform emitted in the direction $(\Theta, \Phi)$, with its two polarizations $h_{+},h_{\times}$, can be decomposed as a superposition of modes as~\cite{Thorne80}
 \be\label{eq:defmodes}
-	h_{+} - i h_{\times} = \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{+\ell} {}_{-2}Y_{\ell m}(\Theta,\Phi) h_{\ell m} \,,
+	h_{+} - i h_{\times} = \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} {}_{-2}Y_{\ell m}(\Theta,\Phi) h_{\ell m} \,,
 \ee
 where the ${}_{-2}Y_{\ell m}(\Theta,\Phi)$ are spin-weighted spherical harmonics.
+For a non-precessing system, with a fixed orbital plane, the individual modes have the additional symmetry property $h_{\ell -m} = (-1)^{\ell} h_{\ell m}^{*}$.
 Since we will work in the Fourier domain, it will be useful to write the contributions of the individual modes to the Fourier transforms of the polarizations as
@@
 \begin{align}
-	\tilde{h}_{+}(f) &= \frac{1}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{+\ell} \left( {}_{-2}Y_{\ell m} \widetilde{h_{\ell m}}(f) + {}_{-2}Y_{\ell m}^{*} \widetilde{h_{\ell m}}(-f)^{*} \right) \,, \\
-	\tilde{h}_{\times}(f) &= \frac{i}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{+\ell} \left( {}_{-2}Y_{\ell m} \widetilde{h_{\ell m}}(f) - {}_{-2}Y_{\ell m}^{*} \widetilde{h_{\ell m}}(-f)^{*} \right) \,,
+	\tilde{h}_{+}(f) &= \frac{1}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left( {}_{-2}Y_{\ell m} \widetilde{h_{\ell m}}(f) + {}_{-2}Y_{\ell m}^{*} \widetilde{h_{\ell m}}(-f)^{*} \right) \,, \\
+	\tilde{h}_{\times}(f) &= \frac{i}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left( {}_{-2}Y_{\ell m} \widetilde{h_{\ell m}}(f) - {}_{-2}Y_{\ell m}^{*} \widetilde{h_{\ell m}}(-f)^{*} \right) \,,
 \end{align}
@@
 where we used $\widetilde{h_{\ell m}^{*}}(f) = \widetilde{h_{\ell m}}(-f)^{*}$.
-In the following, we will denote generically by $h$ an individual mode of the waveform, dropping the mode indices $\ell m$, and we will often refer to the Fourier-domain amplitude $A$ and phase $\Psi$ of the mode, according to
-\be
+An additional approximation often used in waveform models consists in considering that the Fourier transforms $\tilde{h}_{\ell m}$ have support only on one side of the spectrum, either for positive of for negative frequencies depending on the sign of $m$.
+This holds in particular within the stationary phase approximation, as explained below.
+With our sign convention~\eqref{eq:defFT}, this approximation reads
+\begin{align}\label{eq:approxnegativef}
+	\tilde{h}_{\ell m} (f) &\simeq 0 \text{ for } f<0, \; m>0 \nn\,,\\
+	\tilde{h}_{\ell m} (f) &\simeq 0 \text{ for } f>0, \; m<0 \,.
+\end{align}
+
+In the following, we will drop the mode indices and the sum over modes, and we will denote generically by $h$ an individual mode of the waveform\footnote{This notation for a generic individual mode is not to be confused with the often encountered notation $h\equiv h_{+} - i h_{\times}$.}.
+Note that, if we are considering real signals $s(t)$ (such as $h_{+}$, $h_{\times}$), we can focus for instance on the positive-frequencies range and reconstruct the negative frequencies as $\tilde{s}(-f) = \tilde{s}(f)^{*}$.
+For positive frequencies $f>0$, we will use for $h$ the notation
+\begin{align}\label{eq:notationhsignm}
+	h(t) &\equiv h_{\ell m}(t) \,, \quad \tilde{h}(f) \equiv \tilde{h}_{\ell m}(f) \text{ for } m>0 \,, \nn\\
+	h(t) &\equiv h_{\ell m}^{*}(t) \,, \quad \tilde{h}(f) \equiv \tilde{h}_{\ell m}(-f)^{*} \text{ for } m<0 \,.
+\end{align}
+We will further decompose $\tilde{h}$ into a Fourier-domain amplitude $A$ and a phase $\Psi$ according to
+\be\label{eq:defAPsi}
 	\tilde{h}(f) \equiv A(f) e^{-i\Psi(f)} \,.
 \ee
+
 Since we intend to work in the Fourier domain, we will use PhenomD waveforms~\cite{Khan+15}.
@@
 \ee
-where the tilde here does not indicate a Fourier transform operation but simply that this quantity is an effective time defined from the Fourier-domain waveform.
 This definition is a generalization of the time-to-frequency correspondence at the heart of the SPA, as will be discussed in Sec.~\ref{subsec:linkSPA}, but it is worth noting that here it is defined entirely from the Fourier-domain waveform.
@@
 We now turn to the transposition of~\eqref{eq:yslr} to the Fourier-domain.
-We will choose to write the response directly in terms of spin-weighted spherical modes $h_{\ell m}$, whose Fourier transforms are assumed to have a smooth amplitude and phase.
-Since the orbital delay due to the orbit around the Sun is common to all $y_{slr}$ observables, we will decompose the response in the orbital delay and the constellation-centered response as
-\begin{subequations}
-\begin{align}
-	h^{\rm TT}_{O} (t) &\equiv h^{\rm TT}(t-\hat{k}\cdot p_{O}) \,, \label{eq:dfresponseO} \\
-	y_{slr} &= \frac{1}{2} \frac{1}{1 - \hat{k}\cdot n_{l}} n_{l}\cdot \left[ h_{O}^{\rm TT}(t - \hat{k}\cdot p^{L}_{s}) - h_{O}^{\rm TT}(t - \hat{k}\cdot p^{L}_{r}) \right] \cdot n_{l}\,, \label{eq:dfresponsec}
-\end{align}
-\end{subequations}
+Since the orbital delay due to the orbit around the Sun is common to all $y_{slr}$ observables, we will decompose the full instrument response in two steps.
+The first step consists simply in applying the varying time delay to bring the signal from the SSB to the center of the constellation.
+We write this orbital time delay as
+\be\label{eq:defresponseO}
+	h^{\rm TT}_{O} (t) = h^{\rm TT}(t-\hat{k}\cdot p_{O}) \,.
+\ee
+The second step consists in the remaining, constellation-centered response, which we write as
+\be\label{eq:dfresponsec}
+	y_{slr} = \frac{1}{2} \frac{1}{1 - \hat{k}\cdot n_{l}} n_{l}\cdot \left[ h_{O}^{\rm TT}(t - L - \hat{k}\cdot p^{L}_{s}) - h_{O}^{\rm TT}(t - \hat{k}\cdot p^{L}_{r}) \right] \cdot n_{l}\,,
+\ee
 where we introduced the positions of the spacecrafts relative to the center of the constellation, $p^{L}_{A} \equiv p_{A} - p_{O}$.
 
+We will decompose the full signal in the contributions of the individual spin-weighted spherical modes $h_{\ell m}$, whose Fourier transforms are assumed to have a smooth amplitude and phase.
+First, we define the matrices $P_{+},P_{\times}$ such that, in the sense of matrices,
+\be
+	h^{\rm TT} = h_{+}P_{+} + h_{\times}P_{\times} \,.
+\ee
+Assuming that the approximation~\eqref{eq:approxnegativef} applies, we focus only on positive frequencies and consider a mode contribution $h$ following~\eqref{eq:notationhsignm}.
+We can define a complex matrix $P$ incorporating the spin-weighted spherical harmonic constant factor as
+\be
+	P = 
+	\begin{cases} 
+	\frac{1}{2} {}_{-2}Y_{\ell m} \left( P_{+} + i P_{\times} \right) \text{ for } m>0\,,\\
+	\frac{1}{2} {}_{-2}Y_{\ell m}^{*} \left( P_{+} - i P_{\times} \right) \text{ for } m<0\,.
+	\end{cases}
+\ee
+
 When using the rigid approximation for the constellation, a particular simplification occurs when writing the transfer function at leading order in the delays.
-First, we define the matrices $P^{+},P^{\times}$ such that
-\be
-	h^{\rm TT} = h_{+}P_{+} + h_{\times}P_{\times} \,,
-\ee
-in the sense of matrices.
-The spin-weighted spherical harmonics factors in~\eqref{eq:hpcfrommodes} are mere constants of time, so we can also define a geometric projection factor $P_{slr}^{\ell m}$ for each individual mode.
-Under the assumption that we can neglect the negative frequencies for the modes $m>0$ and the positive frequencies for $m<0$ (which is also used for precessing waveforms), we have for $f>0$ and $m>0$
-%\begin{subequations}
-\begin{align}
-	P_{slr}^{\ell m} &= \frac{1}{4} \frac{1}{1-\hat{k}\cdot n_{l}} {}_{-2}Y_{\ell m} n_{l}\cdot \left( P_{+} - i P_{\times} \right) \cdot n_{l} \text{ for } m>0 \,,
-\end{align}
-%\end{subequations}
-the transfer function for~\eqref{eq:dfresponsec} reads in the local approximation
-\be
-	\calT_{slr}^{\ell m} = \frac{i \pi f L}{2} \sinc \left[ \pi f L\left(1-\hat{k}\cdot n_{l} \right) \right] \exp\left[ \pi f \left( L + \hat{k}\cdot \left( p_{1}^{L} + p_{2}^{L} \right) \right) \right] {}_{-2}Y_{\ell m} n_{l}\cdot \left( P_{+} - i P_{\times} \right) \cdot n_{l} \text{ for } m>0 \,,
+The Fourier-domain transfer function for~\eqref{eq:dfresponsec} then reads in the local approximation
+\be
+	\calT_{slr}^{\ell m} = \frac{i \pi f L}{2} \sinc \left[ \pi f L\left(1-\hat{k}\cdot n_{l} \right) \right] \exp\left[ i \pi f \left( L + \hat{k}\cdot \left( p_{1}^{L} + p_{2}^{L} \right) \right) \right]  n_{l} \cdot P_{\ell m} \cdot n_{l} \text{ for } m>0 \,,
 \ee
 where we used the fact that, in our rigid approximation for the trajectory of the constellation, $p^{L}_{r} - p^{L}_{s} =  L n_{l}$.
+In the case $m<0$, the formalism can be applied in the same way, except that notations are complicated by the fact that the mode $\tilde{h}_{\ell m}$ has to be evaluated at $-f$.
+In the following, we will limit ourselves to the case $m>0$ for simplicity of presentation.
 
@@
 The approximation that consists in taking the first term in the series, the LLP approximation~\eqref{} which was in particular used for the PhenomP~\cite{} waveform model, gets {\it worse} the further away from merger the signal is.
-[This does not seem to have been clearly stated before, to check].
 We investigate further in Sec.~\ref{} the validity of the treatment of precession used for PhenomP.

### 2016-12-02 EOBNRv2HMROM/GenerateWaveform.c

@@ -42,15 +42,17 @@ Arguments are as follows:\n\
 --------------------------------------------------\n\
 ----- Generation Parameters ----------------------\n\
 --------------------------------------------------\n\
- --nbmode              Number of modes of radiation to generate (1-5, default=5)\n\
+ --nbmode              Number of modes of radiation to generate (1-5, default=1)\n\
  --minf                Minimal frequency, ignore if 0 (Hz, default=0) - will use first frequency covered by the ROM if higher\n\
  --maxf                Maximal frequency, ignore if 0 (Hz, default=0) - will use last frequency covered by the ROM if lower\n\
  --deltatobs           Observation duration (years, default=2)\n\
  --tagextpn            Tag to allow PN extension of the waveform at low frequencies (default=1)\n\
  --Mfmatch             When PN extension allowed, geometric matching frequency: will use ROM above this value. If <=0, use ROM down to the lowest covered frequency (default=0.)\n\
  --taggenwave          Tag choosing the wf format: hlm (default: downsampled modes, Amp/Phase form),  h22TD (IFFT of h22, Amp/Phase form - currently not supported for higher modes), hphcFD (hlm interpolated and summed, Re/Im form), hphcTD (IFFT of hphcFD, Re/Im form)\n\
- --f1windowend         If generating h22TD/hphcTD, start frequency for windowing at the end - set to 0 to ignore and use 0.995*fHighROM, where fHighROM is the highest frequency covered by the ROM (Hz, default=0)\n\
- --f2windowend         If generating h22TD/hphcTD, stop frequency for windowing at the end - set to 0 to ignore and use fHighROM, where fHighROM is the highest frequency covered by the ROM (Hz, default=0)\n\
+ --f1windowbeg         If generating h22TD/hphcTD, start frequency for windowing at the beginning - set to 0 to ignore and use max(fstartobs, fLowROM, minf), where fLowROM is either the lowest frequency covered by the ROM or simply minf if PN extension is used (Hz, default=0)\n\
[…]

### 2016-12-02 EOBNRv2HMROM/GenerateWaveform.h

@@ -63,15 +63,16 @@ typedef struct tagGenWaveParams {
   double deltatobs;          /* Observation duration (years, default=2) */
   int tagextpn;              /* Tag to allow PN extension of the waveform at low frequencies */
   double Mfmatch;            /* When PN extension allowed, geometric matching frequency: will use ROM above this value. If <=0, use ROM down to the lowest covered frequency */
-  int nbmode;                /* number of modes to generate (starting with 22) - defaults to 5 (all modes) */
+  int nbmode;                /* Number of modes to generate (starting with 22) - defaults to 1 (22 only) */
   int taggenwave;            /* Tag selecting the desired output format */
-  double f1windowend;        /* If generating hphcTD, start frequency for windowing at the end - set to 0 to ignore and use 0.995*fHighROM, where fHighROM is the highest frequency covered by the ROM (Hz, default=0) */
-  double f2windowend;        /* If generating hphcTD, stop frequency for windowing at the end - set to 0 to ignore and use fHighROM, where fHighROM is the highest frequency covered by the ROM (Hz, default=0) */
+  double f1windowbeg;        /* If generating h22TD/hphcTD, start frequency for windowing at the beginning - set to 0 to ignore and use max(fstartobs, fLowROM, minf), where fLowROM is either the lowest frequency covered by the ROM or simply minf if PN extension is used (Hz, default=0) */
+  double f2windowbeg;        /* If generating h22TD/hphcTD, stop frequency for windowing at the beginning - set to 0 to ignore and use 1.1*f1windowbeg (Hz, default=0) */
+  double f1windowend;        /* If generating h22TD/hphcTD, start frequency for windowing at the end - set to 0 to ignore and use 0.995*f2windowend (Hz, default=0) */
+  double f2windowend;        /* If generating h22TD/hphcTD, stop frequency for windowing at the end - set to 0 to ignore and use min(maxf, fHighROM), where fHighROM is the highest frequency covered by the ROM (Hz, default=0) */
   int binaryout;             /* Tag for outputting the data in gsl binary form instead of text (default 0) */
   char outdir[256];          /* Path for the output directory */
   char outfile[256];         /* Path for the output file */
 } GenWaveParams;
-
 
 #if 0
 { /* so that editors will match succeeding brace */

### 2016-12-02 LISAinference/LISAlikelihood.c

@@ -12,7 +12,7 @@
 /* This program works compatibly with LISAinference, but takes a list of 9 param values (in order as below) and only
    computes the likelihood at that parameter point.
 */
-noMPI=1;
+int noMPI=1;
 
 
 int main(int argc, char *argv[])

### 2016-12-02 LISAsim/GenerateTDIFD.c

@@ -59,7 +59,7 @@ Arguments are as follows:\n\
  --tagtdi              Tag choosing the set of TDI variables to use (default TDIAETXYZ)\n\
  --taggenwave          Tag choosing the wf format: TDIhlm (default: downsampled mode contributions to TDI in Amp/Phase form), TDIFD (hlm interpolated and summed accross modes), h22FFT and yslrFFT (used with fomrtditdfile -- loads time series and FFT)\n\
  --restorescaledfactor Option to restore the factors scaled out of TDI observables (default: false)\n	\
- --fromtditdfile       Option for loading time series for TDI observables and FFTing (default: false)\n\
+ --fromtdfile          Option for loading time series and FFTing (default: false)\n\
  --nsamplesinfile      Number of lines of input file when loading TDI time series from file\n\
  --binaryin            Tag for loading the data in gsl binary form instead of text (default false)\n\
  --binaryout           Tag for outputting the data in gsl binary form instead of text (default false)\n\
@@ -96,7 +96,7 @@ Arguments are as follows:\n\
     params->tagtdi = TDIXYZ;
     params->taggenwave = TDIhlm;
     params->restorescaledfactor = 0;
-    params->fromtditdfile = 0;
+    params->fromtdfile = 0;
     params->nsamplesinfile = 0;    /* No default; has to be provided */
     strcpy(params->indir, "");   /* No default; has to be provided */
     strcpy(params->infile, "");  /* No default; has to be provided */
@@ -152,8 +152,8 @@ Arguments are as follows:\n\
 	  params->taggenwave = ParseGenTDIFDtag(argv[++i]);
         } else if (strcmp(argv[i], "--restorescaledfactor") == 0) {
             params->restorescaledfactor = 1;
-        } else if (strcmp(argv[i], "--fromtditdfile") == 0) {
-            params->fromtditdfile = 1;
+        } else if (strcmp(argv[i], "--fromtdfile") == 0) {
+            params->fromtdfile = 1;
         } else if (strcmp(argv[i], "--nsamplesinfile") == 0) {
             params->nsamplesinfile = atoi(argv[++i]);
[…]

### 2016-12-02 LISAsim/GenerateTDIFD.h

@@ -75,7 +75,7 @@ typedef struct tagGenTDIFDparams {
   int tagtdi;                /* Tag selecting the desired output format */
   int taggenwave;            /* Tag selecting the desired output format */
   int restorescaledfactor;   /* If 1, restore the factors that were scaled out of TDI observables */
-  int fromtditdfile;         /* Tag for loading time series for TDI observables and FFTing */
+  int fromtdfile;            /* Option for loading time series and FFTing (default: false) */
   int nsamplesinfile;        /* Number of lines of input file */
   int binaryin;              /* Tag for loading the data in gsl binary form instead of text (default false) */
   int binaryout;             /* Tag for outputting the data in gsl binary form instead of text (default false) */

### 2016-12-02 LISAsim/GenerateTDITD.c

@@ -122,29 +122,6 @@ static void Read_Wave_hphcTD( RealTimeSeries** hptd, RealTimeSeries** hctd, cons
   /* Clean up */
   gsl_matrix_free(inmatrix);
 }
-/* Read waveform time series in Amp/Phase form for a single mode h22TD */
-/* NOTE: assumes a single mode amp/phase in the input file */
-static void Read_Wave_h22TD( AmpPhaseTimeSeries** h22td, const char dir[], const char file[], const int nsamples, int binary)
-{
-  /* Initalize and read input */
-  gsl_matrix* inmatrix =  gsl_matrix_alloc(nsamples, 3);
-  if (!binary) Read_Text_Matrix(dir, file, inmatrix);
-  else Read_Matrix(dir, file, inmatrix);
-
-  /* Initialize structures */
-  AmpPhaseTimeSeries_Init(h22td, nsamples);
-
-  /* Set values */
-  gsl_vector_view timesview = gsl_matrix_column(inmatrix, 0);
-  gsl_vector_view ampview = gsl_matrix_column(inmatrix, 1);
-  gsl_vector_view phaseview = gsl_matrix_column(inmatrix, 2);
-  gsl_vector_memcpy((*h22td)->times, &timesview.vector);
-  gsl_vector_memcpy((*h22td)->h_amp, &ampview.vector);
-  gsl_vector_memcpy((*h22td)->h_phase, &phaseview.vector);
-
-  /* Clean up */
-  gsl_matrix_free(inmatrix);
-}
 /* Output TDI 3 channels TD - real time series */
 /* NOTE: assumes 3 channels with same times */
 static void Write_TDITD(const char dir[], const char file[], RealTimeSeries* TDI1, RealTimeSeries* TDI2, RealTimeSeries* TDI3, int binary)
@@ -164,39 +141,7 @@ static void Write_TDITD(const char dir[], const char file[], RealTimeSeries* TDI
   if(!binary) Write_Text_Matrix(dir, file, outmatrix);
   else Write_Matrix(dir, file, outmatrix);
 }
-/* Output waveform time series in Amp/Phase form for a single mode h22TD */
-static void Write_h22TD(const char dir[], const char file[], AmpPhaseTimeSeries* h22td, int binary)
-{
-  /* Initialize output */
-  /* NOTE: assumes identical times for all 3 TDI observables */
-  int nbtimes = h22td->times->size;
[…]

### 2016-12-02 tools/fft.c

@@ -166,17 +166,31 @@ int FFTTimeSeries(ReImFrequencySeries** freqseries, ReImTimeSeries* timeseries, 
   int nzeros = (int) pow(2, ((int) ceil(log(n)/log(2))) + nzeropad) - n; /* Here defined with ceil, but with floor in IFFT */
   int N = n + nzeros;
 
+  //
+  printf("N: %d\n", N);
+
   /* Compute input TD values, with windowing */
   int nbptswindowbeg = (int) ceil(twindowbeg/deltat) + 1;
   int nbptswindowend = (int) ceil(twindowend/deltat) + 1;
+
+  //
+  printf("twindowbeg,twindowend, deltat: %g, %g, %g\n", twindowbeg,twindowend,deltat);
+  printf("nbptswindowbeg,nbptswindowend: %d, %d\n", nbptswindowbeg,nbptswindowend);
+
   double t1windowbeg = times[0];
   double t2windowbeg = times[nbptswindowbeg-1];
   double t1windowend = times[n-nbptswindowend];
   double t2windowend = times[n-1];
   double deltatwindowbeg = t2windowbeg - t1windowbeg;
   double deltatwindowend = t2windowend - t1windowend;
+  //
+  printf("N: %d\n", N);
+
   double* htdreal = timeseries->h_real->data;
   double* htdimag = timeseries->h_imag->data;
+  //
+  printf("N: %d\n", N);
+
   fftw_complex* in = (fftw_complex*) fftw_malloc(sizeof(fftw_complex) * N);
   for (int i=0; i<nbptswindowbeg; i++) {
     in[i] = WindowFunctionRight(times[i], t1windowbeg, deltatwindowbeg) * (htdreal[i] + I*htdimag[i]);
@@ -188,6 +202,9 @@ int FFTTimeSeries(ReImFrequencySeries** freqseries, ReImTimeSeries* timeseries, 
     in[i] = WindowFunctionLeft(times[i], t2windowend, deltatwindowend) * (htdreal[i] + I*htdimag[i]);
   }
 
+  //
+  printf("N: %d\n", N);
+
   /* FFT - uses flipped convention (i.e. h(f) = int e^(+2ipift)h(t)) */
   /* Represented here by the use of FFTW_BACKWARD (plus sign in the exp) */
   fftw_plan p;
@@ -196,8 +213,14 @@ int FFTTimeSeries(ReImFrequencySeries** freqseries, ReImTimeSeries* timeseries, 
[…]

### 2016-12-02 tools/struct.c

@@ -50,16 +50,18 @@ void Err_Handler(const char *reason, const char *file, int line, int gsl_errno) 
 /* Functions to read binary data from files */
 int Read_Vector(const char dir[], const char fname[], gsl_vector *v) {
   char *path=malloc(strlen(dir)+64);
-
   sprintf(path,"%s/%s", dir, fname);
   FILE *f = fopen(path, "rb");
   if (!f) {
-      return(FAILURE);
+    fprintf(stderr, "Error reading data from %s\n", path);
+    free(path);
+    return(FAILURE);
   }
   int ret = gsl_vector_fread(f, v);
   if (ret != 0) {
-      fprintf(stderr, "Error reading data from %s.\n",path);
-      return(FAILURE);
+    fprintf(stderr, "Error reading data from %s.\n",path);
+    free(path);
+    return(FAILURE);
   }
   fclose(f);
   free(path);
@@ -67,16 +69,18 @@ int Read_Vector(const char dir[], const char fname[], gsl_vector *v) {
 }
 int Read_Matrix(const char dir[], const char fname[], gsl_matrix *m) {
   char *path=malloc(strlen(dir)+64);
-
   sprintf(path,"%s/%s", dir, fname);
   FILE *f = fopen(path, "rb");
   if (!f) {
-      return(FAILURE);
+    fprintf(stderr, "Error reading data from %s\n", path);
+    free(path);
+    return(FAILURE);
   }
   int ret = gsl_matrix_fread(f, m);
   if (ret != 0) {
-      fprintf(stderr, "Error reading data from %s.\n", path);
-      return(FAILURE);
+    fprintf(stderr, "Error reading data from %s\n", path);
+    free(path);
+    return(FAILURE);
   }
   fclose(f);
   free(path);
@@ -85,16 +89,18 @@ int Read_Matrix(const char dir[], const char fname[], gsl_matrix *m) {
 /* Functions to read text data from files */
 int Read_Text_Vector(const char dir[], const char fname[], gsl_vector *v) {
   char *path=malloc(strlen(dir)+64);
-
   sprintf(path,"%s/%s", dir, fname);
   FILE *f = fopen(path, "rb");
   if (!f) {
-      return(FAILURE);
+    fprintf(stderr, "Error reading data from %s\n", path);
+    free(path);
+    return(FAILURE);
   }
[…]

### 2016-12-02 tools/struct.h

@@ -211,6 +211,18 @@ void RealTimeSeries_Init(
 	 const int n );                /* length of the time series */
 void RealTimeSeries_Cleanup(RealTimeSeries* timeseries);
 
+/***********************************************************************/
+/**************** I/O functions for internal structures ****************/
+
+/* Note: at the moment, requires external input for the number of lines in the data */
+int Read_AmpPhaseTimeSeries(AmpPhaseTimeSeries** timeseries, const char dir[], const char file[], const int nblines, const int binary);
+int Read_ReImTimeSeries(ReImTimeSeries** timeseries, const char dir[], const char file[], const int nblines, const int binary);
+int Read_RealTimeSeries(RealTimeSeries** timeseries, const char dir[], const char file[], const int nblines, const int binary);
+int Write_ReImFrequencySeries(const char dir[], const char file[], ReImFrequencySeries* freqseries, const int binary);
+int Write_RealTimeSeries(const char dir[], const char file[], RealTimeSeries* timeseries, int binary);
+int Write_AmpPhaseTimeSeries(const char dir[], const char file[], AmpPhaseTimeSeries* timeseries, int binary);
+
+
 #if 0
 { /* so that editors will match succeeding brace */
 #elif defined(__cplusplus)

### 2016-12-02 tools/waveform.c

@@ -524,36 +524,47 @@ int RestrictFDReImFrequencySeries(
 /* WARNING : found significant accumulating difference with the numpy unwrap function - to be understood */
 int UnwrapPhase(
   gsl_vector* phaseout,    /* Output: unwrapped phase vector, already allocated */
-  gsl_vector*  phasein)    /* Input: phase vector */
+  gsl_vector* phasein)     /* Input: phase vector */
 {
   size_t N = phasein->size;
-  double* diff = (double*) malloc(sizeof(double) * N);
-  double* diffcorr = (double*) malloc(sizeof(double) * N);
-  double* cumsum = (double*) malloc(sizeof(double) * N);
   double* p = phasein->data;
+  double* pmod = (double*) malloc(sizeof(double) * N);
+  int* jumps = (int*) malloc(sizeof(int) * N);
+  int* cumul = (int*) malloc(sizeof(int) * N);
 
-  /* Incremental phase variation - indexed from 1 to N-1 */
-  diff[0] = 0;
+  /* Compute phase mod 2pi (shifted to be between -pi and pi) */
   for(int i=1; i<N; i++) {
-    diff[i] = p[i+1] - p[i];
+    pmod[i] = p[i] - floor((p[i] + PI) / (2*PI))*(2*PI) - PI;
   }
 
-  /* Compute phase variation mod 2pi (shifted to be between -pi and pi) */
+  /* Identify jumps */
+  jumps[0] = 0;
+  double d = 0.;
   for(int i=1; i<N; i++) {
-    diff[i] = (diff[i] + PI) - floor((diff[i] + PI) / (2*PI))*(2*PI) - PI;
+    d = pmod[i] - pmod[i-1];
+    if(d<PI) jumps[i] = 1;
+    else if(d>PI) jumps[i] = -1;
+    else jumps[i] = 0;
   }
 
-  /* One could preserve sign of variation if diff[i] = +pi (maps to -pi here) - here we simply assume that the phase jumps are strictly smaller than +-pi, which is the limiting case for the reconstruction to work  */
+  /* Cumulative of the jump sequence */
+  int c = 0;
+  cumul[0] = 0;
+  for(int i=1; i<N; i++) {
+    c += jumps[i];
+    cumul[i] = c;
+  }
 
-  /* Cumulative sum of diff */
[…]

### 2016-12-02 tools/waveform.h

@@ -139,6 +139,10 @@ int UnwrapPhase(
 int ReImTimeSeries_ToAmpPhase(
   AmpPhaseTimeSeries** timeseriesout,             /* Output: Amp/Phase time series */
   ReImTimeSeries* timeseriesin);                  /* Input: Re/Im time series */
+/* Function to convert a time series from Amp/Phase form to Re/Im form */
+int AmpPhaseTimeSeries_ToReIm(
+  ReImTimeSeries** timeseriesout,                 /* Output: Re/Im time series */
+  AmpPhaseTimeSeries* timeseriesin);              /* Input: Amp/Phase time series */
 
 /***************** Spin weighted spherical harmonics ****************/
 

### 2017-01-09 docs/paper_method.tex

@@
 
-\title{Modulations and delays in Fourier domain: precessing binaries and LISA-type detector response}
-
-\author{A. A. Author}
-\affiliation{U. U. University}
-
-%\author{John G. Baker}
-%\affiliation{Gravitational Astrophysics Laboratory, NASA Goddard Space Flight Center, 8800 Greenbelt Rd., Greenbelt, MD 20771, USA}
-%\author{Alessandra Buonanno}
-%\affiliation{Department of Physics, University of Maryland, College Park, MD 20742, USA}
-%\affiliation{Max Planck Institute for Gravitational Physics (Albert Einstein Institute), Am M\"uhlenberg 1, Potsdam-Golm, 14476, Germany}
-%\author{Philip. B. Graff}
-%\affiliation{Department of Physics, University of Maryland, College Park, MD 20742, USA}
-%\affiliation{Gravitational Astrophysics Laboratory, NASA Goddard Space Flight Center, 8800 Greenbelt Rd., Greenbelt, MD 20771, USA}
-%\author{Sylvain Marsat}
-%\affiliation{Max Planck Institute for Gravitational Physics (Albert Einstein Institute), Am M\"uhlenberg 1, Potsdam-Golm, 14476, Germany}
+\title{Modulations of gravitational-wave signals: a Fourier-domain approach for the response of LISA-type detectors and for precessing binaries}
+
+\author{Sylvain Marsat}
+\affiliation{Max Planck Institute for Gravitational Physics (Albert Einstein Institute), Am M\"uhlenberg 1, Potsdam-Golm, 14476, Germany}
+\author{John G. Baker}
+\affiliation{Gravitational Astrophysics Laboratory, NASA Goddard Space Flight Center, 8800 Greenbelt Rd., Greenbelt, MD 20771, USA}
 
@@
 
-[Abstract]
+We present a Fourier-domain based approach to the problem of the modulation of gravitational wave signals.
+The two main contexts where modulations are important for gravitational-wave astronomy: (i) the time-dependent response of space-based instruments such as LISA; and (ii) the transposition in Fourier domain of the frame decomposition of precessing waveforms, where the precessing waveform is approximated by rotating a non-precessing waveform to follow the precession of the orbital plane.
+Our results recover and generalize previous approaches based on an extension of the stationary phase approximation, and can be applied to full inspiral-merger-ringdown signals.
+In the case of the response LISA-type detectors, we show that higher-order corrections remain small except at the two extremes of the detector's frequency band, and we show that including them reduce the errors.
+In the case of precessing binaries, we perform a limited exploration of the merger-ringdown range based on a toy model for the precession.
+We find that, depending on the post-merger frame velocity, higher-order corrections can become quantitatively large and even challenge our perturbative formalism.
+We develop a direct convolution approach to capture these post-merger features.
+However, in that case these corrections appear mainly as amplitude corrections and give raise to a small unfaithfulness.
 
@@
 
-[Introduction]
+With the recent detections of gravitational-wave signals emitted by the coalescences of binary black holes, announced by the LSC collaboration after the first obervation run of the two LIGO instruments~\cite{LIGO-theevent-2016,LIGO-christmasevent-2016,LIGO-O1BBH-2016}, gravitational-wave astronomy has entered its observational era.
+With the recent start of LIGO's second observation run, and later the extension of the ground-based detectors network with Virgo~\cite{Virgo}, KAGRA~\cite{KAGRA} and LIGO-India~\cite{INDIGO}, many more detections from compact object coalescences are to come, with binary black holes being expected to be the dominant sources.
+Moreover, the European Space Agency has issued a mission call for its L3 mission, adressing the science theme ``The Gravitational Universe'' with a tentative launch in 2034, possibly opening the way for a space-borne observatory based on the concept of the Laser Interferometer Space Antenna (LISA)~\cite{LISA09}.
+Among other sources of gravitational waves, LISA will be able to detect and characterize binary black hole coalescences to cosmological distances over a wide range of masses, from high-redshift observations of supermassive black hole binaries with $M\sim 10^{7} \Msol$ down to the observation of LIGO-type sources with $M\sim 10^{2} \Msol$~\cite{Sesana16}.
+
+The Bayesian analysis and parameter recovery of gravitational-wave signals may require millions of evaluations of waveform templates to sample the posterior probability distribution, which poses a computational challenge for templates including consistently the inspiralling phase of the evolution, the merger and the ringdown (IMR).
+State-of-the-art IMR templates incorporate information from both the perturbative results of post-Newtonian theory covering the inspiral (see e.g.~\cite{BlanchetLiving}) and from numerical relativity simulations covering the end of the inspiral and the merger-ringdown phase~\cite{}.
+The two main approaches currently used to produce complete waveform models are (i) the Effective-One-Body (EOB) approach~\cite{BD99}, which incorporates a resummation of post-Newtonian results and a calibration to numerical relativity, and which has been extended to include both spins of the two bodies~\cite{Taracchini+13, Pan+13, Bohe+16} (ii) phenomenological templates, which extend post-Newtonian signals by postulating a Fourier-domain analytic expression for the amplitude and phase and by fitting coefficients against numerical relativity waveforms~\cite{Husa+15, Khan+15}, and which include a simplified model for precession~\cite{Hannam+13}.
+The second type of models produce directly the amplitude and the phase of the Fourier-domain waveform in the form of a piecewise analytic fit.
+EOB waveforms, by contrast, require to integrate ordinary differential equations in the time-domain, which can be time-consuming.
+For the non-precessing case, Reduced Order Models (ROM), also called surrogate models, have been developed to speed up considerably the waveform generation, directly in the Fourier domain, without losing accuracy~\cite{Field+13, Puerrer14, Bohe+16}.
+
+A complete representation of the spin effects in fast IMR templates still remains a frontier of gravitational wave signal modelling.
+In presence of misaligned spins, the system will endure precession of the orbital plane as it evolves, leading to a modulation of the signal that is more pronunced for observers that see the system as close to edge-on~\cite{ACST94}.
+Precessing EOB templates can represent all six degrees of freedom of the two spins~\cite{Pan+13, BTB16} but still require a costly time-domain integration; while precessing phenomenological templates~\cite{Hannam+13} use effective spin parameters representing only the dominant effects of the spins.
+To model the effect of precession on the emitted waveform, it has been proposed~\cite{BCV03b, BCPTV05, Schmidt+10, OShaughnessy+11, Boyle+11} to decompose precessing waveforms by performing a time-dependent rotation, following the precession of the orbital plane.
+The resulting waveform in the rotated frame is then close to a non-precessing waveform, an approximation that is used both to build the inspiral part of precessing EOB waveforms~\cite{Pan+13} and in the construction of precessing phenomenological waveforms~\cite{Hannam+13}.
+To carry over this approach to Fourier-domain waveforms, one needs to understand how to translate the time-domain modulation created by the frame rotation into a Fourier-domain modulation.
+The most straightforward method consists in using a time-to-frequency correspondence such as the one given in the Stationary Phase Approximation (SPA), a good approximation for the inspiralling phase.
+This is what is used implicitly in Ref.~\cite{Hannam+13}, with the difference that directly frequency-dependent expressions for the modulation are used, which are then extended to also cover high frequencies in the merger part of the signal.
+In Ref.~\cite{Klein+14}, a formalism is introduced to go beyond the SPA and compute more accurately the modulation in the Fourier domain; however, this formalism still relies on the SPA for the underlying signal.
+How to implement
+
+Another frontier of gravitational wave signal modelling is prospective parameter recovery for LISA observations.
+In the past, LISA data analysis studies for comparable-mass binaries used mainly simplified, inspiral-only models~\cite{}, and Fisher-matrix estimates instead of a full Bayesian analysis.
+In Ref.~\cite{Klein+15}, an improved model was used, that included the effects of both spins including the corrections beyond the SPA investigated in~\cite{Klein+14}; the limitation to inspiral-only waveforms was compensated by SNR reweighting using full IMR waveforms.
+The response of the LISA instrument, by contrast with the response of ground-based detectors, has the specificity of being time-dependent.
+For a signal accumulating over several months, the motion and change of orientation of the detector constellation along its orbit lead to time variability in the response of the detector, which is crucial for the localization of the gravitational-wave source in the sky.
+More precisely, these effects can be separated as a modulation coming from the rotation of the constellation, and a delay coming from the motion of the spacecrafts in the wavefront of the incoming wave.
+
+The purpose of this paper is to propose a formalism allowing to transpose directly in the Fourier domain the processing of a signal through a time-domain modulation and delay, while retaining the compactness of a Fourier-domain amplitude and phase representation of the signals.
+This allows us to address both the issue of a direct Fourier-domain response of the LISA instrument, as well as the issue of Fourier-domain precession modulation for IMR signals.
+
+[plan]
+The plan of the paper is as follows.
+In Sec.~\ref{sec:motivation}, we will introduce our notations and conventions for Fourier-domain waveforms, motivate the problem of Fourier-domain modulation and give the relevant timescales for both the response of LISA-type detectors and the modulation of precessing signals.
+In Sec.~\ref{sec:formalism}, we present our general formalism, give its leading order approximation as well as higher-order corrections, and introduce new timescales based on the Fourier-domain signal.
+In Sec.~\ref{sec:quadphasedelay}, we refine the previous results for both the quadratic-in-phase corrections and the treatment of the delays, and make the connection to previous approaches of the problem.
+We then apply our formalism to the response of the LISA detector in Sec.~\ref{sec:LISA}, and to the case of signals from precessing binaries in Sec~\ref{sec:precession}.
+We discuss and summarize our results in Sec.~\ref{sec:discussion}.
 
@@
 \be\label{eq:defFT}
-	\tilde{h}(f) =  \int \ud t \, e^{+2i\pi f t} h(t) \,.
+	\tilde{h}(f) = \mathrm{FT}[h](f) =  \int \ud t \, e^{+2i\pi f t} h(t) \,.
 \ee
@@
 In the following, we will drop the mode indices and the sum over modes, and we will denote generically by $h$ an individual mode of the waveform\footnote{This notation for a generic individual mode is not to be confused with the often encountered notation $h\equiv h_{+} - i h_{\times}$.}.
-Note that, if we are considering real signals $s(t)$ (such as $h_{+}$, $h_{\times}$), we can focus for instance on the positive-frequencies range and reconstruct the negative frequencies as $\tilde{s}(-f) = \tilde{s}(f)^{*}$.
-For positive frequencies $f>0$, we will use for $h$ the notation
+Note that, if we are considering real signals $s(t)$ (such as $h_{+}$, $h_{\times}$, or any detector output), we can focus  on the positive-frequencies range and reconstruct the negative frequencies as $\tilde{s}(-f) = \tilde{s}(f)^{*}$.
+For positive frequencies $f>0$, we will use the generic notation $h$ as
 \begin{align}\label{eq:notationhsignm}
@@
 The vertical line represents $\omega_{22}^{\rm peak}/2\pi$, the frequency corresponding to the time-domain frequency of the waveform at the peak amplitude.
-The starting frequency $Mf_{min} = 2.6\times 10^{-4}$ is appropriate for a 2-years observation of a $M=10^{6} M_{\odot}$ system by LISA, or for an observation by LIGO/Virgo of a $M=5.4 M_{\odot}$ system entering the instrument band at $10\Hz$.}
+The starting frequency $Mf_{min} = 2.6\times 10^{-4}$ is appropriate for a 2-years observation of a $M=10^{6} M_{\odot}$ system by LISA, or for an observation by LIGO/Virgo of a $M=5.4 M_{\odot}$ system entering the instrument band at $10\Hz$.
+[add residuals of interpolation on $~300$ points, to illustrate the compression of Amp/Phase FD waveforms ?
+or show different spins/mass ratios ?]}
   \label{fig:ampphase}
@@
 where $v=(G m \pi f/c^{3})^{1/3}$, $D$ is the luminosity distance to the observer, and where $t_{0}, \phi_{0}$ are constants.
-It is also customary to rewrite the above relations in terms of the chirp mass $\calM \equiv m\nu^{3/5}$, which is the only mass combination arising at the leading PN order.
+It is also customary to rewrite the above relations in terms of the chirp mass $\calM \equiv m\nu^{3/5}$, which is the only mass combination arising in the phase at the leading PN order.
 
@@
 The first- and second-generation TDI observables are then built as combinations of these basic building blocks, evaluated at delayed times.
-Since these delays will take a simple form in Fourier domain in our approximation, we will focus on the $y_{slr}$ observables.
+Since in our rigid approximation these delays will take a simple form in Fourier domain, we will focus in this paper on the $y_{slr}$ observables.
 We will give in Sec.~\ref{subsec:modelLISA} more details on the response of LISA-type detectors, on the TDI observables, and on the approximations that enter the derivation of the basic response~\eqref{eq:yslr} above.
@@
 The vertical line shows the instantaneous frequency at the peak, and the fact that the curve $t_{f}$ does not pass exactly by the crossing of the two lines reflects the fact that the link between time-domain frequency and Fourier-domain f is not accurate anymore in the merger part of the signal.
-Note that $t_{f}$ does not go beyond $20M$ after the peak, and is not monotonous.}
+Note that $t_{f}$ does not go beyond $20M$ after the peak, and is not monotonous.
+[show different spins/mass ratios ?]}
   \label{fig:tf}
@@
 The right panel shows the merger and ringdown part of the signal in more details, the time-domain frequency at merger being represented by the vertical line.
-The derivatives from which these timescales are built encounter zero-crossings in the high-frequency range.}
+The derivatives from which these timescales are built encounter zero-crossings in the high-frequency range.
+[if using PhenomD, discuss discontinuity of third derivative]}
   \label{fig:TfTA}
@@
 The second equality holds provided that the complex coefficients $a_{N,k}$ are a solution of the $N+1$-dimensional linear system
-\be\label{eq:combsystem}
+\be\label{eq:stencilsystem}
 	(-i\epsilon)^{p} (2p-1)!! = \sum\limits_{k=0}^{N} a_{N,k} k^{2p} \quad \text{for } p=0,\dots,N \,.
 \ee
-The two solutions for the comb coefficients in the two cases $\epsilon = \pm 1$ are simply related by a complex conjugation.
+The two solutions for the stencil coefficients in the two cases $\epsilon = \pm 1$ are simply related by a complex conjugation.
 
@@
 Note that the choice of the samples $\tf \pm k\Tf$ to approximate the successive derivatives of the expansion by finite differences is by no means unique.
+However, we remark that its simple structure leads to an anlytical expression for the stencil coefficients $a_{N,k}$, which will be given in~\ref{app:vandermonde}.
 In principle, one could have made any ansatz of samples for the series expansion to be matched on.
 One cannot expect all choices to work equally well when applied to noisy data, and the problems of the numerical size of the coefficients and of the overall numerical stability of the scheme must be investigated numerically.
-
-However, we remark that the transformation~\eqref{eq:resumquadphase} between the two series above has a special structure, and the resulting linear system can be written as a Vandermonde system.
-Defining the Vandermonde matrix $V(x_{0},\dots,x_{N})$ as $V_{ij} = (x_{j})^{i}$ (with the convention $0^{0} = 1$), setting $x_{j} = j^{2}$, and $b_{p} \equiv (-i\epsilon)^{p}(2p-1)!!$, the linear system~\eqref{eq:combsystem} becomes
-\be
-	b_{p} = \sum\limits_{k=0}^{N} V_{p,k} a_{N,k} \,.
-\ee
-Now, since the Vandermonde system solves the Lagrange polynomial interpolation problem, we know that $({}^{t}V^{-1})_{ij}$ is the coefficient of $X^{i}$ in the polynomial $\prod_{k\neq j} (X-x_{k})/(x_{j} - x_{k})$.
-If we introduce the symmetric polynomials $\sigma$ such that
-\be	
-	\sigma(m, \{x_{1}, \dots, x_{n}\}) = \sum\limits_{1\leq i_{1}<\dots<i_{m}\leq n} x_{i_{1}}\dots x_{i_{m}} \,,
-\ee
-we obtain for the inverse of the matrix $V$
-\be
-	V^{-1}_{ij} = \frac{1}{\prod_{k\neq i} (i^{2} - k^{2})} (-1)^{N-j} \sigma(N-j, \{0,\dots,N^{2}\}\backslash \{i^{2}\}) \,.
-\ee
-This allows us to write a closed-form expression for the $a_{N,k}$, for every finite order $N$:
-\be
-	a_{N,k} = \sum\limits_{p=0}^{N} (-i\epsilon)^{p}(2p-1)!!
-\frac{(-1)^{N-p}}{\prod_{j\neq k} (j^{2}-k^{2})} \sum\limits_{\substack{ 0 \leq i_{1} < \dots < i_{N-p} \leq N \\ i_{1}, \dots, i_{N-p} \neq k}} i_{1}^{2}\dots i_{N-p}^{2}
-\ee
-The combinatorial explosion of the number of terms in the symmetric polynomials, however, makes this expression impractical to evaluate for $N$ beyond a few dozen.
 We leave a more complete investigation of the asymptotics and numerical stability of this resummation approach for future work.
@@
 The upper left panel corresponds to the orbital delay part of the response~\eqref{eq:dfresponseO}, and the upper right panel corresponds to the LISA-centered response~\eqref{eq:dfresponsec}.
-The central line and interval are the mean and $1\sigma$ standard deviation of the logarithm of $\epsilon$ over 400 random values for the position in the sky, inclination and polarization.}
+The central line and interval are the mean and $1\sigma$ standard deviation of the logarithm of $\epsilon$ over 400 random values for the position in the sky, inclination and polarization.
+[change middle mass to $3\times 10^{4}$ - clarify mismatch between these $\epsilon$ and the actual size of corrections]}
   \label{fig:fomLISA}
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2y12Lreal.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2y12Limag.pdf}
+  \caption{Normalized error in the real and imaginary part of the transfer function for the $y_{132}$ basic observable, for $M=10^{2} \Msol$.}
+  \label{fig:prectoymodel}
+\end{figure}
+
 Fig.~\ref{fig:tdifdA} shows the amplitude and phase errors in the Fourier-domain local response when compared to the FFT of the time-domain response~\eqref{eq:yslr}.
@@
 \section{Application to waveforms from precessing binaries}
-\label{sec:Precession}
+\label{sec:precession}
 
@@
 The vertical line indicates the time of merger.
-Dashed black lines indicate the enforced asymptotic behaviour: frame rotation with angular velocity $\omega_{220} -\omega_{210}$ in Case II, and asymptotically constant $\beta$.}
+In the left panel the dashed black lines indicate the enforced asymptotic behaviour in Case II, where the frame rotates with constant angular velocity $\omega_{220} -\omega_{210}$.
+In the right panel, $\beta$ is asymptotically constant.}
   \label{fig:prectoymodel}
@@
 \ee
-where $\mathrm{FT}$ stands sfor the Fourier transform.
+where $\mathrm{FT}$ stands for the Fourier transform.
 Fig.~\ref{fig:prectransfer} gives the amplitude and phase of the transfer functions $\calT_{22, 2 m'}$ for the two cases I and II.
@@
 The left panels show the amplitude and the right panels the phase difference with the FFT.
-Note that the vertical scales are not uniform.}
+Note that the vertical scales are not uniform.
+[is it ok to show only 21 and 22 ?
+include the other modes which show problems ?]}
   \label{fig:precerrorcaseI}
@@
 Conventions are the same as in Fig.~\ref{fig:precerrorcaseI} but for Case II, with a more rapid post-merger frame precession.
-Note that the vertical scales are not uniform.}
+Note that the vertical scales are not uniform.
+[is it ok to show only 21 and 22 ?
+include the other modes which show problems ?]}
   \label{fig:precerrorcaseII}
@@
   \includegraphics[width=.48\linewidth]{plots/precmmasympt.pdf}
-  \caption{Mismatches computed between the exact FFT and the various approximations respresented in Figs.~\ref{fig:precerrorcaseI} and~\ref{fig:precerrorcaseII}.}
+  \caption{Mismatches computed between the exact FFT and the various approximations respresented in Figs.~\ref{fig:precerrorcaseI} and~\ref{fig:precerrorcaseII}.
+[log scale for masses ?
+discuss the fact that in case II, leading order is accidentally better than Psi 2]}
   \label{fig:precmm}
@@
 \section{Summary and Conclusions}
-\label{sec:conc}
+\label{sec:discussion}
 
@@
 
-\section{Appendix}
-\label{app:}
+\section{Explicit expression for the stencil coefficients}
+\label{app:vandermonde}
+
+In this Appendix, we exploit the special structure of the transformation~\eqref{eq:resumquadphase} between the two series to rewrite the resulting linear system as a Vandermonde system.
+Defining the Vandermonde matrix $V(x_{0},\dots,x_{N})$ as $V_{ij} = (x_{j})^{i}$ (with the convention $0^{0} = 1$), setting $x_{j} = j^{2}$, and $b_{p} \equiv (-i\epsilon)^{p}(2p-1)!!$, the linear system~\eqref{eq:stencilsystem} becomes
+\be
+	b_{p} = \sum\limits_{k=0}^{N} V_{p,k} a_{N,k} \,.
+\ee
+Now, since the Vandermonde system can be seen as a reformulation of the polynomial interpolation problem, we know that $({}^{t}V^{-1})_{ij}$ is the coefficient of $X^{i}$ in the Lagrange interpolation polynomial $\prod_{k\neq j} (X-x_{k})/(x_{j} - x_{k})$.
+If we introduce the symmetric polynomials $\sigma$ such that
+\be	
+	\sigma(m, \{x_{1}, \dots, x_{n}\}) = \sum\limits_{1\leq i_{1}<\dots<i_{m}\leq n} x_{i_{1}}\dots x_{i_{m}} \,,
+\ee
+we obtain for the inverse of the matrix $V$
+\be
+	V^{-1}_{ij} = \frac{1}{\prod_{k\neq i} (i^{2} - k^{2})} (-1)^{N-j} \sigma(N-j, \{0,\dots,N^{2}\}\backslash \{i^{2}\}) \,.
+\ee
+This allows us to write a closed-form expression for the $a_{N,k}$, for every finite order $N$:
+\be
+	a_{N,k} = \sum\limits_{p=0}^{N} (-i\epsilon)^{p}(2p-1)!!
+\frac{(-1)^{N-p}}{\prod_{j\neq k} (j^{2}-k^{2})} \sum\limits_{\substack{ 0 \leq i_{1} < \dots < i_{N-p} \leq N \\ i_{1}, \dots, i_{N-p} \neq k}} i_{1}^{2}\dots i_{N-p}^{2}
+\ee
+The combinatorial explosion of the number of terms in the symmetric polynomials, however, makes this expression rather impractical to evaluate for $N$ beyond a few dozen.
+In practice, one can simply solve the original linear system~\eqref{eq:stencilsystem} for each order one is interested in, and an actual analysis is likely to use only stencils with a limited size ($\leq 10$).
 

### 2017-01-26 docs/paper_method.tex
Update Makefile for Datura, and small changes to the paper.

@@
 With the recent detections of gravitational-wave signals emitted by the coalescences of binary black holes, announced by the LSC collaboration after the first obervation run of the two LIGO instruments~\cite{LIGO-theevent-2016,LIGO-christmasevent-2016,LIGO-O1BBH-2016}, gravitational-wave astronomy has entered its observational era.
-With the recent start of LIGO's second observation run, and later the extension of the ground-based detectors network with Virgo~\cite{Virgo}, KAGRA~\cite{KAGRA} and LIGO-India~\cite{INDIGO}, many more detections from compact object coalescences are to come, with binary black holes being expected to be the dominant sources.
-Moreover, the European Space Agency has issued a mission call for its L3 mission, adressing the science theme ``The Gravitational Universe'' with a tentative launch in 2034, possibly opening the way for a space-borne observatory based on the concept of the Laser Interferometer Space Antenna (LISA)~\cite{LISA09}.
+With the recent start of LIGO's second observation run, and later the extension of the ground-based detectors network with Virgo~\cite{Virgo}, KAGRA~\cite{KAGRA} and LIGO-India~\cite{INDIGO}, many more detections from compact object coalescences are to come, with binary black holes being expected to be the dominant sources~\cite{}.
+Moreover, the European Space Agency has issued a mission call for its L3 mission, adressing the science theme ``The Gravitational Universe'' with a tentative launch around 2034, possibly opening the way for a space-borne observatory based on the concept of the Laser Interferometer Space Antenna (LISA)~\cite{LISA09}.
 Among other sources of gravitational waves, LISA will be able to detect and characterize binary black hole coalescences to cosmological distances over a wide range of masses, from high-redshift observations of supermassive black hole binaries with $M\sim 10^{7} \Msol$ down to the observation of LIGO-type sources with $M\sim 10^{2} \Msol$~\cite{Sesana16}.
 
-The Bayesian analysis and parameter recovery of gravitational-wave signals may require millions of evaluations of waveform templates to sample the posterior probability distribution, which poses a computational challenge for templates including consistently the inspiralling phase of the evolution, the merger and the ringdown (IMR).
-State-of-the-art IMR templates incorporate information from both the perturbative results of post-Newtonian theory covering the inspiral (see e.g.~\cite{BlanchetLiving}) and from numerical relativity simulations covering the end of the inspiral and the merger-ringdown phase~\cite{}.
-The two main approaches currently used to produce complete waveform models are (i) the Effective-One-Body (EOB) approach~\cite{BD99}, which incorporates a resummation of post-Newtonian results and a calibration to numerical relativity, and which has been extended to include both spins of the two bodies~\cite{Taracchini+13, Pan+13, Bohe+16} (ii) phenomenological templates, which extend post-Newtonian signals by postulating a Fourier-domain analytic expression for the amplitude and phase and by fitting coefficients against numerical relativity waveforms~\cite{Husa+15, Khan+15}, and which include a simplified model for precession~\cite{Hannam+13}.
+Data analysis for gravitational-wave observations of compact binary coalescences require accurate models (or templates) for the signals, both for ensuring efficient detections of signals that may be buried in instrumental noise, and to extract the physical parameters of the source in subsequent analysis.
+This has prompted an important and ongoing effort to improve gravitational wave templates.
+In parallel of the improvement of the accuracy of the templates, an crucial element of the analysis is the computational efficiency of template generation, as Bayesian analysis for parameter estimation of gravitational-wave signals may require millions of evaluations of the likelihood function to sample the posterior probability distribution.
+Two important axis of improvement of the templates are the coverage of all phases of the evolution of the system, with the inspiral, merger and ringdown (IMR), and the inclusion of the effects of the spins of the compact objects.
+
+State-of-the-art IMR templates need to combine information from the perturbative results of post-Newtonian theory covering the inspiral (see e.g.~\cite{BlanchetLiving}) and from numerical relativity simulations covering the end of the inspiral and the merger-ringdown phase (see e.g.~\cite{Pfeiffer12}).
+The Effective-One-Body (EOB) approach~\cite{BD99} does so by incorporating a resummation of post-Newtonian results and a calibration to numerical relativity, and has been extended to include both spins of the two bodies~\cite{Taracchini+13, Pan+13, Bohe+16}.
+Phenomenological templates, which extend post-Newtonian signals by postulating a Fourier-domain analytic expression for the amplitude and phase and by fitting coefficients against numerical relativity waveforms~\cite{Husa+15, Khan+15}, and can include a simplified model for precession~\cite{Hannam+13}.
 The second type of models produce directly the amplitude and the phase of the Fourier-domain waveform in the form of a piecewise analytic fit.
 EOB waveforms, by contrast, require to integrate ordinary differential equations in the time-domain, which can be time-consuming.
-For the non-precessing case, Reduced Order Models (ROM), also called surrogate models, have been developed to speed up considerably the waveform generation, directly in the Fourier domain, without losing accuracy~\cite{Field+13, Puerrer14, Bohe+16}.
+For the non-precessing case, Reduced Order Models (ROM), also called surrogate models, have been developed to speed up considerably the waveform generation, without losing accuracy~\cite{Field+13, Puerrer14, Bohe+16}.
+Both approaches are therefore able to produce efficiently IMR waveforms.
+An important aspect is that the resulting waveforms can be represented in the Fourier domain by an amplitude and phase for each mode, with only a few hundred samples~\cite{Puerrer14}.
 
 A complete representation of the spin effects in fast IMR templates still remains a frontier of gravitational wave signal modelling.
-In presence of misaligned spins, the system will endure precession of the orbital plane as it evolves, leading to a modulation of the signal that is more pronunced for observers that see the system as close to edge-on~\cite{ACST94}.
-Precessing EOB templates can represent all six degrees of freedom of the two spins~\cite{Pan+13, BTB16} but still require a costly time-domain integration; while precessing phenomenological templates~\cite{Hannam+13} use effective spin parameters representing only the dominant effects of the spins.
+In presence of misaligned spins, the system will endure a precession of the orbital plane as it evolves, leading to modulations of the signal~\cite{ACST94}.
+Precessing EOB templates can represent all six degrees of freedom of the two spins~\cite{Pan+13, BTB16} but still require a costly time-domain integration; while precessing phenomenological templates~\cite{Hannam+13} decrease the dimensionnality by using effective spin parameters representing only the dominant effects of the spins.
 To model the effect of precession on the emitted waveform, it has been proposed~\cite{BCV03b, BCPTV05, Schmidt+10, OShaughnessy+11, Boyle+11} to decompose precessing waveforms by performing a time-dependent rotation, following the precession of the orbital plane.
-The resulting waveform in the rotated frame is then close to a non-precessing waveform, an approximation that is used both to build the inspiral part of precessing EOB waveforms~\cite{Pan+13} and in the construction of precessing phenomenological waveforms~\cite{Hannam+13}.
-To carry over this approach to Fourier-domain waveforms, one needs to understand how to translate the time-domain modulation created by the frame rotation into a Fourier-domain modulation.
-The most straightforward method consists in using a time-to-frequency correspondence such as the one given in the Stationary Phase Approximation (SPA), a good approximation for the inspiralling phase.
-This is what is used implicitly in Ref.~\cite{Hannam+13}, with the difference that directly frequency-dependent expressions for the modulation are used, which are then extended to also cover high frequencies in the merger part of the signal.
-In Ref.~\cite{Klein+14}, a formalism is introduced to go beyond the SPA and compute more accurately the modulation in the Fourier domain; however, this formalism still relies on the SPA for the underlying signal.
-How to implement
-
-Another frontier of gravitational wave signal modelling is prospective parameter recovery for LISA observations.
-In the past, LISA data analysis studies for comparable-mass binaries used mainly simplified, inspiral-only models~\cite{}, and Fisher-matrix estimates instead of a full Bayesian analysis.
-In Ref.~\cite{Klein+15}, an improved model was used, that included the effects of both spins including the corrections beyond the SPA investigated in~\cite{Klein+14}; the limitation to inspiral-only waveforms was compensated by SNR reweighting using full IMR waveforms.
+The resulting waveform in the rotated frame is then close to a non-precessing waveform, an approximation that is used in the construction of the inspiral part of precessing EOB waveforms~\cite{Pan+13} and in the construction of precessing phenomenological waveforms~\cite{Hannam+13}.
+For Fourier-domain waveforms, one needs to understand how to translate the time-domain modulation created by the frame rotation into a Fourier-domain modulation.
+
+The Stationary Phase Approximation (SPA) (see e.g.~\cite{Thorne300, CF94}) is a common and very good approximation to compute the Fourier transform of non-precessing signals in the inspiralling phase, but applying it directly to the modulated signal is prone to pathologies.
+In Ref.~\cite{KCY13,KCY14}, a formalism (called shifted uniform asymptotics or SUA) was introduced to go beyond the SPA and compute more accurately the modulation in the Fourier domain; however, this formalism still relies on the SPA for the underlying signal, and is as such limited to inspiralling signals.
+Another intuitive approach consists in using a time-to-frequency correspondence, such as the one given by the SPA, to directly carry over the modulation to the Fourier domain.
+This corresponds, in a sense to be detailed below, to the zeroth-order approximation of the SUA, and it is used implicitly\footnote{With the difference that directly frequency-dependent expressions for the modulation are used, which are then extended to also cover high frequencies in the merger part of the signal, as we will discuss below.} for the simplified precession of the phenomenological waveforms of Ref.~\cite{Hannam+13}.
+Consistently extending these previous approaches to the merger-ringdown part of the signal is one of the objectives of this paper.
+
+Prospective parameter recovery for LISA observations of comparable-mass binary black hole coalescences is another frontier of gravitational wave signal modelling, becoming a very timely topic with the upcoming effort for the ESA L3 mission characterization in the next few years.
+In the past, many LISA data analysis studies used mainly post-Newtonian, inspiral-only models, and were limited to the Fisher-matrix formalism instead of full Bayesian analysis (among numerous works, see e.g.
+\cite{Cutler97,LISAPE09}).
+However, the Fisher matrix analysis, although considerably cheaper to perform, has known shortcomings including at high SNR~\cite{Vallisneri08,Cokelaer08,Rodriguez+13}.
+In Ref.~\cite{Klein+15}, an improved model was used, that included the effects of both spins including the corrections beyond the SPA investigated in~\cite{KCY14}; the limitation to inspiral-only waveforms was compensated by SNR reweighting using full IMR waveforms.
+
 The response of the LISA instrument, by contrast with the response of ground-based detectors, has the specificity of being time-dependent.
@@
 
-\subsection{Waveforms in the Fourier domain and stationary phase approximation}
-\label{subsec:SPA}
+\subsection{Notations and conventions}
+\label{subsec:notations}
 
@@
 
-We recall below an approximation widely used to model gravitational waves signals emitted by inspiralling binaries, the stationary phase approximation (thereafter SPA).
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Stationary phase approximation}
+\label{subsec:SPA}
+
+We recall here an approximation widely used to model gravitational waves signals emitted by inspiralling binaries, the stationary phase approximation (thereafter SPA).
 It applies in general to chirping signals, and we refer the reader to~\ref{} for details.
@@
 The approximation then applies to signals verifying the conditions
-%
 \be\label{eq:conditionSPA}
@@
 \ee
-%
 Since the integral~\eqref{eq:defFT} defining the Fourier transform is rapidly oscillatory unless the term $2\pi f t$ cancels the evolution of $-2\varphi(t)$, its support is well centered around the point of stationary phase.
 This determines the time-to-frequency correspondence in the stationary phase approximation, and leads to the definition of this time as an implicit function of frequency by the relation
-%
 \be\label{eq:deftfSPA}
@@
 \ee
-% 
 For a chirping signal of increasing phase, $\omega>0$ and $\dot{\omega}>0$, there is a unique point of stationary phase, located in the positive frequency range $f>0$.
@@
 For this Newtonian inspiral, applying the SPA gives for the $22$ mode $\tilde{h}(f) = A_{\rm N}(f)e^{-i\Psi_{\rm N}(f)}$ with
-%
 \begin{subequations}
@@
 \end{subequations}
-%
 where $v=(G m \pi f/c^{3})^{1/3}$, $D$ is the luminosity distance to the observer, and where $t_{0}, \phi_{0}$ are constants.
@@
 Defining
-%
 \be
@@
 \ee
-%
-we have
-%
+our purpose is to derive the Fourier transform $\tilde{s}(f)$, expressed by means of a Fourier-domain transfer function $\calT$ such that
+\be\label{eq:deftransfer}
+	\tilde{s}(f) \equiv \calT(f) \tilde{h}(f) \,. 
+\ee 
+We can first write
 \be
@@
 \ee
-%
 and
-%
 \begin{align}
@@
 \end{align}
-%
-We can rewrite the last equation as a kind of convolution with a frequency-dependent Kernel, according to
-%
+The last equation can then be rewritten as a kind of convolution with a frequency-dependent Kernel, according to
 \be\label{eq:FDkernel}
-	\tilde{s}(f) = \int \ud f' \, \tilde{h}(f-f') \tilde{g}(f-f',f') \,,
-\ee
-%
-where we introduced the frequency-dependent function of time $g(f,t)$, and its Fourier transform in the auxiliary frequency $f'$, denoted by $\tilde{g}(f,f')$, as
-%
+	\tilde{s}(f) = \int \ud f' \, \tilde{h}(f-f') \tilde{G}(f-f',f') \,,
+\ee
+where we introduced the frequency-dependent function of time $G(f,t)$, and its Fourier transform in the auxiliary frequency $f'$, denoted by $\tilde{g}(f,f')$, as
 \begin{subequations}\label{eq:defg}
 \begin{align}
-	g(f,t) &= e^{-2i\pi f d(t)} F(t) \,, \\
-	\tilde{g}(f,f') &= \int \ud t \, e^{2i\pi f' t} e^{-2i\pi f d(t)} F(t) \,.
+	G(f,t) &= e^{-2i\pi f d(t)} F(t) \,, \\
+	\tilde{G}(f,f') &= \int \ud t \, e^{2i\pi f' t} e^{-2i\pi f d(t)} F(t) \,.
 \end{align}
 \end{subequations}
-%
-In the case of precessing binaries, where the delays $d(t)$ will be absent, the frequency-dependence of $g$ will vanish and this function will simply reduce to the modulation $F(t)$.
-
-In our approach, two assumptions will be crucial to the approximation:
+In the absence of delays $d(t)$, as in the case of precessing binaries, the function $G$ loses its frequency-dependence and becomes a modulation in the form of a function of time $F(t)$, and the result~\eqref{eq:FDkernel} above reduces to the usual convolution theorem for the Fourier transform.
+
+In our approach, two assumptions will be crucial for our approximations to be valid:
 \begin{itemize}
-	\item the time scales entering the signal $h(t)$ are well separated from the ones entering the modulation and the delay, so that $\tilde{g}(f-f',f')$ has support limited to $[-f_{\rm max},f_{\rm max}]$, with $f_{\rm max}$ small compared the characteristic scale for the variation of the Fourier-domain signal;
+	\item the time scales entering the signal $h(t)$ are well separated from the ones entering the modulation and the delay, so that $\tilde{G}(f-f',f')$ has support limited to $[-f_{\rm max},f_{\rm max}]$, with $f_{\rm max}$ small compared the characteristic scale for the variation of the Fourier-domain signal;
 	\item the original signal has a smooth amplitude and phase in the Fourier-domain, slowly varying over the scale $f_{\rm max}$, so that a local treatment in $f'$ of the convolution in~\eqref{eq:FDkernel} will be possible.
@@
 \begin{align}
-	\tilde{s}(f) &= \tilde{h}(f) \int \ud f' \, \exp\left[ i f' \frac{\ud \Psi}{\ud f} \right] \tilde{g}(f,f') \nn\\
-	&= \tilde{h}(f) g\left( f, -\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) \,,
+	\tilde{s}(f) &= \tilde{h}(f) \int \ud f' \, \exp\left[ i f' \frac{\ud \Psi}{\ud f} \right] \tilde{G}(f,f') \nn\\
+	&= \tilde{h}(f) G\left( f, -\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) \,,
 \end{align}
@@
 \ee
-This definition is a generalization of the time-to-frequency correspondence at the heart of the SPA, as will be discussed in Sec.~\ref{subsec:linkSPA}, but it is worth noting that here it is defined entirely from the Fourier-domain waveform.
+This definition is a generalization of the time-to-frequency correspondence at the heart of the SPA, as will be discussed in Sec.~\ref{subsec:linkSPA}, but it is worth noting that here it is defined entirely from the Fourier-domain waveform, and independently of the SPA being valid or not for the underlying signal $\tilde{h}$.
+As such, this time-of-frequency function also extends to the merger-ringdown part of the signals.
 Fig.~\ref{fig:tf} shows the time function $\tf$ for the example of an equal-mass waveform.
@@
 Using this notation, we have simply
-%
 \be\label{eq:local}
-	\tilde{s}_{\rm local}(f) = \tilde{h}(f) g(f, \tf) \,.
-\ee
-%
+	\calT_{\rm local}(f) = G(f, \tf) \,.
+\ee
 The interpretation of this approximation is straightforward: the signal is simply multiplied by the response function evaluated at the time $\tf$.
@@
 If $h'(t) = h(t -  \Delta t)$ [but the modulation and delays are left unchanged (which is relevant for space-based detectors, not for precessing binaries)], $\Psi'(f) = \Psi(f) - 2\pi f \Delta t$, so that
-%
 \be\label{eq:localshifted}
-	\tilde{s}'_{\rm local}(f) = \tilde{h}'(f) g(f, \tf + \Delta t) \,. 
-\ee
-%
+	\tilde{s}'_{\rm local}(f) = \tilde{h}'(f) G(f, \tf + \Delta t) \,. 
+\ee
 This observation also justifies that the first derivative of the Fourier-domain phase cannot, in general, be considered a small quantity, as it can take arbitrary values for arbitrary shifts in time of the signal.
@@
 We will formally Taylor-expand $\tilde{h}(f-f')$ in $f'$ as follows: we write
-%
 \begin{subequations}
@@
 	A(f-f') &= A(f) \sum\limits_{q\geq 0} \frac{(-1)^{q}}{q!} {f'}^{q} \frac{1}{A}\frac{\ud^{q} A}{\ud f^{q}} \,, \label{eq:expandA}\\
-	\tilde{g}(f-f', f') &= \sum\limits_{r\geq 0} \frac{(-1)^{r}}{r!} {f'}^{r} \frac{\partial^{r} }{\partial f^{r}}  \tilde{g}(f,f') \label{eq:expandg}
+	\tilde{G}(f-f', f') &= \sum\limits_{r\geq 0} \frac{(-1)^{r}}{r!} {f'}^{r} \frac{\partial^{r} }{\partial f^{r}}  \tilde{G}(f,f') \label{eq:expandg}
 \end{align}
 \end{subequations}
-%
 This expansion will only be usable in practice when we will be able to keep only a few terms in each of these expansions.
 We therefore limited the expansion in $\Psi$ to a finite order $N$, to keep the formula readable.
-Notice that we expand $\tilde{g}(f-f',f')$ in $f'$ only in its first argument.
+Notice that we expand $\tilde{G}(f-f',f')$ in $f'$ only in its first argument.
 We can also interchange the $f$-derivatives of $g$ with its Fourier transform.
@@
 \be
-	\int \ud f'\, {f'}^{n} \frac{\partial^{m} }{\partial f^{m}}  \tilde{g}(f,f') e^{-2i\pi f' \tf} = \frac{1}{(-2i\pi)^{n}} \left( \frac{\partial^{m} }{\partial f^{m}} \frac{\partial^{n} }{\partial t^{n}} g \right)(f,\tf) \,,
+	\int \ud f'\, {f'}^{n} \frac{\partial^{m} }{\partial f^{m}}  \tilde{G}(f,f') e^{-2i\pi f' \tf} = \frac{1}{(-2i\pi)^{n}} \left( \frac{\partial^{m} }{\partial f^{m}} \frac{\partial^{n} }{\partial t^{n}} G \right)(f,\tf) \,,
 \ee
@@
 \begin{align}
-	\tilde{s}(f) &= \tilde{h}(f) \sum\limits_{q\geq 0} \sum\limits_{r\geq 0} \sum\limits_{k_{2}\geq 0} \dots \sum\limits_{k_{N}\geq 0} \frac{(-1)^{q}}{q!} \frac{(-1)^{r}}{r!} \prod\limits_{p=2}^{N} \frac{1}{k_{p}!}\left( \frac{(-1)^{p}(-i)}{p!} \frac{\ud^{p}\Psi}{\ud f^{p}}\right)^{k_{p}} \frac{1}{A} \frac{\ud^{q} A}{\ud f ^{q}} \frac{1}{(-2i\pi)^m} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{m} }{\partial t^{m}} g \right)(f,\tf) \,,
+	\calT(f) &= \sum\limits_{q\geq 0} \sum\limits_{r\geq 0} \sum\limits_{k_{2}\geq 0} \dots \sum\limits_{k_{N}\geq 0} \frac{(-1)^{q}}{q!} \frac{(-1)^{r}}{r!} \prod\limits_{p=2}^{N} \frac{1}{k_{p}!}\left( \frac{(-1)^{p}(-i)}{p!} \frac{\ud^{p}\Psi}{\ud f^{p}}\right)^{k_{p}} \frac{1}{A} \frac{\ud^{q} A}{\ud f ^{q}} \frac{1}{(-2i\pi)^m} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{m} }{\partial t^{m}} G \right)(f,\tf) \,,
 \end{align}
@@
 This rather cumbersome expression will not be used as such, but is easy to simplify for special cases.
-For instance, if we expand only the frequency-dependence of $g$ as in~\eqref{eq:expandg}, we obtain
+For instance, if we expand only the frequency-dependence of $G$ as in~\eqref{eq:expandg}, we obtain
 \be\label{eq:taylordelay}
-	\tilde{s}(f) = \tilde{h}(f) \sum\limits_{r\geq 0} \frac{1}{(2i\pi)^{r}r!} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{r} }{\partial t^{r}} g \right)(f,\tf)
-\ee
-which interestingly looks like a Taylor expansion of $g$, but with joint derivatives in frequency and time. Similarly, the expansion~\eqref{eq:expandA} gives
+	\calT(f) = \sum\limits_{r\geq 0} \frac{1}{(2i\pi)^{r}r!} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{r} }{\partial t^{r}} G \right)(f,\tf)
+\ee
+which interestingly looks like a Taylor expansion of $G$, but with joint derivatives in frequency and time. Similarly, the expansion~\eqref{eq:expandA} gives
 \be\label{eq:resultdfA}
-	\tilde{s}(f) = \tilde{h}(f) \sum\limits_{q\geq 0} \frac{1}{(2i\pi)^{q}q!} \frac{1}{A} \frac{\ud^{q} A}{\ud f ^{q}}  \left( \frac{\partial^{q} }{\partial t^{q}} g \right)(f,\tf)
+	\calT(f) = \sum\limits_{q\geq 0} \frac{1}{(2i\pi)^{q}q!} \frac{1}{A} \frac{\ud^{q} A}{\ud f ^{q}}  \left( \frac{\partial^{q} }{\partial t^{q}} G \right)(f,\tf)
 \ee
@@
 \be\label{eq:resultdffPsi}
-	\tilde{s}(f) = \tilde{h}(f) \sum\limits_{p\geq 0} \frac{1}{p!} \left( \frac{i}{8\pi^{2}}\frac{\ud^{2} \Psi}{\ud f^{2}} \right)^{p} \left( \frac{\partial^{2p} }{\partial t^{2p}} g \right)(f, \tf) \,.
-\ee
-Indeed, this correction will be the most significant correction beyond leading order in the applications that we will consider in Secs.~\ref{} and~\ref{}, and this particular approximation will turn out to be closely related to the approach of~\cite{Klein+14}.
+	\calT(f) = \sum\limits_{p\geq 0} \frac{1}{p!} \left( \frac{i}{8\pi^{2}}\frac{\ud^{2} \Psi}{\ud f^{2}} \right)^{p} \left( \frac{\partial^{2p} }{\partial t^{2p}} G \right)(f, \tf) \,.
+\ee
+Indeed, this correction will be the most significant correction beyond leading order in the applications that we will consider in Secs.~\ref{} and~\ref{}, and this particular approximation will turn out to be closely related to the approach of~\cite{KCY14}.
 
@@
 \begin{align}
-	\epsilon_{\Psi 2} &\equiv \frac{1}{4\pi^{2}} \left| \frac{\ud^{2} \Psi}{\ud f^{2}} \frac{1}{g} \frac{\partial^{2} g}{\partial t^{2}}(f, \tf) \right| \,, \\
-	\epsilon_{A 1} &\equiv \frac{1}{2\pi} \left| \frac{1}{A}\frac{\ud A}{\ud f} \frac{1}{g} \frac{\partial g}{\partial t}(f, \tf) \right| \,, \\
-	\epsilon_{A 2} &\equiv \frac{1}{4\pi^{2}} \left| \frac{1}{A}\frac{\ud^{2} A}{\ud f^{2}} \frac{1}{g} \frac{\partial^{2} g}{\partial t^{2}}(f, \tf) \right| \,, \\
-	\epsilon_{d} &\equiv \frac{1}{2\pi} \left| \frac{1}{g} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{r} }{\partial t^{r}} g \right)(f, \tf) \right| \,.
+	\epsilon_{\Psi 2} &\equiv \frac{1}{4\pi^{2}} \left| \frac{\ud^{2} \Psi}{\ud f^{2}} \frac{1}{G} \frac{\partial^{2} G}{\partial t^{2}}(f, \tf) \right| \,, \\
+	\epsilon_{A 1} &\equiv \frac{1}{2\pi} \left| \frac{1}{A}\frac{\ud A}{\ud f} \frac{1}{G} \frac{\partial G}{\partial t}(f, \tf) \right| \,, \\
+	\epsilon_{A 2} &\equiv \frac{1}{4\pi^{2}} \left| \frac{1}{A}\frac{\ud^{2} A}{\ud f^{2}} \frac{1}{G} \frac{\partial^{2} G}{\partial t^{2}}(f, \tf) \right| \,, \\
+	\epsilon_{d} &\equiv \frac{1}{2\pi} \left| \frac{1}{G} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{r} }{\partial t^{r}} G \right)(f, \tf) \right| \,.
 \end{align}
@@
 We found in both the LISA case and in the precession case that the third and higher derivatives of both the amplitude and phase are negligible.
-Note that the figure of merit $\epsilon_{d}$, introduced to represent the effect of the frequency dependence in the convolution kernel~\eqref{eq:FDkernel}, is signal-independent and depends only on $g$.
+Note that the figure of merit $\epsilon_{d}$, introduced to represent the effect of the frequency dependence in the convolution kernel~\eqref{eq:FDkernel}, is signal-independent and depends only on $G$.
 
@@
 \begin{align}\label{eq:resultdffPsiTf}
-	 \tilde{s}(f) &= \tilde{h}(f) \sum\limits_{p\geq 0} \frac{(-i\epsilon)^{p}}{2^{p}p!} \Tf^{2p} \left( \frac{\partial^{2p} }{\partial t^{2p}} g \right)(f, \tf) \,,
+	 \calT(f) &= \sum\limits_{p\geq 0} \frac{(-i\epsilon)^{p}}{2^{p}p!} \Tf^{2p} \left( \frac{\partial^{2p} }{\partial t^{2p}} G \right)(f, \tf) \,,
 \end{align}
@@
 
-The expansion in amplitude corrections~\eqref{eq:resultdfA} also leads to the natural introduction of a set of other timescales related to the successive derivatives of te amplitude.
+The expansion in amplitude corrections~\eqref{eq:resultdfA} also leads to the natural introduction of a set of other timescales related to the successive derivatives of the amplitude.
 In an analogous manner to the definition~\eqref{eq:defTf} of the timescale $T_{f}$, we can define
@@
 \ee
-where we included an absolute value to accomodate the possible sign changes in the right-hand side. In practice, only the first few of these timescales will be relevant.
+where we included an absolute value to accomodate the possible sign changes in the right-hand side.
+In practice, only the first few of these timescales will be relevant ($T_{A1}$ and $T_{A2}$ in the analysis of this paper).
 
@@
 
-We can now compare our result to the approach of Ref.~\cite{Klein+14}, where the authors extend the SPA in a formalism called the shifted uniform asymptotic expansion (SUA).
+We can now compare our result to the approach of Ref.~\cite{KCY14}, where the authors extend the SPA in a formalism called the shifted uniform asymptotic expansion (SUA).
 The derivation includes an expansion of the modulation in Bessel functions, a study of the displacement of the stationary phase point induced by the precession, and a resummation of the result in the form of time derivatives.
-The main result of~\cite{Klein+14}, their equation~(34), reads exactly like~\eqref{eq:resultdffPsiTf} with the identifications $\tilde{H}_{corr}(f)\rightarrow \tilde{s}(f)/\tilde{h}(f)$ for the transfer function, $T\rightarrow \Tf$ for the radiation-reaction timescale and $e^{-i\delta\phi} \rightarrow F$ for the modulation function (which is restricted in their framework to include only the phase of the modulation, its amplitude being treated jointly with the amplitude of the signal).
-Thus, we are able to make a direct correspondence between the treatment of~\cite{Klein+14} and our framework, which provides us with a simple and alternative derivation of their result.
-Namely, the result of~\cite{Klein+14} corresponds to the approximation~\eqref{eq:expandPsi}, where in the Fourier-domain convolution the phase of the signal is expanded to quadratic order and the amplitude is not expanded.
+The main result of~\cite{KCY14}, their equation~(34), reads exactly like~\eqref{eq:resultdffPsiTf} with the identifications $\tilde{H}_{corr}(f)\rightarrow \tilde{s}(f)/\tilde{h}(f)$ for the transfer function, $T\rightarrow \Tf$ for the radiation-reaction timescale and $e^{-i\delta\phi} \rightarrow F$ for the modulation function (which is restricted in their framework to include only the phase of the modulation, its amplitude being treated jointly with the amplitude of the signal).
+Thus, we are able to make a direct correspondence between the treatment of~\cite{KCY14} and our framework, which provides us with a simple and alternative derivation of their result.
+Namely, the result of~\cite{KCY14} corresponds to the approximation~\eqref{eq:expandPsi}, where in the Fourier-domain convolution the phase of the signal is expanded to quadratic order and the amplitude is not expanded.
 
@@
 In this Section, we focus on the second-derivative corrective term in~\eqref{}.
-We will restrict to the case of a pure modulation, $d=0$ and $g(f,t) = F(t)$.
-The authors of~\cite{Klein+14} proposed a resummation scheme.
+We will restrict to the case of a pure modulation, $d=0$ and $G(f,t) = F(t)$ [go back to more general $G(f,t)$ ?].
+The authors of~\cite{KCY14} proposed a resummation scheme.
 Indeed, the result~\eqref{} looks like a symmetrized Taylor expansion, except for the factors $i^{p}$ and $1/p!$ instead of $1/(2p)!$.
-After having truncated the sum at some finite $N$, one can write
-%
+After having truncated the sum at some finite $N$, one can write (following~\cite{KCY14})
 \begin{align}\label{eq:resumquadphase}
-	\tilde{s}(f) &\simeq \tilde{h}(f) \sum\limits_{p = 0}^{N} \frac{(-i\epsilon\Tf^{2})^{p}}{2^{p}p!} \partial_{t}^{2p}F(\tf) \nn\\
-	&\simeq \tilde{h}(f) \sum\limits_{p= 0}^{N} \sum\limits_{k=0}^{N} a_{N,k}\frac{(k\Tf)^{2p}}{(2p)!}  \partial_{t}^{2p}F(\tf) \nn\\
-	&\simeq \tilde{h}(f) \frac{1}{2}\sum\limits_{k=0}^{N} a_{N,k} \left( F(\tf + k\Tf) + F(\tf-k\Tf) \right)\,,
-\end{align}
-%
-where we applied a truncation approximation at the first and last stage.
+	\calT(f) &\simeq \sum\limits_{p = 0}^{N} \frac{(-i\epsilon\Tf^{2})^{p}}{2^{p}p!} \partial_{t}^{2p}F(\tf) \nn\\
+	&\simeq \sum\limits_{p= 0}^{N} \sum\limits_{k=0}^{N} a_{N,k}\frac{(k\Tf)^{2p}}{(2p)!}  \partial_{t}^{2p}F(\tf) \nn\\
+	&\simeq \frac{1}{2}\sum\limits_{k=0}^{N} a_{N,k} \left( F(\tf + k\Tf) + F(\tf-k\Tf) \right)\,,
+\end{align}
+where one applies a truncation approximation at the first and last stage.
 The second equality holds provided that the complex coefficients $a_{N,k}$ are a solution of the $N+1$-dimensional linear system
@@
 %[TO CORRECT AND REWRITE]
-%The rewriting of~\eqref{eq:resultdffPsiTf} by means of an integral transform leads by analogy to another way of formulating the connection to the results of Ref.~\cite{Klein+14}.
+%The rewriting of~\eqref{eq:resultdffPsiTf} by means of an integral transform leads by analogy to another way of formulating the connection to the results of Ref.~\cite{KCY14}.
 If we consider the inspiralling phase of a precessing binary, the SPA applies to the signal $h(t) = a(t)e^{-2i\varphi(t)}$.
@@
 %\ee
-%We see that, within the context of the SPA, the result~\eqref{} for the quadratic term in the phase, equivalent to the result of~\cite{Klein+14}, can be recovered by a time-domain approach.
+%We see that, within the context of the SPA, the result~\eqref{} for the quadratic term in the phase, equivalent to the result of~\cite{KCY14}, can be recovered by a time-domain approach.
 However, our formalism is somewhat more general, as it does not make use of the SPA for the underlying signal $\tilde{h}(f)$ and can in principle be used for the merger-ringdown part of the signal as well.
@@
 Now, the most relevant correction in~\eqref{eq:delaycorrleading} comes from the phase factor at high frequencies, where the factors $2\pi f d_{O}$ and $2\pi f d_{c}$ give a magnification by a factor of, respectively, $3.10^{3}$ and $10^{2}$ at $1\Hz$.
-Ignoring the other corrections, we thus arrive at the approximate form
-\be
-	\tilde{s}(f) \simeq \tilde{h}(f) F(t_{f})\exp\left[ -2i\pi f d(t_{f}) (1-\dot{d}(t_{f})) \right] \,.
+Ignoring the other corrections, we thus arrive at the approximate form for the transfer function
+\be
+	\calT(f) \simeq F(t_{f})\exp\left[ -2i\pi f d(t_{f}) (1-\dot{d}(t_{f})) \right] \,.
 \ee
@@
 \be
-	\tilde{s}(f) = \tilde{h}(f) \sum\limits_{k=0}^{+\infty} \frac{1}{(2i\pi)^{k} k!} \frac{1}{A} \frac{\ud^{k}A}{\ud f^{k}}  \calF^{\epsilon}_{\Tf} \left[ \frac{\ud^{k}}{\ud \tau^{k}} \left( \frac{F(\tau - d(\tau))}{1+\dot{d}(\tau)} e^{-2i\pi f d(\tau)(1-\dot{d}(\tau))} \right) \right] (\tf) \,.
+	\calT(f) = \sum\limits_{k=0}^{+\infty} \frac{1}{(2i\pi)^{k} k!} \frac{1}{A} \frac{\ud^{k}A}{\ud f^{k}}  \calF^{\epsilon}_{\Tf} \left[ \frac{\ud^{k}}{\ud \tau^{k}} \left( \frac{F(\tau - d(\tau))}{1+\dot{d}(\tau)} e^{-2i\pi f d(\tau)(1-\dot{d}(\tau))} \right) \right] (\tf) \,.
 \ee
@@
   \centering
-  \includegraphics[width=.48\linewidth]{plots/tdifdAamp.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/tdifdAphase.pdf}
-  \caption{Fourier-domain amplitude and phase timescales, as defined in~\eqref{eq:defTf} and~\eqref{eq:defTA}, in geometric units for an equal-mass, non-spinning system.
-The right panel shows the merger and ringdown part of the signal in more details, the time-domain frequency at merger being represented by the vertical line.
-The derivatives from which these timescales are built encounter zero-crossings in the high-frequency range.}
-  \label{fig:tdifdA}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2y12Lreal.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2y12Limag.pdf}
-  \caption{Normalized error in the real and imaginary part of the transfer function for the $y_{132}$ basic observable, for $M=10^{2} \Msol$.}
+  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e7dOamp.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e7dOphase.pdf}
+  %
+  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e7dOamp.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e7dOphase.pdf}
+  \caption{Error in the transfer function for the orbital delay $d_{O}$, for $M=10^{7} \Msol$.}
+  \label{fig:prectoymodel}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e7y12Lamp.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e7y12Lphase.pdf}
+  %
+  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e7y12LcorrPsi.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e7y12LcorrAd.pdf}
+  \caption{Error in the transfer function for the basic observable $y_{12}$, for $M=10^{7} \Msol$.}
+  \label{fig:prectoymodel}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e2dOamp.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e2dOphase.pdf}
+  %
+  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2dOamp.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2dOphase.pdf}
+  \caption{Transfer function and reconstruction error for the orbital delay $d_{O}$, for $M=10^{2} \Msol$.}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e2y12Lamp.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e2y12LReIm.pdf}
+  %
+  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2y12LcorrPsi.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2y12LcorrAd.pdf}
+  \caption{Transfer function and reconstruction error for the basic observable $y_{12}$, for $M=10^{2} \Msol$.}
   \label{fig:prectoymodel}

### 2017-01-26 LISAinference/LISAinference.c

@@ -165,6 +165,13 @@ void getLogLike(double *Cube, int *ndim, int *npars, double *lnew, void *context
     return;
   }
 
+  /* If zerolikelihood option is activated, simply set result to 0 for all parameters choices within prior */
+  /* Useful for testing purposes, to just sample from the prior */
+  if(globalparams->zerolikelihood) {
+    *lnew = 0;
+    return;
+  }
+
   /* Cube values here have the order of the 9 physical parameters (fixed): m1, m2, tRef, dist, phase, inc, lambda, beta, pol */
   LISAParams templateparams;
   templateparams.m1 = Cube[0];

### 2017-01-26 LISAinference/LISAutils.c

@@ -144,6 +144,7 @@ Arguments are as follows:\n\
  --tagint              Tag choosing the integrator: 0 for Fresnel (default), 1 for linear integration\n\
  --tagtdi              Tag choosing the set of TDI variables to use (default TDIAETXYZ)\n\
  --nbptsoverlap        Number of points to use for linear integration (default 32768)\n\
+ --zerolikelihood      Zero out the likelihood to sample from the prior for testing purposes (default 0)\n\
 \n\
 --------------------------------------------------\n\
 ----- Prior Boundary Settings --------------------\n\
@@ -223,6 +224,7 @@ Syntax: --PARAM-min\n\
     globalparams->tagint = 0;
     globalparams->tagtdi = TDIAETXYZ;
     globalparams->nbptsoverlap = 32768;
+    globalparams->zerolikelihood = 0;
 
     /* set default values for the prior limits */
     prior->deltaT = 3600.;
@@ -326,6 +328,8 @@ Syntax: --PARAM-min\n\
             globalparams->tagtdi = ParseTDItag(argv[++i]);
         } else if (strcmp(argv[i], "--nbptsoverlap") == 0) {
             globalparams->nbptsoverlap = atoi(argv[++i]);
+        } else if (strcmp(argv[i], "--zerolikelihood") == 0) {
+            globalparams->zerolikelihood = 1;
         } else if (strcmp(argv[i], "--deltaT") == 0) {
             prior->deltaT = atof(argv[++i]);
         } else if (strcmp(argv[i], "--comp-min") == 0) {
@@ -512,6 +516,7 @@ int print_parameters_to_file_LISA(
   fprintf(f, "tagint:       %d\n", globalparams->tagint);
   fprintf(f, "tagtdi:       %d\n", globalparams->tagtdi); //Translation back from enum to string not implemented yet
   fprintf(f, "nbptsoverlap: %d\n", globalparams->nbptsoverlap);
+  fprintf(f, "zerolikelihood: %d\n", globalparams->zerolikelihood);
   fprintf(f, "-----------------------------------------------\n");
   fprintf(f, "\n");
 
@@ -566,17 +571,17 @@ int print_parameters_to_file_LISA(
   fprintf(f, "-----------------------------------------------\n");
   fprintf(f, "Run parameters:\n");
[…]

### 2017-01-26 LISAinference/LISAutils.h

@@ -55,6 +55,7 @@ typedef struct tagLISAGlobalParams {
   int tagint;                /* Tag choosing the integrator: 0 for wip (default), 1 for linear integration */
   int tagtdi;                /* Tag choosing the TDI variables to use */
   int nbptsoverlap;          /* Number of points to use in loglinear overlaps (default 32768) */
+  int zerolikelihood;        /* Tag to zero out the likelihood, to sample from the prior for testing purposes (default 0) */
 } LISAGlobalParams;
 
 typedef struct tagLISASignalCAmpPhase

### 2017-01-26 LISAsim/LISAgeometry.h

@@ -49,7 +49,7 @@ extern "C" {
 typedef enum TDItag {
   delayO,                  /* Orbital delay */
   y12L,                    /* Constellation-only response y12L */
-  y12,                     /* Complete response y12 (includees orbital delay) */
+  y12,                     /* Complete response y12 (includes orbital delay) */
   TDIXYZ,
   TDIalphabetagamma,
   TDIAETXYZ,

### 2017-01-26 python/flare.py

@@ -18,7 +18,8 @@ ireport=9
 FisherRunFailCount=0
 noRun=False
 all_params_file=False
-ROM_DATA_PATH="/discover/nobackup/jgbaker/GW-DA/flare/ROMdata/q1-12_Mfmin_0.0003940393857519091"
+#ROM_DATA_PATH="/discover/nobackup/jgbaker/GW-DA/flare/ROMdata/q1-12_Mfmin_0.0003940393857519091"
+ROM_DATA_PATH="/Users/marsat/src/flare/ROMdata/q1-12_Mfmin_0.0003940393857519091"
 
 def set_flare_flags(snr,params):
     flags=""
@@ -27,9 +28,10 @@ def set_flare_flags(snr,params):
     #flags+=" --tagint 0" #1 for Fresnel integration(default), vs gridded quadrature"
     #flags+=" --tagint 1" --nbptsoverlap 8192" #gridded quadrature
     flags+=" --deltatobs 5.0" #duration in years of LISA observation
+    #flags+=" --deltatobs 1.0" #duration in years of LISA observation
     #flags+=" --minf 1e-4" #minimun frequency included in analysis
     flags+=" --minf 3e-6" #minimun frequency included in analysis
-    #flags+=" --nbmodeinj 1 --nbmodetemp 1" #for no higher modes in injection and template
+    flags+=" --nbmodeinj 1 --nbmodetemp 1" #for no higher modes in injection and template
     if(snr>0):
         flags+=" --snr "+str(snr)+" --rescale-distprior" #fixing SNR (rescales distance)
     flags+=" --comp-min 1e5 --comp-max 1e8" #min/max for component mass prior ranges
@@ -75,7 +77,7 @@ def set_mcmc_flags(outroot,ptN):
     if(ptN>0):
         flags += " --pt_n="+str(ptN) #else default is 20
     flags += " --pt_swap_rate=0.10"   #rate of temp swaps (or default 0.01)
-    flags += " --pt_evolve_rate=0.01" #rate at which temps are allowed to evolve 
+    flags += " --pt_evolve_rate=0.01" #rate at which temps are allowed to evolve
 
     flags += " --pt_reboot_rate=0.0001 --pt_reboot_every=10000 --pt_reboot_grace=50000" #Somewhat hacky trick to avoid chains getting stuck.  Not sure whether we need this.
     #stopping criteria
[…]

### 2017-03-28 docs/paper_method.tex
Edits to the paper.
Also added more TDItags allowed for SNR computation.

@@
 
-\title{Modulations of gravitational-wave signals: a Fourier-domain approach for the response of LISA-type detectors and for precessing binaries}
+\title{Fourier-domain modulations and delays of gravitational-wave signals}
 
@@
 
-We present a Fourier-domain based approach to the problem of the modulation of gravitational wave signals.
-The two main contexts where modulations are important for gravitational-wave astronomy: (i) the time-dependent response of space-based instruments such as LISA; and (ii) the transposition in Fourier domain of the frame decomposition of precessing waveforms, where the precessing waveform is approximated by rotating a non-precessing waveform to follow the precession of the orbital plane.
-Our results recover and generalize previous approaches based on an extension of the stationary phase approximation, and can be applied to full inspiral-merger-ringdown signals.
-In the case of the response LISA-type detectors, we show that higher-order corrections remain small except at the two extremes of the detector's frequency band, and we show that including them reduce the errors.
+We present a Fourier-domain based approach modulations and delays of gravitational wave signals, a problem which arises in two different contexts.
+For space-based detectors like LISA, the orbital motion of the detector introduces a time-dependency in the response of the detector, consisting of both a modulation and a varying delay.
+In the context of signals from precessing spinning binary systems, a useful tool for building models of the waveform consists in representing the signal as a time-dependent rotation of a quasi-non-precessing waveform.
+In both cases, being able to compute transfer functions for these effects directly in the Fourier domain would allow to gain performance for data analysis applications by using fast frequency-domain waveforms.
+Our results recover previous approaches based on an extension of the stationary phase approximation, generalize them by including delays and higher-order corrections, and can in principle be applied to full inspiral-merger-ringdown signals.
+In the LISA case, we find that higher-order corrections remain small except at the two extremes of the detector's frequency band, where we show that including them reduces the errors.
 In the case of precessing binaries, we perform a limited exploration of the merger-ringdown range based on a toy model for the precession.
-We find that, depending on the post-merger frame velocity, higher-order corrections can become quantitatively large and even challenge our perturbative formalism.
-We develop a direct convolution approach to capture these post-merger features.
-However, in that case these corrections appear mainly as amplitude corrections and give raise to a small unfaithfulness.
+We find that, depending on the post-merger frame velocity, higher-order corrections can become quantitatively large to the point of challenging our perturbative formalism, while affecting mainly the amplitude and giving raise to a limited unfaithfulness.
+We develop an alternative direct convolution approach to represent accurately these post-merger features.
 
@@
 
-With the recent detections of gravitational-wave signals emitted by the coalescences of binary black holes, announced by the LSC collaboration after the first obervation run of the two LIGO instruments~\cite{LIGO-theevent-2016,LIGO-christmasevent-2016,LIGO-O1BBH-2016}, gravitational-wave astronomy has entered its observational era.
+With the recent detections of gravitational-wave signals emitted by the coalescences of binary black holes, announced by the LVC collaboration after the first obervation run of the two LIGO instruments~\cite{LIGO-theevent-2016,LIGO-christmasevent-2016,LIGO-O1BBH-2016}, gravitational-wave astronomy has entered its observational era.
 With the recent start of LIGO's second observation run, and later the extension of the ground-based detectors network with Virgo~\cite{Virgo}, KAGRA~\cite{KAGRA} and LIGO-India~\cite{INDIGO}, many more detections from compact object coalescences are to come, with binary black holes being expected to be the dominant sources~\cite{}.
-Moreover, the European Space Agency has issued a mission call for its L3 mission, adressing the science theme ``The Gravitational Universe'' with a tentative launch around 2034, possibly opening the way for a space-borne observatory based on the concept of the Laser Interferometer Space Antenna (LISA)~\cite{LISA09}.
-Among other sources of gravitational waves, LISA will be able to detect and characterize binary black hole coalescences to cosmological distances over a wide range of masses, from high-redshift observations of supermassive black hole binaries with $M\sim 10^{7} \Msol$ down to the observation of LIGO-type sources with $M\sim 10^{2} \Msol$~\cite{Sesana16}.
-
-Data analysis for gravitational-wave observations of compact binary coalescences require accurate models (or templates) for the signals, both for ensuring efficient detections of signals that may be buried in instrumental noise, and to extract the physical parameters of the source in subsequent analysis.
+Moreover, the European Space Agency has issued a mission call for its L3 space mission with a tentative launch around 2034, adressing the science theme ``The Gravitational Universe'', possibly opening the way for a space-borne observatory based on the concept of the Laser Interferometer Space Antenna (LISA)~\cite{LISA09,LISA17}.
+Among other sources of gravitational waves, LISA will be able to detect and characterize comparable-mass binary black hole coalescences to cosmological distances over a wide range of masses, from high-redshift observations of supermassive black hole binaries with $M\sim 10^{7} \Msol$ down to the observation of LIGO-type sources with $M\sim 10^{2} \Msol$~\cite{Sesana16}.
+
+Data analysis for gravitational-wave observations of compact binary coalescences require accurate models (or templates) for the signals, both for ensuring efficient detections of signals that may be buried in instrumental noise, and to extract the physical parameters of the source in a subsequent analysis.
 This has prompted an important and ongoing effort to improve gravitational wave templates.
-In parallel of the improvement of the accuracy of the templates, an crucial element of the analysis is the computational efficiency of template generation, as Bayesian analysis for parameter estimation of gravitational-wave signals may require millions of evaluations of the likelihood function to sample the posterior probability distribution.
-Two important axis of improvement of the templates are the coverage of all phases of the evolution of the system, with the inspiral, merger and ringdown (IMR), and the inclusion of the effects of the spins of the compact objects.
+Two important qualitative features for the improvement of the templates are the coverage of all phases of the evolution of the system, including the inspiral, merger and ringdown (IMR), which became possible thanks to the breakthrough of numerical relativity~\cite{Pretorius05,Baker+06,Campanelli+06}, and the inclusion of the effects of the spins of the compact objects, that can lead to a precession of the orbital plane~\cite{ACST94,K95}.
+Besides accuracy, the computational efficiency of template generation is crucial, as Bayesian analysis for parameter estimation of gravitational-wave signals, as was performed for the LIGO detections~\cite{LIGO-theeventPE-2016,LIGO-O1BBH-2016}, may require millions of evaluations of the likelihood function to sample the posterior probability distribution.
 
@@
 The Effective-One-Body (EOB) approach~\cite{BD99} does so by incorporating a resummation of post-Newtonian results and a calibration to numerical relativity, and has been extended to include both spins of the two bodies~\cite{Taracchini+13, Pan+13, Bohe+16}.
-Phenomenological templates, which extend post-Newtonian signals by postulating a Fourier-domain analytic expression for the amplitude and phase and by fitting coefficients against numerical relativity waveforms~\cite{Husa+15, Khan+15}, and can include a simplified model for precession~\cite{Hannam+13}.
-The second type of models produce directly the amplitude and the phase of the Fourier-domain waveform in the form of a piecewise analytic fit.
+Phenomenological templates extend post-Newtonian models by postulating an analytic ansatz for the Fourier-domain amplitude and phase, fitting coefficients against numerical relativity waveforms~\cite{Husa+15, Khan+15}, and can include a simplified model for precession~\cite{Hannam+13}.
+The second type of models produce directly the amplitude and the phase of the Fourier-domain waveform in the form of a piecewise analytic function.
 EOB waveforms, by contrast, require to integrate ordinary differential equations in the time-domain, which can be time-consuming.
 For the non-precessing case, Reduced Order Models (ROM), also called surrogate models, have been developed to speed up considerably the waveform generation, without losing accuracy~\cite{Field+13, Puerrer14, Bohe+16}.
-Both approaches are therefore able to produce efficiently IMR waveforms.
-An important aspect is that the resulting waveforms can be represented in the Fourier domain by an amplitude and phase for each mode, with only a few hundred samples~\cite{Puerrer14}.
+Both approaches are therefore able to produce efficiently IMR Fourier-domain waveforms.
+An important aspect is that the resulting waveforms can be represented by an amplitude and phase for each mode, with only a few hundred samples~\cite{Puerrer14}.
 
 A complete representation of the spin effects in fast IMR templates still remains a frontier of gravitational wave signal modelling.
-In presence of misaligned spins, the system will endure a precession of the orbital plane as it evolves, leading to modulations of the signal~\cite{ACST94}.
-Precessing EOB templates can represent all six degrees of freedom of the two spins~\cite{Pan+13, BTB16} but still require a costly time-domain integration; while precessing phenomenological templates~\cite{Hannam+13} decrease the dimensionnality by using effective spin parameters representing only the dominant effects of the spins.
+In presence of misaligned spins, the system will endure a precession of the orbital plane as it evolves, leading to modulations of the signal as seen by the observer~\cite{ACST94}.
+Precessing EOB templates incorporate all six parameters of the two spins~\cite{Pan+13, BTB16} but still require a costly time-domain integration; while precessing phenomenological templates~\cite{Hannam+13} decrease the dimensionnality by using effective spin parameters representing only the dominant effects of the spins.
 To model the effect of precession on the emitted waveform, it has been proposed~\cite{BCV03b, BCPTV05, Schmidt+10, OShaughnessy+11, Boyle+11} to decompose precessing waveforms by performing a time-dependent rotation, following the precession of the orbital plane.
-The resulting waveform in the rotated frame is then close to a non-precessing waveform, an approximation that is used in the construction of the inspiral part of precessing EOB waveforms~\cite{Pan+13} and in the construction of precessing phenomenological waveforms~\cite{Hannam+13}.
-For Fourier-domain waveforms, one needs to understand how to translate the time-domain modulation created by the frame rotation into a Fourier-domain modulation.
-
-The Stationary Phase Approximation (SPA) (see e.g.~\cite{Thorne300, CF94}) is a common and very good approximation to compute the Fourier transform of non-precessing signals in the inspiralling phase, but applying it directly to the modulated signal is prone to pathologies.
-In Ref.~\cite{KCY13,KCY14}, a formalism (called shifted uniform asymptotics or SUA) was introduced to go beyond the SPA and compute more accurately the modulation in the Fourier domain; however, this formalism still relies on the SPA for the underlying signal, and is as such limited to inspiralling signals.
+The resulting waveform in the rotated frame can then be modelled by a non-precessing waveform, an approximation which is used both in the construction of the inspiral part of precessing EOB waveforms~\cite{Pan+13} and in the construction of precessing phenomenological waveforms~\cite{Hannam+13}.
+To follow this modelling approach and create efficiently Fourier-domain waveforms, one needs to understand how to translate the time-domain modulation created by the frame rotation into a Fourier-domain transfer function, which is our first motivation.
+
+Prospective parameter recovery studies for LISA observations of comparable-mass binary black hole coalescences aim at characterizing the scientific performance of the future experiment.
+They become a timely topic with the upcoming effort for the ESA L3 mission characterization in the next few years.
+However, the recent progresses in waveform modelling, most notably the availability of cheap IMR waveforms, have not yet made their way to the field of LISA data analysis studies.
+In the past, such studies used mainly post-Newtonian, inspiral-only models, and were limited to the Fisher-matrix formalism instead of full Bayesian analysis (among numerous works, see e.g.
+\cite{Cutler97,LISAPE09}).
+Refs.~\cite{McWilliams+09,McWilliams+10,McWilliams+11} explored the effects of including the merger and ringdown, keeping the mass parameters fixed.
+However, the Fisher matrix analysis, although considerably cheaper to perform, is valid only in the limit of a high signal to noise ratio (SNR) and has known shortcomings~\cite{Vallisneri08}.
+Several implementations of Bayesian analyses were developed~\cite{CP06,Wickham+06,Babak+08,GP09,Feroz+09,Petiteau+10,PC13} (see also~\cite{MLDC4} and references therein), with Ref.~\cite{PC15} proposing a direct comparison of Fisher and Bayes analyses, but they were limited to inspiral-only waveforms.
+One exception was Ref.~\cite{Babak+08} where IMR signals were used in a Bayesian study, but there the masses of the black holes were kept fixed.
+In the recent Ref.~\cite{Klein+15}, the limitation to inspiral-only precessing post-Newtonian waveforms was compensated by an SNR reweighting procedure.
+Thus, due mainly to their high computational cost, full-fledged Bayesian parameter estimation simulations for IMR signals are still missing.
+
+By contrast with ground-based detectors, the response of a LISA-type instrument has the specificity of being time-dependent~\cite{Cutler97}.
+For signals possibly accumulating over several months, the motion and change of orientation of the detector constellation along its orbit lead to significant time variability in the form of a modulation and a varying delay, which contain information about the localization of the gravitational-wave source in the sky.
+This complicated detector response increases the computational cost of parameter-estimation analyses.
+To leverage the performance of state-of-the-art Fourier-domain IMR templates~\cite{BTB16,Khan+15}, one needs to be able to process the signals through the time-dependent response of the dectector while staying in the Fourier domain, which is our second motivation.
+
+The purpose of this paper is to propose a formalism allowing to transpose directly in the Fourier domain the processing of a signal through a time-domain modulation and delay, while retaining the compactness of a Fourier-domain amplitude and phase representation of the signals.
+This will allow us to address both the issue of the Fourier-domain response of the LISA instrument, as well as the issue of Fourier-domain precession modulation for IMR signals from precessing binaries.
+
+In previous works, the Stationary Phase Approximation (SPA) (see e.g.~\cite{Thorne300, CF94}) has been often used for this purpose.
+The SPA is a common approximation to compute the Fourier transform of non-precessing signals during the inspiralling phase.
+In the case of precessing binaries, applying it directly to the modulated signal is however prone to pathologies.
+In Ref.~\cite{KCY13,KCY14}, a formalism (called shifted uniform asymptotics or SUA) was introduced to go beyond the SPA and compute more accurately the modulation in the Fourier domain; however, this formalism still relies on the SPA for the underlying precessing-frame signal, and is as such limited to inspiralling signals.
 Another intuitive approach consists in using a time-to-frequency correspondence, such as the one given by the SPA, to directly carry over the modulation to the Fourier domain.
-This corresponds, in a sense to be detailed below, to the zeroth-order approximation of the SUA, and it is used implicitly\footnote{With the difference that directly frequency-dependent expressions for the modulation are used, which are then extended to also cover high frequencies in the merger part of the signal, as we will discuss below.} for the simplified precession of the phenomenological waveforms of Ref.~\cite{Hannam+13}.
-Consistently extending these previous approaches to the merger-ringdown part of the signal is one of the objectives of this paper.
-
-Prospective parameter recovery for LISA observations of comparable-mass binary black hole coalescences is another frontier of gravitational wave signal modelling, becoming a very timely topic with the upcoming effort for the ESA L3 mission characterization in the next few years.
-In the past, many LISA data analysis studies used mainly post-Newtonian, inspiral-only models, and were limited to the Fisher-matrix formalism instead of full Bayesian analysis (among numerous works, see e.g.
-\cite{Cutler97,LISAPE09}).
-However, the Fisher matrix analysis, although considerably cheaper to perform, has known shortcomings including at high SNR~\cite{Vallisneri08,Cokelaer08,Rodriguez+13}.
-In Ref.~\cite{Klein+15}, an improved model was used, that included the effects of both spins including the corrections beyond the SPA investigated in~\cite{KCY14}; the limitation to inspiral-only waveforms was compensated by SNR reweighting using full IMR waveforms.
-
-The response of the LISA instrument, by contrast with the response of ground-based detectors, has the specificity of being time-dependent.
-For a signal accumulating over several months, the motion and change of orientation of the detector constellation along its orbit lead to time variability in the response of the detector, which is crucial for the localization of the gravitational-wave source in the sky.
-More precisely, these effects can be separated as a modulation coming from the rotation of the constellation, and a delay coming from the motion of the spacecrafts in the wavefront of the incoming wave.
-
-The purpose of this paper is to propose a formalism allowing to transpose directly in the Fourier domain the processing of a signal through a time-domain modulation and delay, while retaining the compactness of a Fourier-domain amplitude and phase representation of the signals.
-This allows us to address both the issue of a direct Fourier-domain response of the LISA instrument, as well as the issue of Fourier-domain precession modulation for IMR signals.
-
-[plan]
+This corresponds, in a sense to be detailed below, to the zeroth-order approximation of the SUA, and it is used implicitly\footnote{With the difference that directly frequency-dependent expressions for the modulation are used, which are then extended to also cover high frequencies in the merger-ringdown part of the signal, as we will discuss below.} for the simplified precession of the phenomenological waveforms of Ref.~\cite{Hannam+13}.
+In the case of the LISA response, the SPA provides a natural map from time-domain (for the orbit) to Fourier-domain (for the signal)~\cite{Cutler97}.
+Ref.~\cite{Klein+15} included the orbital motion of the detector in the SUA treatment.
+Consistently extending these previous approaches to the merger-ringdown part of the signals, including the delays in the LISA case, is part of the objectives of this paper.
+
 The plan of the paper is as follows.
-In Sec.~\ref{sec:motivation}, we will introduce our notations and conventions for Fourier-domain waveforms, motivate the problem of Fourier-domain modulation and give the relevant timescales for both the response of LISA-type detectors and the modulation of precessing signals.
+In Sec.~\ref{sec:motivation}, we will introduce our notations and conventions for Fourier-domain waveforms, motivate the problem of Fourier-domain modulation and introduce the relevant timescales for both the response of LISA-type detectors and the modulation of precessing signals.
 In Sec.~\ref{sec:formalism}, we present our general formalism, give its leading order approximation as well as higher-order corrections, and introduce new timescales based on the Fourier-domain signal.

### 2017-03-28 EOBNRv2HMROM/EOBNRv2HMROM.c

@@ -455,6 +455,7 @@ int EOBNRv2HMROMCore(
       /* We use the SPA formula tf = -(1/2pi)*dPsi/df to estimate the correspondence between frequency and time */
       /* The frequency corresponding to the 22 peak is omega22peak/2pi, with omega22peak taken from the fit to NR in Pan&al 1106 EOBNRv2HM paper */
       double f22peak = fmin(omega22peakOfq(q)/(2*PI), Mf_ROM_max_ref); /* We ensure we evaluate the spline within its range */
+      /* Note : twopishifttime is almost 0 (order 1e-8) by construction for the 22 mode, so it does not intervene here */
       tpeak22estimate = -1./(2*PI) * gsl_spline_eval_deriv(spline_phi22, f22peak, accel_phi22);
       /* Determine the change in phase (to be propagated to all modes) required to have phi22(fRef) = 2*phiRef */
       phase_change_ref = 2*phiRef + (gsl_spline_eval(spline_phi22, fRef_geom, accel_phi22) - (twopishifttime - 2*PI*tpeak22estimate + 2*PI*deltatRef_geom) * fRef_geom - shiftphase);
@@ -543,7 +544,7 @@ int SimEOBNRv2HMROM(
   double Mtot_sec = Mtot * MTSUN_SI; /* Total mass in seconds */
 
   if ( q > q_max ) {
-    //printf( "Error - %s: q out of range!\nEOBNRv2HMROM is only available for a mass ratio in the range q <= %g.\n", __func__, q_max); 
+    //printf( "Error - %s: q out of range!\nEOBNRv2HMROM is only available for a mass ratio in the range q <= %g.\n", __func__, q_max);
     return FAILURE;
   }
 

### 2017-03-28 LISAinference/ComputeLISASNR.c

@@ -198,6 +198,18 @@ static void Read_Text_TDITD3Chan( RealTimeSeries** TDI1, RealTimeSeries** TDI2, 
   gsl_matrix_free(inmatrix);
 }
 
+/***************** Function to control that the TDI tag is allowed *****************/
+/* Only tdi allowed : A,E,T (built from X,Y,Z), either individually or all three together */
+static int AllowedTDItag(TDItag tag) {
+  int ret = 0;
+  if(tag==TDIAETXYZ) ret = 1;
+  else if(tag==TDIAXYZ) ret = 1;
+  else if(tag==TDIEXYZ) ret = 1;
+  else if(tag==TDITXYZ) ret = 1;
+  else ret = 0;
+  return ret;
+}
+
 /***************** Main program *****************/
 
 int main(int argc, char *argv[])
@@ -213,7 +225,7 @@ int main(int argc, char *argv[])
   parse_args_ComputeLISASNR(argc, argv, params);
 
   /* NOTE: supports only AET(XYZ) (orthogonal) */
-  if(!(params->tagtdi==TDIAETXYZ)) {
+  if(!(AllowedTDItag(params->tagtdi))) {
     printf("Error in ComputeLISASNR: TDI tag not recognized.\n");
     exit(1);
   }
@@ -247,8 +259,8 @@ int main(int argc, char *argv[])
       RestrictFDReImFrequencySeries(&TDI3FFTrestr, TDI3FFT, params->minf, __LISASimFD_Noise_fHigh);
 
       /* Compute SNR with linear integration, weighting with non-rescaled noise functions */
-      /* Note: assumes same lengths for all FD FFT frequency seriess */
-      /* Note: not-rescaled noise functions */
+      /* Note: assumes same lengths for all FD FFT frequency series */
+      /* Note: non-rescaled noise functions */
       int sizeA = TDI1FFTrestr->freq->size;
       int sizeE = TDI2FFTrestr->freq->size;
       int sizeT = TDI3FFTrestr->freq->size;

### 2017-03-28 LISAsim/LISAFDresponse.c

@@ -269,6 +269,9 @@ int LISASimFDResponseTDI3Chan(
     gsl_vector* amp_imag = freqseries->amp_imag;
     gsl_vector* phase = freqseries->phase;
     int len = (int) freq->size;
+
+    //
+    //printf("len: %d\n", len);
     //
     //if(l==2&&m==1) printf("no resampling 21\n");
     //      if(l==2&&m==1) {for(int i=0; i<len; i++) printf("%d, %g, %g, %g, %g\n", i, gsl_vector_get(freq, i), gsl_vector_get(amp_real, i), gsl_vector_get(amp_imag, i), gsl_vector_get(phase, i));};
@@ -317,9 +320,11 @@ int LISASimFDResponseTDI3Chan(
     int resampled = 0; /* Keeps track of wether or not we resampled and allocated new resources we need to free */
     double maxfsignal = gsl_vector_get(freq, len-1);
     double fHigh = fmin(maxf, maxfsignal);
+    /* BEWARE : as Mathematica tests show, this is way too pessimistic - normally deltaf=0.002Hz sould work */
+    /* To be investigated */
     double fHigh_log_samp = 0.002;
     double deltaflineartarget = 0.00002; //1e-5 better, but slower
-    int ifmax = len-1; /* last index to be covered in original sampling overall*/
+    int ifmax = len-1; /* last index to be covered in original sampling overall */
     while((gsl_vector_get(freq, ifmax)>fHigh) && ifmax>0) ifmax--;
     int imaxlogsampling = ifmax; /* last index to be covered with original sampling */
     while((gsl_vector_get(freq, imaxlogsampling)>fHigh_log_samp) && imaxlogsampling>0) imaxlogsampling--;
@@ -332,7 +337,9 @@ int LISASimFDResponseTDI3Chan(
     //printf("%d\n", imaxlogsampling);
     //printf("maxf? %g, %g, %g\n", maxf, maxfsignal, fHigh);
     //printf(" test: %g, %i\n", ((fHigh - gsl_vector_get(freq, imaxlogsampling))/deltaflineartarget),ifmax-imaxlogsampling );
-    if((fHigh>fHigh_log_samp) && ((fHigh - gsl_vector_get(freq, imaxlogsampling))/deltaflineartarget)>ifmax-imaxlogsampling ) { /* condition to check if the linear sampling will add points - if not, do nothing */
+
[…]

### 2017-06-07 docs/paper_method.tex

@@
 
-A useful representation of the gravitational waveform is given by its decomposition in spin-weighted spherical.
+A useful representation of the gravitational waveform is given by its decomposition in spin-weighted spherical harmonics.
 The waveform emitted in the direction $(\Theta, \Phi)$, with its two polarizations $h_{+},h_{\times}$, can be decomposed as a superposition of modes as~\cite{Thorne80}
@@
 where one applies a truncation approximation at the first and last stage.
-The second equality holds provided that the complex coefficients $a_{N,k}$ are a solution of the $N+1$-dimensional linear system
+The second equality holds provided that the complex coefficients $a_{N,k}$ are a solution of the $N+1$-dimensional linear system~\cite{KCY14}
 \be\label{eq:stencilsystem}

### 2017-09-05 docs/paper_method.tex
Added option for simplified likelihood. Edits to the paper and updated plots.
The simplified likelihood is for frozen LISA, in the low-f approximation. The response becomes purely external for fixed masses. At setup the appropriate prefactor is computed once and for all. The likelihood computation becomes a very fast evaluation of a closed-form analytical expression.

@@
-\documentclass[aps,showpacs,%twocolumn,
-prd,superscriptaddress,nofootinbib]{revtex4}
+\documentclass[aps,showpacs,twocolumn,
+prd,superscriptaddress,nofootinbib]{revtex4-1}
 
@@
 \usepackage{ulem}
+\usepackage{tabularx}
 
@@
 %\captionsetup{compatibility=false}
+
+\usepackage{color}
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+%\usepackage[colorlinks=true,
+%            linkcolor=green,
+%            urlcolor=blue,
+%            citecolor=red]{hyperref}
+\usepackage[colorlinks=true,
+            linkcolor=YellowOrange,
+            urlcolor=RoyalBlue,
+            citecolor=RedViolet]{hyperref}
 
@@
 \newcommand\calT{{\mathcal{T}}}
+\newcommand\calD{{\mathcal{D}}}
+\newcommand\calA{{\mathcal{A}}}
 \newcommand\bfx{\mathbf{x}}
@@
 
+\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
+\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
+
+\newcommand{\SM}[1]{{\color{Red} #1}}
+
 \begin{document}
@@
 With the recent detections of gravitational-wave signals emitted by the coalescences of binary black holes, announced by the LVC collaboration after the first obervation run of the two LIGO instruments~\cite{LIGO-theevent-2016,LIGO-christmasevent-2016,LIGO-O1BBH-2016}, gravitational-wave astronomy has entered its observational era.
-With the recent start of LIGO's second observation run, and later the extension of the ground-based detectors network with Virgo~\cite{Virgo}, KAGRA~\cite{KAGRA} and LIGO-India~\cite{INDIGO}, many more detections from compact object coalescences are to come, with binary black holes being expected to be the dominant sources~\cite{}.
+With the recent start of LIGO's second observation run, and later the extension of the ground-based detectors network with Virgo~\cite{Virgo}, KAGRA~\cite{KAGRA} and LIGO-India~\cite{INDIGO}, many more detections from compact object coalescences are to come, with binary black holes being expected to be the dominant sources~\cite{} \SM{[C]}.
 Moreover, the European Space Agency has issued a mission call for its L3 space mission with a tentative launch around 2034, adressing the science theme ``The Gravitational Universe'', possibly opening the way for a space-borne observatory based on the concept of the Laser Interferometer Space Antenna (LISA)~\cite{LISA09,LISA17}.
@@
 In Sec.~\ref{sec:motivation}, we will introduce our notations and conventions for Fourier-domain waveforms, motivate the problem of Fourier-domain modulation and introduce the relevant timescales for both the response of LISA-type detectors and the modulation of precessing signals.
-In Sec.~\ref{sec:formalism}, we present our general formalism, give its leading order approximation as well as higher-order corrections, and introduce new timescales based on the Fourier-domain signal.
-In Sec.~\ref{sec:quadphasedelay}, we refine the previous results for both the quadratic-in-phase corrections and the treatment of the delays, and make the connection to previous approaches of the problem.
+In Sec.~\ref{sec:formalism}, we present our general formalism, give its leading order approximation as well as higher-order corrections, introduce new timescales based on the Fourier-domain signal, refine the previous results for both the quadratic-in-phase corrections and the treatment of the delays, and make the connection to previous approaches of the problem.
 We then apply our formalism to the response of the LISA detector in Sec.~\ref{sec:LISA}, and to the case of signals from precessing binaries in Sec~\ref{sec:precession}.
@@
 \ee
-where the ${}_{-2}Y_{\ell m}(\Theta,\Phi)$ are spin-weighted spherical harmonics.
-For a non-precessing system, with a fixed orbital plane, the individual modes have the additional symmetry property $h_{\ell -m} = (-1)^{\ell} h_{\ell m}^{*}$.
+where the ${}_{-2}Y_{\ell m}(\Theta,\Phi)$ are spin-weighted spherical harmonics. For a non-precessing system, with a fixed orbital plane, the individual modes have the additional symmetry property
+\be\label{eq:symmetryhlminusm}
+	h_{\ell -m} = (-1)^{\ell} h_{\ell m}^{*} \,.
+\ee
 Since we will work in the Fourier domain, it will be useful to write the contributions of the individual modes to the Fourier transforms of the polarizations as
@@
 \begin{align}
-	\tilde{h}_{+}(f) &= \frac{1}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left( {}_{-2}Y_{\ell m} \widetilde{h_{\ell m}}(f) + {}_{-2}Y_{\ell m}^{*} \widetilde{h_{\ell m}}(-f)^{*} \right) \,, \\
-	\tilde{h}_{\times}(f) &= \frac{i}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left( {}_{-2}Y_{\ell m} \widetilde{h_{\ell m}}(f) - {}_{-2}Y_{\ell m}^{*} \widetilde{h_{\ell m}}(-f)^{*} \right) \,,
+	\tilde{h}_{+}(f) &= \frac{1}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left( {}_{-2}Y_{\ell m} \tilde{h}_{\ell m}(f) + {}_{-2}Y_{\ell m}^{*} \tilde{h}_{\ell m}(-f)^{*} \right) \,, \\
+	\tilde{h}_{\times}(f) &= \frac{i}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left( {}_{-2}Y_{\ell m} \tilde{h}_{\ell m}(f) - {}_{-2}Y_{\ell m}^{*} \tilde{h}_{\ell m}(-f)^{*} \right) \,,
 \end{align}
 \end{subequations}
-where we used $\widetilde{h_{\ell m}^{*}}(f) = \widetilde{h_{\ell m}}(-f)^{*}$.
+where we used $\widetilde{h_{\ell m}^{*}}(f) = \tilde{h}_{\ell m}(-f)^{*}$.
 An additional approximation often used in waveform models consists in considering that the Fourier transforms $\tilde{h}_{\ell m}$ have support only on one side of the spectrum, either for positive of for negative frequencies depending on the sign of $m$.
@@
 With our sign convention~\eqref{eq:defFT}, this approximation reads
-\begin{align}\label{eq:approxnegativef}
+\begin{align}\label{eq:zeronegativef}
 	\tilde{h}_{\ell m} (f) &\simeq 0 \text{ for } f<0, \; m>0 \nn\,,\\
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.48\linewidth]{plots/Af.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/Psif.pdf}
-  \caption{Fourier-domain amplitude and phase in geometric units for an equal-mass, non-spinning system.
+%\begin{figure*}
+%  \centering
+%  \includegraphics[width=.48\linewidth]{plots/Af.pdf}
+%  \hspace{0.2cm}
+%  \includegraphics[width=.48\linewidth]{plots/Psif.pdf}
+%  \caption{Fourier-domain amplitude and phase in geometric units for an equal-mass, non-spinning system.
 The vertical line represents $\omega_{22}^{\rm peak}/2\pi$, the frequency corresponding to the time-domain frequency of the waveform at the peak amplitude.
@@
 [add residuals of interpolation on $~300$ points, to illustrate the compression of Amp/Phase FD waveforms ?
-or show different spins/mass ratios ?]}
-  \label{fig:ampphase}
-\end{figure}
-
-We recall that, with our sign convention, the effect of a positive shift in time $h_{\Delta t} (t) = h(t- \Delta t)$ is
-\be
-	\tilde{h}_{\Delta t}(f) = e^{2i\pi f \Delta t} \tilde{h}(f) \,.
+or show different spins/mass ratios ?] \SM{[Plot perhaps not necessary ?]}}
+%  \label{fig:ampphase}
+%\end{figure*}
+
+We recall that, with our sign convention, the effect of a positive shift in time $h' (t) = h(t- \Delta t)$ is
+\be
+	\tilde{h}'(f) = e^{2i\pi f \Delta t} \tilde{h}(f) \,.
 \ee
@@
 Using the conditions above, one can formally expand the signal around $t_{f}^{\rm SPA}$ to quadratic order in time, according to
-\be
-	\tilde{h}_{\rm SPA} (f) \simeq a(t_{f}^{\rm SPA}) \exp\left[2i\pi f t_{f}^{\rm SPA}-2i\varphi(t_{f}^{\rm SPA}) \right]  \int \ud t \, e^{-i \dot{\omega} (t_{f}^{\rm SPA}) (t-t_{f}^{\rm SPA})^{2}} \,.
-\ee
+\begin{align}
+	\tilde{h}_{\rm SPA} (f) &\simeq a(t_{f}^{\rm SPA}) \exp\left[2i\pi f t_{f}^{\rm SPA}-2i\varphi(t_{f}^{\rm SPA}) \right] \nn\\
+	& \qquad \cdot \int \ud t \, e^{-i \dot{\omega} (t_{f}^{\rm SPA}) (t-t_{f}^{\rm SPA})^{2}} \,.
+\end{align}
 Note that to be able to treat the amplitude as a constant in the integral above, we used the third condition in~\eqref{eq:conditionSPA}.
@@
 	A_{N}(f) &= \frac{G^{2}m^{2} \pi}{Dc^{5}} \sqrt{\frac{2\nu}{3}} v^{-7/2}\,,\\
-	\Psi_{\rm SPA}^{\rm N}(f) &= 2\pi f t_{0} - \phi_{0} + \frac{3}{128\nu v^{5}} \,, 
+	\Psi_{\rm SPA}^{\rm N}(f) &= \phi_{0} - 2\pi f t_{0} - \frac{3}{128\nu v^{5}} \,, 
 \end{align}
 \end{subequations}
+\SM{[Check sign of $\Psi$ ? Should have $<0$ 2nd der.] }
 where $v=(G m \pi f/c^{3})^{1/3}$, $D$ is the luminosity distance to the observer, and where $t_{0}, \phi_{0}$ are constants.
@@
 We refer the reader to~\cite{Vallisneri04} for a comparative account on these various conventions.
-In this work, we will use the conventions of~\cite{Vallisneri04}, which were also used in the Mock LISA Data Challenges (MLDC)~\cite{}.
+In this work, we will use the conventions of~\cite{Vallisneri04}, which were also used in the Mock LISA Data Challenges (MLDC)~\cite{} \SM{[C]}.
 
@@
 Written in terms of the fractional laser frequency shifts between two spacecrafts, the elementary response of the detector reads~\cite{EW75, RCP04, Vallisneri04}
-\be\label{eq:yslr}
-	y_{slr} \equiv \frac{\nu_{r} - \nu_{s}}{\nu} = \frac{1}{2} \frac{n_{l}^{i}n_{l}^{j}}{1 - \hat{k}\cdot n_{l}} \left[ h_{ij}^{\rm TT}(t - \hat{k}\cdot p_{s}) - h_{ij}^{\rm TT}(t - \hat{k}\cdot p_{r}) \right]
-\ee
+\begin{align}\label{eq:yslr}
+	y_{slr} &\equiv \frac{\nu_{r} - \nu_{s}}{\nu} \nn\\
+	&= \frac{1}{2} \frac{n_{l}^{i}n_{l}^{j}}{1 - \hat{k}\cdot n_{l}} \left[ h_{ij}^{\rm TT}(t - \hat{k}\cdot p_{s}) - h_{ij}^{\rm TT}(t - \hat{k}\cdot p_{r}) \right] \,.
+\end{align}
 As detailed below in Sec.~\ref{}, we will use a rigid instantaneous model, approximating the geometry of the constellation by a moving equilateral triangle (neglecting the flexing of the arms induced by corrections in the orbits), and evaluating all geometric factors in~\eqref{eq:yslr} at a single time $t$ (neglecting pointing-ahead corrections).
@@
 We will give in Sec.~\ref{subsec:modelLISA} more details on the response of LISA-type detectors, on the TDI observables, and on the approximations that enter the derivation of the basic response~\eqref{eq:yslr} above.
+\SM{[explicit expressions for TDI]}
 
@@
 \be
-	s(t) = F(t) h(t + d(t))
+	s(t) = F(t) h(t + d(t)) \,.
 \ee
@@
 
-It was suggested by a number of authors~\cite{} that the effect of the precession can be modeled, to a good approximation, by a time-dependent rotation of a non-precessing waveform.
+It was suggested by a number of authors~\cite{} \SM{[C]} that the effect of the precession can be modeled, to a good approximation, by a time-dependent rotation of a non-precessing waveform.
 In this approach, one introduces a precessing frame following the evolution of the plane of the orbit, and approximates the waveform in this precessing frame by using a non-precessing model.
-Different prescriptions have been proposed for the construction of the precessing frame itself~\cite{}.
+Different prescriptions have been proposed for the construction of the precessing frame itself~\cite{} \SM{[C]}.
 
 If $(\alpha, \beta, \gamma)$ are the Euler angles relating the precessing frame to the inertial frame in the $(z,y,z)$ convention, the modes in the inertial frame $h_{\ell m}^{\rm I}$ are then related to the modes in the precessing frame $h_{\ell m}^{\rm P}$ by~\cite{Goldberg+67}
-\be\label{eq:wignerrot}
-	h_{\ell m}^{\rm I} = \sum\limits_{m=-\ell}^{\ell} D^{\ell *}_{m'm} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm P} \,,
-\ee
-where the $D^{\ell}_{mm'}$ are Wigner matrices~\cite{}. Notice that there is no mixing of the modes with different values of $\ell$.
+\be\label{eq:wignerrotintro}
+	h_{\ell m}^{\rm I} = \sum\limits_{m=-\ell}^{\ell} \calD^{\ell *}_{mm'} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm P} \,,
+\ee
+where the $\calD^{\ell}_{mm'}$ are Wigner matrices~\cite{} \SM{[C]}. Notice that there is no mixing of the modes with different values of $\ell$.
 
@@
 where the modulation function $F(t)$ is given by a Wigner matrix and depends on time through the Euler angles $(\alpha, \beta, \gamma)(t)$.
-We will discuss more in Sec.~\ref{subsec:modelPrec} the limitations of the assumptions used to perform the decomposition above.
+We will discuss more in Sec.~\ref{subsec:modelPrec} \SM{[R]} the limitations of the assumptions used to perform the decomposition above.
 
@@
 The separation between the precessional timescale and the orbital timescale will be crucial for our analysis.
-In the limit of low frequencies, we will see in Sec.~\ref{} below that, although these timescales become more and more separated, the decrease in the chirping rate gives raise to a corrective contribution that does not vanish in this limit.
-In the other limit, as the system gets close to merger, the separation of timescales becomes weaker, and we will explore in Sec.~\ref{} the application of our formalism to the merger and ringdown phase.
+In the limit of low frequencies, we will see in Sec.~\ref{} \SM{[R]} below that, although these timescales become more and more separated, the decrease in the chirping rate gives raise to a corrective contribution that does not vanish in this limit.
+In the other limit, as the system gets close to merger, the separation of timescales becomes weaker, and we will explore in Sec.~\ref{} \SM{[R]} the application of our formalism to the merger and ringdown phase.
 
@@
 \ee
-where we introduced the frequency-dependent function of time $G(f,t)$, and its Fourier transform in the auxiliary frequency $f'$, denoted by $\tilde{g}(f,f')$, as
+where we introduced the frequency-dependent function of time $G(f,t)$, and its Fourier transform in the auxiliary frequency $f'$, denoted by $\tilde{G}(f,f')$, as
 \begin{subequations}\label{eq:defg}
@@
 \begin{itemize}
-	\item the time scales entering the signal $h(t)$ are well separated from the ones entering the modulation and the delay, so that $\tilde{G}(f-f',f')$ has support limited to $[-f_{\rm max},f_{\rm max}]$, with $f_{\rm max}$ small compared the characteristic scale for the variation of the Fourier-domain signal;
+	\item the time scales entering the signal $h(t)$ are well separated from the ones entering the modulation and the delay, so that $\tilde{G}(f-f',f')$ has support limited to $|f'| < f_{\rm max}$, with $f_{\rm max}$ small compared the characteristic scale for the variation of the Fourier-domain signal;
 	\item the original signal has a smooth amplitude and phase in the Fourier-domain, slowly varying over the scale $f_{\rm max}$, so that a local treatment in $f'$ of the convolution in~\eqref{eq:FDkernel} will be possible.
@@
 
-\subsection{Leading order: the local approximation}
+\subsection{Leading order: the local approximation \SM{[Find a better name for this approximation ?]}}
 \label{subsec:LLP}
@@
   \centering
-  \includegraphics[width=.48\linewidth]{plots/tf.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/tfzoom.pdf}
-  \caption{Time-to-frequency correspondence, as defined directly from the phase of the Fourier-domain signal as in~\eqref{eq:deftf}, in geometric units for an equal-mass, non-spinning system.
-The right panel shows the merger and ringdown part of the signal in more details.
-The time axis is such that the time-domain amplitude peaks at $t=0$.
-The vertical line shows the instantaneous frequency at the peak, and the fact that the curve $t_{f}$ does not pass exactly by the crossing of the two lines reflects the fact that the link between time-domain frequency and Fourier-domain f is not accurate anymore in the merger part of the signal.
-Note that $t_{f}$ does not go beyond $20M$ after the peak, and is not monotonous.
-[show different spins/mass ratios ?]}
+  \includegraphics[width=.98\linewidth]{plots/tf.pdf}
+  \caption{Time-to-frequency correspondence $t_{f}$, as defined directly from the phase of the Fourier-domain signal as in~\eqref{eq:deftf}, in geometric units.
+We show only the high-frequency part of the signal, corresponding to the merger region.
+The PhenomD waveforms have been aligned such that the time-domain amplitude (obtained by an IFFT) peaks at $t=0$.
+The two aligned spins are equal, with components of $0.95$ ($++$, blue), $0$ ($00$, red), and $-0.95$ ($--$, yellow), for mass ratios $q=1$ (full line) and $q=8$ (dashed).
+The vertical lines shows the time-domain instantaneous frequency at the peak $\omega^{22}_{\rm peak}/(2\pi)$, again measured from an IFFT, and the fact that the curves $t_{f}$ do not pass exactly by the crossing with the $t=0$ horizontal line reflects the fact that the link between time-domain frequency and Fourier-domain $f$ is not accurate anymore in the merger part of the signal.
+Note that $t_{f}$ does not go beyond $~20M$ after the peak, and is not monotonous.
+\SM{[add color to the vertical lines]} \SM{[add dots to highlights the crossings]}}
   \label{fig:tf}
@@
 Using this notation, we have simply
-\be\label{eq:local}
+\be\label{eq:transferlocal}
 	\calT_{\rm local}(f) = G(f, \tf) \,.
@@
 The interpretation of this approximation is straightforward: the signal is simply multiplied by the response function evaluated at the time $\tf$.
+For precessing binary, this corresponds to the approximation used for the PhenomP approximant~\cite{Hannam+13}, as will be discussed in Sec.~\ref{}.
 Additionally, when delays are present, the frequency-dependence of the response is evaluated directly at $f$.
-For precessing binary, this corresponds to the approximation used for the PhenomP approximant~\cite{Hannam+13}, as will be discussed in Sec.~\ref{}.
-
-It is worth noting that a shift in time of the time-domain signal will, by construction, be appropriately propagated to~\eqref{eq:local}.
+
+It is worth noting that a shift in time of the time-domain signal will, by construction, be appropriately propagated to~\eqref{eq:tranferlocal}.
 If $h'(t) = h(t -  \Delta t)$ [but the modulation and delays are left unchanged (which is relevant for space-based detectors, not for precessing binaries)], $\Psi'(f) = \Psi(f) - 2\pi f \Delta t$, so that
@@
 This observation also justifies that the first derivative of the Fourier-domain phase cannot, in general, be considered a small quantity, as it can take arbitrary values for arbitrary shifts in time of the signal.
-The leading order approximation corresponds to neglecting the curvature in $\Psi(f)$, and yields a result~\eqref{eq:local} where the passage to the Fourier domain is done by simply using a time-to-frequency correspondence~\eqref{eq:deftf}, which motivates the name ``local approximation''.
+The leading order approximation corresponds to neglecting the curvature in $\Psi(f)$, and yields a result~\eqref{eq:transferlocal} where the passage to the Fourier domain is done by simply using a time-to-frequency correspondence~\eqref{eq:deftf}, which motivates the name ``local approximation''.
 
@@
 
-We now extend our approach beyond the LLP approximation.
+We now extend our approach beyond the local approximation.
 We will give a formal derivation of the  general formula for the higher-order corrections in~\eqref{eq:FDkernel}, and then specialize it to the special cases of interest.
@@
 
+\begin{widetext}
 If we further expand the exponentials to obtain a pure $f'$-expansion, by applying the derivative rule
+\allowdisplaybreaks
 \be
@@
 \ee
-we arrive at the formal expansion
+we arrive at the formal combined expansion
 \begin{align}
@@
 where $m = p+r+2k_{2}+\dots+N k_{N}$.
+\end{widetext}
 
@@
 \be\label{eq:taylordelay}
-	\calT(f) = \sum\limits_{r\geq 0} \frac{1}{(2i\pi)^{r}r!} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{r} }{\partial t^{r}} G \right)(f,\tf)
+	\calT_{\rm delay}(f) = \sum\limits_{r\geq 0} \frac{1}{(2i\pi)^{r}r!} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{r} }{\partial t^{r}} G \right)(f,\tf)
 \ee
@@
 \be\label{eq:resultdfA}
-	\calT(f) = \sum\limits_{q\geq 0} \frac{1}{(2i\pi)^{q}q!} \frac{1}{A} \frac{\ud^{q} A}{\ud f ^{q}}  \left( \frac{\partial^{q} }{\partial t^{q}} G \right)(f,\tf)
+	\calT_{\rm amp}(f) = \sum\limits_{q\geq 0} \frac{1}{(2i\pi)^{q}q!} \frac{1}{A} \frac{\ud^{q} A}{\ud f ^{q}}  \left( \frac{\partial^{q} }{\partial t^{q}} G \right)(f,\tf)
 \ee
@@
 \be\label{eq:resultdffPsi}
-	\calT(f) = \sum\limits_{p\geq 0} \frac{1}{p!} \left( \frac{i}{8\pi^{2}}\frac{\ud^{2} \Psi}{\ud f^{2}} \right)^{p} \left( \frac{\partial^{2p} }{\partial t^{2p}} G \right)(f, \tf) \,.
-\ee
-Indeed, this correction will be the most significant correction beyond leading order in the applications that we will consider in Secs.~\ref{} and~\ref{}, and this particular approximation will turn out to be closely related to the approach of~\cite{KCY14}.
-
-In the following, it will be convenient to introduce figures of merit of these Taylor-like expansions. Thus we define
+	\calT_{\rm phase}(f) = \sum\limits_{p\geq 0} \frac{1}{p!} \left( \frac{i}{8\pi^{2}}\frac{\ud^{2} \Psi}{\ud f^{2}} \right)^{p} \left( \frac{\partial^{2p} }{\partial t^{2p}} G \right)(f, \tf) \,.
+\ee
+Indeed, this correction will be shown in Sec.~\ref{subsec:resumquadphase} to be closely related to the previous approach of~\cite{KCY14}.
+
+In the following, it will be convenient to introduce \SM{[think of removing ``figure of merit'' as it is not quite the word]} error estimates for these Taylor-like expansions, that we will take to be the magnitude of the first ignored term in the series.
+Thus we define
 \begin{subequations}\label{eq:deffom}
@@
 
+\SM{[interpretation of $\epsilon$'s as ratios of timescales - next subsec]}
+
+\SM{[stress that these are errors in the relative sense - so less important in RD than it seems]}
+
 We found in both the LISA case and in the precession case that the third and higher derivatives of both the amplitude and phase are negligible.
-Note that the figure of merit $\epsilon_{d}$, introduced to represent the effect of the frequency dependence in the convolution kernel~\eqref{eq:FDkernel}, is signal-independent and depends only on $G$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Signal-dependent timescales and relation to the stationary phase approximation}
+\SM{[maybe to be recalled below]} Note that the figure of merit $\epsilon_{d}$, introduced to represent the effect of the frequency dependence in the convolution kernel~\eqref{eq:FDkernel}, is signal-independent and depends only on $G$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Signal-dependent timescales and relation to the stationary phase approximation and SUA}
 \label{subsec:linkSPA}
 
-\begin{figure}
+\begin{figure*}
   \centering
@@
   \label{fig:TfTA}
-\end{figure}
+\end{figure*}
 
@@
 \begin{align}\label{eq:resultdffPsiTf}
-	 \calT(f) &= \sum\limits_{p\geq 0} \frac{(-i\epsilon)^{p}}{2^{p}p!} \Tf^{2p} \left( \frac{\partial^{2p} }{\partial t^{2p}} G \right)(f, \tf) \,,
-\end{align}
-with $\epsilon = -\mathrm{sgn}(\ud^{2}\Psi/\ud f^{2} )$.
+	 \calT_{\rm phase}(f) &= \sum\limits_{p\geq 0} \frac{(-i\epsilon)^{p}}{2^{p}p!} \Tf^{2p} \left( \frac{\partial^{2p} }{\partial t^{2p}} G \right)(f, \tf) \,,
+\end{align}
+where $\epsilon = -\mathrm{sgn}(\ud^{2}\Psi/\ud f^{2} )$ is $1$ in the inspiral.
 
@@
 The derivation includes an expansion of the modulation in Bessel functions, a study of the displacement of the stationary phase point induced by the precession, and a resummation of the result in the form of time derivatives.
-The main result of~\cite{KCY14}, their equation~(34), reads exactly like~\eqref{eq:resultdffPsiTf} with the identifications $\tilde{H}_{corr}(f)\rightarrow \tilde{s}(f)/\tilde{h}(f)$ for the transfer function, $T\rightarrow \Tf$ for the radiation-reaction timescale and $e^{-i\delta\phi} \rightarrow F$ for the modulation function (which is restricted in their framework to include only the phase of the modulation, its amplitude being treated jointly with the amplitude of the signal).
+The main result of~\cite{KCY14}, their equation~(34), reads exactly like~\eqref{eq:resultdffPsiTf} with the identifications $\tilde{H}_{corr}(f)\rightarrow \calT(f)$ for the transfer function, $T\rightarrow \Tf$ for the radiation-reaction timescale and $e^{-i\delta\phi} \rightarrow F$ for the modulation function (which is restricted in their framework to include only the phase of the modulation, its amplitude being treated jointly with the amplitude of the signal).
 Thus, we are able to make a direct correspondence between the treatment of~\cite{KCY14} and our framework, which provides us with a simple and alternative derivation of their result.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Quadratic phase corrections and delays}
-\label{sec:quadphasedelay}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Resummation of derivatives for quadratic term in the phase}
+\SM{[add: $\sim$ The relevant scaling for the terms in the series~\eqref{eq:resultdffPsiTf}, for the inspiral phase where the SPA applies, is therefore
+\be
+	\Tf^{2p} \partial_{t}^{2p} F = \calO(-1) \,,
+\ee]}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\section{Quadratic phase corrections and delays}
+%\label{sec:quadphasedelay}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Quadratic term in the phase}
 \label{subsec:resumquadphase}
@@
 The two solutions for the stencil coefficients in the two cases $\epsilon = \pm 1$ are simply related by a complex conjugation.
+\SM{[give the rational number expressions of $a_{N,k}$ for the first values of $N$; add derivation in appendix]}
 
@@
 
-\subsection{Quadratic term in the phase and Fresnel transform}
+\subsection{Relation to the Fresnel transform}
 \label{subsec:fresneltransform}
@@
 where we recall that $\epsilon = -\mathrm{sgn}(\ud ^{2} \Psi/\ud f^{2})$.
-The integral on $f'$ is a simple complex Gaussian integral, and results in a complex Gaussian integral over time where we recognize a Fresnel transform~\cite{} of the function $F$.
+The integral on $f'$ is a simple complex Gaussian integral, and results in a complex Gaussian integral over time where we recognize a Fresnel transform~\cite{} \SM{[C]} of the function $F$.
 We introduce for the Fresnel transform the notation
@@
 The Fresnel transform is an integral transform that shares some properties with the Fourier transform.
-It intervenes in particular in the theory of diffraction~\cite{}.
+It intervenes in particular in the theory of diffraction~\cite{} \SM{[C]}.
 Contrarily to the Fourier transform, however, it is localized in the sense that the part of the integral that is centered around $t_{0}$ contributes predominantly, due to the cancelling oscillations far from $t_{0}$.
@@
 As a verification, upon performing a formal Taylor expansion in time of $F(t)$ around $\tf$ in the above integral and integrating term by term the result\footnote{Note that the intermediate integrals produced by such a Taylor expansion are formally divergent.
-One can regularize them for instance by introducing a small imaginary part in $\tau$ that and sending it to $0$ at the end of the computation.}, one readily recovers our previous result~\eqref{eq:resultdffPsiTf}.
+One can regularize them for instance by introducing a small imaginary part in $\tau$ that is sent it to $0$ at the end of the computation.}, one readily recovers our previous result~\eqref{eq:resultdffPsiTf}.
 Now, the resummation of~\eqref{eq:resumquadphase} can be rephrased as a quadrature rule for computing the Fresnel integral~\eqref{eq:defFresnel}, according to
@@
 Due to the complex character of the kernel in~\eqref{eq:defFresnel}, the induced inner product is not positive definite and some key properties are lost, such as the construction of the optimal Gaussian nodes from the zeros of the orthonormalized polynomials.
-For more details on quadrature rules when applied to integrals with oscillatory kernels, we refer to~\cite{}.
+For more details on quadrature rules when applied to integrals with oscillatory kernels, we refer to~\cite{} \SM{[C]}.
 
@@
 Keeping as before only the first-order terms in $\dot{d}$ (and neglecting its higher derivatives), we can write
+\begin{widetext}
 \begin{align}
@@
 \end{align}
+\end{widetext}
 where we used a change of variable $\tau = t_{d}(t)$. We see that the result can again be expressed as a Fresnel transform.
@@
 Thus, we obtain our most general result for an exact treatment of the quadratic term in the phase and for an improved treatment of the delays as:
-\be
+\begin{widetext}
+\be\label{eq:transferfinal}
 	\calT(f) = \sum\limits_{k=0}^{+\infty} \frac{1}{(2i\pi)^{k} k!} \frac{1}{A} \frac{\ud^{k}A}{\ud f^{k}}  \calF^{\epsilon}_{\Tf} \left[ \frac{\ud^{k}}{\ud \tau^{k}} \left( \frac{F(\tau - d(\tau))}{1+\dot{d}(\tau)} e^{-2i\pi f d(\tau)(1-\dot{d}(\tau))} \right) \right] (\tf) \,.
 \ee
+\end{widetext}
 Each term in this series (in practice, only the first few will be relevant) can in turn be approximated by using the stencil formula~\eqref{eq:Fresnelstencil}.
@@
 
-The frequency-shift response for a single link~\eqref{eq:yslr} was derived in Ref.~\cite{EW75} (see also~\cite{} for a more modern version).
+The frequency-shift response for a single link~\eqref{eq:yslr} was derived in Ref.~\cite{EW75} (see also~\cite{} \SM{[C]} for a more modern version).
 Several assumptions enter the result as written in~\eqref{eq:yslr}: (i) effects of the order $v/c$ are neglected, including for instance the special relativistic Doppler effect created by the relative speeds of the spacecrafts on their orbits (ii) the propagation is assumed to take place in a flat spacetime, perturbed only by the gravitational wave; thus the gravitational redshift as well as the deflection of light created by the gravitational potential of the Sun is ignored (iii) all geometric factors are evaluated at a single time, whereas one should consider the beam as propagating from the position of the first spacecraft at the time of emission to the position of the second scapecraft at the time of reception, leading to a pointing-ahead effect.
@@
 
-Although all the effects (i)-(v) are neglected here, keeping track of these corrections is crucial for the purpose of noise cancellations, and led to the development of new generations of TDI observables~\cite{}.
+Although all the effects (i)-(v) are neglected here, keeping track of these corrections is crucial for the purpose of noise cancellations, and led to the development of new generations of TDI observables~\cite{} \SM{[C]}.
 
@@
 \ee
-Assuming that the approximation~\eqref{eq:approxnegativef} applies, we focus only on positive frequencies and consider a mode contribution $h$ following~\eqref{eq:notationhsignm}.
+Assuming that the approximation~\eqref{eq:zeronegativef} applies, we focus only on positive frequencies and consider a mode contribution $h$ following~\eqref{eq:notationhsignm}.
 We can define a complex matrix $P$ incorporating the spin-weighted spherical harmonic constant factor as
@@
 The Fourier-domain transfer function for~\eqref{eq:dfresponsec} then reads in the local approximation
-\be
-	\calT_{slr}^{\ell m} = \frac{i \pi f L}{2} \sinc \left[ \pi f L\left(1-\hat{k}\cdot n_{l} \right) \right] \exp\left[ i \pi f \left( L + \hat{k}\cdot \left( p_{1}^{L} + p_{2}^{L} \right) \right) \right]  n_{l} \cdot P_{\ell m} \cdot n_{l} \text{ for } m>0 \,,
-\ee
+\begin{align}
+	\calT_{slr} &= \frac{i \pi f L}{2} \sinc \left[ \pi f L\left(1-\hat{k}\cdot n_{l} \right) \right] \nn\\
+	& \qquad \cdot \exp\left[ i \pi f \left( L + \hat{k}\cdot \left( p_{1}^{L} + p_{2}^{L} \right) \right) \right]  n_{l} \cdot P_{\ell m} \cdot n_{l} \,,
+\end{align}
 where we used the fact that, in our rigid approximation for the trajectory of the constellation, $p^{L}_{r} - p^{L}_{s} =  L n_{l}$.
@@
 \label{subsec:sizecorrLISA}
+
+\SM{[add order-of-magnitude estimate for ET]}
+
+\SM{[check if there is something to be done with recent paper on ET modulations]}
+
+\SM{[add order-of-magnitude estimate for EMRIS]}
+
+\SM{[possibly add order-of-magnitude estimate for geocentric-LISA ?]}
+
+\SM{[justify somewhere smallness of cubic-in-phase term]}
+
+\SM{[crude estimate of the target accuracy level: 1/SNR ?]}
 
@@
 
-\begin{figure}
+\begin{figure*}
   \centering
@@
 The central line and interval are the mean and $1\sigma$ standard deviation of the logarithm of $\epsilon$ over 400 random values for the position in the sky, inclination and polarization.
-[change middle mass to $3\times 10^{4}$ - clarify mismatch between these $\epsilon$ and the actual size of corrections]}
+\SM{[change middle mass to $3\times 10^{4}$ ?
+- clarify mismatch between these $\epsilon$ and the actual size of corrections]} \SM{[got to 6 plots, completely separating orbital and constellation response]}}
   \label{fig:fomLISA}
-\end{figure}
+\end{figure*}
 
@@
   \label{fig:prectoymodel}
-\end{figure}
-
-\begin{figure}
+\end{figure*}
+
+\begin{figure*}
   \centering
@@
   \caption{Transfer function and reconstruction error for the orbital delay $d_{O}$, for $M=10^{2} \Msol$.}
-\end{figure}
-
-\begin{figure}
+\end{figure*}
+
+\begin{figure*}
   \centering
@@
   \label{fig:prectoymodel}
-\end{figure}
+\end{figure*}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Simplified model of precessing IMR waveform}
+\SM{[explain somewhere the fact that the modulation has support for negative f, convolution has support to larger f - pretty important for direct convolution]}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Precession and frame decomposition}
+\label{subsec:precdef}
+
+We will now turn to the application of our formalism to the decomposition of signals from precessing binaries.
+As we already discussed in Sec.~\ref{subsec:modulationPrec}, in the presence of spin components that are not aligned with the orbital angular momentum, an inspiralling binary system will undergo precession of its orbital plane~\cite{Apostolatos+94, Kidder95}.
+This is a crucial effect to be taken into account in the modelling of such signals.
+
+As proposed by several authors~\cite{} \SM{[C]}, if one performs the mode decomposition of the waveform~\eqref{eq:defmodes} not in a fixed inertial frame but rather in a time-dependent, rotating frame that follows the plane of the orbit, it is possible to restore much of the structure of a non-precessing waveform.
+In particular, one recovers qualitatively the hierarchy of mode amplitudes that prevails for non-precessing systems, with the modes $h_{22}$ and $h_{2,-2}$ being dominant.
+
+We define $(\alpha, \beta, \gamma)$ are the Euler angles relating the precessing frame to the inertial frame in the $(z,y,z)$ convention.
+The $z$ axis of the inertial frame is chosen to be the direction of $\bm{J}$, the total angular momentum, which is almost constant\footnote{Except in the cases of transitional precession~\cite{Apostolatos94}, where for high mass ratios and large antialigned spins the orbital angular momentum and the spin can almost cancel each other.}.
+The first two angles, $\alpha$ and $\beta$, are the two spherical angles tracking the direction of dominant radiation reaction, which during the inspiral follows essentially the normal to the orbital plane.
+The last angle $\gamma$ parametrizes the remaining freedom of rotation around this radiation axis.
+An important element of the construction of the precessing frame is the fixing of this third degree of freedom, for which we use the minimal rotation condition~\cite{} \SM{[C]}
+\be\label{eq:gammadot}
+	\dot{\gamma} = -\dot{\alpha}\cos \beta \,.
+\ee
+Quite generically, and in particular in the case of simple precession~\cite{Apostolatos+94, Kidder95} but also in the model we will take for the post-merger precession, the radiation axis precesses on a cone around the direction of $\bm{J}$, with $\alpha$ increasing, while $\beta$, the opening angle of the precession cone, is slowly varying.
+
+The modes in the inertial frame $h_{\ell m}^{\rm I}$ and the modes in the precessing frame $h_{\ell m}^{\rm P}$ are then related by~\cite{Goldberg+67}
+\begin{subequations}
+\label{eq:wignerrot}
+\begin{align}
+	h_{\ell m}^{\rm I} = \sum\limits_{m=-\ell}^{\ell} \calD^{\ell *}_{mm'} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm P} \,, \\
+	h_{\ell m}^{\rm P} = \sum\limits_{m=-\ell}^{\ell} \calD^{\ell }_{m'm} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm I} \,,
+\end{align}
+\end{subequations}
+where the coefficients $\calD^{\ell}_{mm'}$ are given by Wigner D-matrices~\cite{} \SM{[C]}, which take the expression
+\be\label{eq:defWignerD}
+	\calD^{\ell}_{mm'} (\alpha, \beta, \gamma) = e^{im \alpha} d^{\ell}_{mm'}(\beta) e^{im' \gamma}\,,
+\ee
+where the Wigner d-matrix $d^{\ell}_{mm'}(\beta)$ takes the form of a polynomial in $\cos (\beta/2)$, $\sin (\beta/2)$, and acts as an amplitude for the modulation function.
+We refer to App.~\ref{app:wigner} for explicit expressions.
+
+An important qualitative observation is that, for small to moderate opening angles, the modulation functions show a clear hierarchy in amplitude.
+This situation is quite generic, since large misalignments between $\bm{J}$ and $\bm{\ell}$ can only be reached with large spins and large mass ratios.
+From the explicit expressions~\eqref{} \SM{[R]} and~\eqref{} \SM{[R]}, one can see that $d^{\ell}_{mm'}(\beta)$ is greatly suppressed in the limit $\beta \rightarrow 0$ when increasing $|m-m'|$.
+Intuitively, this means that, in the limit of a small misalignment between frames, the rotation produces mainly mode contributions with the same mode number.
+Similarly, when $\beta$ is constant or slowly varying, \eqref{eq:gammadot} gives $\gamma \simeq - \alpha \cos\beta$, and we have $\calD^{\ell *}_{mm'} (\alpha, \beta, \gamma) \propto e^{i(m' \cos\beta - m) \alpha}$ with $\cos\beta \simeq 1$ for small $\beta$.
+This shows that the modulation for $m=m'$ has a suppressed phase, while increasing $|m-m'|$ increases the magnitude of the modulation phase.
+Thus, modulation functions for distant mode contributions (for instance from the $h^{\rm P}_{22}$ mode all the way to the $h^{\rm I}_{2,-2}$ mode) have larger phase evolutions, but smaller amplitudes.
+
+The relations~\eqref{eq:wignerrot} do not mix different values of $\ell$.
+Furthermore, rotations leave invariant the combined square amplitude $\calA_{\ell}$ for each $\ell$, as well as the total square amplitude:
+\be
+	\calA^{2} = \sum\limits_{\ell \geq 2}\calA_{\ell}^{2} = \sum\limits_{\ell \geq 2}\sum\limits_{m=-\ell}^{\ell} |h_{\ell m}|^{2} \,.
+\ee
+One can therefore use these amplitudes (in our case, limited to $\ell = 2$) to define a frame-independent peak amplitude of the precessing waveform.
+
+Several constructions for the precessing frame have been proposed~\cite{} \SM{[C]}.
+In the following, we will use a co-precessing frame based on the dominant eigenvector of the matrix representing the action of the angular momentum operator acting on the waveform modes, proposed in~\cite{} \SM{[C]} (see App.~\ref{app:wigner} for more details).
+
+When approximating the precessing-frame waveform by a non-precessing one, one can use the assumption (exactly valid for non-precessing systems, see~\eqref{eq:symmetryhlminusm} \SM{[R]})
+\be\label{eq:approxsymmetryhlminusm}
+	h^{\rm P}_{\ell,-m} \simeq (-1)^{\ell} h^{\rm P*}_{\ell,m} \,.
+\ee
+Similarly, for Fourier-domain models, one can neglect either the negative or positive frequency band in the Fourier transform of precessing-frame modes (as in~\eqref{eq:zeronegativef})
+\begin{align}\label{eq:approxzeronegativef}
+	\tilde{h}_{\ell m}^{\rm P} (f) &\simeq 0 \text{ for } f<0, \; m>0 \nn\,,\\
+	\tilde{h}_{\ell m}^{\rm P} (f) &\simeq 0 \text{ for } f>0, \; m<0 \,,
+\end{align}
+and neglect altogether the $m=0$ modes, $\tilde{h}_{\ell 0}^{\rm P} (f) \simeq 0$.
+
+We will introduce mode-by-mode transfer functions $\calT^{\ell}_{mm'}$,
+\be\label{eq:defprectransfer}
+	\mathrm{FT} \left[ \calD^{\ell *}_{mm'} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm P} \right] (f) \equiv \calT^{\ell}_{mm'}(f) \tilde{h}_{\ell m'}^{\rm P} (f)
+\ee
+such that the complete Fourier-domain inertial-frame waveform will be given as the sum of these individual mode contributions,
+\be\label{eq:defprechIsum}
+	\tilde{h}_{\ell m}^{\rm I} (f) = \sum\limits_{m'=-\ell}^{\ell} \calT^{\ell}_{mm'}(f) \tilde{h}_{\ell m'}^{\rm P} (f) \,.
+\ee
+
+When using the assumptions~\eqref{eq:approxsymmetryhlminusm}-\eqref{eq:approxzeronegativef}, one can derive a symmetry relation in the transfer functions themselves.
+From the explicit expression of the Wigner matrices, we have indeed (see App.~\ref{app:wigner})
+\be
+	\calD^{\ell *}_{-m,-m'} = (-1)^{m+m'}\calD^{\ell}_{mm'} \,.
+\ee
+Since for a function $g$ we have in general for its conjugate
+\be
+	\mathrm{FT}[g^{*}](f) = \tilde{g}(-f)^{*} \,,
+\ee
+we can write, using~\eqref{eq:approxsymmetryhlminusm},
+\be
+	\mathrm{FT}[\calD^{\ell *}_{mm'} h_{\ell m'}^{\rm P}](-f)^{*} = (-1)^{\ell+m+m'} \mathrm{FT}[\calD^{\ell *}_{-m,-m'} h_{\ell, -m'}^{\rm P}](f) \,,
+\ee
+or, for transfer functions,
+\be
+	\calT^{\ell}_{mm'}(f) = (-1)^{m+m'} \calT^{\ell *}_{-m,-m'}(-f) \,.
+\ee
+This means that such a model is required to cover only the positive frequency band and the values $m'>0$, the negative frequencies being deduced from
+\begin{align}
+	\tilde{h}^{\rm I}_{\ell m}(f) &= \sum_{m'>0} \calT^{\ell}_{mm'}(f) \tilde{h}^{\rm P}_{\ell, m'}(f) \text{ for } f>0 \,, \nn\\
+	\tilde{h}^{\rm I}_{\ell m}(f) &= \sum_{m'>0} (-1)^{\ell+m+m'} \calT^{\ell *}_{-m,m'}(-f) \tilde{h}^{\rm P *}_{\ell, m'}(-f) \text{ for } f<0 \,.
+\end{align}
+When including only the leading harmonic $h^{\rm P}_{22}$ in the precessing-frame waveform (as is done for instance in PhenomP~\cite{} \SM{[C]}), we have for $f<0$
+\begin{subequations}
+\begin{align}
+	\calT^{2}_{m,-2}(f) &= (-1)^{m} \calT^{\ell *}_{-m,2}(-f) \,, \\
+	\tilde{h}^{\rm I}_{2 m}(f) &= (-1)^{m} \tilde{h}^{\rm I *}_{2,-m}(-f) \,. \label{eq:symmetryhIfor22only}
+\end{align}
+\end{subequations}
+In general, when including a superposition of modes with different values of $m'$ in the model (for instance, the modes $h^{\rm P}_{22}$ and $h^{\rm P}_{21}$), this symmetry exists only at the level of the mode contributions, and there is no relation such as~\eqref{eq:symmetryhIfor22only}  at the level of $\tilde{h}^{\rm I}_{\ell m}$, as the complete mode reconstruction includes factors $(-1)^{\ell+m+m'}$ in the sum.
+
+In the notations of our formalism of Sec.~\ref{sec:formalism}, 
+\be\label{eq:defmodulationprec}
+	s(t) = F(t) h(t) \,,
+\ee
+with $s$ being a mode contribution to $h^{\rm I}_{\ell m}$, $h$ one of the precessing-frame modes $h^{\rm P}_{\ell m'}$, and the modulation $F$ being a Wigner matrix $\calD^{\ell *}_{mm'}$.
+In the absence of delays, the convolution with a frequency-dependent kernel~\eqref{eq:FDkernel} becomes a simple convolution in the Fourier-domain,
+\be\label{eq:convolution}
+	\tilde{s} (f) = \int \ud f' \, \tilde{h}(f-f') \tilde{F} (f') \,.
+\ee
+Our objective is therefore to compute these convolution integrals to obtain the Fourier-domain transfer functions $\calT^{\ell}_{mm'}$.
+We note an important point about the structure of these comvolution integrals.
+In presence of the usual behaviour of precession around an almost constant $\bm{J}$, $\alpha$ is monotonously increasing and the modulation functions have the phase evolution $\calD^{\ell *}_{mm'} (\alpha, \beta, \gamma) \propto e^{i(m' \cos\beta - m) \alpha}$.
+For the 22 mode in the P-frame, $m'=2$, the coefficient of $\alpha$ is close to $0$ for small $\beta$ for $m=2$, and increasingly positive as we go down from $m=1$ to $m=-2$.
+In our Fourier convention~\eqref{eq:defFT}, the Fourier transforms $\tilde{F}$ thus have most of their support on negative frequencies.
+This means that the support of the convolution integral~\eqref{eq:convolution} extends to the right side for $\tilde{h}$, i.e.
+$f-f' > f$.
+This point will become important when considering the high-frequency part of the signal.
+
+\subsection{Previous approaches}
+\label{subsec:precpreviousapproaches}
+
+The above decomposition~\eqref{eq:wignerrot}, and simplifying assumptions~\eqref{eq:approxsymmetryhlminusm}-\eqref{eq:approxzeronegativef} for the precessing-frame waveform, have been widely used for the existing models of waveforms from precessing binaries.
+For the inspiral phase of the SEOBNRv3 model~\cite{} \SM{[C]}, the frame is assumed to follow the instantaneous orbital plane and the waveform in this frame is represented by a spin-aligned expression, with spin-aligned components that are left to vary according to the precessing dynamics.
+The ringdown phase of the signal is treated differently, the waveform being first rotated back to an inertial frame and then matched to a superposition of quasi-normal modes in this final frame.
+The PhenomP model~\cite{} \SM{[C]} uses the decomposition~\eqref{eq:wignerrot} above (although directly in the Fourier domain, as we will detail below) with a precessing-frame waveform given by a spin-aligned PhenomD waveform~\cite{} \SM{[C]}.
+The SUA waveforms~\cite{} \SM{[C]}, focused on the inspiral phase, also use the decomposition~\eqref{eq:wignerrot} and a post-Newtonian non-precessing model for the modes $h_{\ell m}^{\rm P}$.
+
+To account for previous approaches to the problem of producing precessing waveforms directly in the Fourier domain, avoiding the use of a FFT \SM{[check acronym]}, it is useful to introduce schematically three levels of approximation (dropping the mode indices):
+\begin{itemize}
+	\item unstable SPA: $\calT(f) \tilde{h}^{\rm P}(f) = \mathrm{SPA}\left[ \calD^{*} h^{\rm P} \right](f)$,
+	\item SPA, $0^{\text{th}}$-order SUA, PhenomP: $\calT(f) = \calD^{*}(t_{f}^{\rm SPA}) $,
+	\item SUA: $\calT(f) = \frac{1}{2} \sum\limits_{k} a_{k} \calD^{*}(t_{f}^{\rm SPA} \pm k T_{f}^{\rm SPA})$.
+\end{itemize}
+
+In the first option, one applies directly the SPA, as described in Sec.~\ref{subsec:SPA}, to the product of the modulation and the signal.
+This is known~\cite{} \SM{[C]} to lead to possible pathologies, since the prefactor $1/\sqrt{\ddot{\phi}_{\rm prec}}$ can blow up due the precessing contributions to the phase, and is not used in waveform modelling.
+
+The second option corresponds to simply multiplying the Fourier-domain signal by the modulation function evaluated at the time $t_{f}^{\rm SPA}$ as defined in~\eqref{eq:deftfSPA}.
+In the SUA formalism of Ref.~\cite{} \SM{[C]}, this amounts to using a stencil reduced to a single point, i.e.
+using $a_{0,0} = 1$ in~\eqref{} \SM{[R]}.
+It is equivalent to the leading order of our formalism, as explained in Sec.~\ref{} \SM{[R]}, with the difference that we use a more general definition of $t_{f}$ (see~\eqref{eq:deftf}) that extends through the merger and ringdown.
+\SM{[Mention O'Shaugnessy single-spin FD waveforms]}.
+The PhenomP model~\cite{} \SM{[C]} implictly uses this level of approximation, although with the notable difference that a frequency-based expression of the modulation is used, allowing to cover the whole frequency band.
+The precessing-frame evolution is modelled using post-Newtonian expressions for the Euler angles $(\alpha^{\rm PN}, \beta^{\rm PN}, \gamma^{\rm PN}) (\omega)$, that are obtained as functions of the orbital frequency $\omega$ in the limit of a small opening angle of the precession cone and at the spin-orbit level~\cite{} \SM{[C]}.
+Using the SPA-inspired correspondence~\eqref{} \SM{[R]} between the orbital and Fourier-domain frequency, one then writes
+\be\label{eq:precPhenomP}
+	\calT^{\ell}_{mm'}(f) = \calD^{\ell *}_{mm'}(\alpha^{\rm PN}, \beta^{\rm PN}, \gamma^{\rm PN}) \left( \omega=\pi f\right) \,.
+\ee
+When the SPA is valid for the underlying signal $h^{\rm P}_{\ell m}$, the condition $\omega = \pi f$ is equivalent by definition to $t=t^{\rm SPA}_{f}$, but writing things in terms of frequency gives an effective prescription that can cover the whole Fourier-domain frequency band, including the merger and ringdown.
+Note that this procedure amounts to using PN \SM{[check acronym]} results outside of their range of validity, since in the merger-ringdown phase both the PN perturbative treatment and the SPA approximation break down.
+In particular, seen as a function of time, $\omega^{\rm PN}(t)$ blows up at merger.
+In practice, one finds however that the frequency-based expressions~\eqref{eq:precPhenomP} are mildly varying as a function of frequency, and the agreement of the model with numerical relativity waveforms is good~\cite{} \SM{[C]}.
+
+In the SUA formalism proposed in~\cite{} \SM{[C]}, one extends the SPA to incorporate the leading-order correction during the inspiral phase.
+The transfer function is then computed by evaluating the modulation on a stencil of times centered around the time-of-frequency given by the SPA, following
+\begin{widetext}
+\be
+	\calT(f) = \frac{1}{2} \sum\limits_{k=0}^{N} a_{N,k} \left[\calD^{\ell *}_{mm'}(t_{f}^{\rm SPA} + k T_{f}^{\rm SPA}) + \calD^{\ell *}_{mm'}(t_{f}^{\rm SPA} - k T_{f}^{\rm SPA}) \right] \,.
+\ee
+\end{widetext}
+The definition of $T_{f}^{\rm SPA}$ was given in~\ref{} \SM{[R]}.
+As described in Sec.~\ref{} \SM{[R]}, this approach is equivalent to our formalism when keeping only the quadratic phase correction, but is a priori limited to the inspiral phase of the signal through the definitions of the times $t^{\rm SPA}_{f}$ and $T^{\rm SPA}_{f}$.
+For inspiral waveforms, Ref.~\cite{} \SM{[C]} showed that this treatment improves the accuracy with respect to the local approximation.
+
+The time-domain SEOB~\cite{} \SM{[C]} model includes the modes $h^{\rm P}_{2m}$ with $|m|=2$ and $|m|=1$ for the inspiral, and the ringdown constructed in the inertial frame contains a superposition of all modes $h^{\rm I}_{2 m}$ by construction.
+In the PhenomP~\cite{} \SM{[C]} model, only the dominant harmonics $h^{\rm P}_{22}$ and $h^{\rm P}_{2,-2}$ were included in the precessing-frame waveform.
+The SUA~\cite{} \SM{[C]} model can be constructed using any PN non-precessing model for an inspiral-only waveform, including higher harmonics.
+\SM{[Check what is implemented in LAL ?
+HM or not ?]}
+
+\subsection{Simplified model for precessing IMR waveforms}
 \label{subsec:precmodel}
 
-[explain inertial frame/precessing frame]
-[explain Gatech formula for post-merger precession]
-
-Here we will only consider the inertial-frame modes induced by the precessing-frame dominant 22 mode.
-
-[introduce phenomP]
-[explain model with smoothing for the angles with its two versions]
-[qualify how bad the model is with respect to full NR]
-
-In the following, since our purpose here is not to provide a full-fledged model for precessing signals in the Fourier domain, we will not explore the full parameter space and we will limit our analysis to a simplified toy model for precession.
-Namely, we will approximate the Euler angles during the inspiral phase by using the PN expressions used in PhenomP~\cite{Hannam+13}, and we will adopt two alternatives asymptotic behaviours post-merger.
-In the first one (thereafter Case I), we simply send smoothly all three Euler angles to constants.
-In the second one (Case II), we send the opening angle of the precession cone $\beta$ to a constant, but we reproduce the behaviour describe letting the frame rotate around the direction of the final total angular momentum $\hat{\bm{J}}_{\rm final}$ with an angular velocity of $\Omega_{\rm frame} = \omega_{220} - \omega_{210}$.
-Fig.~\ref{fig:prectoymodel} shows the Euler angles evolution for Cases I and II.
-
-\begin{figure}
+\SM{[qualify how bad the model is with respect to full NR]}
+
+\SM{[explain shortcomings of pure SEOB in post-merger, problem of smoothness]}
+
+We now present our toy model for IMR waveforms of precessing binaries, which we will use in the following to illustrate the accuracy of the Fourier-domain transfer functions at various orders of approximation.
+Since our purpose in this article is not to propose a full-fledged waveform model, but rather to investigate wether the formalism of Sec.~\ref{sec:formalism} is relevant to tackle the merger part of the signals, and wether its application reduces the errors in the reconstruction of transfer functions, we will allow for a number of idealizations and simplifications in this toy model.
+
+Although the frame decomposition and precessing-frame prescription is well understood for the inspiral, where the precessing frame roughly follows the orbital plane of the binary (see~\ref{} \SM{[R]}), we wish to investigate the merger and ringdown part of the signal as well, and need a way to extend this intuitive picture.
+
+\SM{[acronym QNM]}
+
+For this extension, one needs to define a precessing frame after merger, when the orbital plane has lost any meaning.
+This can be done naturally when using a frame prescription that is not tied to the dynamics but directly extracted from the gravitational waveform, such as~\eqref{} \SM{[R]}.
+Investigating the post-merger behaviour of this frame requires numerical relativity simulations of precessing binaries.
+Ref.~\cite{} \SM{[C]} proposed a qualitative model where the frame vector essentially keeps precessing around the final $\bm{J}$, but transitions to a faster precession rate.
+More precisely, they observe that the interpretation of the post-merger waveform as an excitation of a superposition of quasinormal modes suggests that the frame should evolve according to the difference of the leading QNM frequencies.
+In their simulations, they find indeed a precession rate around the final total angular momentum $\hat{\bm{J}}_{\rm final}$, asymptoting to the angular velocity
+\be\label{eq:OmegaframeQNM}
+	\Omega_{\rm frame} = \omega_{220}^{\rm QNM} - \omega_{210}^{\rm QNM} \,.
+\ee
+The behaviour of the opening angle of the precessing cone $\beta$, however, is harder to interpret.
+Although it usually shows some transition at merger (to a lower value), it does not show an exponential decay as the interpretation above would suggest.
+We found that overall this picture is at least qualitatively correct for the precessing waveforms	 currently available in the SXS catalog.
+\SM{[allowed to make such a vague statement about catalogue waveforms ?]}
+
+Note however that this model should be taken with caution, since the parameter space currently accessible to numerical simulations of precessing systems is limited in spin mass ratio.
+The behaviour of systems with a close-to-extremal remnant spin largely remains to be explored.
+In addition, systems with large mass ratios and large anti-aligned spins on the primary black hole can have a total angular momentum dominated by the primary spin instead of the orbital angular momentum (e.g.
+in the case of transitional precession~\cite{} \SM{[C]}).
+This can lead to the remnant spin being anti-aligned with respect to the precessing frame axis prior to merger.
+This could conceivably lead to a different structure in the excitation of the quasi-nomal modes, and to a qualitatively different frame behaviour.
+Assessing the limits of validity of this model would require a more extensive coverage of the parameter space by numerical relativity simulations and would go beyond the scope our present study.
+
+The basic ingredients of our model \SM{[apologize for the Frankenstein-ism !]} will be:
+\begin{itemize}
+	\item PhenomD~\cite{} \SM{[C]} for the precessing-frame waveform;
+	\item SEOB for Euler angles during the inspiral;
+	\item an effective extension of the Euler angles post-merger based on~\eqref{eq:OmegaframeQNM}.
+\end{itemize}
+For the precessing-frame waveform, in keeping with the approximation~\eqref{} \SM{[R]} for the precessing-frame waveform we will use a slightly modified version of the PhenomD waveform model~\cite{} \SM{[C]}, a Fourier-domain model that covers the inpiral, merger and ringdown for binaries with generic aligned spins.
+The waveform is produced as piecewise analytical expressions\footnote{The PhenomD model is publicly available in the LIGO Algortitm Library (LAL).}.
+We limit ourselves to the dominant harmonics $h^{\rm P}_{22}$ and $h^{\rm P}_{2,-2}$ (see~\cite{} \SM{[C]} for a recent extension of the model to include higher harmonics).
+Using a Fourier-domain approximant that is smooth by construction avoids the problems arising from the tapering time-domain waveforms, as we will detail below, and will allow us to easily take Fourier-domain derivatives of the waveform.
+Since the PhenomD amplitude and phase are only of class $C^{1}$, to avoid spurious discontinuities we introduce decaying corrective terms of the form $(f-f_{\rm join})^{2} e^{-\lambda(f-f_{\rm join})^{2}}$ on each side of $f_{\rm join}$, the junction frequency between two bands.
+Here $\lambda$ is a large parameter ensuring that the modification remains local.
+The modificiations have very little effect overall but artificially enforce the continuity of second derivatives.
+
+For the frame trajectory in the inspiral phase, we extract the dominant-eigenvector frame as defined in~\eqref{} \SM{[R]} from an SEOB waveform.
+Since the SEOB waveform is constructed by using the instantaneous orbital plane of the dynamics as the precessing frame, the frame extraction procedure simply gives back the orbital plane (except very close to merger where the waveform starts to smoothly transition to the post-merger QNM superposition).
+The frame thus presents oscillatory features of small amplitude at the orbital timescale, due to the nutation of the orbital plane \cite{} \SM{[C]}.
+
+For the extension of the frame trajectory post-merger, we do not rely on the PhenomP procedure as it is merely effective.
+Furthermore, we found that the ringdown attachment procedure, as implemented in SEOB, leads to frame trajectories that show oscillations in all the Euler angles when the spins are anti-aligned.
+These oscillations increase in amplitude as the spins go to larger misalignments, and are already present well beneath the regime of transitional precession.
+Since we checked that the qualitative behaviour~\eqref{eq:OmegaframeQNM} holds rather well for numerical relativity simulations in regions of the parameter space where SEOB starts to show these oscillations, and since the SEOB post-merger ringdown attachment is an effective representation, we chose to use an artificial post-merger extension of the frame, following
+\be\label{eq:alphaQNM}
+	\alpha_{\rm post-merger}(t) = \alpha(t_{\rm peak}) + (\omega_{220}^{\rm QNM} - \omega_{210}^{\rm QNM})(t-t_{\rm peak}) \,.
+\ee
+Since we do not have yet a quantitative model of the behaviour of the opening angle of the precessing cone $\beta$, we simply set it to a constant value post-merger, taken to be its value at merger.
+The third Euler angle $\gamma$ is extended using the minimal rotation condition~\eqref{} \SM{[R]}.
+For the QNM frequencies as a function of the final spin, we use values tabulated in Ref.~\cite{} \SM{[C]}.
+The final spin is taken to be the same as in the SEOB code~\cite{} \SM{[C]}, where the final-spin fit formula of Ref.~\cite{} \SM{[C]}, built for spin-aligned systems, is applied to the spin components \SM{[check that and complete]}.
+Our simple extension is an idealization, especially for the angle $\beta$, and is not to be considered quantitatively accurate.
+However, it allows us to capture what we believe to be the dominant post-merger behaviour, in the sense that the frame is set to rotate with a velocity of the right order of magnitude, which is crucial for us since we want to investigate how the separation between the orbital (or $h^{\rm P}_{\ell m}$) timescale and the precessing timescale is challenged post-merger.
+We show in Fig~\ref{} the frequencies $\omega_{220}^{\rm QNM}$, $\omega_{210}^{\rm QNM}$ and their difference $\Omega_{\rm frame}$.
+The behaviour for near-to-extremal spin of the remnant would require more exploration, as the QNM frequencies vary a lot when approaching extramality.
+\SM{[add plot of QNM frequency and frequency diff according to final spin]}.
+Fig.~\ref{fig:prectoymodel} presents a comparison of the frame trajectory at merger for the SEOB, NR waveforms and for our model.
+The example catalog waveform is SXS:BBH:0058~\cite{} \SM{[C]}, which presents a rather high mass ratio of $q=5$, with a single spin in-plane of $\chi_{1} = 0.5$.
+The NR and SEOB post-merger frames disagree mainly due to a different value of the final spin $\chi_{f}$\footnote{Work is underway to update the fit formulas used to compute $\chi_{f}$ in the SEOB code, which is expected to reduce these disagreement~\cite{}.
+\SM{[C]} \SM{[work in preparation by SEOB people, essentially already done]}}, which leads to a different $\Omega_{\rm frame}$.
+Given their different value of $\chi_{f}$, however, both agree with the qualitative prescription~\eqref{eq:OmegaframeQNM}.
+The oscillations developing after $t\sim 50 M$ occur in a regime where the overall waveform amplitude has decayed to small values, and are not important in that regard.
+Our model follows well the SEOB behaviour for $\alpha$ and $\gamma$, but departs more for $\beta$ as it ignores its variation at merger.
+
+To combine our various waveform ingredients, we first ensure that the PhenomD and SEOB waveforms are aligned in time, by performing an inverse Fourier transform of PhenomD and applying a time shift to both waveforms so that the peak amplitude occurs at $t=0$ (taken for SEOB to be the $\ell=2$ combined square amplitude~\eqref{}, and for PhenomD just the peak of the 22 mode for PhenomD).
+We apply to the resulting frame trajectory a Gaussian filtering following the same procedure as the one used for numerical-relativity surrogate of Ref.~\cite{} \SM{[]}: the scale of the filter is set to a fixed width in waveform phase \SM{[detail what we used exactly]}.
+This filtering ensures a continuity of the modulation function and their derivatives at the point of attachment.
+It also eliminates the nutation of the frame at the orbital timescale, which are generically less pronounced in the frame extracted following~\eqref{} \SM{[R]} from numerical relativity simulations than for the normal to the orbital plane.
+
+Note an important caveat of our approach: the application of the perturbative formalism presented in Sec.~\ref{sec:formalism} relies on the smoothness of both the time-domain modulation and the Fourier-domain signal.
+For the time-domain modulation, the assumption of the separation of timescales~\eqref{} \SM{[R]} relies on the identification of a characeteristic frequency $\Omega_{\rm prec}$, which is based on the time derivative of $F$.
+It is always possible to add and remove arbitrarily small oscillations in $F$ that would give arbitrarily large contributions to its derivatives, hence affecting any local computation of $\Omega_{\rm prec}$.
+One can consider that our approach is to be applied to such a slowly varying approximation of $F$, that can be obtained by filtering away rapidly-varying, small-amplitude features.
+However, we should keep in mind that any smoothing procedure could possibly eliminate  some information present in the original waveforms.
+
+For the smoothness of the signal in the Fourier domain, one has to keep in mind that, for waveforms computed starting with a time-domain waveform, Gibbs oscillations can be mitigated by properly tapering the waveform in time domain before performing the FFT, however taking Fourier-domain derivatives magnifies this oscillations, to the point that they would dominate among others the timescale $T_{f}$ (unless one takes much longer waveforms to push this problem to lower frequencies).
+This is avoided in our case by working with a pure Fourier-domain PhenomD waveform that is smooth by construction, but would become important if one were to apply our formalism to the FFT of NR or SEOB waveforms.
+
+Another caveat is that for the precessing-frame waveform, the assumption~\eqref{eq:approxsymmetryhlminusm} is only approximate.
+Mode asymmetries are present in the late inspiral and merger phases of precessing binaries, as shown by numerical relativity simulations (see e.g.~\cite{} \SM{[C]}).
+These asymmetries play a role in the emission of linear momentum and in the resulting kick of the black hole remnant, and they also result in amplitude and phase variations on the orbital timescale.
+
+In the following, we will illustrate our formalism by applying it to three examples.
+The parameters of these example cases are summarized in Table~\ref{table:precparams}.
+We consider a mass ratio of $q=4$, close-to-maximal spins of $\chi_{1} = \chi_{2} = 0.95$, and vary both spins misalignmnent angles to be $\pi/6$ (close to spin-aligned), $\pi/2$ (spins in the plane) and $5\pi/6$ (close to anti-aligned spins).
+Table~\ref{table:precparams} also shows the remnant spins that are internally computed in both the SEOB and the PhenomD/PhenomP codes, as well as the QNM frequencies and the resulting frame precession frequency $\Omega_{\rm frame}$.
+Note that the case $--$ shows a larger variation in the direction of $\bm{J}$, and the spin of the remnant is almost 0, due to the fact that the spin is large and antialigned, for a large mass ratio.
+This example therefore serves the purpose of lying at the edge of validity of the picture of standard precession along a cone around an almost-fixed direction, and the standard hierarchy between mode contributions is not valid in this case.
+\SM{[soundness of trajectories, particularly in this case, are to be checked again]}
+
+\begin{figure*}
   \centering
-  \includegraphics[width=.48\linewidth]{plots/prectoymodelalphagamma.pdf}
+  \includegraphics[width=.45\linewidth]{plots/eulerSXS0058alphagamma.pdf}
   \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/prectoymodelbeta.pdf}
-  \caption{Evolution of Euler angles $(\alpha, \beta, \gamma)$ for our two toy models.
+  \includegraphics[width=.45\linewidth]{plots/eulerSXS0058beta.pdf}
+  \caption{Evolution of Euler angles $(\alpha, \beta, \gamma)$ near merger for the example waveform SXS:BBH:0058~\cite{} \SM{[C]}.
+Full, dashed and dotted lines represent the NR, SEOB and extended SEOB waveforms respectively.
+The vertical line at $t=0$ indicates the time of merger.
+In the left panel the full and dashed black lines indicate the asymptotic behaviour~\eqref{eq:OmegaframeQNM} for the rotation of the frame around the direction of the final $\bm{J}$.
+The difference between NR and SEOB there is due to the value of the final spin, with $\chi_{f}^{\rm NR} = 0.54$ and $\chi_{f}^{\rm SEOB} = 0.42$, which leads to a different $\Omega_{\rm frame}$.
+In the right panel, $\beta$ is asymptotically constant in our toy model.
+\SM{[Add amplitude/$\omega$ plots comparing NR/SEOB, and possibly PhenomD for the 22 mode ?
+Add Euler angles produced by IFFT of PhenomP ?]}}
+  \label{fig:prectoymodel}
+\end{figure*}
+
+%\begin{figure*}
+%  \centering
+%  \includegraphics[width=.48\linewidth]{plots/prectoymodelalphagamma.pdf}
+%  \hspace{0.2cm}
+%  \includegraphics[width=.48\linewidth]{plots/prectoymodelbeta.pdf}
+%  \caption{Evolution of Euler angles $(\alpha, \beta, \gamma)$ for our two toy models.
 Dashed lines represent Case I and solid lines Case II.
@@
 In the right panel, $\beta$ is asymptotically constant.}
-  \label{fig:prectoymodel}
-\end{figure}
-
-These simplifications comes with obvious limitations.
-Notably, they produce modulation functions $F(t)$ that are very smooth by construction.
-However, these two toy models will allow us to explore the separation of timescales in the merger-ringdown part of the signal, and to highlight how the perturbative analysis can be challenged by the rapid rotation of the radiation frame after merger, even in this idealized model.
-
-In the absence of delays, we have a simple modulation of the signal by a function $F(t)$, as given in~\eqref{eq:defmodulationprec} above.
-The convolution with a frequency-dependent kernel~\eqref{eq:FDkernel} becomes a simple convolution in the Fourier-domain,
-\be\label{eq:convolution}
-	\tilde{s} (f) = \int \ud f' \, \tilde{h}(f-f') \tilde{F} (f') \,.
-\ee
-
-It will be convenient to use the notation
-\be\label{eq:defFlmlmp}
-	F_{\ell m, \ell m'} \equiv D^{\ell *}_{m'm} (\alpha,\beta,\gamma)
-\ee
-which represents the modulation function corresponding to the contribution of each precessing-frame mode $h^{P}_{\ell m}$ mode to the inertial-frame mode $h^{I}_{\ell m'}$ (see Eq.~\eqref{eq:wignerrot}).
-One can also define Fourier-domain tranfer functions for each mode contribution, according to
-\be\label{eq:defTlmlmp}
-	\mathrm{FT} \left[F_{\ell m, \ell m'} h \right] (f) \equiv \calT_{\ell m, \ell m'} (f) \tilde{h}(f) \,,
-\ee
-where $\mathrm{FT}$ stands for the Fourier transform.
-Fig.~\ref{fig:prectransfer} gives the amplitude and phase of the transfer functions $\calT_{22, 2 m'}$ for the two cases I and II.
-As expected, the difference between the two toy models for the post-merger frame precession is localized to the high frequency range.
-One can see a clear hierarchy in the amplitude of the inertial-frame modes generated by the precessing-frame 22 mode.
-This hierarchy come from the Wigner geometric factors and is dependent on the opening angle of the precession cone, or second Euler angle $\beta$, which in turn is dependent on the magnitude of the misaligned spins in the system.
-Case II shows stronger amplitude features in the merger and ringdown.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.48\linewidth]{plots/prectransferAcaseI.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/prectransferPsicaseI.pdf}
-  %
-  \includegraphics[width=.48\linewidth]{plots/prectransferAcaseII.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/prectransferPsicaseII.pdf}  
-  \caption{Amplitude and phase of the Fourier-domain transfer fonctions $\calT_{22,2m'}$ as defined in~\eqref{eq:defTlmlmp} for individual mode contributions.
+%  \label{fig:prectoymodel}
+%\end{figure*}
+
+\SM{[Check the mismatch for $--$ between Jf from dynamics and chif used in SEOB internals to do the ringdown attachment - check that the frame trajectory makes sense by comparing to orbital plane.]}
+
+%In Eq.~\eqref{eq:defmodulationprec} above, the modulation function $F$ evolves on the precessional timescale, which evolves throughout the inspiral.
+The separation between the precessional timescale and the orbital timescale will be crucial for our analysis.
+In the limit of low frequencies, we will see in Sec.~\ref{} below that, although these timescales become more and more separated, the decrease in the chirping rate gives raise to a corrective contribution that does not vanish in this limit.
+In the other limit, as the system gets close to merger, the separation of timescales becomes weaker, and we will explore in Sec.~\ref{} the application of our formalism to the merger and ringdown phase.
+
+\begin{table}[t]
+\begin{ruledtabular}\caption{Parameters of of the three example cases that we use to illustrate our formalism.
+The three cases differ by their spin alignment angles $\theta_{A} = \bm{\chi}_{A} \cdot \hat{\bm{L}}_{i}$, the spins being in the orbital plane, almost aligned and almost anti-aligned.
+The table gives also the final spins magnitudes (for SEOB as well as PhenomD for comparison), angles between the initial and final angular momenta $\theta(\bm{J}_{i}, \bm{J}_{f})$, as well as QNM frequencies and frame rotation velocity used for the post-merger extension.
+\SM{[To be updated with longer waveforms, 10Hz 10Msol.]} \SM{[add misalignment angles $(\ell, \bm{J}_{i})$]}}\label{table:precparams}
+%\begin{tabular}{C{.2\columnwidth}C{.24\columnwidth}C{.15\columnwidth}}
+\begin{tabular}{ccccccc}\label{table:precexamples}
+	$f_{\rm min}$ & $ M_{\rm min} $ & $q$ & $\chi_{1}$ & $\chi_{2}$ & $ \phi_{1} $ & $ \phi_{2} $ \\
+	\hline
+	$20\mathrm{Hz}$ & $20\Msol$ & $ 4.0 $ & $ 0.95 $ & $ 0.95 $ & $0$ & $\pi/2$ \\
+	\hline\hline
+	Case && ++ && $pp$ && $--$ \\
+	\hline
+	$\theta_{1}$ && $\pi/6$ && $\pi/2$ && $5\pi/6$ \\
+	$\theta_{2}$ && $\pi/6$ && $\pi/2$ && $5\pi/6$ \\
+	\hline
+	$\chi_{f}$ && $0.89$ && $0.50$ && $0.02$ \\
+	$\chi_{f}^{\rm Ph}$ && $0.90$ && $0.47$ && $0.002$ \\
+	$\theta(\bm{J}_{i}, \bm{J}_{f})$ && $0.02$ && $0.04$ && $0.27$ \\
+	$M \omega_{220}^{\rm QNM}$ && $0.66$ && $0.47$ && $0.377$ \\
+	$M \omega_{210}^{\rm QNM}$ && $0.51$ && $0.42$ && $0.375$ \\
+	$M \Omega_{\rm frame}$ && $0.15$ && $0.04$ && $0.002$ \\
+\end{tabular}
+\end{ruledtabular}
+\end{table}
+
+%\begin{table}[t]
+%\begin{ruledtabular}\caption{Parameters of the three example cases that we use to illustrate our formalism.}\label{table:precparams}
+%%\begin{tabular}{C{.2\columnwidth}C{.24\columnwidth}C{.15\columnwidth}}
+%\begin{tabular}{cccc}
+%	Case & pp & $++$ & $--$ \\
+%	\hline
+%	$q$ & $4.0$ & $4.0$ & $4.0$ \\
+%	$\chi_{1}$ & $0.95$ & $0.95$ & $0.95$ \\
+%	$\chi_{2}$ & $0.95$ & $0.95$ & $0.95$ \\
+%	$\theta_{1}$ & $\pi/2$ & $\pi/6$ & $5\pi/6$ \\
+%	$\theta_{2}$ & $\pi/2$ & $\pi/6$ & $5\pi/6$ \\
+%	$\phi_{2}$ & $0.0$ & $0.0$ & $0.0$ \\
+%	$\phi_{2}$ & $\pi/2$ & $\pi/2$ & $\pi/2$
+%\end{tabular}
+%\end{ruledtabular}
+%\end{table}
+
+\SM{[We could add plots of the 3 frame trajectories, and of the FD transfer functions in Amp/Phase form.
+But the trajectories are not very informative except for - - (where our model might not be very solid) and for showing oscillations in $\beta$.
+The reconstruction errors plots already show the amplitude of the modes themselves, giving an idea of the hierarchy; qhat is the missing is the phase in Tlmmp, which is not very informative.]}
+
+%\begin{figure*}
+%  \centering
+%  \includegraphics[width=.48\linewidth]{plots/prectransferAcaseI.pdf}
+%  \hspace{0.2cm}
+%  \includegraphics[width=.48\linewidth]{plots/prectransferPsicaseI.pdf}
+%  %
+%  \includegraphics[width=.48\linewidth]{plots/prectransferAcaseII.pdf}
+%  \hspace{0.2cm}
+%  \includegraphics[width=.48\linewidth]{plots/prectransferPsicaseII.pdf}  
+%  \caption{Amplitude and phase of the Fourier-domain transfer fonctions $\calT_{22,2m'}$ as defined in~\eqref{eq:defTlmlmp} for individual mode contributions.
 The left panels show the amplitude and the right panels the phase.}
-  \label{fig:prectransfer}
-\end{figure}
+%  \label{fig:prectransfer}
+%\end{figure*}
 
@@
 
-We investigate now the post-Newtonian order counting of the expansion~\eqref{eq:resultdffPsiTf} during the inspiralling phase of a precessing binary, with modulation $F$ as in~\eqref{}.
-Using the notation $\calO(n) = \calO(c^{-n})$, we have for quasi-cirular inspirals that $\dot{\omega} = \calO(5)$, corresponding to the radiation reaction effect at the 2.5PN order.
-The mode mixing will be induced by the rotation of the orbital angular momentum $\bm{L}$ around the total angular momentum $\bm{J}$, with the angular velocity for this precession $\bm{\Omega} = \calO(2)$.
-The relevant scaling for the terms in the series~\eqref{eq:resultdffPsiTf}, for the inspiral phase where the SPA applies, is therefore
-\be
-	\Tf^{2p} \partial_{t}^{2p} F = \calO(-1) \,,
-\ee
-meaning that the expansion becomes {\it less} well behaved for {\it lower} frequencies.
-The presence of a $1/p!$ gives a formal sense to the expansion if the successive derivatives of $F(t)$ are appropriately bounded, but one may have to keep a large number of terms in the sum.
-The approximation that consists in taking the first term in the series, the LLP approximation~\eqref{} which was in particular used for the PhenomP~\cite{} waveform model, gets {\it worse} the further away from merger the signal is.
-We investigate further in Sec.~\ref{} the validity of the treatment of precession used for PhenomP.
-
-Notice that, for the amplitude derivatives in~\eqref{eq:expandA} and the higher order derivatives of the phase in~\eqref{eq:expandPsi}, the same PN scaling arguments give however the opposite, more usual behaviour that the terms beyond the first one appear as corrections becoming less and less important at lower frequencies.
-For instance, within the SPA we obtain
+\SM{[Add full expressions for the $\epsilon$ quantities]}
+
+\SM{[justify somewhere smallness of cubic-in-phase term]}
+
+We now investigate the magnitude of higher-order corrections in the case of precessing binaries, using the error estimates $\epsilon_{\Psi 2}$, $\epsilon_{A1}$ and $\epsilon_{A 2}$ defined in~\eqref{eq:deffom}.
+We would like to stress that these figures are to be understood as merely order-of-magnitude estimates, since they compute the error of each approximation by gauging the magnitude of the first corrective term in a Taylor series.
+In particular, they rely on computing derivatives of the time-domain modulation, which can be sensitive to oscillatory features.
+
+First, we would like to stress that the modulation functions $\calD^{\ell *}_{mm'}$. As shown in~\ref{app:Wigner}, their phase dependence is.
+
+For the inspiral phase of the dynamics, one can read the quantities $\epsilon_{\Psi 2}$, $\epsilon_{A1}$ and $\epsilon_{A 2}$ from the outcome of post-Newtonian computations.
+To obtain a qualitative picture of the timescales involved, here we will consider the most simple setting for precession in the post-Newtonian framework: at leading PN order and for a single-spin system, both the orbital angular momentum $\bm{L}$ and the spin vector $\bm{S}$ undergo what is called simple precession around the total angular momentum, $\bm{J}$ (see e.g.~\cite{Apostolatos+94, Kidder95}).
+By neglecting the slow evolution of the orbit under radiation reaction, ignoring higher-order corrections that are quadratic and higher in the spins, and orbit-averaging the precession equations (thus eliminating nutations in the precession of the orbital plane, which occur at the orbital timescale), the trajectories of $\bm{L}$ and $\bm{S}$ become cones around the direction of $\bm{J}$, with a constant precession velocity $\Omega_{\rm prec}$ that we will use to define the timescale of variations of the modulation functions $F$.
+Note that the presence of two spins complicates the evolution of the system, but the picture of a precession cone for $\bm{\ell}$ remains approximately valid (see~\cite{} \SM{[C]} for a classification of the morphologies of precessing trajectories).
+
+In the following, we will use the notations $m=m_{1}+m_{2}$, $\nu=m_{1}m_{2}/m^{2}$, $\delta = (m_{1}-m_{2})/m$, $|\bm{S}_{A}|=Gm_{A}^{2} \chi_{A}$ for $A=1,2$.
+We take the convention $m_{1} \geq m_{2}$.
+We define $\bm{\ell}$ as the unit vector normal to the orbital plane, and assume that only the more massive object has a spin, so that $\bm{S}_{2} = 0$.
+A factor $c$ is scaled out of the spins so that $S_{A} = G m_{A}^{2} \chi$, with $\chi$ the dimensionless spin between $0$ and $1$ for Kerr black holes.
+We will use the notation $v = (G m \omega/c^{3})^{1/3}$ with $\omega$ the orbital frequency, which is equivalent to the frequency-domain definition $v=(G \pi m f/c^{3})^{1/3}$ when the SPA applies.
+\SM{[Check redundancy with notations of previous sections]} The Newtonian order orbital angular momentum is $\bm{L} = L_{N} \bm{\ell}$, with
+\be\label{eq:defLN}
+	L_{N} = \frac{G m^{2} \nu}{c v} \,.
+\ee
+Since we neglect radiation reaction, $\bm{J}$ is treated a constant that we use to set the z-axis so that $\bm{J} = J \bm{e}_{z}$, and we decompose the spin in its aligned and perpendicular components as $\bm{S}_{1} = S_{1}^{z} \bm{e}_{z} + \bm{S}_{1}^{\perp}$.
+At leading order, the precession equations read
+\begin{align}
+	\dot{\bm{S}}_{1} &= \bm{\Omega}_{1} \times \bm{S}_{1} \,, \nn\\
+	\dot{\bm{\ell}} &= - \frac{1}{c L_{N}}  \dot{\bm{S}}_{1}\,, \nn\\
+\end{align}
+with the spin precession velocity \cite{} \SM{[C]}
+\be
+	\bm{\Omega}_{1} = \Omega_{1} \bm{\ell} = \frac{c^{3}}{G m} \left( \frac{3}{4} - \frac{\nu}{2} + \frac{3\delta}{4} \right) v^{5} \bm{\ell} \,.
+\ee
+which is formally a 1PN quantity.
+Decomposing the vectors in their in-plane component and projection on $z$, and using $\delta^{2} = 1-4\nu$ to make explicit the overall scaling in $\nu$, we obtain for the frame precession velocity (defined such that $\dot{\alpha} = \Omega_{\rm prec}$)
+\be
+	\Omega_{\rm prec} = \frac{c^{4}}{G^{2} m^{3}} \frac{7+\delta}{2(1+\delta)} v^{6} J\,.
+\ee
+For convenience, we will use the notations
 \begin{subequations}
 \begin{align}
-	\frac{\ud^{3} \Psi}{\ud f^{3}} \frac{1}{F} \partial_{t}^{3} F &= \calO(6) \,, \\
-	\frac{1}{A}\frac{\ud A}{\ud f} \frac{1}{F} \partial_{t} F &= \calO(2) \,.
+	\Lambda &\equiv \frac{7+\delta}{4(1+\delta)} \,, \\
+	\xi &\equiv 1 + \frac{v S_{1}^{z}}{G m^{2} \nu} \,.
 \end{align}
 \end{subequations}
-
-\begin{figure}
+$\Lambda$ is a mass ratio-dependent factor chosen to be of order 1, varying from $7/4$ for equal masses to $1$ in the test-mass limit.
+The factor $\xi$ shows how the contribution of the spin to the total angular momentum affects the precession rate.
+With these notations,
+\be\label{eq:Omegaprec}
+	\Omega_{\rm prec} = \frac{2c^{3} \nu}{G m} \Lambda \xi v^{5} \,.
+\ee
+If precession effects are in general larger for larger mass ratios, as the precession cones widens, the overall scaling of $\nu$ in~\eqref{eq:Omegaprec} above ($\Lambda$ has a mild variation with masses, which is also decreasing with increasing mass ratio) shows that, as long as the orbital angular momentum still dominates the spin, high mass ratio systems have also a slower precession rate.
+
+We now turn to the Euler angles and modulation functions $\calD^{\ell *}_{mm'}(\alpha, \beta, \gamma)$, given explicitly in~\eqref{eq:defWignerD}.
+In the case of simple precession, the opening angle of the precession cone, $\beta$, evolves solely due to radiation reaction.
+If we treat it as a constant, the minimal rotation condition relates $\gamma$ to $\alpha$ by $\gamma  = -\alpha \cos \beta$ (up to a constant), and the only variable part of the modulation functions~\eqref{} \SM{[R]} are the phases.
+The constant rate of rotation around $\bm{J}$ translates into $\dot{\alpha} = \Omega_{\rm prec}$.
+From the closure relation $\bm{J} = \bm{L} + \bm{S}_{1}$, we have
+\be\label{eq:betaconst}
+	\cos \beta = \sqrt{1 - \left( \frac{S_{1}^{\perp}}{c L_{N}} \right)^{2}}\,.
+\ee
+For each mode contribution $\calD^{\ell *}_{mm'}$, we have
+\be
+	\calD^{\ell *}_{mm'} (\alpha, \beta, \gamma) \propto e^{i(m' \cos\beta - m) \alpha}
+\ee
+which gives an additional geometric factors of $(m' \cos\beta - m)$.
+Under our assumptions of simple precession, ignoring radiation reaction, and using the leading-order expressions~\eqref{} \SM{[R]} for the Fourier-domain timescales, we finally obtain the following error estimates for individual modes \SM{[restore the square in epsilonA2]}:
+\begin{subequations}
+\begin{align}
+	\epsilon_{\Psi 2} &= \frac{5\nu}{96} \Lambda^{2} \xi^{2} (m' \cos\beta - m)^{2} v^{-1} \,, \\
+	\epsilon_{A 1} &=  \frac{7\nu}{6} \Lambda \xi |m' \cos\beta - m| v^{2} \,, \\
+	\epsilon_{A 2} &= \frac{91 \nu^{2}}{72} \Lambda^{2} \xi^{2} (m' \cos\beta - m)^{2} v^{4} \,.
+\end{align}
+\end{subequations}
+As noted in Section~\ref{} \SM{[]}, $\epsilon_{\Psi 2}$ has a factor $v^{-1}$, formally at $-0.5$PN order, which means that the relative size of this correction grows towards smaller frequencies, away from merger.
+Remember however that the opening angle of the precession cone, giving the overall normalization for precession effects in the waveform, also goes to 0 as $v$ in that limit.
+The amplitude error estimates $\epsilon_{A1}$, $\epsilon_{A2}$, by contrast, have the more usual behaviour of PN corrections (of order 1PN and 2PN respectively) growing towards merger.
+The geometric factors $(m' \cos\beta - m)$ show that higher $|m-m'|$ mode contributions are harder to model, with larger $\epsilon$, although they also have smaller amplitudes for small $\beta$.
+The overall $\nu$ scalings indicate a better separation of timescales for larger mass ratio systems.
+These expressions also show that, since the total angular momentum appear as a factor in $\Omega_{\rm prec}$, higher-order corrections will be larger for spin-aligned systems than for antialigned spins \SM{[language: antialigned or anti-aligned ?]}, an effect that becomes significant for large mass ratios.
+Note that when $\xi$ gets close to 0, it means that the system is in the transitional precession range, with the spin of the primary compensating the orbital angular momentum, and our analysis based on simple precession is not valid anymore.
+
+The previous expressions for the error estimate only aim at providing a crude order-of-magnitude picture.
+We now present a more complete numerical computation of these quantities for our post-merger extended precession model presented in Sec.~\ref{subsec:precmodel}, for the three examples summarized in Table.~\ref{table:precexamples}.
+Here and in the following we present only the $h^{\rm I}_{2 m}$ mode contributions induced by $h^{\rm P}_{22}$, knowing that the ones induced by $h^{\rm P}_{2,-2}$ can be deduced from~\eqref{} \SM{[R]}.
+We compute the Fourier-domain based timescales $T_{f}$, $T_{A1}$ and $T_{A2}$ from the analytic derivatives of the PhenomD phase and amplitude (modified to be of class $C^{2}$).
+The time derivatives of the time-domain modulation are computed from an interpolating cubic spline.
+The results are shown in Fig.~\ref{fig:fomprec}, using the frequency at merger, shown by the vertical lines, to give an idea of the separation between the inspiral and post-merger phases.
+The first two cases, $++$ and $pp$, have \SM{[come back to this after having checked trajectories]} the usual behaviour of precession on a cone, with a larger opening angle for $pp$ due to the larger in-plane spin components.
+After merger, the case $++$ shows a much faster frame rotation than the $pp$ case, as shown by the values of $\Omega_{\rm frame}$ in Table~\ref{table:precexamples}, due to the larger remnant spin that incorporates the aligned spin components.
+The case $--$ is qualitatively different, with a larger variation of $\bm{J}$ in direction, and a large and varying $\beta$.
+Since the $\epsilon$ quantities are the magnitude of the first or second term in a Taylor series, we will take $\epsilon \sim 1$ as an indication of the breakdown of a perturbative treatment.
+We recall an important point about Fig.~\ref{fig:fomprec}: since the error estimates are computed on a mode-by-mode basis and normalized by the transfer function for this mode, the amplitude hierarchy between mode contributions is erased here.
+What is plotted is an error estimate in relative sense, and one should keep in mind that higher $|m-m'|$ modes, which we will find harder to model accurately, can be very subdominant in the final waveform.
+
+The resulting magnitude of the error estimates in Fig.~\ref{fig:fomprec} is roughly in agreement with the PN-inspired computation~\ref{eq:} \SM{[R]} above for the inspiral \SM{[check if the PN is quantitatively ok -- overlay it on the plot ?]}.
+For the inspiral, the magnitude of the quadratic phase term $\epsilon_{\Psi 2}$ has a negative slope $v^{-1}=(Mf)^{-1/3}$ \SM{[check the slope on the figure]}, with a clear hierarchy that is due to the prefactor of $\alpha$ in the phase of the modulation, as explained above.
+This hierarchy is not respected for $--$.
+The mode $m=2$, the most important to model once the amplitudes are refactored in, is mainly in the perturbative regime, but we see that the perturbative treatment can be challenged for modes $m\leq 0$ and also for $m=1$ in the case $++$, especially for the post-merger.
+The amplitude corrections are found to be in the perturbative regime for all cases during the inspiral, however both the $++$ and $pp$ cases show a sharp increase of $\epsilon_{A1}$, $\epsilon_{A2}$ post-merger.
+The case $--$, with its small remnant spin and mild post-merger frame rotation, shows no such increase.
+
+Overall, the conclusion to be drawn of Fig.~\ref{} is that for the $++$ and $pp$ cases we expect the perturbative approach to be applicable only during the inspiral, with a possible breakdown for the post-merger phase, especially for the case $++$ with its fast post-merger frame rotation.
+The higher $|m-m'|$ modes are expected to be more challenging to compute accurately, including during the inspiral.
+To make this picture quantitative, we need a full comparison of the signals processed at different orders of approximation against a numerical Fourier transform, which will be presented in Sec.~\ref{subsec:precerror} below.
+
+\SM{[Pb: scaling of FOM might be dominated here by SS oscillations, notably for 2nd der. ? To check - check also geometric scaling from mode to mode]}
+
+\begin{figure*}
   \centering
-  \includegraphics[width=.48\linewidth]{plots/precfom22caseI.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/precfom21caseI.pdf}
+  \includegraphics[width=.32\linewidth]{plots/fom_pp_Psi2.pdf}
+  \hspace{0cm}
+  \includegraphics[width=.32\linewidth]{plots/fom_pp_A1.pdf}
+  \hspace{0cm}
+  \includegraphics[width=.32\linewidth]{plots/fom_pp_A2.pdf}
   %
-  \includegraphics[width=.48\linewidth]{plots/precfom22caseII.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/precfom21caseII.pdf}
+  \includegraphics[width=.32\linewidth]{plots/fom_++_Psi2.pdf}
+  \hspace{0cm}
+  \includegraphics[width=.32\linewidth]{plots/fom_++_A1.pdf}
+  \hspace{0cm}
+  \includegraphics[width=.32\linewidth]{plots/fom_++_A2.pdf}
+  %
+  \includegraphics[width=.32\linewidth]{plots/fom_--_Psi2.pdf}
+  \hspace{0cm}
+  \includegraphics[width=.32\linewidth]{plots/fom_--_A1.pdf}
+  \hspace{0cm}
+  \includegraphics[width=.32\linewidth]{plots/fom_--_A2.pdf}
   \caption{Figures of merit of the approximation as defined in~\eqref{eq:deffom} for the precession modulations.
+\SM{[transpose, compact the columns to share the x axis, fix legends, add horizontal line for $10^{0}$]}}
+  \label{fig:fomprec}
+\end{figure*}
+
+%\begin{figure*}
+%  \centering
+%  \includegraphics[width=.48\linewidth]{plots/precfom22caseI.pdf}
+%  \hspace{0.2cm}
+%  \includegraphics[width=.48\linewidth]{plots/precfom21caseI.pdf}
+%  %
+%  \includegraphics[width=.48\linewidth]{plots/precfom22caseII.pdf}
+%  \hspace{0.2cm}
+%  \includegraphics[width=.48\linewidth]{plots/precfom21caseII.pdf}
+%  \caption{Figures of merit of the approximation as defined in~\eqref{eq:deffom} for the precession modulations.
 The precessing-frame waveform is taken to be the $22$ mode of an equal-mass, non-spinning system.
@@
 The amplitude corrections become non-negligible after merger, and become large enough to challenge a perturbative treatment in Case II for $F_{22,21}$ (notice the change of scale for the lower right panel).}
-  \label{fig:TfTA}
-\end{figure}
+%  \label{fig:TfTA}
+%\end{figure*}
 
@@
 \subsection{Direct convolution approach for the merger-ringdown phase}
-\label{subsec:directconvolution}
+\label{subsec:convolution}
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/trig_amp.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/trig_phase.pdf}
+  \caption{Amplitude (left panel) and phase (right panel), compared to its trigonometric polynomial representation. \SM{[Add different aligned spins here]} \SM{[color for vertical lines at fpeak]}}
+  \label{fig:trig}
+\end{figure*}
+
+\SM{[possibly move this to the formalism section ? but plots for our three examples]}
 
@@
 Defining $\Delta f \equiv 2 (f_{\rm max} - f_{0})$, we write
+\begin{widetext}
 \be\label{eq:defhsym}
@@
 	\begin{cases} 
-		\exp\left[- i \left( \Psi(f_{0}) - 2i\pi (f-f_{0}) t_{p} \right) \right] \tilde{h}(f) \,,  &\text{ for } f \in [f_{0}, f_{0} + \Delta f /2] \\
+		\exp\left[i \Psi(f_{0}) - 2i\pi (f-f_{0}) t_{p} \right] \tilde{h}(f) \,,  &\text{ for } f \in [f_{0}, f_{0} + \Delta f /2] \\
 	\tilde{h}_{\rm sym}(2 f_{0} - f)^{*} \,,  &\text{ for } f \in [f_{0} - \Delta f/2, f_{0}]
@@
 \ee
+\end{widetext}
 
@@
 This construction is intimately related to the DFT, but we recall it for completeness.
-For a periodic function $f(x)$ defined on $x\in [x_{0}, x_{0} + \Delta x]$, and represented by $N$ samples $x_{j} = \ov{x} + j \Delta x/N$ (with $N$ chosen to satisfy the Nyquist criterion), we can build a trigonometric interpolant $P(x)$ as
+For a periodic function $f(x)$ defined on $x\in [x_{0}, x_{0} + \Delta x]$, and represented by $N$ samples $x_{j} = \ov{x} + j \Delta x/N$ (with $N$ chosen to satisfy the Nyquist criterion \SM{[historically should be called Shannon ?]}), we can build a trigonometric interpolant $P(x)$ as
 \be
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.48\linewidth]{plots/trigamp.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/trigphase.pdf}
-  \caption{Amplitude (left panel) and phase (right panel) of the high-frequency part of $\tilde{h}_{\rm sym}$, compared to its trigonometric polynomial representation.}
-  \label{fig:trig}
-\end{figure}
-
 Now, assuming that we wish to compute~\eqref{eq:convolution} for a frequency $f$ such that the extent of the convolution does not extend beyond $[f_{1}, f_{\rm max}]$, we can approximate
@@
 When inserting this representation~\eqref{eq:defhsym} and~\eqref{eq:hsymtrigo} into~\eqref{eq:convolution}, we obtain
-\be\label{eq:resultdirectconvol}
-	\tilde{s}(f) = e^{-i \Psi(f_{0})} e^{2i\pi (f-f_{0}) t_{p}} \sum\limits_{k=-M}^{+M} (-1)^{k} c_{k} e^{2i\pi k \frac{f-f_{0}}{\Delta f}} F(t_{p} + k\delta t) \,, 
-\ee
+\begin{align}\label{eq:resultdirectconvol}
+	\tilde{s}(f) &= e^{-i \Psi(f_{0})} e^{2i\pi (f-f_{0}) t_{p}} \nn\\
+	& \qquad \cdot\sum\limits_{k=-M}^{+M} (-1)^{k} c_{k} e^{2i\pi k \frac{f-f_{0}}{\Delta f}} F(t_{p} + k\delta t) \,, 
+\end{align}
 where we defined the time sampling $\delta t \equiv 1/\Delta f$.
@@
 
+\SM{[update performance - plots were made with double number of points]}
+
+\SM{[explain what this all means translated in terms of sampling in the time-domain -- better relate this approach to simply IFFT]}
+
+\SM{[for plot illustrating reconstruction trigo of amplitude : what normalization to pick for A FD ?]}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{figure}
+We now assess the accuracy of the transfer function computation, applying the formalism of Secs.~\ref{} and \ref{} to our three examples listed in Table~\ref{table:precexamples} and comparing the result to a reference numerical computation.
+
+To numerically compute the transfer functions, we first have to perform an IFFT of our Fourier-domain P-frame waveform, to obtain $h^{\rm P}_{22}$ as a function of time.
+To mitigate the effect of the necessary tapering of the waveform when computing this numerical inverse Fourier transform, we apply a Planck-window tapering on the range $f\in 16-20 \text{Hz}$, using a total mass of $M=20 \Msol$.
+\SM{[Update this to $f_{\rm min} = 10\text{Hz}$]}.
+In parallel, we generate an SEOB waveform with the appropriate length in time.
+Both are aligned to peak at $t=0$, we build the post-merger extended modulation functions as explained in Sec.~\ref{} \SM{[R]}, and we compute the Fourier-domain modes $h^{\rm I}_{2m}$ with an FFT.
+The transfer functions are then computed using~\eqref{eq:defprectransfer}.
+
+In the following, we will either present the errors obtained with the formalism of Sec.~\ref{sec:formalism} applied to the whole frequency band, including the merger and ringdown part of the signal, or cover the high-frequency range with the convolution treatment of Sec.~\ref{subsec:convolution}, with a Planckian window on $Mf\in [Mf_{0}, Mf_{1}]$ with $f_{0}=$ \SM{[complete: definition of f0, f1 ?]} for a smooth transition between the two formalisms.
+
+The figures Fig.~\ref{fig:precerrors++}, Fig.~\ref{fig:precerrorspp} and Fig.~\ref{fig:precerrors--} display three successive approximations: $\{\Psi:0,A:0\}$, which is simply the leading-order transfer function~\eqref{eq:transferlocal}, ignoring all the corrections; $\{\Psi:2,A:2\}$, which incorporates both the phase corrections~\eqref{eq:Fresnelstencil}, using a stencil size $N=3$, and the amplitude corrections up to second order in~\eqref{eq:transferfinal}; and $\{\Psi:2,A:2,\text{conv}\}$, which is the same as the previous for the inspiral but uses the convolution formalism of Sec.~\ref{}.
+They show the Fourier-domain amplitude for each of the modes, both exact and from the reconstruction, the relative errors in amplitude, and the errors in phase.
+The errors are given in a relative sense, and the amplitude allow to visualize the mode hierarchy, and give an idea of how much the relative errors matter in an absolute sense.
+
+The $++$ case, as expected from the analysis of the previous section, is the most challenging, with large relative errors.
+Applying higher-order corrections improves the accuracy in the inspiral, but the most difficult modes $m=-1$ and $m=-2$ still show large errors.
+We also find that higher-order corrections do not reduce the errors in the merger-ringdown region, indicating a breakdown of the perturbative formalism as indicated by Fig.~\ref{fig:fomprec}.
+Using the convolution treatment improves the main modes $m=2$, $m=1$, but not the most challenging modes $m=-1$ and $m=-2$, which can be seen to depart from the perturbative treatment before the range covered by the convolution.
+Note however that, as we will show below by computing the full unfaithfulness these large errors for the higher modes do not affect much the full waveform.
+\SM{[explain how the convolution formalism fails - convolution support to the right, in low-amp region, not well controlled]}
+
+For the $pp$ case, we find a clearer picture.
+Errors are increasing to larger $|m-m'|$, higher-order corrections improve the reconstruction but are unsufficient for the merger-ringdown, and the convolution works for all modes on the frequency range where it is applied.
+The errors at the very high end of the frequency band occur when the overall amplitudes are low, and are therefore unimportant.
+
+For the $--$ case, the mode hierarchy is not respected, as the frame trajectory does not quite follow the picture of simple precession.
+The precession velocity is much milder, leading to smaller errors.
+Apart from some amplitude errors in the merger-ringdown region, we find that in this case even the leading-order treatment gives very reasonable results.
+
+In order to illustrate what these errors really mean for the analysis of GW signals, we also compute the unfaithfulness (or mismatch) between various orders of approximation and the exact, numerical result.
+This unfaithfulness figure, although giving a simplified view of waveform inaccuracies, is commonly used to quantify disagreements between template families and to compare them to numerical relativity waveforms.
+For real signals $h_{1}$ and $h_{2}$, one introduces the usual noise-weighted scalar product
+\be\label{eq:defoverlap}
+	\left( h_{1} | h_{2} \right) = 4\text{Re} \int_{0}^{+\infty} \ud f \frac{\tilde{h}_{1} \tilde{h}_{2}^{*}}{S_{n}(f)} \,,
+\ee
+with $S_{n}(f)$ the noise power spectral density, for which we use the aLIGO ZDHP noise curve~\cite{} \SM{[R]}.
+
+For spin-aligned waveforms, the unfaithfulness is customarily defined as the overlap~\eqref{eq:defoverlap} above optimized over the time at coalescence and global phase of the signal.
+In the precessing case, different prescriptions coexist, differing by the extrinsic parameters that are optimized over.
+Here we adopt the same prescription that was used in the construction of a numerical relativity surrogate model for precessing waveforms~\cite{} \SM{[R]}.
+We refer to the Appendix.~\SM{[R]} of~\cite{} for more details \SM{[complete a bit here]}.
+
+Here we take our illustrative model of Sec.~\ref{} \SM{[R]} as the reference waveform, thus ignoring the inaccuracy introduced by our approximate treatment of post-merger precession and focusing only on the passage from time-domain to Fourier-domain.
+We show the unfaithfulness obtained for 3 inclination angles between the line-of-sight and the direction of the final $\bm{J}$, $0$ (face-on), $\pi/3$ and $\pi/2$ (edge-on), using different approximations (note that we consider more options than for the preceding error plots).
+The total mass ranges from $20\Msol$ (where the inspiral dominates) \SM{[to be updated to $10 \Msol$]} to $400\Msol$ (where the merger-ringdown phase dominates).
+The symbol $\Psi:n$ with $n=0$ or $n=2$ means that the phase corrections~\eqref{eq:Fresnelstencil} are either ignored or included with a stencil size $N=3$.
+The symbol $A:n$ with $n=0,1,2$ indicates the order at which the amplitude corrections in~\eqref{eq:transferfinal} are included.
+Additionally, for all these approximation orders we either apply the perturbative formalism of Sec.~\ref{sec:formalism} to the whole frequency band (symbol ``no conv''), or treat the high-frequency part with the convolution formalism of Sec.~\ref{subsec:convolution} (symbol ``conv'').
+
+Note that the order of approximation $\{\Psi:0,A:0\}$ corresponds to the leading-order formula~\eqref{eq:transferlocal}, which is close (but not identical; see the discussion in Sec.~\ref{} \SM{[R]}) to the PhenomP~\cite{} \SM{[C]} treatment.
+The order of approximation $\{\Psi:2,A:0\}$ is equivalent to formalism of Ref.~\cite{KCY14} with a stencil order of $N=3$, except that the latter formalism is based on an SPA representation of the P-frame waveform, which we bypass by keeping to the Fourier-domain, as discussed above in Sec.~\ref{} \SM{[R]}.
+
+For the $++$ case, the face-on computation (where only the $h^{\rm I}_{22}$ mode contributes) shows good agreement even at leading order.
+By contrast, with $\pi/3$ or $\pi/2$ inclination the mismatch reaches $10^{-2}$ for high masses, even with the perturbative corrections.
+Using the convolution limits the mismatch, and the inclusion of amplitude corrections makes a clear improvement.
+
+For the $pp$ case, both the perturbative corrections and the convolution make a clear improvement.
+One can see clearly that the convolution becomes unimportant at low masses, as the inspiral dominates here.
+
+The $--$ case does not show the same behaviour with inclination as the other ones, as we measure the inclination angle with respect to the final direction of $\bm{J}$, which moves during the inspiral.
+The mismatches are small in this case.
+
+Overall, we find that all the corrections we investigated, perturbative in phase, perturbative in amplitude, and the convolution treatment, are relveant in lowering the unfaithfulness.
+The only exception are second-order amplitude corrections that make little difference, and can even give worse resluts than first order in some cases.
+Using all the tools at our hands, we are able to keep the unfaithfulness of our three example waveforms at below $2.10^{-3}$ for all masses and inclinations.
+
+\begin{figure*}
   \centering
-  \includegraphics[width=.48\linewidth]{plots/precerrorA22caseI.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/precerrorPsi22caseI.pdf}
-  %
-  \includegraphics[width=.48\linewidth]{plots/precerrorA21caseI.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/precerrorPsi21caseI.pdf}  
-  \caption{Errors in the Fourier-domain transfer fonctions $\calT_{22,22}$ and $\calT_{22,21}$ as defined in~\eqref{eq:defTlmlmp}, for Case I and for various approximations.
+  \includegraphics[width=.32\linewidth]{plots/precerrors_++_00_noconv_col.pdf}
+  \hspace{0.cm}
+  \includegraphics[width=.32\linewidth]{plots/precerrors_++_32_noconv_col.pdf}
+  \hspace{0.cm}
+  \includegraphics[width=.32\linewidth]{plots/precerrors_++_32_conv_col.pdf}
+  \caption{Amplitude of the inertial-frame modes $h^{I}_{2m}$ and errors in the Fourier-domain transfer fonctions $\calT^{2}_{m2}$ for the configuration $++$.
+The columns show the orders of approximation $\{\Psi:0,A:0\}$, $\{\Psi:2,A:2\}$, $\{\Psi:2,A:2,\text{conv}\}$.
+\SM{[Compact the columns to share the x axis, add vertical line for merger frequency and shaded area for stitching pert/conv]} \SM{[check: where is the phase of the orange, mode 20 ?]}}
+  \label{fig:precerrors++}
+\end{figure*}
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=.32\linewidth]{plots/precerrors_pp_00_noconv_col.pdf}
+  \hspace{0.cm}
+  \includegraphics[width=.32\linewidth]{plots/precerrors_pp_32_noconv_col.pdf}
+  \hspace{0.cm}
+  \includegraphics[width=.32\linewidth]{plots/precerrors_pp_32_conv_col.pdf}
+  \caption{Amplitude of the inertial-frame modes $h^{I}_{2m}$ and errors in the Fourier-domain transfer fonctions $\calT^{2}_{m2}$ for the configuration $pp$.
+The columns show the orders of approximation $\{\Psi:0,A:0\}$, $\{\Psi:2,A:2\}$, $\{\Psi:2,A:2,\text{conv}\}$.
+\SM{[Compact the columns to share the x axis, add vertical line for merger frequency and shaded area for stitching pert/conv]}}
+  \label{fig:precerrorspp}
+\end{figure*}
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=.32\linewidth]{plots/precerrors_--_00_noconv_col.pdf}
+  \hspace{0.cm}
+  \includegraphics[width=.32\linewidth]{plots/precerrors_--_32_noconv_col.pdf}
+  \hspace{0.cm}
+  \includegraphics[width=.32\linewidth]{plots/precerrors_--_32_conv_col.pdf}
+  \caption{Amplitude of the inertial-frame modes $h^{I}_{2m}$ and errors in the Fourier-domain transfer fonctions $\calT^{2}_{m2}$ for the configuration $--$.
+The columns show the orders of approximation $\{\Psi:0,A:0\}$, $\{\Psi:2,A:2\}$, $\{\Psi:2,A:2,\text{conv}\}$.
+\SM{[Compact the columns to share the x axis, add vertical line for merger frequency and shaded area for stitching pert/conv.
+Check the post-merger frame trajectory for - -.]}}
+  \label{fig:precerrors--}
+\end{figure*}
+
+%\begin{figure*}
+%  \centering
+%  \includegraphics[width=.48\linewidth]{plots/precerrorA22caseI.pdf}
+%  \hspace{0.2cm}
+%  \includegraphics[width=.48\linewidth]{plots/precerrorPsi22caseI.pdf}
+%  %
+%  \includegraphics[width=.48\linewidth]{plots/precerrorA21caseI.pdf}
+%  \hspace{0.2cm}
+%  \includegraphics[width=.48\linewidth]{plots/precerrorPsi21caseI.pdf}  
+%  \caption{Errors in the Fourier-domain transfer fonctions $\calT_{22,22}$ and $\calT_{22,21}$ as defined in~\eqref{eq:defTlmlmp}, for Case I and for various approximations.
 The black curve is the exact result provided by the FFT.
@@
 include the other modes which show problems ?]}
-  \label{fig:precerrorcaseI}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.48\linewidth]{plots/precerrorA22caseII.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/precerrorPsi22caseII.pdf}
-  %
-  \includegraphics[width=.48\linewidth]{plots/precerrorA21caseII.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/precerrorPsi21caseII.pdf}  
-  \caption{Errors in the Fourier-domain transfer fonctions $\calT_{22,22}$ and $\calT_{22,21}$.
+%  \label{fig:precerrorcaseI}
+%\end{figure*}
+
+%\begin{figure*}
+%  \centering
+%  \includegraphics[width=.48\linewidth]{plots/precerrorA22caseII.pdf}
+%  \hspace{0.2cm}
+%  \includegraphics[width=.48\linewidth]{plots/precerrorPsi22caseII.pdf}
+%  %
+%  \includegraphics[width=.48\linewidth]{plots/precerrorA21caseII.pdf}
+%  \hspace{0.2cm}
+%  \includegraphics[width=.48\linewidth]{plots/precerrorPsi21caseII.pdf}  
+%  \caption{Errors in the Fourier-domain transfer fonctions $\calT_{22,22}$ and $\calT_{22,21}$.
 Conventions are the same as in Fig.~\ref{fig:precerrorcaseI} but for Case II, with a more rapid post-merger frame precession.
@@
 include the other modes which show problems ?]}
-  \label{fig:precerrorcaseII}
-\end{figure}
-
-Figs.~\ref{fig:precerrorcaseI} and~\ref{fig:precerrorcaseII} show the amplitude and the phase error of the various reconstructions of the Fourier-domain transfer functions $\calT_{22,22}$ and $\calT_{22,21}$.
-Alongside with the exact result provided by the FFT of the time-domain precession modulation, we show the result obtained with the leading-order approximation (the local approximation) and with.
-We also show the result for the direct convolution approach presented in Sec.~\ref{subsec:directconvolution}.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.48\linewidth]{plots/precmmsimple.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/precmmasympt.pdf}
-  \caption{Mismatches computed between the exact FFT and the various approximations respresented in Figs.~\ref{fig:precerrorcaseI} and~\ref{fig:precerrorcaseII}.
+%  \label{fig:precerrorcaseII}
+%\end{figure*}
+
+\SM{[say how the problems with reconstruction notably for ++ are probably caused by the convolution extending to the right and not to the left]}
+
+%\begin{figure*}
+%  \centering
+%  \includegraphics[width=.48\linewidth]{plots/precmmsimple.pdf}
+%  \hspace{0.2cm}
+%  \includegraphics[width=.48\linewidth]{plots/precmmasympt.pdf}
+%  \caption{Mismatches computed between the exact FFT and the various approximations respresented in Figs.~\ref{fig:precerrorcaseI} and~\ref{fig:precerrorcaseII}.
 [log scale for masses ?
 discuss the fact that in case II, leading order is accidentally better than Psi 2]}
-  \label{fig:precmm}
-\end{figure}
+%  \label{fig:precmm}
+%\end{figure*}
+
+\begin{figure*}
+  \centering
+  %
+  \includegraphics[width=.32\linewidth]{plots/mismatches_++_inc0.pdf}
+  \hspace{0.1cm}
+  \includegraphics[width=.32\linewidth]{plots/mismatches_++_incpi3.pdf}
+  \hspace{0.1cm}
+  \includegraphics[width=.32\linewidth]{plots/mismatches_++_incpi2.pdf}
+  %
+  \includegraphics[width=.32\linewidth]{plots/mismatches_pp_inc0.pdf}
+  \hspace{0.1cm}
+  \includegraphics[width=.32\linewidth]{plots/mismatches_pp_incpi3.pdf}
+  \hspace{0.1cm}
+  \includegraphics[width=.32\linewidth]{plots/mismatches_pp_incpi2.pdf}
+  %
+  \includegraphics[width=.32\linewidth]{plots/mismatches_--_inc0.pdf}
+  \hspace{0.1cm}
+  \includegraphics[width=.32\linewidth]{plots/mismatches_--_incpi3.pdf}
+  \hspace{0.1cm}
+  \includegraphics[width=.32\linewidth]{plots/mismatches_--_incpi2.pdf}
+  \caption{Unfaithfulness.
+Computed here with $f_{\rm min} = 20\mathrm{Hz}$, for masses between 20 and 400 $\Msol$, using the standard aLIGO noise curve \SM{[reference for that curve]}.
+\SM{[Extend computation to fmin=10 and Mmin=10]} }
+  \label{fig:precunfaithfulness}
+\end{figure*}
 
@@
 
-\section{Explicit expression for the stencil coefficients}
-\label{app:vandermonde}
-
-In this Appendix, we exploit the special structure of the transformation~\eqref{eq:resumquadphase} between the two series to rewrite the resulting linear system as a Vandermonde system.
-Defining the Vandermonde matrix $V(x_{0},\dots,x_{N})$ as $V_{ij} = (x_{j})^{i}$ (with the convention $0^{0} = 1$), setting $x_{j} = j^{2}$, and $b_{p} \equiv (-i\epsilon)^{p}(2p-1)!!$, the linear system~\eqref{eq:stencilsystem} becomes
-\be
-	b_{p} = \sum\limits_{k=0}^{N} V_{p,k} a_{N,k} \,.
-\ee
-Now, since the Vandermonde system can be seen as a reformulation of the polynomial interpolation problem, we know that $({}^{t}V^{-1})_{ij}$ is the coefficient of $X^{i}$ in the Lagrange interpolation polynomial $\prod_{k\neq j} (X-x_{k})/(x_{j} - x_{k})$.
-If we introduce the symmetric polynomials $\sigma$ such that
-\be	
-	\sigma(m, \{x_{1}, \dots, x_{n}\}) = \sum\limits_{1\leq i_{1}<\dots<i_{m}\leq n} x_{i_{1}}\dots x_{i_{m}} \,,
-\ee
-we obtain for the inverse of the matrix $V$
-\be
-	V^{-1}_{ij} = \frac{1}{\prod_{k\neq i} (i^{2} - k^{2})} (-1)^{N-j} \sigma(N-j, \{0,\dots,N^{2}\}\backslash \{i^{2}\}) \,.
-\ee
-This allows us to write a closed-form expression for the $a_{N,k}$, for every finite order $N$:
-\be
-	a_{N,k} = \sum\limits_{p=0}^{N} (-i\epsilon)^{p}(2p-1)!!
-\frac{(-1)^{N-p}}{\prod_{j\neq k} (j^{2}-k^{2})} \sum\limits_{\substack{ 0 \leq i_{1} < \dots < i_{N-p} \leq N \\ i_{1}, \dots, i_{N-p} \neq k}} i_{1}^{2}\dots i_{N-p}^{2}
-\ee
-The combinatorial explosion of the number of terms in the symmetric polynomials, however, makes this expression rather impractical to evaluate for $N$ beyond a few dozen.
-In practice, one can simply solve the original linear system~\eqref{eq:stencilsystem} for each order one is interested in, and an actual analysis is likely to use only stencils with a limited size ($\leq 10$).
+\section{Conventions for the Wigner matrices and precessing frame}
+\label{app:wigner}
+
+In this Appendix, we summarize our conventions for the Wigner matrices.
+
+\SM{[This differs from G. Faye's definition by a transposition $m\leftrightarrow m'$]}
+\be\label{eq:defWignerDapp}
+	\calD^{\ell}_{mm'} (\alpha, \beta, \gamma) = e^{im \alpha} d^{\ell}_{mm'}(\beta) e^{im' \gamma}\,,
+\ee
+\SM{[check]}
+\begin{widetext}
+\be\label{eq:defWignerdapp}
+	d^{\ell}_{mm'}(\beta) = \sum\limits_{k=k_{\rm min}}^{k_{\rm max}} \frac{(-1)^{k}}{k!} \frac{\sqrt{(l+m)!
+(l-m)!
+(l+m')!
+(l-m')!}}{(l+m-k)!
+(l-m'-k)!
+(k-m+m')!} \left( \cos\frac{\beta}{2} \right)^{2\ell+m-m'-2k} \left( \sin\frac{\beta}{2} \right)^{2k-m+m'}\,,
+\ee
+\end{widetext}
+where the boundaries of the sum, $k_{\rm min} = \mathrm{max}(0, m-m')$ and $k_{\rm max} = \mathrm{min}(\ell+m, \ell-m')$, can also be read by enforcing that the arguments of the factorials must be non-negative.
+
+\SM{[add here explicitly the dominant-eigenvector construction of the P-frame]}
 

### 2017-09-05 LISAinference/LISAinference.c

@@ -251,7 +251,7 @@ void getLogLike(double *Cube, int *ndim, int *npars, double *lnew, void *context
   templateparams.nbmode = globalparams->nbmodetemp; /* Using the global parameter for the number of modes in templates */
 
   /* Note: context points to a LISAContext structure containing a LISASignal* */
-  if(globalparams->tagint==0) {
+  if((globalparams->tagint==0) && (!globalparams->tagsimplelikelihood)) {
     LISAInjectionCAmpPhase* injection = ((LISAInjectionCAmpPhase*) context);
 
     //TESTING
@@ -262,7 +262,7 @@ void getLogLike(double *Cube, int *ndim, int *npars, double *lnew, void *context
     //printf("time Likelihood: %g\n", (double) (tend-tbeg)/CLOCKS_PER_SEC);
     //
   }
-  else if(globalparams->tagint==1) {
+  else if((globalparams->tagint==1) && (!globalparams->tagsimplelikelihood)) {
     LISAInjectionReIm* injection = ((LISAInjectionReIm*) context);
 
     //TESTING
@@ -272,6 +272,10 @@ void getLogLike(double *Cube, int *ndim, int *npars, double *lnew, void *context
     //tend = clock();
     //printf("time Likelihood: %g\n", (double) (tend-tbeg)/CLOCKS_PER_SEC);
     //
+  }
+  else if(globalparams->tagsimplelikelihood) {
+    SimpleLikelihoodPrecomputedValues* injection = ((SimpleLikelihoodPrecomputedValues*) context);
+    *lnew = CalculateLogLSimpleLikelihood(injection, &templateparams);
   }
 }
 

### 2017-09-05 LISAinference/LISAinference_common.c

@@ -24,6 +24,9 @@ void addendum(int argc, char *argv[],LISARunParams *runParams, int *ndim, int *n
   memset(priorParams, 0, sizeof(LISAPrior));
   addparams = (LISAAddParams*) malloc(sizeof(LISAAddParams));
   memset(addparams, 0, sizeof(LISAAddParams));
+  /* This structure is used only for storing precomputed values for the simple likelihood */
+  simplelikelihoodinjvals = (SimpleLikelihoodPrecomputedValues*) malloc(sizeof(SimpleLikelihoodPrecomputedValues));
+  memset(simplelikelihoodinjvals, 0, sizeof(SimpleLikelihoodPrecomputedValues));
 
   /* Parse commandline to read parameters of injection - copy the number of modes demanded for the injection */
   parse_args_LISA(argc, argv, injectedparams, globalparams, priorParams, runParams, addparams);
@@ -92,6 +95,12 @@ void addendum(int argc, char *argv[],LISARunParams *runParams, int *ndim, int *n
     report_LISAParams(injectedparams);
   }
 
+  /* If using simple likelihood, initialize precomputed values - note that the other initializations for the injection are done anyway, but will be ignored */
+  /* Note: the optional distance adjustment to a given snr is done above using the response as given by responseapprox, not the simplified response */
+  if(globalparams->tagsimplelikelihood) {
+    LISAComputeSimpleLikelihoodPrecomputedValues(simplelikelihoodinjvals, injectedparams);
+  }
+
   /* Print SNR */
   if (myid == 0) {
     printf("Total SNR: %g\n", SNR123);
@@ -116,11 +125,14 @@ void addendum(int argc, char *argv[],LISARunParams *runParams, int *ndim, int *n
   if(myid == 0) printf("logZtrue = %lf\n", *logZtrue-logZdata);
 
   /* Set the context pointer */
-  if(globalparams->tagint==0) {
+  if((globalparams->tagint==0) && (!globalparams->tagsimplelikelihood)) {
     *contextp = injectedsignalCAmpPhase;
   }
-  else if(globalparams->tagint==1) {
+  else if((globalparams->tagint==1) && (!globalparams->tagsimplelikelihood)) {
     *contextp = injectedsignalReIm;
+  }
+  else if(globalparams->tagsimplelikelihood) {
+    *contextp = simplelikelihoodinjvals;
   }
 
   *nPar = 9;	  /* Total no. of parameters including free & derived parameters */

### 2017-09-05 LISAinference/LISAutils.c

@@ -7,6 +7,7 @@ LISAParams* injectedparams = NULL;
 LISAGlobalParams* globalparams = NULL;
 LISAPrior* priorParams = NULL;
 LISAAddParams* addparams = NULL;
+SimpleLikelihoodPrecomputedValues* simplelikelihoodinjvals = NULL;
 
 /***************** Pasring string to choose what masses set to sample for *****************/
 
@@ -164,7 +165,7 @@ Arguments are as follows:\n\
  --minf                Minimal frequency (Hz, default=0) - when set to 0, use the lowest frequency where the detector noise model is trusted __LISASimFD_Noise_fLow (set somewhat arbitrarily)\n\
  --maxf                Maximal frequency (Hz, default=1Hz) - when set to 0, use the highest frequency where the detector noise model is trusted __LISASimFD_Noise_fHigh (set somewhat arbitrarily)\n\
  --tagextpn            Tag to allow PN extension of the waveform at low frequencies (default=1)\n\
- --tagtRefatLISA         Tag to allow t0 to specify signal offset time at LISA guiding center rather than at SSB (default=0)\n\
+ --tagtRefatLISA       Tag to allow t0 to specify signal offset time at LISA guiding center rather than at SSB (default=0)\n\
  --Mfmatch             When PN extension allowed, geometric matching frequency: will use ROM above this value. If <=0, use ROM down to the lowest covered frequency (default=0.)\n\
  --nbmodeinj           Number of modes of radiation to use for the injection (1-5, default=5)\n\
  --nbmodetemp          Number of modes of radiation to use for the templates (1-5, default=5)\n\
@@ -175,6 +176,7 @@ Arguments are as follows:\n\
  --zerolikelihood      Zero out the likelihood to sample from the prior for testing purposes (default 0)\n\
  --frozenLISA          Freeze the orbital configuration to the time of peak of the injection (default 0)\n\
  --responseapprox      Approximation in the GAB and orb response - choices are full (full response, default), lowfL (keep orbital delay frequency-dependence but simplify constellation response) and lowf (simplify constellation and orbital response) - WARNING : at the moment noises are not consistent, and TDI combinations from the GAB are unchanged\n\
[…]

### 2017-09-05 LISAinference/LISAutils.h

@@ -70,12 +70,13 @@ typedef struct tagLISAGlobalParams {
   int nbmodeinj;             /* number of modes to include in the injection (starting with 22) - defaults to 5 (all modes) */
   int nbmodetemp;            /* number of modes to include in the templates (starting with 22) - defaults to 5 (all modes) */
   int tagint;                /* Tag choosing the integrator: 0 for wip (default), 1 for linear integration */
-  TDItag tagtdi;                /* Tag choosing the TDI variables to use */
+  TDItag tagtdi;             /* Tag choosing the TDI variables to use */
   int nbptsoverlap;          /* Number of points to use in loglinear overlaps (default 32768) */
   LISAconstellation *variant;  /* A structure defining the LISA constellation features */
   int zerolikelihood;        /* Tag to zero out the likelihood, to sample from the prior for testing purposes (default 0) */
   int frozenLISA;            /* Freeze the orbital configuration to the time of peak of the injection (default 0) */
   ResponseApproxtag responseapprox;    /* Approximation in the GAB and orb response - choices are full (full response, default), lowfL (keep orbital delay frequency-dependence but simplify constellation response) and lowf (simplify constellation and orbital response) - WARNING : at the moment noises are not consistent, and TDI combinations from the GAB are unchanged */
+  int tagsimplelikelihood;   /* Tag to use simplified, frozen-LISA and lowf likelihood where mode overlaps are precomputed - can only be used when the masses and time (tL) are pinned to injection values (Note: when using --snr, distance adjustment done using responseapprox, not the simple response) */
 } LISAGlobalParams;
 
 typedef struct tagLISASignalCAmpPhase
@@ -201,6 +202,13 @@ typedef struct tagLISAAddParams {
   char outfile[256];         /* Output file for LISAlikelihood */
 } LISAAddParams;
 
+/* Saved precomputed values for the injection , when using the simplified frozenLISA lowf likelihood - here for now assumes 22 mode only */
+typedef struct tagSimpleLikelihoodPrecomputedValues {
+  double normalization;
+  double complex sa;
+  double complex se;
+} SimpleLikelihoodPrecomputedValues;
+
 /************ Functions for LISA parameters, injection, likelihood, prior ************/
 
[…]

### 2017-09-05 LISAsim/LISAFDresponse.c

@@ -474,14 +474,14 @@ int LISASimFDResponseTDI3Chan(
       /* Phase term due to the R-delay, including correction to first order */
       double phase=variant->OrbitOmega*tforb + variant->OrbitPhi0 - lambda;
       double OrbitRoC=variant->OrbitR/C_SI;
-      //double phaseRdelay = -2*PI*R_SI/C_SI*f*cos(beta)*cos(Omega_SI*tf - lambda) * (1 + R_SI/C_SI*Omega_SI*sin(Omega_SI*tf - lambda));
+      //double phaseRdelay = -2*PI*R_SI/C_SI*f*cos(beta)*cos(Omega_SI*tf - lambda) * (1 + R_SI/C_SI*cos(beta)*Omega_SI*sin(Omega_SI*tf - lambda));
       double phaseRdelay;
       if(tagtRefatLISA==0){//delay so that tinj=torb refers to arrival time at SSB
 	phaseRdelay = -2*PI*OrbitRoC*f*cos(beta)*cos(phase) * (1 + cos(beta)*OrbitRoC*variant->OrbitOmega*sin(phase));
       } else {
 	//In this version we delay so that tinj=torb is relative to LISAcenter arrival time, so there is no orbital delay when tf = tinj
 	//If the original delay realized td=t+d(t), now we want td=t+d(t)-d(t0), that we we change d(t) -> d(t) + d(t0)
-	//Then with the approximation d(td)=d(t)(1-ddot(t)), 
+	//Then with the approximation d(td)=d(t)(1-ddot(t)),
 	double phase0=variant->OrbitOmega*torb+variant->OrbitPhi0-lambda;
 	phaseRdelay =
 	  - 2*PI*OrbitRoC*f*cos(beta)*( cos(phase) -cos(phase0) ) * (1 + cos(beta)*OrbitRoC*variant->OrbitOmega*sin(phase));

### 2017-09-05 LISAsim/LISAgeometry.c

@@ -473,7 +473,8 @@ void SetCoeffsG(const double lambda, const double beta, const double psi) {
 /* Conventions changed: now MLDC conventions */
 
 /* Function evaluating G21, combining the two polarization with the spherical harmonics factors */
-double complex G21mode(const LISAconstellation *variant, const double f, const double t, const double complex Yfactorplus, const double complex Yfactorcross) {
+double complex G21mode(const LISAconstellation *variant, const double f, const double t, const double complex Yfactorplus, const double complex Yfactorcross)
+{
 
   double phase=variant->ConstOmega*t + variant->ConstPhi0;
 
@@ -496,7 +497,8 @@ double complex G21mode(const LISAconstellation *variant, const double f, const d
   return I*PI*f*variant->ConstL/C_SI * (n3Pn3plus*Yfactorplus + n3Pn3cross*Yfactorcross) * sinc( PI*f*variant->ConstL/C_SI * (1.+kn3)) * cexp( I*PI*f*variant->ConstL/C_SI * (1.+kp1plusp2) );
 }
 /* Function evaluating G12, combining the two polarization with the spherical harmonics factors */
-double complex G12mode(const LISAconstellation *variant, const double f, const double t, const double complex Yfactorplus, const double complex Yfactorcross) {
+double complex G12mode(const LISAconstellation *variant, const double f, const double t, const double complex Yfactorplus, const double complex Yfactorcross)
+{
 
   double phase=variant->ConstOmega*t + variant->ConstPhi0;
 
@@ -520,7 +522,8 @@ double complex G12mode(const LISAconstellation *variant, const double f, const d
   return I*PI*f*variant->ConstL/C_SI * (n3Pn3plus*Yfactorplus + n3Pn3cross*Yfactorcross) * sinc( PI*f*variant->ConstL/C_SI * (1.-kn3)) * cexp( I*PI*f*variant->ConstL/C_SI * (1.+kp1plusp2) );
 }
 /* Function evaluating G32, combining the two polarization with the spherical harmonics factors */
-double complex G32mode(const LISAconstellation *variant, const double f, const double t, const double complex Yfactorplus, const double complex Yfactorcross) {
+double complex G32mode(const LISAconstellation *variant, const double f, const double t, const double complex Yfactorplus, const double complex Yfactorcross)
+{
 
   double phase=variant->ConstOmega*t + variant->ConstPhi0;
 
@@ -544,7 +547,8 @@ double complex G32mode(const LISAconstellation *variant, const double f, const d
[…]

### 2017-09-05 LISAsim/LISAgeometry.h

@@ -87,7 +87,7 @@ typedef struct tagLISAconstellation {
   double ConstPhi0;
   double ConstL;
   enum LISANoiseType noise;
-}LISAconstellation;
+} LISAconstellation;
 extern LISAconstellation LISA2017;
 extern LISAconstellation LISA2010;
 extern LISAconstellation slowOrbitLISA;
@@ -95,7 +95,7 @@ extern LISAconstellation tinyOrbitLISA;
 extern LISAconstellation fastOrbitLISA;
 extern LISAconstellation bigOrbitLISA;
 
-  
+
 /**************************************************/
 /**************** Prototypes **********************/
 
@@ -123,7 +123,7 @@ double complex G23mode(const LISAconstellation *LISAvariant, const double f, con
 double complex G13mode(const LISAconstellation *LISAvariant, const double f, const double t, const double complex Yfactorplus, const double complex Yfactorcross);
 double complex G31mode(const LISAconstellation *LISAvariant, const double f, const double t, const double complex Yfactorplus, const double complex Yfactorcross);
 int EvaluateGABmode(
-  const LISAconstellation *LISAvariant, 
+  const LISAconstellation *LISAvariant,
   double complex* G12,                     /* Output for G12 */
   double complex* G21,                     /* Output for G21 */
   double complex* G23,                     /* Output for G23 */
@@ -142,7 +142,7 @@ int EvaluateGABmode(
 /* Note: in case only one channel is considered, amplitudes for channels 2 and 3 are simply set to 0 */
 /* (allows minimal changes from the old structure that assumed KTV A,E,T - but probably not optimal) */
 int EvaluateTDIfactor3Chan(
-  const LISAconstellation *variant,    /* Description of LISA variant */ 
+  const LISAconstellation *variant,    /* Description of LISA variant */
   double complex* factor1,                       /* Output for factor for TDI channel 1 */
   double complex* factor2,                       /* Output for factor for TDI channel 2 */
   double complex* factor3,                       /* Output for factor for TDI channel 3 */
@@ -169,7 +169,7 @@ int EvaluateTDIfactor3Chan(
 /* The factors scaled out, parallel what is done for the noise functions */
 /* Note: in case only one channel is considered, factors for channels 2 and 3 are simply set to 0 */
 int ScaledTDIfactor3Chan(
-  const LISAconstellation *variant,    /* Description of LISA variant */ 
+  const LISAconstellation *variant,    /* Description of LISA variant */
   double complex* factor1,                       /* Output for factor for TDI factor 1 */
[…]

### 2017-09-08 docs/paper_method.tex
Edits to the paper.

@@
 
-\section{Fourier-domain waveforms and separation of timescales}
+\section{Fourier-domain waveforms, modulations and delays}
 \label{sec:motivation}
@@
 \ee
-where the ${}_{-2}Y_{\ell m}(\Theta,\Phi)$ are spin-weighted spherical harmonics. For a non-precessing system, with a fixed orbital plane, the individual modes have the additional symmetry property
+where the ${}_{-2}Y_{\ell m}(\Theta,\Phi)$ are spin-weighted spherical harmonics~\cite{Goldberg+67}. Conversely, the individual polarizations are obtained from the individual modes as
+\begin{subequations}
+\begin{align}
+	h_{+} = \frac{1}{2} \sum\limits_{\ell, m} \left[ {}_{-2}Y_{\ell m}h_{\ell m} + {}_{-2}Y_{\ell m}^{*} h_{\ell m}^{*} \right] \,,\\
+	h_{\times} = \frac{i}{2} \sum\limits_{\ell, m} \left[ {}_{-2}Y_{\ell m}h_{\ell m} - {}_{-2}Y_{\ell m}^{*} h_{\ell m}^{*} \right] \,.
+\end{align}
+\end{subequations}
+
+For a non-precessing system, with a fixed orbital plane, the individual modes have the additional symmetry property
 \be\label{eq:symmetryhlminusm}
-	h_{\ell -m} = (-1)^{\ell} h_{\ell m}^{*} \,.
+	h_{\ell, -m} = (-1)^{\ell} h_{\ell m}^{*} \,.
 \ee
@@
 \begin{align}
-	\tilde{h}_{+}(f) &= \frac{1}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left( {}_{-2}Y_{\ell m} \tilde{h}_{\ell m}(f) + {}_{-2}Y_{\ell m}^{*} \tilde{h}_{\ell m}(-f)^{*} \right) \,, \\
-	\tilde{h}_{\times}(f) &= \frac{i}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left( {}_{-2}Y_{\ell m} \tilde{h}_{\ell m}(f) - {}_{-2}Y_{\ell m}^{*} \tilde{h}_{\ell m}(-f)^{*} \right) \,,
+	\tilde{h}_{+}(f) &= \frac{1}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left[ {}_{-2}Y_{\ell m} \tilde{h}_{\ell m}(f) + {}_{-2}Y_{\ell m}^{*} \tilde{h}_{\ell m}(-f)^{*} \right] \,, \\
+	\tilde{h}_{\times}(f) &= \frac{i}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left[ {}_{-2}Y_{\ell m} \tilde{h}_{\ell m}(f) - {}_{-2}Y_{\ell m}^{*} \tilde{h}_{\ell m}(-f)^{*} \right] \,,
 \end{align}
@@
 
-We recall here an approximation widely used to model gravitational waves signals emitted by inspiralling binaries, the stationary phase approximation (thereafter SPA).
+We recall here an approximation widely used to model gravitational waves signals emitted by inspiralling binaries, the Stationary Phase Approximation (thereafter SPA).
 It applies in general to chirping signals, and we refer the reader to~\ref{} for details.
@@
 In order to keep close to the notations used in the gravitational wave literature, we introduced a factor of 2 in the phase of the wave, which is appropriate for the dominant 22 harmonic of the signal.
-The approximation then applies to signals verifying the conditions
+The approximation then applies to signals verifying the conditions \SM{[say that the third is sometimes missing ?]}
 \be\label{eq:conditionSPA}
@@
 \end{subequations}
+As for the leading-order time-domain amplitude of the $22$ mode, we have~\cite{} \SM{[C]}
+\be\label{eq:a22N}
+	a_{22}^{\rm N} (t) = \frac{2 G m \nu x}{D c^{2}} \sqrt{\frac{16 \pi}{5}} \,,
+\ee
+where we set $x = (G M\omega/c^{3})^{2/3}$ and where $D$ is the luminosity distance to the observer.
 For this Newtonian inspiral, applying the SPA gives for the $22$ mode $\tilde{h}(f) = A_{\rm N}(f)e^{-i\Psi_{\rm N}(f)}$ with
@@
 \SM{[Check sign of $\Psi$ ? Should have $<0$ 2nd der.] }
-where $v=(G m \pi f/c^{3})^{1/3}$, $D$ is the luminosity distance to the observer, and where $t_{0}, \phi_{0}$ are constants.
+where we used the notation $v=(G m \pi f/c^{3})^{1/3}$ ($x=v^{2}$ with the SPA correspondence~\eqref{eq:deftfSPA}), and where $t_{0}, \phi_{0}$ are constants.
 It is also customary to rewrite the above relations in terms of the chirp mass $\calM \equiv m\nu^{3/5}$, which is the only mass combination arising in the phase at the leading PN order.
@@
 For spinning compact objects, the presence of the spins leads to the precession of the orbit~\cite{Apostolatos+94, Kidder95}, which can have large effects on the waveform.
-In particular, it breaks the planar symmetry of the gravitational wave emission, and leads to amplitude modulations that are especially important for edge-on observers.
+In particular, it breaks the planar symmetry of the gravitational wave emission, and leads to modulations that are especially important for systems that are observed edge-on.
 
@@
 
-In Eq.~\eqref{eq:defmodulationprec} above, the modulation function $F$ evolves on the precessional timescale, which evolves throughout the inspiral.
+In Eq.~\eqref{eq:defmodulationprec} above, the modulation function $F$ has time variations on the precessional timescale, which evolves throughout the inspiral.
 The separation between the precessional timescale and the orbital timescale will be crucial for our analysis.
@@
 \begin{itemize}
-	\item the time scales entering the signal $h(t)$ are well separated from the ones entering the modulation and the delay, so that $\tilde{G}(f-f',f')$ has support limited to $|f'| < f_{\rm max}$, with $f_{\rm max}$ small compared the characteristic scale for the variation of the Fourier-domain signal;
-	\item the original signal has a smooth amplitude and phase in the Fourier-domain, slowly varying over the scale $f_{\rm max}$, so that a local treatment in $f'$ of the convolution in~\eqref{eq:FDkernel} will be possible.
+	\item the time scales entering the signal $h(t)$ are well separated from the ones entering the modulation and the delay, so that $\tilde{G}(f-f',f')$ has support limited to $|f'| < f'_{\rm max}$, with $f'_{\rm max}$ small compared the characteristic scale for the variation of the Fourier-domain signal;
+	\item the original signal has a smooth amplitude and phase in the Fourier-domain, slowly varying over the scale $f'_{\rm max}$, so that a local treatment in $f'$ of the convolution in~\eqref{eq:FDkernel} will be possible.
 \end{itemize}
@@
 As such, this time-of-frequency function also extends to the merger-ringdown part of the signals.
-Fig.~\ref{fig:tf} shows the time function $\tf$ for the example of an equal-mass waveform.
+Fig.~\ref{fig:tf} shows the behaviour of the time function $\tf$ around merger for six example waveforms, for mass ratios $q=1$ and $q=8$ and for aligned spin components $\chi=0.95,0.,-0.95$.
 In particular, one should note that $\tf$ is not monotonously increasing with frequency anymore when reaching in the high-frequency part of the waveform, corresponding to the ringdown.
@@
 The interpretation of this approximation is straightforward: the signal is simply multiplied by the response function evaluated at the time $\tf$.
-For precessing binary, this corresponds to the approximation used for the PhenomP approximant~\cite{Hannam+13}, as will be discussed in Sec.~\ref{}.
+For precessing binaries, this corresponds to the approximation used for the PhenomP approximant~\cite{Hannam+13}, as will be discussed in Sec.~\ref{}.
 Additionally, when delays are present, the frequency-dependence of the response is evaluated directly at $f$.
@@
 
-We will formally Taylor-expand $\tilde{h}(f-f')$ in $f'$ as follows: we write
+We will formally Taylor-expand $\tilde{h}(f-f')$ in $f'$ as follows:
 \begin{subequations}
@@
 	A(f-f') &= A(f) \sum\limits_{q\geq 0} \frac{(-1)^{q}}{q!} {f'}^{q} \frac{1}{A}\frac{\ud^{q} A}{\ud f^{q}} \,, \label{eq:expandA}\\
-	\tilde{G}(f-f', f') &= \sum\limits_{r\geq 0} \frac{(-1)^{r}}{r!} {f'}^{r} \frac{\partial^{r} }{\partial f^{r}}  \tilde{G}(f,f') \label{eq:expandg}
+	\tilde{G}(f-f', f') &= \sum\limits_{r\geq 0} \frac{(-1)^{r}}{r!} {f'}^{r} \frac{\partial^{r} }{\partial f^{r}}  \tilde{G}(f,f') \label{eq:expandg} \,.
 \end{align}
@@
 
-This rather cumbersome expression will not be used as such, but is easy to simplify for special cases.
+This rather cumbersome expansion will not be used as such, but takes a more intuitive form for special cases.
 For instance, if we expand only the frequency-dependence of $G$ as in~\eqref{eq:expandg}, we obtain
@@
 
-In the following, it will be convenient to introduce \SM{[think of removing ``figure of merit'' as it is not quite the word]} error estimates for these Taylor-like expansions, that we will take to be the magnitude of the first ignored term in the series.
+In the following, it will be convenient to introduce \SM{[think of removing ``figure of merit'' throughout the text as it is not quite the word]} error estimates for these Taylor-like expansions, that we will take to be the magnitude of the first ignored term in the series.
 Thus we define
@@
 \ee
-we have
+we have \SM{[add more details on Fourier-domain derivatives of SPA amp and phase]}
 \be
@@
 \end{widetext}
-Each term in this series (in practice, only the first few will be relevant) can in turn be approximated by using the stencil formula~\eqref{eq:Fresnelstencil}.
+Each term in this series (in practice, only the first two or three will be relevant) can in turn be approximated by using the stencil formula~\eqref{eq:Fresnelstencil}.
 
@@
 
-The frequency-shift response for a single link~\eqref{eq:yslr} was derived in Ref.~\cite{EW75} (see also~\cite{} \SM{[C]} for a more modern version).
+The frequency-shift response for a single link~\eqref{eq:yslr} was derived in Ref.~\cite{EW75} (see also~\cite{} \SM{[C]} for a more modern derivation).
 Several assumptions enter the result as written in~\eqref{eq:yslr}: (i) effects of the order $v/c$ are neglected, including for instance the special relativistic Doppler effect created by the relative speeds of the spacecrafts on their orbits (ii) the propagation is assumed to take place in a flat spacetime, perturbed only by the gravitational wave; thus the gravitational redshift as well as the deflection of light created by the gravitational potential of the Sun is ignored (iii) all geometric factors are evaluated at a single time, whereas one should consider the beam as propagating from the position of the first spacecraft at the time of emission to the position of the second scapecraft at the time of reception, leading to a pointing-ahead effect.
 
-Additionally, we limit ourselves to a rigid model for the detector, namely we assume that the constellation remains in an equilateral configuration with fixed armlengths $L=5.10^{6}\mathrm{km}$.
+Additionally, we limit ourselves to a rigid model for the detector, namely we assume that the constellation remains in an equilateral configuration with fixed armlengths $L=2.5\times 10^{6}\mathrm{km}$, as proposed in~\cite{} \SM{[C]}.
 This simplified orbits neglect (iv) the effect of excentricity of the individual Keplerian orbits, which introduces corrections at the level $e^{2}$ (v) the effect of gravitational perturbations coming from other celestial bodies, such as the Earth, the quadrupole of the Sun, and the other planets.
 
-Although all the effects (i)-(v) are neglected here, keeping track of these corrections is crucial for the purpose of noise cancellations, and led to the development of new generations of TDI observables~\cite{} \SM{[C]}.
-
-We now turn to the transposition of~\eqref{eq:yslr} to the Fourier-domain.
+Note that although we can neglect all the effects (i)-(v) for our present study, keeping track of these corrections is crucial for the purpose of noise cancellations, and led to the development of new generations of TDI observables~\cite{} \SM{[C]}.
+
 Since the orbital delay due to the orbit around the Sun is common to all $y_{slr}$ observables, we will decompose the full instrument response in two steps.
-The first step consists simply in applying the varying time delay to bring the signal from the SSB to the center of the constellation.
-We write this orbital time delay as
+The first step consists simply in applying the varying time delay to bring the signal from the SSB to the center of the LISA triangular constellation.
+For $h=h^{\rm TT}$ the transverse-traceless gravitational waveform in matrix form, we write this orbital time delay as
 \be\label{eq:defresponseO}
-	h^{\rm TT}_{O} (t) = h^{\rm TT}(t-\hat{k}\cdot p_{O}) \,.
+	h_{O} (t) = h(t-\hat{k}\cdot p_{O}) \,.
 \ee
 The second step consists in the remaining, constellation-centered response, which we write as
-\be\label{eq:dfresponsec}
-	y_{slr} = \frac{1}{2} \frac{1}{1 - \hat{k}\cdot n_{l}} n_{l}\cdot \left[ h_{O}^{\rm TT}(t - L - \hat{k}\cdot p^{L}_{s}) - h_{O}^{\rm TT}(t - \hat{k}\cdot p^{L}_{r}) \right] \cdot n_{l}\,,
-\ee
+\begin{align}\label{eq:dfresponsec}
+	y_{slr} &= \frac{1}{2} \frac{1}{1 - \hat{k}\cdot n_{l}} \nn\\
+	& \qquad \cdot n_{l}\cdot \left[ h_{O}(t - L - \hat{k}\cdot p^{L}_{s}) - h_{O}(t - \hat{k}\cdot p^{L}_{r}) \right] \cdot n_{l}\,,
+\end{align}
 where we introduced the positions of the spacecrafts relative to the center of the constellation, $p^{L}_{A} \equiv p_{A} - p_{O}$.
@@
 \be
-	h^{\rm TT} = h_{+}P_{+} + h_{\times}P_{\times} \,.
+	h = h^{\rm TT} = h_{+}P_{+} + h_{\times}P_{\times} \,.
 \ee
@@
 
+We now turn to the transposition of~\eqref{eq:yslr} to the Fourier-domain.
 When using the rigid approximation for the constellation, a particular simplification occurs when writing the transfer function at leading order in the delays.
-The Fourier-domain transfer function for~\eqref{eq:dfresponsec} then reads in the local approximation
+The Fourier-domain transfer function for~\eqref{eq:dfresponsec} then reads in the local approximation~\eqref{eq:transferlocal}
 \begin{align}
@@
 \SM{[crude estimate of the target accuracy level: 1/SNR ?]}
+
+\SM{[distinguish orbital from constellation at the level of orders of magnitude estimates: factor 2pif*d when taking derivatives of the transfer function for the pure delay]}
 
@@
 
+\SM{[We have to say why we take a non-spinning, equal-mass system - say why the method should also work for precessing signals]}
+
+\SM{[Write explicitly what the orders A1, d1 mean]}
+
 \begin{figure*}
@@
 We define $(\alpha, \beta, \gamma)$ are the Euler angles relating the precessing frame to the inertial frame in the $(z,y,z)$ convention.
-The $z$ axis of the inertial frame is chosen to be the direction of $\bm{J}$, the total angular momentum, which is almost constant\footnote{Except in the cases of transitional precession~\cite{Apostolatos94}, where for high mass ratios and large antialigned spins the orbital angular momentum and the spin can almost cancel each other.}.
+The $z$ axis of the inertial frame is chosen to be the direction of $\bm{J}$, the total angular momentum, which is almost constant\footnote{Except in the cases of transitional precession~\cite{Apostolatos+94}, where for high mass ratios and large antialigned spins the orbital angular momentum and the spin can almost cancel each other.}.
 The first two angles, $\alpha$ and $\beta$, are the two spherical angles tracking the direction of dominant radiation reaction, which during the inspiral follows essentially the normal to the orbital plane.
@@
 Table~\ref{table:precparams} also shows the remnant spins that are internally computed in both the SEOB and the PhenomD/PhenomP codes, as well as the QNM frequencies and the resulting frame precession frequency $\Omega_{\rm frame}$.
-Note that the case $--$ shows a larger variation in the direction of $\bm{J}$, and the spin of the remnant is almost 0, due to the fact that the spin is large and antialigned, for a large mass ratio.
+Note that the case $--$ shows a larger variation in the direction of $\bm{J}$, and the spin of the remnant is almost 0, due to the fact that the spin is large and antialigned, with a large mass ratio.
 This example therefore serves the purpose of lying at the edge of validity of the picture of standard precession along a cone around an almost-fixed direction, and the standard hierarchy between mode contributions is not valid in this case.
@@
 But the trajectories are not very informative except for - - (where our model might not be very solid) and for showing oscillations in $\beta$.
-The reconstruction errors plots already show the amplitude of the modes themselves, giving an idea of the hierarchy; qhat is the missing is the phase in Tlmmp, which is not very informative.]}
+The reconstruction errors plots already show the amplitude of the modes themselves, giving an idea of the hierarchy; what is the missing is the phase in Tlmmp, which is not very informative.]}
 
@@
 \be
-	\bm{\Omega}_{1} = \Omega_{1} \bm{\ell} = \frac{c^{3}}{G m} \left( \frac{3}{4} - \frac{\nu}{2} + \frac{3\delta}{4} \right) v^{5} \bm{\ell} \,.
+	\bm{\Omega}_{1} = \Omega_{1} \bm{\ell} = \frac{c^{3}}{G m} \left( \frac{3}{4} + \frac{\nu}{2} - \frac{3\delta}{4} \right) v^{5} \bm{\ell} \,.
 \ee
@@
 which gives an additional geometric factors of $(m' \cos\beta - m)$.
-Under our assumptions of simple precession, ignoring radiation reaction, and using the leading-order expressions~\eqref{} \SM{[R]} for the Fourier-domain timescales, we finally obtain the following error estimates for individual modes \SM{[restore the square in epsilonA2]}:
+Under our assumptions of simple precession, ignoring radiation reaction, and using the leading-order expressions~\eqref{} \SM{[R]} for the Fourier-domain timescales, we finally obtain the following error estimates for individual modes:
 \begin{subequations}
@@
 \end{subequations}
-As noted in Section~\ref{} \SM{[]}, $\epsilon_{\Psi 2}$ has a factor $v^{-1}$, formally at $-0.5$PN order, which means that the relative size of this correction grows towards smaller frequencies, away from merger.
+As noted in Section~\ref{} \SM{[R]}, $\epsilon_{\Psi 2}$ has a factor $v^{-1}$, formally at $-0.5$PN order, which means that the relative size of this correction grows towards smaller frequencies, away from merger.
 Remember however that the opening angle of the precession cone, giving the overall normalization for precession effects in the waveform, also goes to 0 as $v$ in that limit.
@@
   \includegraphics[width=.48\linewidth]{plots/trig_phase.pdf}
-  \caption{Amplitude (left panel) and phase (right panel), compared to its trigonometric polynomial representation. \SM{[Add different aligned spins here]} \SM{[color for vertical lines at fpeak]}}
-  \label{fig:trig}
+  \caption{Amplitude (left panel) and phase (right panel), compared to its trigonometric polynomial representation for the three cases listed in Table~\ref{table:precexamples}.
+The vertical lines indicate the merger frequencies.
+\SM{[color for vertical lines at fpeak]}}
+  \label{fig:trigopoly}
 \end{figure*}
@@
 
-We consider the high-frequency part of the signal above some frequency $f_{0}$, up to some maximal frequency $f_{\rm max}$ where the Fourier-domain amplitude of the signal has decayed to a negligible level.
-In our example of the equal-mass, non-spinning waveform, we take $[Mf_{0}, Mf_{\rm max}] = [0.05, 0.16]$.
-First, we eliminate a constant and a linear term in the phase by choosing another frequency central to the high-frequency range we want to represent, which we take to be $Mf_{p} = 0.1$.
-We can think of $f_{p}$ as representing roughly the frequency at merger, and of the associated time $t_{p}\equiv \tf(f_{p})$ as an estimate of the time of merger.
+We consider the high-frequency part of the signal above some frequency $f_{0}$, above which the signal has limited amplitude and phase evolution, up to some maximal frequency $f_{\rm max}$ where the Fourier-domain amplitude of the signal has decayed to a negligible level.
+In practice, we define $f_{\rm knee}$ as the peak of $f^{2}A(f)$, representing the onset of the decay in amplitude, and roughly corresponding to $\omega_{22}^{\rm QNM}/\pi$.
+We set $f_{0}\equiv 2/3f_{\rm knee}$, which is close to the merger frequency, and $f_{\rm max}$ is chosen so that the amplitude is $10^{-4}$ of the amplitude at $f_{\rm knee}$.
+We also eliminate a constant and a linear term in the phase by choosing another frequency central to the high-frequency range we want to represent, which we take to be $f_{p} = f_{\rm knee}^{2/3} f_{\rm max}^{1/3}$ with corresponding time $t_{p}\equiv \tf(f_{p})$.
 Note that the method should only be weakly sensitive to the precise choice of $f_{p}$ and $t_{p}$.
 
-Additionaly, we taper the amplitude to arrive at a flat amplitude at $f_{0}$, at the price of introducing a deviation from the amplitude of the original signal.
-We localize this deviation to $f\leq f_{1}$, with $Mf_{1} = 0.06$ in our example, and in general the trigonometric representation will be accurate to the original signal only for $f\geq f_{1}$.
-This tapering is built in practice as the discrete integral of a cosine window function.
+Additionaly, we taper the amplitude to artificially flatten the amplitude at $f_{0}$, at the price of introducing a deviation from the amplitude of the original signal.
+This tapering is built in practice as the discrete integral of a cosine window function on just the first two samples.
+We taper to 0 the last three samples before $f_{\rm max}$.
 To ensure good continuity properties, we perform an artificial symmetrization to a fictitious range $f\in [f_{0} - (f_{\rm max} - f_{0}), f_{0}]$, by imposing symmetric amplitudes and phases about $f_{0}$.
@@
 This construction is intimately related to the DFT, but we recall it for completeness.
-For a periodic function $f(x)$ defined on $x\in [x_{0}, x_{0} + \Delta x]$, and represented by $N$ samples $x_{j} = \ov{x} + j \Delta x/N$ (with $N$ chosen to satisfy the Nyquist criterion \SM{[historically should be called Shannon ?]}), we can build a trigonometric interpolant $P(x)$ as
+For a periodic function $F(x)$ defined on $x\in [x_{0}, x_{0} + \Delta x]$, and represented by $N$ samples $x_{j} = \ov{x} + j \Delta x/N$ (with $N$ chosen to satisfy the Nyquist criterion \SM{[historically should be called Shannon ?]}), we can build a trigonometric interpolant $P(x)$ as
 \be
@@
 \ee
-that will satisfy the system $P(x_{j}) = f(x_{j})$ for $j=0,\dots, N-1$.
+that will satisfy the system $P(x_{j}) = F(x_{j})$ for $j=0,\dots, N-1$.
 Here we set $M=N/2$ (assuming N is even).
-The coefficients $c_{k}$ are related to the coefficients of the inverse DFT of the series for $f$.
+The coefficients $c_{k}$ are related to the coefficients of the inverse DFT of the series for $F$.
 If we set $\omega \equiv e^{2i\pi/N}$ and define
 \be
-	y_{k} = \frac{1}{N} \sum\limits_{j=0}^{N-1} f(x_{j}) \omega^{jk} \,,
-\ee
-which is the expression of the inverse DFT in our sign convention~\eqref{}, the coefficients $c_{k}$ are given by
+	y_{k} = \frac{1}{N} \sum\limits_{j=0}^{N-1} F(x_{j}) \omega^{jk} \,,
+\ee
+which is the expression of the inverse DFT in our sign convention~\eqref{eq:defFT}, the coefficients $c_{k}$ are given by
 \begin{align}\label{eq:ckyk}
@@
 where the condition $c_{M} = c_{-M}$ is an arbitrary condition enforced to match the number of degrees of freedom.
+In practice, we also 0-pad by a factor of 2 to high frequencies.
 Fig~\ref{fig:convolsymmetrized} shows both the original high-frequency part of the signal and its reconstruction as a trigonometric polynomial.
 
-Now, assuming that we wish to compute~\eqref{eq:convolution} for a frequency $f$ such that the extent of the convolution does not extend beyond $[f_{1}, f_{\rm max}]$, we can approximate
+Now, assuming that we wish to compute~\eqref{eq:convolution} for a frequency $f$ such that the extent of the convolution does not extend beyond $[f_{0}, 2f_{\rm max} - f_{0}]$, we can approximate
 \be\label{eq:hsymtrigo}
@@
 
+The crucial point in this approach is that the convolution integral in~\eqref{eq:convolution} has support mainly on $f'<0$, which means to the right of $f$ in Fig.~\ref{fig:trigopoly} as discussed in Sec.~\ref{subsec:} \SM{[R]}.
+This means that the trigonometric representation of $\tilde{h}(f)$ can be used to compute the convolution almost all the way down to $f_{0}$, effects of the tapering aside.
+
 In terms of performance, the implementation of~\eqref{eq:ykDFT} and~\eqref{eq:resultdirectconvol} above is expected to have a very reasonable cost.
-We obtained results with a good accuracy, as shown in Fig.~\ref{fig:convolsymmetrized}, by including $M=32$ samples ($N=64$ samples in total for the artificially symmetrized signal).
+We obtained results with a good accuracy, as shown in Fig.~\ref{fig:trigopoly}, by including $M=64$ samples ($32$ useful samples before 0-padding, $N=128$ samples in total for the artificially symmetrized signal).
 The calculation described here is done only once for a given waveform and modulation function.
-The number of samples is of the same order of magnitude as the number of (logarithmically-spaced) frequency samples used to represent the Fourier-domain waveform with its amplitude and phase in the first place, thus the cost is comparable to the Taylor-like approach presented in Sec.~\ref{sec:quadphasedelay}.
+The number of samples is of the same order of magnitude as the number of (logarithmically-spaced) frequency samples used to represent the Fourier-domain waveform with its amplitude and phase in the first place, thus the cost is comparable to the Taylor-like approach presented in Sec.~\ref{sec:quadphasedelay} \SM{[R]}.
 
@@
 
-In the following, we will either present the errors obtained with the formalism of Sec.~\ref{sec:formalism} applied to the whole frequency band, including the merger and ringdown part of the signal, or cover the high-frequency range with the convolution treatment of Sec.~\ref{subsec:convolution}, with a Planckian window on $Mf\in [Mf_{0}, Mf_{1}]$ with $f_{0}=$ \SM{[complete: definition of f0, f1 ?]} for a smooth transition between the two formalisms.
+In the following, we will either present the errors obtained with the formalism of Sec.~\ref{sec:formalism} applied to the whole frequency band, including the merger and ringdown part of the signal, or cover the high-frequency range with the convolution treatment of Sec.~\ref{subsec:convolution}, with a Planckian window on $Mf\in [Mf_{0}, Mf_{\rm knee}]$ for a smooth transition between the two formalisms.
 
@@
 They show the Fourier-domain amplitude for each of the modes, both exact and from the reconstruction, the relative errors in amplitude, and the errors in phase.
-The errors are given in a relative sense, and the amplitude allow to visualize the mode hierarchy, and give an idea of how much the relative errors matter in an absolute sense.
+The errors are given in a relative sense, and the amplitude allows to visualize the mode hierarchy, and give an idea of how much the relative errors matter in an absolute sense.
 
@@
 \be\label{eq:defoverlap}
-	\left( h_{1} | h_{2} \right) = 4\text{Re} \int_{0}^{+\infty} \ud f \frac{\tilde{h}_{1} \tilde{h}_{2}^{*}}{S_{n}(f)} \,,
+	\left( h_{1} | h_{2} \right) = 4\text{Re} \int_{0}^{+\infty} \ud f \frac{\tilde{h}_{1}(f) \tilde{h}_{2}^{*}(f)}{S_{n}(f)} \,,
 \ee
@@
 
-For spin-aligned waveforms, the unfaithfulness is customarily defined as the overlap~\eqref{eq:defoverlap} above optimized over the time at coalescence and global phase of the signal.
+For spin-aligned waveforms, the unfaithfulness is customarily defined as the normalized overlap~\eqref{eq:defoverlap} above optimized over the time at coalescence and global phase of the signal.
 In the precessing case, different prescriptions coexist, differing by the extrinsic parameters that are optimized over.
@@
 Overall, we find that all the corrections we investigated, perturbative in phase, perturbative in amplitude, and the convolution treatment, are relveant in lowering the unfaithfulness.
-The only exception are second-order amplitude corrections that make little difference, and can even give worse resluts than first order in some cases.
-Using all the tools at our hands, we are able to keep the unfaithfulness of our three example waveforms at below $2.10^{-3}$ for all masses and inclinations.
+The only exception are second-order amplitude corrections that make little difference, and can even give worse results than first-order in some cases.
+Using all the tools at our hands, we are able to keep the unfaithfulness of our three example waveforms below $2.10^{-3}$ for all masses and inclinations.
 

### 2017-09-11 docs/paper_method.tex
I rearranged the introductory part as discussed and made other minor edits.

@@
 \newcommand{\SM}[1]{{\color{Red} #1}}
+\newcommand{\jgb}[1]{{\color{DarkGreen} #1}}
 
@@
 \affiliation{Gravitational Astrophysics Laboratory, NASA Goddard Space Flight Center, 8800 Greenbelt Rd., Greenbelt, MD 20771, USA}
+\affiliation{Joint Space-Science Institute, University of Maryland, College Park, MD 20742, USA}
 
@@
 In the context of signals from precessing spinning binary systems, a useful tool for building models of the waveform consists in representing the signal as a time-dependent rotation of a quasi-non-precessing waveform.
-In both cases, being able to compute transfer functions for these effects directly in the Fourier domain would allow to gain performance for data analysis applications by using fast frequency-domain waveforms.
-Our results recover previous approaches based on an extension of the stationary phase approximation, generalize them by including delays and higher-order corrections, and can in principle be applied to full inspiral-merger-ringdown signals.
+In both cases, being able to compute transfer functions for these effects directly in the Fourier domain may enable performance improvements for data analysis applications by using fast frequency-domain waveforms.
+Our results generalize previous approaches based on the stationary phase approximation, extending them by including delays and higher-order corrections while being applicable to a broader class of signals including full inspiral-merger-ringdown waveforms.
 In the LISA case, we find that higher-order corrections remain small except at the two extremes of the detector's frequency band, where we show that including them reduces the errors.
 In the case of precessing binaries, we perform a limited exploration of the merger-ringdown range based on a toy model for the precession.
-We find that, depending on the post-merger frame velocity, higher-order corrections can become quantitatively large to the point of challenging our perturbative formalism, while affecting mainly the amplitude and giving raise to a limited unfaithfulness.
-We develop an alternative direct convolution approach to represent accurately these post-merger features.
+We find that, depending on the post-merger frame velocity, higher-order corrections can become quantitatively large to the point of challenging our perturbative formalism, while affecting mainly the amplitude and giving rise to limited unfaithfulness.
+We further present an alternative direct convolution approach to accurately represent these post-merger features.
 
@@
 
-With the recent detections of gravitational-wave signals emitted by the coalescences of binary black holes, announced by the LVC collaboration after the first obervation run of the two LIGO instruments~\cite{LIGO-theevent-2016,LIGO-christmasevent-2016,LIGO-O1BBH-2016}, gravitational-wave astronomy has entered its observational era.
-With the recent start of LIGO's second observation run, and later the extension of the ground-based detectors network with Virgo~\cite{Virgo}, KAGRA~\cite{KAGRA} and LIGO-India~\cite{INDIGO}, many more detections from compact object coalescences are to come, with binary black holes being expected to be the dominant sources~\cite{} \SM{[C]}.
+With the recent detections of gravitational-wave signals emitted by the coalescences of binary black holes, announced by the LVC collaboration during the first and second obervation runs of the two LIGO instruments~\cite{LIGO-theevent-2016,LIGO-christmasevent-2016,LIGO-O1BBH-2016}\jgb{[update refs]}, gravitational-wave astronomy has entered its observational era.
+With the conclusion of LIGO's second observation run, and the recent extension of the ground-based detectors network with Virgo~\cite{Virgo} ( and later also KAGRA~\cite{KAGRA} and LIGO-India~\cite{INDIGO}), many more detections from compact object coalescences are to come, with binary black holes being expected to be the dominant sources~\cite{} \SM{[C]}.
+
 Moreover, the European Space Agency has issued a mission call for its L3 space mission with a tentative launch around 2034, adressing the science theme ``The Gravitational Universe'', possibly opening the way for a space-borne observatory based on the concept of the Laser Interferometer Space Antenna (LISA)~\cite{LISA09,LISA17}.
@@
 Data analysis for gravitational-wave observations of compact binary coalescences require accurate models (or templates) for the signals, both for ensuring efficient detections of signals that may be buried in instrumental noise, and to extract the physical parameters of the source in a subsequent analysis.
-This has prompted an important and ongoing effort to improve gravitational wave templates.
-Two important qualitative features for the improvement of the templates are the coverage of all phases of the evolution of the system, including the inspiral, merger and ringdown (IMR), which became possible thanks to the breakthrough of numerical relativity~\cite{Pretorius05,Baker+06,Campanelli+06}, and the inclusion of the effects of the spins of the compact objects, that can lead to a precession of the orbital plane~\cite{ACST94,K95}.
-Besides accuracy, the computational efficiency of template generation is crucial, as Bayesian analysis for parameter estimation of gravitational-wave signals, as was performed for the LIGO detections~\cite{LIGO-theeventPE-2016,LIGO-O1BBH-2016}, may require millions of evaluations of the likelihood function to sample the posterior probability distribution.
-
-State-of-the-art IMR templates need to combine information from the perturbative results of post-Newtonian theory covering the inspiral (see e.g.~\cite{BlanchetLiving}) and from numerical relativity simulations covering the end of the inspiral and the merger-ringdown phase (see e.g.~\cite{Pfeiffer12}).
+The the greater sensitivity of LISA and other future instruments will require further increases in the accuracy of computationally efficient signal templates.
+This motivates an important and ongoing effort to improve gravitational wave templates.
+Two important focus areas for template improvements are need to cover all phases of the evolution of the system, including the inspiral, merger and ringdown (IMR), which became possible thanks to the breakthrough of numerical relativity~\cite{Pretorius05,Baker+06,Campanelli+06}, and the inclusion of the effects of the spins of the compact objects, that can lead to a precession of the orbital plane~\cite{ACST94,K95}.
+Besides accuracy, the computational efficiency of template generation is crucial.
+Bayesian analysis for parameter estimation of gravitational-wave signals, as was performed for the LIGO detections~\cite{LIGO-theeventPE-2016,LIGO-O1BBH-2016}, may require millions of evaluations of the likelihood function to sample the posterior probability distribution.
+
+The GW community is making progress assembling higher-fidelity tools for these coming challenges.
+State-of-the-art IMR templates combine information from the perturbative results of post-Newtonian theory covering the inspiral (see e.g.~\cite{BlanchetLiving}) and from numerical relativity simulations covering the end of the inspiral and the merger-ringdown phase (see e.g.~\cite{Pfeiffer12}).
 The Effective-One-Body (EOB) approach~\cite{BD99} does so by incorporating a resummation of post-Newtonian results and a calibration to numerical relativity, and has been extended to include both spins of the two bodies~\cite{Taracchini+13, Pan+13, Bohe+16}.
 Phenomenological templates extend post-Newtonian models by postulating an analytic ansatz for the Fourier-domain amplitude and phase, fitting coefficients against numerical relativity waveforms~\cite{Husa+15, Khan+15}, and can include a simplified model for precession~\cite{Hannam+13}.
-The second type of models produce directly the amplitude and the phase of the Fourier-domain waveform in the form of a piecewise analytic function.
-EOB waveforms, by contrast, require to integrate ordinary differential equations in the time-domain, which can be time-consuming.
+The second type of models produce directly the amplitude and the phase of the Fourier-domain waveform components in the form of a piecewise analytic function.
+EOB waveforms, by contrast, require integration of ordinary differential equations in the time-domain, which can be time-consuming.
 For the non-precessing case, Reduced Order Models (ROM), also called surrogate models, have been developed to speed up considerably the waveform generation, without losing accuracy~\cite{Field+13, Puerrer14, Bohe+16}.
-Both approaches are therefore able to produce efficiently IMR Fourier-domain waveforms.
+Both approaches are therefore able to efficiently produce IMR Fourier-domain waveforms.
 An important aspect is that the resulting waveforms can be represented by an amplitude and phase for each mode, with only a few hundred samples~\cite{Puerrer14}.
 
-A complete representation of the spin effects in fast IMR templates still remains a frontier of gravitational wave signal modelling.
-In presence of misaligned spins, the system will endure a precession of the orbital plane as it evolves, leading to modulations of the signal as seen by the observer~\cite{ACST94}.
+A complete representation of spin effects in fast IMR templates still remains a frontier of gravitational wave signal modelling.
+In presence of misaligned spins, the system will endure precession of the orbital plane as it evolves, leading to modulations of the signal as seen by the observer~\cite{ACST94}.
 Precessing EOB templates incorporate all six parameters of the two spins~\cite{Pan+13, BTB16} but still require a costly time-domain integration; while precessing phenomenological templates~\cite{Hannam+13} decrease the dimensionnality by using effective spin parameters representing only the dominant effects of the spins.
@@
 The resulting waveform in the rotated frame can then be modelled by a non-precessing waveform, an approximation which is used both in the construction of the inspiral part of precessing EOB waveforms~\cite{Pan+13} and in the construction of precessing phenomenological waveforms~\cite{Hannam+13}.
-To follow this modelling approach and create efficiently Fourier-domain waveforms, one needs to understand how to translate the time-domain modulation created by the frame rotation into a Fourier-domain transfer function, which is our first motivation.
+To follow this modelling approach and efficiently create Fourier-domain waveforms, one needs to understand how to translate the time-domain modulations created by the frame rotation into a Fourier-domain transfer function, which is our first motivation.
 
 Prospective parameter recovery studies for LISA observations of comparable-mass binary black hole coalescences aim at characterizing the scientific performance of the future experiment.
-They become a timely topic with the upcoming effort for the ESA L3 mission characterization in the next few years.
+Although LISA\ref{LISAMissionWP} is still at least a decade from launch, the ESA L3 mission is rapidly progressing into its early stages where mission design choice must be informed by science impact analysis.
 However, the recent progresses in waveform modelling, most notably the availability of cheap IMR waveforms, have not yet made their way to the field of LISA data analysis studies.
@@
 Thus, due mainly to their high computational cost, full-fledged Bayesian parameter estimation simulations for IMR signals are still missing.
-
-By contrast with ground-based detectors, the response of a LISA-type instrument has the specificity of being time-dependent~\cite{Cutler97}.
-For signals possibly accumulating over several months, the motion and change of orientation of the detector constellation along its orbit lead to significant time variability in the form of a modulation and a varying delay, which contain information about the localization of the gravitational-wave source in the sky.
-This complicated detector response increases the computational cost of parameter-estimation analyses.
-To leverage the performance of state-of-the-art Fourier-domain IMR templates~\cite{BTB16,Khan+15}, one needs to be able to process the signals through the time-dependent response of the dectector while staying in the Fourier domain, which is our second motivation.
-
-The purpose of this paper is to propose a formalism allowing to transpose directly in the Fourier domain the processing of a signal through a time-domain modulation and delay, while retaining the compactness of a Fourier-domain amplitude and phase representation of the signals.
+\jgb{[Maybe there is too much detail in this paragraph?]}
+
+In contrast with current ground-based detectors, which are typically sensitive to chirping binaries for less than a second, LISA signals may accumulating over months or years.
+The response of a LISA-type instrument is thus time-dependent~\cite{Cutler97}.
+The motion and change of orientation of the detector constellation along its orbit lead to significant time variability in the form of a modulation and a varying delay.
+These effeccts then convey information about the localization of the gravitational-wave source in the sky.
+The complicated detector response increase the computational cost of parameter-estimation analyses.
+To leverage the performance of state-of-the-art Fourier-domain IMR templates~\cite{BTB16,Khan+15}, one needs to be able to process the signals through the time-dependent response of the dectector while staying in the Fourier domain.
+%, which is our second motivation.
+
+The purpose of this paper is to introduce a formalism for efficiently processing signals through a time-domain modulation and delay within the Fourier domain, while retaining the compactness of a Fourier-domain amplitude and phase representation of the signals.
 This will allow us to address both the issue of the Fourier-domain response of the LISA instrument, as well as the issue of Fourier-domain precession modulation for IMR signals from precessing binaries.
@@
 The plan of the paper is as follows.
-In Sec.~\ref{sec:motivation}, we will introduce our notations and conventions for Fourier-domain waveforms, motivate the problem of Fourier-domain modulation and introduce the relevant timescales for both the response of LISA-type detectors and the modulation of precessing signals.
+In Sec.~\ref{sec:motivation}, \sout{we will introduce our notations and conventions for Fourier-domain waveforms}, motivate the problem of Fourier-domain modulation and introduce the relevant timescales for both the response of LISA-type detectors and the modulation of precessing signals.
 In Sec.~\ref{sec:formalism}, we present our general formalism, give its leading order approximation as well as higher-order corrections, introduce new timescales based on the Fourier-domain signal, refine the previous results for both the quadratic-in-phase corrections and the treatment of the delays, and make the connection to previous approaches of the problem.
@@
 
-\subsection{Notations and conventions}
-\label{subsec:notations}
-
-The convention we will be using for the Fourier transform of a signal $h(t)$ is
-\be\label{eq:defFT}
-	\tilde{h}(f) = \mathrm{FT}[h](f) =  \int \ud t \, e^{+2i\pi f t} h(t) \,.
-\ee
-Notice that this sign convention is not the most frequently used in the literature.
-We chose it to ensure that, with the conventions of~\cite{BlanchetLiving}, spin-weighted spherical modes $h_{\ell m}$ with $m>0$ will have support mostly for positive frequencies.
-One can revert to the more usual convention by taking $f\rightarrow -f$.
-
-A useful representation of the gravitational waveform is given by its decomposition in spin-weighted spherical harmonics.
-The waveform emitted in the direction $(\Theta, \Phi)$, with its two polarizations $h_{+},h_{\times}$, can be decomposed as a superposition of modes as~\cite{Thorne80}
-\be\label{eq:defmodes}
-	h_{+} - i h_{\times} = \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} {}_{-2}Y_{\ell m}(\Theta,\Phi) h_{\ell m} \,,
-\ee
-where the ${}_{-2}Y_{\ell m}(\Theta,\Phi)$ are spin-weighted spherical harmonics~\cite{Goldberg+67}. Conversely, the individual polarizations are obtained from the individual modes as
-\begin{subequations}
-\begin{align}
-	h_{+} = \frac{1}{2} \sum\limits_{\ell, m} \left[ {}_{-2}Y_{\ell m}h_{\ell m} + {}_{-2}Y_{\ell m}^{*} h_{\ell m}^{*} \right] \,,\\
-	h_{\times} = \frac{i}{2} \sum\limits_{\ell, m} \left[ {}_{-2}Y_{\ell m}h_{\ell m} - {}_{-2}Y_{\ell m}^{*} h_{\ell m}^{*} \right] \,.
-\end{align}
-\end{subequations}
-
-For a non-precessing system, with a fixed orbital plane, the individual modes have the additional symmetry property
-\be\label{eq:symmetryhlminusm}
-	h_{\ell, -m} = (-1)^{\ell} h_{\ell m}^{*} \,.
-\ee
-Since we will work in the Fourier domain, it will be useful to write the contributions of the individual modes to the Fourier transforms of the polarizations as
-\begin{subequations}\label{eq:hpcfrommodes}
-\begin{align}
-	\tilde{h}_{+}(f) &= \frac{1}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left[ {}_{-2}Y_{\ell m} \tilde{h}_{\ell m}(f) + {}_{-2}Y_{\ell m}^{*} \tilde{h}_{\ell m}(-f)^{*} \right] \,, \\
-	\tilde{h}_{\times}(f) &= \frac{i}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left[ {}_{-2}Y_{\ell m} \tilde{h}_{\ell m}(f) - {}_{-2}Y_{\ell m}^{*} \tilde{h}_{\ell m}(-f)^{*} \right] \,,
-\end{align}
-\end{subequations}
-where we used $\widetilde{h_{\ell m}^{*}}(f) = \tilde{h}_{\ell m}(-f)^{*}$.
-An additional approximation often used in waveform models consists in considering that the Fourier transforms $\tilde{h}_{\ell m}$ have support only on one side of the spectrum, either for positive of for negative frequencies depending on the sign of $m$.
-This holds in particular within the stationary phase approximation, as explained below.
-With our sign convention~\eqref{eq:defFT}, this approximation reads
-\begin{align}\label{eq:zeronegativef}
-	\tilde{h}_{\ell m} (f) &\simeq 0 \text{ for } f<0, \; m>0 \nn\,,\\
-	\tilde{h}_{\ell m} (f) &\simeq 0 \text{ for } f>0, \; m<0 \,.
-\end{align}
-
-In the following, we will drop the mode indices and the sum over modes, and we will denote generically by $h$ an individual mode of the waveform\footnote{This notation for a generic individual mode is not to be confused with the often encountered notation $h\equiv h_{+} - i h_{\times}$.}.
-Note that, if we are considering real signals $s(t)$ (such as $h_{+}$, $h_{\times}$, or any detector output), we can focus  on the positive-frequencies range and reconstruct the negative frequencies as $\tilde{s}(-f) = \tilde{s}(f)^{*}$.
-For positive frequencies $f>0$, we will use the generic notation $h$ as
-\begin{align}\label{eq:notationhsignm}
-	h(t) &\equiv h_{\ell m}(t) \,, \quad \tilde{h}(f) \equiv \tilde{h}_{\ell m}(f) \text{ for } m>0 \,, \nn\\
-	h(t) &\equiv h_{\ell m}^{*}(t) \,, \quad \tilde{h}(f) \equiv \tilde{h}_{\ell m}(-f)^{*} \text{ for } m<0 \,.
-\end{align}
-We will further decompose $\tilde{h}$ into a Fourier-domain amplitude $A$ and a phase $\Psi$ according to
-\be\label{eq:defAPsi}
-	\tilde{h}(f) \equiv A(f) e^{-i\Psi(f)} \,.
-\ee
-
-Since we intend to work in the Fourier domain, we will use PhenomD waveforms~\cite{Khan+15}.
-These phenomenological waveforms provide analytic fits for both the Fourier-domain amplitude and phase of the 22 mode, with fit coefficients that were calibrated on EOB/NR hybrids [describe smoothing across intervals].
-Fig.~\ref{fig:ampphase} shows an example of amplitude and phase in geometric units for an equal-mass binary.
-
-%\begin{figure*}
-%  \centering
-%  \includegraphics[width=.48\linewidth]{plots/Af.pdf}
-%  \hspace{0.2cm}
-%  \includegraphics[width=.48\linewidth]{plots/Psif.pdf}
-%  \caption{Fourier-domain amplitude and phase in geometric units for an equal-mass, non-spinning system.
-The vertical line represents $\omega_{22}^{\rm peak}/2\pi$, the frequency corresponding to the time-domain frequency of the waveform at the peak amplitude.
-The starting frequency $Mf_{min} = 2.6\times 10^{-4}$ is appropriate for a 2-years observation of a $M=10^{6} M_{\odot}$ system by LISA, or for an observation by LIGO/Virgo of a $M=5.4 M_{\odot}$ system entering the instrument band at $10\Hz$.
-[add residuals of interpolation on $~300$ points, to illustrate the compression of Amp/Phase FD waveforms ?
-or show different spins/mass ratios ?] \SM{[Plot perhaps not necessary ?]}}
-%  \label{fig:ampphase}
-%\end{figure*}
-
-We recall that, with our sign convention, the effect of a positive shift in time $h' (t) = h(t- \Delta t)$ is
-\be
-	\tilde{h}'(f) = e^{2i\pi f \Delta t} \tilde{h}(f) \,.
-\ee
-If we also allow for a global shift in phase of the signal, which corresponds to a rotation of the unit vector pointing towards the observer in the frame of the gravitational waves source, a constant term can also be added (note that the effect of such a rotation gives different phase changes $e^{\pm i m \delta \phi}$ for the different modes of the radiation).
-We see that both a constant and a linear term in the Fourier-domain phase can be adjusted by changing extrinsic parameters.
-Thus, as is well known, apart from the relative phases of the modes for waveforms that include higher harmonics, the intrinsic information on the phasing is contained in the second derivative of the Fourier-domain phase $\Psi(f)$.
+\subsection{Basic considerations}
+\label{subsec:basic}
+
+We can jointly consider challenges in the contexts of both the LISA response and GW precession applications in terms of a more general problem which covers both domains.
+In each case must determine the Fourier transform of a signal $h(t)$ to which we apply a varying delay $d(t)$ before multiplication by a modulation function $F(t)$,
+\be
+\label{eq:delay-mod-defs}
+	h_{d}(t) = h(t+d(t)) \,, \quad s(t) = F(t)h_{d}(t) \,,
+\ee
+and seek an efficient way to commpute the Fourier transform $\tilde{s}(f)$, expressed by means of a Fourier-domain transfer function $\calT$ such that
+\be\label{eq:deftransfer}
+	\tilde{s}(f) \equiv \calT(f) \tilde{h}(f) \,. 
+\ee
+
+In the case of precessing binaries, the delays will be absent and the modulation functions will be the time-dependent Wigner coefficients applied to rotate the waveform from a precessing frame, where it verifies a restored mode hierarchy, to an inertial frame where the observations take place.
+In the case of a LISA-type detector, the signal $h(t)$ will simply be the waveform in a fixed heliocentric frame, and the delays will come from the motion of each detector in the wave front, while the modulation will represent the time-variation in the detector orientation.
+\jgb{MOVE:Both can be considered as periodic functions with a period of a year, which allows for an additional rewriting in terms of Fourier series instead of Fourier transforms, as will be developed in Sec.~\ref{}.NOTE: Probably move other technical partso f the par.
+as well.}
+
+\jgb{In full generality computing the transfer function $\calT(f)$ will require a convolution, a costly integral (numerically a sum) over the full frequency domain for each value of $f$.
+For our context though, and perhaps for other applications, there are some properties of $h(t)$, $d(t)$ and $F(t)$ which we can exploit for a more efficient computation.}
+
+\jgb{First consider the waveforms.
+There is a separation between the timescales of orbital motion and radiation-reaction.
+This leads to a general feature of many GW signals, including binary inspirals and black hole mergers, that the signal is relatively localized in time-frequency.
+As far as $h(f)$ is concerned then, for given $f$, only a limited range of times should be relevant in $d(t)$ and $F(t)$ so that we may expect to find a more efficient treatment for the transfer function.}
+
+\jgb{As is well known (refs) this is especially true for non-precessing systems when the signal is decomposed in spin-weighted spherical harmonic components with respect to the source orientation angles.
+In amplitude-phase form the individual spherical harmonic components of even long-lasting signals may be treated as smoothly varying functions of frequency.
+An important consequence is that these signal components can be accurately represented in terms of a relatively coarsely sampled frequency grid, but we will also use this property in transfer function treatment.
+(See Appendix~\ref{sec:notation} for details of our notation and conventions for Fourier transforms and spherical harmonics.)}
+
+\jgb{Next consider the delay and modulation functions.
+Again we can identify a separation of time scales between the modulation or delay and the gravitational-wave period.
+This is either the precession timescale or for the LISA response the annual-orbital motion timescale for the constellation.}
+
+\jgb{Our objective is to find an efficient approximate treatment for $\calT(f)$ which allows us to exploit these properties in a way which is extensible to the high-accuracies which will be required by LISA and other future gravitational-wave instruments, yet which widely applicable to GW analysis, without unnecessary limiting assumptions.
+Before introducing developing our formalism, though, we review the salient features of our application problems.}
+
+\subsection{Instrumental modulations and delays for LISA-type detectors}
+\label{subsec:modulationLISA}
+
+The response of a detector of the LISA type to an incident gravitational wave can be written in two different, equivalent forms, in terms of phase or frequency measurements.
+Here we will work with the second representation, which will prove more convenient for our purposes.
+Moreover, various notation and conventions have been used in the literature to label the spacecrafts and describe their orbits.
+We refer the reader to~\cite{Vallisneri04} for a comparative account on these various conventions.
+In this work, we will use the conventions of~\cite{Vallisneri04}, which were also used in the Mock LISA Data Challenges (MLDC)~\cite{} \SM{[C]}.
+
+We use a coordinate system centered on the solar system barycenter (SSB), and we denote the center of the constellation by $O$.
+We introduce the notation $\hat{k}$ for the propagation vector of the gravitational wave, which we denote by $h_{ij}^{\rm TT}(t)$ in transverse-traceless matrix form, as measured at the SSB (thus, at position $p$, $h_{ij}^{\rm TT}(t, p) = h_{ij}^{\rm TT}(t - \hat{k}.p)$).
+We denote by $p_{A}$ ($A=1,2,3$) the position of the spacecrafts and $n_{l}$ the unit vectors of the three links, with the convention that $n_{3}$ points from 1 to 2.
+
+Written in terms of the fractional laser frequency shifts between two spacecrafts, the elementary response of the detector reads~\cite{EW75, RCP04, Vallisneri04}
+\begin{align}\label{eq:yslr}
+	y_{slr} &\equiv \frac{\nu_{r} - \nu_{s}}{\nu} \nn\\
+	&= \frac{1}{2} \frac{n_{l}^{i}n_{l}^{j}}{1 - \hat{k}\cdot n_{l}} \left[ h_{ij}^{\rm TT}(t - \hat{k}\cdot p_{s}) - h_{ij}^{\rm TT}(t - \hat{k}\cdot p_{r}) \right] \,.
+\end{align}
+As detailed below in Sec.~\ref{}, we will use a rigid instantaneous model, approximating the geometry of the constellation by a moving equilateral triangle (neglecting the flexing of the arms induced by corrections in the orbits), and evaluating all geometric factors in~\eqref{eq:yslr} at a single time $t$ (neglecting pointing-ahead corrections).
+
+The first- and second-generation TDI observables are then built as combinations of these basic building blocks, evaluated at delayed times.
+Since in our rigid approximation these delays will take a simple form in Fourier domain, we will focus in this paper on the $y_{slr}$ observables.
+We will give in Sec.~\ref{subsec:modelLISA} more details on the response of LISA-type detectors, on the TDI observables, and on the approximations that enter the derivation of the basic response~\eqref{eq:yslr} above.
+\SM{[explicit expressions for TDI]}
+
+The structure of~\eqref{eq:yslr} is such that one can build the full signal from individual contributions of the form
+\be
+	s(t) = F(t) h(t + d(t)) \,.
+\ee
+Here, $h(t)$ represents one of the individual modes building the full gravitational wave signal (see Eq.~\eqref{eq:defmodes}), $d(t)$ represents the time-varying delays of the form $-\hat{k}\cdot p_{A}(t)$, and $F(t)$ incorporates all the relevant geometric prefactors.
+
+The functions $F(t)$ and $d(t)$ vary on a timescale of one year.
+When neglecting small non-periodic orbital perturbations (such as those caused by the influence of the Earth and of the other planets on the orbits), both $F(t)$ and $d(t)$ are periodic.
+It will also be useful to separate the delays in two types of terms: the first, $d_{O} = -\hat{k}\cdot p_{O}$, relates the waveform at the SSB to the waveform at the center of the constellation $O$, whereas the second, $d_{c}$, represents the various delays between the spacecrafts of the constellation.
+For an orbit radius $R=1\,\mathrm{au}$ and an armlength $L=5.10^{6}\mathrm{km}$, $d_{O} \sim R/c \sim 500\mathrm{s}$ and $d_{c} \sim L/c \sim 17\mathrm{s}$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Precession modulation for spinning binaries}
+\label{subsec:modulationPrec}
+
+For spinning compact objects, the presence of the spins leads to the precession of the orbit~\cite{Apostolatos+94, Kidder95}, which can have large effects on the waveform.
+In particular, it breaks the planar symmetry of the gravitational wave emission, and leads to modulations that are especially important for systems that are observed edge-on.
+
+It was suggested by a number of authors~\cite{} \SM{[C]} that the effect of the precession can be modeled, to a good approximation, by a time-dependent rotation of a non-precessing waveform.
+In this approach, one introduces a precessing frame following the evolution of the plane of the orbit, and approximates the waveform in this precessing frame by using a non-precessing model.
+Different prescriptions have been proposed for the construction of the precessing frame itself~\cite{} \SM{[C]}.
+
+If $(\alpha, \beta, \gamma)$ are the Euler angles relating the precessing frame to the inertial frame in the $(z,y,z)$ convention, the modes in the inertial frame $h_{\ell m}^{\rm I}$ are then related to the modes in the precessing frame $h_{\ell m}^{\rm P}$ by~\cite{Goldberg+67}
+\be\label{eq:wignerrotintro}
+	h_{\ell m}^{\rm I} = \sum\limits_{m=-\ell}^{\ell} \calD^{\ell *}_{mm'} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm P} \,,
+\ee
+where the $\calD^{\ell}_{mm'}$ are Wigner matrices~\cite{} \SM{[C]}. Notice that there is no mixing of the modes with different values of $\ell$.
+
+If we make the assumption that the precessing-frame waveform $h^{\rm P}$ is approximated by a non-precessing model that provides us with a smooth Fourier-domain amplitude and phase, then the problem reduces to computing the Fourier transform of the signal
+\be\label{eq:defmodulationprec}
+	s(t) = F(t) h(t) \,,
+\ee
+where the modulation function $F(t)$ is given by a Wigner matrix and depends on time through the Euler angles $(\alpha, \beta, \gamma)(t)$.
+We will discuss more in Sec.~\ref{subsec:modelPrec} \SM{[R]} the limitations of the assumptions used to perform the decomposition above.
+
+In Eq.~\eqref{eq:defmodulationprec} above, the modulation function $F$ has time variations on the precessional timescale, which evolves throughout the inspiral.
+The separation between the precessional timescale and the orbital timescale will be crucial for our analysis.
+In the limit of low frequencies, we will see in Sec.~\ref{} \SM{[R]} below that, although these timescales become more and more separated, the decrease in the chirping rate gives raise to a corrective contribution that does not vanish in this limit.
+In the other limit, as the system gets close to merger, the separation of timescales becomes weaker, and we will explore in Sec.~\ref{} \SM{[R]} the application of our formalism to the merger and ringdown phase.
 
@@
 
-We recall here an approximation widely used to model gravitational waves signals emitted by inspiralling binaries, the Stationary Phase Approximation (thereafter SPA).
+As a preliminary step, we recall here an approximation widely used for similar purposes in treating gravitational waves signals emitted by inspiralling binaries, the Stationary Phase Approximation (thereafter SPA).
 It applies in general to chirping signals, and we refer the reader to~\ref{} for details.
 One first writes the time-domain signal in amplitude and phase form as $h(t) = a(t) e^{-2i\varphi(t)}$, where for gravitational wave signals $\varphi$ will correspond to the orbital phase, with the orbital frequency being $\omega = \dot{\varphi}$.
-In order to keep close to the notations used in the gravitational wave literature, we introduced a factor of 2 in the phase of the wave, which is appropriate for the dominant 22 harmonic of the signal.
+In order to keep close to the notation used in the gravitational wave literature, we introduced a factor of 2 in the phase of the wave, which is appropriate for the dominant 22 harmonic of the signal.
 The approximation then applies to signals verifying the conditions \SM{[say that the third is sometimes missing ?]}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Instrumental modulations and delays for LISA-type detectors}
-\label{subsec:modulationLISA}
-
-The response of a detector of the LISA type to an incident gravitational wave can be written in two different, equivalent forms, in terms of phase or frequency measurements.
-Here we will work with the second representation, which will prove more convenient for our purposes.
-Moreover, various notations and conventions have been used in the literature to label the spacecrafts and describe their orbits.
-We refer the reader to~\cite{Vallisneri04} for a comparative account on these various conventions.
-In this work, we will use the conventions of~\cite{Vallisneri04}, which were also used in the Mock LISA Data Challenges (MLDC)~\cite{} \SM{[C]}.
-
-We use a coordinate system centered on the solar system barycenter (SSB), and we denote the center of the constellation by $O$.
-We introduce the notation $\hat{k}$ for the propagation vector of the gravitational wave, which we denote by $h_{ij}^{\rm TT}(t)$ in transverse-traceless matrix form, as measured at the SSB (thus, at position $p$, $h_{ij}^{\rm TT}(t, p) = h_{ij}^{\rm TT}(t - \hat{k}.p)$).
-We denote by $p_{A}$ ($A=1,2,3$) the position of the spacecrafts and $n_{l}$ the unit vectors of the three links, with the convention that $n_{3}$ points from 1 to 2.
-
-Written in terms of the fractional laser frequency shifts between two spacecrafts, the elementary response of the detector reads~\cite{EW75, RCP04, Vallisneri04}
-\begin{align}\label{eq:yslr}
-	y_{slr} &\equiv \frac{\nu_{r} - \nu_{s}}{\nu} \nn\\
-	&= \frac{1}{2} \frac{n_{l}^{i}n_{l}^{j}}{1 - \hat{k}\cdot n_{l}} \left[ h_{ij}^{\rm TT}(t - \hat{k}\cdot p_{s}) - h_{ij}^{\rm TT}(t - \hat{k}\cdot p_{r}) \right] \,.
-\end{align}
-As detailed below in Sec.~\ref{}, we will use a rigid instantaneous model, approximating the geometry of the constellation by a moving equilateral triangle (neglecting the flexing of the arms induced by corrections in the orbits), and evaluating all geometric factors in~\eqref{eq:yslr} at a single time $t$ (neglecting pointing-ahead corrections).
-
-The first- and second-generation TDI observables are then built as combinations of these basic building blocks, evaluated at delayed times.
-Since in our rigid approximation these delays will take a simple form in Fourier domain, we will focus in this paper on the $y_{slr}$ observables.
-We will give in Sec.~\ref{subsec:modelLISA} more details on the response of LISA-type detectors, on the TDI observables, and on the approximations that enter the derivation of the basic response~\eqref{eq:yslr} above.
-\SM{[explicit expressions for TDI]}
-
-The structure of~\eqref{eq:yslr} is such that one can build the full signal from individual contributions of the form
-\be
-	s(t) = F(t) h(t + d(t)) \,.
-\ee
-Here, $h(t)$ represents one of the individual modes building the full gravitational wave signal (see Eq.~\eqref{eq:defmodes}), $d(t)$ represents the time-varying delays of the form $-\hat{k}\cdot p_{A}(t)$, and $F(t)$ incorporates all the relevant geometric prefactors.
-
-The functions $F(t)$ and $d(t)$ vary on a timescale of one year.
-When neglecting small non-periodic orbital perturbations (such as those caused by the influence of the Earth and of the other planets on the orbits), both $F(t)$ and $d(t)$ are periodic.
-It will also be useful to separate the delays in two types of terms: the first, $d_{O} = -\hat{k}\cdot p_{O}$, relates the waveform at the SSB to the waveform at the center of the constellation $O$, whereas the second, $d_{c}$, represents the various delays between the spacecrafts of the constellation.
-For an orbit radius $R=1\,\mathrm{au}$ and an armlength $L=5.10^{6}\mathrm{km}$, $d_{O} \sim R/c \sim 500\mathrm{s}$ and $d_{c} \sim L/c \sim 17\mathrm{s}$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Precession modulation for spinning binaries}
-\label{subsec:modulationPrec}
-
-For spinning compact objects, the presence of the spins leads to the precession of the orbit~\cite{Apostolatos+94, Kidder95}, which can have large effects on the waveform.
-In particular, it breaks the planar symmetry of the gravitational wave emission, and leads to modulations that are especially important for systems that are observed edge-on.
-
-It was suggested by a number of authors~\cite{} \SM{[C]} that the effect of the precession can be modeled, to a good approximation, by a time-dependent rotation of a non-precessing waveform.
-In this approach, one introduces a precessing frame following the evolution of the plane of the orbit, and approximates the waveform in this precessing frame by using a non-precessing model.
-Different prescriptions have been proposed for the construction of the precessing frame itself~\cite{} \SM{[C]}.
-
-If $(\alpha, \beta, \gamma)$ are the Euler angles relating the precessing frame to the inertial frame in the $(z,y,z)$ convention, the modes in the inertial frame $h_{\ell m}^{\rm I}$ are then related to the modes in the precessing frame $h_{\ell m}^{\rm P}$ by~\cite{Goldberg+67}
-\be\label{eq:wignerrotintro}
-	h_{\ell m}^{\rm I} = \sum\limits_{m=-\ell}^{\ell} \calD^{\ell *}_{mm'} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm P} \,,
-\ee
-where the $\calD^{\ell}_{mm'}$ are Wigner matrices~\cite{} \SM{[C]}. Notice that there is no mixing of the modes with different values of $\ell$.
-
-If we make the assumption that the precessing-frame waveform $h^{\rm P}$ is approximated by a non-precessing model that provides us with a smooth Fourier-domain amplitude and phase, then the problem reduces to computing the Fourier transform of the signal
-\be\label{eq:defmodulationprec}
-	s(t) = F(t) h(t) \,,
-\ee
-where the modulation function $F(t)$ is given by a Wigner matrix and depends on time through the Euler angles $(\alpha, \beta, \gamma)(t)$.
-We will discuss more in Sec.~\ref{subsec:modelPrec} \SM{[R]} the limitations of the assumptions used to perform the decomposition above.
-
-In Eq.~\eqref{eq:defmodulationprec} above, the modulation function $F$ has time variations on the precessional timescale, which evolves throughout the inspiral.
-The separation between the precessional timescale and the orbital timescale will be crucial for our analysis.
-In the limit of low frequencies, we will see in Sec.~\ref{} \SM{[R]} below that, although these timescales become more and more separated, the decrease in the chirping rate gives raise to a corrective contribution that does not vanish in this limit.
-In the other limit, as the system gets close to merger, the separation of timescales becomes weaker, and we will explore in Sec.~\ref{} \SM{[R]} the application of our formalism to the merger and ringdown phase.
+\jgb{In the Newtonian, low-frequency limit each of the combinations \eqref{eq:conditionsSPA} assumed to be small in the SPA limit indeed vanishes at $\calO{((v/c)^5)}$, but as the system approaches merger, this condition is no-longer satisfied.
+Near-merger however the treatment is not applicable, so that this is not a viable approach for binary systems beyond insprial.
+Nonetheless the SPA, has been a useful workhorse in many gravitational analysis involving frequency domain transfer functions of the form \eqref{eq:deftransfer}.
+The usual approach is simply to replace any time dependencies appearing the transfer function using \eqref{eq:deftfSPA}.
+We will later return to further consider the SPA treatment in the context of a more general formalism.}
 
@@
 
+\jgb{In this section we develop a Fourier-domain formalism for treating time-delays and temporally multiplicative signal transformations, where the period of primary signal oscilations is short compared to the timescale on which the modulating functions vary.}
+
 \subsection{Fourier transform of a modulated and delayed signal}
@@
 
-Let us consider, in general, the problem of determining the Fourier transform of a signal $h(t)$ to which we apply a varying delay $d(t)$ and a multiplication by a modulation function $F(t)$.
-In the case of precessing binaries, the delays will be absent, and the modulation functions will be the time-dependent Wigner coefficients applied to rotate the waveform from a precessing frame, where it verifies a restored mode hierarchy, to an inertial frame where the observations take place.
-In the case of a LISA-type detector, the signal $h(t)$ will simply be the waveform in a fixed heliocentric frame, and the delays will come from the motion of each detector in the wave front, while the modulation will represent the time-variation in the detector orientation.
-Both can be considered as periodic functions with a period of a year, which allows for an additional rewriting in terms of Fourier series instead of Fourier transforms, as will be developed in Sec.~\ref{}.
-
-Defining
-\be
-	h_{d}(t) = h(t+d(t)) \,, \quad s(t) = F(t)h_{d}(t) \,,
-\ee
-our purpose is to derive the Fourier transform $\tilde{s}(f)$, expressed by means of a Fourier-domain transfer function $\calT$ such that
-\be\label{eq:deftransfer}
-	\tilde{s}(f) \equiv \calT(f) \tilde{h}(f) \,. 
-\ee 
-We can first write
-\be
-	h_{d}(t) = \int \ud f \, e^{-2i\pi f (t+d(t))}\tilde{h}(f) \,,
-\ee
+\jgb{We begin by expression our delay and modulation function definitions in \eqref{eq:delay-mod-defs} using the Fourier transform (note conventions in \eqref{eq:defFT})},
+\begin{align}
+h_{d}(t) & =  h(t+d(t)) \nonumber\\
+&=\int \ud f \, e^{-2i\pi f (t+d(t))}\tilde{h}(f) \,,
+\end{align}
 and
 \begin{align}
-	\tilde{s}(f) &= \int \ud t \, e^{2i\pi f t} F(t)  \int \ud f' \, e^{-2i\pi f' (t+d(t))}\tilde{h}(f') \nn\\
+  \tilde{s}(f) &= F(t)h_{d}(t) \nn \\
+  &= \int \ud t \, e^{2i\pi f t} F(t)  \int \ud f' \, e^{-2i\pi f' (t+d(t))}\tilde{h}(f') \nn\\
 	&= \int \ud f' \, \tilde{h}(f-f') \int \ud t \, e^{2i\pi f' t} e^{-2i\pi (f-f') d(t)} F(t) \,.
@@
 	\tilde{G}(f,f') &= \int \ud t \, e^{2i\pi f' t} e^{-2i\pi f d(t)} F(t) \,.
+\label{eq:defGtilde}
 \end{align}
@@
 
-In our approach, two assumptions will be crucial for our approximations to be valid:
+\jgb{A direct computation of \eqref{eq:FDkernel} using \eqref{eq:defGtilde} will generally be computationally demanding, but when $\tilde{G}(f,f')$ has only limited support over $f'$, we can seek an accurate but more efficient approximation.} In our approach, two assumptions will be crucial for our approximations to be valid:
 \begin{itemize}
-	\item the time scales entering the signal $h(t)$ are well separated from the ones entering the modulation and the delay, so that $\tilde{G}(f-f',f')$ has support limited to $|f'| < f'_{\rm max}$, with $f'_{\rm max}$ small compared the characteristic scale for the variation of the Fourier-domain signal;
-	\item the original signal has a smooth amplitude and phase in the Fourier-domain, slowly varying over the scale $f'_{\rm max}$, so that a local treatment in $f'$ of the convolution in~\eqref{eq:FDkernel} will be possible.
+	\item The time scales entering the signal $h(t)$ are well separated from the ones entering the modulation and the delay, so that $\tilde{G}(f-f',f')$ has support limited to $|f'| < f'_{\rm max}$, with $f'_{\rm max}$ small compared the characteristic scale for the variation of the Fourier-domain signal;
+	\item The original signal has a smooth amplitude and phase in the Fourier-domain, slowly varying over the scale $f'_{\rm max}$, so that a local treatment in $f'$ of the convolution in~\eqref{eq:FDkernel} will be possible.
 \end{itemize}
@@
 where we used the fact that, in our rigid approximation for the trajectory of the constellation, $p^{L}_{r} - p^{L}_{s} =  L n_{l}$.
-In the case $m<0$, the formalism can be applied in the same way, except that notations are complicated by the fact that the mode $\tilde{h}_{\ell m}$ has to be evaluated at $-f$.
+In the case $m<0$, the formalism can be applied in the same way, except that notation is complicated by the fact that the mode $\tilde{h}_{\ell m}$ has to be evaluated at $-f$.
 In the following, we will limit ourselves to the case $m>0$ for simplicity of presentation.
@@
 
-In the notations of our formalism of Sec.~\ref{sec:formalism}, 
+In the notation of our formalism of Sec.~\ref{sec:formalism}, 
 \be\label{eq:defmodulationprec}
@@
 
-In the following, we will use the notations $m=m_{1}+m_{2}$, $\nu=m_{1}m_{2}/m^{2}$, $\delta = (m_{1}-m_{2})/m$, $|\bm{S}_{A}|=Gm_{A}^{2} \chi_{A}$ for $A=1,2$.
+In the following, we will use the notation $m=m_{1}+m_{2}$, $\nu=m_{1}m_{2}/m^{2}$, $\delta = (m_{1}-m_{2})/m$, $|\bm{S}_{A}|=Gm_{A}^{2} \chi_{A}$ for $A=1,2$.
 We take the convention $m_{1} \geq m_{2}$.
@@
 \ee
-For convenience, we will use the notations
+For convenience, we will use the notation
 \begin{subequations}
@@
 The factor $\xi$ shows how the contribution of the spin to the total angular momentum affects the precession rate.
-With these notations,
+With this notation,
 \be\label{eq:Omegaprec}
@@
 
+\section{Notation and conventions}
+\label{sec:notation}
+
+The convention we will be using for the Fourier transform of a signal $h(t)$ is
+\be\label{eq:defFT}
+	\tilde{h}(f) = \mathrm{FT}[h](f) =  \int \ud t \, e^{+2i\pi f t} h(t) \,.
+\ee
+Notice that this sign convention is not the most frequently used in the literature.
+We chose it to ensure that, with the conventions of~\cite{BlanchetLiving}, spin-weighted spherical modes $h_{\ell m}$ with $m>0$ will have support mostly for positive frequencies.
+One can revert to the more usual convention by taking $f\rightarrow -f$.
+
+A useful representation of the gravitational waveform is given by its decomposition in spin-weighted spherical harmonics.
+The waveform emitted in the direction $(\Theta, \Phi)$, with its two polarizations $h_{+},h_{\times}$, can be decomposed as a superposition of modes as~\cite{Thorne80}
+\be\label{eq:defmodes}
+	h_{+} - i h_{\times} = \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} {}_{-2}Y_{\ell m}(\Theta,\Phi) h_{\ell m} \,,
+\ee
+where the ${}_{-2}Y_{\ell m}(\Theta,\Phi)$ are spin-weighted spherical harmonics~\cite{Goldberg+67}. Conversely, the individual polarizations are obtained from the individual modes as
+\begin{subequations}
+\begin{align}
+	h_{+} = \frac{1}{2} \sum\limits_{\ell, m} \left[ {}_{-2}Y_{\ell m}h_{\ell m} + {}_{-2}Y_{\ell m}^{*} h_{\ell m}^{*} \right] \,,\\
+	h_{\times} = \frac{i}{2} \sum\limits_{\ell, m} \left[ {}_{-2}Y_{\ell m}h_{\ell m} - {}_{-2}Y_{\ell m}^{*} h_{\ell m}^{*} \right] \,.
+\end{align}
+\end{subequations}
+
+For a non-precessing system, with a fixed orbital plane, the individual modes have the additional symmetry property
+\be\label{eq:symmetryhlminusm}
+	h_{\ell, -m} = (-1)^{\ell} h_{\ell m}^{*} \,.
+\ee
+Since we will work in the Fourier domain, it will be useful to write the contributions of the individual modes to the Fourier transforms of the polarizations as
+\begin{subequations}\label{eq:hpcfrommodes}
+\begin{align}
+	\tilde{h}_{+}(f) &= \frac{1}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left[ {}_{-2}Y_{\ell m} \tilde{h}_{\ell m}(f) + {}_{-2}Y_{\ell m}^{*} \tilde{h}_{\ell m}(-f)^{*} \right] \,, \\
+	\tilde{h}_{\times}(f) &= \frac{i}{2} \sum\limits_{\ell \geq 2} \sum\limits_{m=-\ell}^{\ell} \left[ {}_{-2}Y_{\ell m} \tilde{h}_{\ell m}(f) - {}_{-2}Y_{\ell m}^{*} \tilde{h}_{\ell m}(-f)^{*} \right] \,,
+\end{align}
+\end{subequations}
+where we used $\widetilde{h_{\ell m}^{*}}(f) = \tilde{h}_{\ell m}(-f)^{*}$.
+An additional approximation often used in waveform models consists in considering that the Fourier transforms $\tilde{h}_{\ell m}$ have support only on one side of the spectrum, either for positive of for negative frequencies depending on the sign of $m$.
+This holds in particular within the stationary phase approximation, as explained below.
+With our sign convention~\eqref{eq:defFT}, this approximation reads
+\begin{align}\label{eq:zeronegativef}
+	\tilde{h}_{\ell m} (f) &\simeq 0 \text{ for } f<0, \; m>0 \nn\,,\\
+	\tilde{h}_{\ell m} (f) &\simeq 0 \text{ for } f>0, \; m<0 \,.
+\end{align}
+
+In the following, we will drop the mode indices and the sum over modes, and we will denote generically by $h$ an individual mode of the waveform\footnote{This notation for a generic individual mode is not to be confused with the often encountered notation $h\equiv h_{+} - i h_{\times}$.}.
+Note that, if we are considering real signals $s(t)$ (such as $h_{+}$, $h_{\times}$, or any detector output), we can focus  on the positive-frequencies range and reconstruct the negative frequencies as $\tilde{s}(-f) = \tilde{s}(f)^{*}$.
+For positive frequencies $f>0$, we will use the generic notation $h$ as
+\begin{align}\label{eq:notationhsignm}
+	h(t) &\equiv h_{\ell m}(t) \,, \quad \tilde{h}(f) \equiv \tilde{h}_{\ell m}(f) \text{ for } m>0 \,, \nn\\
+	h(t) &\equiv h_{\ell m}^{*}(t) \,, \quad \tilde{h}(f) \equiv \tilde{h}_{\ell m}(-f)^{*} \text{ for } m<0 \,.
+\end{align}
+We will further decompose $\tilde{h}$ into a Fourier-domain amplitude $A$ and a phase $\Psi$ according to
+\be\label{eq:defAPsi}
+	\tilde{h}(f) \equiv A(f) e^{-i\Psi(f)} \,.
+\ee
+
+Since we intend to work in the Fourier domain, we will use PhenomD waveforms~\cite{Khan+15}.
+These phenomenological waveforms provide analytic fits for both the Fourier-domain amplitude and phase of the 22 mode, with fit coefficients that were calibrated on EOB/NR hybrids [describe smoothing across intervals].
+Fig.~\ref{fig:ampphase} shows an example of amplitude and phase in geometric units for an equal-mass binary.
+
+%\begin{figure*}
+%  \centering
+%  \includegraphics[width=.48\linewidth]{plots/Af.pdf}
+%  \hspace{0.2cm}
+%  \includegraphics[width=.48\linewidth]{plots/Psif.pdf}
+%  \caption{Fourier-domain amplitude and phase in geometric units for an equal-mass, non-spinning system.
+The vertical line represents $\omega_{22}^{\rm peak}/2\pi$, the frequency corresponding to the time-domain frequency of the waveform at the peak amplitude.
+The starting frequency $Mf_{min} = 2.6\times 10^{-4}$ is appropriate for a 2-years observation of a $M=10^{6} M_{\odot}$ system by LISA, or for an observation by LIGO/Virgo of a $M=5.4 M_{\odot}$ system entering the instrument band at $10\Hz$.
+[add residuals of interpolation on $~300$ points, to illustrate the compression of Amp/Phase FD waveforms ?
+or show different spins/mass ratios ?] \SM{[Plot perhaps not necessary ?]}}
+%  \label{fig:ampphase}
+%\end{figure*}
+
+We recall that, with our sign convention, the effect of a positive shift in time $h' (t) = h(t- \Delta t)$ is
+\be
+	\tilde{h}'(f) = e^{2i\pi f \Delta t} \tilde{h}(f) \,.
+\ee
+If we also allow for a global shift in phase of the signal, which corresponds to a rotation of the unit vector pointing towards the observer in the frame of the gravitational waves source, a constant term can also be added (note that the effect of such a rotation gives different phase changes $e^{\pm i m \delta \phi}$ for the different modes of the radiation).
+We see that both a constant and a linear term in the Fourier-domain phase can be adjusted by changing extrinsic parameters.
+Thus, as is well known, apart from the relative phases of the modes for waveforms that include higher harmonics, the intrinsic information on the phasing is contained in the second derivative of the Fourier-domain phase $\Psi(f)$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Conventions for the Wigner matrices and precessing frame}

### 2017-10-20 docs/paper_method.tex
Edits to the paper

@@
             citecolor=RedViolet]{hyperref}
+
+\usepackage{showlabels}
 
@@
 \newcommand{\nn}{\nonumber}
-\newcommand{\hatk}{\hat{k}}
+%\newcommand{\hatk}{\hat{k}}
+\newcommand{\hatk}{k}
 \newcommand{\Hz}{\,\mathrm{Hz}}
@@
 
-Moreover, the European Space Agency has issued a mission call for its L3 space mission with a tentative launch around 2034, adressing the science theme ``The Gravitational Universe'', possibly opening the way for a space-borne observatory based on the concept of the Laser Interferometer Space Antenna (LISA)~\cite{LISA09,LISA17}.
+Moreover, the European Space Agency has issued a mission call \SM{[update]} for its L3 space mission with a tentative launch around 2034, adressing the science theme ``The Gravitational Universe'', possibly opening the way for a space-borne observatory based on the concept of the Laser Interferometer Space Antenna (LISA)~\cite{LISA09,LISA17}.
 Among other sources of gravitational waves, LISA will be able to detect and characterize comparable-mass binary black hole coalescences to cosmological distances over a wide range of masses, from high-redshift observations of supermassive black hole binaries with $M\sim 10^{7} \Msol$ down to the observation of LIGO-type sources with $M\sim 10^{2} \Msol$~\cite{Sesana16}.
@@
 Bayesian analysis for parameter estimation of gravitational-wave signals, as was performed for the LIGO detections~\cite{LIGO-theeventPE-2016,LIGO-O1BBH-2016}, may require millions of evaluations of the likelihood function to sample the posterior probability distribution.
+\SM{[objective: accuracy, not necessary necessary for LIGO now but in the future yes]}
 
@@
 Prospective parameter recovery studies for LISA observations of comparable-mass binary black hole coalescences aim at characterizing the scientific performance of the future experiment.
-Although LISA\ref{LISAMissionWP} is still at least a decade from launch, the ESA L3 mission is rapidly progressing into its early stages where mission design choice must be informed by science impact analysis.
+Although LISA~\cite{LISA17} is still at least a decade from launch, the ESA L3 mission is rapidly progressing into its early stages where mission design choices must be informed by science impact analysis.
+However, the recent progresses in waveform modelling, most notably the availability of fast and accurate IMR waveforms, have not yet made their way to the field of LISA data analysis studies.
+In the past, many such studies used post-Newtonian, inspiral-only models, and were limited to the cheaper Fisher-matrix formalism instead of a full Bayesian analysis (among numerous works, see e.g.
+\cite{Cutler97,LISAPE09}).
+Refs.~\cite{McWilliams+09,McWilliams+10,McWilliams+11} explored the effects of including the merger and ringdown, keeping the mass parameters fixed.
+However, the Fisher matrix analysis, although considerably cheaper to perform, is valid only in the limit of a high signal to noise ratio (SNR).
+Several implementations of Bayesian analyses were developed~\cite{CP06,Wickham+06,Babak+08,GP09,Feroz+09,Petiteau+10,PC13} (see also~\cite{MLDC4} and references therein), with Ref.~\cite{PC15} proposing a direct comparison of Fisher and Bayes analyses, but they were limited to inspiral-only waveforms.
+One exception was Ref.~\cite{Babak+08} where IMR signals were used in a Bayesian study, but there the masses of the black holes were kept fixed.
+In the recent Ref.~\cite{Klein+15}, the limitation to inspiral-only precessing post-Newtonian waveforms was compensated by an SNR reweighting procedure.
+Thus, due mainly to their high computational cost, full-fledged Bayesian parameter estimation simulations for IMR signals are still missing.
+\jgb{[Maybe there is too much detail in this paragraph?]}
+
+%Prospective parameter recovery studies for LISA observations of comparable-mass binary black hole coalescences aim at characterizing the scientific performance of the future experiment.
+Although LISA~\cite{LISA17} is still at least a decade from launch, the ESA L3 mission is rapidly progressing into its early stages where mission design choice must be informed by science impact analysis.
 However, the recent progresses in waveform modelling, most notably the availability of cheap IMR waveforms, have not yet made their way to the field of LISA data analysis studies.
@@
 To leverage the performance of state-of-the-art Fourier-domain IMR templates~\cite{BTB16,Khan+15}, one needs to be able to process the signals through the time-dependent response of the dectector while staying in the Fourier domain.
-%, which is our second motivation.
 
@@
 Another intuitive approach consists in using a time-to-frequency correspondence, such as the one given by the SPA, to directly carry over the modulation to the Fourier domain.
-This corresponds, in a sense to be detailed below, to the zeroth-order approximation of the SUA, and it is used implicitly\footnote{With the difference that directly frequency-dependent expressions for the modulation are used, which are then extended to also cover high frequencies in the merger-ringdown part of the signal, as we will discuss below.} for the simplified precession of the phenomenological waveforms of Ref.~\cite{Hannam+13}.
+This corresponds, in a sense to be detailed below, to the zeroth-order approximation of the SUA, and it is used implicitly\footnote{With the difference that the expressions used for the modulation are given in terms of frequency directly for the inspiral, and are then extended to also cover high frequencies in the merger-ringdown part of the signal, as we will discuss below.} for the simplified precession of the phenomenological waveforms of Ref.~\cite{Hannam+13}.
 In the case of the LISA response, the SPA provides a natural map from time-domain (for the orbit) to Fourier-domain (for the signal)~\cite{Cutler97}.
@@
 The plan of the paper is as follows.
-In Sec.~\ref{sec:motivation}, \sout{we will introduce our notations and conventions for Fourier-domain waveforms}, motivate the problem of Fourier-domain modulation and introduce the relevant timescales for both the response of LISA-type detectors and the modulation of precessing signals.
+In Sec.~\ref{sec:motivation}, we motivate the problem of Fourier-domain modulation and introduce the relevant timescales for both the response of LISA-type detectors and the modulation of precessing signals.
 In Sec.~\ref{sec:formalism}, we present our general formalism, give its leading order approximation as well as higher-order corrections, introduce new timescales based on the Fourier-domain signal, refine the previous results for both the quadratic-in-phase corrections and the treatment of the delays, and make the connection to previous approaches of the problem.
@@
 
+\SM{[introduce three timescales: orbital, Trr, 1/f, and say what is to be compared ]}
+
 \subsection{Instrumental modulations and delays for LISA-type detectors}
@@
 We refer the reader to~\cite{Vallisneri04} for a comparative account on these various conventions.
-In this work, we will use the conventions of~\cite{Vallisneri04}, which were also used in the Mock LISA Data Challenges (MLDC)~\cite{} \SM{[C]}.
+In this work, we will use the conventions of~\cite{Vallisneri04}, which were also used in the Mock LISA Data Challenges (MLDC)~\cite{MLDC4}.
 
 We use a coordinate system centered on the solar system barycenter (SSB), and we denote the center of the constellation by $O$.
-We introduce the notation $\hat{k}$ for the propagation vector of the gravitational wave, which we denote by $h_{ij}^{\rm TT}(t)$ in transverse-traceless matrix form, as measured at the SSB (thus, at position $p$, $h_{ij}^{\rm TT}(t, p) = h_{ij}^{\rm TT}(t - \hat{k}.p)$).
+We introduce the notation $\hatk$ for the propagation vector of the gravitational wave, which we denote by $h_{ij}^{\rm TT}(t)$ in transverse-traceless matrix form, as measured at the SSB (thus, at position $p$, $h_{ij}^{\rm TT}(t, p) = h_{ij}^{\rm TT}(t - \hatk.p)$).
 We denote by $p_{A}$ ($A=1,2,3$) the position of the spacecrafts and $n_{l}$ the unit vectors of the three links, with the convention that $n_{3}$ points from 1 to 2.
@@
 	y_{slr} &\equiv \frac{\nu_{r} - \nu_{s}}{\nu} \nn\\
-	&= \frac{1}{2} \frac{n_{l}^{i}n_{l}^{j}}{1 - \hat{k}\cdot n_{l}} \left[ h_{ij}^{\rm TT}(t - \hat{k}\cdot p_{s}) - h_{ij}^{\rm TT}(t - \hat{k}\cdot p_{r}) \right] \,.
+	&= \frac{1}{2} \frac{n_{l}^{i}n_{l}^{j}}{1 - \hatk\cdot n_{l}} \left[ h_{ij}^{\rm TT}(t - \hatk\cdot p_{s}) - h_{ij}^{\rm TT}(t - \hatk\cdot p_{r}) \right] \,.
 \end{align}
@@
 We will give in Sec.~\ref{subsec:modelLISA} more details on the response of LISA-type detectors, on the TDI observables, and on the approximations that enter the derivation of the basic response~\eqref{eq:yslr} above.
-\SM{[explicit expressions for TDI]}
 
@@
 \ee
-Here, $h(t)$ represents one of the individual modes building the full gravitational wave signal (see Eq.~\eqref{eq:defmodes}), $d(t)$ represents the time-varying delays of the form $-\hat{k}\cdot p_{A}(t)$, and $F(t)$ incorporates all the relevant geometric prefactors.
+Here, $h(t)$ represents one of the individual modes building the full gravitational wave signal (see Eq.~\eqref{eq:defmodes}), $d(t)$ represents the time-varying delays of the form $-\hatk\cdot p_{A}(t)$, and $F(t)$ incorporates all the relevant geometric prefactors.
 
@@
 When neglecting small non-periodic orbital perturbations (such as those caused by the influence of the Earth and of the other planets on the orbits), both $F(t)$ and $d(t)$ are periodic.
-It will also be useful to separate the delays in two types of terms: the first, $d_{O} = -\hat{k}\cdot p_{O}$, relates the waveform at the SSB to the waveform at the center of the constellation $O$, whereas the second, $d_{c}$, represents the various delays between the spacecrafts of the constellation.
-For an orbit radius $R=1\,\mathrm{au}$ and an armlength $L=5.10^{6}\mathrm{km}$, $d_{O} \sim R/c \sim 500\mathrm{s}$ and $d_{c} \sim L/c \sim 17\mathrm{s}$.
+It will also be useful to separate the delays in two types of terms: the first, $d_{O} = -\hatk\cdot p_{O}$, relates the waveform at the SSB to the waveform at the center of the constellation $O$, whereas the second, $d_{c}$, represents the various delays between the spacecrafts of the constellation.
+For an orbit radius $R=1\,\mathrm{au}$ and an armlength $L=2.5\times10^{6}\mathrm{km}$~\cite{LISA17}, $d_{O} \sim R/c \sim 500\mathrm{s}$ and $d_{c} \sim L/c \sim 8\mathrm{s}$.
 
@@
 In order to keep close to the notation used in the gravitational wave literature, we introduced a factor of 2 in the phase of the wave, which is appropriate for the dominant 22 harmonic of the signal.
-The approximation then applies to signals verifying the conditions \SM{[say that the third is sometimes missing ?]}
+The approximation then applies to signals verifying the conditions \SM{[say that the third is sometimes missing ?]} \SM{[add reference to Finn-Chernoff, Cutler-Flanagan where the third one is missing]}
 \be\label{eq:conditionSPA}
@@
 To understand the separation of timescales in the problem, it will be useful to have at hand the leading-order scaling laws for an inspiral (at the Newtonian order in the PN language).
-Although inaccurate for the purpose of waveform modelling, these Newtonian estimates will give a good qualitative description of the order of magnitude of the relevant timescales in the inspiral.
+Although inaccurate for the purpose of waveform modelling, these Newtonian estimates will give useful orders of magnitude of the relevant timescales in the inspiral.
 For a binary with masses $m_{1}, m_{2}$, we define the total mass $m=m_{1}+m_{2}$ and the symmetric mass ratio $\nu = m_{1}m_{2}/m^{2}$.
 Introducing a time of coalescence $t_{c}$, the relations between the orbital frequency and phase and the time to coalescence $t_{c} - t$ are then given by
-\begin{subequations}
+\begin{subequations}\label{eq:omegaphiN}
 \begin{align}
@@
 For this Newtonian inspiral, applying the SPA gives for the $22$ mode $\tilde{h}(f) = A_{\rm N}(f)e^{-i\Psi_{\rm N}(f)}$ with
-\begin{subequations}
-\begin{align}
-	A_{N}(f) &= \frac{G^{2}m^{2} \pi}{Dc^{5}} \sqrt{\frac{2\nu}{3}} v^{-7/2}\,,\\
-	\Psi_{\rm SPA}^{\rm N}(f) &= \phi_{0} - 2\pi f t_{0} - \frac{3}{128\nu v^{5}} \,, 
+\begin{subequations}\label{eq:SPAN}
+\begin{align}
+	A_{N}(f) &= \frac{G^{2}m^{2} \pi}{Dc^{5}} \sqrt{\frac{2\nu}{3}} v^{-7/2}\,, \label{eq:ASPAN}\\
+	\Psi_{\rm SPA}^{\rm N}(f) &= \phi_{0} - 2\pi f t_{0} - \frac{3}{128\nu v^{5}} \,, \label{eq:PsiSPAN} 
 \end{align}
@@
 It is also customary to rewrite the above relations in terms of the chirp mass $\calM \equiv m\nu^{3/5}$, which is the only mass combination arising in the phase at the leading PN order.
+
+\SM{[explain how (11) and (12) allow one to directly check conditions (7)]}
 
@@
 
-\jgb{A direct computation of \eqref{eq:FDkernel} using \eqref{eq:defGtilde} will generally be computationally demanding, but when $\tilde{G}(f,f')$ has only limited support over $f'$, we can seek an accurate but more efficient approximation.} In our approach, two assumptions will be crucial for our approximations to be valid:
+\jgb{A direct computation of \eqref{eq:FDkernel} using \eqref{eq:defGtilde} will generally be computationally demanding, but when $\tilde{G}(f,f')$ has only limited support over $f'$, we can seek an accurate but more efficient approximation.} In our approach, two assumptions will be crucial for our approximations to be valid: \SM{[rewrite: not clear, make clear the 'SPA for the integral' argument]}
 \begin{itemize}
@@
 \end{itemize}
-Although well verified for the case of LISA-type detectors, the first assumption can be a limitation of the approximation in the case of preccesing binaries, and we will carefully explore its validity in Sec.~\ref{}.
+Although well verified for the case of LISA-type detectors, the first assumption can be a limitation of the approximation in the case of precessing binaries, and we will carefully explore its validity in Sec.~\ref{}.
 
@@
 
-\subsection{Leading order: the local approximation \SM{[Find a better name for this approximation ?]}}
+\subsection{Leading order: the local-in-frequency approximation}
 \label{subsec:LLP}
-
-Under the assumptions above, it is possible to see the convolution in the $f'$ variable in~\eqref{eq:FDkernel} as being a local operation, allowing us to expand the signal $\tilde{h}(f-f')$ around $f$.
-A first leading-order approximation, which could be called the local approximation for reasons that will appear clear below, is to discard the $f'$ dependence in the first argument of $\tilde{g}(f-f', f')$, to keep only the leading order in the Fourier-domain amplitude, and to expand the Fourier-domain phase to the first order, according to
-\be
-	\tilde{h}(f-f') \simeq A(f) \exp\left[ -i\left( \Psi(f) - f' \frac{\ud \Psi}{\ud f} \right) \right] \,,
-\ee
-Here and in the rest of the paper, derivatives with respect to the frequency will always be evaluated at $f$.
-Notice that, because of the freedom of adding a linear term to $\Psi(f)$ by simply shifting the signal in time, no assumption can be made on the smallness of the first derivative of the phase, and this is really a leading order approximation.
-
-Plugging this relation in~\eqref{eq:FDkernel}, we obtain
-\begin{align}
-	\tilde{s}(f) &= \tilde{h}(f) \int \ud f' \, \exp\left[ i f' \frac{\ud \Psi}{\ud f} \right] \tilde{G}(f,f') \nn\\
-	&= \tilde{h}(f) G\left( f, -\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) \,,
-\end{align}
-which motivates the introduction of an effective time, seen as a function of frequency,
-\be\label{eq:deftf}
-	\tf \equiv -\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \,,
-\ee
-This definition is a generalization of the time-to-frequency correspondence at the heart of the SPA, as will be discussed in Sec.~\ref{subsec:linkSPA}, but it is worth noting that here it is defined entirely from the Fourier-domain waveform, and independently of the SPA being valid or not for the underlying signal $\tilde{h}$.
-As such, this time-of-frequency function also extends to the merger-ringdown part of the signals.
-Fig.~\ref{fig:tf} shows the behaviour of the time function $\tf$ around merger for six example waveforms, for mass ratios $q=1$ and $q=8$ and for aligned spin components $\chi=0.95,0.,-0.95$.
-In particular, one should note that $\tf$ is not monotonously increasing with frequency anymore when reaching in the high-frequency part of the waveform, corresponding to the ringdown.
-This would forbid an unambiguous definition of an inverse fequency-of-time correspondence $f(t)$.
 
@@
 The vertical lines shows the time-domain instantaneous frequency at the peak $\omega^{22}_{\rm peak}/(2\pi)$, again measured from an IFFT, and the fact that the curves $t_{f}$ do not pass exactly by the crossing with the $t=0$ horizontal line reflects the fact that the link between time-domain frequency and Fourier-domain $f$ is not accurate anymore in the merger part of the signal.
-Note that $t_{f}$ does not go beyond $~20M$ after the peak, and is not monotonous.
+Note that $t_{f}$ does not go beyond $\sim 30M$ after the peak, and is not monotonous.
 \SM{[add color to the vertical lines]} \SM{[add dots to highlights the crossings]}}
@@
 \end{figure}
+
+Under the assumptions above, it is possible to see the convolution in the $f'$ variable in~\eqref{eq:FDkernel} as being a local operation, allowing us to expand the signal $\tilde{h}(f-f')$ around $f$.
+A first leading-order approximation, which could be called the local approximation for reasons that will appear clear below, is to discard the $f'$ dependence in the first argument of $\tilde{g}(f-f', f')$, to keep only the leading order in the Fourier-domain amplitude, and to expand the Fourier-domain phase to the first order, according to
+\be
+	\tilde{h}(f-f') \simeq A(f) \exp\left[ -i\left( \Psi(f) - f' \frac{\ud \Psi}{\ud f} \right) \right] \,,
+\ee
+Here and in the rest of the paper, derivatives with respect to the frequency will always be evaluated at $f$.
+Notice that, because of the freedom of adding a linear term to $\Psi(f)$ by simply shifting the signal in time, no assumption can be made on the smallness of the first derivative of the phase, and this is really a leading order approximation.
+
+Plugging this relation in~\eqref{eq:FDkernel}, we obtain
+\begin{align}
+	\tilde{s}(f) &= \tilde{h}(f) \int \ud f' \, \exp\left[ i f' \frac{\ud \Psi}{\ud f} \right] \tilde{G}(f,f') \nn\\
+	&= \tilde{h}(f) G\left( f, -\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) \,,
+\end{align}
+which motivates the introduction of an effective time, seen as a function of frequency,
+\be\label{eq:deftf}
+	\tf \equiv -\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \,,
+\ee
+This definition is a generalization of the time-to-frequency correspondence at the heart of the SPA, as will be discussed in Sec.~\ref{subsec:linkSPA}, but it is worth noting that here it is defined entirely from the Fourier-domain waveform, and independently of the SPA being valid or not for the underlying signal $\tilde{h}$.
+As such, this time-of-frequency function also extends to the merger-ringdown part of the signals.
+Fig.~\ref{fig:tf} shows the behaviour of the time function $\tf$ around merger for six example waveforms, for mass ratios $q=1$ and $q=8$ and for aligned spin components $\chi=0.95,0.,-0.95$.
+In particular, one should note that $\tf$ is not monotonously increasing with frequency anymore when reaching in the high-frequency part of the waveform, corresponding to the ringdown.
+This would forbid an unambiguous definition of an inverse frequency-of-time correspondence $f(t)$.
 
@@
 
-\subsection{General Taylor expansion in the Fourier domain}
+\subsection{Taylor expansion in the Fourier domain}
 \label{subsec:TaylorFD}
@@
 \begin{widetext}
-If we further expand the exponentials to obtain a pure $f'$-expansion, by applying the derivative rule
+If we further expand the exponentials to obtain a pure $f'$-expansion, by applying the formal derivative rule
 \allowdisplaybreaks
@@
 
-This rather cumbersome expansion will not be used as such, but takes a more intuitive form for special cases.
-For instance, if we expand only the frequency-dependence of $G$ as in~\eqref{eq:expandg}, we obtain
-\be\label{eq:taylordelay}
-	\calT_{\rm delay}(f) = \sum\limits_{r\geq 0} \frac{1}{(2i\pi)^{r}r!} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{r} }{\partial t^{r}} G \right)(f,\tf)
-\ee
-which interestingly looks like a Taylor expansion of $G$, but with joint derivatives in frequency and time. Similarly, the expansion~\eqref{eq:expandA} gives
-\be\label{eq:resultdfA}
-	\calT_{\rm amp}(f) = \sum\limits_{q\geq 0} \frac{1}{(2i\pi)^{q}q!} \frac{1}{A} \frac{\ud^{q} A}{\ud f ^{q}}  \left( \frac{\partial^{q} }{\partial t^{q}} G \right)(f,\tf)
-\ee
-
-Of particular interest for us will be the case where we keep only the first term in~\eqref{eq:expandPsi}, and expand the resulting exponential, which yields
-\be\label{eq:resultdffPsi}
+This cumbersome expansion will not be used as such, but takes a more intuitive form for special cases.
+Of particular interest for us will be the case where we keep only the first term of the series in~\eqref{eq:expandPsi}, and expand the resulting exponential, which yields
+\be\label{eq:resulttaylorPsi}
 	\calT_{\rm phase}(f) = \sum\limits_{p\geq 0} \frac{1}{p!} \left( \frac{i}{8\pi^{2}}\frac{\ud^{2} \Psi}{\ud f^{2}} \right)^{p} \left( \frac{\partial^{2p} }{\partial t^{2p}} G \right)(f, \tf) \,.
@@
 Indeed, this correction will be shown in Sec.~\ref{subsec:resumquadphase} to be closely related to the previous approach of~\cite{KCY14}.
-
-In the following, it will be convenient to introduce \SM{[think of removing ``figure of merit'' throughout the text as it is not quite the word]} error estimates for these Taylor-like expansions, that we will take to be the magnitude of the first ignored term in the series.
-Thus we define
-\begin{subequations}\label{eq:deffom}
-\begin{align}
-	\epsilon_{\Psi 2} &\equiv \frac{1}{4\pi^{2}} \left| \frac{\ud^{2} \Psi}{\ud f^{2}} \frac{1}{G} \frac{\partial^{2} G}{\partial t^{2}}(f, \tf) \right| \,, \\
-	\epsilon_{A 1} &\equiv \frac{1}{2\pi} \left| \frac{1}{A}\frac{\ud A}{\ud f} \frac{1}{G} \frac{\partial G}{\partial t}(f, \tf) \right| \,, \\
-	\epsilon_{A 2} &\equiv \frac{1}{4\pi^{2}} \left| \frac{1}{A}\frac{\ud^{2} A}{\ud f^{2}} \frac{1}{G} \frac{\partial^{2} G}{\partial t^{2}}(f, \tf) \right| \,, \\
-	\epsilon_{d} &\equiv \frac{1}{2\pi} \left| \frac{1}{G} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{r} }{\partial t^{r}} G \right)(f, \tf) \right| \,.
-\end{align}
-\end{subequations}
+The expansion shows that the transfer function is signal-dependent through the second derivative of the phase $\Psi$.
+Similarly, the expansion~\eqref{eq:expandA} gives
+\be\label{eq:resulttaylorA}
+	\calT_{\rm amp}(f) = \sum\limits_{q\geq 0} \frac{1}{(2i\pi)^{q}q!} \frac{1}{A} \frac{\ud^{q} A}{\ud f ^{q}}  \left( \frac{\partial^{q} }{\partial t^{q}} G \right)(f,\tf) \,.
+\ee
+If we expand only the frequency-dependence of $G$ as in~\eqref{eq:expandg}, we obtain
+\be\label{eq:resulttaylordelay}
+	\calT_{\rm delay}(f) = \sum\limits_{r\geq 0} \frac{1}{(2i\pi)^{r}r!} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{r} }{\partial t^{r}} G \right)(f,\tf) \,,
+\ee
+which interestingly looks like a Taylor expansion of $G$, but with joint derivatives in frequency and time.
+This time, the expansion is signal-independent as it only depends on the function $G$ and not on $A$, $\Psi$.
+
+In the following, we will rewrite in a more convenient form the formal Taylor expansions for the quadratic phase effects and the delay effects.
+As will be explained in Sec.~\ref{} \SM{[R]}, we will use these Taylor expansions to build error measures designed to keep track of which of these three types of corrections is important.
 
 \SM{[interpretation of $\epsilon$'s as ratios of timescales - next subsec]}
+
+\SM{[add magnitude of the derivatives of G, like $\Omega^{n}$ with possible factors of $2\pi f d$]}
+
+\SM{[add loose convergence argument with scaling $\partial^{n}_{t} \sim \Omega^{n} $]}
 
@@
 
-\subsection{Signal-dependent timescales and relation to the stationary phase approximation and SUA}
-\label{subsec:linkSPA}
-
-\begin{figure*}
+\subsection{Signal-dependent timescales}
+\label{subsec:timescales}
+
+\begin{figure}
   \centering
-  \includegraphics[width=.48\linewidth]{plots/TfTA.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/TfTAzoom.pdf}
+  \includegraphics[width=.99\linewidth]{plots/TfTA.pdf}
   \caption{Fourier-domain amplitude and phase timescales, as defined in~\eqref{eq:defTf} and~\eqref{eq:defTA}, in geometric units for an equal-mass, non-spinning system.
-The right panel shows the merger and ringdown part of the signal in more details, the time-domain frequency at merger being represented by the vertical line.
-The derivatives from which these timescales are built encounter zero-crossings in the high-frequency range.
-[if using PhenomD, discuss discontinuity of third derivative]}
+The hierarchy of these timescales is roughly the same for higher mass ratio and higher spin systems.
+The time-domain frequency at merger is represented by the vertical line.
+The derivatives from which these timescales are built encounter zero-crossings and change sign in the high-frequency range.
+\SM{[discuss discontinuity of third derivative]} \SM{[change legent to Tf/M]}}
   \label{fig:TfTA}
-\end{figure*}
-
-Considering~\eqref{eq:eq:resultdffPsi} it is natural to define a new timescale, as a function of frequency, as
+\end{figure}
+
+Considering~\eqref{eq:resulttaylorPsi} it is natural to define a new timescale, as a function of frequency, as
 \be\label{eq:defTf}
@@
 \ee
-The interpretation of this timescale for the inspiralling part of the signal, when the SPA applies, will be discussed below. Whe can use this notation to rewrite~\eqref{eq:resultdffPsi} as
+The interpretation of this timescale for the inspiralling part of the signal, when the SPA applies, will be discussed below. Whe can use this notation to rewrite~\eqref{eq:resulttaylorPsi} as
 \begin{align}\label{eq:resultdffPsiTf}
@@
 
-The expansion in amplitude corrections~\eqref{eq:resultdfA} also leads to the natural introduction of a set of other timescales related to the successive derivatives of the amplitude.
+The expansion in amplitude corrections~\eqref{eq:resulttaylorA} also leads to the natural introduction of a set of other timescales related to the successive derivatives of the amplitude.
 In an analogous manner to the definition~\eqref{eq:defTf} of the timescale $T_{f}$, we can define
@@
 where we included an absolute value to accomodate the possible sign changes in the right-hand side.
-In practice, only the first few of these timescales will be relevant ($T_{A1}$ and $T_{A2}$ in the analysis of this paper).
+Although the definition above is for a generic $n\geq 0$, in practice only the first few of these timescales will be relevant.
+In this paper we will use only the first two, $T_{A1}$ and $T_{A2}$.
 
@@
 Similarly, if we define
-\be
+\be\label{eq:TfSPA}
 	T_{f}^{\rm SPA} \equiv \frac{1}{\sqrt{2\dot{\omega}(t_{f})}} \,,
@@
 
-We can now compare our result to the approach of Ref.~\cite{KCY14}, where the authors extend the SPA in a formalism called the shifted uniform asymptotic expansion (SUA).
+It is useful, to estimate the magnitude of the corrections with respect to the leading order as we will do in Secs.~\ref{} and~\ref{} below \SM{[R]}, to give an estimate for these timescales based on the Newtonian expressions~\eqref{eq:SPAN}.
+This gives for the Newtonian radiation-reaction and amplitude timescales:
+\begin{subequations}\label{eq:timescalesN}
+\begin{align}
+	T_{f}^{\rm N} &= \frac{1}{8} \sqrt{\frac{5}{3\nu}} \frac{G m}{c^{3}} v^{-11/2} \,, \label{eq:TfN}\\
+	T_{A1}^{\rm N} &= \frac{7}{6} \frac{1}{2\pi f} \,. \label{eq:TA1N}
+\end{align}
+\end{subequations}
+One can check explicitly that this expression for $T_{f}$ agrees with~\eqref{eq:TfSPA}.
+With a simple power-law amplitude as in~\eqref{eq:ASPAN}, all higher-order amplitude timescales are also simply proportional to $1/f$ and differ from the first only by their numerical factor.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\section{Quadratic phase corrections and delays}
+%\label{sec:quadphasedelay}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Quadratic term in the phase and relation to the SUA}
+\label{subsec:resumquadphase}
+
+\SM{[move all of the SUA comparison here]}
+
+In this Section, we focus on the second-derivative corrective term in~\eqref{}.
+
+If we restrict to the case of a pure modulation, $d=0$ and $G(f,t) = F(t)$ [go back to more general $G(f,t)$ ?], we can relate our result~\eqref{eq:resultdffPsiTf} to the approach of Ref.~\cite{KCY14}, where the authors extend the SPA in a formalism called the shifted uniform asymptotic expansion (SUA).
 The derivation includes an expansion of the modulation in Bessel functions, a study of the displacement of the stationary phase point induced by the precession, and a resummation of the result in the form of time derivatives.
@@
 
-\SM{[add: $\sim$ The relevant scaling for the terms in the series~\eqref{eq:resultdffPsiTf}, for the inspiral phase where the SPA applies, is therefore
-\be
-	\Tf^{2p} \partial_{t}^{2p} F = \calO(-1) \,,
-\ee]}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\section{Quadratic phase corrections and delays}
-%\label{sec:quadphasedelay}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Quadratic term in the phase}
-\label{subsec:resumquadphase}
-
-In this Section, we focus on the second-derivative corrective term in~\eqref{}.
-We will restrict to the case of a pure modulation, $d=0$ and $G(f,t) = F(t)$ [go back to more general $G(f,t)$ ?].
 The authors of~\cite{KCY14} proposed a resummation scheme.
@@
 
-Note that the choice of the samples $\tf \pm k\Tf$ to approximate the successive derivatives of the expansion by finite differences is by no means unique.
-However, we remark that its simple structure leads to an anlytical expression for the stencil coefficients $a_{N,k}$, which will be given in~\ref{app:vandermonde}.
+Note that the choice of the samples $\tf \pm k\Tf$ to approximate the successive derivatives of the expansion by finite differences, even if simple and natural, is by no means unique.
+Its structure leads to simple rational expressions for the stencil coefficients $a_{N,k}$, which will be given in~\ref{app:vandermonde} \SM{[refer to table]}.
 In principle, one could have made any ansatz of samples for the series expansion to be matched on.
@@
 
-In the case of a LISA-type detector response, the presence of the delay $d(t)$ is responsible for the frequency-dependence of the kernel $g$ introduced in~\eqref{eq:defg}.
-We will show below in Sec.~\ref{} that the higher-order correction in the delay can become non-negligible at high frequencies, as measured by the magnitude of the first correction in~\eqref{eq:taylordelay}.
+In the case of a LISA-type detector response, the presence of the delay $d(t)$ is responsible for the frequency-dependence of the kernel $G$ introduced in~\eqref{eq:defg}.
+We will show below in Sec.~\ref{} \SM{[R]}that the higher-order correction in the delay can become non-negligible at high frequencies, as measured by the magnitude of the first correction in~\eqref{eq:taylordelay}.
 This is especially true for the orbital delay related to the motion of the whole detector constellation around the Sun, due to the large baseline for this delay.
@@
 
-For the standard LISA configuration, we have for the orbital delays the scaling $d_{O}\sim R/c \sim 500s$, for the constellation delays the scaling $d_{c}\sim L/c \sim 17s$ (with an additional dependence on angular factors).
-Since the motion of the constellation is periodic with a period of one year and $\Omega_{0} = 2.10^{-7}\mathrm{rad}.s^{-1}$, we have $\dot{d}_{O} \sim \Omega_{0} R/c \sim 10^{-4}$ and $\dot{d}_{c} \sim \Omega_{0} L/c \sim 3.10^{-6}$.
+For the standard LISA configuration, we have for the orbital delays the scaling $d_{O}\sim R \sim 500s$, for the constellation delays the scaling $d_{c}\sim L \sim 8s$ (with $c=1$ and ignoring the dependence on angular factors).
+Since the motion of the constellation is periodic with a period of one year and $\Omega_{0} = 2 \times 10^{-7}\mathrm{rad}.s^{-1}$, we have $\dot{d}_{O} \sim \Omega_{0} R/c \sim 10^{-4}$ and $\dot{d}_{c} \sim \Omega_{0} L/c \sim 1.7 \times 10^{-6}$.
 The smallness of the dimensionless quantity $\dot{d} \ll 1$ (and of its subsequent derivatives) will allow us to treat it perturbatively with a very good approximation, and shows also that the function $t_{d}$ is univalued and that there is no ambiguity in defining its reciprocal $t_{d}^{-1}$.
@@
 \end{widetext}
-where we used a change of variable $\tau = t_{d}(t)$. We see that the result can again be expressed as a Fresnel transform.
-
+where we used a change of variable $\tau = t_{d}(t)$.
+We see that the result can again be expressed as a Fresnel transform.
 When considering amplitude corrections as well, as in~\eqref{}, additional powers of $f'$ can be translated as time derivatives with respect to the variable $\tau$ after performing the change of variables.
-Thus, we obtain our most general result for an exact treatment of the quadratic term in the phase and for an improved treatment of the delays as:
+This produces our most general result given in Eq.~\eqref{eq:transferfinal} below.
+
+\subsection{Executive summary of the formalism}
+
+In this Section, we gather our previous results for the convenience of the reader.
+For a modulation $F$ and delay $d$, so that $s(t) = F(t) h(t+d(t))$, we defined the associated function $G(f,t) = F(t) e^{-2i\pi f d(t)}$ and we obtained the transfer function $\calT (f) = \tilde{s}(f)/\tilde{h}(f)$:
 \begin{widetext}
 \be\label{eq:transferfinal}
-	\calT(f) = \sum\limits_{k=0}^{+\infty} \frac{1}{(2i\pi)^{k} k!} \frac{1}{A} \frac{\ud^{k}A}{\ud f^{k}}  \calF^{\epsilon}_{\Tf} \left[ \frac{\ud^{k}}{\ud \tau^{k}} \left( \frac{F(\tau - d(\tau))}{1+\dot{d}(\tau)} e^{-2i\pi f d(\tau)(1-\dot{d}(\tau))} \right) \right] (\tf) \,.
+	\calT(f) = \sum\limits_{k \geq 0} \frac{(-i)^{k}}{k!} (T_{Ak})^{k} \calF^{\epsilon}_{\Tf} \left[ \frac{\ud^{k}}{\ud \tau^{k}} \left( \frac{F(\tau - d(\tau))}{1+\dot{d}(\tau)} e^{-2i\pi f d(\tau)(1-\dot{d}(\tau))} \right) \right] (\tf) \,.
 \ee
 \end{widetext}
-Each term in this series (in practice, only the first two or three will be relevant) can in turn be approximated by using the stencil formula~\eqref{eq:Fresnelstencil}.
+In each term of this series, the Fresnel transform $\calF^{\epsilon}_{\Tf}$ can in turn be approximated by using the stencil formula~\eqref{eq:Fresnelstencil}.
+The timescales $T_{f}$ and $T_{Ak}$ were defined in~\eqref{eq:defTf} and~\eqref{eq:defTA}, and the generalized time-to-frequency function $t_{f}$ was given in~\eqref{eq:deftf}.
+The delays $d$ are present only in the LISA context, while in the context of precessing binaries we only have to consider the modulation function $F$.
+In practice, only the first few of the terms in this last remaining series expansion are relevant.
+We will investigate several orders of approximation, combining corrections from the phase, amplitude and delays, and explore which ones are relevant for a given level of accuracy.
+We will use symbols of the form $\{N,A,d\}$ to indicate the order of the stencil in~\eqref{eq:Fresnelstencil}, the maximal order of the amplitude correction included, and the inclusion or not of the delay correction.
+
+In the LISA context, due to the smallness of the corrections we will only go up to $k=1$ in~\eqref{eq:transferfinal}, and in computing the remaining time derivatives in~\eqref{eq:transferfinal} we will neglect the second and higher derivatives.
+We will also use $F(\tau - d(\tau)) \simeq (F - d \dot{F})(\tau)$.
+This gives concretely:
+\begin{widetext}
+\begin{subequations}
+\begin{align}
+	\{N,A:0,d:0\}&: \; \calT(f) = \calF^{\epsilon}_{\Tf} \left[ F e^{-2i\pi f d} \right] (t_{f}) \,, \\
+	\{N,A:1,d:0\}&: \; \calT(f) = \calF^{\epsilon}_{\Tf} \left[ \left(F - i T_{A1} \left( \dot{F} - 2i\pi f \dot{d}\right) \right) e^{-2i\pi f d} \right] (t_{f}) \,, \\
+	\{N,A:0,d:1\}&: \; \calT(f) = \calF^{\epsilon}_{\Tf} \left[ \frac{F - d\dot{F}}{1+\dot{d}} e^{-2i\pi f d (1-\dot{d})} \right] (t_{f})\,, \\
+	\{N,A:1,d:1\}&: \; \calT(f) = \calF^{\epsilon}_{\Tf} \left[ \frac{1}{1+\dot{d}}\left(  F - d\dot{F} - i T_{A1} \left( \dot{F} - 2i\pi f \dot{d} \right) \right) e^{-2i\pi f d (1 - \dot{d})} \right] (t_{f})\,.
+\end{align}
+\end{subequations}
+\end{widetext}
+
+In the context of precessing binaries, the delays are absent and we will go up to $k=2$ in~\eqref{eq:transferfinal}. The orders of approximation will be:
+\begin{subequations}
+\begin{align}
+	\{N,A:0\}&: \; \calT(f) = \calF^{\epsilon}_{\Tf} \left[ F \right] (t_{f}) \,, \\
+	\{N,A:1\}&: \; \calT(f) = \calF^{\epsilon}_{\Tf} \left[ \left(F - i T_{A1} \dot{F} \right) \right] (t_{f}) \,, \\
+	\{N,A:2\}&: \; \calT(f) = \calF^{\epsilon}_{\Tf} \left[ \left(F - i T_{A1} \dot{F} - \frac{1}{2} (T_{A2})^{2} \ddot{F} \right) \right] (t_{f}) \,.
+\end{align}
+\end{subequations}
+
+In the following, it will be convenient to introduce \SM{[think of removing ``figure of merit'' throughout the text as it is not quite the word - error measure ?]} error estimates built from the Taylor-like series~\eqref{eq:resulttaylorPsi},~\eqref{eq:resulttaylorA} and~\eqref{eq:resulttaylordelay}.
+We simply define  this error estimates at a certain level of approximation as the magnitude of the first term ignored in the original Taylor series.
+Thus we define \SM{[check if the 1/2!
+was taken into account]}
+\begin{subequations}\label{eq:deffom}
+\begin{align}
+	\epsilon_{\Psi 2} &\equiv \frac{1}{2} T_{f}^{2} \left| \frac{1}{G}\partial_{tt}G \right| \,, \\
+	\epsilon_{A 1} &\equiv T_{A1} \left| \frac{1}{G} \partial_{t} G \right| \,, \\
+	\epsilon_{A 2} &\equiv \frac{1}{2} T_{A2}^{2} \left| \frac{1}{G} \partial_{tt}G \right| \,, \\
+	\epsilon_{d} &\equiv \frac{1}{2\pi} \left| \frac{1}{G} \partial_{tf} G \right| \,,
+\end{align}
+\end{subequations}
+where the function $G$ and its derivatives are evaluated at $(f, t_{f})$.
+As we will detail below, in the presence of delays these derivatives do not obey the simple scaling $\partial_{t}^{n} \rightarrow \Omega_{0}^{n}$, with $\Omega_{0}$ the characteristic frequency for the time variablility in the modulation and delay.
+This is due to the presence of additional dimensionless delay terms of the form $2\pi f d$, which can be larger than 1.
 
@@
 
-\subsection{The response model and TDI observables}
+\subsection{The response model}
 \label{subsec:modelLISA}
 
-In this Section, we expand Sec.~\ref{subsec:} and detail more the model that we use for the response of a LISA-like detector, together with the assumptions used and their limitations.
+In this Section, we expand Sec.~\ref{subsec:} \SM{[R]} and detail more the model that we use for the response of a LISA-like detector, together with the assumptions used and their limitations.
 
@@
 
-Since the orbital delay due to the orbit around the Sun is common to all $y_{slr}$ observables, we will decompose the full instrument response in two steps.
+Since the orbital delay due to the orbit around the Sun is common to all $y_{slr}$ observables, we will decompose the full instrument response~\eqref{} \SM{[R]} in two steps.
 The first step consists simply in applying the varying time delay to bring the signal from the SSB to the center of the LISA triangular constellation.
-For $h=h^{\rm TT}$ the transverse-traceless gravitational waveform in matrix form, we write this orbital time delay as
+For $h^{\rm TT}$ the transverse-traceless gravitational waveform in matrix form, we write this orbital time delay as
 \be\label{eq:defresponseO}
-	h_{O} (t) = h(t-\hat{k}\cdot p_{O}) \,.
+	h_{O} (t) = h(t-\hatk\cdot p_{O}) \,.
 \ee
 The second step consists in the remaining, constellation-centered response, which we write as
-\begin{align}\label{eq:dfresponsec}
-	y_{slr} &= \frac{1}{2} \frac{1}{1 - \hat{k}\cdot n_{l}} \nn\\
-	& \qquad \cdot n_{l}\cdot \left[ h_{O}(t - L - \hat{k}\cdot p^{L}_{s}) - h_{O}(t - \hat{k}\cdot p^{L}_{r}) \right] \cdot n_{l}\,,
+\begin{align}\label{eq:defresponsec}
+	y_{slr} &= \frac{1}{2} \frac{1}{1 - \hatk\cdot n_{l}} \nn\\
+	& \cdot n_{l}\cdot \left[ h_{O}^{\rm TT}(t - L - \hatk\cdot p^{L}_{s}) - h_{O}^{\rm TT}(t - \hatk\cdot p^{L}_{r}) \right] \cdot n_{l}\,,
 \end{align}
@@
 \be
-	h = h^{\rm TT} = h_{+}P_{+} + h_{\times}P_{\times} \,.
+	h^{\rm TT} = h_{+}P_{+} + h_{\times}P_{\times} \,.
 \ee
@@
 \be
-	P = 
+	P_{\ell m} = 
 	\begin{cases} 
@@
 
-We now turn to the transposition of~\eqref{eq:yslr} to the Fourier-domain.
-When using the rigid approximation for the constellation, a particular simplification occurs when writing the transfer function at leading order in the delays.
-The Fourier-domain transfer function for~\eqref{eq:dfresponsec} then reads in the local approximation~\eqref{eq:transferlocal}
-\begin{align}
-	\calT_{slr} &= \frac{i \pi f L}{2} \sinc \left[ \pi f L\left(1-\hat{k}\cdot n_{l} \right) \right] \nn\\
-	& \qquad \cdot \exp\left[ i \pi f \left( L + \hat{k}\cdot \left( p_{1}^{L} + p_{2}^{L} \right) \right) \right]  n_{l} \cdot P_{\ell m} \cdot n_{l} \,,
-\end{align}
-where we used the fact that, in our rigid approximation for the trajectory of the constellation, $p^{L}_{r} - p^{L}_{s} =  L n_{l}$.
+We now turn to the transposition of~\eqref{eq:defresponseO} and~\eqref{eq:dfresponsec} to the Fourier-domain.
+For the leading-order response~\eqref{eq:deflocal}, in what we call the local approximation, applying a pure delay as in~\eqref{eq:defresponseO} translates into Fourier-domain transfer function that is a pure phase factor, proportional to the frequency but also $t_{f}$-dependent: \SM{[check sign]}
+\be\label{eq:transferOlocal}
+	\calT_{O}^{\rm local}(f) = e^{-2i\pi f d_{O}(t_{f})}\,,
+\ee
+with $d_{O} = -\hatk \cdot p_{O}$ the orbital delay.
+When using the leading-order response~\eqref{eq:transferlocal} for the constellation response, the delays take similarly the form of pure phase factors.
+When also using the rigid approximation for the constellation, where the armlengths are fixed, a particular simplification occurs when combining these individual delays.
+The Fourier-domain transfer function then reads
+\begin{align}\label{eq:transferclocal}
+	\calT_{slr}^{\rm local} &= \frac{i \pi f L}{2} \sinc \left[ \pi f L\left(1-\hatk\cdot n_{l} \right) \right] \nn\\
+	& \qquad \cdot \exp\left[ i \pi f \left( L + \hatk\cdot \left( p_{1}^{L} + p_{2}^{L} \right) \right) \right]  n_{l} \cdot P_{\ell m} \cdot n_{l} \,,
+\end{align}
+where all vectors $n_{l}$ and $p_{A}^{L}$ are to be evaluated at time $t_{f}$.
+We used the rigid-arms approximation relation $p^{L}_{r} - p^{L}_{s} =  L n_{l}$.
+The result~\eqref{eq:transferclocal} is familiar, as it was derived in the past taking the point of view of monochromatic signals (see e.g.~\cite{}).
+
 In the case $m<0$, the formalism can be applied in the same way, except that notation is complicated by the fact that the mode $\tilde{h}_{\ell m}$ has to be evaluated at $-f$.
 In the following, we will limit ourselves to the case $m>0$ for simplicity of presentation.
+\SM{[really necessary ?]}
 
 Note that, if the corrections~\eqref{} are included for the constellation delays, this transfer function will not have this simple form anymore as $\dot{d}$ will have a different velocity-dependent expression for the sending and receiving spacecrafts.
+
+It is natural to define two transfer frequencies for the two relevant length scales for the delays, defined such that a wavelength fits within this length scale.
+Geometrical projection factors aside, these scales are the radius of the orbit around the Sun $R=1\text{au}=1.5\times 10^{8} \text{km}$ and the detector armlength $L=2.5\times 10^{6}\text{km}$ (in the configuration proposed in~\cite{} \SM{[C]}).
+This gives
+\begin{subequations}\label{eq:transferfrequencies}
+\begin{align}
+	f_{R} &= 3.2\times10^{-4}\text{Hz} \,\\
+	f_{L} &= 1.9\times 10^{-2}\text{Hz} \,.
+\end{align}
+\end{subequations}
+These transfer frequencies will indicate where in the LISA band a delay term of the form $2\pi f d$ becomes larger than unity.\SM{[check if this appears elsewhere]}
+
+\SM{[comment on the added structure in the high-f response - say a word of f-factor corresponding to discrete derivatives - define low-f response that others have been using - reference to folks also looking at that (marginal) for ET?
+]}
 
@@
 
-Due to the smallness of $\Omega/2\pi \simeq 3.1\times10^{-8}\mathrm{Hz} \ll f$ over all the frequency band we consider, one can expect that . In this Section we try to quantify to which extent.
-
-The magnitude of the first terms in a Taylor expansion in $f_{0}$ is (taking $10^{-5}\mathrm{Hz}$ as a reference starting frequency):
-%
-\begin{widetext}
+\SM{[change notations from O to 0 for orbital response, and check we use notations $\Omega_{0}$, $f_{0}$]}
+
+Due to the smallness of the annual frequency $\Omega_{0}/2\pi \simeq 3.1\times10^{-8}\mathrm{Hz}$ compared to the timescales relevant for the signal over the LISA frequency band, one can expect that the higher-order corrections described in Section~\ref{sec:formalism} are small.
+In this Section we try to quantify to which extent, and give simple estimates of their magnitude.
+
+Using the Newtonian-order expressions~\eqref{eq:timescalesN} for the signal-dependent timescales $T_{f}$ and $T_{A1}$, we can obtain a useful order-of-magnitude estimate for the error measures $\epsilon$ introduced in~\eqref{eq:deffom}.
+Higher-order amplitude terms beyond the first one in~\eqref{eq:resulttaylorA} will turn out to be completely negligible in the LISA case and we will ignore them in the following.
+We will present these estimates separately for the orbital response~\eqref{eq:transferOlocal} and the constellation~\eqref{eq:transferclocal}, as the different baseline of the delays ($R$ or $L$) as well as the presence of a time-varying prefactor $F(t)$ both affect the result.
+As we will see, when estimating the magnitude of the relevant derivatives of $G$ we have to take into account delay factors of the type $2\pi f d$.
+
+We start with the orbital response, which takes the form of a pure delay $G_{0}(f, t) = e^{-2 i \pi f d_{O}(t)}$, and obtain
 \begin{subequations}
 \begin{align}
-	\delta_{A1} &\equiv \left| \frac{1}{A_{N}}\frac{\ud A_{N}}{\ud f} f_0 \right| \simeq 3\times10^{-3} \left(\frac{f}{10^{-5}\Hz}\right)^{-1} \,,\\
-	\delta_{\Psi 2} &\equiv \left| \frac{1}{2} f_0^{2}\frac{\ud^{2} \Psi_{N}}{\ud f^{2}} \right| \simeq 5 \left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-11/3} \,, \\
-	\delta_{\Psi 3} &\equiv \left| \frac{1}{6} f_0^{3}\frac{\ud^{3} \Psi_{N}}{\ud f^{3}} \right| \simeq 2\times 10^{-2} \left( \frac{m}{10^{6}\Msol} \right)^{-5/3} \left( \frac{\nu}{1/4} \right)^{-1} \left( \frac{f}{10^{-5}\Hz} \right)^{-14/3} \,.
+	\frac{1}{G_{0}} \partial_{t} G_{0} &= -2i\pi f \dot{d}_{O}\,,\\
+	\frac{1}{G_{0}} \partial_{tt} G_{0} &= -2i\pi f \ddot{d}_{O} - 4\pi^{2} f^{2} \dot{d}_{O}^{2} \,,\\
+	\frac{1}{G_{0}} \partial_{tf} G_{0} &= -2 i \pi \dot{d}_{O} - 4\pi^{2} f d_{O} \dot{d}_{O} \,,
 \end{align}
 \end{subequations}
-\end{widetext}
-%
-The approximation degrades for lower starting frequencies, and, at a fixed frequency, when the total mass decreases and/or the mass ratio increases.
-However, if we consider only merging binaries, that is to say binaries of which the merger is seen during the observation period of the mission, we have a lower limit for the starting frequency which can be larger than the detector's lowest frequency in band.
-
-From the Newtonian estimate~\eqref{}, the starting frequency is then
-%
-\be
-	f_{\rm start} = 10^{-5}\mathrm{Hz} \left( \frac{m}{8.5\times10^{6}\Msol} \right)^{-5/8} \left( \frac{\nu}{1/4} \right)^{-3/8} \left( \frac{\Delta t}{5 \mathrm{yr}} \right)^{-3/8} \,.
-\ee
-%
-So we see that the ``worst'' total mass is $8.5\times10^6\Msol (\nu/(1/4))^{3/5}$, for 5 years of observation and $f_{\rm low} = 10^{-5}\mathrm{Hz}$ for the entry in band, which are conservative values; for higher masses, part of the signal is cut because it is out-of-band, while for lower masses, the signal is cut because of the limited time of observation before merger, and we have the scaling
-%
-\be
-	|\delta^{2}\Psi |_{f_\mathrm{start}} \propto m^{5/8}\nu^{3/8}\Delta t^{11/8} \,,
-\ee
-%
-which shows that the effect of the truncation of the signal for a finite observation time wins: the approximation becomes better for lower mass and higher mass ratios systems.
-The worse point in the parameter space should be, for these Newtonian estimates, at equal mass and $m=8.5\times 10^{6} \Msol$, with $|\delta^{2}\Psi |_{\mathrm{worst}} \simeq 0.15$.
+
+For the constellation response \SM{[say here that even the one-arm response is actually the difference between two G's, so the error measure is not quite rightly used -- origin of the f factor in front]}, since we wish to write errors in relative sense starting from the single-arm response~\eqref{eq:transferclocal} which is already analogous to a discrete time derivative of the signal, we keep explicit the overall factor $f$ reflecting this derivative.
+\SM{[investigate later how this applies to the discrete time derivatives that define the TDI - here we are allowed to ignore this because they are with a fixed L-delay]} We will write symbolically $G_{c}(f,t) \sim f F(t) e^{-2i \pi f d_{c}(t)}$, where $d_{c}$ represents the delay term and $F(t)$ represents the rest of the geometric factors in~\eqref{eq:transferclocal} where we ignore the $f$-dependence.
+This gives in this case
+\begin{subequations}
+\begin{align}
+	\frac{1}{G_{c}} \partial_{t} G_{c} &\sim -2i\pi f \dot{d}_{c} + \frac{\dot{F}}{F}\,,\\
+	\frac{1}{G_{c}} \partial_{tt} G_{c} &\sim -2i\pi f \ddot{d}_{c} - 4\pi^{2} f^{2} \dot{d}_{c}^{2} - 2i\pi f \dot{d}_{c} \frac{\dot{F}}{F} - \frac{\dot{F}^{2}}{F^{2}} + \frac{\ddot{F}}{F} \,,\\
+	\frac{1}{G_{c}} \partial_{tf} G_{c} &\sim -4 i \pi \dot{d}_{c} - 2i\pi d_{c} \frac{\dot{F}}{F} - 4\pi^{2} f d_{c} \dot{d}_{c} + \frac{1}{f}\frac{\dot{F}}{F}\,.
+\end{align}
+\end{subequations}
+
+In order to obtain simple scalings for the error estimates, we will make the replacements $\partial_{t}^{n} d \sim \Omega_{0}^{n} d$ as well as $\partial_{t}^{n} F \sim \Omega_{0}^{n} F$.
+These scalings are only approximate and the various possible orientation angles can lead to significant variations.
+We will assume $d_{0} \sim R_{\rm eff} $ (recalling that we set $c=1$ \SM{[check if that is consistently done throughout]}) with an effective orbit radius $R_{\rm eff} \equiv 2R/\pi$ representing an average of the geometric projection factor of the gravitational wave propagation vector on the plane of the orbit.
+For the constellation delays, we simply take $d_{c} \sim L$ as there projection effects can lead to variations in both ways.
+We build estimates for the magnitude of these derivatives as follows:
+\begin{subequations}\label{eq:estimatederivorb}
+\begin{align}
+	\left| \frac{1}{G_{0}} \partial_{t} G_{0} \right| &\sim 2 \pi f \Omega_{0} R_{\rm eff}\,,\\
+	\left| \frac{1}{G_{0}} \partial_{tt} G_{0} \right| &\sim \text{max} \left[ 2 \pi f \Omega_{0}^{2} R_{\rm eff} , 4\pi^{2} f^{2} \Omega_{0}^{2} R_{\rm eff}^{2}\right] \,,\\
+	\left| \frac{1}{G_{0}} \partial_{tf} G_{0} \right| &\sim \text{max} \left[ 2 \pi  \Omega_{0} R_{\rm eff}, 4 \pi^{2} f \Omega_{0} R_{\rm eff}^{2} \right] \,,
+\end{align}
+\end{subequations}
+for the orbital response and
+\begin{subequations}\label{eq:estimatederivconst}
+\begin{align}
+	\left| \frac{1}{G_{c}} \partial_{t} G_{c} \right| &\sim \text{max} \left[ \Omega_{0}, 2 \pi f \Omega_{0} L \right] \,,\\
+	\left| \frac{1}{G_{c}} \partial_{tt} G_{c} \right| &\sim \text{max} \left[ \Omega_{0}^{2}, 4 \pi f \Omega_{0}^{2} L, 4\pi^{2} f^{2} \Omega_{0}^{2} L^{2} \right] \,,\\
+	\left| \frac{1}{G_{c}} \partial_{tf} G_{c} \right| &\sim \text{max} \left[ \Omega_{0}/f, 6 \pi \Omega_{0} L, 4\pi^{2} f \Omega_{0}^{2} L^{2} \right] \,,
+\end{align}
+\end{subequations}
+for the constellation response.
+In the presence of different terms, we simply take the maximum of their norms.
+An important point in the above is the presence of delay factors in the forms of powers of $2\pi f d$.
+Thus, the magnitude of the corrections cannot be evaluated with the simple replacement $\partial_{t}^{n} \rightarrow \Omega_{0}^{n}$, and the frequency dependency of the error measures will depend on the frequency being above or below the transfer frequencies $f_{R}$ and $f_{L}$ defined in~\eqref{eq:transferfrequencies}.
+
+The error measure $\epsilon_{d}$ does not depend on signal-dependent timescales and can be directly read off the estimates above, with $\epsilon_{d} = \partial_{tf}G/(2\pi G)$.
+For the error measures $\epsilon_{\Psi}$ and $\epsilon_{A1}$, we can combine the above with the Newtonian timescales given in~\eqref{eq:timescalesN} for the inspiral phase of the signal.
+An important distinction needs to be made there, as the radiation reaction timescale $T_{f}$ grows rapidly towards lower frequencies as $f^{-11/6}$.
+If one considers systems that are arbitrarily far from merging, $T_{f}$ can be made arbitrarily large, leading the error measure $\epsilon_{\Psi}$ to take values arbitrarily larger than 1 where the perturbative treatment we adopted breaks down.
+This limit would describe systems observed in their deep inspiral and very slowly chirping, such as the population of galactic binaries that form one of the science targets of the LISA mission~\cite{} \SM{[C]}.
+Such signals would be very compact in Fourier domain, and are not covered by our formalism that assumes an extended Fourier spectrum.
+We will present in App.~\ref{} \SM{[R]} another version of our formalism based on discrete Fourier coefficients, that could apply better to this case \SM{[check: do we put this appendix or not ?]}.
+On the contrary, if one considers only systems that will merge during or shortly after the observation span of the mission, the starting frequency of the observations will be set by this limited lifetime of the source before merging and we can then bound $\epsilon_{\Psi}$ to be smaller than 1.
+
+In the following we will consider only these merging binaries, which include the supermassive black hole mergers that we expect to detect~\cite{} \SM{[C]}.
+The population of stellar-origin binary black holes, proposed in~\cite{} \SM{[C]}, will have both sources rather far from merger and sources that will merge within a few years after the mission finishes.
+We will take a reference starting frequency corresponding to a time from merger $\Delta t = 10\text{yrs}$.
+Since the corrections turn out to be most important at the beginning of the signal, we will evaluate the timescales at this starting frequency.
+From the Newtonian relation~\eqref{eq:omegaphiN}, we have (using $\nu_{0} = 1/4$ for the equal-mass value): \SM{[check]}
+\be
+	f_{\rm start} = 2.9\times 10^{-5} \text{Hz} \left( \frac{\nu}{\nu_{0}} \right)^{-3/8} \left( \frac{m}{10^{6}M_{\odot}} \right)^{-5/8} \left( \frac{\Delta t}{10 \text{yr}} \right)^{-3/8} \,,
+\ee
+which gives in turn for the timescales
+\begin{subequations}
+\begin{align}
+	T_{f} = 6.8\times10^{-2}\text{yr} \left( \frac{\nu}{\nu_{0}} \right)^{3/16} \left( \frac{m}{10^{6}M_{\odot}} \right)^{5/16} \left( \frac{\Delta t}{10 \text{yr}} \right)^{11/16} \,\\
+	T_{A1} = 2.0\times10^{-4}\text{yr} \left( \frac{\nu}{\nu_{0}} \right)^{3/8} \left( \frac{m}{10^{6}M_{\odot}} \right)^{5/8} \left( \frac{\Delta t}{10 \text{yr}} \right)^{3/8} \,.
+\end{align}
+\end{subequations}
+When inserting these expressions in~\eqref{eq:deffom}, in~\eqref{eq:estimatederivorb} and~\eqref{eq:estimatederivconst} we must also use $f_{\rm start}$ for the possible extra factors of $f$ or $f^{2}$ coming from the delay terms, which can lead to different overall scalings with mass ratio and total mass depending on where the signal starts in the band.
+
+We summarize the results for the error estimates in Fig.~\ref{fig:fomLISA}.
+We show the magnitude of the figures of merit $\epsilon_{\Psi}$, $\epsilon_{A1}$, $\epsilon_{d}$ both for the orbital delay and the constellation response, for an equal-mass system with total masses of $m=10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$ and a starting frequency corresponding to $\Delta t =10 \text{yr}$ of observation.
+Since individual signals can show significant variation depending on the angular parameters, we show an average of $\ln \epsilon$ over the sky position, inclination and polarization, together with $\pm \sigma$ standard deviation, and overlay the estimates resulting from~\eqref{eq:estimatederivorb},~\eqref{eq:estimatederivconst} and~\eqref{eq:timescalesN}.
+Here and in the following, for completeness we show results on the frequency band from $10^{-5}\text{Hz}$ to $1Hz$, while the signals are expected to have little signal-to-noise ratio outside of the band from $10^{-4}\text{Hz}$ to $10^{-1}Hz$, which is sometimes taken as a reference~\cite{} \SM{[C]}.
+
+For $\epsilon_{\Psi}$, we find as expected a steep rise towards lower frequencies, which shows that considering only merging binaries with a limited time remaining before merger is crucial here.
+Due to the different occurences of delay factors, we find a different behaviour between the orbital response, where the initial value of $\epsilon_{\Psi}$ grows towards lower masses, and the constellation response, where $\epsilon_{\Psi}$ grows towards higher masses.
+In both cases, the error measure remains below $\epsilon_{\Psi} \sim 0.1$, where the corrections can be sizeable but well manageable by the perturbative formalism (as we will show in the next Section).
+Not represented in the figure is the behaviour with mass ratio, which also differs: for a total mass of $10^{2} \Msol$, the extra factors of $f_{\rm start}$ due to the delay factors lead to an increase towards larger mass ratios, while for a total mass of $10^{7} \Msol$ we have a decrease towards larger mass ratios.
+
+For $\epsilon_{A1}$, we find that the structure in the waveform close to merger does have a visible but limited impact on the error measure.
+For the orbital response, the error measure remains small for all masses.
+For the constellation response, we find that $\epsilon_{A1}$ can grow up to $\sim 0.01$ at the lower end of the frequency band for high masses.
+Since the timescale $T_{A1}^{\rm N}$ has a simple dependence $\propto 1/f$, these statements should hold for all mass ratios at least during the inspiral.
+
+The $\epsilon_{d}$ is signal-independent, and follows a quite different behaviour for the orbital and constellation responses.
+For the orbital response, $\epsilon_{d_{0}}$ grows for higher frequencies, reaching the $\sim 0.1$ at around $1\text{Hz}$.
+For the constellation response, $\epsilon_{d_{c}}$ is mainly important at lower frequencies, reaching $\sim 0.01$ at around $10^{-5}\text{Hz}$.
+
+Overall, we find that for merging binaries the perturbative approach presented in Sec.~\ref{} \SM{[]} will apply, with the most relevant higher-order corrections being expected to be the phase corrections close to the starting frequency of the signal and the delay correction at high frequencies for the orbital response.
 
@@
   \centering
-  \includegraphics[width=.48\linewidth]{plots/LISAfomdO.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/LISAfomdc.pdf}
+  \includegraphics[width=.32\linewidth]{plots/lisafom_Psiorb.pdf}
+  \includegraphics[width=.32\linewidth]{plots/lisafom_A1orb.pdf}
+  \includegraphics[width=.32\linewidth]{plots/lisafom_dorb.pdf}
   %
-  \includegraphics[width=.48\linewidth]{plots/LISAfomPsi2.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/LISAfomA1.pdf}
-  \caption{Figures of merit of the approximation as defined in~\eqref{eq:deffom}, for an equal-mass, non-spinning system and for total masses of $10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$.
-The upper left panel corresponds to the orbital delay part of the response~\eqref{eq:dfresponseO}, and the upper right panel corresponds to the LISA-centered response~\eqref{eq:dfresponsec}.
+  \includegraphics[width=.32\linewidth]{plots/lisafom_Psiconst.pdf}
+  \includegraphics[width=.32\linewidth]{plots/lisafom_A1const.pdf}
+  \includegraphics[width=.32\linewidth]{plots/lisafom_dconst.pdf}
+  \caption{Error estimates of the approximation as defined in~\eqref{eq:deffom}, for an equal-mass, non-spinning system and for total masses of $10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$.
+The top row corresponds to the orbital delay part of the response~\eqref{eq:dfresponseO}, and the lower row shows the LISA-centered constellation response~\eqref{eq:dfresponsec}.
+The error estimates~\eqref{} \SM{[R]} are shown for the phase, amplitude and delay terms from left to right.
 The central line and interval are the mean and $1\sigma$ standard deviation of the logarithm of $\epsilon$ over 400 random values for the position in the sky, inclination and polarization.
+The starting frequency is set by an observation time of $\Delta t = 10 \text{yrs}$ before merger.
 \SM{[change middle mass to $3\times 10^{4}$ ?
-- clarify mismatch between these $\epsilon$ and the actual size of corrections]} \SM{[got to 6 plots, completely separating orbital and constellation response]}}
+- clarify mismatch between these $\epsilon$ and the actual size of corrections]}}
   \label{fig:fomLISA}
@@
 
-Fig.~\ref{fig:fomLISA} shows the magnitude of the figures of merit $\epsilon$ for different total masses, for the orbital delay, for the rest of the response, for the quadratic-in-phase term and for the first derivative of the amplitude.
-Since individual signals can show significant variation depending on the angular parameters, we show the average over the sky position, inclination and polarization, and $\pm \sigma$ standard deviation.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \SM{[Write explicitly what the orders A1, d1 mean]}
+
+\SM{[argument why we do not compute mismatches]}
+
+\SM{[stress that binaries very far from merger could have arbitrarily large Psi2 corrections]}
+
+Having at hand generic estimates for where the higher-order corrections in the LISA response are expected to be relevant, we now demonstrate that including them indeed reduces the reconstruction error of the Fourier-domain waveform.
+We will use an example PhenomD~\cite{} \SM{[C]} waveform for an equal-mass and non-spinning system, and consider only two values for the total mass, $m=10^{7} \Msol$ and $m=10^{2} \Msol$.
+For this example, we will also focus on the orbital response and on a single-arm constellation response $y_{132}$, as TDI combinations will appear (within the approximations listed in Sec.~\ref{} \SM{[R]}) as linear combinations of these basic observables.
+On one hand, we perform a numerical inverse Fourier transform (IFFT) of the Fourier-domain PhenomD waveform, process the signal through the time-domain response~\eqref{eq:defresponseO} or~\eqref{eq:defresponsec}, and perform a numerical Fourier transform (FFT).
+This gives us the target Frourier-domain waveform, considered to be exact but for the possible artifacts of numerical Fourier transforms (most notably, oscillations induced by the necessary tapering of the signal).
+On the other hand, we process the Fourier-domain signal through the response summarized in Sec.~\ref{} \SM{[R]}, including various higher-order corrections.
+We then compare the output of the two procedures.
+
+We show the results for the orbital response in Figs.~\ref{fig:LISAerrorM1e7orb} and~\ref{fig:LISAerrorM1e2orb} in amplitude and phase form.
+In both cases, in accordance with~\eqref{eq:transferOlocal}, the transfer function is essentially a phase, and contains many more cycles in the low-mass case in the higher frequency band.
+We find that including higher-order corrections does reduce the reconstruction errors, but with the distincitve feature that including only one of the amplitude and delay corrections can make the approximation actually worse, while including both does improve the accuracy.
+In the high-mass case, we also find for all orders of approximation an increase in the errors at the very end.
+This is likely to be due to the fact that the errors in the transfer function are to be interpreted in the relative sense, and the amplitude of the signal is decaying rapidly \SM{[ref to amplitude plot somewhere]} in this post-merger region.
+Although we did not identify the precise cause of this rise in errors, one possible explanation would be tapering-induced oscillations in the numerical Fourier transforms.
+
+For the constellation response, we show the results in Figs.~\ref{fig:LISAerrorM1e7const} and~\ref{fig:LISAerrorM1e2const}, using directly the real and imaginary part of the transfer function.
+We find similar features as for the orbital response: including higher-corrections reduces the errors, but in the high-mass case the amplitude and delay corrections should be added together.
+We also find the same uncontrolled grow of errors at the very end of the signal in the high-mass case, in the region where the amplitude of the signal decays.
 
@@
   \caption{Error in the transfer function for the orbital delay $d_{O}$, for $M=10^{7} \Msol$.}
-  \label{fig:prectoymodel}
+  \label{fig:LISAerrorM1e7orb}
+\end{figure*}
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e2dOamp.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e2dOphase.pdf}
+  %
+  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2dOamp.pdf}
+  \hspace{0.2cm}
+  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2dOphase.pdf}
+  \caption{Transfer function and reconstruction error for the orbital delay $d_{O}$, for $M=10^{2} \Msol$.}
+  \label{fig:LISAerrorM1e2orb}
 \end{figure*}
@@
   \caption{Error in the transfer function for the basic observable $y_{12}$, for $M=10^{7} \Msol$.}
-  \label{fig:prectoymodel}
-\end{figure*}
-
-\begin{figure*}
-  \centering
-  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e2dOamp.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e2dOphase.pdf}
-  %
-  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2dOamp.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2dOphase.pdf}
-  \caption{Transfer function and reconstruction error for the orbital delay $d_{O}$, for $M=10^{2} \Msol$.}
+  \label{fig:LISAerrorM1e7const}
 \end{figure*}
@@
   \caption{Transfer function and reconstruction error for the basic observable $y_{12}$, for $M=10^{2} \Msol$.}
-  \label{fig:prectoymodel}
+  \label{fig:LISAerrorM1e2const}
 \end{figure*}
 
-Fig.~\ref{fig:tdifdA} shows the amplitude and phase errors in the Fourier-domain local response when compared to the FFT of the time-domain response~\eqref{eq:yslr}.
+\SM{[say a word of resampling needed at low and high f, refer to paper PE - discussion section at the end ?]}
 
@@
 
-The basic ingredients of our model \SM{[apologize for the Frankenstein-ism !]} will be:
+The basic ingredients of our model \SM{[apologize for the Frankenstein-ism !]} will be: \SM{[move most of all this to an appendix]}
 \begin{itemize}
@@
 
-First, we would like to stress that the modulation functions $\calD^{\ell *}_{mm'}$. As shown in~\ref{app:Wigner}, their phase dependence is.
+First, we would like to stress that the modulation functions $\calD^{\ell *}_{mm'}$. As shown in~\ref{app:Wigner}, their phase dependence is. \SM{[complete]}
 
@@
 Computed here with $f_{\rm min} = 20\mathrm{Hz}$, for masses between 20 and 400 $\Msol$, using the standard aLIGO noise curve \SM{[reference for that curve]}.
-\SM{[Extend computation to fmin=10 and Mmin=10]} }
+\SM{[Extend computation to fmin=10 and Mmin=10]} \SM{[relate unfaithfulness levels to future detectors in the text]} }
   \label{fig:precunfaithfulness}
@@
 \label{sec:notation}
+
+\SM{[add appendix with discrete Fourier coeffs]}
 
@@
 
-\section{Conventions for the Wigner matrices and precessing frame}
+\section{Wigner matrices and precessing frame}
 \label{app:wigner}

### 2017-10-20 LISAsim/GenerateTDITD.c

@@ -149,7 +149,7 @@ int main(int argc, char *argv[])
 {
   int ret;
 
-  LISAconstellation *variant=&LISA2017;
+  LISAconstellation *variant = &LISA2017;
   /* Initialize structure for parameters */
   GenTDITDparams* params;
   params = (GenTDITDparams*) malloc(sizeof(GenTDITDparams));

### 2017-11-03 docs/paper_method.tex
Edits to the paper - redid the plots for the response.

@@
 
-\usepackage{showlabels}
+%\usepackage{showlabels}
 
@@
 
-It was suggested by a number of authors~\cite{} \SM{[C]} that the effect of the precession can be modeled, to a good approximation, by a time-dependent rotation of a non-precessing waveform.
+It was suggested by a number of authors~\cite{BCV03b, BCPTV05, Schmidt+10, OShaughnessy+11, Boyle+11} that the effect of the precession can be modeled, to a good approximation, by a time-dependent rotation of a non-precessing waveform.
 In this approach, one introduces a precessing frame following the evolution of the plane of the orbit, and approximates the waveform in this precessing frame by using a non-precessing model.
@@
 \end{subequations}
-As for the leading-order time-domain amplitude of the $22$ mode, we have~\cite{} \SM{[C]}
+As for the leading-order time-domain amplitude of the $22$ mode, we have~\cite{BlanchetLiving}
 \be\label{eq:a22N}
@@
 
-Additionally, we limit ourselves to a rigid model for the detector, namely we assume that the constellation remains in an equilateral configuration with fixed armlengths $L=2.5\times 10^{6}\mathrm{km}$, as proposed in~\cite{} \SM{[C]}.
+Additionally, we limit ourselves to a rigid model for the detector, namely we assume that the constellation remains in an equilateral configuration with fixed armlengths $L=2.5\times 10^{6}\mathrm{km}$, as proposed in~\cite{LISA17}.
 This simplified orbits neglect (iv) the effect of excentricity of the individual Keplerian orbits, which introduces corrections at the level $e^{2}$ (v) the effect of gravitational perturbations coming from other celestial bodies, such as the Earth, the quadrupole of the Sun, and the other planets.
@@
 It is natural to define two transfer frequencies for the two relevant length scales for the delays, defined such that a wavelength fits within this length scale.
-Geometrical projection factors aside, these scales are the radius of the orbit around the Sun $R=1\text{au}=1.5\times 10^{8} \text{km}$ and the detector armlength $L=2.5\times 10^{6}\text{km}$ (in the configuration proposed in~\cite{} \SM{[C]}).
+Geometrical projection factors aside, these scales are the radius of the orbit around the Sun $R=1\text{au}=1.5\times 10^{8} \text{km}$ and the detector armlength $L=2.5\times 10^{6}\text{km}$ (in the configuration proposed in~\cite{LISA17}).
 This gives
@@
 If one considers systems that are arbitrarily far from merging, $T_{f}$ can be made arbitrarily large, leading the error measure $\epsilon_{\Psi}$ to take values arbitrarily larger than 1 where the perturbative treatment we adopted breaks down.
-This limit would describe systems observed in their deep inspiral and very slowly chirping, such as the population of galactic binaries that form one of the science targets of the LISA mission~\cite{} \SM{[C]}.
+This limit would describe systems observed in their deep inspiral and very slowly chirping, such as the population of galactic binaries that form one of the science targets of the LISA mission~\cite{LISA17}.
 Such signals would be very compact in Fourier domain, and are not covered by our formalism that assumes an extended Fourier spectrum.
@@
 
-In the following we will consider only these merging binaries, which include the supermassive black hole mergers that we expect to detect~\cite{} \SM{[C]}.
-The population of stellar-origin binary black holes, proposed in~\cite{} \SM{[C]}, will have both sources rather far from merger and sources that will merge within a few years after the mission finishes.
+In the following we will consider only these merging binaries, which include the supermassive black hole mergers that we expect to detect~\cite{LISA17}.
+The population of stellar-origin binary black holes, proposed in~\cite{Sesana16}, will have both sources rather far from merger and sources that will merge within a few years after the mission finishes.
 We will take a reference starting frequency corresponding to a time from merger $\Delta t = 10\text{yrs}$.
@@
 Since individual signals can show significant variation depending on the angular parameters, we show an average of $\ln \epsilon$ over the sky position, inclination and polarization, together with $\pm \sigma$ standard deviation, and overlay the estimates resulting from~\eqref{eq:estimatederivorb},~\eqref{eq:estimatederivconst} and~\eqref{eq:timescalesN}.
-Here and in the following, for completeness we show results on the frequency band from $10^{-5}\text{Hz}$ to $1Hz$, while the signals are expected to have little signal-to-noise ratio outside of the band from $10^{-4}\text{Hz}$ to $10^{-1}Hz$, which is sometimes taken as a reference~\cite{} \SM{[C]}.
+Here and in the following, for completeness we show results on the frequency band from $10^{-5}\text{Hz}$ to $1Hz$, while the signals are expected to have little signal-to-noise ratio outside of the band from $10^{-4}\text{Hz}$ to $10^{-1}Hz$, which is sometimes taken as a reference~\cite{LISA17}.
 
@@
 Having at hand generic estimates for where the higher-order corrections in the LISA response are expected to be relevant, we now demonstrate that including them indeed reduces the reconstruction error of the Fourier-domain waveform.
-We will use an example PhenomD~\cite{} \SM{[C]} waveform for an equal-mass and non-spinning system, and consider only two values for the total mass, $m=10^{7} \Msol$ and $m=10^{2} \Msol$.
+We will use an example PhenomD~\cite{Khan+15,Husa+15} waveform for an equal-mass and non-spinning system, and consider only two values for the total mass, $m=10^{7} \Msol$ and $m=10^{2} \Msol$.
 For this example, we will also focus on the orbital response and on a single-arm constellation response $y_{132}$, as TDI combinations will appear (within the approximations listed in Sec.~\ref{} \SM{[R]}) as linear combinations of these basic observables.
@@
 
-\begin{figure*}
+\begin{figure}
   \centering
-  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e7y12Lamp.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e7y12Lphase.pdf}
+  \includegraphics[width=.98\linewidth]{plots/LISAtransferM1e7y12LReIm.pdf}
   %
-  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e7y12LcorrPsi.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e7y12LcorrAd.pdf}
+  \includegraphics[width=.98\linewidth]{plots/LISAerrorM1e7y12L.pdf}
   \caption{Error in the transfer function for the basic observable $y_{12}$, for $M=10^{7} \Msol$.}
   \label{fig:LISAerrorM1e7const}
-\end{figure*}
-
-\begin{figure*}
+\end{figure}
+
+\begin{figure}
   \centering
-  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e2y12Lamp.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e2y12LReIm.pdf}
+  \includegraphics[width=.98\linewidth]{plots/LISAtransferM1e2y12LReIm.pdf}
   %
-  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2y12LcorrPsi.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2y12LcorrAd.pdf}
+  \includegraphics[width=.98\linewidth]{plots/LISAerrorM1e2y12L.pdf}
   \caption{Transfer function and reconstruction error for the basic observable $y_{12}$, for $M=10^{2} \Msol$.}
   \label{fig:LISAerrorM1e2const}
-\end{figure*}
+\end{figure}
 
@@
 
-As proposed by several authors~\cite{} \SM{[C]}, if one performs the mode decomposition of the waveform~\eqref{eq:defmodes} not in a fixed inertial frame but rather in a time-dependent, rotating frame that follows the plane of the orbit, it is possible to restore much of the structure of a non-precessing waveform.
+As proposed by several authors~\cite{BCV03b, BCPTV05, Schmidt+10, OShaughnessy+11, Boyle+11}, if one performs the mode decomposition of the waveform~\eqref{eq:defmodes} not in a fixed inertial frame but rather in a time-dependent, rotating frame that follows the plane of the orbit, it is possible to restore much of the structure of a non-precessing waveform.
 In particular, one recovers qualitatively the hierarchy of mode amplitudes that prevails for non-precessing systems, with the modes $h_{22}$ and $h_{2,-2}$ being dominant.
@@
 The last angle $\gamma$ parametrizes the remaining freedom of rotation around this radiation axis.
-An important element of the construction of the precessing frame is the fixing of this third degree of freedom, for which we use the minimal rotation condition~\cite{} \SM{[C]}
+An important element of the construction of the precessing frame is the fixing of this third degree of freedom, for which we use the minimal rotation condition~\cite{Boyle+11}
 \be\label{eq:gammadot}
@@
 Several constructions for the precessing frame have been proposed~\cite{} \SM{[C]}.
-In the following, we will use a co-precessing frame based on the dominant eigenvector of the matrix representing the action of the angular momentum operator acting on the waveform modes, proposed in~\cite{} \SM{[C]} (see App.~\ref{app:wigner} for more details).
+In the following, we will use a co-precessing frame based on the dominant eigenvector of the matrix representing the action of the angular momentum operator acting on the waveform modes, proposed in~\cite{OShaughnessy+11} (see App.~\ref{app:wigner} for more details \SM{[R]}).
 
@@
 \end{align}
-When including only the leading harmonic $h^{\rm P}_{22}$ in the precessing-frame waveform (as is done for instance in PhenomP~\cite{} \SM{[C]}), we have for $f<0$
+When including only the leading harmonic $h^{\rm P}_{22}$ in the precessing-frame waveform (as is done for instance in PhenomP~\cite{Hannam+13}), we have for $f<0$
 \begin{subequations}
@@
 The above decomposition~\eqref{eq:wignerrot}, and simplifying assumptions~\eqref{eq:approxsymmetryhlminusm}-\eqref{eq:approxzeronegativef} for the precessing-frame waveform, have been widely used for the existing models of waveforms from precessing binaries.
-For the inspiral phase of the SEOBNRv3 model~\cite{} \SM{[C]}, the frame is assumed to follow the instantaneous orbital plane and the waveform in this frame is represented by a spin-aligned expression, with spin-aligned components that are left to vary according to the precessing dynamics.
+For the inspiral phase of the SEOBNRv3 model~\cite{Pan+13,BTB16}, the frame is assumed to follow the instantaneous orbital plane and the waveform in this frame is represented by a spin-aligned expression, with spin-aligned components that are left to vary according to the precessing dynamics.
 The ringdown phase of the signal is treated differently, the waveform being first rotated back to an inertial frame and then matched to a superposition of quasi-normal modes in this final frame.
-The PhenomP model~\cite{} \SM{[C]} uses the decomposition~\eqref{eq:wignerrot} above (although directly in the Fourier domain, as we will detail below) with a precessing-frame waveform given by a spin-aligned PhenomD waveform~\cite{} \SM{[C]}.
-The SUA waveforms~\cite{} \SM{[C]}, focused on the inspiral phase, also use the decomposition~\eqref{eq:wignerrot} and a post-Newtonian non-precessing model for the modes $h_{\ell m}^{\rm P}$.
+The PhenomP model~\cite{Hannam+13} uses the decomposition~\eqref{eq:wignerrot} above (although directly in the Fourier domain, as we will detail below) with a precessing-frame waveform given by a spin-aligned PhenomD waveform~\cite{Khan+15,Husa+15}.
+The SUA waveforms~\cite{KCY14,Chatziioannou+16,Chatziioannou+17}, focused on the inspiral phase, also use the decomposition~\eqref{eq:wignerrot} and a post-Newtonian non-precessing model for the modes $h_{\ell m}^{\rm P}$.
 
@@
 In the first option, one applies directly the SPA, as described in Sec.~\ref{subsec:SPA}, to the product of the modulation and the signal.
-This is known~\cite{} \SM{[C]} to lead to possible pathologies, since the prefactor $1/\sqrt{\ddot{\phi}_{\rm prec}}$ can blow up due the precessing contributions to the phase, and is not used in waveform modelling.
+This is known (see e.g.~\cite{KCY13}) to lead to possible pathologies, since the prefactor $1/\sqrt{\ddot{\phi}_{\rm prec}}$ can blow up due the precessing contributions to the phase, and is not used in waveform modelling.
 
 The second option corresponds to simply multiplying the Fourier-domain signal by the modulation function evaluated at the time $t_{f}^{\rm SPA}$ as defined in~\eqref{eq:deftfSPA}.
-In the SUA formalism of Ref.~\cite{} \SM{[C]}, this amounts to using a stencil reduced to a single point, i.e.
+In the SUA formalism of Ref.~\cite{KCY14}, this amounts to using a stencil reduced to a single point, i.e.
 using $a_{0,0} = 1$ in~\eqref{} \SM{[R]}.
@@
 \SM{[Mention O'Shaugnessy single-spin FD waveforms]}.
-The PhenomP model~\cite{} \SM{[C]} implictly uses this level of approximation, although with the notable difference that a frequency-based expression of the modulation is used, allowing to cover the whole frequency band.
-The precessing-frame evolution is modelled using post-Newtonian expressions for the Euler angles $(\alpha^{\rm PN}, \beta^{\rm PN}, \gamma^{\rm PN}) (\omega)$, that are obtained as functions of the orbital frequency $\omega$ in the limit of a small opening angle of the precession cone and at the spin-orbit level~\cite{} \SM{[C]}.
+The PhenomP model~\cite{Hannam+13} implictly uses this level of approximation, although with the notable difference that a frequency-based expression of the modulation is used, allowing to cover the whole frequency band.
+The precessing-frame evolution is modelled using post-Newtonian expressions for the Euler angles $(\alpha^{\rm PN}, \beta^{\rm PN}, \gamma^{\rm PN}) (\omega)$, that are obtained as functions of the orbital frequency $\omega$ in the limit of a small opening angle of the precession cone and at the spin-orbit level~\cite{Hannam+13} (see also~\cite{BBF11, MBBB13}).
 Using the SPA-inspired correspondence~\eqref{} \SM{[R]} between the orbital and Fourier-domain frequency, one then writes

### 2017-11-03 EOBNRv2HMROM/GenerateWaveform.c

@@ -300,7 +300,7 @@ int main(int argc, char *argv[])
   if(!(params->tagh22fromfile)) {
     /* Starting frequency corresponding to duration of observation deltatobs */
     if(!(params->deltatobs==0.)) fstartobs = Newtonianfoft(params->m1, params->m2, params->deltatobs);
-      
+
     /* Generate Fourier-domain waveform as a list of hlm modes */
     /* Use TF2 extension, if required to, to arbitrarily low frequencies */
     /* NOTE: at this stage, if no extension is performed, deltatobs and minf are ignored - will start at MfROM */

### 2017-12-18 docs/paper_method.tex
Edits to the paper.

@@
 \newcommand{\tfd}{t_{f}^{d}}
+\newcommand{\tfSPA}{t_{f}^{\rm SPA}}
 
@@
 
-\section{Fourier-domain waveforms, modulations and delays}
+\section{Motivation}
 \label{sec:motivation}
@@
 \ee
+As is well known, GW signals decomposed in spin-weighted spherical harmonic components are smoothly varying functions of frequency in amplitude/phase form.
+An important consequence is that these signal components can be accurately represented in terms of a relatively coarsely sampled frequency grid.
+This property will also extend to the transfer functions, allowing us to keep a compact representation of the full signals.
+(See Appendix~\ref{app:notation} for details of our notation and conventions for Fourier transforms and spherical harmonics.)
 
@@
 In the case of a LISA-type detector, the signal $h(t)$ will simply be the waveform in a fixed heliocentric frame, and the delays will come from the motion of each detector in the wave front, while the modulation will represent the time-variation in the detector orientation.
-\jgb{MOVE:Both can be considered as periodic functions with a period of a year, which allows for an additional rewriting in terms of Fourier series instead of Fourier transforms, as will be developed in Sec.~\ref{}.NOTE: Probably move other technical partso f the par.
-as well.}
-
-\jgb{In full generality computing the transfer function $\calT(f)$ will require a convolution, a costly integral (numerically a sum) over the full frequency domain for each value of $f$.
-For our context though, and perhaps for other applications, there are some properties of $h(t)$, $d(t)$ and $F(t)$ which we can exploit for a more efficient computation.}
-
-\jgb{First consider the waveforms.
-There is a separation between the timescales of orbital motion and radiation-reaction.
-This leads to a general feature of many GW signals, including binary inspirals and black hole mergers, that the signal is relatively localized in time-frequency.
-As far as $h(f)$ is concerned then, for given $f$, only a limited range of times should be relevant in $d(t)$ and $F(t)$ so that we may expect to find a more efficient treatment for the transfer function.}
-
-\jgb{As is well known (refs) this is especially true for non-precessing systems when the signal is decomposed in spin-weighted spherical harmonic components with respect to the source orientation angles.
-In amplitude-phase form the individual spherical harmonic components of even long-lasting signals may be treated as smoothly varying functions of frequency.
-An important consequence is that these signal components can be accurately represented in terms of a relatively coarsely sampled frequency grid, but we will also use this property in transfer function treatment.
-(See Appendix~\ref{sec:notation} for details of our notation and conventions for Fourier transforms and spherical harmonics.)}
-
-\jgb{Next consider the delay and modulation functions.
-Again we can identify a separation of time scales between the modulation or delay and the gravitational-wave period.
-This is either the precession timescale or for the LISA response the annual-orbital motion timescale for the constellation.}
-
-\jgb{Our objective is to find an efficient approximate treatment for $\calT(f)$ which allows us to exploit these properties in a way which is extensible to the high-accuracies which will be required by LISA and other future gravitational-wave instruments, yet which widely applicable to GW analysis, without unnecessary limiting assumptions.
-Before introducing developing our formalism, though, we review the salient features of our application problems.}
-
-\SM{[introduce three timescales: orbital, Trr, 1/f, and say what is to be compared ]}
+
+In full generality computing the transfer function $\calT(f)$ would require a convolution, a costly (discretized) integral over the full frequency domain for each value of $f$.
+For our context though, there are some properties of $h(t)$, $d(t)$ and $F(t)$ which we can exploit for a more efficient computation.
+In particular, we will be able to exploit the separation of the different timescales in the problem.
+
+First, the gravitational waveforms present a clear separation between the timescales of orbital motion and radiation-reaction.
+This leads to a general feature of GW signals from compact binaries, during the inspiral phase but also for black hole mergers, that the signal is relatively localized in time-frequency.
+The localization is particularly clear during the inspiral phase, where the SPA (see Sec.~\ref{subsec:SPA} below) provides an unambiguous time-to-frequency correspondence, however we will find that the applicability of the SPA will not be a limiting factor of our approach.
+
+Second, the delay and modulation functions we consider are much more slowly varying than the GW signal.
+The relevant timescales are either the precession timescale for precessing binaries or the fixed annual-orbital motion timescale for the LISA response.
+This means that the modulation and delay have relatively compact support in the Fourier domain, hence the convolution with the signal will be localized in frequency, which justifies writing its output as a transfer function as in~\eqref{eq:deftransfer}.
+
+Together, these observations let us think that, for a given $f$, only a limited range of times should be relevant in $d(t)$ and $F(t)$ so that we may expect to find a treatment for the transfer function that would be local in time for the modulation and delay.
+Intuitively, for an inspiral signal and in the limit of an extremely slowly varying $d(t)$ and $F(t)$, we should be able to simply evaluate them at the time given by the time-to-frequency correspondence of the SPA.
+Furthermore, a natural quantitative criterion for the applicability of this idea is given by the comparison of the radiation-reaction timescale of the signal with the modulation timescale (possibly augmented by a dimensionless factor when delays are present).
+In other words, the change in frequency of the signal over a characeteristic time of the modulation should be small for the separation of timescales to work.
+
+Our objective is to find an approximate treatment for $\calT(f)$ which allows us to exploit these properties without relying on unnecessary limiting assumptions for the signal (like the limitation of the SPA to inspiral signals), which is extensible to the high-accuracies which will be required by LISA and other future gravitational-wave instruments, while being computationally efficient and widely applicable to GW analysis.
+Before introducing developing our formalism, though, we review the salient features of our application problems.
 
@@
 We refer the reader to~\cite{Vallisneri04} for a comparative account on these various conventions.
-In this work, we will use the conventions of~\cite{Vallisneri04}, which were also used in the Mock LISA Data Challenges (MLDC)~\cite{MLDC4}.
+In this work, we will keep close to the conventions of~\cite{Vallisneri04}, which were also used in the Mock LISA Data Challenges (MLDC)~\cite{MLDC4}.
 
 We use a coordinate system centered on the solar system barycenter (SSB), and we denote the center of the constellation by $O$.
-We introduce the notation $\hatk$ for the propagation vector of the gravitational wave, which we denote by $h_{ij}^{\rm TT}(t)$ in transverse-traceless matrix form, as measured at the SSB (thus, at position $p$, $h_{ij}^{\rm TT}(t, p) = h_{ij}^{\rm TT}(t - \hatk.p)$).
+We introduce the notation $\hatk$ for the propagation vector of the gravitational wave, which we denote by $h_{ij}^{\rm TT}(t)$ in transverse-traceless matrix form, as measured at the SSB (thus, at position $p$, $h_{ij}^{\rm TT}(t, p) = h_{ij}^{\rm TT}(t - \hatk \cdot p)$).
 We denote by $p_{A}$ ($A=1,2,3$) the position of the spacecrafts and $n_{l}$ the unit vectors of the three links, with the convention that $n_{3}$ points from 1 to 2.
@@
 	y_{slr} &\equiv \frac{\nu_{r} - \nu_{s}}{\nu} \nn\\
-	&= \frac{1}{2} \frac{n_{l}^{i}n_{l}^{j}}{1 - \hatk\cdot n_{l}} \left[ h_{ij}^{\rm TT}(t - \hatk\cdot p_{s}) - h_{ij}^{\rm TT}(t - \hatk\cdot p_{r}) \right] \,.
+	&= \frac{1}{2} \frac{n_{l}^{i}n_{l}^{j}}{1 - \hatk\cdot n_{l}} \left[ h_{ij}^{\rm TT}(t - L - \hatk\cdot p_{s}) - h_{ij}^{\rm TT}(t - \hatk\cdot p_{r}) \right] \,.
 \end{align}
 As detailed below in Sec.~\ref{}, we will use a rigid instantaneous model, approximating the geometry of the constellation by a moving equilateral triangle (neglecting the flexing of the arms induced by corrections in the orbits), and evaluating all geometric factors in~\eqref{eq:yslr} at a single time $t$ (neglecting pointing-ahead corrections).
+The additional delay $L$ (we use $c=1$) in the first term represents the light propagation time along the arm between spacecrafts $s$ and $l$.
 
@@
 It will also be useful to separate the delays in two types of terms: the first, $d_{O} = -\hatk\cdot p_{O}$, relates the waveform at the SSB to the waveform at the center of the constellation $O$, whereas the second, $d_{c}$, represents the various delays between the spacecrafts of the constellation.
-For an orbit radius $R=1\,\mathrm{au}$ and an armlength $L=2.5\times10^{6}\mathrm{km}$~\cite{LISA17}, $d_{O} \sim R/c \sim 500\mathrm{s}$ and $d_{c} \sim L/c \sim 8\mathrm{s}$.
+For an orbit radius $R=1\,\mathrm{au}$ and an armlength $L=2.5\times10^{6}\mathrm{km}$~\cite{LISA17}, $d_{O} \sim R/c \simeq 500\mathrm{s}$ and $d_{c} \sim L/c \simeq 8\mathrm{s}$.
 
@@
 It was suggested by a number of authors~\cite{BCV03b, BCPTV05, Schmidt+10, OShaughnessy+11, Boyle+11} that the effect of the precession can be modeled, to a good approximation, by a time-dependent rotation of a non-precessing waveform.
-In this approach, one introduces a precessing frame following the evolution of the plane of the orbit, and approximates the waveform in this precessing frame by using a non-precessing model.
-Different prescriptions have been proposed for the construction of the precessing frame itself~\cite{} \SM{[C]}.
+This allows for a modelling approach where one models separately a precessing frame following the evolution of the plane of the orbit, and approximates the waveform in this precessing frame by using a non-precessing model.
+Different prescriptions have been proposed for the construction of a precessing frame from the waveform itself~\cite{Schmidt+10, OShaughnessy+11, Boyle+11}.
 
@@
 
-As a preliminary step, we recall here an approximation widely used for similar purposes in treating gravitational waves signals emitted by inspiralling binaries, the Stationary Phase Approximation (thereafter SPA).
+As a preliminary step, we recall here an approximation widely used for similar purposes in treating gravitational waves signals emitted by inspiralling binaries, the Stationary Phase Approximation (thereafter SPA, and also sometimes called the steepest-descent method).
 It applies in general to chirping signals, and we refer the reader to~\ref{} for details.
@@
 The approximation then applies to signals verifying the conditions \SM{[say that the third is sometimes missing ?]} \SM{[add reference to Finn-Chernoff, Cutler-Flanagan where the third one is missing]}
-\be\label{eq:conditionSPA}
+\be\label{eq:conditionsSPA}
 	\left| \frac{\dot{a}/a}{\omega} \right| \ll 1\,, \quad \left|\frac{\dot{\omega}}{\omega^{2}} \right| \ll 1\,, \quad \left| \frac{(\dot{a}/a)^{2}}{\dot{\omega}} \right| \ll 1 \,.
@@
 \be\label{eq:deftfSPA}
-	\omega(t_{f}^{\rm SPA}) = \pi  f \,.
+	\omega(\tfSPA) = \pi  f \,.
 \ee
 For a chirping signal of increasing phase, $\omega>0$ and $\dot{\omega}>0$, there is a unique point of stationary phase, located in the positive frequency range $f>0$.
-Using the conditions above, one can formally expand the signal around $t_{f}^{\rm SPA}$ to quadratic order in time, according to
-\begin{align}
-	\tilde{h}_{\rm SPA} (f) &\simeq a(t_{f}^{\rm SPA}) \exp\left[2i\pi f t_{f}^{\rm SPA}-2i\varphi(t_{f}^{\rm SPA}) \right] \nn\\
-	& \qquad \cdot \int \ud t \, e^{-i \dot{\omega} (t_{f}^{\rm SPA}) (t-t_{f}^{\rm SPA})^{2}} \,.
+Using the conditions above, one can formally expand the signal around $\tfSPA$ to quadratic order in time, according to
+\begin{align}
+	\tilde{h}_{\rm SPA} (f) &\simeq a(\tfSPA) \exp\left[2i\pi f \tfSPA-2i\varphi(\tfSPA) \right] \nn\\
+	& \qquad \cdot \int \ud t \, e^{-i \dot{\omega} (\tfSPA) (t-\tfSPA)^{2}} \,.
 \end{align}
 Note that to be able to treat the amplitude as a constant in the integral above, we used the third condition in~\eqref{eq:conditionSPA}.
-The resulting complex Gaussian integral yields\footnote{Note that the expressions given here are valid for the $22$ mode of the waveform.}
+The resulting complex Gaussian integral yields\footnote{Note that the expressions given here are valid for the dominant mode $h_{22}$ of the waveform.
+They can be readily generalized to other modes $h_{\ell m}$ with phase $e^{-im\varphi}$.}
 \begin{subequations}
@@
 	\tilde{h}_{\rm SPA}(f) &= A_{\rm SPA}(f) e^{-i\Psi_{\rm SPA}(f)} \,, \\
-	A_{\rm SPA}(f) &= a(t_{f}^{\rm SPA}) \sqrt{\frac{\pi}{\dot{\omega}(t_{f}^{\rm SPA})}} \,, \\
-	\Psi_{\rm SPA}(f) &= 2\varphi(t_{f}^{\rm SPA}) - 2\pi f t_{f}^{\rm SPA} + \frac{\pi}{4} \,.
+	A_{\rm SPA}(f) &= a(\tfSPA) \sqrt{\frac{\pi}{\dot{\omega}(\tfSPA)}} \,, \\
+	\Psi_{\rm SPA}(f) &= 2\varphi(\tfSPA) - 2\pi f \tfSPA + \frac{\pi}{4} \,. \label{eq:PsiSPA}
 \end{align}
@@
 Although inaccurate for the purpose of waveform modelling, these Newtonian estimates will give useful orders of magnitude of the relevant timescales in the inspiral.
-For a binary with masses $m_{1}, m_{2}$, we define the total mass $m=m_{1}+m_{2}$ and the symmetric mass ratio $\nu = m_{1}m_{2}/m^{2}$.
+For a binary with masses $m_{1}, m_{2}$, we define the total mass $M=m_{1}+m_{2}$ and the symmetric mass ratio $\nu = m_{1}m_{2}/M^{2}$.
 Introducing a time of coalescence $t_{c}$, the relations between the orbital frequency and phase and the time to coalescence $t_{c} - t$ are then given by
@@
 \begin{align}
-	\omega(t) &= \left[ \frac{256\nu}{5c^{5}} (Gm)^{5/3} (t_{c}-t) \right]^{-3/8} \,, \\
-	\varphi(t) &= -\left[ \frac{c^{3}}{G m \nu^{3/5}} (t_{c}-t) \right]^{5/8} \,.
+	\omega(t) &= \left[ \frac{256\nu}{5c^{5}} (GM)^{5/3} (t_{c}-t) \right]^{-3/8} \,, \\
+	\varphi(t) &= -\left[ \frac{c^{3}}{5 G M \nu^{3/5}} (t_{c}-t) \right]^{5/8} \,.
 \end{align}
@@
 \be\label{eq:a22N}
-	a_{22}^{\rm N} (t) = \frac{2 G m \nu x}{D c^{2}} \sqrt{\frac{16 \pi}{5}} \,,
+	a_{22}^{\rm N} (t) = \frac{2 G M \nu x}{D c^{2}} \sqrt{\frac{16 \pi}{5}} \,,
 \ee
@@
 \begin{align}
-	A_{N}(f) &= \frac{G^{2}m^{2} \pi}{Dc^{5}} \sqrt{\frac{2\nu}{3}} v^{-7/2}\,, \label{eq:ASPAN}\\
+	A_{N}(f) &= \frac{G^{2}M^{2} \pi}{Dc^{5}} \sqrt{\frac{2\nu}{3}} v^{-7/2}\,, \label{eq:ASPAN}\\
 	\Psi_{\rm SPA}^{\rm N}(f) &= \phi_{0} - 2\pi f t_{0} - \frac{3}{128\nu v^{5}} \,, \label{eq:PsiSPAN} 
@@
 \end{subequations}
-\SM{[Check sign of $\Psi$ ? Should have $<0$ 2nd der.] }
-where we used the notation $v=(G m \pi f/c^{3})^{1/3}$ ($x=v^{2}$ with the SPA correspondence~\eqref{eq:deftfSPA}), and where $t_{0}, \phi_{0}$ are constants.
-It is also customary to rewrite the above relations in terms of the chirp mass $\calM \equiv m\nu^{3/5}$, which is the only mass combination arising in the phase at the leading PN order.
-
-\SM{[explain how (11) and (12) allow one to directly check conditions (7)]}
-
-\jgb{In the Newtonian, low-frequency limit each of the combinations \eqref{eq:conditionsSPA} assumed to be small in the SPA limit indeed vanishes at $\calO{((v/c)^5)}$, but as the system approaches merger, this condition is no-longer satisfied.
-Near-merger however the treatment is not applicable, so that this is not a viable approach for binary systems beyond insprial.
-Nonetheless the SPA, has been a useful workhorse in many gravitational analysis involving frequency domain transfer functions of the form \eqref{eq:deftransfer}.
+where we used the notation $v=(G M \pi f/c^{3})^{1/3}$ ($x=v^{2}$ with the SPA correspondence~\eqref{eq:deftfSPA}), and where $t_{0}, \phi_{0}$ are constants.
+It is also customary to rewrite the above relations in terms of the chirp mass $\calM \equiv M\nu^{3/5}$, which is the only mass combination characterizing the signal at the leading PN order.
+
+In this Newtonian, low-frequency limit one can check that each of the combinations \eqref{eq:conditionsSPA} indeed vanish at $\calO{(v^5)}$.
+but as the system approaches merger, this condition is no-longer satisfied.
+However, near and after the merger, the SPA treatment is not applicable, as can be seen by the breakdown of the relation~\eqref{eq:deftfSPA} as $\omega$ asymptotes the quasi-normal mode frequency in the ringdown.
+Because of this limitation, we aim at developing a formalism that would not rely on this approximation.
+Nonetheless, the SPA has been a useful workhorse in many gravitational analysis involving frequency domain transfer functions of the form~\eqref{eq:deftransfer}.
 The usual approach is simply to replace any time dependencies appearing the transfer function using \eqref{eq:deftfSPA}.
-We will later return to further consider the SPA treatment in the context of a more general formalism.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Fourier-domain approach to modulations and delays}
+We will later return to further consider the SPA treatment in the context of a more general formalism.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Perturbative Fourier-domain approach to modulations and delays}
 \label{sec:formalism}
@@
 
-\jgb{In this section we develop a Fourier-domain formalism for treating time-delays and temporally multiplicative signal transformations, where the period of primary signal oscilations is short compared to the timescale on which the modulating functions vary.}
+In this section we develop a perturbative Fourier-domain formalism for treating time-delays and temporally multiplicative signal transformations, exploiting the separation of timescales in the problem.
 
@@
 
-\jgb{We begin by expression our delay and modulation function definitions in \eqref{eq:delay-mod-defs} using the Fourier transform (note conventions in \eqref{eq:defFT})},
+We begin by expression our delay and modulation function definitions in \eqref{eq:delay-mod-defs} using the Fourier transform (note our unusal convention~\eqref{eq:defFT}),
 \begin{align}
@@
 \begin{align}
-  \tilde{s}(f) &= F(t)h_{d}(t) \nn \\
+  \tilde{s}(f) &= \mathrm{FT} \left[ F h_{d}\right] (f) \nn \\
   &= \int \ud t \, e^{2i\pi f t} F(t)  \int \ud f' \, e^{-2i\pi f' (t+d(t))}\tilde{h}(f') \nn\\
@@
 \end{align}
-The last equation can then be rewritten as a kind of convolution with a frequency-dependent Kernel, according to
+The last equation can then be rewritten as a generalized convolution integral with a frequency-dependent Kernel, according to
 \be\label{eq:FDkernel}
@@
 	G(f,t) &= e^{-2i\pi f d(t)} F(t) \,, \\
-	\tilde{G}(f,f') &= \int \ud t \, e^{2i\pi f' t} e^{-2i\pi f d(t)} F(t) \,.
+	\tilde{G}(f,f') &= \int \ud t \, e^{2i\pi f' t} G(f,t) \,.
 \label{eq:defGtilde}
@@
 \end{subequations}
-In the absence of delays $d(t)$, as in the case of precessing binaries, the function $G$ loses its frequency-dependence and becomes a modulation in the form of a function of time $F(t)$, and the result~\eqref{eq:FDkernel} above reduces to the usual convolution theorem for the Fourier transform.
-
-\jgb{A direct computation of \eqref{eq:FDkernel} using \eqref{eq:defGtilde} will generally be computationally demanding, but when $\tilde{G}(f,f')$ has only limited support over $f'$, we can seek an accurate but more efficient approximation.} In our approach, two assumptions will be crucial for our approximations to be valid: \SM{[rewrite: not clear, make clear the 'SPA for the integral' argument]}
-\begin{itemize}
-	\item The time scales entering the signal $h(t)$ are well separated from the ones entering the modulation and the delay, so that $\tilde{G}(f-f',f')$ has support limited to $|f'| < f'_{\rm max}$, with $f'_{\rm max}$ small compared the characteristic scale for the variation of the Fourier-domain signal;
-	\item The original signal has a smooth amplitude and phase in the Fourier-domain, slowly varying over the scale $f'_{\rm max}$, so that a local treatment in $f'$ of the convolution in~\eqref{eq:FDkernel} will be possible.
-\end{itemize}
-Although well verified for the case of LISA-type detectors, the first assumption can be a limitation of the approximation in the case of precessing binaries, and we will carefully explore its validity in Sec.~\ref{}.
-
-Notice that these approximations are not exactly the same as the ones usually made to justify the stationary phase approximation (thereafter SPA) for the treatment of the Fourier transform of the inspiral part of the signal.
-Indeed, we will exploit the smoothness of the Fourier-domain $\tilde{h}(f)$ to be able to cover the merger-ringdown part of the signal $h$, where the SPA itself breaks down.
+In the absence of delays $d(t)$, as in the case of precessing binaries, the function $G$ loses its frequency-dependence and becomes a modulation in the form of a function of time $F(t)$, and the result~\eqref{eq:FDkernel} above reduces to the familiar convolution theorem for the Fourier transform.
+
+A direct computation of the generalized convolution~\eqref{eq:FDkernel} using~\eqref{eq:defGtilde} will generally be computationally demanding, but we can exploit the separation of timescales in the problem to seek an accurate but efficient approximation.
+A first argument would be to notice that the Fourier transform $\tilde{G}(f,f')$ should have a compact support, limited to $f' \in [-f_{\rm max}, f_{\rm max}]$ with $f_{\rm max}$ a maximal frequency for the modulation, roughly the inverse of its characteristic timescale.
+This makes the convolution integral~\eqref{eq:FDkernel} localized in frequency, as the waveform is to be evaluated only at frequencies close to $f$.
+We can then approximate $\tilde{h}(f-f')$ by $\tilde{h}(f)$, with a simple Taylor expansion for a perturbative treatment of the difference between the two.
+Doing so, we will recover at leading order a locality in time: the modulation is to be evaluated at a time $\tf$, which will be equivalent to $\tfSPA$ for inspiral signals.
+
+Now, there are two limitations to this straightforward argument.
+The first is that we did not yet specify what $f_{\rm max}$ should be compared with, in order for the approximation to work.
+As shown in Sec.~\ref{subsec:SPA}, the Fourier-domain expressions for the amplitude and phases inspiral signals are steep power-laws.
+We will therefore need a quantitative criterion to ensure $\tilde{h}(f-f')$ does not vary too much on the range $f'\in [-f_{\rm max}, f_{\rm max}]$.
+Second, although clear when considering a fixed characteristic timescale for the modulation and delay (like in the LISA case), the above argument does not apply as such to modulations with a varying timescale (like in the case of precessing binaries, where the precession goes faster when arriving at merger).
+We will show below that what will be relevant is the characteristic timescale at the time $t_{f}$ associated to $f$.
+
+In presence of delays, the above is also complicated by the additional delay phases.
+Depending on the frequency $f$, this can lead $G(f,t)$ to have faster variations that its nominal timescale ($1\mathrm{yr}$ for LISA).
+As we will see in Sec.~\ref{} \SM{[R]}, to decide of the applicability of this approach we will not only need to compare timescales, but also take into account potential dimensionless factors of the form $2\pi f d$.
 
@@
 We show only the high-frequency part of the signal, corresponding to the merger region.
-The PhenomD waveforms have been aligned such that the time-domain amplitude (obtained by an IFFT) peaks at $t=0$.
+The PhenomD waveforms have been aligned such that the time-domain amplitude (obtained by an IFFT \SM{[check acronym]}) peaks at $t=0$.
 The two aligned spins are equal, with components of $0.95$ ($++$, blue), $0$ ($00$, red), and $-0.95$ ($--$, yellow), for mass ratios $q=1$ (full line) and $q=8$ (dashed).
-The vertical lines shows the time-domain instantaneous frequency at the peak $\omega^{22}_{\rm peak}/(2\pi)$, again measured from an IFFT, and the fact that the curves $t_{f}$ do not pass exactly by the crossing with the $t=0$ horizontal line reflects the fact that the link between time-domain frequency and Fourier-domain $f$ is not accurate anymore in the merger part of the signal.
-Note that $t_{f}$ does not go beyond $\sim 30M$ after the peak, and is not monotonous.
+The vertical lines shows the time-domain instantaneous frequency at the peak $\omega^{22}_{\rm peak}/(2\pi)$, again measured from an IFFT, and the fact that the curves $t_{f}$ do not pass exactly by the crossing with the $t=0$ horizontal line reflects the fact that the link between time-domain frequency and Fourier frequency $f$ is only approximate at merger.
+Note that $t_{f}$ increases to at most $\sim 30M$ after the peak, and is not monotonous.
 \SM{[add color to the vertical lines]} \SM{[add dots to highlights the crossings]}}
@@
 
-Under the assumptions above, it is possible to see the convolution in the $f'$ variable in~\eqref{eq:FDkernel} as being a local operation, allowing us to expand the signal $\tilde{h}(f-f')$ around $f$.
-A first leading-order approximation, which could be called the local approximation for reasons that will appear clear below, is to discard the $f'$ dependence in the first argument of $\tilde{g}(f-f', f')$, to keep only the leading order in the Fourier-domain amplitude, and to expand the Fourier-domain phase to the first order, according to
+As a first step, we perform a formal leading-order expansion of the signal $\tilde{h}(f-f')$ around $f$.
+We use the amplitude/phase decomposition~\eqref{eq:defAPsi}, treat the Fourier-domain amplitude $A$ as a constant, expand the Fourier-domain phase $\Psi$ to the first order, and discard the $f'$ dependence in the first argument of $\tilde{G}(f-f', f')$, to obtain
 \be
@@
 \ee
-Here and in the rest of the paper, derivatives with respect to the frequency will always be evaluated at $f$.
-Notice that, because of the freedom of adding a linear term to $\Psi(f)$ by simply shifting the signal in time, no assumption can be made on the smallness of the first derivative of the phase, and this is really a leading order approximation.
+Here and in the rest of the paper, such derivatives with respect to the frequency will always be evaluated at $f$.
 
@@
 \begin{align}
-	\tilde{s}(f) &= \tilde{h}(f) \int \ud f' \, \exp\left[ i f' \frac{\ud \Psi}{\ud f} \right] \tilde{G}(f,f') \nn\\
+	\tilde{s}(f) &\simeq \tilde{h}(f) \int \ud f' \, \exp\left[ i f' \frac{\ud \Psi}{\ud f} \right] \tilde{G}(f,f') \nn\\
 	&= \tilde{h}(f) G\left( f, -\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) \,,
@@
 \be\label{eq:deftf}
-	\tf \equiv -\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \,,
-\ee
-This definition is a generalization of the time-to-frequency correspondence at the heart of the SPA, as will be discussed in Sec.~\ref{subsec:linkSPA}, but it is worth noting that here it is defined entirely from the Fourier-domain waveform, and independently of the SPA being valid or not for the underlying signal $\tilde{h}$.
-As such, this time-of-frequency function also extends to the merger-ringdown part of the signals.
+	\tf \equiv -\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \,.
+\ee
+It is worth noting that a shift in time of the time-domain signal will, by virtue of~\eqref{eq:shifttime}, be appropriately propagated to $t_{f}$.
+Because of the freedom of adding a linear term to $\Psi(f)$ by simply shifting the signal in time, no assumption can be made on the smallness of the first derivative of the phase, and this is really a leading order approximation.
+
+The definition~\eqref{eq:deftf} is a generalization of the time-to-frequency correspondence at the heart of the SPA~\eqref{eq:deftfSPA}.
+Indeed, using~\eqref{eq:deftfSPA} one can verify that the derivative of the SPA phase $\Psi_{\rm SPA}$~\eqref{eq:PsiSPA} with respect to $f$ yields back $\tfSPA$, as
+\be
+	\tfSPA = -\frac{1}{2\pi} \frac{\ud \Psi_{\rm SPA}}{\ud f} \,.
+\ee
+However, our defintion~\eqref{eq:deftf} refers only to the Fourier-domain waveform.
+We do not need to relate the frequency $f$ to a time-domain frequency like the orbital frequency $\omega$, and the defintion is independent of the SPA being valid or not for the underlying signal $\tilde{h}$.
+
+The main advantage of our time-of-frequency function~\eqref{eq:deftf} is that it extends naturally to the merger-ringdown part of the signals.
 Fig.~\ref{fig:tf} shows the behaviour of the time function $\tf$ around merger for six example waveforms, for mass ratios $q=1$ and $q=8$ and for aligned spin components $\chi=0.95,0.,-0.95$.
 In particular, one should note that $\tf$ is not monotonously increasing with frequency anymore when reaching in the high-frequency part of the waveform, corresponding to the ringdown.
-This would forbid an unambiguous definition of an inverse frequency-of-time correspondence $f(t)$.
+Although $t_{f}$ is always a well-defined function of $f$, this non-monotonicity would forbid an unambiguous definition of a reciprocal frequency-of-time function $f(t)$.
 
@@
 \be\label{eq:transferlocal}
-	\calT_{\rm local}(f) = G(f, \tf) \,.
-\ee
-The interpretation of this approximation is straightforward: the signal is simply multiplied by the response function evaluated at the time $\tf$.
-For precessing binaries, this corresponds to the approximation used for the PhenomP approximant~\cite{Hannam+13}, as will be discussed in Sec.~\ref{}.
-Additionally, when delays are present, the frequency-dependence of the response is evaluated directly at $f$.
-
-It is worth noting that a shift in time of the time-domain signal will, by construction, be appropriately propagated to~\eqref{eq:tranferlocal}.
-If $h'(t) = h(t -  \Delta t)$ [but the modulation and delays are left unchanged (which is relevant for space-based detectors, not for precessing binaries)], $\Psi'(f) = \Psi(f) - 2\pi f \Delta t$, so that
-\be\label{eq:localshifted}
-	\tilde{s}'_{\rm local}(f) = \tilde{h}'(f) G(f, \tf + \Delta t) \,. 
-\ee
-This observation also justifies that the first derivative of the Fourier-domain phase cannot, in general, be considered a small quantity, as it can take arbitrary values for arbitrary shifts in time of the signal.
-The leading order approximation corresponds to neglecting the curvature in $\Psi(f)$, and yields a result~\eqref{eq:transferlocal} where the passage to the Fourier domain is done by simply using a time-to-frequency correspondence~\eqref{eq:deftf}, which motivates the name ``local approximation''.
+	\calT_{\rm local}(f) = G(f, \tf) = F(t_{f}) e^{-2i\pi f d(t_{f})}\,.
+\ee
+The interpretation of this approximation is straightforward: the signal is simply multiplied by the response function evaluated at the time $\tf$, the delay phase becoming the same linear phase contribution as one would have in~\eqref{eq:shifttime} with a time shift $d(t_{f})$ treated like a constant.
+The locality in frequency at $f$ for $\tilde{h}$ translates into a locality in time at $t_{f}$ for $F,d$.
 
@@
 
-We now extend our approach beyond the local approximation.
-We will give a formal derivation of the  general formula for the higher-order corrections in~\eqref{eq:FDkernel}, and then specialize it to the special cases of interest.
-
-We will formally Taylor-expand $\tilde{h}(f-f')$ in $f'$ as follows:
-\begin{subequations}
+We now extend our approach beyond the leading-order approximation. A Taylor-expansion of~\eqref{eq:FDkernel} in the variable $f'$ yields:
+\begin{subequations}\label{eq:expandfprime}
 \begin{align}
@@
 	A(f-f') &= A(f) \sum\limits_{q\geq 0} \frac{(-1)^{q}}{q!} {f'}^{q} \frac{1}{A}\frac{\ud^{q} A}{\ud f^{q}} \,, \label{eq:expandA}\\
-	\tilde{G}(f-f', f') &= \sum\limits_{r\geq 0} \frac{(-1)^{r}}{r!} {f'}^{r} \frac{\partial^{r} }{\partial f^{r}}  \tilde{G}(f,f') \label{eq:expandg} \,.
+	\tilde{G}(f-f', f') &= \sum\limits_{r\geq 0} \frac{(-1)^{r}}{r!} {f'}^{r} \frac{\partial^{r} }{\partial f^{r}}  \tilde{G}(f,f') \label{eq:expandG} \,,
 \end{align}
 \end{subequations}
-This expansion will only be usable in practice when we will be able to keep only a few terms in each of these expansions.
-We therefore limited the expansion in $\Psi$ to a finite order $N$, to keep the formula readable.
-Notice that we expand $\tilde{G}(f-f',f')$ in $f'$ only in its first argument.
-We can also interchange the $f$-derivatives of $g$ with its Fourier transform.
-
-\begin{widetext}
-If we further expand the exponentials to obtain a pure $f'$-expansion, by applying the formal derivative rule
+were we used the definition of $t_{f}$ introduced in~\eqref{eq:deftf}.
+In the following, we will only keep the first few terms in each of these expansions, but we keep these expressions in terms of formal series for now.
+Notice that we expand $\tilde{G}(f-f',f')$ in $f'$ only in its first argument, and that we can interchange the $f$-derivatives of $G$ with the Fourier transform operation.
+
+Now, if we further expand the exponential of the phase to obtain a pure $f'$-expansion we can convert the resulting power series in $f'$ into a Taylor series with time derivatives, by applying the formal derivative rule
 \allowdisplaybreaks
 \be
-	\int \ud f'\, {f'}^{n} \frac{\partial^{m} }{\partial f^{m}}  \tilde{G}(f,f') e^{-2i\pi f' \tf} = \frac{1}{(-2i\pi)^{n}} \left( \frac{\partial^{m} }{\partial f^{m}} \frac{\partial^{n} }{\partial t^{n}} G \right)(f,\tf) \,,
-\ee
-we arrive at the formal combined expansion
-\begin{align}
-	\calT(f) &= \sum\limits_{q\geq 0} \sum\limits_{r\geq 0} \sum\limits_{k_{2}\geq 0} \dots \sum\limits_{k_{N}\geq 0} \frac{(-1)^{q}}{q!} \frac{(-1)^{r}}{r!} \prod\limits_{p=2}^{N} \frac{1}{k_{p}!}\left( \frac{(-1)^{p}(-i)}{p!} \frac{\ud^{p}\Psi}{\ud f^{p}}\right)^{k_{p}} \frac{1}{A} \frac{\ud^{q} A}{\ud f ^{q}} \frac{1}{(-2i\pi)^m} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{m} }{\partial t^{m}} G \right)(f,\tf) \,,
-\end{align}
-where $m = p+r+2k_{2}+\dots+N k_{N}$.
-\end{widetext}
-
-This cumbersome expansion will not be used as such, but takes a more intuitive form for special cases.
-Of particular interest for us will be the case where we keep only the first term of the series in~\eqref{eq:expandPsi}, and expand the resulting exponential, which yields
+	\int \ud f'\, {(-2i\pi f')}^{n} \frac{\partial^{m}}{\partial f^{m}} \tilde{G}(f,f') e^{-2i\pi f' \tf} = \frac{\partial^{m} }{\partial f^{m}} \frac{\partial^{n} }{\partial t^{n}} G (f,\tf) \,.
+\ee
+The result is a rather cumbersome expression with multiple sums, that we will not use directly.
+Instead, it will be more instructive to separate the different expansions in~\eqref{eq:expandfprime}.
+Only in Sec.~\ref{subsec:executivesummary} will we come back to combining our results together.
+
+We first consider the effect of the higher-order corrections in~\eqref{eq:expandPsi}.
+We will find that the third and higher derivatives of the phase are always negligible for our purposes, and we will ignore them.
+Keeping only the first term of the series in~\eqref{eq:expandPsi}, corresponding to the second derivative of $\Psi$, and expanding the exponential of the phase, we obtain straightforwardly
 \be\label{eq:resulttaylorPsi}
@@
 \ee
-Indeed, this correction will be shown in Sec.~\ref{subsec:resumquadphase} to be closely related to the previous approach of~\cite{KCY14}.
-The expansion shows that the transfer function is signal-dependent through the second derivative of the phase $\Psi$.
-Similarly, the expansion~\eqref{eq:expandA} gives
+This correction to the transfer function is of particular interest to us.
+It will be quantitatively dominant over the other ones in most contexts, and it will be shown in Sec.~\ref{subsec:resumquadphase} that it generalizes the previous approach of~\cite{KCY14}.
+This result shows that the transfer function is signal-dependent, not only through the time-to-frequency correspondence $t_{f}$ but also through the second derivative of the phase $\Psi$.
+
+Similarly, the expansion of the Fourier-domain amplitude~\eqref{eq:expandA} gives
 \be\label{eq:resulttaylorA}
-	\calT_{\rm amp}(f) = \sum\limits_{q\geq 0} \frac{1}{(2i\pi)^{q}q!} \frac{1}{A} \frac{\ud^{q} A}{\ud f ^{q}}  \left( \frac{\partial^{q} }{\partial t^{q}} G \right)(f,\tf) \,.
-\ee
-If we expand only the frequency-dependence of $G$ as in~\eqref{eq:expandg}, we obtain
+	\calT_{\rm amp}(f) = \sum\limits_{p\geq 0} \frac{1}{(2i\pi)^{p}p!} \frac{1}{A} \frac{\ud^{p} A}{\ud f ^{p}}  \left( \frac{\partial^{p} }{\partial t^{p}} G \right)(f,\tf) \,,
+\ee
+while expanding only the frequency-dependence of $G$ as in~\eqref{eq:expandG} yields
 \be\label{eq:resulttaylordelay}
-	\calT_{\rm delay}(f) = \sum\limits_{r\geq 0} \frac{1}{(2i\pi)^{r}r!} \left( \frac{\partial^{r} }{\partial f^{r}} \frac{\partial^{r} }{\partial t^{r}} G \right)(f,\tf) \,,
-\ee
-which interestingly looks like a Taylor expansion of $G$, but with joint derivatives in frequency and time.
-This time, the expansion is signal-independent as it only depends on the function $G$ and not on $A$, $\Psi$.
-
-In the following, we will rewrite in a more convenient form the formal Taylor expansions for the quadratic phase effects and the delay effects.
-As will be explained in Sec.~\ref{} \SM{[R]}, we will use these Taylor expansions to build error measures designed to keep track of which of these three types of corrections is important.
-
-\SM{[interpretation of $\epsilon$'s as ratios of timescales - next subsec]}
-
-\SM{[add magnitude of the derivatives of G, like $\Omega^{n}$ with possible factors of $2\pi f d$]}
-
-\SM{[add loose convergence argument with scaling $\partial^{n}_{t} \sim \Omega^{n} $]}
-
-\SM{[stress that these are errors in the relative sense - so less important in RD than it seems]}
-
-We found in both the LISA case and in the precession case that the third and higher derivatives of both the amplitude and phase are negligible.
-\SM{[maybe to be recalled below]} Note that the figure of merit $\epsilon_{d}$, introduced to represent the effect of the frequency dependence in the convolution kernel~\eqref{eq:FDkernel}, is signal-independent and depends only on $G$.
+	\calT_{\rm delay}(f) = \sum\limits_{p\geq 0} \frac{1}{(2i\pi)^{p}p!} \left( \frac{\partial^{p} }{\partial f^{p}} \frac{\partial^{p} }{\partial t^{p}} G \right)(f,\tf) \,,
+\ee
+which interestingly looks like a Taylor expansion of $G$, but this time with joint derivatives in frequency and time.
+This last expansion, taken separately from the other corrections, is signal-independent as it only depends on the function $G$ and not on $A$, $\Psi$.
+
+As will be explained in Sec.~\ref{subsec:executivesummary}, we will also use these formal Taylor expansions to build error measures (constructed as the magnitude of the first term ignored in the series, see~\eqref{eq:deffom}) designed to estimate which of these three types of corrections are important to take into account.
 
@@
 The derivatives from which these timescales are built encounter zero-crossings and change sign in the high-frequency range.
-\SM{[discuss discontinuity of third derivative]} \SM{[change legent to Tf/M]}}
+\SM{[discuss discontinuity of third derivative]}\SM{[add vertical line for RD, make merger thick]}}
   \label{fig:TfTA}
@@
 \ee
-The interpretation of this timescale for the inspiralling part of the signal, when the SPA applies, will be discussed below. Whe can use this notation to rewrite~\eqref{eq:resulttaylorPsi} as
+The interpretation of this timescale for the inspiralling part of the signal, when the SPA applies, will be discussed below. We can use this notation to rewrite~\eqref{eq:resulttaylorPsi} as
 \begin{align}\label{eq:resultdffPsiTf}
@@
 
+We can obtain a straighforward physical interpretation of this timescale $\Tf$ by considering inspiral signals for which the SPA is valid. If we define
+\be\label{eq:TfSPA}
+	\Tf^{\rm SPA} \equiv \frac{1}{\sqrt{2\dot{\omega}(\tfSPA)}} \,,
+\ee
+we have, by taking two derivatives of~\eqref{eq:PsiSPA},
+\be
+	\left(\Tf^{\rm SPA} \right)^{2} = -\frac{1}{4\pi^{2}}  \frac{\ud^{2} \Psi_{\rm SPA}}{\ud f^{2}} \,,
+\ee
+where $\ud^{2}\Psi/\ud f^{2} < 0$ in the SPA with our sign conventions.
+The quantity $\Tf^{\rm SPA}$ is clearly a radiation-reaction timescale: the shorter this timescale, the faster the binary chirps to higher frequencies on its quasi-circular inspiral.
+Comparing with~\eqref{eq:defTf} above, we see that our generalized time scale $\Tf$ also reduces to $\Tf^{\rm SPA}$ when the SPA applies.
+
+However, in the same way as our definition~\eqref{eq:deftf} for the time-of-frequency function $t_{f}$ generalizes the SPA definition~\eqref{eq:deftfSPA}, the definition~\eqref{eq:defTf} only refers to the phase of the Fourier-domain signal, does not require to introduce a time-domain frequency like $\omega$, and extends naturally to the merger-ringdown part of the signal.
+In this part of the signal, its physical interpretation as the timescale of radiation reaction is obscured, and the second derivative $\ud^{2}\Psi/\ud f^{2}$ can go through zero and change sign, as shown in Fig.~\ref{fig:TfTA}.
+We include an absolute value in the definition~\eqref{eq:defTf} to allow for this possibility, and keep track of the sign via $\epsilon$.
+
 The expansion in amplitude corrections~\eqref{eq:resulttaylorA} also leads to the natural introduction of a set of other timescales related to the successive derivatives of the amplitude.
-In an analogous manner to the definition~\eqref{eq:defTf} of the timescale $T_{f}$, we can define
+In an analogous manner to the definition~\eqref{eq:defTf} of the timescale $\Tf$, we can define
 \be\label{eq:defTA}
-	\left( T_{An} \right)^{n} \equiv \frac{1}{(2 \pi)^{n}} \frac{1}{A(f)} \left| \frac{\ud^{n} A}{\ud f^{n}} \right| \,,
-\ee
-where we included an absolute value to accomodate the possible sign changes in the right-hand side.
-Although the definition above is for a generic $n\geq 0$, in practice only the first few of these timescales will be relevant.
-In this paper we will use only the first two, $T_{A1}$ and $T_{A2}$.
-
-Of particular interest is the relation between out formal results and the SPA.
-For non-precessing gravitational wave signals, the SPA is a very good approximation through the inspiralling phase~\cite{Droz+99}.
-However, it is important to note that the approximation breaks down when reaching the merger-ringdown part of the signal, as the rapidly varying features of the merger break the locality of the time-to-frequency relation~\eqref{eq:deftfSPA}.
-
-Taking the first derivative of the defining relation~\eqref{eq:deftfSPA} gives
-\be
-	t_{f} = -\frac{1}{2\pi} \frac{\ud \Psi_{\rm SPA}}{\ud f} \,,
-\ee
-which means that our generalized definition~\eqref{eq:deftf} for the effective time $t_{f}$ coincides with $t_{f}^{\rm SPA}$ in the inspiral phase of the signal, where the SPA is applicable.
-Similarly, if we define
-\be\label{eq:TfSPA}
-	T_{f}^{\rm SPA} \equiv \frac{1}{\sqrt{2\dot{\omega}(t_{f})}} \,,
-\ee
-we have \SM{[add more details on Fourier-domain derivatives of SPA amp and phase]}
-\be
-	\left(T_{f}^{\rm SPA} \right)^{2} = -\frac{1}{4\pi^{2}}  \frac{\ud^{2} \Psi_{\rm SPA}}{\ud f^{2}} \,,
-\ee
-where $\ud^{2}\Psi/\ud f^{2} < 0$ in the SPA with our sign conventions.
-Comparing with~\eqref{eq:defTf} above, we see that our generalized time scale $\Tf$ also reduces to $T_{f}^{\rm SPA}$ when the SPA applies.
-However, its definition extends to the merger-ringdown part of the signal, and we have to include an absolute value to allow for a possible change of sign of $\ud^{2}\Psi/\ud f^{2}$ in that regime, as shown in Fig.~\ref{fig:TfTA}.
-
-It is useful, to estimate the magnitude of the corrections with respect to the leading order as we will do in Secs.~\ref{} and~\ref{} below \SM{[R]}, to give an estimate for these timescales based on the Newtonian expressions~\eqref{eq:SPAN}.
-This gives for the Newtonian radiation-reaction and amplitude timescales:
+	\left( T_{Ap} \right)^{p} \equiv \frac{1}{(2 \pi)^{p}} \frac{1}{A(f)} \left| \frac{\ud^{p} A}{\ud f^{p}} \right| \,,
+\ee
+where we included an absolute value to accomodate the possible sign changes in the right-hand side. With this notation, \eqref{} becomes simply
+\be\label{eq:resulttaylorA}
+	\calT_{\rm amp}(f) = \sum\limits_{p\geq 0} \frac{1}{p!} (T_{Ap})^{p}  \left( \partial_{t}^{p} G \right) (f,\tf) \,,
+\ee
+Although the above is written for a generic $p\geq 0$, in practice only the first few of these timescales will be relevant. In this paper we will use only the first two, $T_{A1}$ and $T_{A2}$.
+
+By contrast, the expansion~\eqref{eq:resulttaylordelay} is signal-independent and does not lead to the introduction of new timescales as the coupled time- and frequency derivatives are dimensionless.
+We will see however in Sec.~\ref{subsec:sizecorrLISA} that treating delays requires to take into account dimensionless factors of the type $2\pi f d$.
+
+We can obtain useful estimates for these timescales from the Newtonian expressions~\eqref{eq:SPAN}. This gives for the Newtonian radiation-reaction and amplitude timescales:
 \begin{subequations}\label{eq:timescalesN}
 \begin{align}
-	T_{f}^{\rm N} &= \frac{1}{8} \sqrt{\frac{5}{3\nu}} \frac{G m}{c^{3}} v^{-11/2} \,, \label{eq:TfN}\\
-	T_{A1}^{\rm N} &= \frac{7}{6} \frac{1}{2\pi f} \,. \label{eq:TA1N}
+	\Tf^{\rm N} &= \frac{1}{8} \sqrt{\frac{5}{3\nu}} \frac{G m}{c^{3}} v^{-11/2} \,, \label{eq:TfN}\\
+	T_{A1}^{\rm N} &= \frac{7}{6} \frac{1}{2\pi f}\,, \quad T_{A2}^{\rm N} = \frac{\sqrt{91}}{6} \frac{1}{2\pi f} \,. \label{eq:TA1N-TA2N}
 \end{align}
 \end{subequations}
-One can check explicitly that this expression for $T_{f}$ agrees with~\eqref{eq:TfSPA}.
-With a simple power-law amplitude as in~\eqref{eq:ASPAN}, all higher-order amplitude timescales are also simply proportional to $1/f$ and differ from the first only by their numerical factor.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\section{Quadratic phase corrections and delays}
-%\label{sec:quadphasedelay}
+One can check explicitly that this expression for $\Tf$ agrees with~\eqref{eq:TfSPA}.
+With a simple power-law amplitude as in~\eqref{eq:ASPAN}, all higher-order amplitude timescales are also simply proportional to $1/f$ and differ only by their numerical factor.
+
+We show in Fig.~\ref{fig:TfTA} the timescales $\Tf$, $T_{A1}$, $T_{A2}$ for an equal-mass and non-spinning system.
+In the inspiral, they follow the scalings~\eqref{eq:timescalesN} and $\Tf$ is much larger than the other two.
+For frequencies above the merger frequency, $\Tf$ can go through zero and the amplitude-related timescales become comparable or larger.
 
@@
 
-\SM{[move all of the SUA comparison here]}
-
-In this Section, we focus on the second-derivative corrective term in~\eqref{}.
-
-If we restrict to the case of a pure modulation, $d=0$ and $G(f,t) = F(t)$ [go back to more general $G(f,t)$ ?], we can relate our result~\eqref{eq:resultdffPsiTf} to the approach of Ref.~\cite{KCY14}, where the authors extend the SPA in a formalism called the shifted uniform asymptotic expansion (SUA).
-The derivation includes an expansion of the modulation in Bessel functions, a study of the displacement of the stationary phase point induced by the precession, and a resummation of the result in the form of time derivatives.
-The main result of~\cite{KCY14}, their equation~(34), reads exactly like~\eqref{eq:resultdffPsiTf} with the identifications $\tilde{H}_{corr}(f)\rightarrow \calT(f)$ for the transfer function, $T\rightarrow \Tf$ for the radiation-reaction timescale and $e^{-i\delta\phi} \rightarrow F$ for the modulation function (which is restricted in their framework to include only the phase of the modulation, its amplitude being treated jointly with the amplitude of the signal).
-Thus, we are able to make a direct correspondence between the treatment of~\cite{KCY14} and our framework, which provides us with a simple and alternative derivation of their result.
-Namely, the result of~\cite{KCY14} corresponds to the approximation~\eqref{eq:expandPsi}, where in the Fourier-domain convolution the phase of the signal is expanded to quadratic order and the amplitude is not expanded.
-
-The authors of~\cite{KCY14} proposed a resummation scheme.
-Indeed, the result~\eqref{} looks like a symmetrized Taylor expansion, except for the factors $i^{p}$ and $1/p!$ instead of $1/(2p)!$.
-After having truncated the sum at some finite $N$, one can write (following~\cite{KCY14})
-\begin{align}\label{eq:resumquadphase}
-	\calT(f) &\simeq \sum\limits_{p = 0}^{N} \frac{(-i\epsilon\Tf^{2})^{p}}{2^{p}p!} \partial_{t}^{2p}F(\tf) \nn\\
-	&\simeq \sum\limits_{p= 0}^{N} \sum\limits_{k=0}^{N} a_{N,k}\frac{(k\Tf)^{2p}}{(2p)!}  \partial_{t}^{2p}F(\tf) \nn\\
-	&\simeq \frac{1}{2}\sum\limits_{k=0}^{N} a_{N,k} \left( F(\tf + k\Tf) + F(\tf-k\Tf) \right)\,,
-\end{align}
-where one applies a truncation approximation at the first and last stage.
-The second equality holds provided that the complex coefficients $a_{N,k}$ are a solution of the $N+1$-dimensional linear system~\cite{KCY14}
+If we restrict to the case of a pure modulation, $d=0$ and $G(f,t) = F(t)$, we can relate our result~\eqref{eq:resultdffPsiTf} to the approach of Ref.~\cite{KCY14}, where the authors extend the SPA in a formalism called the Shifted Uniform Asymptotic expansion (SUA) \SM{[look for other reference beyon KCY]}.
+%The derivation includes an expansion of the modulation in Bessel functions, a study of the displacement of the stationary phase point induced by the precession, and a resummation of the result in the form of time derivatives.
+The main intermediate result of~\cite{KCY14}, their equation~(34), reads exactly like~\eqref{eq:resultdffPsiTf} with $\epsilon=1$ and with the identifications $\tilde{H}_{corr}(f)\rightarrow \calT(f)$ for the transfer function, $T\rightarrow \Tf$ for the radiation-reaction timescale and $e^{-i\delta\phi} \rightarrow F$ for the modulation function (which is restricted in their framework to a phase, amplitudes being treated jointly with the amplitude of the signal).
+Thus, our treatment gives a straightforward rederivation of the result of~\cite{KCY14} which corresponds in our framework to the approximation~\eqref{eq:expandPsi}, where in the Fourier-domain convolution the phase of the signal is expanded to quadratic order and the amplitude is not expanded.
+The main difference is that the approach of~\cite{KCY14} still relies on the SPA being valid for the underlying signal.
+
+The authors of~\cite{KCY14} then proposed a resummation scheme for~\eqref{eq:resultdffPsiTf}, using finite differences for the derivatives.
+Indeed, the result~\eqref{eq:resultdffPsiTf} looks like a symmetrized Taylor expansion, except for the factors $i^{p}$ and $1/p!$ instead of $1/(2p)!$.
+Truncating the sum at some finite order $N$, one can write (following~\cite{KCY14})
+\be
+	\calT_{\rm phase}(f) \simeq \sum\limits_{p = 0}^{N} \frac{(-i\epsilon\Tf^{2})^{p}}{2^{p}p!} \partial_{t}^{2p}F(\tf) \simeq \calF_{\Tf}^{N,\epsilon}[F] (\tf) \,,
+\ee
+with the operator $\calF_{T}^{N,\epsilon}$ defined as
+\be
+	\calF_{T}^{N,\epsilon}[F] (t) \equiv \frac{1}{2}\sum\limits_{k=0}^{N} a_{N,k}^{\epsilon} \left( F(t + kT) + F(t - k T) \right) \,,
+\ee
+provided that the complex coefficients $a_{N,k}^{\epsilon}$ are a solution of the $N+1$-dimensional linear system~\cite{KCY14}
 \be\label{eq:stencilsystem}
-	(-i\epsilon)^{p} (2p-1)!! = \sum\limits_{k=0}^{N} a_{N,k} k^{2p} \quad \text{for } p=0,\dots,N \,.
-\ee
+	(-i\epsilon)^{p} (2p-1)!! = \sum\limits_{k=0}^{N} a_{N,k}^{\epsilon} k^{2p} \quad \text{for } p=0,\dots,N \,.
+\ee
+In~\cite{KCY14}, only the case $\epsilon=1$ was considered.
 The two solutions for the stencil coefficients in the two cases $\epsilon = \pm 1$ are simply related by a complex conjugation.
-\SM{[give the rational number expressions of $a_{N,k}$ for the first values of $N$; add derivation in appendix]}
+Explicit expressions for the stencil coefficients $a_{N,k}^{\epsilon}$ are given in App.~\ref{app:stencil} for the first values of $N$.
 
@@
 For more details on quadrature rules when applied to integrals with oscillatory kernels, we refer to~\cite{} \SM{[C]}.
-
-%[TO CORRECT AND REWRITE]
-%The rewriting of~\eqref{eq:resultdffPsiTf} by means of an integral transform leads by analogy to another way of formulating the connection to the results of Ref.~\cite{KCY14}.
-If we consider the inspiralling phase of a precessing binary, the SPA applies to the signal $h(t) = a(t)e^{-2i\varphi(t)}$.
-Refering to~\eqref{} above, if we consider as before that $a(t)$ is slowly variable
-%\be
-%	\tilde{s}(f) \simeq \tilde{h}_{\rm SPA} (f) \simeq a(t_{f}^{\rm SPA}) \exp\left[2i\pi f t_{f}^{\rm SPA}-2i%\varphi(t_{f}^{\rm SPA}) \right]  \int \ud t \, F(t) e^{-i \dot{\omega} (t_{f}^{\rm SPA}) (t-t_{f}^{\rm SPA})^{2}} \,.
-%\ee
-%We see that, within the context of the SPA, the result~\eqref{} for the quadratic term in the phase, equivalent to the result of~\cite{KCY14}, can be recovered by a time-domain approach.
-However, our formalism is somewhat more general, as it does not make use of the SPA for the underlying signal $\tilde{h}(f)$ and can in principle be used for the merger-ringdown part of the signal as well.
 
@@
 
-For the standard LISA configuration, we have for the orbital delays the scaling $d_{O}\sim R \sim 500s$, for the constellation delays the scaling $d_{c}\sim L \sim 8s$ (with $c=1$ and ignoring the dependence on angular factors).
-Since the motion of the constellation is periodic with a period of one year and $\Omega_{0} = 2 \times 10^{-7}\mathrm{rad}.s^{-1}$, we have $\dot{d}_{O} \sim \Omega_{0} R/c \sim 10^{-4}$ and $\dot{d}_{c} \sim \Omega_{0} L/c \sim 1.7 \times 10^{-6}$.
+For the standard LISA configuration, we have for the orbital delays the scaling $d_{O}\sim R/c \sim 500s$, for the constellation delays the scaling $d_{c}\sim L/c \sim 8s$ (with $c=1$ and ignoring the dependence on angular factors).
+Since the motion of the constellation is periodic with a period of one year and $\Omega_{0} \simeq 2 \times 10^{-7}\mathrm{rad}.s^{-1}$, we have $\dot{d}_{O} \sim \Omega_{0} R/c \simeq 10^{-4}$ and $\dot{d}_{c} \sim \Omega_{0} L/c \simeq 1.7 \times 10^{-6}$.
 The smallness of the dimensionless quantity $\dot{d} \ll 1$ (and of its subsequent derivatives) will allow us to treat it perturbatively with a very good approximation, and shows also that the function $t_{d}$ is univalued and that there is no ambiguity in defining its reciprocal $t_{d}^{-1}$.
@@
 
-\subsection{Executive summary of the formalism}
+\subsection{Executive summary of the formalism}\label{subsec:executivesummary}
 
@@
 In each term of this series, the Fresnel transform $\calF^{\epsilon}_{\Tf}$ can in turn be approximated by using the stencil formula~\eqref{eq:Fresnelstencil}.
-The timescales $T_{f}$ and $T_{Ak}$ were defined in~\eqref{eq:defTf} and~\eqref{eq:defTA}, and the generalized time-to-frequency function $t_{f}$ was given in~\eqref{eq:deftf}.
+The timescales $\Tf$ and $T_{Ak}$ were defined in~\eqref{eq:defTf} and~\eqref{eq:defTA}, and the generalized time-to-frequency function $t_{f}$ was given in~\eqref{eq:deftf}.
 The delays $d$ are present only in the LISA context, while in the context of precessing binaries we only have to consider the modulation function $F$.
@@
 
-In the LISA context, due to the smallness of the corrections we will only go up to $k=1$ in~\eqref{eq:transferfinal}, and in computing the remaining time derivatives in~\eqref{eq:transferfinal} we will neglect the second and higher derivatives.
+In the LISA context, due to the smallness of the corrections we will only go up to \SM{[k=2 for prec]} $k=1$ in~\eqref{eq:transferfinal}, and in computing the remaining time derivatives in~\eqref{eq:transferfinal} we will neglect the second and higher derivatives.
 We will also use $F(\tau - d(\tau)) \simeq (F - d \dot{F})(\tau)$.
@@
 
-In the following, it will be convenient to introduce \SM{[think of removing ``figure of merit'' throughout the text as it is not quite the word - error measure ?]} error estimates built from the Taylor-like series~\eqref{eq:resulttaylorPsi},~\eqref{eq:resulttaylorA} and~\eqref{eq:resulttaylordelay}.
+In the following, it will be convenient to introduce error estimates built from the Taylor-like series~\eqref{eq:resulttaylorPsi},~\eqref{eq:resulttaylorA} and~\eqref{eq:resulttaylordelay}.
 We simply define  this error estimates at a certain level of approximation as the magnitude of the first term ignored in the original Taylor series.
@@
 was taken into account]}
+\SM{[interpretation of $\epsilon$'s as ratios of timescales - next subsec]}
+
+\SM{[add magnitude of the derivatives of G, like $\Omega^{n}$ with possible factors of $2\pi f d$]}
+
+\SM{[add loose convergence argument with scaling $\partial^{n}_{t} \sim \Omega^{n} $]}
+
+\SM{[stress that these are errors in the relative sense - so less important in RD than it seems]}
 \begin{subequations}\label{eq:deffom}
 \begin{align}
-	\epsilon_{\Psi 2} &\equiv \frac{1}{2} T_{f}^{2} \left| \frac{1}{G}\partial_{tt}G \right| \,, \\
+	\epsilon_{\Psi 2} &\equiv \frac{1}{2} \Tf^{2} \left| \frac{1}{G}\partial_{tt}G \right| \,, \\
 	\epsilon_{A 1} &\equiv T_{A1} \left| \frac{1}{G} \partial_{t} G \right| \,, \\
@@
 where the function $G$ and its derivatives are evaluated at $(f, t_{f})$.
+Reaching $\epsilon \simeq 1$ will indicate a breakdown of the perturbative approach.
 As we will detail below, in the presence of delays these derivatives do not obey the simple scaling $\partial_{t}^{n} \rightarrow \Omega_{0}^{n}$, with $\Omega_{0}$ the characteristic frequency for the time variablility in the modulation and delay.
@@
 
+We will also \SM{[]}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-As argued in~\cite{Krolak+04}, since we focus on the gravitational-wave contribution to the basic observable, we can use a simplified model for the response, ignoring corrections that are crucial from the point of view of noise cancellations.
+As argued in~\cite{Krolak+04} \SM{[expand a bit; not the only place where this approximation is used]}, since we focus on the gravitational-wave contribution to the basic observable, we can use a simplified model for the response, ignoring corrections that are crucial from the point of view of noise cancellations.
 Our aim is indeed to assess the accuracy of a direct Fourier-domain treatment of the response, thus focusing on the separation of timescales in the problem, keeping in mind that corrections can be added later without affecting the conclusions of the present analysis.
 
-The frequency-shift response for a single link~\eqref{eq:yslr} was derived in Ref.~\cite{EW75} (see also~\cite{} \SM{[C]} for a more modern derivation).
+The frequency-shift response for a single link~\eqref{eq:yslr} was derived in Ref.~\cite{EW75} (see also~\cite{RCP04, Finn08, Cornish09} for more modern derivations).
 Several assumptions enter the result as written in~\eqref{eq:yslr}: (i) effects of the order $v/c$ are neglected, including for instance the special relativistic Doppler effect created by the relative speeds of the spacecrafts on their orbits (ii) the propagation is assumed to take place in a flat spacetime, perturbed only by the gravitational wave; thus the gravitational redshift as well as the deflection of light created by the gravitational potential of the Sun is ignored (iii) all geometric factors are evaluated at a single time, whereas one should consider the beam as propagating from the position of the first spacecraft at the time of emission to the position of the second scapecraft at the time of reception, leading to a pointing-ahead effect.
@@
 
-Note that although we can neglect all the effects (i)-(v) for our present study, keeping track of these corrections is crucial for the purpose of noise cancellations, and led to the development of new generations of TDI observables~\cite{} \SM{[C]}.
-
-Since the orbital delay due to the orbit around the Sun is common to all $y_{slr}$ observables, we will decompose the full instrument response~\eqref{} \SM{[R]} in two steps.
+Note that although we can neglect all the effects (i)-(v) for our present study, keeping track of these corrections is crucial for the purpose of laser noise cancellations, and led to the development of new generations of TDI observables~\cite{} \SM{[C]}.
+
+Since the orbital delay due to the orbit around the Sun is common to all $y_{slr}$ observables, we will decompose the one-arm instrument response~\eqref{} \SM{[R]} in two steps.
 The first step consists simply in applying the varying time delay to bring the signal from the SSB to the center of the LISA triangular constellation.
@@
 \be\label{eq:defresponseO}
-	h_{O} (t) = h(t-\hatk\cdot p_{O}) \,.
+	h_{O}^{\rm TT} (t) = h^{\rm TT}(t-\hatk\cdot p_{O}) \,.
 \ee
@@
 
-Note that, if the corrections~\eqref{} are included for the constellation delays, this transfer function will not have this simple form anymore as $\dot{d}$ will have a different velocity-dependent expression for the sending and receiving spacecrafts.
+Note that, if the corrections~\eqref{} are included for the constellation delays, this transfer function will not have this simple form anymore, as $\dot{d}$ will have a different velocity-dependent expression for the sending and receiving spacecrafts.
 
@@
 \end{subequations}
-These transfer frequencies will indicate where in the LISA band a delay term of the form $2\pi f d$ becomes larger than unity.\SM{[check if this appears elsewhere]}
+These transfer frequencies will indicate where in the LISA band a dimensionless delay term of the form $2\pi f d$ becomes larger than unity. \SM{[check if this appears elsewhere]}
 
@@
 
-Using the Newtonian-order expressions~\eqref{eq:timescalesN} for the signal-dependent timescales $T_{f}$ and $T_{A1}$, we can obtain a useful order-of-magnitude estimate for the error measures $\epsilon$ introduced in~\eqref{eq:deffom}.
+Using the Newtonian-order expressions~\eqref{eq:timescalesN} for the signal-dependent timescales $\Tf$ and $T_{A1}$, we can obtain a useful order-of-magnitude estimate for the error measures $\epsilon$ introduced in~\eqref{eq:deffom}.
 Higher-order amplitude terms beyond the first one in~\eqref{eq:resulttaylorA} will turn out to be completely negligible in the LISA case and we will ignore them in the following.
@@
 For the error measures $\epsilon_{\Psi}$ and $\epsilon_{A1}$, we can combine the above with the Newtonian timescales given in~\eqref{eq:timescalesN} for the inspiral phase of the signal.
-An important distinction needs to be made there, as the radiation reaction timescale $T_{f}$ grows rapidly towards lower frequencies as $f^{-11/6}$.
-If one considers systems that are arbitrarily far from merging, $T_{f}$ can be made arbitrarily large, leading the error measure $\epsilon_{\Psi}$ to take values arbitrarily larger than 1 where the perturbative treatment we adopted breaks down.
+An important distinction needs to be made there, as the radiation reaction timescale $\Tf$ grows rapidly towards lower frequencies as $f^{-11/6}$.
+If one considers systems that are arbitrarily far from merging, $\Tf$ can be made arbitrarily large, leading the error measure $\epsilon_{\Psi}$ to take values arbitrarily larger than 1 where the perturbative treatment we adopted breaks down.
 This limit would describe systems observed in their deep inspiral and very slowly chirping, such as the population of galactic binaries that form one of the science targets of the LISA mission~\cite{LISA17}.
@@
 \begin{align}
-	T_{f} = 6.8\times10^{-2}\text{yr} \left( \frac{\nu}{\nu_{0}} \right)^{3/16} \left( \frac{m}{10^{6}M_{\odot}} \right)^{5/16} \left( \frac{\Delta t}{10 \text{yr}} \right)^{11/16} \,\\
+	\Tf = 6.8\times10^{-2}\text{yr} \left( \frac{\nu}{\nu_{0}} \right)^{3/16} \left( \frac{m}{10^{6}M_{\odot}} \right)^{5/16} \left( \frac{\Delta t}{10 \text{yr}} \right)^{11/16} \,\\
 	T_{A1} = 2.0\times10^{-4}\text{yr} \left( \frac{\nu}{\nu_{0}} \right)^{3/8} \left( \frac{m}{10^{6}M_{\odot}} \right)^{5/8} \left( \frac{\Delta t}{10 \text{yr}} \right)^{3/8} \,.
@@
 \end{subequations}
-When inserting these expressions in~\eqref{eq:deffom}, in~\eqref{eq:estimatederivorb} and~\eqref{eq:estimatederivconst} we must also use $f_{\rm start}$ for the possible extra factors of $f$ or $f^{2}$ coming from the delay terms, which can lead to different overall scalings with mass ratio and total mass depending on where the signal starts in the band.
+\SM{[change notation for total mass everywhere to $M$]} When inserting these expressions in~\eqref{eq:deffom}, in~\eqref{eq:estimatederivorb} and~\eqref{eq:estimatederivconst} we must also use $f_{\rm start}$ for the possible extra factors of $f$ or $f^{2}$ coming from the delay terms, which can lead to different overall scalings with mass ratio and total mass depending on where the signal starts in the band.
 
 We summarize the results for the error estimates in Fig.~\ref{fig:fomLISA}.
-We show the magnitude of the figures of merit $\epsilon_{\Psi}$, $\epsilon_{A1}$, $\epsilon_{d}$ both for the orbital delay and the constellation response, for an equal-mass system with total masses of $m=10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$ and a starting frequency corresponding to $\Delta t =10 \text{yr}$ of observation.
+We show the magnitude of the error estimates $\epsilon_{\Psi}$, $\epsilon_{A1}$, $\epsilon_{d}$ both for the orbital delay and the constellation response, for an equal-mass system with total masses of $m=10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$ and a starting frequency corresponding to $\Delta t =10 \text{yr}$ of observation.
 Since individual signals can show significant variation depending on the angular parameters, we show an average of $\ln \epsilon$ over the sky position, inclination and polarization, together with $\pm \sigma$ standard deviation, and overlay the estimates resulting from~\eqref{eq:estimatederivorb},~\eqref{eq:estimatederivconst} and~\eqref{eq:timescalesN}.
@@
 
+\SM{[say that we take a large number of samples to eliminate spline interpolation errors]}
+
 Having at hand generic estimates for where the higher-order corrections in the LISA response are expected to be relevant, we now demonstrate that including them indeed reduces the reconstruction error of the Fourier-domain waveform.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Alternative approach for periodic modulations and delays}
+
+In the LISA case, the modulations and delays entering~\eqref{eq:yslr} are periodic, with a period of $f_{0} = 1/\mathrm{yr} \simeq 3.2\times 10^{-8}\mathrm{Hz}$.
+If the above perturbative formalism cannot be applied, we can still exploit this periodic structure .
+\SM{[]}
+\be
+	G(f,t) = \sum_{n} c_{n}(f) e^{-in\Omega_{0}t} \,.
+\ee
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	\item unstable SPA: $\calT(f) \tilde{h}^{\rm P}(f) = \mathrm{SPA}\left[ \calD^{*} h^{\rm P} \right](f)$,
-	\item SPA, $0^{\text{th}}$-order SUA, PhenomP: $\calT(f) = \calD^{*}(t_{f}^{\rm SPA}) $,
-	\item SUA: $\calT(f) = \frac{1}{2} \sum\limits_{k} a_{k} \calD^{*}(t_{f}^{\rm SPA} \pm k T_{f}^{\rm SPA})$.
+	\item SPA, $0^{\text{th}}$-order SUA, PhenomP: $\calT(f) = \calD^{*}(\tfSPA) $,
+	\item SUA: $\calT(f) = \frac{1}{2} \sum\limits_{k} a_{k} \calD^{*}(\tfSPA \pm k \Tf^{\rm SPA})$.
 \end{itemize}
@@
 
-The second option corresponds to simply multiplying the Fourier-domain signal by the modulation function evaluated at the time $t_{f}^{\rm SPA}$ as defined in~\eqref{eq:deftfSPA}.
+The second option corresponds to simply multiplying the Fourier-domain signal by the modulation function evaluated at the time $\tfSPA$ as defined in~\eqref{eq:deftfSPA}.
 In the SUA formalism of Ref.~\cite{KCY14}, this amounts to using a stencil reduced to a single point, i.e.
@@
 \be
-	\calT(f) = \frac{1}{2} \sum\limits_{k=0}^{N} a_{N,k} \left[\calD^{\ell *}_{mm'}(t_{f}^{\rm SPA} + k T_{f}^{\rm SPA}) + \calD^{\ell *}_{mm'}(t_{f}^{\rm SPA} - k T_{f}^{\rm SPA}) \right] \,.
+	\calT(f) = \frac{1}{2} \sum\limits_{k=0}^{N} a_{N,k} \left[\calD^{\ell *}_{mm'}(\tfSPA + k \Tf^{\rm SPA}) + \calD^{\ell *}_{mm'}(\tfSPA - k \Tf^{\rm SPA}) \right] \,.
 \ee
 \end{widetext}
-The definition of $T_{f}^{\rm SPA}$ was given in~\ref{} \SM{[R]}.
+The definition of $\Tf^{\rm SPA}$ was given in~\ref{} \SM{[R]}.
 As described in Sec.~\ref{} \SM{[R]}, this approach is equivalent to our formalism when keeping only the quadratic phase correction, but is a priori limited to the inspiral phase of the signal through the definitions of the times $t^{\rm SPA}_{f}$ and $T^{\rm SPA}_{f}$.
@@
 
-For the smoothness of the signal in the Fourier domain, one has to keep in mind that, for waveforms computed starting with a time-domain waveform, Gibbs oscillations can be mitigated by properly tapering the waveform in time domain before performing the FFT, however taking Fourier-domain derivatives magnifies this oscillations, to the point that they would dominate among others the timescale $T_{f}$ (unless one takes much longer waveforms to push this problem to lower frequencies).
+For the smoothness of the signal in the Fourier domain, one has to keep in mind that, for waveforms computed starting with a time-domain waveform, Gibbs oscillations can be mitigated by properly tapering the waveform in time domain before performing the FFT, however taking Fourier-domain derivatives magnifies this oscillations, to the point that they would dominate among others the timescale $\Tf$ (unless one takes much longer waveforms to push this problem to lower frequencies).
 This is avoided in our case by working with a pure Fourier-domain PhenomD waveform that is smooth by construction, but would become important if one were to apply our formalism to the FFT of NR or SEOB waveforms.
@@
 
-In the following, we will use the notation $m=m_{1}+m_{2}$, $\nu=m_{1}m_{2}/m^{2}$, $\delta = (m_{1}-m_{2})/m$, $|\bm{S}_{A}|=Gm_{A}^{2} \chi_{A}$ for $A=1,2$.
+In the following, we will use the notation \SM{[use M instead of m]} $m=m_{1}+m_{2}$, $\nu=m_{1}m_{2}/m^{2}$, $\delta = (m_{1}-m_{2})/m$, $|\bm{S}_{A}|=Gm_{A}^{2} \chi_{A}$ for $A=1,2$.
 We take the convention $m_{1} \geq m_{2}$.
@@
 Here and in the following we present only the $h^{\rm I}_{2 m}$ mode contributions induced by $h^{\rm P}_{22}$, knowing that the ones induced by $h^{\rm P}_{2,-2}$ can be deduced from~\eqref{} \SM{[R]}.
-We compute the Fourier-domain based timescales $T_{f}$, $T_{A1}$ and $T_{A2}$ from the analytic derivatives of the PhenomD phase and amplitude (modified to be of class $C^{2}$).
+We compute the Fourier-domain based timescales $\Tf$, $T_{A1}$ and $T_{A2}$ from the analytic derivatives of the PhenomD phase and amplitude (modified to be of class $C^{2}$).
 The time derivatives of the time-domain modulation are computed from an interpolating cubic spline.
@@
   \includegraphics[width=.32\linewidth]{plots/fom_--_A2.pdf}
-  \caption{Figures of merit of the approximation as defined in~\eqref{eq:deffom} for the precession modulations.
+  \caption{Error estimates of the approximation as defined in~\eqref{eq:deffom} for the precession modulations.
 \SM{[transpose, compact the columns to share the x axis, fix legends, add horizontal line for $10^{0}$]}}
@@
 \section{Notation and conventions}
-\label{sec:notation}
+\label{app:notation}
 
@@
 One can revert to the more usual convention by taking $f\rightarrow -f$.
+
+The effect of a shift in time of the time-domain signal translates into a linear phase contribution added to the Fourier-domain signal.
+For $h_{\Delta t}(t) \equiv h(t+\Delta t)$ with $\Delta t$ a constant, in our convention we have
+\be\label{eq:shifttime}
+	\tilde{h}_{\Delta t} (f) = e^{-2i\pi f \Delta t} \tilde{h}(f) \,.
+\ee
 

### 2017-12-18 tools/waveform.c

@@ -566,8 +566,8 @@ int UnwrapPhase(
   int* cumul = (int*) malloc(sizeof(int) * N);
 
   /* Compute phase mod 2pi (shifted to be between -pi and pi) */
-  for(int i=1; i<N; i++) {
-    pmod[i] = p[i] - floor((p[i] + PI) / (2*PI))*(2*PI) - PI;
+  for(int i=0; i<N; i++) {
+    pmod[i] = p[i] - floor((p[i] + PI) / (2*PI))*(2*PI);
   }
 
   /* Identify jumps */

### 2017-12-19 docs/paper_method.tex
Edits to the paper.

@@
 When neglecting small non-periodic orbital perturbations (such as those caused by the influence of the Earth and of the other planets on the orbits), both $F(t)$ and $d(t)$ are periodic.
-It will also be useful to separate the delays in two types of terms: the first, $d_{O} = -\hatk\cdot p_{O}$, relates the waveform at the SSB to the waveform at the center of the constellation $O$, whereas the second, $d_{c}$, represents the various delays between the spacecrafts of the constellation.
-For an orbit radius $R=1\,\mathrm{au}$ and an armlength $L=2.5\times10^{6}\mathrm{km}$~\cite{LISA17}, $d_{O} \sim R/c \simeq 500\mathrm{s}$ and $d_{c} \sim L/c \simeq 8\mathrm{s}$.
+It will also be useful to separate the delays in two types of terms: the first, $d_{0} = -\hatk\cdot p_{0}$, relates the waveform at the SSB to the waveform at the center of the constellation $O$, whereas the second, $d_{c}$, represents the various delays between the spacecrafts of the constellation.
+For an orbit radius $R=1\,\mathrm{au}$ and an armlength $L=2.5\times10^{6}\mathrm{km}$~\cite{LISA17}, $d_{0} \sim R/c \simeq 500\mathrm{s}$ and $d_{c} \sim L/c \simeq 8\mathrm{s}$.
 
@@
 where we introduced the frequency-dependent function of time $G(f,t)$, and its Fourier transform in the auxiliary frequency $f'$, denoted by $\tilde{G}(f,f')$, as
-\begin{subequations}\label{eq:defg}
+\begin{subequations}\label{eq:defG}
 \begin{align}
@@
 	\tilde{G}(f,f') &= \int \ud t \, e^{2i\pi f' t} G(f,t) \,.
-\label{eq:defGtilde}
 \end{align}
@@
 
-A direct computation of the generalized convolution~\eqref{eq:FDkernel} using~\eqref{eq:defGtilde} will generally be computationally demanding, but we can exploit the separation of timescales in the problem to seek an accurate but efficient approximation.
+A direct computation of the generalized convolution~\eqref{eq:FDkernel} using~\eqref{eq:defG} will generally be computationally demanding, but we can exploit the separation of timescales in the problem to seek an accurate but efficient approximation.
 A first argument would be to notice that the Fourier transform $\tilde{G}(f,f')$ should have a compact support, limited to $f' \in [-f_{\rm max}, f_{\rm max}]$ with $f_{\rm max}$ a maximal frequency for the modulation, roughly the inverse of its characteristic timescale.
@@
 \ee
-where we included an absolute value to accomodate the possible sign changes in the right-hand side. With this notation, \eqref{} becomes simply
+where we included an absolute value to accomodate the possible sign changes in the right-hand side. With this notation, \eqref{eq:resulttaylorPsi} becomes simply
 \be\label{eq:resulttaylorA}
@@
 By contrast, the expansion~\eqref{eq:resulttaylordelay} is signal-independent and does not lead to the introduction of new timescales as the coupled time- and frequency derivatives are dimensionless.
-We will see however in Sec.~\ref{subsec:sizecorrLISA} that treating delays requires to take into account dimensionless factors of the type $2\pi f d$.
+We will see however in Sec.~\ref{subsec:lisafom} that treating delays requires to take into account dimensionless factors of the type $2\pi f d$.
 
@@
 For frequencies above the merger frequency, $\Tf$ can go through zero and the amplitude-related timescales become comparable or larger.
+
+In the following, we will compare these signal-dependent timescales to the timescales present in the modulations and delays.
+In the case of signals from precessing binaries, the precession timescale decreases as the system gets closer to merger, as will be discussed in details in Sec.~\ref{subsec:sizecorrPrec} below.
+In the case of the response of a LISA-like detectors, the modulation and delay evolve with a fixed timescale of one year, as will be detailed in Sec.~\ref{subsec:lisafom}.
 
@@
 Truncating the sum at some finite order $N$, one can write (following~\cite{KCY14})
-\be
+\be\label{eq:stencilresult}
 	\calT_{\rm phase}(f) \simeq \sum\limits_{p = 0}^{N} \frac{(-i\epsilon\Tf^{2})^{p}}{2^{p}p!} \partial_{t}^{2p}F(\tf) \simeq \calF_{\Tf}^{N,\epsilon}[F] (\tf) \,,
@@
 with the operator $\calF_{T}^{N,\epsilon}$ defined as
-\be
+\be\label{eq:stencilfresnel}
 	\calF_{T}^{N,\epsilon}[F] (t) \equiv \frac{1}{2}\sum\limits_{k=0}^{N} a_{N,k}^{\epsilon} \left( F(t + kT) + F(t - k T) \right) \,,
@@
 An immediate advantage of this reformulation is its improved numerical stability.
-For the signal of precessing binaries, although one should expect an ideal scaling $\partial_{t}^{p} F \sim \Omega^{p} F$ with $\Omega$ the precession frequency, it can be extremely hard to control numerical derivatives of $F$ (which in any realistic model will be only provided as sampled at discrete times) well enough to enforce this scaling.
-Here, however, one simply evaluates the original smooth function $F$ at shifted times.
-
-Note that the choice of the samples $\tf \pm k\Tf$ to approximate the successive derivatives of the expansion by finite differences, even if simple and natural, is by no means unique.
-Its structure leads to simple rational expressions for the stencil coefficients $a_{N,k}$, which will be given in~\ref{app:vandermonde} \SM{[refer to table]}.
-In principle, one could have made any ansatz of samples for the series expansion to be matched on.
-One cannot expect all choices to work equally well when applied to noisy data, and the problems of the numerical size of the coefficients and of the overall numerical stability of the scheme must be investigated numerically.
-We leave a more complete investigation of the asymptotics and numerical stability of this resummation approach for future work.
+In waveform modelling applications, it can be bery hard to control high-order numerical derivatives of the modulation.
+Here, one simply evaluates the original smooth modulation function at shifted times.
 
@@
 
-We can give an alternative interpretation of~\eqref{eq:resultdffPsi} and of Sec.~\ref{subsec:resumquadphase} above by using an integral approach.
-For simplicity of notation, here we will again keep to the case of precessing binaries, with only a modulation $F$ and no delays, $d=0$.
+We can give an alternative interpretation of the expansion~\eqref{eq:resultdffPsiTf} and of Sec.~\ref{subsec:resumquadphase} in terms of the Fresnel transform.
+For simplicity of notation, here we keep to the case of a pure modulation $F$ with no delays, $d=0$.
 We will reintroduce the delays below in Sec.~\ref{subsec:delays}.
-To obtain~\eqref{eq:resultdffPsi}, we expanded fully the exponential as a Taylor series.
+To obtain~\eqref{eq:resultdffPsiTf}, we expanded fully the exponential of the phase as a Taylor series.
 If it is not expanded, however, we have
@@
 \ee
-to accomodate for the possible sign change represented by $\epsilon$. The integral~\eqref{eq:integralquadphase} then reads
+to accomodate for the possible sign change represented by $\epsilon$. The integral~\eqref{eq:integralquadphase} then gives for the transfer function
 \be\label{eq:resultFresnel}
-	\tilde{s}(f) = \tilde{h}(f) \calF^{\epsilon}_{\Tf}[F](\tf) \,.
-\ee
-
-The Fresnel transform is an integral transform that shares some properties with the Fourier transform.
-It intervenes in particular in the theory of diffraction~\cite{} \SM{[C]}.
-Contrarily to the Fourier transform, however, it is localized in the sense that the part of the integral that is centered around $t_{0}$ contributes predominantly, due to the cancelling oscillations far from $t_{0}$.
+	\calT_{\rm phase}(f) = \calF^{\epsilon}_{\Tf}[F](\tf) \,.
+\ee
+
+The Fresnel transform~\cite{} \SM{[C]} is an integral transform that shares some properties with the Fourier transform.
+Contrarily to the latter, it is localized in the sense that the part of the integral that is centered around $t_{0}$ contributes predominantly, due to the cancelling oscillations far from $t_{0}$.
+The parameter $\tau$ determines how local the transform is.
+In the limit $\tau\rightarrow 0$, fast oscillations away from the central value $t_{0}$ will cancel out, leading to the integral taking the value $F(t_{0})$.
+For large values of $\tau$, by contrast, the integral~\eqref{eq:defFresnel} has an extended support.
 Note also that only the part of the function $F$ that is symmetric about $t_{0}$ contributes to the integral in~\eqref{eq:defFresnel}.
 
-The parameter $\tau$ of the transform plays the role of a focusing parameter, determining how local the transform is.
-In the limit $\tau\rightarrow 0$, fast oscillations away from the central value $t_{0}$ will cancel out, leading to the integral taking the value $F(t_{0})$.
-For large values of $\tau$, however, the integral~\eqref{eq:defFresnel} has an extended support.
-Transposed in terms of the timescale $\Tf$, since we have seen that $\Tf$ can be interpreted as the radiation reaction timescale in the SPA regime, this means that a faster-chirping signal ($\Tf$ small) will have a Fresnel transform that is more focused, whereas a slower-chirping signal ($\Tf$ large) will have a Fresnel transform that is more extended.
-How much the Fresnel transform is focused in time is to be put in relation with how fast the function $F(t)$ in the integrand is varying.
-In the case of signals from precessing binaries, the precession timescale evolves as the system gets closer to merger, as will be discussed in details in~\ref{subsec:sizecorrPrec} below.
-In the case of the response of a LISA-like detectors, the modulation evolves with a timescale of one year, as will be detailed in Sec.\ref{subsec:sizecorrLISA}.
-
-Although it is interesting to be able to relate our results to a known transformation used in signal processing, whether this opens avenues for more efficient computation methods remains to be investigated, for instance by performing a decomposition of the modulation $F$ in chirplets, which have analytically computable Fresnel transforms.
-Note that in the result~\eqref{eq:resultFresnel}, both the scale $\Tf$ and the central time $\tf$ are functions of the frequency $f$.
-
-As a verification, upon performing a formal Taylor expansion in time of $F(t)$ around $\tf$ in the above integral and integrating term by term the result\footnote{Note that the intermediate integrals produced by such a Taylor expansion are formally divergent.
-One can regularize them for instance by introducing a small imaginary part in $\tau$ that is sent it to $0$ at the end of the computation.}, one readily recovers our previous result~\eqref{eq:resultdffPsiTf}.
-Now, the resummation of~\eqref{eq:resumquadphase} can be rephrased as a quadrature rule for computing the Fresnel integral~\eqref{eq:defFresnel}, according to
-\be\label{eq:Fresnelstencil}
-	\calF_{\tau}^{\epsilon} [F](t_{0}) \simeq \frac{1}{2} \sum\limits_{k=0}^{N} a_{N,k}^{\epsilon} \left( F(t_{0} + k\tau) + F(t_{0} - k\tau) \right)\,,
-\ee
-with $a_{N,k}^{1} = a_{N,k}$ and $a_{N,k}^{-1} = a_{N,k}^{*}$.
-Formally, if one allows for polynomial integrands in~\eqref{eq:defFresnel} (which can only be given sense by introducing a regularization), Eq.~\eqref{eq:Fresnelstencil} amounts to the building a quadrature rule for the particular choice of nodes $t_{0} \pm k \tau$, which is exact (with a regularization) if $F$ is a symmetric polynomial of degree $\leq 2N$.
-Note however that this quadrature rule does not fall into the class of Gaussian quadratures, which provide $N$-point rules that give an exact representation of the integral for polynomials of degree $2N-1$.
-Due to the complex character of the kernel in~\eqref{eq:defFresnel}, the induced inner product is not positive definite and some key properties are lost, such as the construction of the optimal Gaussian nodes from the zeros of the orthonormalized polynomials.
+In our result~\eqref{eq:resultFresnel}, both the scale $\Tf$ and the central time $\tf$ are functions of the frequency $f$.
+Since we have seen that $\Tf$ can be interpreted as the radiation reaction timescale in the SPA regime, this means that a faster-chirping signal ($\Tf$ small) will have a Fresnel transform that is more focused, whereas a slower-chirping signal ($\Tf$ large) will have a Fresnel transform that is more extended.
+In turn, this is to be put in relation with how fast the function $F(t)$ in the integrand is varying.
+In Sec.~\ref{subsec:executivesummary}, we will build an estimate for the magnitude of these phase corrections by comparing of the radiation-reaction timescale to the timescale of variation of the modulation.
+
+Thus, the previous result~\eqref{eq:stencilresult}-\eqref{eq:stencilfresnel} can be rephrased as a quadrature rule, and the stencil $\calF_{T}^{N,\epsilon}$ is an quadrature approximation of the Fresnel transfrom $\calF^{\epsilon}_{T}$.
+If one allows for polynomial integrands\footnote{Note that such integrals with a polynomial integrand are formally divergent.
+One can regularize them for instance by introducing a small imaginary part in $\tau$ that is sent it to $0$ at the end of the computation.} in~\eqref{eq:defFresnel}, using the stencil~\eqref{eq:stencilfresnel} amounts to building a quadrature rule for the particular choice of nodes $\tf \pm k \Tf$, which is exact (with a regularization) if $F$ is a symmetric polynomial of degree $\leq 2N$.
+As a verification, performing a formal Taylor expansion in time of $F(t)$ around $\tf$ in the integral~\eqref{eq:defFresnel} and integrating term by term yields back~\eqref{eq:resultdffPsiTf}.
+
+Note however that this quadrature rule does not fall into the class of Gaussian quadratures, due to the complex character of the kernel in~\eqref{eq:defFresnel}.
+The induced inner product is not positive definite and some key properties are lost, such as the construction of the optimal Gaussian nodes from the zeros of the orthonormalized polynomials.
 For more details on quadrature rules when applied to integrals with oscillatory kernels, we refer to~\cite{} \SM{[C]}.
+Note that the choice of a stencil with quadrature nodes $\tf \pm k\Tf$, even if natural, is by no means unique, and other choices would have led to different stencils.
+The formulation of the result~\eqref{eq:resultdffPsiTf} as a Fresnel transform~\eqref{eq:resultFresnel} opens the way for future investigations of different numerical approaches to the problem.
 
@@
 
-In the case of a LISA-type detector response, the presence of the delay $d(t)$ is responsible for the frequency-dependence of the kernel $G$ introduced in~\eqref{eq:defg}.
-We will show below in Sec.~\ref{} \SM{[R]}that the higher-order correction in the delay can become non-negligible at high frequencies, as measured by the magnitude of the first correction in~\eqref{eq:taylordelay}.
-This is especially true for the orbital delay related to the motion of the whole detector constellation around the Sun, due to the large baseline for this delay.
-
-Here we give an alternative treatment of the delays which is an alternative to the mere Taylor expansion of~\eqref{eq:taylordelay}.
-It is useful to look first at the case where one neglects the signal-related amplitude and phase corrections as given in~\eqref{eq:resultdfA} and~\eqref{eq:resultdffPsiTf}, but keeps the frequency-dependence in the convolution kernel in~\eqref{eq:defg}.
+In the case of a LISA-type detector response, the presence of the delay $d(t)$ is responsible for the frequency-dependence of the kernel $G$ introduced in~\eqref{eq:defG}.
+Here we give a direct treatment of the delays which is an alternative to the mere Taylor expansion of~\eqref{eq:resulttaylordelay}.
+As a first illustration, we leave aside amplitude and phase corrections as given in~\eqref{eq:resulttaylorA} and~\eqref{eq:resulttaylorPsi}.
 The response then depends on the signal only by the time-to-frequency correspondence and reads
@@
 \end{equation}
-where we recall that we defined $t_{f} \equiv -1/(2\pi)\ud \Psi/\ud f$.
-This motivates the introduction of the delayed time function $t_{d}:t \mapsto t+d(t)$, and of the reciprocal function $t_{d}^{-1}$, and of a modified time-to-frequency correspondence $\tfd = t_{d}^{-1}(\tf)$, defined implicitly by
+with $\tf$ defined in~\eqref{eq:deftf}.
+This motivates a change of variable to the delayed time function $t_{d}:t \mapsto t+d(t)$.
+We also define the reciprocal function $t_{d}^{-1}$, and a modified time-to-frequency correspondence $\tfd = t_{d}^{-1}(\tf)$, defined implicitly by
 \be
@@
 \ee
-The above integral can then be formally computed as
+The above integral gives then for the transfer function
 \begin{align}\label{eq:delaycorrleading}
-	\tilde{s}(f) &= \tilde{h}(f) \int \ud t \, F(t) e^{-2i\pi f d(t)} \delta(t + d(t) - t_{f}) \nn \\
-	&= \tilde{h}(f) F(t_{f}^{d}) \frac{e^{-2i\pi f d(t_{f}^{d})}}{1+\dot{d}(t_{f}^{d})}
-\end{align}
-
-For the standard LISA configuration, we have for the orbital delays the scaling $d_{O}\sim R/c \sim 500s$, for the constellation delays the scaling $d_{c}\sim L/c \sim 8s$ (with $c=1$ and ignoring the dependence on angular factors).
-Since the motion of the constellation is periodic with a period of one year and $\Omega_{0} \simeq 2 \times 10^{-7}\mathrm{rad}.s^{-1}$, we have $\dot{d}_{O} \sim \Omega_{0} R/c \simeq 10^{-4}$ and $\dot{d}_{c} \sim \Omega_{0} L/c \simeq 1.7 \times 10^{-6}$.
-The smallness of the dimensionless quantity $\dot{d} \ll 1$ (and of its subsequent derivatives) will allow us to treat it perturbatively with a very good approximation, and shows also that the function $t_{d}$ is univalued and that there is no ambiguity in defining its reciprocal $t_{d}^{-1}$.
+	\calT(f) &= \int \ud t \, F(t) e^{-2i\pi f d(t)} \delta(t + d(t) - t_{f}) \nn \\
+	&= F(t_{f}^{d}) \frac{e^{-2i\pi f d(t_{f}^{d})}}{1+\dot{d}(t_{f}^{d})} \,.
+\end{align}
+
+For the proposed LISA configuration~\cite{LISA17}, we have for the orbital delays the scaling $d_{0}\sim R/c \simeq 500s$, and for the constellation delays the scaling $d_{c}\sim L/c \simeq 8s$ (ignoring the dependence on angular factors).
+Since the motion of the constellation is periodic with a period of one year, with a frequency $\Omega_{0} \simeq 2 \times 10^{-7}\mathrm{rad}.s^{-1}$, we have $\dot{d}_{0} \sim \Omega_{0} R/c \simeq 10^{-4}$ and $\dot{d}_{c} \sim \Omega_{0} L/c \simeq 1.7 \times 10^{-6}$.
+The smallness of the dimensionless quantity $\dot{d} \ll 1$ (and of its subsequent derivatives) will allow us to treat it perturbatively with a very good approximation, and shows also that the function $t_{d}$ is univalued and that there is no ambiguity in defining the reciprocal $t_{d}^{-1}$.
 
@@
 \end{align}
-Now, the most relevant correction in~\eqref{eq:delaycorrleading} comes from the phase factor at high frequencies, where the factors $2\pi f d_{O}$ and $2\pi f d_{c}$ give a magnification by a factor of, respectively, $3.10^{3}$ and $10^{2}$ at $1\Hz$.
-Ignoring the other corrections, we thus arrive at the approximate form for the transfer function
+Now, the most relevant correction in~\eqref{eq:delaycorrleading} comes from the phase factor at high frequencies, where the factors $2\pi f d_{0}$ and $2\pi f d_{c}$ give a magnification factors reaching respectively $3.10^{3}$ and $10^{2}$ at $1\Hz$.
+Ignoring the other corrections, we thus arrive at the following form for the dominant delay correction in the transfer function:
 \be
@@
 \ee
-Notice that this leading-order correction in the treatment of the delays affects purely the phase of the signal.
+This first correction beyond the leading order is signal-independent and affects purely the phase of the output signal.
 
@@
 \end{align}
-\end{widetext}
 where we used a change of variable $\tau = t_{d}(t)$.
 We see that the result can again be expressed as a Fresnel transform.
-When considering amplitude corrections as well, as in~\eqref{}, additional powers of $f'$ can be translated as time derivatives with respect to the variable $\tau$ after performing the change of variables.
-This produces our most general result given in Eq.~\eqref{eq:transferfinal} below.
-
-\subsection{Executive summary of the formalism}\label{subsec:executivesummary}
-
-In this Section, we gather our previous results for the convenience of the reader.
-For a modulation $F$ and delay $d$, so that $s(t) = F(t) h(t+d(t))$, we defined the associated function $G(f,t) = F(t) e^{-2i\pi f d(t)}$ and we obtained the transfer function $\calT (f) = \tilde{s}(f)/\tilde{h}(f)$:
-\begin{widetext}
+When considering amplitude corrections as well, as in~\eqref{eq:resulttaylorA}, additional powers of $f'$ can be translated as time derivatives with respect to the variable $\tau$ after performing the change of variables.
+This produces the final result:
 \be\label{eq:transferfinal}
@@
 \end{widetext}
-In each term of this series, the Fresnel transform $\calF^{\epsilon}_{\Tf}$ can in turn be approximated by using the stencil formula~\eqref{eq:Fresnelstencil}.
+
+\subsection{Executive summary of the formalism}\label{subsec:executivesummary}
+
+In this Section, we gather our previous results for the convenience of the reader, and explain how we will use them in practice.
+For a modulation $F$ and delay $d$, so that $s(t) = F(t) h(t+d(t))$, we obtained the transfer function $\calT (f) = \tilde{s}(f)/\tilde{h}(f)$ given in~\eqref{eq:transferfinal}.
+
+In this result, the Fresnel transform $\calF^{\epsilon}_{\Tf}$ can in turn be approximated by the stencil $\calF^{N,\epsilon}_{\Tf}$ using the formula~\eqref{eq:stencilfresnel}.
 The timescales $\Tf$ and $T_{Ak}$ were defined in~\eqref{eq:defTf} and~\eqref{eq:defTA}, and the generalized time-to-frequency function $t_{f}$ was given in~\eqref{eq:deftf}.
 The delays $d$ are present only in the LISA context, while in the context of precessing binaries we only have to consider the modulation function $F$.
-In practice, only the first few of the terms in this last remaining series expansion are relevant.
+In practice, only the first few of the terms in the series expansion are relevant.
 We will investigate several orders of approximation, combining corrections from the phase, amplitude and delays, and explore which ones are relevant for a given level of accuracy.
-We will use symbols of the form $\{N,A,d\}$ to indicate the order of the stencil in~\eqref{eq:Fresnelstencil}, the maximal order of the amplitude correction included, and the inclusion or not of the delay correction.
-
-In the LISA context, due to the smallness of the corrections we will only go up to \SM{[k=2 for prec]} $k=1$ in~\eqref{eq:transferfinal}, and in computing the remaining time derivatives in~\eqref{eq:transferfinal} we will neglect the second and higher derivatives.
+We will use symbols of the form $\{N,A,d\}$ to indicate the order of the stencil in~\eqref{eq:stencilfresnel}, the maximal order of the amplitude correction included, and the inclusion or not of the delay corrections.
+
+In the LISA context, due to the smallness of the corrections we will only go up to $k=1$ in~\eqref{eq:transferfinal}, and in computing the remaining time derivatives in~\eqref{eq:transferfinal} we will neglect the second and higher derivatives.
 We will also use $F(\tau - d(\tau)) \simeq (F - d \dot{F})(\tau)$.
@@
 
-In the context of precessing binaries, the delays are absent and we will go up to $k=2$ in~\eqref{eq:transferfinal}. The orders of approximation will be:
+In the context of precessing binaries, the delays are absent and we will go up to $k=2$ in~\eqref{eq:transferfinal}. The tranfer function at different orders of approximation will be:
 \begin{subequations}
@@
 We simply define  this error estimates at a certain level of approximation as the magnitude of the first term ignored in the original Taylor series.
-Thus we define \SM{[check if the 1/2!
-was taken into account]}
-\SM{[interpretation of $\epsilon$'s as ratios of timescales - next subsec]}
-
-\SM{[add magnitude of the derivatives of G, like $\Omega^{n}$ with possible factors of $2\pi f d$]}
-
-\SM{[add loose convergence argument with scaling $\partial^{n}_{t} \sim \Omega^{n} $]}
-
-\SM{[stress that these are errors in the relative sense - so less important in RD than it seems]}
+Thus we define, with $G(f,t) = F(t) e^{-2i\pi f d(t)}$,
 \begin{subequations}\label{eq:deffom}
@@
 \end{subequations}
-where the function $G$ and its derivatives are evaluated at $(f, t_{f})$.
-Reaching $\epsilon \simeq 1$ will indicate a breakdown of the perturbative approach.
-As we will detail below, in the presence of delays these derivatives do not obey the simple scaling $\partial_{t}^{n} \rightarrow \Omega_{0}^{n}$, with $\Omega_{0}$ the characteristic frequency for the time variablility in the modulation and delay.
-This is due to the presence of additional dimensionless delay terms of the form $2\pi f d$, which can be larger than 1.
-
-We will also \SM{[]}
+where the function $G$ and its derivatives are evaluated at $(f, t_{f})$. Reaching $\epsilon \simeq 1$ will indicate a breakdown of the perturbative approach.
+
+The physical interpretation of these error measures is clear: when the modulation and delay obey the simple scaling $\partial_{t}^{n} \rightarrow \Omega^{n}$, with $\Omega_{0}$ a characteristic frequency, the phase and amplitude error estimates are simply ratios of timescales.
+For instance, with this scaling $\epsilon_{\Psi 2} \sim T_{f}^{2}\Omega^{2}/2$, so that the approximation will work well when the radiation-reaction timescale is shorter than the characteristic timescale of the modulation.
+However, as we will show in Sec.~\ref{subsec:lisafom} this simple picture will need to be refined in the presence of delays, due to the presence of additional dimensionless factors of the form $2\pi f d$, which can be larger than 1.
+
+On top of this perturbative formalism, in the LISA case we will also develop another approach exploiting the periodicity of the modulation and delays (see Sec.~\ref{subsec:comblisa}), while in the case of precessing binaries we will use a trigonometric polynomial approach to represent the merger-ringdown part of the signal (see Sec.~\ref{subsec:trigopoly})
 
@@
 \be\label{eq:defresponseO}
-	h_{O}^{\rm TT} (t) = h^{\rm TT}(t-\hatk\cdot p_{O}) \,.
+	h_{0}^{\rm TT} (t) = h^{\rm TT}(t-\hatk\cdot p_{0}) \,.
 \ee
@@
 	y_{slr} &= \frac{1}{2} \frac{1}{1 - \hatk\cdot n_{l}} \nn\\
-	& \cdot n_{l}\cdot \left[ h_{O}^{\rm TT}(t - L - \hatk\cdot p^{L}_{s}) - h_{O}^{\rm TT}(t - \hatk\cdot p^{L}_{r}) \right] \cdot n_{l}\,,
-\end{align}
-where we introduced the positions of the spacecrafts relative to the center of the constellation, $p^{L}_{A} \equiv p_{A} - p_{O}$.
+	& \cdot n_{l}\cdot \left[ h_{0}^{\rm TT}(t - L - \hatk\cdot p^{L}_{s}) - h_{0}^{\rm TT}(t - \hatk\cdot p^{L}_{r}) \right] \cdot n_{l}\,,
+\end{align}
+where we introduced the positions of the spacecrafts relative to the center of the constellation, $p^{L}_{A} \equiv p_{A} - p_{0}$.
 
@@
 \be\label{eq:transferOlocal}
-	\calT_{O}^{\rm local}(f) = e^{-2i\pi f d_{O}(t_{f})}\,,
-\ee
-with $d_{O} = -\hatk \cdot p_{O}$ the orbital delay.
+	\calT_{0}^{\rm local}(f) = e^{-2i\pi f d_{0}(t_{f})}\,,
+\ee
+with $d_{0} = -\hatk \cdot p_{0}$ the orbital delay.
 When using the leading-order response~\eqref{eq:transferlocal} for the constellation response, the delays take similarly the form of pure phase factors.
@@
 We used the rigid-arms approximation relation $p^{L}_{r} - p^{L}_{s} =  L n_{l}$.
-The result~\eqref{eq:transferclocal} is familiar, as it was derived in the past taking the point of view of monochromatic signals (see e.g.~\cite{}).
+\SM{[rewrite this]} The result~\eqref{eq:transferlocal} is familiar, as it was derived in the past taking the point of view of monochromatic signals (see e.g.~\cite{} \SM{[C]}).
 
@@
 \subsection{Estimates for the magnitude of higher-order corrections}
-\label{subsec:sizecorrLISA}
+\label{subsec:lisafom}
 
@@
 
-We start with the orbital response, which takes the form of a pure delay $G_{0}(f, t) = e^{-2 i \pi f d_{O}(t)}$, and obtain
+We start with the orbital response, which takes the form of a pure delay $G_{0}(f, t) = e^{-2 i \pi f d_{0}(t)}$, and obtain
 \begin{subequations}
 \begin{align}
-	\frac{1}{G_{0}} \partial_{t} G_{0} &= -2i\pi f \dot{d}_{O}\,,\\
-	\frac{1}{G_{0}} \partial_{tt} G_{0} &= -2i\pi f \ddot{d}_{O} - 4\pi^{2} f^{2} \dot{d}_{O}^{2} \,,\\
-	\frac{1}{G_{0}} \partial_{tf} G_{0} &= -2 i \pi \dot{d}_{O} - 4\pi^{2} f d_{O} \dot{d}_{O} \,,
+	\frac{1}{G_{0}} \partial_{t} G_{0} &= -2i\pi f \dot{d}_{0}\,,\\
+	\frac{1}{G_{0}} \partial_{tt} G_{0} &= -2i\pi f \ddot{d}_{0} - 4\pi^{2} f^{2} \dot{d}_{0}^{2} \,,\\
+	\frac{1}{G_{0}} \partial_{tf} G_{0} &= -2 i \pi \dot{d}_{0} - 4\pi^{2} f d_{0} \dot{d}_{0} \,,
 \end{align}
@@
   \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e7dOphase.pdf}
-  \caption{Error in the transfer function for the orbital delay $d_{O}$, for $M=10^{7} \Msol$.}
+  \caption{Error in the transfer function for the orbital delay $d_{0}$, for $M=10^{7} \Msol$.}
   \label{fig:LISAerrorM1e7orb}
@@
   \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2dOphase.pdf}
-  \caption{Transfer function and reconstruction error for the orbital delay $d_{O}$, for $M=10^{2} \Msol$.}
+  \caption{Transfer function and reconstruction error for the orbital delay $d_{0}$, for $M=10^{2} \Msol$.}
   \label{fig:LISAerrorM1e2orb}
@@
 
-\subsection{Alternative approach for periodic modulations and delays}
+\subsection{Direct approach for periodic modulations and delays}\label{subsec:comblisa}
 
@@
 \subsection{Direct convolution approach for the merger-ringdown phase}
-\label{subsec:convolution}
+\label{subsec:trigopoly}
 
@@
 
-In the following, we will either present the errors obtained with the formalism of Sec.~\ref{sec:formalism} applied to the whole frequency band, including the merger and ringdown part of the signal, or cover the high-frequency range with the convolution treatment of Sec.~\ref{subsec:convolution}, with a Planckian window on $Mf\in [Mf_{0}, Mf_{\rm knee}]$ for a smooth transition between the two formalisms.
+In the following, we will either present the errors obtained with the formalism of Sec.~\ref{sec:formalism} applied to the whole frequency band, including the merger and ringdown part of the signal, or cover the high-frequency range with the convolution treatment of Sec.~\ref{subsec:trigopoly}, with a Planckian window on $Mf\in [Mf_{0}, Mf_{\rm knee}]$ for a smooth transition between the two formalisms.
 
@@
 The symbol $A:n$ with $n=0,1,2$ indicates the order at which the amplitude corrections in~\eqref{eq:transferfinal} are included.
-Additionally, for all these approximation orders we either apply the perturbative formalism of Sec.~\ref{sec:formalism} to the whole frequency band (symbol ``no conv''), or treat the high-frequency part with the convolution formalism of Sec.~\ref{subsec:convolution} (symbol ``conv'').
+Additionally, for all these approximation orders we either apply the perturbative formalism of Sec.~\ref{sec:formalism} to the whole frequency band (symbol ``no conv''), or treat the high-frequency part with the convolution formalism of Sec.~\ref{subsec:trigopoly} (symbol ``conv'').
 
@@
 
+\section{Explicit expression for the stencil coefficients}
+\label{app:stencil}
+
+\begin{table}[t]
+\begin{ruledtabular}\caption{Stencil coefficients $a_{N,k}^{\epsilon}$ entering the formula~\eqref{eq:stencilfresnel}, given here for $\epsilon=1$.
+The case $\epsilon=-1$ is obtained by complex conjugation.}\label{table:stencil}
+\begin{tabular}{c|cccccc}
+	$N \backslash k$ & $0$ & $1$ & $2$ & $3$ & $4$ & $5$ \\
+	\hline
+	$0$ & $1$ & - & - & - & - & - \\
+	$1$ & $1+i$ & $-i$ & - & - & - & - \\
+	$2$ & $\frac{1+5i}{4}$ & $\frac{3-4i}{3}$ & $\frac{-3+i}{12}$ & - & - & - \\
+	$3$ & $\frac{17 i-3}{18}$ & $\frac{13-7 i}{8}$ & $\frac{-5-i}{10}$ & $\frac{11 i+15}{360}$ & - & - \\
+	$4$ & $\frac{185 i-69}{288}$ & $\frac{209-47 i}{120}$ & $\frac{-41 i-67}{120}$ & $\frac{251 i+147}{2520}$ & $\frac{-29 i-7}{3360}$ & \\
+	$5$ & $\frac{1669 i-690}{3600}$ & $\frac{2393-135 i}{1440}$ & $\frac{-645 i-646}{1260}$ & $\frac{3295 i+831}{20160}$ & $\frac{26-345 i}{15120}$ & $\frac{429 i-115}{302400}$ \\
+\end{tabular}
+\end{ruledtabular}
+\end{table}
+
+In this Appendix, we give explicit expressions for the stencil coefficients entering~\eqref{eq:stencilfresnel}.
+In this work we used only low order stencils with usually $N\leq 5$, and one can simply numerically invert of the system~\eqref{eq:stencilsystem}, as this operation has to be done only once.
+For future reference, we give here exact expressions, as the specific choice of samples at $\pm kT$, makes the system~\eqref{eq:stencilsystem} take the form of a Vandermonde system.
+Defining the Vandermonde matrix $V(x_{0},\dots,x_{N})$ as $V_{ij} = (x_{j})^{i}$, setting $x_{j} = j^{2}$, and $b_{p} \equiv (-i\epsilon)^{p}(2p-1)!!$, the linear system~\eqref{eq:stencilsystem} becomes
+\be
+	b_{p} = \sum\limits_{k=0}^{N} V_{p,k} a_{N,k}^{\epsilon} \,.
+\ee
+%Now, since the Vandermonde system can be seen as a reformulation of the polynomial interpolation problem, we know that $({}^{t}V^{-1})_{ij}$ is the coefficient of $X^{i}$ in the Lagrange interpolation polynomial $\prod_{k\neq j} (X-x_{k})/(x_{j} - x_{k})$.
+If we introduce the symmetric polynomials $\sigma$ such that
+%\be	
+%	\sigma(m, \{x_{1}, \dots, x_{n}\}) = \sum\limits_{1\leq i_{1}<\dots<i_{m}\leq n} x_{i_{1}}\dots x_{i_{m}} \,,
+%\ee
+%we obtain for the inverse of the matrix $V$
+%\be
+%	V^{-1}_{ij} = \frac{1}{\prod_{k\neq i} (i^{2} - k^{2})} (-1)^{N-j} \sigma(N-j, \{0,\dots,N^{2}\}\backslash \{i^{2}\}) \,.
+%\ee
+The expression of the inverse of a Vandermonde matrix in terms of symmetric polynomials then allows us to write a closed-form (but rather impractical) expression for the $a_{N,k}^{\epsilon}$, for every finite order $N$:
+\begin{align}
+	a_{N,k}^{\epsilon} &= \frac{1}{\prod\limits_{\substack{q=0 \\ q\neq k}}^{N} (q^{2}-k^{2})} \sum\limits_{p=0}^{N} (i\epsilon)^{p}(2p-1)!!
+\nn\\ & \quad \cdot \sum\limits_{\substack{ 0 \leq j_{1} < \dots < j_{N-p} \leq N \\ j_{1}, \dots, j_{N-p} \neq k}} j_{1}^{2}\dots j_{N-p}^{2}
+\end{align}
+The two cases $\epsilon\pm 1$ correspond simply to a complex conjugation of the coefficients $a_{N,k}^{\epsilon}$.
+Table~\ref{table:stencil} gives the resulting rational values for these coefficients for $N\leq 5$.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-12-20 docs/paper_method.tex
Edits to the paper.

@@
 \be\label{eq:stencilresult}
-	\calT_{\rm phase}(f) \simeq \sum\limits_{p = 0}^{N} \frac{(-i\epsilon\Tf^{2})^{p}}{2^{p}p!} \partial_{t}^{2p}F(\tf) \simeq \calF_{\Tf}^{N,\epsilon}[F] (\tf) \,,
-\ee
-with the operator $\calF_{T}^{N,\epsilon}$ defined as
+	\calT_{\rm phase}(f) \simeq \sum\limits_{p = 0}^{N} \frac{(-i\epsilon\Tf^{2})^{p}}{2^{p}p!} \partial_{t}^{2p}F(\tf) \simeq \calF^{N}_{\Tf, \epsilon}[F] (\tf) \,,
+\ee
+with the operator $\calF_{T, \epsilon}^{N}$ defined as
 \be\label{eq:stencilfresnel}
-	\calF_{T}^{N,\epsilon}[F] (t) \equiv \frac{1}{2}\sum\limits_{k=0}^{N} a_{N,k}^{\epsilon} \left( F(t + kT) + F(t - k T) \right) \,,
+	\calF_{T, \epsilon}^{N}[F] (t) \equiv \frac{1}{2}\sum\limits_{k=0}^{N} a_{N,k}^{\epsilon} \left( F(t + kT) + F(t - k T) \right) \,,
 \ee
@@
 \be\label{eq:Fresnelsign}
-	\calF^{\epsilon}_{\tau}[F](t_{0}) \equiv
+	\calF_{\tau, \epsilon}[F](t_{0}) \equiv
 \begin{cases}
@@
 \be\label{eq:resultFresnel}
-	\calT_{\rm phase}(f) = \calF^{\epsilon}_{\Tf}[F](\tf) \,.
+	\calT_{\rm phase}(f) = \calF_{\Tf, \epsilon}[F](\tf) \,.
 \ee
@@
 
-Thus, the previous result~\eqref{eq:stencilresult}-\eqref{eq:stencilfresnel} can be rephrased as a quadrature rule, and the stencil $\calF_{T}^{N,\epsilon}$ is an quadrature approximation of the Fresnel transfrom $\calF^{\epsilon}_{T}$.
+Thus, the previous result~\eqref{eq:stencilresult}-\eqref{eq:stencilfresnel} can be rephrased as a quadrature rule, and the stencil $\calF_{T, \epsilon}^{N}$ is an quadrature approximation of the Fresnel transfrom $\calF_{T, \epsilon}$.
 If one allows for polynomial integrands\footnote{Note that such integrals with a polynomial integrand are formally divergent.
@@
 \be\label{eq:transferfinal}
-	\calT(f) = \sum\limits_{k \geq 0} \frac{(-i)^{k}}{k!} (T_{Ak})^{k} \calF^{\epsilon}_{\Tf} \left[ \frac{\ud^{k}}{\ud \tau^{k}} \left( \frac{F(\tau - d(\tau))}{1+\dot{d}(\tau)} e^{-2i\pi f d(\tau)(1-\dot{d}(\tau))} \right) \right] (\tf) \,.
+	\calT(f) = \sum\limits_{k \geq 0} \frac{(-i)^{k}}{k!} (T_{Ak})^{k} \calF_{\Tf, \epsilon} \left[ \frac{\ud^{k}}{\ud \tau^{k}} \left( \frac{F(\tau - d(\tau))}{1+\dot{d}(\tau)} e^{-2i\pi f d(\tau)(1-\dot{d}(\tau))} \right) \right] (\tf) \,.
 \ee
@@
 
-In this result, the Fresnel transform $\calF^{\epsilon}_{\Tf}$ can in turn be approximated by the stencil $\calF^{N,\epsilon}_{\Tf}$ using the formula~\eqref{eq:stencilfresnel}.
+In this result, the Fresnel transform $\calF^{\epsilon}_{\Tf}$ can in turn be approximated by the stencil $\calF^{N}_{\Tf, \epsilon}$ using the formula~\eqref{eq:stencilfresnel}.
 The timescales $\Tf$ and $T_{Ak}$ were defined in~\eqref{eq:defTf} and~\eqref{eq:defTA}, and the generalized time-to-frequency function $t_{f}$ was given in~\eqref{eq:deftf}.
@@
 \begin{align}
-	\{N,A:0,d:0\}&: \; \calT(f) = \calF^{\epsilon}_{\Tf} \left[ F e^{-2i\pi f d} \right] (t_{f}) \,, \\
-	\{N,A:1,d:0\}&: \; \calT(f) = \calF^{\epsilon}_{\Tf} \left[ \left(F - i T_{A1} \left( \dot{F} - 2i\pi f \dot{d}\right) \right) e^{-2i\pi f d} \right] (t_{f}) \,, \\
-	\{N,A:0,d:1\}&: \; \calT(f) = \calF^{\epsilon}_{\Tf} \left[ \frac{F - d\dot{F}}{1+\dot{d}} e^{-2i\pi f d (1-\dot{d})} \right] (t_{f})\,, \\
-	\{N,A:1,d:1\}&: \; \calT(f) = \calF^{\epsilon}_{\Tf} \left[ \frac{1}{1+\dot{d}}\left(  F - d\dot{F} - i T_{A1} \left( \dot{F} - 2i\pi f \dot{d} \right) \right) e^{-2i\pi f d (1 - \dot{d})} \right] (t_{f})\,.
+	\{N,A:0,d:0\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ F e^{-2i\pi f d} \right] (t_{f}) \,, \\
+	\{N,A:1,d:0\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ \left(F - i T_{A1} \left( \dot{F} - 2i\pi f \dot{d}\right) \right) e^{-2i\pi f d} \right] (t_{f}) \,, \\
+	\{N,A:0,d:1\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ \frac{F - d\dot{F}}{1+\dot{d}} e^{-2i\pi f d (1-\dot{d})} \right] (t_{f})\,, \\
+	\{N,A:1,d:1\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ \frac{1}{1+\dot{d}}\left(  F - d\dot{F} - i T_{A1} \left( \dot{F} - 2i\pi f \dot{d} \right) \right) e^{-2i\pi f d (1 - \dot{d})} \right] (t_{f})\,.
 \end{align}
@@
 \begin{align}
-	\{N,A:0\}&: \; \calT(f) = \calF^{\epsilon}_{\Tf} \left[ F \right] (t_{f}) \,, \\
-	\{N,A:1\}&: \; \calT(f) = \calF^{\epsilon}_{\Tf} \left[ \left(F - i T_{A1} \dot{F} \right) \right] (t_{f}) \,, \\
-	\{N,A:2\}&: \; \calT(f) = \calF^{\epsilon}_{\Tf} \left[ \left(F - i T_{A1} \dot{F} - \frac{1}{2} (T_{A2})^{2} \ddot{F} \right) \right] (t_{f}) \,.
+	\{N,A:0\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ F \right] (t_{f}) \,, \\
+	\{N,A:1\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ \left(F - i T_{A1} \dot{F} \right) \right] (t_{f}) \,, \\
+	\{N,A:2\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ \left(F - i T_{A1} \dot{F} - \frac{1}{2} (T_{A2})^{2} \ddot{F} \right) \right] (t_{f}) \,.
 \end{align}
@@
 \end{subequations}
-where the function $G$ and its derivatives are evaluated at $(f, t_{f})$. Reaching $\epsilon \simeq 1$ will indicate a breakdown of the perturbative approach.
+where the function $G$ and its derivatives are evaluated at $(f, t_{f})$. Reaching $\epsilon \simeq 1$ will indicate a breakdown of the perturbative approach. \SM{[expand a bit]}
 
@@
 Higher-order amplitude terms beyond the first one in~\eqref{eq:resulttaylorA} will turn out to be completely negligible in the LISA case and we will ignore them in the following.
-We will present these estimates separately for the orbital response~\eqref{eq:transferOlocal} and the constellation~\eqref{eq:transferclocal}, as the different baseline of the delays ($R$ or $L$) as well as the presence of a time-varying prefactor $F(t)$ both affect the result.
+We will present these estimates separately for the orbital response~\eqref{eq:transferOlocal} and the constellation~\eqref{eq:transferclocal}, as the different baseline of the delays (orbital radius $R$ or armlength $L$) as well as the presence of a time-varying prefactor $F(t)$ both affect the result.
 As we will see, when estimating the magnitude of the relevant derivatives of $G$ we have to take into account delay factors of the type $2\pi f d$.
@@
 
-\subsection{Direct approach for periodic modulations and delays}\label{subsec:comblisa}
-
-In the LISA case, the modulations and delays entering~\eqref{eq:yslr} are periodic, with a period of $f_{0} = 1/\mathrm{yr} \simeq 3.2\times 10^{-8}\mathrm{Hz}$.
-If the above perturbative formalism cannot be applied, we can still exploit this periodic structure .
+\subsection{Slowly chirping binaries and Direct approach for periodic modulations and delays}\label{subsec:comblisa}
+
+In the LISA case, the modulations and delays entering~\eqref{eq:yslr} are periodic, with a period of one year and a frequency $f_{0} = 1/\mathrm{yr} \simeq 3.2\times 10^{-8}\mathrm{Hz}$.
+If the above perturbative formalism cannot be applied, we can still exploit this periodic structure to transform the .
 \SM{[]}

### 2017-12-28 docs/paper_method.tex
Edits to the paper.

@@
 \newcommand{\Hz}{\,\mathrm{Hz}}
+\newcommand{\yr}{\,\mathrm{yr}}
 \newcommand{\sinc}{\,\mathrm{sinc}}
@@
 
-We use a coordinate system centered on the solar system barycenter (SSB), and we denote the center of the constellation by $O$.
+We use a coordinate system centered on the solar system barycenter (SSB), and represent the center of the constellation by the vector $p_{0}$.
 We introduce the notation $\hatk$ for the propagation vector of the gravitational wave, which we denote by $h_{ij}^{\rm TT}(t)$ in transverse-traceless matrix form, as measured at the SSB (thus, at position $p$, $h_{ij}^{\rm TT}(t, p) = h_{ij}^{\rm TT}(t - \hatk \cdot p)$).
-We denote by $p_{A}$ ($A=1,2,3$) the position of the spacecrafts and $n_{l}$ the unit vectors of the three links, with the convention that $n_{3}$ points from 1 to 2.
+We denote by $p_{A}$ ($A=1,2,3$) the position of the individual spacecrafts and $n_{l}$ the unit vectors of the three links, with the convention that $n_{3}$ points from 1 to 2.
 
@@
 \end{align}
-As detailed below in Sec.~\ref{}, we will use a rigid instantaneous model, approximating the geometry of the constellation by a moving equilateral triangle (neglecting the flexing of the arms induced by corrections in the orbits), and evaluating all geometric factors in~\eqref{eq:yslr} at a single time $t$ (neglecting pointing-ahead corrections).
+We will use a rigid instantaneous model, approximating the geometry of the constellation by a moving equilateral triangle (neglecting the flexing of the arms induced by corrections in the orbits), and evaluating all geometric factors in~\eqref{eq:yslr} at a single time $t$ (neglecting pointing-ahead corrections).
 The additional delay $L$ (we use $c=1$) in the first term represents the light propagation time along the arm between spacecrafts $s$ and $l$.
@@
 
-The functions $F(t)$ and $d(t)$ vary on a timescale of one year.
+The functions $F(t)$ and $d(t)$ vary on a timescale of one year, with a frequency $f_{0} = 1/\mathrm{yr} \simeq 3.169\times10^{-8} \mathrm{Hz}$ (we will also use $\Omega_{0} = 2\pi f_{0}$).
 When neglecting small non-periodic orbital perturbations (such as those caused by the influence of the Earth and of the other planets on the orbits), both $F(t)$ and $d(t)$ are periodic.
-It will also be useful to separate the delays in two types of terms: the first, $d_{0} = -\hatk\cdot p_{0}$, relates the waveform at the SSB to the waveform at the center of the constellation $O$, whereas the second, $d_{c}$, represents the various delays between the spacecrafts of the constellation.
+It will also be useful to separate the delays in two types of terms: the first, $d_{0} = -\hatk\cdot p_{0}$, relates the waveform at the SSB to the waveform at the center of the constellation, whereas the second, $d_{c}$, represents the various delays between the spacecrafts of the constellation.
 For an orbit radius $R=1\,\mathrm{au}$ and an armlength $L=2.5\times10^{6}\mathrm{km}$~\cite{LISA17}, $d_{0} \sim R/c \simeq 500\mathrm{s}$ and $d_{c} \sim L/c \simeq 8\mathrm{s}$.
@@
 \end{subequations}
-where the function $G$ and its derivatives are evaluated at $(f, t_{f})$. Reaching $\epsilon \simeq 1$ will indicate a breakdown of the perturbative approach. \SM{[expand a bit]}
+where the function $G$ and its derivatives are evaluated at $(f, t_{f})$.
+For $\epsilon \ll 1$, the perturbative approach applies and $\epsilon$ can be used as an estimate for the magnitude of the effect.
+Reaching $\epsilon \sim 1$ will indicate a breakdown of the perturbative approach.
 
@@
 
+In this Section, we apply the formalism of Sec.~\ref{sec:formalism} to the Fourier-domain response of a LISA-like detector, and assess the accuracy of our approach at various levels of approximation.
+We also propose an alternative approach for slowly-chipring signals.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In this Section, we expand Sec.~\ref{subsec:} \SM{[R]} and detail more the model that we use for the response of a LISA-like detector, together with the assumptions used and their limitations.
-
-As argued in~\cite{Krolak+04} \SM{[expand a bit; not the only place where this approximation is used]}, since we focus on the gravitational-wave contribution to the basic observable, we can use a simplified model for the response, ignoring corrections that are crucial from the point of view of noise cancellations.
-Our aim is indeed to assess the accuracy of a direct Fourier-domain treatment of the response, thus focusing on the separation of timescales in the problem, keeping in mind that corrections can be added later without affecting the conclusions of the present analysis.
+We begin by detailing the model that we use for the response of a LISA-like detector, together with the assumptions used and their limitations.
+Since our aim is to assess the accuracy of our direct Fourier-domain treatment of the response, we can focus on the gravitational-wave contribution to the basic observables.
+We therefore use a somewhat simplified model for the time-domain response, ignoring corrections that are crucial from the point of view of noise cancellations, keeping in mind that the response model can be enriched later without affecting the conclusions of the present analysis.
 
@@
 
-Additionally, we limit ourselves to a rigid model for the detector, namely we assume that the constellation remains in an equilateral configuration with fixed armlengths $L=2.5\times 10^{6}\mathrm{km}$, as proposed in~\cite{LISA17}.
+Additionally, we limit ourselves to a rigid model for the orbits of the constellation, namely we assume that the constellation remains in an equilateral configuration with fixed armlengths $L=2.5\times 10^{6}\mathrm{km}$ (the latest configuration~\cite{LISA17}).
 This simplified orbits neglect (iv) the effect of excentricity of the individual Keplerian orbits, which introduces corrections at the level $e^{2}$ (v) the effect of gravitational perturbations coming from other celestial bodies, such as the Earth, the quadrupole of the Sun, and the other planets.
@@
 Note that although we can neglect all the effects (i)-(v) for our present study, keeping track of these corrections is crucial for the purpose of laser noise cancellations, and led to the development of new generations of TDI observables~\cite{} \SM{[C]}.
-
-Since the orbital delay due to the orbit around the Sun is common to all $y_{slr}$ observables, we will decompose the one-arm instrument response~\eqref{} \SM{[R]} in two steps.
-The first step consists simply in applying the varying time delay to bring the signal from the SSB to the center of the LISA triangular constellation.
+This simplified response has been abundantly used in previous studies (see e.g.~\cite{} \SM{[C]}).
+
+It is natural to split the response~\eqref{eq:yslr} into two steps: the orbital delay related to the orbit around the Sun of the whole constellation on one side, and the constellation response on the other side.
+The baselines for the delays are indeed very different in the two cases.
+Geometrical projection factors aside, we have for the orbit around the Sun $R=1\text{au}=1.5\times 10^{8} \text{km}$, while the detector armlength is $L=2.5\times 10^{6}\text{km}$ (in the configuration proposed in~\cite{LISA17}).
+It is natural to define two transfer frequencies for the two relevant length scales for the delays, defined such that a wavelength fits within this length scale, i.e.
+$2\pi f d = 1$.
+This gives
+\begin{subequations}\label{eq:transferfrequencies}
+\begin{align}
+	f_{R} &= 3.2\times10^{-4}\Hz \,,\\
+	f_{L} &= 1.9\times 10^{-2}\Hz \,.
+\end{align}
+\end{subequations}
+The LISA response behaves qualitatively differently on the frequency bands separated by these transfer frequencies.
+
+The first step consists simply in applying the varying time delay to bring the signal from the SSB to the center of the LISA triangular constellation, common to all $y_{slr}$ observables.
 For $h^{\rm TT}$ the transverse-traceless gravitational waveform in matrix form, we write this orbital time delay as
-\be\label{eq:defresponseO}
-	h_{0}^{\rm TT} (t) = h^{\rm TT}(t-\hatk\cdot p_{0}) \,.
-\ee
-The second step consists in the remaining, constellation-centered response, which we write as
-\begin{align}\label{eq:defresponsec}
+\be\label{eq:defresponse0}
+	h_{0}^{\rm TT} (t) = h^{\rm TT}(t-\hatk\cdot p_{0}) \,,
+\ee
+with $p_{0}$ the position of the constellation center, which follows the Earth orbit around the Sun. The second step consists in the remaining, constellation-centered response, which we write as
+\begin{align}\label{eq:defresponseL}
 	y_{slr} &= \frac{1}{2} \frac{1}{1 - \hatk\cdot n_{l}} \nn\\
@@
 
-We will decompose the full signal in the contributions of the individual spin-weighted spherical modes $h_{\ell m}$, whose Fourier transforms are assumed to have a smooth amplitude and phase.
+As described in App.~\ref{app:notation} We will decompose the full signal in the contributions of the individual spin-weighted spherical modes $h_{\ell m}$, whose Fourier transforms are assumed to have a smooth amplitude and phase.
 First, we define the matrices $P_{+},P_{\times}$ such that, in the sense of matrices,
@@
 \ee
-Assuming that the approximation~\eqref{eq:zeronegativef} applies, we focus only on positive frequencies and consider a mode contribution $h$ following~\eqref{eq:notationhsignm}.
-We can define a complex matrix $P$ incorporating the spin-weighted spherical harmonic constant factor as
+We focus only on positive frequencies.
+Assuming that the approximation~\eqref{eq:zeronegativef} applies, we consider a single mode contribution, either $h=h_{\ell m}$ for $m>0$ or $h=h_{\ell m}^{*}$ for $m<0$ (following~\eqref{eq:notationhsignm}).
+For each mode we define a complex matrix $P_{\ell m}$ incorporating the spin-weighted spherical harmonic constant factor as
 \be
@@
 
-We now turn to the transposition of~\eqref{eq:defresponseO} and~\eqref{eq:dfresponsec} to the Fourier-domain.
-For the leading-order response~\eqref{eq:deflocal}, in what we call the local approximation, applying a pure delay as in~\eqref{eq:defresponseO} translates into Fourier-domain transfer function that is a pure phase factor, proportional to the frequency but also $t_{f}$-dependent: \SM{[check sign]}
-\be\label{eq:transferOlocal}
+We now turn to the transposition of~\eqref{eq:defresponse0} and~\eqref{eq:defresponseL} to the Fourier domain.
+For the leading-order response~\eqref{eq:transferlocal}, applying a pure delay as in~\eqref{eq:defresponse0} translates into Fourier-domain transfer function common to all modes, that is a pure phase factor, proportional to the frequency but also $t_{f}$-dependent:
+\be\label{eq:transfer0local}
 	\calT_{0}^{\rm local}(f) = e^{-2i\pi f d_{0}(t_{f})}\,,
 \ee
-with $d_{0} = -\hatk \cdot p_{0}$ the orbital delay.
-When using the leading-order response~\eqref{eq:transferlocal} for the constellation response, the delays take similarly the form of pure phase factors.
-When also using the rigid approximation for the constellation, where the armlengths are fixed, a particular simplification occurs when combining these individual delays.
-The Fourier-domain transfer function then reads
-\begin{align}\label{eq:transferclocal}
-	\calT_{slr}^{\rm local} &= \frac{i \pi f L}{2} \sinc \left[ \pi f L\left(1-\hatk\cdot n_{l} \right) \right] \nn\\
-	& \qquad \cdot \exp\left[ i \pi f \left( L + \hatk\cdot \left( p_{1}^{L} + p_{2}^{L} \right) \right) \right]  n_{l} \cdot P_{\ell m} \cdot n_{l} \,,
-\end{align}
-where all vectors $n_{l}$ and $p_{A}^{L}$ are to be evaluated at time $t_{f}$.
-We used the rigid-arms approximation relation $p^{L}_{r} - p^{L}_{s} =  L n_{l}$.
-\SM{[rewrite this]} The result~\eqref{eq:transferlocal} is familiar, as it was derived in the past taking the point of view of monochromatic signals (see e.g.~\cite{} \SM{[C]}).
-
-In the case $m<0$, the formalism can be applied in the same way, except that notation is complicated by the fact that the mode $\tilde{h}_{\ell m}$ has to be evaluated at $-f$.
-In the following, we will limit ourselves to the case $m>0$ for simplicity of presentation.
-\SM{[really necessary ?]}
-
-Note that, if the corrections~\eqref{} are included for the constellation delays, this transfer function will not have this simple form anymore, as $\dot{d}$ will have a different velocity-dependent expression for the sending and receiving spacecrafts.
-
-It is natural to define two transfer frequencies for the two relevant length scales for the delays, defined such that a wavelength fits within this length scale.
-Geometrical projection factors aside, these scales are the radius of the orbit around the Sun $R=1\text{au}=1.5\times 10^{8} \text{km}$ and the detector armlength $L=2.5\times 10^{6}\text{km}$ (in the configuration proposed in~\cite{LISA17}).
-This gives
-\begin{subequations}\label{eq:transferfrequencies}
-\begin{align}
-	f_{R} &= 3.2\times10^{-4}\text{Hz} \,\\
-	f_{L} &= 1.9\times 10^{-2}\text{Hz} \,.
-\end{align}
-\end{subequations}
-These transfer frequencies will indicate where in the LISA band a dimensionless delay term of the form $2\pi f d$ becomes larger than unity. \SM{[check if this appears elsewhere]}
-
-\SM{[comment on the added structure in the high-f response - say a word of f-factor corresponding to discrete derivatives - define low-f response that others have been using - reference to folks also looking at that (marginal) for ET?
-]}
+with $d_{0} = -\hatk \cdot p_{0}$ the delay associated to the orbit around the Sun.
+If $(\lambda, \beta)$ are the ecliptic longitude and latitude of the source in the sky, and if the orbital phase is set to $0$ at $t=0$, this delay has the simple expression
+\be\label{eq:delay0}
+	d_{0}(t) = -R \cos\beta \cos\left(\Omega_{0}t - \lambda\right)\,.
+\ee
+
+For the constellation response~\eqref{eq:defresponseL}, we can separate the two delays by writing
+\begin{align}\label{eq:decomposeGslr}
+	F_{slr}^{L}(t) &= \frac{1}{2} \frac{1}{1 - \hatk\cdot n_{l}(t)} n_{l}(t) \cdot P_{\ell m} \cdot n_{l} (t) \,,\nn\\
+	d_{s,L}(t) &= -L - k\cdot p_{s}^{L}(t) \,, \quad d_{r}(t) = - k\cdot p_{r}^{L}(t) \,,\nn\\
+	G_{slr}^{L}(f,t) &=  F_{slr}^{L}(t) \left( e^{-2i\pi f d_{s,L}(t)} - e^{-2i\pi f d_{r}(t)} \right) \,.
+\end{align}
+When also using the rigid approximation for the constellation, where the armlengths are fixed, a particular simplification occurs when combining these individual delays, thanks to the relation $p^{L}_{r} - p^{L}_{s} =  L n_{l}$.
+In the local approximation~\eqref{eq:transferlocal}, the Fourier-domain transfer function then reads
+\begin{align}\label{eq:transferLlocal}
+	\calT_{slr}^{L, \mathrm{local}} &= \frac{i \pi f L}{2} \sinc \left[ \pi f L\left(1-\hatk\cdot n_{l} \right) \right] \nn\\
+	& \quad \cdot \exp\left[ i \pi f \left( L + \hatk\cdot \left( p_{1}^{L} + p_{2}^{L} \right) \right) \right]  n_{l} \cdot P_{\ell m} \cdot n_{l} \,,
+\end{align}
+where all vectors $n_{l}$ and $p_{A}^{L}$ are to be evaluated at the time $t_{f}$.
+This result~\eqref{eq:transferlocal} is familiar, as it was derived in the past taking the point of view of monochromatic signals (see e.g.~\cite{} \SM{[C]}).
+
+Note that, if the corrections of Sec.~\ref{subsec:delays} are included for the constellation delays, the transfer function will not have this simple form anymore, as $\dot{d}$ will have a different velocity-dependent expression for the sending and receiving spacecrafts.
+One must then handle separately $d_{s,L}$ and $d_{r}$ in~\eqref{eq:decomposeGslr}.
+
+The response~\eqref{eq:transfer0local}-\eqref{eq:delay0} takes a simple analytic form, but the phase contribution of this delay is significant on most of the frequency band and can be large for $f \gg f_{R}$.
+
+The response~\eqref{eq:transferLlocal} can be interpreted as the Fourier-domain translation of a discrete derivative taken on the waveform.
+The factor shows that the amplitude of the response is $\propto f$ in the low-frequency limit $f\ll f_{L}$, where the other factors are essentially unity.
+For $f\gtrsim f_{L}$, the $\sinc$ and the phase of the exponential give raise to additional structure in the response, including zero-crossings when the projected armlength corresponds to a wavelength.
+From~\eqref{eq:transferLlocal}, an expansion for small $f\ll f_{L}$ yields back a Fourier-domain analog of the low-frequency approximation of the response~\cite{Cutler97, RCP04}, which amounts to having two LIGO-type interferometers turned by $\pi/4$ and set in motion.
 
@@
 
-\SM{[add order-of-magnitude estimate for ET]}
-
-\SM{[check if there is something to be done with recent paper on ET modulations]}
-
-\SM{[add order-of-magnitude estimate for EMRIS]}
-
-\SM{[possibly add order-of-magnitude estimate for geocentric-LISA ?]}
-
-\SM{[justify somewhere smallness of cubic-in-phase term]}
-
-\SM{[crude estimate of the target accuracy level: 1/SNR ?]}
-
-\SM{[distinguish orbital from constellation at the level of orders of magnitude estimates: factor 2pif*d when taking derivatives of the transfer function for the pure delay]}
-
-\SM{[change notations from O to 0 for orbital response, and check we use notations $\Omega_{0}$, $f_{0}$]}
-
-Due to the smallness of the annual frequency $\Omega_{0}/2\pi \simeq 3.1\times10^{-8}\mathrm{Hz}$ compared to the timescales relevant for the signal over the LISA frequency band, one can expect that the higher-order corrections described in Section~\ref{sec:formalism} are small.
-In this Section we try to quantify to which extent, and give simple estimates of their magnitude.
-
-Using the Newtonian-order expressions~\eqref{eq:timescalesN} for the signal-dependent timescales $\Tf$ and $T_{A1}$, we can obtain a useful order-of-magnitude estimate for the error measures $\epsilon$ introduced in~\eqref{eq:deffom}.
-Higher-order amplitude terms beyond the first one in~\eqref{eq:resulttaylorA} will turn out to be completely negligible in the LISA case and we will ignore them in the following.
-We will present these estimates separately for the orbital response~\eqref{eq:transferOlocal} and the constellation~\eqref{eq:transferclocal}, as the different baseline of the delays (orbital radius $R$ or armlength $L$) as well as the presence of a time-varying prefactor $F(t)$ both affect the result.
-As we will see, when estimating the magnitude of the relevant derivatives of $G$ we have to take into account delay factors of the type $2\pi f d$.
+We will now use the approximate error measures $\epsilon$ introduced in~\eqref{eq:deffom} to estimate, for each type of correction (phase, amplitude, delay), the level of errors in the transfer function induced by ignoring this correction.
+
+As we will see, the smallness of the annual frequency $f_{0} \simeq 3.1\times10^{-8}\mathrm{Hz}$, compared to the timescales relevant for the signal, will mean that the separation of timescales is very good in the LISA case.
+To obtain an order-of-magnitude estimate for the error measures $\epsilon$~\eqref{eq:deffom}, we will use the Newtonian-order expressions~\eqref{eq:timescalesN} for the signal-dependent timescales $\Tf$ and $T_{A1}$.
+Higher-order amplitude terms beyond the first one in~\eqref{eq:resulttaylorA}, as well as phase terms beyond the second derivative in~\eqref{eq:expandPsi}, will turn out to be completely negligible in the LISA case and we will ignore them in the following.
+It is useful to separate the orbital response~\eqref{eq:transfer0local} and the constellation response~\eqref{eq:transferLlocal}, as the different baseline of the delays (orbital radius $R$ or armlength $L$) as well as the presence of a time-varying prefactor $F(t)$ both affect the result.
+When estimating the magnitude of the relevant derivatives of $G$, we have to take into account dimensionless delay factors of the type $2\pi f d$.
 
@@
 	\frac{1}{G_{0}} \partial_{tt} G_{0} &= -2i\pi f \ddot{d}_{0} - 4\pi^{2} f^{2} \dot{d}_{0}^{2} \,,\\
-	\frac{1}{G_{0}} \partial_{tf} G_{0} &= -2 i \pi \dot{d}_{0} - 4\pi^{2} f d_{0} \dot{d}_{0} \,,
+	\frac{1}{G_{0}} \partial_{tf} G_{0} &= -2 i \pi \dot{d}_{0} - 4\pi^{2} f d_{0} \dot{d}_{0} \,.
 \end{align}
@@
 
-For the constellation response \SM{[say here that even the one-arm response is actually the difference between two G's, so the error measure is not quite rightly used -- origin of the f factor in front]}, since we wish to write errors in relative sense starting from the single-arm response~\eqref{eq:transferclocal} which is already analogous to a discrete time derivative of the signal, we keep explicit the overall factor $f$ reflecting this derivative.
-\SM{[investigate later how this applies to the discrete time derivatives that define the TDI - here we are allowed to ignore this because they are with a fixed L-delay]} We will write symbolically $G_{c}(f,t) \sim f F(t) e^{-2i \pi f d_{c}(t)}$, where $d_{c}$ represents the delay term and $F(t)$ represents the rest of the geometric factors in~\eqref{eq:transferclocal} where we ignore the $f$-dependence.
-This gives in this case
+Since the one-arm constellation response~\eqref{eq:decomposeGslr}-\eqref{eq:transferLlocal} is analogous to a discrete time derivative of the signal, it is appropriate to keep explicit an overall factor $f$ reflecting this structure.
+Hence we write symbolically $G_{L}(f,t) \sim f F(t) e^{-2i \pi f d_{L}(t)}$, where $d_{L}$ represents a delay term and $F(t)$ represents the rest of the geometric factors in~\eqref{eq:transferLlocal} where we ignore the $f$-dependence.
+This gives
 \begin{subequations}
 \begin{align}
-	\frac{1}{G_{c}} \partial_{t} G_{c} &\sim -2i\pi f \dot{d}_{c} + \frac{\dot{F}}{F}\,,\\
-	\frac{1}{G_{c}} \partial_{tt} G_{c} &\sim -2i\pi f \ddot{d}_{c} - 4\pi^{2} f^{2} \dot{d}_{c}^{2} - 2i\pi f \dot{d}_{c} \frac{\dot{F}}{F} - \frac{\dot{F}^{2}}{F^{2}} + \frac{\ddot{F}}{F} \,,\\
-	\frac{1}{G_{c}} \partial_{tf} G_{c} &\sim -4 i \pi \dot{d}_{c} - 2i\pi d_{c} \frac{\dot{F}}{F} - 4\pi^{2} f d_{c} \dot{d}_{c} + \frac{1}{f}\frac{\dot{F}}{F}\,.
+	\frac{1}{G_{L}} \partial_{t} G_{L} &\sim -2i\pi f \dot{d}_{c} + \frac{\dot{F}}{F}\,,\\
+	\frac{1}{G_{L}} \partial_{tt} G_{L} &\sim -2i\pi f \ddot{d}_{c} - 4\pi^{2} f^{2} \dot{d}_{c}^{2} - 2i\pi f \dot{d}_{c} \frac{\dot{F}}{F} - \frac{\dot{F}^{2}}{F^{2}} + \frac{\ddot{F}}{F} \,,\\
+	\frac{1}{G_{L}} \partial_{tf} G_{L} &\sim -4 i \pi \dot{d}_{c} - 2i\pi d_{c} \frac{\dot{F}}{F} - 4\pi^{2} f d_{c} \dot{d}_{c} + \frac{1}{f}\frac{\dot{F}}{F}\,.
 \end{align}
@@
 In order to obtain simple scalings for the error estimates, we will make the replacements $\partial_{t}^{n} d \sim \Omega_{0}^{n} d$ as well as $\partial_{t}^{n} F \sim \Omega_{0}^{n} F$.
-These scalings are only approximate and the various possible orientation angles can lead to significant variations.
-We will assume $d_{0} \sim R_{\rm eff} $ (recalling that we set $c=1$ \SM{[check if that is consistently done throughout]}) with an effective orbit radius $R_{\rm eff} \equiv 2R/\pi$ representing an average of the geometric projection factor of the gravitational wave propagation vector on the plane of the orbit.
-For the constellation delays, we simply take $d_{c} \sim L$ as there projection effects can lead to variations in both ways.
-We build estimates for the magnitude of these derivatives as follows:
+These scalings are only approximate as different orientation angles can lead to significant variations.
+To represent the average of the geometric projection factor of the gravitational wave propagation vector on the plane of the orbit, we will assume $d_{0} \sim R_{\rm eff} \equiv 2R/\pi $.
+For the constellation delays, we simply take $d_{L} \sim L$ as there projection effects can lead to variations in both ways.
+The resulting estimates for the magnitude of these derivatives are
 \begin{subequations}\label{eq:estimatederivorb}
@@
 \begin{align}
-	\left| \frac{1}{G_{c}} \partial_{t} G_{c} \right| &\sim \text{max} \left[ \Omega_{0}, 2 \pi f \Omega_{0} L \right] \,,\\
-	\left| \frac{1}{G_{c}} \partial_{tt} G_{c} \right| &\sim \text{max} \left[ \Omega_{0}^{2}, 4 \pi f \Omega_{0}^{2} L, 4\pi^{2} f^{2} \Omega_{0}^{2} L^{2} \right] \,,\\
-	\left| \frac{1}{G_{c}} \partial_{tf} G_{c} \right| &\sim \text{max} \left[ \Omega_{0}/f, 6 \pi \Omega_{0} L, 4\pi^{2} f \Omega_{0}^{2} L^{2} \right] \,,
+	\left| \frac{1}{G_{L}} \partial_{t} G_{L} \right| &\sim \text{max} \left[ \Omega_{0}, 2 \pi f \Omega_{0} L \right] \,,\\
+	\left| \frac{1}{G_{L}} \partial_{tt} G_{L} \right| &\sim \text{max} \left[ \Omega_{0}^{2}, 4 \pi f \Omega_{0}^{2} L, 4\pi^{2} f^{2} \Omega_{0}^{2} L^{2} \right] \,,\\
+	\left| \frac{1}{G_{L}} \partial_{tf} G_{L} \right| &\sim \text{max} \left[ \Omega_{0}/f, 6 \pi \Omega_{0} L, 4\pi^{2} f \Omega_{0}^{2} L^{2} \right] \,,
 \end{align}
 \end{subequations}
-for the constellation response.
-In the presence of different terms, we simply take the maximum of their norms.
+for the constellation response. In the presence of different terms, we simply take the maximum of their norms.
+
 An important point in the above is the presence of delay factors in the forms of powers of $2\pi f d$.
-Thus, the magnitude of the corrections cannot be evaluated with the simple replacement $\partial_{t}^{n} \rightarrow \Omega_{0}^{n}$, and the frequency dependency of the error measures will depend on the frequency being above or below the transfer frequencies $f_{R}$ and $f_{L}$ defined in~\eqref{eq:transferfrequencies}.
+Thus, the magnitude of the corrections cannot be evaluated with the simple replacement $\partial_{t}^{n} \rightarrow \Omega_{0}^{n}$, and the frequency dependency of the error measures $\epsilon$ will depend on the frequency being above or below the transfer frequencies $f_{R}$ and $f_{L}$ defined in~\eqref{eq:transferfrequencies}.
 
@@
 Since the corrections turn out to be most important at the beginning of the signal, we will evaluate the timescales at this starting frequency.
-From the Newtonian relation~\eqref{eq:omegaphiN}, we have (using $\nu_{0} = 1/4$ for the equal-mass value): \SM{[check]}
-\be
-	f_{\rm start} = 2.9\times 10^{-5} \text{Hz} \left( \frac{\nu}{\nu_{0}} \right)^{-3/8} \left( \frac{m}{10^{6}M_{\odot}} \right)^{-5/8} \left( \frac{\Delta t}{10 \text{yr}} \right)^{-3/8} \,,
-\ee
-which gives in turn for the timescales
+From the Newtonian relation~\eqref{eq:omegaphiN}, we have (using $\nu_{0} = 1/4$ for the equal-mass value): \SM{[check]} \SM{[use Mchirp]}
+\be
+	f_{\rm start} = 2.9\times 10^{-5} \Hz \left( \frac{\nu}{\nu_{0}} \right)^{-3/8} \left( \frac{m}{10^{6}M_{\odot}} \right)^{-5/8} \left( \frac{\Delta t}{10 \yr} \right)^{-3/8} \,,
+\ee
+which gives in turn for the timescales \SM{[use Mchirp]}
 \begin{subequations}
 \begin{align}
-	\Tf = 6.8\times10^{-2}\text{yr} \left( \frac{\nu}{\nu_{0}} \right)^{3/16} \left( \frac{m}{10^{6}M_{\odot}} \right)^{5/16} \left( \frac{\Delta t}{10 \text{yr}} \right)^{11/16} \,\\
-	T_{A1} = 2.0\times10^{-4}\text{yr} \left( \frac{\nu}{\nu_{0}} \right)^{3/8} \left( \frac{m}{10^{6}M_{\odot}} \right)^{5/8} \left( \frac{\Delta t}{10 \text{yr}} \right)^{3/8} \,.
+	\Tf = 6.8\times10^{-2}\yr \left( \frac{\nu}{\nu_{0}} \right)^{3/16} \left( \frac{m}{10^{6}M_{\odot}} \right)^{5/16} \left( \frac{\Delta t}{10 \yr} \right)^{11/16} \,\\
+	T_{A1} = 2.0\times10^{-4}\yr \left( \frac{\nu}{\nu_{0}} \right)^{3/8} \left( \frac{m}{10^{6}M_{\odot}} \right)^{5/8} \left( \frac{\Delta t}{10 \yr} \right)^{3/8} \,.
 \end{align}
@@
 We summarize the results for the error estimates in Fig.~\ref{fig:fomLISA}.
-We show the magnitude of the error estimates $\epsilon_{\Psi}$, $\epsilon_{A1}$, $\epsilon_{d}$ both for the orbital delay and the constellation response, for an equal-mass system with total masses of $m=10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$ and a starting frequency corresponding to $\Delta t =10 \text{yr}$ of observation.
-Since individual signals can show significant variation depending on the angular parameters, we show an average of $\ln \epsilon$ over the sky position, inclination and polarization, together with $\pm \sigma$ standard deviation, and overlay the estimates resulting from~\eqref{eq:estimatederivorb},~\eqref{eq:estimatederivconst} and~\eqref{eq:timescalesN}.
-Here and in the following, for completeness we show results on the frequency band from $10^{-5}\text{Hz}$ to $1Hz$, while the signals are expected to have little signal-to-noise ratio outside of the band from $10^{-4}\text{Hz}$ to $10^{-1}Hz$, which is sometimes taken as a reference~\cite{LISA17}.
+We show the magnitude of the error estimates $\epsilon_{\Psi}$, $\epsilon_{A1}$, $\epsilon_{d}$ both for the orbital delay and the constellation response, for an equal-mass system with total masses of $m=10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$ and a starting frequency corresponding to $\Delta t =10 \yr$ of observation.
+Since individual signals can show significant variation depending on the angular parameters, we show a geometric average of $\epsilon$ over the sky position, inclination and polarization, together with $\pm \sigma$ geometric standard deviation, and overlay the estimates resulting from~\eqref{eq:estimatederivorb},~\eqref{eq:estimatederivconst} and~\eqref{eq:timescalesN}.
+Here and in the following, for completeness we show results on the frequency band from $10^{-5}\Hz$ to $1\Hz$, while the signals are expected to have little signal-to-noise ratio outside of the band from $10^{-4}\Hz$ to $10^{-1}\Hz$, which is sometimes taken as a reference~\cite{LISA17}.
 
@@
 The $\epsilon_{d}$ is signal-independent, and follows a quite different behaviour for the orbital and constellation responses.
-For the orbital response, $\epsilon_{d_{0}}$ grows for higher frequencies, reaching the $\sim 0.1$ at around $1\text{Hz}$.
-For the constellation response, $\epsilon_{d_{c}}$ is mainly important at lower frequencies, reaching $\sim 0.01$ at around $10^{-5}\text{Hz}$.
+For the orbital response, $\epsilon_{d_{0}}$ grows for higher frequencies, reaching the $\sim 0.1$ at around $1\Hz$.
+For the constellation response, $\epsilon_{d_{c}}$ is mainly important at lower frequencies, reaching $\sim 0.01$ at around $10^{-5}\Hz$.
 
@@
 For this example, we will also focus on the orbital response and on a single-arm constellation response $y_{132}$, as TDI combinations will appear (within the approximations listed in Sec.~\ref{} \SM{[R]}) as linear combinations of these basic observables.
-On one hand, we perform a numerical inverse Fourier transform (IFFT) of the Fourier-domain PhenomD waveform, process the signal through the time-domain response~\eqref{eq:defresponseO} or~\eqref{eq:defresponsec}, and perform a numerical Fourier transform (FFT).
+On one hand, we perform a numerical inverse Fourier transform (IFFT) of the Fourier-domain PhenomD waveform, process the signal through the time-domain response~\eqref{eq:defresponse0} or~\eqref{eq:defresponseL}, and perform a numerical Fourier transform (FFT).
 This gives us the target Frourier-domain waveform, considered to be exact but for the possible artifacts of numerical Fourier transforms (most notably, oscillations induced by the necessary tapering of the signal).
@@
 We show the results for the orbital response in Figs.~\ref{fig:LISAerrorM1e7orb} and~\ref{fig:LISAerrorM1e2orb} in amplitude and phase form.
-In both cases, in accordance with~\eqref{eq:transferOlocal}, the transfer function is essentially a phase, and contains many more cycles in the low-mass case in the higher frequency band.
+In both cases, in accordance with~\eqref{eq:transfer0local}, the transfer function is essentially a phase, and contains many more cycles in the low-mass case in the higher frequency band.
 We find that including higher-order corrections does reduce the reconstruction errors, but with the distincitve feature that including only one of the amplitude and delay corrections can make the approximation actually worse, while including both does improve the accuracy.
@@
 
-\subsection{Slowly chirping binaries and Direct approach for periodic modulations and delays}\label{subsec:comblisa}
+\subsection{Slowly chirping binaries and direct approach for periodic modulations and delays}\label{subsec:comblisa}
 
@@
 \ee
+
+This leads to the following generalized discrete convolution: \SM{[]}
+\be
+	\tilde{s}(f) = \sum_{n \in \mathbb{Z}} c_{n}(f - n f_{0}) \tilde{h} (f - n f_{0}) \,.
+\ee
+
+Although straightforward and generic, this writing of the Fourier-domain response function will be in general more expansive to compute than the perturbative transfer function approach of Sec.~\ref{sec:formalism}.
 
@@
 In the SUA formalism proposed in~\cite{} \SM{[C]}, one extends the SPA to incorporate the leading-order correction during the inspiral phase.
-The transfer function is then computed by evaluating the modulation on a stencil of times centered around the time-of-frequency given by the SPA, following
+The transfer function is then computed by evaluating the modulation on a stencil of times centered around the time-of-frequency given by the SPA, following \SM{[say in words:tf-tfSPA]}
 \begin{widetext}
@@
 
-Thanks to the separation of timescales between the precessional timescale and the orbital timescale, for suffciently high frequencies the convolution~\eqref{eq:convolution} will have a support that will be entirely comprised in the high-frequency part of the signal $\tilde{h}(f)$, which is rather featureless and slowly varying as a function of $f$.
+Thanks to the separation of timescales between the precessional timescale and the orbital timescale, for sufficiently high frequencies the convolution~\eqref{eq:convolution} will have a support that will be entirely comprised in the high-frequency part of the signal $\tilde{h}(f)$, which is rather featureless and slowly varying as a function of $f$.
 Taking advantage of this smoothness, we will adopt a trigonometric polynomial representation for $\tilde{h}(f)$.
@@
 To numerically compute the transfer functions, we first have to perform an IFFT of our Fourier-domain P-frame waveform, to obtain $h^{\rm P}_{22}$ as a function of time.
-To mitigate the effect of the necessary tapering of the waveform when computing this numerical inverse Fourier transform, we apply a Planck-window tapering on the range $f\in 16-20 \text{Hz}$, using a total mass of $M=20 \Msol$.
-\SM{[Update this to $f_{\rm min} = 10\text{Hz}$]}.
+To mitigate the effect of the necessary tapering of the waveform when computing this numerical inverse Fourier transform, we apply a Planck-window tapering on the range $f\in 16-20 \Hz$, using a total mass of $M=20 \Msol$.
+\SM{[Update this to $f_{\rm min} = 10\Hz$]}.
 In parallel, we generate an SEOB waveform with the appropriate length in time.
@@
 Fig.~\ref{fig:ampphase} shows an example of amplitude and phase in geometric units for an equal-mass binary.
+\SM{[remove that and point to appendix on interpolation errors]}
 
@@
 In this work we used only low order stencils with usually $N\leq 5$, and one can simply numerically invert of the system~\eqref{eq:stencilsystem}, as this operation has to be done only once.
-For future reference, we give here exact expressions, as the specific choice of samples at $\pm kT$, makes the system~\eqref{eq:stencilsystem} take the form of a Vandermonde system.
+For future reference, we give here exact expressions, as the specific choice of samples at $\pm kT$ gives to the system~\eqref{eq:stencilsystem} the form of a Vandermonde system.
 Defining the Vandermonde matrix $V(x_{0},\dots,x_{N})$ as $V_{ij} = (x_{j})^{i}$, setting $x_{j} = j^{2}$, and $b_{p} \equiv (-i\epsilon)^{p}(2p-1)!!$, the linear system~\eqref{eq:stencilsystem} becomes

### 2017-12-29 docs/paper_method.tex
Edits to the paper.
Added plot Mc-Deltat for the analytical estimate of epsilon_Psi for LISA.

@@
 \newcommand{\Msol}{M_{\odot}}
+\newcommand{\Mchirp}{M_{c}}
 \newcommand{\tf}{t_{f}}
@@
 where we set $x = (G M\omega/c^{3})^{2/3}$ and where $D$ is the luminosity distance to the observer.
-For this Newtonian inspiral, applying the SPA gives for the $22$ mode $\tilde{h}(f) = A_{\rm N}(f)e^{-i\Psi_{\rm N}(f)}$ with
+For this Newtonian inspiral, applying the SPA gives for the $22$ mode $\tilde{h}(f) = A_{\rm N}(f)e^{-i\Psi_{\rm N}(f)}$ with \SM{[check that total mass convention is $M$ everywhere]}
 \begin{subequations}\label{eq:SPAN}
@@
 where we used the notation $v=(G M \pi f/c^{3})^{1/3}$ ($x=v^{2}$ with the SPA correspondence~\eqref{eq:deftfSPA}), and where $t_{0}, \phi_{0}$ are constants.
-It is also customary to rewrite the above relations in terms of the chirp mass $\calM \equiv M\nu^{3/5}$, which is the only mass combination characterizing the signal at the leading PN order.
+It is also customary to rewrite the above relations in terms of the chirp mass $\Mchirp \equiv M\nu^{3/5}$, which is the only mass combination characterizing the signal at the leading PN order.
 
@@
 	\{N,A:0\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ F \right] (t_{f}) \,, \\
-	\{N,A:1\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ \left(F - i T_{A1} \dot{F} \right) \right] (t_{f}) \,, \\
-	\{N,A:2\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ \left(F - i T_{A1} \dot{F} - \frac{1}{2} (T_{A2})^{2} \ddot{F} \right) \right] (t_{f}) \,.
+	\{N,A:1\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ F - i T_{A1} \dot{F} \right] (t_{f}) \,, \\
+	\{N,A:2\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ F - i T_{A1} \dot{F} - \frac{1}{2} (T_{A2})^{2} \ddot{F} \right] (t_{f}) \,.
 \end{align}
@@
 
-We will now use the approximate error measures $\epsilon$ introduced in~\eqref{eq:deffom} to estimate, for each type of correction (phase, amplitude, delay), the level of errors in the transfer function induced by ignoring this correction.
-
-As we will see, the smallness of the annual frequency $f_{0} \simeq 3.1\times10^{-8}\mathrm{Hz}$, compared to the timescales relevant for the signal, will mean that the separation of timescales is very good in the LISA case.
-To obtain an order-of-magnitude estimate for the error measures $\epsilon$~\eqref{eq:deffom}, we will use the Newtonian-order expressions~\eqref{eq:timescalesN} for the signal-dependent timescales $\Tf$ and $T_{A1}$.
-Higher-order amplitude terms beyond the first one in~\eqref{eq:resulttaylorA}, as well as phase terms beyond the second derivative in~\eqref{eq:expandPsi}, will turn out to be completely negligible in the LISA case and we will ignore them in the following.
-It is useful to separate the orbital response~\eqref{eq:transfer0local} and the constellation response~\eqref{eq:transferLlocal}, as the different baseline of the delays (orbital radius $R$ or armlength $L$) as well as the presence of a time-varying prefactor $F(t)$ both affect the result.
-When estimating the magnitude of the relevant derivatives of $G$, we have to take into account dimensionless delay factors of the type $2\pi f d$.
-
-We start with the orbital response, which takes the form of a pure delay $G_{0}(f, t) = e^{-2 i \pi f d_{0}(t)}$, and obtain
-\begin{subequations}
-\begin{align}
-	\frac{1}{G_{0}} \partial_{t} G_{0} &= -2i\pi f \dot{d}_{0}\,,\\
-	\frac{1}{G_{0}} \partial_{tt} G_{0} &= -2i\pi f \ddot{d}_{0} - 4\pi^{2} f^{2} \dot{d}_{0}^{2} \,,\\
-	\frac{1}{G_{0}} \partial_{tf} G_{0} &= -2 i \pi \dot{d}_{0} - 4\pi^{2} f d_{0} \dot{d}_{0} \,.
-\end{align}
-\end{subequations}
-
-Since the one-arm constellation response~\eqref{eq:decomposeGslr}-\eqref{eq:transferLlocal} is analogous to a discrete time derivative of the signal, it is appropriate to keep explicit an overall factor $f$ reflecting this structure.
-Hence we write symbolically $G_{L}(f,t) \sim f F(t) e^{-2i \pi f d_{L}(t)}$, where $d_{L}$ represents a delay term and $F(t)$ represents the rest of the geometric factors in~\eqref{eq:transferLlocal} where we ignore the $f$-dependence.
-This gives
-\begin{subequations}
-\begin{align}
-	\frac{1}{G_{L}} \partial_{t} G_{L} &\sim -2i\pi f \dot{d}_{c} + \frac{\dot{F}}{F}\,,\\
-	\frac{1}{G_{L}} \partial_{tt} G_{L} &\sim -2i\pi f \ddot{d}_{c} - 4\pi^{2} f^{2} \dot{d}_{c}^{2} - 2i\pi f \dot{d}_{c} \frac{\dot{F}}{F} - \frac{\dot{F}^{2}}{F^{2}} + \frac{\ddot{F}}{F} \,,\\
-	\frac{1}{G_{L}} \partial_{tf} G_{L} &\sim -4 i \pi \dot{d}_{c} - 2i\pi d_{c} \frac{\dot{F}}{F} - 4\pi^{2} f d_{c} \dot{d}_{c} + \frac{1}{f}\frac{\dot{F}}{F}\,.
-\end{align}
-\end{subequations}
-
-In order to obtain simple scalings for the error estimates, we will make the replacements $\partial_{t}^{n} d \sim \Omega_{0}^{n} d$ as well as $\partial_{t}^{n} F \sim \Omega_{0}^{n} F$.
-These scalings are only approximate as different orientation angles can lead to significant variations.
-To represent the average of the geometric projection factor of the gravitational wave propagation vector on the plane of the orbit, we will assume $d_{0} \sim R_{\rm eff} \equiv 2R/\pi $.
-For the constellation delays, we simply take $d_{L} \sim L$ as there projection effects can lead to variations in both ways.
-The resulting estimates for the magnitude of these derivatives are
-\begin{subequations}\label{eq:estimatederivorb}
-\begin{align}
-	\left| \frac{1}{G_{0}} \partial_{t} G_{0} \right| &\sim 2 \pi f \Omega_{0} R_{\rm eff}\,,\\
-	\left| \frac{1}{G_{0}} \partial_{tt} G_{0} \right| &\sim \text{max} \left[ 2 \pi f \Omega_{0}^{2} R_{\rm eff} , 4\pi^{2} f^{2} \Omega_{0}^{2} R_{\rm eff}^{2}\right] \,,\\
-	\left| \frac{1}{G_{0}} \partial_{tf} G_{0} \right| &\sim \text{max} \left[ 2 \pi  \Omega_{0} R_{\rm eff}, 4 \pi^{2} f \Omega_{0} R_{\rm eff}^{2} \right] \,,
-\end{align}
-\end{subequations}
-for the orbital response and
-\begin{subequations}\label{eq:estimatederivconst}
-\begin{align}
-	\left| \frac{1}{G_{L}} \partial_{t} G_{L} \right| &\sim \text{max} \left[ \Omega_{0}, 2 \pi f \Omega_{0} L \right] \,,\\
-	\left| \frac{1}{G_{L}} \partial_{tt} G_{L} \right| &\sim \text{max} \left[ \Omega_{0}^{2}, 4 \pi f \Omega_{0}^{2} L, 4\pi^{2} f^{2} \Omega_{0}^{2} L^{2} \right] \,,\\
-	\left| \frac{1}{G_{L}} \partial_{tf} G_{L} \right| &\sim \text{max} \left[ \Omega_{0}/f, 6 \pi \Omega_{0} L, 4\pi^{2} f \Omega_{0}^{2} L^{2} \right] \,,
-\end{align}
-\end{subequations}
-for the constellation response. In the presence of different terms, we simply take the maximum of their norms.
-
-An important point in the above is the presence of delay factors in the forms of powers of $2\pi f d$.
-Thus, the magnitude of the corrections cannot be evaluated with the simple replacement $\partial_{t}^{n} \rightarrow \Omega_{0}^{n}$, and the frequency dependency of the error measures $\epsilon$ will depend on the frequency being above or below the transfer frequencies $f_{R}$ and $f_{L}$ defined in~\eqref{eq:transferfrequencies}.
-
-The error measure $\epsilon_{d}$ does not depend on signal-dependent timescales and can be directly read off the estimates above, with $\epsilon_{d} = \partial_{tf}G/(2\pi G)$.
-For the error measures $\epsilon_{\Psi}$ and $\epsilon_{A1}$, we can combine the above with the Newtonian timescales given in~\eqref{eq:timescalesN} for the inspiral phase of the signal.
-An important distinction needs to be made there, as the radiation reaction timescale $\Tf$ grows rapidly towards lower frequencies as $f^{-11/6}$.
-If one considers systems that are arbitrarily far from merging, $\Tf$ can be made arbitrarily large, leading the error measure $\epsilon_{\Psi}$ to take values arbitrarily larger than 1 where the perturbative treatment we adopted breaks down.
-This limit would describe systems observed in their deep inspiral and very slowly chirping, such as the population of galactic binaries that form one of the science targets of the LISA mission~\cite{LISA17}.
-Such signals would be very compact in Fourier domain, and are not covered by our formalism that assumes an extended Fourier spectrum.
-We will present in App.~\ref{} \SM{[R]} another version of our formalism based on discrete Fourier coefficients, that could apply better to this case \SM{[check: do we put this appendix or not ?]}.
-On the contrary, if one considers only systems that will merge during or shortly after the observation span of the mission, the starting frequency of the observations will be set by this limited lifetime of the source before merging and we can then bound $\epsilon_{\Psi}$ to be smaller than 1.
-
-In the following we will consider only these merging binaries, which include the supermassive black hole mergers that we expect to detect~\cite{LISA17}.
-The population of stellar-origin binary black holes, proposed in~\cite{Sesana16}, will have both sources rather far from merger and sources that will merge within a few years after the mission finishes.
-We will take a reference starting frequency corresponding to a time from merger $\Delta t = 10\text{yrs}$.
-Since the corrections turn out to be most important at the beginning of the signal, we will evaluate the timescales at this starting frequency.
-From the Newtonian relation~\eqref{eq:omegaphiN}, we have (using $\nu_{0} = 1/4$ for the equal-mass value): \SM{[check]} \SM{[use Mchirp]}
-\be
-	f_{\rm start} = 2.9\times 10^{-5} \Hz \left( \frac{\nu}{\nu_{0}} \right)^{-3/8} \left( \frac{m}{10^{6}M_{\odot}} \right)^{-5/8} \left( \frac{\Delta t}{10 \yr} \right)^{-3/8} \,,
-\ee
-which gives in turn for the timescales \SM{[use Mchirp]}
-\begin{subequations}
-\begin{align}
-	\Tf = 6.8\times10^{-2}\yr \left( \frac{\nu}{\nu_{0}} \right)^{3/16} \left( \frac{m}{10^{6}M_{\odot}} \right)^{5/16} \left( \frac{\Delta t}{10 \yr} \right)^{11/16} \,\\
-	T_{A1} = 2.0\times10^{-4}\yr \left( \frac{\nu}{\nu_{0}} \right)^{3/8} \left( \frac{m}{10^{6}M_{\odot}} \right)^{5/8} \left( \frac{\Delta t}{10 \yr} \right)^{3/8} \,.
-\end{align}
-\end{subequations}
-\SM{[change notation for total mass everywhere to $M$]} When inserting these expressions in~\eqref{eq:deffom}, in~\eqref{eq:estimatederivorb} and~\eqref{eq:estimatederivconst} we must also use $f_{\rm start}$ for the possible extra factors of $f$ or $f^{2}$ coming from the delay terms, which can lead to different overall scalings with mass ratio and total mass depending on where the signal starts in the band.
-
-We summarize the results for the error estimates in Fig.~\ref{fig:fomLISA}.
-We show the magnitude of the error estimates $\epsilon_{\Psi}$, $\epsilon_{A1}$, $\epsilon_{d}$ both for the orbital delay and the constellation response, for an equal-mass system with total masses of $m=10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$ and a starting frequency corresponding to $\Delta t =10 \yr$ of observation.
-Since individual signals can show significant variation depending on the angular parameters, we show a geometric average of $\epsilon$ over the sky position, inclination and polarization, together with $\pm \sigma$ geometric standard deviation, and overlay the estimates resulting from~\eqref{eq:estimatederivorb},~\eqref{eq:estimatederivconst} and~\eqref{eq:timescalesN}.
-Here and in the following, for completeness we show results on the frequency band from $10^{-5}\Hz$ to $1\Hz$, while the signals are expected to have little signal-to-noise ratio outside of the band from $10^{-4}\Hz$ to $10^{-1}\Hz$, which is sometimes taken as a reference~\cite{LISA17}.
-
-For $\epsilon_{\Psi}$, we find as expected a steep rise towards lower frequencies, which shows that considering only merging binaries with a limited time remaining before merger is crucial here.
-Due to the different occurences of delay factors, we find a different behaviour between the orbital response, where the initial value of $\epsilon_{\Psi}$ grows towards lower masses, and the constellation response, where $\epsilon_{\Psi}$ grows towards higher masses.
-In both cases, the error measure remains below $\epsilon_{\Psi} \sim 0.1$, where the corrections can be sizeable but well manageable by the perturbative formalism (as we will show in the next Section).
-Not represented in the figure is the behaviour with mass ratio, which also differs: for a total mass of $10^{2} \Msol$, the extra factors of $f_{\rm start}$ due to the delay factors lead to an increase towards larger mass ratios, while for a total mass of $10^{7} \Msol$ we have a decrease towards larger mass ratios.
-
-For $\epsilon_{A1}$, we find that the structure in the waveform close to merger does have a visible but limited impact on the error measure.
-For the orbital response, the error measure remains small for all masses.
-For the constellation response, we find that $\epsilon_{A1}$ can grow up to $\sim 0.01$ at the lower end of the frequency band for high masses.
-Since the timescale $T_{A1}^{\rm N}$ has a simple dependence $\propto 1/f$, these statements should hold for all mass ratios at least during the inspiral.
-
-The $\epsilon_{d}$ is signal-independent, and follows a quite different behaviour for the orbital and constellation responses.
-For the orbital response, $\epsilon_{d_{0}}$ grows for higher frequencies, reaching the $\sim 0.1$ at around $1\Hz$.
-For the constellation response, $\epsilon_{d_{c}}$ is mainly important at lower frequencies, reaching $\sim 0.01$ at around $10^{-5}\Hz$.
-
-Overall, we find that for merging binaries the perturbative approach presented in Sec.~\ref{} \SM{[]} will apply, with the most relevant higher-order corrections being expected to be the phase corrections close to the starting frequency of the signal and the delay correction at high frequencies for the orbital response.
+\begin{figure}
+  \centering
+  \includegraphics[width=.9\linewidth]{plots/lisafom_Psi_McDeltat.pdf}
+  \caption{\SM{[To be checked and redone]} Contour levels for the analytical estimate of the error measure $\epsilon_{\Psi}$ at the starting frequency, as a function of chirp mass $\Mchirp$ and time to coalescence $\Delta t$.
+Blue corresponds to the orbital response and red to the constellation response.
+The colored shaded areas indicate regions where $\epsilon_{\Psi} \geq 1$, where the perturbative formalism is expected to break down.
+In the gray area, $f_{\rm start}$ given in~\eqref{eq:fstartN} is lower than the lowest in-band frequency $f_{\rm min} = 10^{-5}\Hz$, so that the signal starts at $f_{\rm min}$ and $\epsilon_{\Psi}$ becomes independent of $\Delta t$.}
+  \label{fig:lisafomPsiMcDeltat}
+\end{figure}
 
@@
   \includegraphics[width=.32\linewidth]{plots/lisafom_dconst.pdf}
-  \caption{Error estimates of the approximation as defined in~\eqref{eq:deffom}, for an equal-mass, non-spinning system and for total masses of $10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$.
-The top row corresponds to the orbital delay part of the response~\eqref{eq:dfresponseO}, and the lower row shows the LISA-centered constellation response~\eqref{eq:dfresponsec}.
-The error estimates~\eqref{} \SM{[R]} are shown for the phase, amplitude and delay terms from left to right.
-The central line and interval are the mean and $1\sigma$ standard deviation of the logarithm of $\epsilon$ over 400 random values for the position in the sky, inclination and polarization.
+  \caption{Error estimates as defined in~\eqref{eq:deffom}, for an equal-mass, non-spinning system and for total masses $M=10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$.
+The top row corresponds to the orbital delay part of the response~\eqref{eq:defresponse0}, and the lower row shows the LISA-centered constellation response~\eqref{eq:defresponseL}.
+The error measures $\epsilon_{\Psi}$ for the phase corrections, $\epsilon_{A1}$ for the amplitude, $\epsilon_{d}$ for the delay are shown from left to right.
+The central line and interval are the mean and $1\sigma$ standard deviation of the logarithm of $\epsilon$ computed with numerical derivatives over 400 random values for the position in the sky, inclination and polarization.
 The starting frequency is set by an observation time of $\Delta t = 10 \text{yrs}$ before merger.
-\SM{[change middle mass to $3\times 10^{4}$ ?
-- clarify mismatch between these $\epsilon$ and the actual size of corrections]}}
+We overlay in dashed the analytical estimates obtained from from~\eqref{eq:timescalesNfstart} and~\eqref{eq:estimatederivorb}-\eqref{eq:estimatederivconst}.}
   \label{fig:fomLISA}
 \end{figure*}
+
+We will now use the approximate error measures $\epsilon$ introduced in~\eqref{eq:deffom} to estimate, for each type of correction (phase, amplitude, delay), the level of errors in the transfer function induced by ignoring this correction.
+
+To obtain an order-of-magnitude estimate for the error measures $\epsilon$~\eqref{eq:deffom}, we will use the Newtonian-order expressions~\eqref{eq:timescalesN} for the signal-dependent timescales $\Tf$ and $T_{A1}$.
+Higher-order amplitude terms beyond the first one in~\eqref{eq:resulttaylorA}, as well as phase terms beyond the second derivative in~\eqref{eq:expandPsi}, will turn out to be completely negligible in the LISA case and we will ignore them in the following.
+It is useful to separate the orbital response~\eqref{eq:transfer0local} and the constellation response~\eqref{eq:transferLlocal}, as the different baseline of the delays (orbital radius $R$ or armlength $L$) as well as the presence of a time-varying prefactor $F(t)$ both affect the result.
+When estimating the magnitude of the relevant derivatives of $G$, we have to take into account dimensionless delay factors of the type $2\pi f d$.
+
+We start with the orbital response, which takes the form of a pure delay $G_{0}(f, t) = e^{-2 i \pi f d_{0}(t)}$, and obtain
+\begin{subequations}
+\begin{align}
+	\frac{1}{G_{0}} \partial_{t} G_{0} &= -2i\pi f \dot{d}_{0}\,,\\
+	\frac{1}{G_{0}} \partial_{tt} G_{0} &= -2i\pi f \ddot{d}_{0} - 4\pi^{2} f^{2} \dot{d}_{0}^{2} \,,\\
+	\frac{1}{G_{0}} \partial_{tf} G_{0} &= -2 i \pi \dot{d}_{0} - 4\pi^{2} f d_{0} \dot{d}_{0} \,.
+\end{align}
+\end{subequations}
+
+Since the one-arm constellation response~\eqref{eq:decomposeGslr}-\eqref{eq:transferLlocal} is analogous to a discrete time derivative of the signal, it is appropriate to keep explicit an overall factor $f$ reflecting this structure.
+Hence we write symbolically $G_{L}(f,t) \sim f F(t) e^{-2i \pi f d_{L}(t)}$, where $d_{L}$ represents a delay term and $F(t)$ represents the rest of the geometric factors in~\eqref{eq:transferLlocal} where we ignore the $f$-dependence.
+This gives
+\begin{subequations}
+\begin{align}
+	\frac{1}{G_{L}} \partial_{t} G_{L} &\sim -2i\pi f \dot{d}_{c} + \frac{\dot{F}}{F}\,,\\
+	\frac{1}{G_{L}} \partial_{tt} G_{L} &\sim -2i\pi f \ddot{d}_{c} - 4\pi^{2} f^{2} \dot{d}_{c}^{2} - 2i\pi f \dot{d}_{c} \frac{\dot{F}}{F} - \frac{\dot{F}^{2}}{F^{2}} + \frac{\ddot{F}}{F} \,,\\
+	\frac{1}{G_{L}} \partial_{tf} G_{L} &\sim -4 i \pi \dot{d}_{c} - 2i\pi d_{c} \frac{\dot{F}}{F} - 4\pi^{2} f d_{c} \dot{d}_{c} + \frac{1}{f}\frac{\dot{F}}{F}\,.
+\end{align}
+\end{subequations}
+
+In order to obtain simple scalings for the error estimates, we will make the replacements $\partial_{t}^{n} d \sim \Omega_{0}^{n} d$ as well as $\partial_{t}^{n} F \sim \Omega_{0}^{n} F$.
+These scalings are only approximate as different orientation angles can lead to significant variations.
+To represent the average of the geometric projection factor of the gravitational wave propagation vector on the plane of the orbit, we will assume $d_{0} \sim R_{\rm eff} \equiv 2R/\pi $.
+For the constellation delays, we simply take $d_{L} \sim L$ as there projection effects can lead to variations in both ways.
+The resulting estimates for the magnitude of these derivatives are
+\begin{subequations}\label{eq:estimatederivorb}
+\begin{align}
+	\left| \frac{1}{G_{0}} \partial_{t} G_{0} \right| &\sim 2 \pi f \Omega_{0} R_{\rm eff}\,,\\
+	\left| \frac{1}{G_{0}} \partial_{tt} G_{0} \right| &\sim \text{max} \left[ 2 \pi f \Omega_{0}^{2} R_{\rm eff} , 4\pi^{2} f^{2} \Omega_{0}^{2} R_{\rm eff}^{2}\right] \,,\\
+	\left| \frac{1}{G_{0}} \partial_{tf} G_{0} \right| &\sim \text{max} \left[ 2 \pi  \Omega_{0} R_{\rm eff}, 4 \pi^{2} f \Omega_{0} R_{\rm eff}^{2} \right] \,,
+\end{align}
+\end{subequations}
+for the orbital response and
+\begin{subequations}\label{eq:estimatederivconst}
+\begin{align}
+	\left| \frac{1}{G_{L}} \partial_{t} G_{L} \right| &\sim \text{max} \left[ \Omega_{0}, 2 \pi f \Omega_{0} L \right] \,,\\
+	\left| \frac{1}{G_{L}} \partial_{tt} G_{L} \right| &\sim \text{max} \left[ \Omega_{0}^{2}, 4 \pi f \Omega_{0}^{2} L, 4\pi^{2} f^{2} \Omega_{0}^{2} L^{2} \right] \,,\\
+	\left| \frac{1}{G_{L}} \partial_{tf} G_{L} \right| &\sim \text{max} \left[ \Omega_{0}/f, 6 \pi \Omega_{0} L, 4\pi^{2} f \Omega_{0}^{2} L^{2} \right] \,,
+\end{align}
+\end{subequations}
+for the constellation response. In the presence of different terms, we simply take the maximum of their norms.
+
+An important point in the above is the presence of delay factors in the forms of powers of $2\pi f d$.
+Thus, the magnitude of the corrections cannot be evaluated with the simple replacement $\partial_{t}^{n} \rightarrow \Omega_{0}^{n}$, and the frequency dependency of the error measures $\epsilon$ will depend on the frequency being above or below the transfer frequencies $f_{R}$ and $f_{L}$ defined in~\eqref{eq:transferfrequencies}.
+The error measure $\epsilon_{d}$ does not depend on signal-dependent timescales and can be directly read off the estimates above, with $\epsilon_{d} = |\partial_{tf}G/(2\pi G)|$.
+For the error measures $\epsilon_{\Psi}$ and $\epsilon_{A1}$, we can combine the above derivatives with the Newtonian timescales given in~\eqref{eq:timescalesN} for the inspiral phase of the signal.
+
+The starting frequency of the signal will play an important role.
+As a function of the time remaining before merger $\Delta t$, from the Newtonian relation~\eqref{eq:omegaphiN} we have (using the chirp mass $\Mchirp = M \nu^{3/5}$):
+\be\label{eq:fstartN}
+	f_{\rm start} = 1.75\times 10^{-5} \Hz \left( \frac{\Mchirp}{10^{6}M_{\odot}} \right)^{-5/8} \left( \frac{\Delta t}{10 \yr} \right)^{-3/8} \,.
+\ee
+This gives in turn for the Newtonian timescales~\eqref{eq:timescalesN}
+\begin{subequations}\label{eq:timescalesNfstart}
+\begin{align}
+	\Tf^{\rm N} &= 8.78\times10^{-2}\yr \left( \frac{\Mchirp}{10^{6}M_{\odot}} \right)^{5/16} \left( \frac{\Delta t}{10 \yr} \right)^{11/16} \,, \\
+	T_{A1}^{\rm N} &= 3.37\times10^{-4}\yr \left( \frac{\Mchirp}{10^{6}M_{\odot}} \right)^{5/8} \left( \frac{\Delta t}{10 \yr} \right)^{3/8} \,.
+\end{align}
+\end{subequations}
+Note that if we consider that the frequency band of the instrument starts at a minimal frequency $f_{\rm min}$, for sufficiently high $\Mchirp$ this entry in the frequency band will mark the beginning of the signal, instead of $f_{\rm start}$ which becomes irrelevant.
+For $f_{\rm min} = 10^{-5}\Hz$ and $\Delta t \leq 10\yr$, this is the case for $\Mchirp \geq 2.45\times 10^{6}\Msol $.
+Thus, our final analytical estimates for the $\epsilon$ error measures are built by inserting~\eqref{eq:timescalesNfstart} and~\eqref{eq:estimatederivorb}-\eqref{eq:estimatederivconst} in~\eqref{eq:deffom}, while ensuring the frequency cut $10^{-5}\Hz \leq f \leq 1\Hz$.
+
+An important distinction needs to be made between what we will call merging binaries, systems close to coalescence that will merge during the LISA mission lifetime or a few years later, and slowly-chirping binaries still in the deep inspiral phase, of which we observe only a small snapshot in Fourier domain as they do not sweep to the end of the frequency band.
+The massive black hole binaries (MBH) that will be observed by LISA~\cite{LISA17} fall within the first category, with their merger in band, while the proposed population of stellar-origin black hole binaries (SOBH)~\cite{Sesana16}, with masses comparable to the LIGO/Virgo detections, will comprise both merging binaries, i.e.
+exiting the LISA band towards larger frequencies during observations, and slowly-chirping binaries hundereds or thousands of years away from merger.
+Galactic binaries (GB) fall entirely within the second category, and are very close to monochromatic over the lifetime of LISA~\cite{LISA17} \SM{[C: complete]}.
+
+We turn first to the slowly-chirping binaries.
+For these systems, we will focus on $\epsilon_{\Psi}$, which will be the most important error measure.
+Fig.~\ref{fig:lisafomPsiMcDeltat} shows contour levels for the value of our simple analytical estimate for $\epsilon_{\Psi}$ at the beginning of the observations, as a function of the chirp mass and time to merger, for both the orbital and the constellation response.
+The limit $\epsilon_{\Psi} = 1$ is used to single out areas where the perturbative treatment of Sec.~\ref{sec:formalism} is expected to break down, although the precise location of this boundary will vary, depending on the accuracy level required and on the orientation angles.
+We find that, in the LIGO/Virgo mass range $10^{1}-10^{2} \Msol$, we reach the limit $\epsilon_{\Psi}=1$ first for the orbital response, for time-to-merger values that we expect to observe~\cite{Sesana16}.
+We will introduce in Sec~\ref{subsec:comblisa} an alternative approach to deal with those signals.
+For completeness, we also show the result for intermediate and massive black hole systems, although we expect to observe mainly merging binaries for those masses.
+
+We now turn to the case of merging binaries, computing this time the three error measures $\epsilon_{\Psi}$, $\epsilon_{A1}$ and $\epsilon_{d}$.
+To go beyond the crude analytical estimates built from~\eqref{eq:timescalesNfstart} and~\eqref{eq:estimatederivorb}-\eqref{eq:estimatederivconst}, we use numerical derivatives for the timescales~\eqref{eq:defTf}-\eqref{eq:defTA} and for the derivatives in~\eqref{eq:deffom}.
+We summarize the results in Fig.~\ref{fig:fomLISA}, both for the orbital delay and the constellation response, for an equal-mass system with total masses of $M=10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$ and a starting frequency corresponding to $\Delta t =10 \yr$ of observation.
+Since individual signals can show significant variation depending on the angular parameters, for the full computation with numerical derivatives we show a geometric average of $\epsilon$ over the sky position, inclination and polarization, together with $\pm 1\sigma$ geometric standard deviation.
+Here and in the following, for completeness we show results on the frequency band from $10^{-5}\Hz$ to $1\Hz$, while the signals are expected to have little signal-to-noise ratio outside of the band from $10^{-4}\Hz$ to $10^{-1}\Hz$, which is sometimes taken as a reference~\cite{LISA17}.
+
+For $\epsilon_{\Psi}$, we find as expected from~\eqref{eq:TfN} a steep rise towards lower frequencies, which shows that considering merging binaries with a limited time to coalescence is crucial here.
+Due to the different occurences of delay factors, we find a different behaviour between the orbital response, where for a fixed mass ratio the initial value of $\epsilon_{\Psi}$ grows towards lower masses, and the constellation response, where $\epsilon_{\Psi}$ grows towards higher masses.
+In both cases, the error measure remains below $\epsilon_{\Psi} \sim 0.1$, where the corrections are still well manageable by the perturbative formalism (as we will show in Sec.~\ref{subsec:errorsLISA}).
+
+For $\epsilon_{A1}$, we find that the structure in the waveform close to merger does have a visible but limited impact on the error measure.
+For the orbital response, the error measure remains small for all masses.
+For the constellation response, we find that $\epsilon_{A1}$ can grow up to $\sim 0.01$ at the lower end of the frequency band for high masses.
+
+The $\epsilon_{d}$ is signal-independent, and follows a quite different behaviour for the orbital and constellation responses.
+For the orbital response, $\epsilon_{d_{0}}$ grows for higher frequencies, reaching the $\sim 0.1$ at around $1\Hz$.
+For the constellation response, $\epsilon_{d_{c}}$ is mainly important at lower frequencies, reaching $\sim 0.01$ at around $10^{-5}\Hz$.
+
+Overall, we find that for merging binaries the perturbative approach presented in Sec.~\ref{sec:formalism} will be applicable, as the smallness of the annual frequency $f_{0} \simeq 3.1\times10^{-8}\mathrm{Hz}$, compared to the timescales relevant for the signal, ensures a good separation of timescales on the whole LISA frequency band.
+The most relevant higher-order corrections are expected to be the phase corrections close to the starting frequency of the signal, and the delay correction at high frequencies for the orbital response.
+For slowly chirping binaries, we have identified regions in the parameter space where the perturbative formalism of Sec.~\ref{sec:formalism} will break down, requiring an alternative approach presented below in Sec.~\ref{subsec:comblisa}.
 

### 2018-01-02 docs/paper_method.tex
Edits to the paper.

@@
 
+\SM{[talk about slowly chirping systems in the intro]}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Depending on the frequency $f$, this can lead $G(f,t)$ to have faster variations that its nominal timescale ($1\mathrm{yr}$ for LISA).
-As we will see in Sec.~\ref{} \SM{[R]}, to decide of the applicability of this approach we will not only need to compare timescales, but also take into account potential dimensionless factors of the form $2\pi f d$.
+As we will see in Sec.~\ref{} \SM{[R]}, to decide of the applicability of this approach we will not only need to compare timescales, but also take into account dimensionless factors of the form $2\pi f d$.
 
@@
 
-The definition~\eqref{eq:deftf} is a generalization of the time-to-frequency correspondence at the heart of the SPA~\eqref{eq:deftfSPA}.
+The definition~\eqref{eq:deftf} is a straightforward generalization of the time-to-frequency correspondence at the heart of the SPA~\eqref{eq:deftfSPA}.
 Indeed, using~\eqref{eq:deftfSPA} one can verify that the derivative of the SPA phase $\Psi_{\rm SPA}$~\eqref{eq:PsiSPA} with respect to $f$ yields back $\tfSPA$, as
@@
 
-If we restrict to the case of a pure modulation, $d=0$ and $G(f,t) = F(t)$, we can relate our result~\eqref{eq:resultdffPsiTf} to the approach of Ref.~\cite{KCY14}, where the authors extend the SPA in a formalism called the Shifted Uniform Asymptotic expansion (SUA) \SM{[look for other reference beyon KCY]}.
+If we restrict to the case of a pure modulation, $d=0$ and $G(f,t) = F(t)$, we can relate our result~\eqref{eq:resultdffPsiTf} to the approach of Ref.~\cite{KCY14}, where the authors extend the SPA in a formalism called the Shifted Uniform Asymptotic expansion (SUA) \SM{[look for other reference beyond KCY]}.
 %The derivation includes an expansion of the modulation in Bessel functions, a study of the displacement of the stationary phase point induced by the precession, and a resummation of the result in the form of time derivatives.
@@
 \end{align}
-When also using the rigid approximation for the constellation, where the armlengths are fixed, a particular simplification occurs when combining these individual delays, thanks to the relation $p^{L}_{r} - p^{L}_{s} =  L n_{l}$.
-In the local approximation~\eqref{eq:transferlocal}, the Fourier-domain transfer function then reads
+When also using the rigid approximation for the constellation, where the armlengths are fixed, a particular simplification occurs when combining these individual delays, thanks to the relation $p^{L}_{r} - p^{L}_{s} =  L n_{l}$:
+\begin{align}\label{eq:GslrL}
+	G_{slr}^{L}(f,t) &= \frac{i \pi f L}{2} \sinc \left[ \pi f L\left(1-\hatk\cdot n_{l} \right) \right] \nn\\
+	& \quad \cdot \exp\left[ i \pi f \left( L + \hatk\cdot \left( p_{1}^{L} + p_{2}^{L} \right) \right) \right]  n_{l} \cdot P_{\ell m} \cdot n_{l} \,,
+\end{align}
+with all time-dependent vectors evaluated at $t$. In the local approximation~\eqref{eq:transferlocal}, the Fourier-domain transfer function then reads
 \begin{align}\label{eq:transferLlocal}
-	\calT_{slr}^{L, \mathrm{local}} &= \frac{i \pi f L}{2} \sinc \left[ \pi f L\left(1-\hatk\cdot n_{l} \right) \right] \nn\\
-	& \quad \cdot \exp\left[ i \pi f \left( L + \hatk\cdot \left( p_{1}^{L} + p_{2}^{L} \right) \right) \right]  n_{l} \cdot P_{\ell m} \cdot n_{l} \,,
-\end{align}
-where all vectors $n_{l}$ and $p_{A}^{L}$ are to be evaluated at the time $t_{f}$.
-This result~\eqref{eq:transferlocal} is familiar, as it was derived in the past taking the point of view of monochromatic signals (see e.g.~\cite{} \SM{[C]}).
+	\calT_{slr}^{L, \mathrm{local}}(f) &= G_{slr}^{L}(f, \tf) \,.
+\end{align}
+This leading-order result~\eqref{eq:transfer0local}-\eqref{eq:transferLlocal} is familiar, as it was derived in the past taking the point of view of monochromatic signals (see e.g.~\cite{} \SM{[C]}).
 
@@
 
-The response~\eqref{eq:transferLlocal} can be interpreted as the Fourier-domain translation of a discrete derivative taken on the waveform.
+The response~\eqref{eq:GslrL}-\eqref{eq:transferLlocal} can be interpreted as the Fourier-domain translation of a discrete derivative taken on the waveform.
 The factor shows that the amplitude of the response is $\propto f$ in the low-frequency limit $f\ll f_{L}$, where the other factors are essentially unity.
@@
 From~\eqref{eq:transferLlocal}, an expansion for small $f\ll f_{L}$ yields back a Fourier-domain analog of the low-frequency approximation of the response~\cite{Cutler97, RCP04}, which amounts to having two LIGO-type interferometers turned by $\pi/4$ and set in motion.
+See Fig.~\ref{} and ~\ref{} \SM{[R]} for examples of LISA transfer functions.
+
+For the gravitational waveform itself, we will use the PhenomD model~\cite{Khan+15,Husa+15}, which provides Fourier-domain inspiral-merger-ringdown waveforms for aligned spins.
 
@@
 
-\subsection{Error control for the Fourier-domain response}
+\subsection{Errors in the Fourier-domain response for merging binaries}
 \label{subsec:errorsLISA}
@@
 
-Having at hand generic estimates for where the higher-order corrections in the LISA response are expected to be relevant, we now demonstrate that including them indeed reduces the reconstruction error of the Fourier-domain waveform.
-We will use an example PhenomD~\cite{Khan+15,Husa+15} waveform for an equal-mass and non-spinning system, and consider only two values for the total mass, $m=10^{7} \Msol$ and $m=10^{2} \Msol$.
-For this example, we will also focus on the orbital response and on a single-arm constellation response $y_{132}$, as TDI combinations will appear (within the approximations listed in Sec.~\ref{} \SM{[R]}) as linear combinations of these basic observables.
+Having at hand estimates for the relevance of higher-order corrections in the LISA response, we now demonstrate that including these higher-order corrections does indeed reduce the reconstruction error of the Fourier-domain transfer function.
+We will consider an equal-mass and non-spinning system, with only two values for the total mass, $M=10^{7} \Msol$ and $M=10^{2} \Msol$.
+For this example, we will also focus on the orbital response and on a single-arm constellation response $y_{132}$, as TDI combinations will appear (within the approximations listed in Sec.~\ref{subsec:modelLISA}) as linear combinations of these basic observables.
+The orientation angles chosen for this example are $[\lambda,\beta,\iota,\psi] = [\pi/4,\pi/3,\pi/3,0]$.
+
 On one hand, we perform a numerical inverse Fourier transform (IFFT) of the Fourier-domain PhenomD waveform, process the signal through the time-domain response~\eqref{eq:defresponse0} or~\eqref{eq:defresponseL}, and perform a numerical Fourier transform (FFT).
-This gives us the target Frourier-domain waveform, considered to be exact but for the possible artifacts of numerical Fourier transforms (most notably, oscillations induced by the necessary tapering of the signal).
-On the other hand, we process the Fourier-domain signal through the response summarized in Sec.~\ref{} \SM{[R]}, including various higher-order corrections.
+This gives us the target Fourier-domain waveform, considered to be exact but for possible numerical artifacts (most notably, oscillations induced by the necessary tapering of the signal).
+On the other hand, we process the Fourier-domain signal through the response summarized in Sec.~\ref{subsec:executivesummary}, including various higher-order corrections.
 We then compare the output of the two procedures.
@@
 In both cases, in accordance with~\eqref{eq:transfer0local}, the transfer function is essentially a phase, and contains many more cycles in the low-mass case in the higher frequency band.
-We find that including higher-order corrections does reduce the reconstruction errors, but with the distincitve feature that including only one of the amplitude and delay corrections can make the approximation actually worse, while including both does improve the accuracy.
+We find that including higher-order corrections does reduce the reconstruction errors, but with the distinctive feature that including only one of the amplitude and delay corrections can make the error actually worse, while including both does improve the accuracy.
 In the high-mass case, we also find for all orders of approximation an increase in the errors at the very end.
@@
   \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e7dOphase.pdf}
-  \caption{Error in the transfer function for the orbital delay $d_{0}$, for $M=10^{7} \Msol$.}
+  \caption{Error in the transfer function for the orbital delay $d_{0}$, for $M=10^{7} \Msol$. \SM{[Add line for RD freq]}}
   \label{fig:LISAerrorM1e7orb}
@@
   \includegraphics[width=.98\linewidth]{plots/LISAerrorM1e7y12L.pdf}
-  \caption{Error in the transfer function for the basic observable $y_{12}$, for $M=10^{7} \Msol$.}
+  \caption{Error in the transfer function for the basic observable $y_{12}$, for $M=10^{7} \Msol$. \SM{[Add line for RD freq]}}
   \label{fig:LISAerrorM1e7const}
@@
 
-In the LISA case, the modulations and delays entering~\eqref{eq:yslr} are periodic, with a period of one year and a frequency $f_{0} = 1/\mathrm{yr} \simeq 3.2\times 10^{-8}\mathrm{Hz}$.
+In the LISA case, the modulations and delays entering~\eqref{eq:yslr} are periodic, with a period of one year and a frequency $f_{0} = \Omega_{0}/2\pi = 1/\mathrm{yr} \simeq 3.2\times 10^{-8}\mathrm{Hz}$.
 If the above perturbative formalism cannot be applied, we can still exploit this periodic structure to transform the .
 \SM{[]}
-\be
-	G(f,t) = \sum_{n} c_{n}(f) e^{-in\Omega_{0}t} \,.
+\be\label{eq:Gdiscretefourier}
+	G(f,t) = \sum_{n} c_{n}(f) e^{-in\Omega_{0}t} \,,
+\ee
+with frequency-dependent discrete Fourier coefficients given by the integrals (we recall our Fourier convention~\eqref{eq:defFT})
+\be\label{eq:defcn}
+	c_{n}(f) = \frac{\Omega_{0}}{2\pi} \int_{0}^{\frac{2\pi}{\Omega_{0}}} \ud t \, e^{i n \Omega_{0} t} G(f,t) \,.
 \ee
@@
 Although straightforward and generic, this writing of the Fourier-domain response function will be in general more expansive to compute than the perturbative transfer function approach of Sec.~\ref{sec:formalism}.
+
+Note that for slowly-chirping binaries, the large features in the orbital transfer functions (see Fig.~\ref{} \SM{[R]} and Fig.~\ref{} \SM{[R]}) start to dominate signal-dependent derivatives.
+For the clarity of the presentation, we still do two separate comparisons for the orbital response~\eqref{} \SM{[R]} and the constellation response~\eqref{} \SM{[R]}, while the actual treatment would require to process the signal through the two steps of the response at once, instead of sequentially.
+This would result in replacing the discrete comb~\eqref{eq:defcn} by a double sum, however the extent of the constellation $c_{n}$ is rather limited in $n$.
 
@@
 The vertical lines indicate the merger frequencies.
-\SM{[color for vertical lines at fpeak]}}
+\SM{[color for vertical lines at fpeak]} \SM{[add bottom panel with residual ?]}}
   \label{fig:trigopoly}
@@
 \ee
-which is the expression of the inverse DFT in our sign convention~\eqref{eq:defFT}, the coefficients $c_{k}$ are given by
+which is the expression of the inverse DFT in our sign convention~\eqref{eq:defFT}, the coefficients $c_{k}$ are given by \SM{[move generic trigopoly things to the appendix, as it is common with comb approach for LISA]}
 \begin{align}\label{eq:ckyk}
@@
 
-For the $--$ case, the mode hierarchy is not respected, as the frame trajectory does not quite follow the picture of simple precession.
+For the $--$ case, the mode hierarchy is not respected, as the frame trajectory does not quite follow the picture of simple precession \SM{[mention angle Ji-Jf]}.
 The precession velocity is much milder, leading to smaller errors.
@@
 Thus, as is well known, apart from the relative phases of the modes for waveforms that include higher harmonics, the intrinsic information on the phasing is contained in the second derivative of the Fourier-domain phase $\Psi(f)$.
+\SM{[rewrite this paragraph]}
 

### 2018-01-23 docs/paper_method.tex
Edits to the paper, mainly plots.

@@
 
-We now turn to the transposition of~\eqref{eq:defresponse0} and~\eqref{eq:defresponseL} to the Fourier domain.
-For the leading-order response~\eqref{eq:transferlocal}, applying a pure delay as in~\eqref{eq:defresponse0} translates into Fourier-domain transfer function common to all modes, that is a pure phase factor, proportional to the frequency but also $t_{f}$-dependent:
+We now turn to the transposition of~\eqref{eq:defresponse0} and~\eqref{eq:defresponseL} to the Fourier domain. Applying a pure delay as in~\eqref{eq:defresponse0} translates into
+\be\label{eq:G0}
+	G_{0}(f, t) = e^{-2i\pi f d_{0}(t)} \,,
+\ee
+with $d_{0} = -\hatk \cdot p_{0}$ the delay associated to the orbit around the Sun.
+For the leading-order response~\eqref{eq:transferlocal}, this gives a Fourier-domain transfer function common to all modes, that is a pure phase factor, proportional to the frequency but also $t_{f}$-dependent:
 \be\label{eq:transfer0local}
-	\calT_{0}^{\rm local}(f) = e^{-2i\pi f d_{0}(t_{f})}\,,
-\ee
-with $d_{0} = -\hatk \cdot p_{0}$ the delay associated to the orbit around the Sun.
-If $(\lambda, \beta)$ are the ecliptic longitude and latitude of the source in the sky, and if the orbital phase is set to $0$ at $t=0$, this delay has the simple expression
+	\calT_{0}^{\rm local}(f) = G_{0}(f, \tf)\,.
+\ee
+If $(\lambda, \beta)$ are the ecliptic longitude and latitude of the source in the sky, and if the orbital phase is set by convention to $0$ at $t=0$, the orbital delay has the simple expression
 \be\label{eq:delay0}
@@
   \centering
-  \includegraphics[width=.32\linewidth]{plots/lisafom_Psiorb.pdf}
-  \includegraphics[width=.32\linewidth]{plots/lisafom_A1orb.pdf}
-  \includegraphics[width=.32\linewidth]{plots/lisafom_dorb.pdf}
-  %
-  \includegraphics[width=.32\linewidth]{plots/lisafom_Psiconst.pdf}
-  \includegraphics[width=.32\linewidth]{plots/lisafom_A1const.pdf}
-  \includegraphics[width=.32\linewidth]{plots/lisafom_dconst.pdf}
+%  \includegraphics[width=.32\linewidth]{plots/lisafom_Psiorb.pdf}
+%  \includegraphics[width=.32\linewidth]{plots/lisafom_A1orb.pdf}
+%  \includegraphics[width=.32\linewidth]{plots/lisafom_dorb.pdf}
+%  %
+%  \includegraphics[width=.32\linewidth]{plots/lisafom_Psiconst.pdf}
+%  \includegraphics[width=.32\linewidth]{plots/lisafom_A1const.pdf}
+%  \includegraphics[width=.32\linewidth]{plots/lisafom_dconst.pdf}
+  \includegraphics[width=.98\linewidth]{plots/lisafom_py.pdf}
   \caption{Error estimates as defined in~\eqref{eq:deffom}, for an equal-mass, non-spinning system and for total masses $M=10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$.
@@
 The limit $\epsilon_{\Psi} = 1$ is used to single out areas where the perturbative treatment of Sec.~\ref{sec:formalism} is expected to break down, although the precise location of this boundary will vary, depending on the accuracy level required and on the orientation angles.
-We find that, in the LIGO/Virgo mass range $10^{1}-10^{2} \Msol$, we reach the limit $\epsilon_{\Psi}=1$ first for the orbital response, for time-to-merger values that we expect to observe~\cite{Sesana16}.
+We also single out the high-mass region where the start of the signal is set by its entry in the frequency band, taken to be $10^{-5}\Hz$, so that there is no dependence on the time to merger anymore.
+We find that, in the LIGO/Virgo mass range $10^{1}-10^{2} \Msol$, we reach the limit $\epsilon_{\Psi}=1$ first for the orbital response, for time-to-merger values within the expected observed range~\cite{Sesana16}.
 We will introduce in Sec~\ref{subsec:comblisa} an alternative approach to deal with those signals.
-For completeness, we also show the result for intermediate and massive black hole systems, although we expect to observe mainly merging binaries for those masses.
+For completeness, we also show in Fig.~\ref{fig:lisafomPsiMcDeltat} the result for intermediate and massive black hole systems, although we expect to observe mainly merging systems for this mass range.
 
@@
 For this example, we will also focus on the orbital response and on a single-arm constellation response $y_{132}$, as TDI combinations will appear (within the approximations listed in Sec.~\ref{subsec:modelLISA}) as linear combinations of these basic observables.
-The orientation angles chosen for this example are $[\lambda,\beta,\iota,\psi] = [\pi/4,\pi/3,\pi/3,0]$.
+The orientation angles chosen for these examples are $[\lambda,\beta,\iota,\psi] = [\pi/4,\pi/3,\pi/3,0]$.
 
@@
 
-\begin{figure*}
+\begin{figure}
   \centering
-  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e7dOamp.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e7dOphase.pdf}
-  %
-  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e7dOamp.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e7dOphase.pdf}
+%  \includegraphics[width=.8\linewidth]{plots/LISAtransferM1e7dOamp.pdf}
+%  \hspace{0.1cm}
+%  \includegraphics[width=.8\linewidth]{plots/LISAtransferM1e7dOphase.pdf}
+%  \hspace{0.1cm}
+%  \includegraphics[width=.8\linewidth]{plots/LISAerrorM1e7dO.pdf}
+  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e7dO_py.pdf}
   \caption{Error in the transfer function for the orbital delay $d_{0}$, for $M=10^{7} \Msol$. \SM{[Add line for RD freq]}}
   \label{fig:LISAerrorM1e7orb}
-\end{figure*}
-
-\begin{figure*}
+\end{figure}
+
+\begin{figure}
   \centering
-  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e2dOamp.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/LISAtransferM1e2dOphase.pdf}
-  %
-  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2dOamp.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/LISAerrorM1e2dOphase.pdf}
+%  \includegraphics[width=.8\linewidth]{plots/LISAtransferM1e2dOamp.pdf}
+%  \hspace{0.1cm}
+%  \includegraphics[width=.8\linewidth]{plots/LISAtransferM1e2dOphase.pdf}
+%  \hspace{0.1cm}
+%  \includegraphics[width=.8\linewidth]{plots/LISAerrorM1e2dO.pdf}
+  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e2dO_py.pdf}
   \caption{Transfer function and reconstruction error for the orbital delay $d_{0}$, for $M=10^{2} \Msol$.}
   \label{fig:LISAerrorM1e2orb}
-\end{figure*}
+\end{figure}
 
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/LISAtransferM1e7y12LReIm.pdf}
-  %
-  \includegraphics[width=.98\linewidth]{plots/LISAerrorM1e7y12L.pdf}
+  %\includegraphics[width=.98\linewidth]{plots/LISAtransferM1e7y12LReIm.pdf}
+  %\includegraphics[width=.98\linewidth]{plots/LISAerrorM1e7y12L.pdf}
+  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e7y12L_py.pdf}
   \caption{Error in the transfer function for the basic observable $y_{12}$, for $M=10^{7} \Msol$. \SM{[Add line for RD freq]}}
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/LISAtransferM1e2y12LReIm.pdf}
+  %\includegraphics[width=.98\linewidth]{plots/LISAtransferM1e2y12LReIm.pdf}
   %
-  \includegraphics[width=.98\linewidth]{plots/LISAerrorM1e2y12L.pdf}
-  \caption{Transfer function and reconstruction error for the basic observable $y_{12}$, for $M=10^{2} \Msol$.}
+  %\includegraphics[width=.98\linewidth]{plots/LISAerrorM1e2y12L.pdf}
+  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e2y12L_py.pdf}
+  \caption{Transfer function and reconstruction error for the basic observable $y_{12}$, for $M=10^{2} \Msol$. \SM{[check $\overline{\calT}$ is defined]}}
   \label{fig:LISAerrorM1e2const}
@@
 
-In the LISA case, the modulations and delays entering~\eqref{eq:yslr} are periodic, with a period of one year and a frequency $f_{0} = \Omega_{0}/2\pi = 1/\mathrm{yr} \simeq 3.2\times 10^{-8}\mathrm{Hz}$.
-If the above perturbative formalism cannot be applied, we can still exploit this periodic structure to transform the .
-\SM{[]}
+\begin{figure}
+  \centering
+  \includegraphics[width=.98\linewidth]{plots/LISACombExtent_py.pdf}
+  \caption{Extent of the Fourier comb corresponding to truncation levels $\eta=10^{-12}$, $10^{-6}$, $10^{-3}$ according to the criterion~\eqref{eq:criteriontruncationcomb}.
+Blue corresponds to the orbital delay comb~\eqref{}, while red corresponds to the one-arm constellation response~\eqref{} \SM{[R]}.
+The result is shown averaged over 100 random orientation angles.}
+  \label{fig:lisacombextent}
+\end{figure}
+
+\begin{figure}
+  \centering
+%  \includegraphics[width=.8\linewidth]{plots/LISAtransferSOBHepsPsi1d0ampcorr.pdf}
+%  \hspace{0.1cm}
+%  \includegraphics[width=.8\linewidth]{plots/LISAtransferSOBHepsPsi1d0phasecorr.pdf}
+%  \hspace{0.1cm}
+%  \includegraphics[width=.8\linewidth]{plots/LISAerrorSOBHepsPsi1d0.pdf}
+  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi1d0corr_py.pdf}
+  \caption{Error in the transfer function for the orbital delay $d_{0}$, for a SOBH with $\epsilon_{\Psi}\sim 1$.}
+  \label{fig:}
+\end{figure}
+
+\begin{figure}
+  \centering
+%  \includegraphics[width=.8\linewidth]{plots/LISAtransferSOBHepsPsi10d0ampcorr.pdf}
+%  \hspace{0.1cm}
+%  \includegraphics[width=.8\linewidth]{plots/LISAtransferSOBHepsPsi10d0phasecorr.pdf}
+%  \hspace{0.1cm}
+%  \includegraphics[width=.8\linewidth]{plots/LISAerrorSOBHepsPsi10d0.pdf}
+  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi10d0corr_py.pdf}
+  \caption{Error in the transfer function for the orbital delay $d_{0}$, for a SOBH with $\epsilon_{\Psi}\sim 10$.}
+  \label{fig:}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\linewidth]{plots/LISAtransferSOBHepsPsi1y12LReIm.pdf}
+  \hspace{0.1cm}
+  \includegraphics[width=.8\linewidth]{plots/LISAerrorSOBHepsPsi1y12L.pdf}
+  \caption{Error in the transfer function for the constellation response $y_{132}$, for a SOBH with $\epsilon_{\Psi}\sim 1$.}
+  \label{fig:}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\linewidth]{plots/LISAtransferSOBHepsPsi10y12LReIm.pdf}
+  \hspace{0.1cm}
+  \includegraphics[width=.8\linewidth]{plots/LISAerrorSOBHepsPsi10y12L.pdf}
+  \caption{Error in the transfer function for the constellation response $y_{132}$, for a SOBH with $\epsilon_{\Psi}\sim 10$.}
+  \label{fig:}
+\end{figure}
+
+As explained in Sec.~\ref{subsec:lisafom}, slowly-chirping binaries can be problematic for the perturbative formalism of Sec.~\ref{sec:formalism}.
+Indeed, when far enough from merger, SOBH systems~\cite{Sesana16} can have error estimates $\epsilon_{\Psi}$ reaching and exceeding $1$, as shown in Fig.~\ref{fig:lisafomPsiMcDeltat}.
+Here we investigate these sources in more details, and propose an alternative treatment for their response.
+
+First, we should mention that these signals share similarities, at least qualitatively, with galactic binaries.
+They are far from merger, slowly chirping, and span only a narrow frequency band over the course of the LISA mission lifetime.
+For the galactic binaries, an accurate and very efficient numerical treatment of the response has been proposed in~\cite{} \SM{[C]} and widely used in applications (see e.g.~\cite{} \SM{[C]}).
+This approach is referred to as the fast-slow decomposition, or heterodyning.
+If the gravitational wave signal extends only on the narrow frequency band $f \in [f_{*}, (1+\eta) f_{*}]$ with $\eta \ll 1$, then scaling out a carrier frequency of the signal by multpilying by $e^{-2 i \pi f_{*} t}$ will eliminate most of the time variability, allowing to process the signal through the time-domain response and to take a FFT with a Nyquist frequency shifted from $(1+\eta) f_{*}$ to $\eta f_{*}$, i.e.
+with a much smaller number of samples.
+This multiplication is simply equivalent to a shift in frequency domain, which can be restored after the numerical FFT has been computed.
+The efficiency of this approach is contingent to the smallness of $\eta$.
+For galactic binaries, this quantity is typically \SM{[complete]}.
+For SOBH systems, however, $\eta$ will take a continuous set of values from roughly $10^{-4}$ to $1$ \SM{[check values]}.
+Although the systems for which $\eta$ is the largest are also the easiest to treat with the perturbative formalism of Sec~\ref{sec:formalism}, we propose here yet a third method for dealing with potentially troublesome systems, leaving for the future a more detailed study of SOBH systems as a population, and the investigation of the precise boundaries and oberlap areas of the three methods as well as their respective computational costs.
+
+This third approach exploits the fact that, in the LISA case, the modulations and delays entering~\eqref{eq:yslr} are periodic, with a period of one year and a frequency $f_{0} = \Omega_{0}/2\pi = 1/\mathrm{yr} \simeq 3.2\times 10^{-8}\mathrm{Hz}$.
+For any given frequency, $G(f,t)$ is periodic in time, so that~\eqref{eq:defG} becomes a discrete Fourier series
 \be\label{eq:Gdiscretefourier}
-	G(f,t) = \sum_{n} c_{n}(f) e^{-in\Omega_{0}t} \,,
+	G(f,t) = \sum_{n \in \mathbb{Z}} c_{n}(f) e^{-in\Omega_{0}t} \,,
 \ee
@@
 
-This leads to the following generalized discrete convolution: \SM{[]}
-\be
+In the case of the orbital delay, with $G_{0}(f,t)$ given by~\eqref{eq:G0} and~\eqref{eq:delay0}, the particularly simple expression of the delay gives an analytic expression for the coefficients $c_{n}^{0}$ in terms of Bessel functions of the first kind, as
+\be\label{eq:cn0}
+	c_{n}^{0} = i^{n} e^{i n \lambda} J_{n} \left[ -2 \pi f R \cos \beta \right]\,.
+\ee
+For the constellation response, by contrast, to our knowledge the coefficients $c_{n}^{L}$ do not admit such a simple close-form expression, and they must be computed numerically using the integrand given by~\eqref{eq:GslrL}.
+Truncating~\eqref{eq:Gdiscretefourier} to a finite order $N$, this computation reduces to a DFT \SM{[acronym]}, and the $c_{n}$ coefficients are given by~\eqref{eq:ykDFT} and~\eqref{eq:ckyk}.
+
+Inserting~\eqref{eq:Gdiscretefourier} into the convolution~\eqref{eq:FDkernel} leads to the following generalized discrete convolution:
+\be\label{eq:transferdiscreteconvolution}
 	\tilde{s}(f) = \sum_{n \in \mathbb{Z}} c_{n}(f - n f_{0}) \tilde{h} (f - n f_{0}) \,.
 \ee
+In practice, this sum is to be truncated at $|n|\leq N$, for some finite order $N$ determining the accuracy of the approximation.
+To assess the expected truncation error, we use a simple criterion based on the $L^{1}$-norm of the $\{c_{n}\}$ sequence.
+For a given target truncation error $\eta$, we define the truncation $N(f, \eta)$ as the smallest integer such that
+\be\label{eq:criteriontruncationcomb}
+	\sum_{|n| > N(f, \eta)} |c_{n}(f)| < \eta \,.
+\ee
+The truncation order $N(f,\eta)$ is frequency-dependent, and also depends on the orientation angles.
+For $c_{n}^{L}$, it also determines the number of function evaluations and the cost of the DFT.
+For $c_{n}^{0}$, the expression~\eqref{eq:cn0} can provide an asymptotic estimate.
+For large values of $n$ we have indeed the equivalent $| J_{n}[z] | \sim (e z/2n)^{n}/\sqrt{2\pi n} \; \left[ n\rightarrow + \infty \right]$~\cite{} \SM{[C: dlmf?]}, so that a conservative estimate for the truncation order (almost independent of $\eta$) is given by $N \geq  |e \pi f R \cos \beta|$.
+Fig.~\ref{fig:lisacombextent} shows numerical computations for $N(f,\eta)$, averaged over orientation angles, for both the constellation and the orbital response and for the truncation levels  $\eta=10^{-12}$, $10^{-6}$, $10^{-3}$.
+We see that, especially at high frequencies, the orbital response requires more coefficients than the constellation response due to the longer baseline of the delay.
+At the frequency $f=0.1\Hz$, we have for the orbital response $N \simeq 200$.
+
+An important point is that, since the signals from slowly-chirping binaries extend on a narrow frequency band, for both responses the frequency-dependence of $c_{n}^{0}(f)$, $c_{n}^{L}(f)$ will be very mild, so that the coefficients can be computed at two or three frequencies and interpolated in-between.
+The computational cost of this approach is set by the truncation order $N$, and the number of intermediate frequencies where the transfer function resulting from the sum~\eqref{eq:transferdiscreteconvolution} must be computed for later interpolation, with the necessary waveform evaluations \SM{[possibly refer to appendix on interpolation errors here]}.
+We leave for future work a proper assessment of the computational performance of this method compared to the others \SM{[say a bit more ?]}.
 
@@
 
-The convention we will be using for the Fourier transform of a signal $h(t)$ is
-\be\label{eq:defFT}
-	\tilde{h}(f) = \mathrm{FT}[h](f) =  \int \ud t \, e^{+2i\pi f t} h(t) \,.
-\ee
+The convention we will be using for the Fourier transform of a signal $h(t)$ and its inverse is
+\begin{subequations}
+\label{eq:defFT}
+\begin{align}
+	\tilde{h}(f) &= \mathrm{FT}[h](f) =  \int \ud t \, e^{+2i\pi f t} h(t) \,, \\
+	h(t) &= \mathrm{IFT}[\tilde{h}](t) =  \int \ud f \, e^{-2i\pi f t} \tilde{h}(f) \,.
+\end{align}
+\end{subequations}
 Notice that this sign convention is not the most frequently used in the literature.
@@
 An additional approximation often used in waveform models consists in considering that the Fourier transforms $\tilde{h}_{\ell m}$ have support only on one side of the spectrum, either for positive of for negative frequencies depending on the sign of $m$.
-This holds in particular within the stationary phase approximation, as explained below.
+This holds in particular within the stationary phase approximation, as explained below \SM{[change here]}.
 With our sign convention~\eqref{eq:defFT}, this approximation reads
@@
 \SM{[rewrite this paragraph]}
+
+In our convention, the link between the Discrete Fourier Transform (DFT) \SM{[check acronym]} and the trigonometric polynomial representation of a function goes as follows.
+For a periodic function $F(x)$ defined on $x\in [x_{0}, x_{0} + \Delta x]$, and represented by $N$ samples $x_{j} = \ov{x} + j \Delta x/N$, $j=0,\dots,N-1$, with $N$ large enough to satisfy, at least approximately, the Nyquist criterion, we can build a trigonometric interpolant $P(x)$ as
+\be
+	P(x) = \sum\limits_{k=-M}^{+M} c_{k} e^{2i\pi k \frac{x-x_{0}}{\Delta x}} \,,
+\ee
+that will satisfy the system $P(x_{j}) = F(x_{j})$ for $j=0,\dots, N-1$.
+Here we set $M=N/2$, assuming N is even.
+This trigonometric polynomial representation is equivalent to truncating the formal Fourier series, representing the full signal, to a finite order $M$.
+The coefficients $c_{k}$, in the full series, are defined as integrals over the interval of $F$ weighted by the appropriate exponential (as in~\eqref{eq:defcn}).
+In both interpretations, either the truncated approximation of the Fourier series or  the trigonometric interpolation formulation, these coefficients are related to the coefficients of the inverse DFT.
+If we set $\omega \equiv e^{2i\pi/N}$ and define
+\be\label{eq:ykDFT}
+	y_{k} = \frac{1}{N} \sum\limits_{j=0}^{N-1} F(x_{j}) \omega^{jk} \,,
+\ee
+which is the expression of the inverse DFT in our sign convention~\eqref{eq:defFT}, the coefficients $c_{k}$ are given by
+\begin{align}\label{eq:ckyk}
+	c_{k} &= y_{k} \text{ for } k=0,\dots, M-1 \,, \nn\\
+	c_{k} &= y_{k+N} \text{ for } k=-M+1,\dots, -1 \,, \nn\\
+	c_{M} &= c_{-M} = \frac{y_{M}}{2} \,,
+\end{align}
+where the condition $c_{M} = c_{-M}$ is an arbitrary condition enforced to match the number of degrees of freedom.
+In practice, a good representation of the Fourier series of the signal is achieved when the truncation order $M$ is sufficient so that the coefficients $c_{|n|\geq M}$ become negligibly small.
 

### 2018-01-26 docs/paper_method.tex
Edits to the paper, mostly plots.
Updates on the paper.

@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/tf.pdf}
+  \includegraphics[width=.98\linewidth]{plots/tf_py.pdf}
   \caption{Time-to-frequency correspondence $t_{f}$, as defined directly from the phase of the Fourier-domain signal as in~\eqref{eq:deftf}, in geometric units.
@@
   \centering
-  \includegraphics[width=.99\linewidth]{plots/TfTA.pdf}
+  %\includegraphics[width=.99\linewidth]{plots/TfTA.pdf}
+  \includegraphics[width=.98\linewidth]{plots/TfTA_py.pdf}
   \caption{Fourier-domain amplitude and phase timescales, as defined in~\eqref{eq:defTf} and~\eqref{eq:defTA}, in geometric units for an equal-mass, non-spinning system.
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e7dO_py.pdf}
-  \caption{Error in the transfer function for the orbital delay $d_{0}$, for $M=10^{7} \Msol$. \SM{[Add line for RD freq]}}
+  \caption{Error in the transfer function for the orbital delay $d_{0}$, for $M=10^{7} \Msol$.}
   \label{fig:LISAerrorM1e7orb}
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e7y12L_py.pdf}
-  \caption{Error in the transfer function for the basic observable $y_{12}$, for $M=10^{7} \Msol$. \SM{[Add line for RD freq]}}
+  \caption{Error in the transfer function for the basic observable $y_{12}$, for $M=10^{7} \Msol$.}
   \label{fig:LISAerrorM1e7const}
@@
   \includegraphics[width=.98\linewidth]{plots/LISACombExtent_py.pdf}
-  \caption{Extent of the Fourier comb corresponding to truncation levels $\eta=10^{-12}$, $10^{-6}$, $10^{-3}$ according to the criterion~\eqref{eq:criteriontruncationcomb}.
-Blue corresponds to the orbital delay comb~\eqref{}, while red corresponds to the one-arm constellation response~\eqref{} \SM{[R]}.
+  \caption{Extent of the Fourier-domain comb entering the convolution~\eqref{eq:transferdiscreteconvolution}, determined from the criterion~\eqref{eq:criteriontruncationcomb} with truncation levels $\eta=10^{-12}$, $10^{-6}$, $10^{-3}$.
+Blue corresponds to the orbital delay comb coefficients $c_{n}^{0}$ given by~\eqref{eq:cn0}, and red to the constellation comb $c_{n}^{L}$ for the response~\eqref{eq:GslrL}.
 The result is shown averaged over 100 random orientation angles.}
@@
   \centering
-  \includegraphics[width=.8\linewidth]{plots/LISAtransferSOBHepsPsi1y12LReIm.pdf}
-  \hspace{0.1cm}
-  \includegraphics[width=.8\linewidth]{plots/LISAerrorSOBHepsPsi1y12L.pdf}
+%  \includegraphics[width=.8\linewidth]{plots/LISAtransferSOBHepsPsi1y12LReIm.pdf}
+%  \hspace{0.1cm}
+%  \includegraphics[width=.8\linewidth]{plots/LISAerrorSOBHepsPsi1y12L.pdf}
+  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi1y12L_py.pdf}
   \caption{Error in the transfer function for the constellation response $y_{132}$, for a SOBH with $\epsilon_{\Psi}\sim 1$.}
@@
   \centering
-  \includegraphics[width=.8\linewidth]{plots/LISAtransferSOBHepsPsi10y12LReIm.pdf}
-  \hspace{0.1cm}
-  \includegraphics[width=.8\linewidth]{plots/LISAerrorSOBHepsPsi10y12L.pdf}
+%  \includegraphics[width=.8\linewidth]{plots/LISAtransferSOBHepsPsi10y12LReIm.pdf}
+%  \hspace{0.1cm}
+%  \includegraphics[width=.8\linewidth]{plots/LISAerrorSOBHepsPsi10y12L.pdf}
+  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi10y12L_py.pdf}
   \caption{Error in the transfer function for the constellation response $y_{132}$, for a SOBH with $\epsilon_{\Psi}\sim 10$.}
@@
 Indeed, when far enough from merger, SOBH systems~\cite{Sesana16} can have error estimates $\epsilon_{\Psi}$ reaching and exceeding $1$, as shown in Fig.~\ref{fig:lisafomPsiMcDeltat}.
-Here we investigate these sources in more details, and propose an alternative treatment for their response.
-
-First, we should mention that these signals share similarities, at least qualitatively, with galactic binaries.
+Here we investigate these sources in more details, and propose an alternative treatment for their instrument response.
+
+First, we should mention that these signals share similarities with the galactic binaries that will provide numerous quasi-monochromatic signals in the LISA band~\cite{LISA17}.
 They are far from merger, slowly chirping, and span only a narrow frequency band over the course of the LISA mission lifetime.
 For the galactic binaries, an accurate and very efficient numerical treatment of the response has been proposed in~\cite{} \SM{[C]} and widely used in applications (see e.g.~\cite{} \SM{[C]}).
-This approach is referred to as the fast-slow decomposition, or heterodyning.
+This treatment is referred to as the fast-slow decomposition, or heterodyning approach.
 If the gravitational wave signal extends only on the narrow frequency band $f \in [f_{*}, (1+\eta) f_{*}]$ with $\eta \ll 1$, then scaling out a carrier frequency of the signal by multpilying by $e^{-2 i \pi f_{*} t}$ will eliminate most of the time variability, allowing to process the signal through the time-domain response and to take a FFT with a Nyquist frequency shifted from $(1+\eta) f_{*}$ to $\eta f_{*}$, i.e.
@@
 For SOBH systems, however, $\eta$ will take a continuous set of values from roughly $10^{-4}$ to $1$ \SM{[check values]}.
-Although the systems for which $\eta$ is the largest are also the easiest to treat with the perturbative formalism of Sec~\ref{sec:formalism}, we propose here yet a third method for dealing with potentially troublesome systems, leaving for the future a more detailed study of SOBH systems as a population, and the investigation of the precise boundaries and oberlap areas of the three methods as well as their respective computational costs.
-
-This third approach exploits the fact that, in the LISA case, the modulations and delays entering~\eqref{eq:yslr} are periodic, with a period of one year and a frequency $f_{0} = \Omega_{0}/2\pi = 1/\mathrm{yr} \simeq 3.2\times 10^{-8}\mathrm{Hz}$.
-For any given frequency, $G(f,t)$ is periodic in time, so that~\eqref{eq:defG} becomes a discrete Fourier series
+Although the systems for which $\eta$ is the largest are also the easiest to treat with the perturbative formalism of Sec~\ref{sec:formalism}, we propose here yet a third method for dealing with potentially troublesome systems.
+We leave for the future a more detailed study of SOBH systems as a population, and the investigation of the precise boundaries and overlap areas of the three methods as well as their respective computational costs.
+
+The third approach we propose here exploits the fact that, in the LISA case, the modulations and delays entering~\eqref{eq:yslr} are periodic, with a period of one year and a frequency $f_{0} = \Omega_{0}/2\pi = 1/\mathrm{yr} \simeq 3.2\times 10^{-8}\mathrm{Hz}$.
+For any given frequency $f$, $G(f,t)$ is periodic in time, so that~\eqref{eq:defG} becomes a discrete Fourier series
 \be\label{eq:Gdiscretefourier}
@@
 \ee
-with frequency-dependent discrete Fourier coefficients given by the integrals (we recall our Fourier convention~\eqref{eq:defFT})
+with frequency-dependent discrete Fourier coefficients (that we will also call comb coefficients) given by the integrals (we recall our Fourier convention~\eqref{eq:defFT})
 \be\label{eq:defcn}
@@
 \ee
+Thus, computing the Fourier-domain response now requires to convolve the signal with a discrete comb with frequency-dependent coefficients $c_{n}$.
 In practice, this sum is to be truncated at $|n|\leq N$, for some finite order $N$ determining the accuracy of the approximation.
@@
 For $c_{n}^{L}$, it also determines the number of function evaluations and the cost of the DFT.
-For $c_{n}^{0}$, the expression~\eqref{eq:cn0} can provide an asymptotic estimate.
+For $c_{n}^{0}$, the expression~\eqref{eq:cn0} can provide an asymptotic bound on the width of the comb.
 For large values of $n$ we have indeed the equivalent $| J_{n}[z] | \sim (e z/2n)^{n}/\sqrt{2\pi n} \; \left[ n\rightarrow + \infty \right]$~\cite{} \SM{[C: dlmf?]}, so that a conservative estimate for the truncation order (almost independent of $\eta$) is given by $N \geq  |e \pi f R \cos \beta|$.
@@
 \SM{[explain somewhere the fact that the modulation has support for negative f, convolution has support to larger f - pretty important for direct convolution]}
+
+\SM{[cite GWFrames]}
 
@@
 
-\begin{figure*}
+\begin{figure}
   \centering
-  \includegraphics[width=.45\linewidth]{plots/eulerSXS0058alphagamma.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.45\linewidth]{plots/eulerSXS0058beta.pdf}
+%  \includegraphics[width=.45\linewidth]{plots/eulerSXS0058alphagamma.pdf}
+%  \hspace{0.2cm}
+%  \includegraphics[width=.45\linewidth]{plots/eulerSXS0058beta.pdf}
+  \includegraphics[width=.98\linewidth]{plots/eulerSXS0058_py.pdf}
   \caption{Evolution of Euler angles $(\alpha, \beta, \gamma)$ near merger for the example waveform SXS:BBH:0058~\cite{} \SM{[C]}.
@@
   \label{fig:prectoymodel}
-\end{figure*}
+\end{figure}
 
@@
 The columns show the orders of approximation $\{\Psi:0,A:0\}$, $\{\Psi:2,A:2\}$, $\{\Psi:2,A:2,\text{conv}\}$.
-\SM{[Compact the columns to share the x axis, add vertical line for merger frequency and shaded area for stitching pert/conv]} \SM{[check: where is the phase of the orange, mode 20 ?]}}
+\SM{[Compact the columns to share the x axis, add vertical line for merger frequency and shaded area for stitching pert/conv]} \SM{[check: where is the phase of the orange, mode 20 ?]} \SM{[add vertical lines for merger and RD freqs]}}
   \label{fig:precerrors++}
@@
 The columns show the orders of approximation $\{\Psi:0,A:0\}$, $\{\Psi:2,A:2\}$, $\{\Psi:2,A:2,\text{conv}\}$.
-\SM{[Compact the columns to share the x axis, add vertical line for merger frequency and shaded area for stitching pert/conv]}}
+\SM{[Compact the columns to share the x axis, add vertical line for merger frequency and shaded area for stitching pert/conv]} \SM{[add vertical lines for merger and RD freqs]}}
   \label{fig:precerrorspp}
@@
 \SM{[Compact the columns to share the x axis, add vertical line for merger frequency and shaded area for stitching pert/conv.
-Check the post-merger frame trajectory for - -.]}}
+Check the post-merger frame trajectory for - -.]} \SM{[add vertical lines for merger and RD freqs]}}
   \label{fig:precerrors--}
@@
   \centering
-  %
-  \includegraphics[width=.32\linewidth]{plots/mismatches_++_inc0.pdf}
-  \hspace{0.1cm}
-  \includegraphics[width=.32\linewidth]{plots/mismatches_++_incpi3.pdf}
-  \hspace{0.1cm}
-  \includegraphics[width=.32\linewidth]{plots/mismatches_++_incpi2.pdf}
-  %
-  \includegraphics[width=.32\linewidth]{plots/mismatches_pp_inc0.pdf}
-  \hspace{0.1cm}
-  \includegraphics[width=.32\linewidth]{plots/mismatches_pp_incpi3.pdf}
-  \hspace{0.1cm}
-  \includegraphics[width=.32\linewidth]{plots/mismatches_pp_incpi2.pdf}
-  %
-  \includegraphics[width=.32\linewidth]{plots/mismatches_--_inc0.pdf}
-  \hspace{0.1cm}
-  \includegraphics[width=.32\linewidth]{plots/mismatches_--_incpi3.pdf}
-  \hspace{0.1cm}
-  \includegraphics[width=.32\linewidth]{plots/mismatches_--_incpi2.pdf}
+%  %
+%  \includegraphics[width=.32\linewidth]{plots/mismatches_++_inc0.pdf}
+%  \hspace{0.1cm}
+%  \includegraphics[width=.32\linewidth]{plots/mismatches_++_incpi3.pdf}
+%  \hspace{0.1cm}
+%  \includegraphics[width=.32\linewidth]{plots/mismatches_++_incpi2.pdf}
+%  %
+%  \includegraphics[width=.32\linewidth]{plots/mismatches_pp_inc0.pdf}
+%  \hspace{0.1cm}
+%  \includegraphics[width=.32\linewidth]{plots/mismatches_pp_incpi3.pdf}
+%  \hspace{0.1cm}
+%  \includegraphics[width=.32\linewidth]{plots/mismatches_pp_incpi2.pdf}
+%  %
+%  \includegraphics[width=.32\linewidth]{plots/mismatches_--_inc0.pdf}
+%  \hspace{0.1cm}
+%  \includegraphics[width=.32\linewidth]{plots/mismatches_--_incpi3.pdf}
+%  \hspace{0.1cm}
+%  \includegraphics[width=.32\linewidth]{plots/mismatches_--_incpi2.pdf}
+  \includegraphics[width=.98\linewidth]{plots/precMM_py.pdf}
   \caption{Unfaithfulness.
@@
 Since we intend to work in the Fourier domain, we will use PhenomD waveforms~\cite{Khan+15}.
-These phenomenological waveforms provide analytic fits for both the Fourier-domain amplitude and phase of the 22 mode, with fit coefficients that were calibrated on EOB/NR hybrids [describe smoothing across intervals].
+These phenomenological waveforms provide analytic fits for both the Fourier-domain amplitude and phase of the 22 mode, with fit coefficients that were calibrated on EOB/NR hybrids \SM{[describe smoothing across intervals ?]}.
 Fig.~\ref{fig:ampphase} shows an example of amplitude and phase in geometric units for an equal-mass binary.

### 2018-01-30 docs/paper_method.tex
Edits to the paper plots.

@@
   \centering
-  \includegraphics[width=.32\linewidth]{plots/fom_pp_Psi2.pdf}
-  \hspace{0cm}
-  \includegraphics[width=.32\linewidth]{plots/fom_pp_A1.pdf}
-  \hspace{0cm}
-  \includegraphics[width=.32\linewidth]{plots/fom_pp_A2.pdf}
-  %
-  \includegraphics[width=.32\linewidth]{plots/fom_++_Psi2.pdf}
-  \hspace{0cm}
-  \includegraphics[width=.32\linewidth]{plots/fom_++_A1.pdf}
-  \hspace{0cm}
-  \includegraphics[width=.32\linewidth]{plots/fom_++_A2.pdf}
-  %
-  \includegraphics[width=.32\linewidth]{plots/fom_--_Psi2.pdf}
-  \hspace{0cm}
-  \includegraphics[width=.32\linewidth]{plots/fom_--_A1.pdf}
-  \hspace{0cm}
-  \includegraphics[width=.32\linewidth]{plots/fom_--_A2.pdf}
-  \caption{Error estimates of the approximation as defined in~\eqref{eq:deffom} for the precession modulations.
-\SM{[transpose, compact the columns to share the x axis, fix legends, add horizontal line for $10^{0}$]}}
+%  \includegraphics[width=.32\linewidth]{plots/fom_pp_Psi2.pdf}
+%  \hspace{0cm}
+%  \includegraphics[width=.32\linewidth]{plots/fom_pp_A1.pdf}
+%  \hspace{0cm}
+%  \includegraphics[width=.32\linewidth]{plots/fom_pp_A2.pdf}
+%  %
+%  \includegraphics[width=.32\linewidth]{plots/fom_++_Psi2.pdf}
+%  \hspace{0cm}
+%  \includegraphics[width=.32\linewidth]{plots/fom_++_A1.pdf}
+%  \hspace{0cm}
+%  \includegraphics[width=.32\linewidth]{plots/fom_++_A2.pdf}
+%  %
+%  \includegraphics[width=.32\linewidth]{plots/fom_--_Psi2.pdf}
+%  \hspace{0cm}
+%  \includegraphics[width=.32\linewidth]{plots/fom_--_A1.pdf}
+%  \hspace{0cm}
+%  \includegraphics[width=.32\linewidth]{plots/fom_--_A2.pdf}
+  \includegraphics[width=.98\linewidth]{plots/precfom_py.pdf}
+  \caption{Error estimates of the approximation as defined in~\eqref{eq:deffom} for the precession modulations. \SM{[complete caption]}}
   \label{fig:fomprec}
@@
   \centering
-  \includegraphics[width=.32\linewidth]{plots/precerrors_++_00_noconv_col.pdf}
-  \hspace{0.cm}
-  \includegraphics[width=.32\linewidth]{plots/precerrors_++_32_noconv_col.pdf}
-  \hspace{0.cm}
-  \includegraphics[width=.32\linewidth]{plots/precerrors_++_32_conv_col.pdf}
+%  \includegraphics[width=.32\linewidth]{plots/precerrors_++_00_noconv_col.pdf}
+%  \hspace{0.cm}
+%  \includegraphics[width=.32\linewidth]{plots/precerrors_++_32_noconv_col.pdf}
+%  \hspace{0.cm}
+%  \includegraphics[width=.32\linewidth]{plots/precerrors_++_32_conv_col.pdf}
+  \includegraphics[width=.98\linewidth]{plots/precerror_++_py.pdf}
   \caption{Amplitude of the inertial-frame modes $h^{I}_{2m}$ and errors in the Fourier-domain transfer fonctions $\calT^{2}_{m2}$ for the configuration $++$.
 The columns show the orders of approximation $\{\Psi:0,A:0\}$, $\{\Psi:2,A:2\}$, $\{\Psi:2,A:2,\text{conv}\}$.
-\SM{[Compact the columns to share the x axis, add vertical line for merger frequency and shaded area for stitching pert/conv]} \SM{[check: where is the phase of the orange, mode 20 ?]} \SM{[add vertical lines for merger and RD freqs]}}
+\SM{[Compact the columns to share the x axis, add vertical line for merger frequency and shaded area for stitching pert/conv]} \SM{[check: where is the phase of the orange, mode 20 ?]} \SM{[add vertical lines for merger and RD freqs]} \SM{[scaling of the amplitude ?]}}
   \label{fig:precerrors++}
@@
   \centering
-  \includegraphics[width=.32\linewidth]{plots/precerrors_pp_00_noconv_col.pdf}
-  \hspace{0.cm}
-  \includegraphics[width=.32\linewidth]{plots/precerrors_pp_32_noconv_col.pdf}
-  \hspace{0.cm}
-  \includegraphics[width=.32\linewidth]{plots/precerrors_pp_32_conv_col.pdf}
+%  \includegraphics[width=.32\linewidth]{plots/precerrors_pp_00_noconv_col.pdf}
+%  \hspace{0.cm}
+%  \includegraphics[width=.32\linewidth]{plots/precerrors_pp_32_noconv_col.pdf}
+%  \hspace{0.cm}
+%  \includegraphics[width=.32\linewidth]{plots/precerrors_pp_32_conv_col.pdf}
+  \includegraphics[width=.98\linewidth]{plots/precerror_pp_py.pdf}
   \caption{Amplitude of the inertial-frame modes $h^{I}_{2m}$ and errors in the Fourier-domain transfer fonctions $\calT^{2}_{m2}$ for the configuration $pp$.
@@
   \centering
-  \includegraphics[width=.32\linewidth]{plots/precerrors_--_00_noconv_col.pdf}
-  \hspace{0.cm}
-  \includegraphics[width=.32\linewidth]{plots/precerrors_--_32_noconv_col.pdf}
-  \hspace{0.cm}
-  \includegraphics[width=.32\linewidth]{plots/precerrors_--_32_conv_col.pdf}
+%  \includegraphics[width=.32\linewidth]{plots/precerrors_--_00_noconv_col.pdf}
+%  \hspace{0.cm}
+%  \includegraphics[width=.32\linewidth]{plots/precerrors_--_32_noconv_col.pdf}
+%  \hspace{0.cm}
+%  \includegraphics[width=.32\linewidth]{plots/precerrors_--_32_conv_col.pdf}
+  \includegraphics[width=.98\linewidth]{plots/precerror_--_py.pdf}
   \caption{Amplitude of the inertial-frame modes $h^{I}_{2m}$ and errors in the Fourier-domain transfer fonctions $\calT^{2}_{m2}$ for the configuration $--$.

### 2018-02-08 docs/paper_method.tex
Edits to the paper, added option constant-eff, added minerva build.

@@
 
-\SM{[say a word of resampling needed at low and high f, refer to paper PE - discussion section at the end ?]}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Blue corresponds to the orbital delay comb coefficients $c_{n}^{0}$ given by~\eqref{eq:cn0}, and red to the constellation comb $c_{n}^{L}$ for the response~\eqref{eq:GslrL}.
-The result is shown averaged over 100 random orientation angles.}
+The result is shown after averaging over 100 random orientation angles.}
   \label{fig:lisacombextent}
@@
 \ee
-For the constellation response, by contrast, to our knowledge the coefficients $c_{n}^{L}$ do not admit such a simple close-form expression, and they must be computed numerically using the integrand given by~\eqref{eq:GslrL}.
+By contrast, for the constellation response, to our knowledge the coefficients $c_{n}^{L}$ do not admit such a simple close-form expression, and they must be computed numerically using the integrand given by~\eqref{eq:GslrL}.
 Truncating~\eqref{eq:Gdiscretefourier} to a finite order $N$, this computation reduces to a DFT \SM{[acronym]}, and the $c_{n}$ coefficients are given by~\eqref{eq:ykDFT} and~\eqref{eq:ckyk}.
@@
   \centering
-  \includegraphics[width=.48\linewidth]{plots/trig_amp.pdf}
-  \hspace{0.2cm}
-  \includegraphics[width=.48\linewidth]{plots/trig_phase.pdf}
+%  \includegraphics[width=.48\linewidth]{plots/trig_amp.pdf}
+%  \hspace{0.2cm}
+%  \includegraphics[width=.48\linewidth]{plots/trig_phase.pdf}
+  \includegraphics[width=.98\linewidth]{plots/trigopoly_py.pdf}
   \caption{Amplitude (left panel) and phase (right panel), compared to its trigonometric polynomial representation for the three cases listed in Table~\ref{table:precexamples}.
 The vertical lines indicate the merger frequencies.
-\SM{[color for vertical lines at fpeak]} \SM{[add bottom panel with residual ?]}}
+\SM{[color for vertical lines at fpeak]} \SM{[add bottom panel with residual ?]} \SM{[check definition of A]}}
   \label{fig:trigopoly}
@@
 
-\SM{[add appendix with discrete Fourier coeffs]}
+\SM{[say a word of resampling needed at low and high f for LISA, and in general interpolation errors, appendix or refer to paper PE ?]}
+
+\SM{[clean up part on discrete Fourier coeffs]}
 

### 2018-02-08 LISAinference/LISAinference.c

@@ -449,58 +449,58 @@ int main(int argc, char *argv[])
 
 	/* set the MultiNest sampling parameters */
 
-	int mmodal = runParams.mmodal;			/* do mode separation? */
+	int mmodal = runParams.mmodal;		/* do mode separation? */
 
-	int ceff = 0;					/* run in constant efficiency mode? */
+	int ceff = runParams.consteff;		/* run in constant efficiency mode? */
 
 	int nlive = runParams.nlive;			/* number of live points */
 
-	double efr = runParams.eff;			/* set the required efficiency */
+	double efr = runParams.eff;		  	/* set the required efficiency */
 
-	double tol = runParams.tol;			/* tol, defines the stopping criteria */
+	double tol = runParams.tol;			  /* tol, defines the stopping criteria */
 
-	/* int ndim = 9; */				        /*dimensionality (no. of free parameters) */
-	/* int nPar = 9; */					/* total no. of parameters including free & derived parameters */
+	/* int ndim = 9; */				        /* dimensionality (no. of free parameters) */
+	/* int nPar = 9; */					      /* total no. of parameters including free & derived parameters */
 
-	int nClsPar = runParams.nclspar;            	/* no. of parameters to do mode separation on */
+	int nClsPar = runParams.nclspar;  /* no. of parameters to do mode separation on */
 
-	int updInt = 50;				/* after how many iterations feedback is required & the output files should be updated */
-							/* note: posterior files are updated & dumper routine is called after every updInt*10 iterations */
+	int updInt = 50;			           	/* after how many iterations feedback is required & the output files should be updated */
+							                      /* note: posterior files are updated & dumper routine is called after every updInt*10 iterations */
 
-	double Ztol = runParams.ztol;	       		/* all the modes with logZ < Ztol are ignored */
+	double Ztol = runParams.ztol;	    /* all the modes with logZ < Ztol are ignored */
 
-	int maxModes = runParams.maxcls;		/* expected max no. of modes (used only for memory allocation) */
+	int maxModes = runParams.maxcls;	/* expected max no. of modes (used only for memory allocation) */
 
-	int pWrap[ndim];				/* which parameters to have periodic boundary conditions? */
-	for(i = 0; i < ndim; i++) { /* If non-default limiting values have been set for lambda, phase, pol, do not treat them as periodic */
+	int pWrap[ndim];			           	/* which parameters to have periodic boundary conditions? */
[…]

### 2018-02-08 LISAinference/LISAinference_common.c

@@ -31,9 +31,13 @@ void addendum(int argc, char *argv[],LISARunParams *runParams, int *ndim, int *n
   /* Parse commandline to read parameters of injection - copy the number of modes demanded for the injection */
   parse_args_LISA(argc, argv, injectedparams, globalparams, priorParams, runParams, addparams);
   injectedparams->nbmode = globalparams->nbmodeinj;
-
   //int notLISAlike=strstr(argv[0],"LISAlike")==0;
   if(myid == 0 && runParams->writeparams /*&& notLISAlike*/) print_parameters_to_file_LISA(injectedparams, globalparams, priorParams, runParams);
+  if(myid == 0) {
+    printf("Injected params\n");
+    report_LISAParams(injectedparams);
+  }
+
   /* Initialize the data structure for the injection */
   LISAInjectionCAmpPhase* injectedsignalCAmpPhase = NULL;
   LISAInjectionReIm* injectedsignalReIm = NULL;
@@ -51,10 +55,8 @@ void addendum(int argc, char *argv[],LISARunParams *runParams, int *ndim, int *n
   else if(globalparams->tagint==1) {
     LISAGenerateInjectionReIm(injectedparams, globalparams->minf, globalparams->nbptsoverlap, 1, injectedsignalReIm); /* Use here logarithmic sampling as a default */
   }
-  printf("Injected params\n");
-  report_LISAParams(injectedparams);
 
-  /* Define SNR */
+  /* Compute SNR */
   double SNR123, SNR1, SNR2, SNR3;
   if(globalparams->tagint==0) {
     SNR123 = sqrt(injectedsignalCAmpPhase->TDI123ss);
@@ -67,8 +69,7 @@ void addendum(int argc, char *argv[],LISARunParams *runParams, int *ndim, int *n
   }
 
   /* Rescale distance to match SNR */
-  //
-  printf("isnan(priorParams->snr_target) : %d\n", isnan(priorParams->snr_target));
+  //printf("isnan(priorParams->snr_target) : %d\n", isnan(priorParams->snr_target));
   if (!isnan(priorParams->snr_target)) {
     printf("SNR=%g\n",SNR123);
     if (myid == 0) printf("Rescaling the distance to obtain a network SNR of %g\n", priorParams->snr_target);
@@ -91,8 +92,10 @@ void addendum(int argc, char *argv[],LISARunParams *runParams, int *ndim, int *n
[…]

### 2018-02-08 LISAinference/LISAutils.c

@@ -7,6 +7,7 @@ LISAParams* injectedparams = NULL;
 LISAGlobalParams* globalparams = NULL;
 LISAPrior* priorParams = NULL;
 LISAAddParams* addparams = NULL;
+double logZdata = 0.;
 SimpleLikelihoodPrecomputedValues* simplelikelihoodinjvals = NULL;
 
 /***************** Pasring string to choose what masses set to sample for *****************/
@@ -225,6 +226,7 @@ Syntax: --PARAM-min\n\
 --------------------------------------------------\n\
  --eff                 Target efficiency of sampling (default=0.1)\n\
  --tol                 Tolerance for evidence calculation convergence (default=0.5)\n\
+ --consteff            Option to use constant efficiency mode\n\
  --nlive               Number of live points for sampling (default=1000)\n\
  --bambi               Use BAMBI's neural network logL learning (no option, default off)\n\
  --resume              Resume from a previous run (no option, default off)\n\
@@ -337,6 +339,7 @@ Syntax: --PARAM-min\n\
     /* set default values for the run settings */
     run->eff = 0.1;
     run->tol = 0.5;
+    run->consteff = 0;
     run->nlive = 1000;
     run->writeparams = 1;
     strcpy(run->outroot, "chains/LISAinference_");
@@ -527,6 +530,8 @@ Syntax: --PARAM-min\n\
             run->eff = atof(argv[++i]);
         } else if (strcmp(argv[i], "--tol") == 0) {
             run->tol = atof(argv[++i]);
+        } else if (strcmp(argv[i], "--consteff") == 0) {
+            run->consteff = 1;
         } else if (strcmp(argv[i], "--nlive") == 0) {
             run->nlive = atoi(argv[++i]);
         } else if (strcmp(argv[i], "--bambi") == 0) {
@@ -749,6 +754,7 @@ int print_parameters_to_file_LISA(
   fprintf(f, "-----------------------------------------------\n");
   fprintf(f, "eff:            %g\n", run->eff);
   fprintf(f, "tol:            %g\n", run->tol);
+  fprintf(f, "consteff:       %d\n", run->consteff);
   fprintf(f, "nlive:          %d\n", run->nlive);
   fprintf(f, "bambi:          %d\n", run->bambi);
   fprintf(f, "resume:         %d\n", run->resume);
@@ -779,7 +785,6 @@ int print_rescaleddist_to_file_LISA(
[…]

### 2018-02-08 LISAinference/LISAutils.h

@@ -169,6 +169,7 @@ typedef struct tagLISAPrior {
 typedef struct tagLISARunParams {
   double eff;                /* target efficiency (default 0.1) */
   double tol;                /* logZ tolerance (default 0.5) */
+  int    consteff;           /* constant efficiency mode (default 0) */
   int    nlive;              /* number of live points (default 1000) */
   int    writeparams;        /* Write params - if 1, write run parameters to file (default 1) */
   char   outroot[200];       /* output root (default "chains/LISAinference_") */
@@ -231,9 +232,9 @@ int print_rescaleddist_to_file_LISA(
   LISAGlobalParams* globalparams,
   LISAPrior* prior,
   LISARunParams* run);
+int print_snrlogZ_to_file_LISA(LISARunParams* run, double SNR, double logZ);
 /* Function printing injection/signal parameters to stdout */
-void report_LISAParams(
-  LISAParams* params);
+void report_LISAParams(LISAParams* params);
 
 /* Initialization and clean-up for LISASignal structures */
 void LISASignalCAmpPhase_Cleanup(LISASignalCAmpPhase* signal);
@@ -334,7 +335,7 @@ extern LISAParams* injectedparams;
 extern LISAGlobalParams* globalparams;
 extern LISAPrior* priorParams;
 extern LISAAddParams* addparams;
-double logZdata;
+extern double logZdata; /* TODO: not used */
 extern SimpleLikelihoodPrecomputedValues* simplelikelihoodinjvals;
 
 #if 0

### 2018-02-13 docs/paper_method.tex
Extensive edits in section III to aid clarity of the presentation.

@@
 A direct computation of the generalized convolution~\eqref{eq:FDkernel} using~\eqref{eq:defG} will generally be computationally demanding, but we can exploit the separation of timescales in the problem to seek an accurate but efficient approximation.
-A first argument would be to notice that the Fourier transform $\tilde{G}(f,f')$ should have a compact support, limited to $f' \in [-f_{\rm max}, f_{\rm max}]$ with $f_{\rm max}$ a maximal frequency for the modulation, roughly the inverse of its characteristic timescale.
+\jgb{We will explore approximations which allow the integral in~\eqref{eq:FDkernel} to be computed, so that the signal may be expressed the form of~\eqref{eq:deftransfer}, with the modulation and delay reduced to a multiplicative transfer function $\mathcal T(f)$.}
+
+\jgb{For example, under appropriate conditions with slowly varying modulations and delays, we can expect that} the Fourier transform $\tilde{G}(f,f')$ should have a compact support, limited to $f' \in [-f_{\rm max}, f_{\rm max}]$ with $f_{\rm max}$ a maximal frequency for the modulation, roughly the inverse of its characteristic timescale.
 This makes the convolution integral~\eqref{eq:FDkernel} localized in frequency, as the waveform is to be evaluated only at frequencies close to $f$.
 We can then approximate $\tilde{h}(f-f')$ by $\tilde{h}(f)$, with a simple Taylor expansion for a perturbative treatment of the difference between the two.
-Doing so, we will recover at leading order a locality in time: the modulation is to be evaluated at a time $\tf$, which will be equivalent to $\tfSPA$ for inspiral signals.
-
-Now, there are two limitations to this straightforward argument.
+Doing so, we will recover at leading order a locality in time: the response is approximately reduced to an evaluation of the modulation and delay at a representative signal-dependent time $\tf$, which will be equivalent to $\tfSPA$ for inspiral signals.
+
+Now consider the limitations to this straightforward argument.
 The first is that we did not yet specify what $f_{\rm max}$ should be compared with, in order for the approximation to work.
@@
 
-In presence of delays, the above is also complicated by the additional delay phases.
-Depending on the frequency $f$, this can lead $G(f,t)$ to have faster variations that its nominal timescale ($1\mathrm{yr}$ for LISA).
-As we will see in Sec.~\ref{} \SM{[R]}, to decide of the applicability of this approach we will not only need to compare timescales, but also take into account dimensionless factors of the form $2\pi f d$.
+In presence of delays, the above is also complicated by the additional delay phases in the signal Fourier transform.
+Depending on the frequency $f$, this can lead $G(f,t)$ to have faster variations than its nominal timescale ($1\mathrm{yr}$ for LISA).
+As we will see in Sec.~\ref{} \SM{[R]}, \jgb{to derive useful approximations we will not only need to compare modulation timescales, but also to constrain dimensionless factors of the form $2\pi f d$.}
+
+\jgb{In the next subsections we will examine the approximations necessary to allow a local-in-frequency response treatment, and consider low-order corrections which can improve accuracy in cases where the approriate conditions are more marginally realized.}
 
@@
 As a first step, we perform a formal leading-order expansion of the signal $\tilde{h}(f-f')$ around $f$.
-We use the amplitude/phase decomposition~\eqref{eq:defAPsi}, treat the Fourier-domain amplitude $A$ as a constant, expand the Fourier-domain phase $\Psi$ to the first order, and discard the $f'$ dependence in the first argument of $\tilde{G}(f-f', f')$, to obtain
-\be
-	\tilde{h}(f-f') \simeq A(f) \exp\left[ -i\left( \Psi(f) - f' \frac{\ud \Psi}{\ud f} \right) \right] \,,
-\ee
-Here and in the rest of the paper, such derivatives with respect to the frequency will always be evaluated at $f$.
-
-Plugging this relation in~\eqref{eq:FDkernel}, we obtain
+We use the amplitude/phase decomposition~\eqref{eq:defAPsi}, treat the Fourier-domain amplitude $A$ as a constant, expand the Fourier-domain phase $\Psi$ to the first order, and discard the $f'$ dependence in the first argument of $\tilde{G}(f-f', f')$.
+
+For the signal, then, we have
+\be
+	\tilde{h}(f-f') \simeq A(f) \exp\left[ -i\left( \Psi(f) - f' \frac{\ud \Psi}{\ud f} \right) \right] \,,\label{eq:leadingorderwf}
+\ee
+where here and in the rest of the paper, such derivatives with respect to the frequency will always be evaluated at $f$. Plugging this relation into~\eqref{eq:FDkernel}, we obtain
 \begin{align}
 	\tilde{s}(f) &\simeq \tilde{h}(f) \int \ud f' \, \exp\left[ i f' \frac{\ud \Psi}{\ud f} \right] \tilde{G}(f,f') \nn\\
-	&= \tilde{h}(f) G\left( f, -\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) \,,
-\end{align}
-which motivates the introduction of an effective time, seen as a function of frequency,
+	&= \tilde{h}(f) G\left( f, -\frac{1}{2\pi} \frac{\ud \Psi}{\ud f} \right) \,,\label{eq:leadingorderresponse}
+\end{align}
+\jgb{which we can think of as a local evaluation of the kernel function $G(f,t)$ at a frequency-dependent effective time}
 \be\label{eq:deftf}
@@
 Fig.~\ref{fig:tf} shows the behaviour of the time function $\tf$ around merger for six example waveforms, for mass ratios $q=1$ and $q=8$ and for aligned spin components $\chi=0.95,0.,-0.95$.
-In particular, one should note that $\tf$ is not monotonously increasing with frequency anymore when reaching in the high-frequency part of the waveform, corresponding to the ringdown.
-Although $t_{f}$ is always a well-defined function of $f$, this non-monotonicity would forbid an unambiguous definition of a reciprocal frequency-of-time function $f(t)$.
-
-Using this notation, we have simply
+In particular, one should note that $\tf$ is not monotonically increasing with frequency anymore after reaching in the high-frequency part of the waveform, corresponding to the ringdown.
+\jgb{As long as the Fourier phase is differentiable $t_{f}$ is a well-defined function of $f$.
+While its non-monotonicity would forbid an unambiguous definition of a reciprocal frequency-of-time function $f(t)$, like that in the SPA, no such function will be needed in our treatment.}
+
+\jgb{With \eqref{eq:leadingorderresponse} we have brought the modulated and delayed signal in to the form~\eqref{eq:deftransfer} with transfer function}
 \be\label{eq:transferlocal}
@@
 
-We now extend our approach beyond the leading-order approximation. A Taylor-expansion of~\eqref{eq:FDkernel} in the variable $f'$ yields:
+\jgb{If the effective width of of the kernel function $G(f,f’)$ is not quite negligible compared to the scale of significant variations with $f$ it can be useful to extend our approach beyond the leading-order approximation.
+With the waveform represented in phase-and-amplitude form~\eqref{eq:A9}, the elements of~\eqref{eq:FDkernel} may be Taylor expanded in the variable $f'$:
 \begin{subequations}\label{eq:expandfprime}
@@
 	\Psi(f-f') &= \Psi(f) + 2\pi f' \tf + \sum\limits_{p\geq 2}^{N} \frac{(-1)^{p}}{p!} {f'}^{p} \frac{\ud^{p} \Psi}{\ud f^{p}} \,, \label{eq:expandPsi}\\
-	A(f-f') &= A(f) \sum\limits_{q\geq 0} \frac{(-1)^{q}}{q!} {f'}^{q} \frac{1}{A}\frac{\ud^{q} A}{\ud f^{q}} \,, \label{eq:expandA}\\
-	\tilde{G}(f-f', f') &= \sum\limits_{r\geq 0} \frac{(-1)^{r}}{r!} {f'}^{r} \frac{\partial^{r} }{\partial f^{r}}  \tilde{G}(f,f') \label{eq:expandG} \,,
+	A(f-f') &= A(f)+A(f) \sum\limits_{q\geq 1} \frac{(-1)^{q}}{q!} {f'}^{q} \frac{1}{A}\frac{\ud^{q} A}{\ud f^{q}} \,, \label{eq:expandA}\\
+	\tilde{G}(f-f', f') &=\tilde G(f,f')+ \sum\limits_{r\geq 1} \frac{(-1)^{r}}{r!} {f'}^{r} \frac{\partial^{r} }{\partial f^{r}}  \tilde{G}(f,f') \label{eq:expandG} \,,
 \end{align}
 \end{subequations}
-were we used the definition of $t_{f}$ introduced in~\eqref{eq:deftf}.
-In the following, we will only keep the first few terms in each of these expansions, but we keep these expressions in terms of formal series for now.
-Notice that we expand $\tilde{G}(f-f',f')$ in $f'$ only in its first argument, and that we can interchange the $f$-derivatives of $G$ with the Fourier transform operation.
-
-Now, if we further expand the exponential of the phase to obtain a pure $f'$-expansion we can convert the resulting power series in $f'$ into a Taylor series with time derivatives, by applying the formal derivative rule
+using the definition of $t_{f}$ introduced in~\eqref{eq:deftf}.}
+
+\jgb{The leading order transfer function~\eqref{eq:transferlocal} is obtained by leaving off all the terms in the sums.
+In the following, we will consider the resulting transfer functions when keeping some of the next few terms in each of these expansions.}
+Notice that we expand $\tilde{G}(f-f',f')$ in $f'$ only in its first argument, and that we can commute the $f$-derivatives of $G$ with the Fourier transform operation.
+
+\jgb{In some cases, we will also expand the exponential of the phase
+\be
+\exp\left[ -i\Psi(f) \right] \simeq 1+\sum_{j\geq1}\frac{(-i\Psi(f))^j}{j!}
+\label{eq:expandexp}
+\ee
+to obtain a pure $f'$-expansion. The resulting power series in $f'$ can then be recast as a temporal Taylor series}, by applying the formal derivative rule
 \allowdisplaybreaks
@@
 \ee
-The result is a rather cumbersome expression with multiple sums, that we will not use directly.
-Instead, it will be more instructive to separate the different expansions in~\eqref{eq:expandfprime}.
-Only in Sec.~\ref{subsec:executivesummary} will we come back to combining our results together.
+The fully expanded result is a rather cumbersome expression with multiple sums, that we will not use directly.
+Instead, it will be more instructive to separately consider the different expansions in~\eqref{eq:expandfprime}.
+Later, in Sec.~\ref{subsec:executivesummary} will we come back to combining our results together.
 
@@
 We will find that the third and higher derivatives of the phase are always negligible for our purposes, and we will ignore them.
-Keeping only the first term of the series in~\eqref{eq:expandPsi}, corresponding to the second derivative of $\Psi$, and expanding the exponential of the phase, we obtain straightforwardly
+Keeping only the first term of the \jgb{sum in~\eqref{eq:expandPsi}, corresponding to the second derivative of $\Psi$, with just the leading terms from \eqref{eq:expandA} and \eqref{eq:expandG}, and expanding the phase exponential \eqref{eq:expandexp} so that the result can be cast as a Taylor series in time} , we obtain straightforwardly
 \be\label{eq:resulttaylorPsi}
@@
 
-Similarly, the expansion of the Fourier-domain amplitude~\eqref{eq:expandA} gives
+Similarly, applying the expansion of the Fourier-domain amplitude~\eqref{eq:expandA} while preserving only the leading order terms in \eqref{eq:expandPhi} and \eqref{eq:expandG} gives
 \be\label{eq:resulttaylorA}
-	\calT_{\rm amp}(f) = \sum\limits_{p\geq 0} \frac{1}{(2i\pi)^{p}p!} \frac{1}{A} \frac{\ud^{p} A}{\ud f ^{p}}  \left( \frac{\partial^{p} }{\partial t^{p}} G \right)(f,\tf) \,,
-\ee
-while expanding only the frequency-dependence of $G$ as in~\eqref{eq:expandG} yields
+	\calT_{\rm amp}(f) = \sum\limits_{p\geq 0} \frac{1}{(2i\pi)^{p}p!} \frac{1}{A} \frac{\ud^{p} A}{\ud f ^{p}}  \left( \frac{\partial^{p} }{\partial t^{p}} G \right)(f,\tf) \,.
+\ee
+
+Lastly, expanding only the frequency-dependence of $G$ as in~\eqref{eq:expandG} yields
 \be\label{eq:resulttaylordelay}
@@
 \ee
+\jgb{[It seems like it might be better to label this version of the transfer function as ``kernel'' rather than ``delay'' since $G$ here is general. Do we ultimately use this at all?]}
 which interestingly looks like a Taylor expansion of $G$, but this time with joint derivatives in frequency and time.
-This last expansion, taken separately from the other corrections, is signal-independent as it only depends on the function $G$ and not on $A$, $\Psi$.
+This last expansion, taken separately from the other corrections, is signal-independent as it only depends on the kernel function $G$ and not on $A$, $\Psi$.
 
@@
 
-Considering~\eqref{eq:resulttaylorPsi} it is natural to define a new timescale, as a function of frequency, as
+\jgb{When considering the impact on the transfer function from including the next higher-order phase term, that is the difference between~\eqref{eq:resulttaylorPsi} and~\eqref{eq:transferlocal}, it is natural to define a new timescale, as a function of frequency,
 \be\label{eq:defTf}
@@
 \ee
-The interpretation of this timescale for the inspiralling part of the signal, when the SPA applies, will be discussed below. We can use this notation to rewrite~\eqref{eq:resulttaylorPsi} as
+If the correction is small, this timescale should be small compared to the time-scale of variations in the modulation and delay encoded in $G(f,t)$.}
+We can use this notation to rewrite~\eqref{eq:resulttaylorPsi} as
 \begin{align}\label{eq:resultdffPsiTf}
@@
 
-We can obtain a straighforward physical interpretation of this timescale $\Tf$ by considering inspiral signals for which the SPA is valid. If we define
+We can obtain a straighforward physical interpretation of this timescale $\Tf$ by considering inspiral signals for which the SPA is valid.
+\jgb{In that case
+\be
+	\left(\Tf^{\rm SPA} \right)^{2} = -\frac{1}{4\pi^{2}}  \frac{\ud^{2} \Psi_{\rm SPA}}{\ud f^{2}} \,,
+\ee
+where $\ud^{2}\Psi/\ud f^{2} < 0$ in the SPA with our sign conventions.
+Then, taking two derivatives of~\eqref{eq:PsiSPA}, we find
 \be\label{eq:TfSPA}
-	\Tf^{\rm SPA} \equiv \frac{1}{\sqrt{2\dot{\omega}(\tfSPA)}} \,,
-\ee
-we have, by taking two derivatives of~\eqref{eq:PsiSPA},
-\be
-	\left(\Tf^{\rm SPA} \right)^{2} = -\frac{1}{4\pi^{2}}  \frac{\ud^{2} \Psi_{\rm SPA}}{\ud f^{2}} \,,
-\ee
-where $\ud^{2}\Psi/\ud f^{2} < 0$ in the SPA with our sign conventions.
-The quantity $\Tf^{\rm SPA}$ is clearly a radiation-reaction timescale: the shorter this timescale, the faster the binary chirps to higher frequencies on its quasi-circular inspiral.
-Comparing with~\eqref{eq:defTf} above, we see that our generalized time scale $\Tf$ also reduces to $\Tf^{\rm SPA}$ when the SPA applies.
-
-However, in the same way as our definition~\eqref{eq:deftf} for the time-of-frequency function $t_{f}$ generalizes the SPA definition~\eqref{eq:deftfSPA}, the definition~\eqref{eq:defTf} only refers to the phase of the Fourier-domain signal, does not require to introduce a time-domain frequency like $\omega$, and extends naturally to the merger-ringdown part of the signal.
+	\Tf^{\rm SPA} \equiv \frac{1}{\sqrt{2\dot{\omega}(\tfSPA)}} \,.
+\ee
+Thus, when the SPA applies, $\Tf$ corresponds to} the radiation-reaction timescale: the shorter this timescale, the faster the binary chirps to higher frequencies on its quasi-circular inspiral.
+
+However, in the same way that our definition~\eqref{eq:deftf} for the time-of-frequency function $t_{f}$ generalizes the SPA definition~\eqref{eq:deftfSPA}, the definition~\eqref{eq:defTf} only refers to the phase of the Fourier-domain signal and does not require introducing a time-domain frequency like $\omega$.
+This defintion thus extends naturally to the merger-ringdown part of the signal.
 In this part of the signal, its physical interpretation as the timescale of radiation reaction is obscured, and the second derivative $\ud^{2}\Psi/\ud f^{2}$ can go through zero and change sign, as shown in Fig.~\ref{fig:TfTA}.
@@
 
-The expansion in amplitude corrections~\eqref{eq:resulttaylorA} also leads to the natural introduction of a set of other timescales related to the successive derivatives of the amplitude.
+\jgb{Next consider the impact on the transfer function~\eqref{eq:resulttaylorA} from amplitude corrections beyond leading order.}  This series also leads to the natural introduction of a set of another set of timescales related to the successive derivatives of the amplitude.
 In an analogous manner to the definition~\eqref{eq:defTf} of the timescale $\Tf$, we can define
@@
 \ee
-where we included an absolute value to accomodate the possible sign changes in the right-hand side. With this notation, \eqref{eq:resulttaylorPsi} becomes simply
-\be\label{eq:resulttaylorA}
+where we included an absolute value to accomodate the possible sign changes in the right-hand side. With this notation, \eqref{eq:resulttaylorA} becomes simply
+\be\label{eq:resultATA}
 	\calT_{\rm amp}(f) = \sum\limits_{p\geq 0} \frac{1}{p!} (T_{Ap})^{p}  \left( \partial_{t}^{p} G \right) (f,\tf) \,,
@@
 
-By contrast, the expansion~\eqref{eq:resulttaylordelay} is signal-independent and does not lead to the introduction of new timescales as the coupled time- and frequency derivatives are dimensionless.
-We will see however in Sec.~\ref{subsec:lisafom} that treating delays requires to take into account dimensionless factors of the type $2\pi f d$.
-
-We can obtain useful estimates for these timescales from the Newtonian expressions~\eqref{eq:SPAN}. This gives for the Newtonian radiation-reaction and amplitude timescales:
+\jgb{By contrast, the impact of higher-order terms in the kernel function on the transfer function}~\eqref{eq:resulttaylordelay} is signal-independent.
+It does not lead to the introduction of new timescales since the coupled time and frequency derivatives are dimensionless.
+We will see however in Sec.~\ref{subsec:lisafom} that treating delays \jgb{does require limiting} dimensionless factors of the type $2\pi f d$.
+
+We can obtain useful estimates for these timescales from the leading order post-Newtonian expressions~\eqref{eq:SPAN}. The leading-order radiation-reaction and amplitude timescales are:
 \begin{subequations}\label{eq:timescalesN}
@@
 In the inspiral, they follow the scalings~\eqref{eq:timescalesN} and $\Tf$ is much larger than the other two.
-For frequencies above the merger frequency, $\Tf$ can go through zero and the amplitude-related timescales become comparable or larger.
+For frequencies above the merger frequency, $\Tf$ can go through zero while the amplitude-related timescales become comparable or larger.
 
@@
 
-If we restrict to the case of a pure modulation, $d=0$ and $G(f,t) = F(t)$, we can relate our result~\eqref{eq:resultdffPsiTf} to the approach of Ref.~\cite{KCY14}, where the authors extend the SPA in a formalism called the Shifted Uniform Asymptotic expansion (SUA) \SM{[look for other reference beyond KCY]}.
+If we restrict to the case of a pure undelayed modulation, $d=0$ and $G(f,t) = F(t)$, we can relate our result~\eqref{eq:resultdffPsiTf} \jgb{for the transfer function including up to quadratic phase terms, with treatment} of Ref.~\cite{KCY14}, where the authors extend the SPA in a formalism called the Shifted Uniform Asymptotic expansion (SUA) \SM{[look for other reference beyond KCY]}.
 %The derivation includes an expansion of the modulation in Bessel functions, a study of the displacement of the stationary phase point induced by the precession, and a resummation of the result in the form of time derivatives.
@@
 \ee
-provided that the complex coefficients $a_{N,k}^{\epsilon}$ are a solution of the $N+1$-dimensional linear system~\cite{KCY14}
+where the complex coefficients $a_{N,k}^{\epsilon}$ are a solution of the $N+1$-dimensional linear system~\cite{KCY14}
 \be\label{eq:stencilsystem}
@@
 
+\jgb{In practice, our quadratic-in-phase treatment will apply this formalism with [something about the range of N?]. }
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-We can give an alternative interpretation of the expansion~\eqref{eq:resultdffPsiTf} and of Sec.~\ref{subsec:resumquadphase} in terms of the Fresnel transform.
+We can give an alternative interpretation of the \jgb{quadratic-phase} expansion~\eqref{eq:resultdffPsiTf} and of Sec.~\ref{subsec:resumquadphase} in terms of the Fresnel transform.
 For simplicity of notation, here we keep to the case of a pure modulation $F$ with no delays, $d=0$.
 We will reintroduce the delays below in Sec.~\ref{subsec:delays}.
-To obtain~\eqref{eq:resultdffPsiTf}, we expanded fully the exponential of the phase as a Taylor series.
-If it is not expanded, however, we have
+To obtain~\eqref{eq:resultdffPsiTf}, we expanded the phase exponential using~\eqref{eq:expandexp}.
+\jgb{Without this expansion, though}, we have
 \be\label{eq:integralquadphase}
@@
 where we recall that $\epsilon = -\mathrm{sgn}(\ud ^{2} \Psi/\ud f^{2})$.
-The integral on $f'$ is a simple complex Gaussian integral, and results in a complex Gaussian integral over time where we recognize a Fresnel transform~\cite{} \SM{[C]} of the function $F$.
-We introduce for the Fresnel transform the notation
+The integral over $f'$ is a simple complex Gaussian integral \jgb{which can be carried out explicitly.
+The result is a complex Gaussian integral over time which we recognize as a Fresnel transform~\cite{} \SM{[C]} of the function $F$.
+For the Fresnel transform, we introduce the notation}
 \be\label{eq:defFresnel}
@@
 Since we have seen that $\Tf$ can be interpreted as the radiation reaction timescale in the SPA regime, this means that a faster-chirping signal ($\Tf$ small) will have a Fresnel transform that is more focused, whereas a slower-chirping signal ($\Tf$ large) will have a Fresnel transform that is more extended.
-In turn, this is to be put in relation with how fast the function $F(t)$ in the integrand is varying.
+\jgb{The Fresnel width must then be compared} with how fast the function $F(t)$ in the integrand is varying.
 In Sec.~\ref{subsec:executivesummary}, we will build an estimate for the magnitude of these phase corrections by comparing of the radiation-reaction timescale to the timescale of variation of the modulation.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Treatment of the delays}
+\jgb{Going forward, the Fresnel transform provides a convenient way to non-perturbatively express the effect of the quadratic phase term on the transfer function, but practical compuations will rely on the approximations $\calF_{T, \epsilon}^{N}[F] (t)$ in \eqref{eq:stencilfresnel}.[Check that I have this right.
+]}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Response treatment including delays}
 \label{subsec:delays}
@@
 In the case of a LISA-type detector response, the presence of the delay $d(t)$ is responsible for the frequency-dependence of the kernel $G$ introduced in~\eqref{eq:defG}.
-Here we give a direct treatment of the delays which is an alternative to the mere Taylor expansion of~\eqref{eq:resulttaylordelay}.
-As a first illustration, we leave aside amplitude and phase corrections as given in~\eqref{eq:resulttaylorA} and~\eqref{eq:resulttaylorPsi}.
-The response then depends on the signal only by the time-to-frequency correspondence and reads
+\jgb{In practice we will not use the Taylor expansion of the kernel function in~\eqref{eq:resulttaylordelay}, and will instead expand the delay itself as a slowly varying constant.}
+
+\jgb{As a first step, we consider the transfer function when we keep only the leading order waveform~\eqref{eq:leadingorderwf}, leaving off additional amplitude and phase terms in~\eqref{eq:resulttaylorA} and~\eqref{eq:resulttaylorPsi}, but treat the kernel function nonperturbatively.} The response then depends on the signal only by the time-to-frequency correspondence and reads
 \begin{equation}
@@
 This motivates a change of variable to the delayed time function $t_{d}:t \mapsto t+d(t)$.
-We also define the reciprocal function $t_{d}^{-1}$, and a modified time-to-frequency correspondence $\tfd = t_{d}^{-1}(\tf)$, defined implicitly by
+\jgb{Assuming the delay does not vary too quickly, we can} also define the reciprocal function $t_{d}^{-1}$, and a modified time-to-frequency correspondence $\tfd = t_{d}^{-1}(\tf)$, defined implicitly by
 \be
@@
 \end{align}
+\jgb{The difference with~\eqref{eq:transferlocal}, the extra factor comprising the denominator, shows the effect of using here the non-perturbative kernel function rather than just the leading order term in~\eqref{eq:expandG}.}
 
 For the proposed LISA configuration~\cite{LISA17}, we have for the orbital delays the scaling $d_{0}\sim R/c \simeq 500s$, and for the constellation delays the scaling $d_{c}\sim L/c \simeq 8s$ (ignoring the dependence on angular factors).
-Since the motion of the constellation is periodic with a period of one year, with a frequency $\Omega_{0} \simeq 2 \times 10^{-7}\mathrm{rad}.s^{-1}$, we have $\dot{d}_{0} \sim \Omega_{0} R/c \simeq 10^{-4}$ and $\dot{d}_{c} \sim \Omega_{0} L/c \simeq 1.7 \times 10^{-6}$.
+Since the motion of the constellation is anually periodic, with a frequency $\Omega_{0} \simeq 2 \times 10^{-7}\mathrm{rad}.s^{-1}$, we have $\dot{d}_{0} \sim \Omega_{0} R/c \simeq 10^{-4}$ and $\dot{d}_{c} \sim \Omega_{0} L/c \simeq 1.7 \times 10^{-6}$.
 The smallness of the dimensionless quantity $\dot{d} \ll 1$ (and of its subsequent derivatives) will allow us to treat it perturbatively with a very good approximation, and shows also that the function $t_{d}$ is univalued and that there is no ambiguity in defining the reciprocal $t_{d}^{-1}$.
@@
 \end{align}
-where we used a change of variable $\tau = t_{d}(t)$.
-We see that the result can again be expressed as a Fresnel transform.
-When considering amplitude corrections as well, as in~\eqref{eq:resulttaylorA}, additional powers of $f'$ can be translated as time derivatives with respect to the variable $\tau$ after performing the change of variables.
+where we used a change of variable $\tau = t_{d}(t)$. We see that the result can again be expressed as a Fresnel transform.
+
+Finally, when considering amplitude corrections as well, as in~\eqref{eq:resulttaylorA}, additional powers of $f'$ can be translated as time derivatives with respect to the variable $\tau$ after performing the change of variables.
 This produces the final result:
@@
 \ee
+\jgb{This expression provides a summary of the perturbative treatment we will apply in later sections, with the additional understanding that the Frensel transform is treated approximately using \eqref{eq:fresnelstencil}.[Check this]}
 \end{widetext}
 
-\subsection{Executive summary of the formalism}\label{subsec:executivesummary}
+\subsection{Summary of the formalism}\label{subsec:executivesummary}
 
@@
 We will use symbols of the form $\{N,A,d\}$ to indicate the order of the stencil in~\eqref{eq:stencilfresnel}, the maximal order of the amplitude correction included, and the inclusion or not of the delay corrections.
+\jgb{[The meaning of this notation is unclear.
+Some more work may be needed in this subsection.]}
 

### 2018-02-14 docs/paper_method.tex
Significant edits of the Intro and moderate edits of Sec II.

@@
 
-With the recent detections of gravitational-wave signals emitted by the coalescences of binary black holes, announced by the LVC collaboration during the first and second obervation runs of the two LIGO instruments~\cite{LIGO-theevent-2016,LIGO-christmasevent-2016,LIGO-O1BBH-2016}\jgb{[update refs]}, gravitational-wave astronomy has entered its observational era.
-With the conclusion of LIGO's second observation run, and the recent extension of the ground-based detectors network with Virgo~\cite{Virgo} ( and later also KAGRA~\cite{KAGRA} and LIGO-India~\cite{INDIGO}), many more detections from compact object coalescences are to come, with binary black holes being expected to be the dominant sources~\cite{} \SM{[C]}.
-
-Moreover, the European Space Agency has issued a mission call \SM{[update]} for its L3 space mission with a tentative launch around 2034, adressing the science theme ``The Gravitational Universe'', possibly opening the way for a space-borne observatory based on the concept of the Laser Interferometer Space Antenna (LISA)~\cite{LISA09,LISA17}.
-Among other sources of gravitational waves, LISA will be able to detect and characterize comparable-mass binary black hole coalescences to cosmological distances over a wide range of masses, from high-redshift observations of supermassive black hole binaries with $M\sim 10^{7} \Msol$ down to the observation of LIGO-type sources with $M\sim 10^{2} \Msol$~\cite{Sesana16}.
+\jgb{With the unprecedented recent gravitational-wave detections of coalescencing binary black holes and binary neutron stars, announced by the LIGO-Virgo collaboration~\cite{LIGO-theevent-2016,LIGO-christmasevent-2016,LIGO-O1BBH-2016}[update refs], gravitational-wave astronomy has entered its observational era.
+As LIGO prepares for even more sensitive observation runs, and with the recent expansion of the ground-based detectors network with Virgo~\cite{Virgo} ( and eventually also KAGRA~\cite{KAGRA} and LIGO-India~\cite{INDIGO}), observations of such compact object coalescences are expected at ever-increasing rate, with binary black holes likely to be the dominant sources~\cite{} \SM{[C]}.}
+
+  \jgb{Moreover, the European Space Agency has recently selected the Laser Interferometer Space Antenna (LISA)\cite{LISA17} to realize the ``Gravitational Universe'' science theme\cite{GravitationalUniverse2013} as the 3rd large space mission of its Cosmic Vision program, with a tentative launch around 2034.
+LISA will be able to detect and characterize, among several important gravitational wave source targets, comparable-mass binary black hole coalescences from cosmological distances over a wide range of masses.
+These will range from high-redshift observations of supermassive black hole binaries with $M\sim 10^{7} \Msol$ down to the observation of LIGO-type sources with $M\sim 10^{2} \Msol$~\cite{Sesana16}.}
 
   Data analysis for gravitational-wave observations of compact binary coalescences require accurate models (or templates) for the signals, both for ensuring efficient detections of signals that may be buried in instrumental noise, and to extract the physical parameters of the source in a subsequent analysis.
-The the greater sensitivity of LISA and other future instruments will require further increases in the accuracy of computationally efficient signal templates.
-This motivates an important and ongoing effort to improve gravitational wave templates.
+Bayesian analysis for parameter estimation of gravitational-wave signals, as was performed for the LIGO detections~\cite{LIGO-theeventPE-2016,LIGO-O1BBH-2016}, may require millions of evaluations of the likelihood function to sample the posterior probability distribution.
+\SM{[objective: accuracy, not necessary necessary for LIGO now but in the future yes]} The greater sensitivity of LISA and other future instruments will require further increases in the accuracy and computational efficiency of signal templates.
+
+  %This motivates an important and ongoing effort to improve gravitational wave templates.
 Two important focus areas for template improvements are need to cover all phases of the evolution of the system, including the inspiral, merger and ringdown (IMR), which became possible thanks to the breakthrough of numerical relativity~\cite{Pretorius05,Baker+06,Campanelli+06}, and the inclusion of the effects of the spins of the compact objects, that can lead to a precession of the orbital plane~\cite{ACST94,K95}.
 Besides accuracy, the computational efficiency of template generation is crucial.
-Bayesian analysis for parameter estimation of gravitational-wave signals, as was performed for the LIGO detections~\cite{LIGO-theeventPE-2016,LIGO-O1BBH-2016}, may require millions of evaluations of the likelihood function to sample the posterior probability distribution.
-\SM{[objective: accuracy, not necessary necessary for LIGO now but in the future yes]}
-
-The GW community is making progress assembling higher-fidelity tools for these coming challenges.
+
+\jgb{The GW community is making progress assembling higher-fidelity tools for these coming challenges.
 State-of-the-art IMR templates combine information from the perturbative results of post-Newtonian theory covering the inspiral (see e.g.~\cite{BlanchetLiving}) and from numerical relativity simulations covering the end of the inspiral and the merger-ringdown phase (see e.g.~\cite{Pfeiffer12}).
-The Effective-One-Body (EOB) approach~\cite{BD99} does so by incorporating a resummation of post-Newtonian results and a calibration to numerical relativity, and has been extended to include both spins of the two bodies~\cite{Taracchini+13, Pan+13, Bohe+16}.
-Phenomenological templates extend post-Newtonian models by postulating an analytic ansatz for the Fourier-domain amplitude and phase, fitting coefficients against numerical relativity waveforms~\cite{Husa+15, Khan+15}, and can include a simplified model for precession~\cite{Hannam+13}.
-The second type of models produce directly the amplitude and the phase of the Fourier-domain waveform components in the form of a piecewise analytic function.
-EOB waveforms, by contrast, require integration of ordinary differential equations in the time-domain, which can be time-consuming.
-For the non-precessing case, Reduced Order Models (ROM), also called surrogate models, have been developed to speed up considerably the waveform generation, without losing accuracy~\cite{Field+13, Puerrer14, Bohe+16}.
-Both approaches are therefore able to efficiently produce IMR Fourier-domain waveforms.
-An important aspect is that the resulting waveforms can be represented by an amplitude and phase for each mode, with only a few hundred samples~\cite{Puerrer14}.
+Approaches to template construction include phenomenological templates postulating an analytic ansatz for the Fourier-domain amplitude and phase~\cite{Husa+15, Khan+15,Hannam+13}, and the more physically motived Effective-One-Body (EOB) approach~\cite{BD99,Taracchini+13, Pan+13, Bohe+16}.
+Where needed, Reduced Order Models (ROM), also called surrogate models, have been developed to considerably speed up waveform generation, without losing accuracy~\cite{Field+13, Puerrer14, Bohe+16}.
+These tools provide efficient non-precessing IMR Fourier-domain waveforms.
+Importantly for this work, the resulting waveforms can be represented by an amplitude and phase for each mode, with only a few hundred samples~\cite{Puerrer14}.[JB:I cut this down significantly.]} Our challenge will be to accurately complete the representation of instrumental signals including precession effects and the LISA instrumental response and precession effects without sacrificing the computational efficiency enabled by these fast waveform templates.
 
@@
 In presence of misaligned spins, the system will endure precession of the orbital plane as it evolves, leading to modulations of the signal as seen by the observer~\cite{ACST94}.
-Precessing EOB templates incorporate all six parameters of the two spins~\cite{Pan+13, BTB16} but still require a costly time-domain integration; while precessing phenomenological templates~\cite{Hannam+13} decrease the dimensionnality by using effective spin parameters representing only the dominant effects of the spins.
-To model the effect of precession on the emitted waveform, it has been proposed~\cite{BCV03b, BCPTV05, Schmidt+10, OShaughnessy+11, Boyle+11} to decompose precessing waveforms by performing a time-dependent rotation, following the precession of the orbital plane.
+%Precessing EOB templates incorporate all six parameters of the two spins~\cite{Pan+13, BTB16} but still require a costly time-domain integration; while precessing phenomenological templates~\cite{Hannam+13} decrease the dimensionnality by using effective spin parameters representing only the dominant effects of the spins.
+\jgb{[JB:I cut this down too.]}
+
+A promising approach to modeling the effect of precession on the emitted waveform, it has been proposed~\cite{BCV03b, BCPTV05, Schmidt+10, OShaughnessy+11, Boyle+11} to decompose precessing waveforms by performing a time-dependent rotation, following the precession of the orbital plane.
 The resulting waveform in the rotated frame can then be modelled by a non-precessing waveform, an approximation which is used both in the construction of the inspiral part of precessing EOB waveforms~\cite{Pan+13} and in the construction of precessing phenomenological waveforms~\cite{Hannam+13}.
@@
 
-Prospective parameter recovery studies for LISA observations of comparable-mass binary black hole coalescences aim at characterizing the scientific performance of the future experiment.
+Beyond a fast representation of the incident waveform, observational analyses also require transforming the signals through some instrumental response.
+For short duration mergers such as LIGO  has detected, this can be treated by a simple multiplier.
+For future instruments though, the instrumental response will be more complicated.
+
+%Prospective parameter recovery studies for LISA observations of comparable-mass binary black hole coalescences aim at characterizing the scientific performance of the future experiment.
 Although LISA~\cite{LISA17} is still at least a decade from launch, the ESA L3 mission is rapidly progressing into its early stages where mission design choices must be informed by science impact analysis.
@@
 
-In contrast with current ground-based detectors, which are typically sensitive to chirping binaries for less than a second, LISA signals may accumulating over months or years.
+\jgb{Whereas LIGO and Virgo are typically sensitive to chirping binaries for a minute or less, LISA signals may accumulating over months or years.
 The response of a LISA-type instrument is thus time-dependent~\cite{Cutler97}.
 The motion and change of orientation of the detector constellation along its orbit lead to significant time variability in the form of a modulation and a varying delay.
-These effeccts then convey information about the localization of the gravitational-wave source in the sky.
-The complicated detector response increase the computational cost of parameter-estimation analyses.
-To leverage the performance of state-of-the-art Fourier-domain IMR templates~\cite{BTB16,Khan+15}, one needs to be able to process the signals through the time-dependent response of the dectector while staying in the Fourier domain.
+These effects then convey information about the localization of the gravitational-wave source in the sky.
+Direct implementation of the detector response is straightforward\cite{SynthLISA,LISACode,LISASim} but at prohibitive computational cost for parameter-estimation analyses.
+To leverage the performance of state-of-the-art Fourier-domain IMR templates~\cite{BTB16,Khan+15}, we must efficiently process the signals through the time-dependent response of the dectector while staying in the Fourier domain.}
 
@@
 
-In previous works, the Stationary Phase Approximation (SPA) (see e.g.~\cite{Thorne300, CF94}) has been often used for this purpose.
-The SPA is a common approximation to compute the Fourier transform of non-precessing signals during the inspiralling phase.
-In the case of precessing binaries, applying it directly to the modulated signal is however prone to pathologies.
-In Ref.~\cite{KCY13,KCY14}, a formalism (called shifted uniform asymptotics or SUA) was introduced to go beyond the SPA and compute more accurately the modulation in the Fourier domain; however, this formalism still relies on the SPA for the underlying precessing-frame signal, and is as such limited to inspiralling signals.
-Another intuitive approach consists in using a time-to-frequency correspondence, such as the one given by the SPA, to directly carry over the modulation to the Fourier domain.
-This corresponds, in a sense to be detailed below, to the zeroth-order approximation of the SUA, and it is used implicitly\footnote{With the difference that the expressions used for the modulation are given in terms of frequency directly for the inspiral, and are then extended to also cover high frequencies in the merger-ringdown part of the signal, as we will discuss below.} for the simplified precession of the phenomenological waveforms of Ref.~\cite{Hannam+13}.
+In previous works focused on gravitational-wave inspirals, the Stationary Phase Approximation (SPA) (see e.g.~\cite{Thorne300, CF94}) has been often used for this purpose.
+While the SPA is a common approximation to compute the Fourier transform of non-precessing signals during the inspiraling phase, it is not applicable for IMR waveforms.
+
+\jgb{In the case of precessing binaries, applying the SPA directly to the modulated signal is also prone to pathologies.
+In Ref.~\cite{KCY13,KCY14}, a formalism (called shifted uniform asymptotics or SUA) was introduced to go beyond the SPA and compute more accurately the modulation in the Fourier domain; however, this formalism still relies on the SPA for the underlying precessing-frame signal, and is as such limited to inspiraling signals.
+\textbf{[JB: Still uncertain about the rest of this paragraph.]} Most directly, a time-to-frequency correspondence, such as the one given by the SPA, can be exploited to directly carry over the precision modulation to the Fourier domain.} This corresponds, in a sense to be detailed below, to the zeroth-order approximation of the SUA, and it is used implicitly\footnote{With the difference that the expressions used for the modulation are given in terms of frequency directly for the inspiral, and are then extended to also cover high frequencies in the merger-ringdown part of the signal, as we will discuss below.} for the simplified precession of the phenomenological waveforms of Ref.~\cite{Hannam+13}.
+
 In the case of the LISA response, the SPA provides a natural map from time-domain (for the orbit) to Fourier-domain (for the signal)~\cite{Cutler97}.
@@
 
+We seek to overcome two limitations in such previous approaches.
+The first is that  SPA-based methods are not applicable to IMR waveforms.
+Second, there is often no clear way to improve the accuracy of these methods beyond the intuitive leading order treatment in order to meet the high-accuracy needs of future detectors.
+Our approach exploits separation of time-scales approximations, based on a general treatment directly in the Fourier domain of slowly varying delays and amplitude modulations for chirping waveforms with slowly varying amplitude and phase.
+
 The plan of the paper is as follows.
-In Sec.~\ref{sec:motivation}, we motivate the problem of Fourier-domain modulation and introduce the relevant timescales for both the response of LISA-type detectors and the modulation of precessing signals.
+In Sec.~\ref{sec:motivation}, we provide a general presentation of problem of Fourier-domain modulation and introduce the relevant timescales for both the response of LISA-type detectors and the modulation of precessing signals.
 In Sec.~\ref{sec:formalism}, we present our general formalism, give its leading order approximation as well as higher-order corrections, introduce new timescales based on the Fourier-domain signal, refine the previous results for both the quadratic-in-phase corrections and the treatment of the delays, and make the connection to previous approaches of the problem.
@@
 
-\SM{[talk about slowly chirping systems in the intro]}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Motivation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{GW signals in the frequency domain}
 \label{sec:motivation}
@@
 
-\subsection{Basic considerations}
-\label{subsec:basic}
-
-We can jointly consider challenges in the contexts of both the LISA response and GW precession applications in terms of a more general problem which covers both domains.
-In each case must determine the Fourier transform of a signal $h(t)$ to which we apply a varying delay $d(t)$ before multiplication by a modulation function $F(t)$,
+%\subsection{Basic considerations}
+%\label{subsec:basic}
+
+\jgb{To motivate our signal treatment we consider challenges of both the LISA response and GW precession applications in terms of a more general problem which covers both domains.
+In each case must determine the Fourier transform a signal $\tilde s(f)$ which results from processing some incident waveform $h(t)$ through some combination of modulations and delays.}
+
+In the general problem, we apply a time-varying delay $d(t)$ to the waveform followed by a multiplicative modulation function $F(t)$,
 \be
 \label{eq:delay-mod-defs}
-	h_{d}(t) = h(t+d(t)) \,, \quad s(t) = F(t)h_{d}(t) \,,
-\ee
-and seek an efficient way to commpute the Fourier transform $\tilde{s}(f)$, expressed by means of a Fourier-domain transfer function $\calT$ such that
+	h_{d}(t) = h(t+d(t)) \,, \quad s(t) = F(t)h_{d}(t) \,.
+\ee
+We then seek an \textit{efficient} way to compute the Fourier transform $\tilde{s}(f)$, expressed by means of a Fourier-domain transfer function $\calT$ such that
 \be\label{eq:deftransfer}
@@
 \ee
-As is well known, GW signals decomposed in spin-weighted spherical harmonic components are smoothly varying functions of frequency in amplitude/phase form.
+
+As is well known, GW signals decomposed in spin-weighted spherical harmonic components are smoothly varying functions of frequency in amplitude/phase form\cite{smoothAmpPhaseRefs}.
 An important consequence is that these signal components can be accurately represented in terms of a relatively coarsely sampled frequency grid.
@@
 
-In the case of precessing binaries, the delays will be absent and the modulation functions will be the time-dependent Wigner coefficients applied to rotate the waveform from a precessing frame, where it verifies a restored mode hierarchy, to an inertial frame where the observations take place.
-In the case of a LISA-type detector, the signal $h(t)$ will simply be the waveform in a fixed heliocentric frame, and the delays will come from the motion of each detector in the wave front, while the modulation will represent the time-variation in the detector orientation.
+In the case of precessing binaries, the delays will be absent and the modulation functions will be the time-dependent Wigner coefficients applied to rotate the waveform from an \jgb{precessing orbital frame, in which waveform modes exhibit smooth amplitude and phase variation}, to an inertial frame where the observations take place.
+In the case of a LISA-type detector, the signal $h(t)$ will simply be the waveform in a fixed heliocentric frame, and the delays will come from the motion of each detector against the wave front, while the modulation will represent the time-variation in the detector orientation.
 
@@
 
-Together, these observations let us think that, for a given $f$, only a limited range of times should be relevant in $d(t)$ and $F(t)$ so that we may expect to find a treatment for the transfer function that would be local in time for the modulation and delay.
-Intuitively, for an inspiral signal and in the limit of an extremely slowly varying $d(t)$ and $F(t)$, we should be able to simply evaluate them at the time given by the time-to-frequency correspondence of the SPA.
-Furthermore, a natural quantitative criterion for the applicability of this idea is given by the comparison of the radiation-reaction timescale of the signal with the modulation timescale (possibly augmented by a dimensionless factor when delays are present).
+Together, these observations lead us to expect that, for a given $f$, only a limited range of times should be relevant in $d(t)$ and $F(t)$ so that we may expect to find a treatment for the transfer function that would be local in time for the modulation and delay.
+This general idea has already been applied for inspiral signals in the limit of an extremely slowly varying $d(t)$ and $F(t)$, where intuitively we should be able to simply evaluate them at the time given by the time-to-frequency correspondence of the SPA.
+
+Furthermore, a natural quantitative criterion for the applicability of this idea is given by the comparison of the radiation-reaction timescale of the signal with the modulation timescale.
 In other words, the change in frequency of the signal over a characeteristic time of the modulation should be small for the separation of timescales to work.
@@
 Our objective is to find an approximate treatment for $\calT(f)$ which allows us to exploit these properties without relying on unnecessary limiting assumptions for the signal (like the limitation of the SPA to inspiral signals), which is extensible to the high-accuracies which will be required by LISA and other future gravitational-wave instruments, while being computationally efficient and widely applicable to GW analysis.
-Before introducing developing our formalism, though, we review the salient features of our application problems.
+In preparation for developing our formalism we first review the salient features of our application problems.
 
@@
 Here we will work with the second representation, which will prove more convenient for our purposes.
-Moreover, various notation and conventions have been used in the literature to label the spacecrafts and describe their orbits.
+Moreover, various notation and conventions have been used in the literature to label the spacecraft and describe their orbits.
 We refer the reader to~\cite{Vallisneri04} for a comparative account on these various conventions.
@@
 We introduce the notation $\hatk$ for the propagation vector of the gravitational wave, which we denote by $h_{ij}^{\rm TT}(t)$ in transverse-traceless matrix form, as measured at the SSB (thus, at position $p$, $h_{ij}^{\rm TT}(t, p) = h_{ij}^{\rm TT}(t - \hatk \cdot p)$).
-We denote by $p_{A}$ ($A=1,2,3$) the position of the individual spacecrafts and $n_{l}$ the unit vectors of the three links, with the convention that $n_{3}$ points from 1 to 2.
-
-Written in terms of the fractional laser frequency shifts between two spacecrafts, the elementary response of the detector reads~\cite{EW75, RCP04, Vallisneri04}
+We denote by $p_{A}$ ($A=1,2,3$) the position of the individual spacecraft and $n_{l}$ the unit vectors of the three links, with the convention that $n_{3}$ points from 1 to 2.
+
+Written in terms of the fractional laser frequency shifts between two spacecraft, the elementary response of the detector reads~\cite{EW75, RCP04, Vallisneri04}
 \begin{align}\label{eq:yslr}
@@
 \end{align}
-We will use a rigid instantaneous model, approximating the geometry of the constellation by a moving equilateral triangle (neglecting the flexing of the arms induced by corrections in the orbits), and evaluating all geometric factors in~\eqref{eq:yslr} at a single time $t$ (neglecting pointing-ahead corrections).
-The additional delay $L$ (we use $c=1$) in the first term represents the light propagation time along the arm between spacecrafts $s$ and $l$.
+We will use a rigid instantaneous model, approximating the geometry of the constellation by a moving equilateral triangle (neglecting the flexing of the arms induced by corrections in the orbits), and evaluating all geometric factors in~\eqref{eq:yslr} at a single time $t$ (neglecting point-ahead corrections).
+The additional delay $L$ (we use $c=1$) in the first term represents the light propagation time along the arm between spacecraft $s$ and $l$.
 
@@
 Since in our rigid approximation these delays will take a simple form in Fourier domain, we will focus in this paper on the $y_{slr}$ observables.
-We will give in Sec.~\ref{subsec:modelLISA} more details on the response of LISA-type detectors, on the TDI observables, and on the approximations that enter the derivation of the basic response~\eqref{eq:yslr} above.
+Sec.~\ref{subsec:modelLISA} provides more details on the response of LISA-type detectors, on the TDI observables, and on the approximations that enter the derivation of the basic response~\eqref{eq:yslr} above.
 
@@
 
-The functions $F(t)$ and $d(t)$ vary on a timescale of one year, with a frequency $f_{0} = 1/\mathrm{yr} \simeq 3.169\times10^{-8} \mathrm{Hz}$ (we will also use $\Omega_{0} = 2\pi f_{0}$).
+For the LISA response, the functions $F(t)$ and $d(t)$ vary on a timescale of one year, with a frequency $f_{0} = 1/\mathrm{yr} \simeq 3.169\times10^{-8} \mathrm{Hz}$ (we will also use $\Omega_{0} = 2\pi f_{0}$).
 When neglecting small non-periodic orbital perturbations (such as those caused by the influence of the Earth and of the other planets on the orbits), both $F(t)$ and $d(t)$ are periodic.
-It will also be useful to separate the delays in two types of terms: the first, $d_{0} = -\hatk\cdot p_{0}$, relates the waveform at the SSB to the waveform at the center of the constellation, whereas the second, $d_{c}$, represents the various delays between the spacecrafts of the constellation.
-For an orbit radius $R=1\,\mathrm{au}$ and an armlength $L=2.5\times10^{6}\mathrm{km}$~\cite{LISA17}, $d_{0} \sim R/c \simeq 500\mathrm{s}$ and $d_{c} \sim L/c \simeq 8\mathrm{s}$.
+It will also be useful to separate the delays in two types of terms: the first, $d_{0} = -\hatk\cdot p_{0}$, relates the waveform at the SSB to the waveform at the center of the constellation, whereas the second, $d_{c}$, represents the various delays between the spacecraft of the constellation.
+\jgb{Concretely, we assume the basic reference design parameters of the LISA mission\cite{LISA17} recently selected by the European Space Agency(ESA), with orbit radius $R=1\,\mathrm{au}$ and armlength $L=2.5\times10^{6}\mathrm{km}$, and with $d_{0} \sim R/c \simeq 500\mathrm{s}$ and $d_{c} \sim L/c \simeq 8\mathrm{s}$.}
 
@@
 
-For spinning compact objects, the presence of the spins leads to the precession of the orbit~\cite{Apostolatos+94, Kidder95}, which can have large effects on the waveform.
-In particular, it breaks the planar symmetry of the gravitational wave emission, and leads to modulations that are especially important for systems that are observed edge-on.
-
-It was suggested by a number of authors~\cite{BCV03b, BCPTV05, Schmidt+10, OShaughnessy+11, Boyle+11} that the effect of the precession can be modeled, to a good approximation, by a time-dependent rotation of a non-precessing waveform.
-This allows for a modelling approach where one models separately a precessing frame following the evolution of the plane of the orbit, and approximates the waveform in this precessing frame by using a non-precessing model.
+For spinning compact objects, \jgb{angular momentum interactions typically} lead to the precession of the orbit~\cite{Apostolatos+94, Kidder95}, which can have large effects on the waveform.
+In particular, it breaks the planar symmetry of the gravitational wave emission, causing modulations that are especially important for systems that are observed edge-on.
+
+A number of authors~\cite{BCV03b, BCPTV05, Schmidt+10, OShaughnessy+11, Boyle+11} have suggested that the effect of the precession can be modeled, to a good approximation, by a time-dependent rotation of a effectively non-precessing waveform.
+This allows for a modelling approach where one separately models a precessing frame following the evolution of the plane of the orbit, and approximates the waveform in this precessing frame by using a effective non-precessing model.
 Different prescriptions have been proposed for the construction of a precessing frame from the waveform itself~\cite{Schmidt+10, OShaughnessy+11, Boyle+11}.
@@
 
-As a preliminary step, we recall here an approximation widely used for similar purposes in treating gravitational waves signals emitted by inspiralling binaries, the Stationary Phase Approximation (thereafter SPA, and also sometimes called the steepest-descent method).
+As a preliminary step, we recall here an approximation widely used for similar purposes in treating gravitational waves signals emitted by inspiraling binaries, the Stationary Phase Approximation (thereafter SPA, and also sometimes called the steepest-descent method).
 It applies in general to chirping signals, and we refer the reader to~\ref{} for details.
@@
 
-To understand the separation of timescales in the problem, it will be useful to have at hand the leading-order scaling laws for an inspiral (at the Newtonian order in the PN language).
-Although inaccurate for the purpose of waveform modelling, these Newtonian estimates will give useful orders of magnitude of the relevant timescales in the inspiral.
+To understand the separation of timescales in the problem, it will be useful to have at hand the leading-order scaling laws for an inspiral (labeled the Newtonian order in the PN language).
+Although inaccurate for the purpose of waveform modelling, these leading-order estimates will give useful orders of magnitude of the relevant timescales in the inspiral.
 For a binary with masses $m_{1}, m_{2}$, we define the total mass $M=m_{1}+m_{2}$ and the symmetric mass ratio $\nu = m_{1}m_{2}/M^{2}$.
@@
 In this Newtonian, low-frequency limit one can check that each of the combinations \eqref{eq:conditionsSPA} indeed vanish at $\calO{(v^5)}$.
-but as the system approaches merger, this condition is no-longer satisfied.
-However, near and after the merger, the SPA treatment is not applicable, as can be seen by the breakdown of the relation~\eqref{eq:deftfSPA} as $\omega$ asymptotes the quasi-normal mode frequency in the ringdown.
+but as the system approaches merger, these condition are no-longer satisfied.
+Near and after the merger, the SPA treatment is not applicable.
+%, as can be seen by the breakdown of the relation~\eqref{eq:deftfSPA} as $\omega$ asymptotes the quasi-normal mode frequency in the ringdown.
 Because of this limitation, we aim at developing a formalism that would not rely on this approximation.
-Nonetheless, the SPA has been a useful workhorse in many gravitational analysis involving frequency domain transfer functions of the form~\eqref{eq:deftransfer}.
+Nonetheless, the SPA has been a useful workhorse in many gravitational analyses involving frequency domain transfer functions of the form~\eqref{eq:deftransfer}.
 The usual approach is simply to replace any time dependencies appearing the transfer function using \eqref{eq:deftfSPA}.
-We will later return to further consider the SPA treatment in the context of a more general formalism.
+\jgb{We will reference the SPA treatment, as a familiar touchstone, as we develop a more general formalism.}
 
@@
 The frequency-shift response for a single link~\eqref{eq:yslr} was derived in Ref.~\cite{EW75} (see also~\cite{RCP04, Finn08, Cornish09} for more modern derivations).
-Several assumptions enter the result as written in~\eqref{eq:yslr}: (i) effects of the order $v/c$ are neglected, including for instance the special relativistic Doppler effect created by the relative speeds of the spacecrafts on their orbits (ii) the propagation is assumed to take place in a flat spacetime, perturbed only by the gravitational wave; thus the gravitational redshift as well as the deflection of light created by the gravitational potential of the Sun is ignored (iii) all geometric factors are evaluated at a single time, whereas one should consider the beam as propagating from the position of the first spacecraft at the time of emission to the position of the second scapecraft at the time of reception, leading to a pointing-ahead effect.
+Several assumptions enter the result as written in~\eqref{eq:yslr}: (i) effects of the order $v/c$ are neglected, including for instance the special relativistic Doppler effect created by the relative speeds of the spacecraft on their orbits (ii) the propagation is assumed to take place in a flat spacetime, perturbed only by the gravitational wave; thus the gravitational redshift as well as the deflection of light created by the gravitational potential of the Sun is ignored (iii) all geometric factors are evaluated at a single time, whereas one should consider the beam as propagating from the position of the first spacecraft at the time of emission to the position of the second scapecraft at the time of reception, leading to a pointing-ahead effect.
 
@@
 \end{align}
-where we introduced the positions of the spacecrafts relative to the center of the constellation, $p^{L}_{A} \equiv p_{A} - p_{0}$.
+where we introduced the positions of the spacecraft relative to the center of the constellation, $p^{L}_{A} \equiv p_{A} - p_{0}$.
 
@@
 
-Note that, if the corrections of Sec.~\ref{subsec:delays} are included for the constellation delays, the transfer function will not have this simple form anymore, as $\dot{d}$ will have a different velocity-dependent expression for the sending and receiving spacecrafts.
+Note that, if the corrections of Sec.~\ref{subsec:delays} are included for the constellation delays, the transfer function will not have this simple form anymore, as $\dot{d}$ will have a different velocity-dependent expression for the sending and receiving spacecraft.
 One must then handle separately $d_{s,L}$ and $d_{r}$ in~\eqref{eq:decomposeGslr}.
@@
 We will now turn to the application of our formalism to the decomposition of signals from precessing binaries.
-As we already discussed in Sec.~\ref{subsec:modulationPrec}, in the presence of spin components that are not aligned with the orbital angular momentum, an inspiralling binary system will undergo precession of its orbital plane~\cite{Apostolatos+94, Kidder95}.
+As we already discussed in Sec.~\ref{subsec:modulationPrec}, in the presence of spin components that are not aligned with the orbital angular momentum, an inspiraling binary system will undergo precession of its orbital plane~\cite{Apostolatos+94, Kidder95}.
 This is a crucial effect to be taken into account in the modelling of such signals.

### 2018-02-14 LISAinference/ComputeLISASNR.c
Changes to the noise and utility executables.
- updated noise to include the 2017 proposal noise used in the LDC (NOTE: made the new default)
- updated interface for ComputeLISASNR and GenerateTDIFD

@@ -214,10 +214,14 @@ static int AllowedTDItag(TDItag tag) {
 
 int main(int argc, char *argv[])
 {
-  //These are set by command line in other programs but fixed here.
-  LISAconstellation *variant=&LISA2017;
-  int tRefatLISA=0;
-    
+  /* These global parameters are set by command line in other programs but fixed here. */
+  LISAconstellation *variant = &LISAProposal;
+  int tagtRefatLISA = 0;
+  int tagsimplelikelihood = 0;
+  int zerolikelihood = 0;
+  int frozenLISA = 0;
+  ResponseApproxtag responseapprox = full;
+
   double SNR = 0;
 
   /* Initialize structure for parameters */
@@ -272,13 +276,13 @@ int main(int argc, char *argv[])
       gsl_vector* noisevaluesE = gsl_vector_alloc(sizeE);
       gsl_vector* noisevaluesT = gsl_vector_alloc(sizeT);
       for(int i=0; i<sizeA; i++) {
-	gsl_vector_set(noisevaluesA, i, SnAXYZNoRescaling(variant, gsl_vector_get(TDI1FFTrestr->freq, i)));
+        gsl_vector_set(noisevaluesA, i, SnAXYZNoRescaling(variant, gsl_vector_get(TDI1FFTrestr->freq, i)));
       }
       for(int i=0; i<sizeE; i++) {
-	gsl_vector_set(noisevaluesE, i, SnEXYZNoRescaling(variant, gsl_vector_get(TDI2FFTrestr->freq, i)));
+        gsl_vector_set(noisevaluesE, i, SnEXYZNoRescaling(variant, gsl_vector_get(TDI2FFTrestr->freq, i)));
       }
       for(int i=0; i<sizeT; i++) {
-	gsl_vector_set(noisevaluesT, i, SnTXYZNoRescaling(variant, gsl_vector_get(TDI3FFTrestr->freq, i)));
+        gsl_vector_set(noisevaluesT, i, SnTXYZNoRescaling(variant, gsl_vector_get(TDI3FFTrestr->freq, i)));
       }
       double SNRA2 = FDOverlapReImvsReIm(TDI1FFTrestr, TDI1FFTrestr, noisevaluesA);
       double SNRE2 = FDOverlapReImvsReIm(TDI2FFTrestr, TDI2FFTrestr, noisevaluesE);
@@ -318,10 +322,15 @@ int main(int argc, char *argv[])
[…]

### 2018-02-14 LISAinference/LISAutils.c

@@ -173,7 +173,7 @@ Arguments are as follows:\n\
  --tagint              Tag choosing the integrator: 0 for Fresnel (default), 1 for linear integration\n\
  --tagtdi              Tag choosing the set of TDI variables to use (default TDIAETXYZ)\n\
  --nbptsoverlap        Number of points to use for linear integration (default 32768)\n\
- --variant             String representing the variant of LISA to be applied (default LISA2017)\n\
+ --variant             String representing the variant of LISA to be applied (default LISAProposal)\n\
  --zerolikelihood      Zero out the likelihood to sample from the prior for testing purposes (default 0)\n\
  --frozenLISA          Freeze the orbital configuration to the time of peak of the injection (default 0)\n\
  --responseapprox      Approximation in the GAB and orb response - choices are full (full response, default), lowfL (keep orbital delay frequency-dependence but simplify constellation response) and lowf (simplify constellation and orbital response) - WARNING : at the moment noises are not consistent, and TDI combinations from the GAB are unchanged\n\
@@ -278,7 +278,7 @@ Syntax: --PARAM-min\n\
     globalparams->tagint = 0;
     globalparams->tagtdi = TDIAETXYZ;
     globalparams->nbptsoverlap = 32768;
-    globalparams->variant = &LISA2017;
+    globalparams->variant = &LISAProposal;
     globalparams->zerolikelihood = 0;
     globalparams->frozenLISA = 0;
     globalparams->responseapprox = full;
@@ -558,15 +558,16 @@ Syntax: --PARAM-min\n\
             run->seed = 1;
         } else if (strcmp(argv[i], "--variant") == 0) {
 	  i++;
-	  if (strcmp(argv[i], "LISA2017") == 0) globalparams->variant=&LISA2017;
-	  else if (strcmp(argv[i], "LISA2010") == 0) globalparams->variant=&LISA2010;
-	  else if (strcmp(argv[i], "fastOrbitLISA") == 0) globalparams->variant=&fastOrbitLISA;
-	  else if (strcmp(argv[i], "slowOrbitLISA") == 0) globalparams->variant=&slowOrbitLISA;
-	  else if (strcmp(argv[i], "tinyOrbitLISA") == 0) globalparams->variant=&tinyOrbitLISA;
[…]

### 2018-02-14 LISAsim/GenerateTDIFD.c

@@ -355,10 +355,10 @@ static void Write_TDIhlm(const char dir[], const char file[], ListmodesCAmpPhase
 
 int main(int argc, char *argv[])
 {
-  //These are set by command line in other programs but fixed here.
-  LISAconstellation *variant=&LISA2017;
-  int tRefatLISA=0;
-    
+  /* These global parameters are set by command line in other programs but fixed here. */
+  LISAconstellation *variant = &LISAProposal;
+  int tagtRefatLISA = 0;
+
   /* Initialize structure for parameters */
   GenTDIFDparams* params;
   params = (GenTDIFDparams*) malloc(sizeof(GenTDIFDparams));
@@ -527,7 +527,7 @@ int main(int argc, char *argv[])
     ListmodesCAmpPhaseFrequencySeries* listTDI2= NULL;
     ListmodesCAmpPhaseFrequencySeries* listTDI3= NULL;
     /* torb is the orbital reference time (role played by injection tRef in the inference) */
-    LISASimFDResponseTDI3Chan(tRefatLISA, variant, &listhlm, &listTDI1, &listTDI2, &listTDI3, params->torb, params->lambda, params->beta, params->inclination, params->polarization, params->m1, params->m2, params->maxf, params->tagtdi, params->frozenLISA, params->responseapprox);
+    LISASimFDResponseTDI3Chan(tagtRefatLISA, variant, &listhlm, &listTDI1, &listTDI2, &listTDI3, params->torb, params->lambda, params->beta, params->inclination, params->polarization, params->m1, params->m2, params->maxf, params->tagtdi, params->frozenLISA, params->responseapprox);
 
     /* If asked for it, rescale the complex amplitudes to include the factors that were scaled out of TDI observables */
     if(params->restorescaledfactor) {

### 2018-02-14 LISAsim/GenerateTDITD.c

@@ -149,7 +149,8 @@ int main(int argc, char *argv[])
 {
   int ret;
 
-  LISAconstellation *variant = &LISA2017;
+  LISAconstellation *variant = &LISAProposal;
+  
   /* Initialize structure for parameters */
   GenTDITDparams* params;
   params = (GenTDITDparams*) malloc(sizeof(GenTDITDparams));

### 2018-02-14 LISAsim/LISAFDresponse.c

@@ -477,14 +477,13 @@ int LISASimFDResponseTDI3Chan(
       //double phaseRdelay = -2*PI*R_SI/C_SI*f*cos(beta)*cos(Omega_SI*tf - lambda) * (1 + R_SI/C_SI*cos(beta)*Omega_SI*sin(Omega_SI*tf - lambda));
       double phaseRdelay;
       if(tagtRefatLISA==0){//delay so that tinj=torb refers to arrival time at SSB
-	phaseRdelay = -2*PI*OrbitRoC*f*cos(beta)*cos(phase) * (1 + cos(beta)*OrbitRoC*variant->OrbitOmega*sin(phase));
+        phaseRdelay = -2*PI*OrbitRoC*f*cos(beta)*cos(phase) * (1 + cos(beta)*OrbitRoC*variant->OrbitOmega*sin(phase));
       } else {
-	//In this version we delay so that tinj=torb is relative to LISAcenter arrival time, so there is no orbital delay when tf = tinj
-	//If the original delay realized td=t+d(t), now we want td=t+d(t)-d(t0), that we we change d(t) -> d(t) + d(t0)
-	//Then with the approximation d(td)=d(t)(1-ddot(t)),
-	double phase0=variant->OrbitOmega*torb+variant->OrbitPhi0-lambda;
-	phaseRdelay =
-	  - 2*PI*OrbitRoC*f*cos(beta)*( cos(phase) -cos(phase0) ) * (1 + cos(beta)*OrbitRoC*variant->OrbitOmega*sin(phase));
+        //In this version we delay so that tinj=torb is relative to LISAcenter arrival time, so there is no orbital delay when tf = tinj
+        //If the original delay realized td=t+d(t), now we want td=t+d(t)-d(t0), that we we change d(t) -> d(t) + d(t0)
+        //Then with the approximation d(td)=d(t)(1-ddot(t)),
+        double phase0 = variant->OrbitOmega*torb + variant->OrbitPhi0 - lambda;
+        phaseRdelay = -2*PI*OrbitRoC*f*cos(beta)*( cos(phase)-cos(phase0) ) * (1 + cos(beta)*OrbitRoC*variant->OrbitOmega*sin(phase));
       }
       if(responseapprox==lowf) { /* In the full low-f approximation, ignore this delay term */
         phaseRdelay = 0.;
@@ -503,8 +502,8 @@ int LISASimFDResponseTDI3Chan(
[…]

### 2018-02-14 LISAsim/LISAgeometry.c

@@ -45,14 +45,14 @@ struct tagLISAconstellation {
 }
 */
 
-LISAconstellation LISA2010 = {
+LISAconstellation LISAProposal = {
   EarthOrbitOmega_SI,
   0,
   AU_SI,
   EarthOrbitOmega_SI,
   0,
-  5e9,
-  LISA2010noise
+  2.5e9,
+  LISAProposalnoise
 };
 
 LISAconstellation LISA2017 = {
@@ -63,6 +63,16 @@ LISAconstellation LISA2017 = {
   0,
   2.5e9,
   LISA2017noise
+};
+
+LISAconstellation LISA2010 = {
+  EarthOrbitOmega_SI,
+  0,
+  AU_SI,
+  EarthOrbitOmega_SI,
+  0,
+  5e9,
+  LISA2010noise
 };
 
 LISAconstellation slowOrbitLISA = {
@@ -639,8 +649,8 @@ int EvaluateGABmode(
   const int tagdelayR,                     /* Tag: when 1, include the phase term of the R-delay */
   const ResponseApproxtag responseapprox)  /* Tag to select possible low-f approximation level in FD response */
 {
-  double phase=variant->ConstOmega*t + variant->ConstPhi0;
-
+  double phase = variant->ConstOmega*t + variant->ConstPhi0;
+  
   /* Precompute array of sine/cosine */
   for(int j=0; j<4; j++) {
     cosarray[j] = cos((j+1) * phase);

### 2018-02-14 LISAsim/LISAgeometry.h

@@ -74,10 +74,11 @@ typedef enum ResponseApproxtag {
 /**************************************************/
 /************** LISAconstellation *****************/
 
-enum LISANoiseType{
+typedef enum LISANoiseType{
   LISA2010noise,
-  LISA2017noise
-} ;
+  LISA2017noise,
+  LISAProposalnoise
+} LISANoiseType;
 
 typedef struct tagLISAconstellation {
   double OrbitOmega;
@@ -86,8 +87,9 @@ typedef struct tagLISAconstellation {
   double ConstOmega;
   double ConstPhi0;
   double ConstL;
-  enum LISANoiseType noise;
+  LISANoiseType noise;
 } LISAconstellation;
+extern LISAconstellation LISAProposal;
 extern LISAconstellation LISA2017;
 extern LISAconstellation LISA2010;
 extern LISAconstellation slowOrbitLISA;

### 2018-02-14 LISAsim/LISAnoise.c

@@ -42,7 +42,6 @@
 
 /* Proof mass and optic noises - f in Hz */
 /* Taken from (4) in McWilliams&al_0911 */
-
 static double SpmLISA2010(const double f) {
   double invf2 = 1./(f*f);
   //return 2.5e-48 * invf2 * sqrt(1. + 1e-8*invf2);
@@ -52,18 +51,16 @@ static double SpmLISA2010(const double f) {
   double invf8=invf2*invf2*invf2*invf2;
   //Here we add an eyeball approximation based on 4yrs integration with L3LISAReferenceMission looking at a private comm from Neil Cornish 2016.11.12
   double WDWDnoise=5000.0/sqrt(1e-21*invf8 + invf2 + 3e28/invf8)*SaccelFF*invf2;
-  return SaccelFF * invf2 * sqrt(1. + 1e-8*invf2) + WDWDnoise; 
+  return SaccelFF * invf2 * sqrt(1. + 1e-8*invf2) + WDWDnoise;
 }
-
 static double SopLISA2010(const double f) {
   const double Dop=2.0e-11; //Optical path noise in m/rtHz (Standard LISA)
   const double SopFF=Dop*Dop*4.0*PI*PI/C_SI/C_SI; //f^2 coeff for OP frac-freq noise PSD.  Yields 1.76e-37 for Dop=2e-11.
   return SopFF * f * f;
 }
 
-
-// Proof mass and optical noises - f in Hz
-// L3 Reference Mission, from Petiteau LISA-CST-TN-0001
+/* Proof mass and optical noises - f in Hz */
+/* L3 Reference Mission, from Petiteau LISA-CST-TN-0001 */
 static double SpmLISA2017(const double f) {
   double invf2 = 1./(f*f);
   //double invf4=invf2*invf2;
@@ -85,7 +82,6 @@ static double SpmLISA2017(const double f) {
   double Spm_vel = ( Saccel_red + Sloc + S4yrWDWD );
   return Spm_vel / C2;//finally convert from velocity noise to fractional-frequency doppler noise.
 }
-
 static double SopLISA2017(const double f) {
   //double invf2 = 1./(f*f);
   const double twopi2=4.0*PI*PI;
[…]

### 2018-03-01 docs/paper_method.tex
Changes to the noise and utility executables.
- updated noise to include the 2017 proposal noise used in the LDC (NOTE: made the new default)
- updated interface for ComputeLISASNR and GenerateTDIFD
More edits through IV.B

@@
 In Ref.~\cite{KCY13,KCY14}, a formalism (called shifted uniform asymptotics or SUA) was introduced to go beyond the SPA and compute more accurately the modulation in the Fourier domain; however, this formalism still relies on the SPA for the underlying precessing-frame signal, and is as such limited to inspiraling signals.
-\textbf{[JB: Still uncertain about the rest of this paragraph.]} Most directly, a time-to-frequency correspondence, such as the one given by the SPA, can be exploited to directly carry over the precision modulation to the Fourier domain.} This corresponds, in a sense to be detailed below, to the zeroth-order approximation of the SUA, and it is used implicitly\footnote{With the difference that the expressions used for the modulation are given in terms of frequency directly for the inspiral, and are then extended to also cover high frequencies in the merger-ringdown part of the signal, as we will discuss below.} for the simplified precession of the phenomenological waveforms of Ref.~\cite{Hannam+13}.
+\textbf{[JB: Did I capture your thoughts correctly in the following.]}
+The simplified treatment of the precession response in the phenomenological waveforms in Ref.~\cite{Hannam+13} takes another approach, treating the precession modulation in the frequency domain by directly associating the Fourier frequency with (a multiple of) the post-Newtonain orbital frequency.
+As we will explain, this corresponds to the zeroth-order approximation of the SUA.}
 
@@
 We begin by detailing the model that we use for the response of a LISA-like detector, together with the assumptions used and their limitations.
-Since our aim is to assess the accuracy of our direct Fourier-domain treatment of the response, we can focus on the gravitational-wave contribution to the basic observables.
+Since our aim is to assess the accuracy of our direct Fourier-domain treatment of the response, we can focus on the gravitational-wave contribution to the basic single-linkobservables.
 We therefore use a somewhat simplified model for the time-domain response, ignoring corrections that are crucial from the point of view of noise cancellations, keeping in mind that the response model can be enriched later without affecting the conclusions of the present analysis.
@@
 
-Additionally, we limit ourselves to a rigid model for the orbits of the constellation, namely we assume that the constellation remains in an equilateral configuration with fixed armlengths $L=2.5\times 10^{6}\mathrm{km}$ (the latest configuration~\cite{LISA17}).
-This simplified orbits neglect (iv) the effect of excentricity of the individual Keplerian orbits, which introduces corrections at the level $e^{2}$ (v) the effect of gravitational perturbations coming from other celestial bodies, such as the Earth, the quadrupole of the Sun, and the other planets.
+Additionally, we limit ourselves to a rigid model for the orbits of the constellation, namely we assume that the constellation remains in an equilateral configuration with fixed armlengths.
+These simplified orbits neglect (iv) effects of order $e^{2}$ from the eccentricity of the individual Keplerian orbits, (v) the effect of gravitational perturbations coming from other celestial bodies, such as the Earth, the quadrupole of the Sun, and the other planets.
 
@@
 
-It is natural to split the response~\eqref{eq:yslr} into two steps: the orbital delay related to the orbit around the Sun of the whole constellation on one side, and the constellation response on the other side.
+It is natural to split the response~\eqref{eq:yslr} into two steps: first the orbital delay related to the orbit around the Sun of the whole constellation, and then the constellation response.
 The baselines for the delays are indeed very different in the two cases.
@@
 \end{subequations}
-The LISA response behaves qualitatively differently on the frequency bands separated by these transfer frequencies.
-
-The first step consists simply in applying the varying time delay to bring the signal from the SSB to the center of the LISA triangular constellation, common to all $y_{slr}$ observables.
+The LISA response behaves qualitatively differently on the frequency bands separated by these transfer frequencies. \jgb{[\bf I don't understand the meaning or intent of the previous sentence]}
+
+\jgb{The first stage of the response, the orbital delay, consists simply in applying the varying time delay to bring the wavefront sampling point from the SSB reference to the center} of the LISA triangular constellation, common to all $y_{slr}$ observables.
 For $h^{\rm TT}$ the transverse-traceless gravitational waveform in matrix form, we write this orbital time delay as
@@
 \ee
-with $p_{0}$ the position of the constellation center, which follows the Earth orbit around the Sun. The second step consists in the remaining, constellation-centered response, which we write as
+with $p_{0}$ the position of the constellation center, which follows the Earth orbit around the Sun.
+The second \jgb{stage of the response calculation comprises the remaining, constellation-centered response.
+For the single-link contribution to the response, for the laser link from spacecraft $s$ to spacecraft $r$ along a path in direction $n_l=$ [{\bf Check that I got the notation right}], we write}
 \begin{align}\label{eq:defresponseL}
@@
 \end{align}
-where we introduced the positions of the spacecraft relative to the center of the constellation, $p^{L}_{A} \equiv p_{A} - p_{0}$.
+where we reference the positions of the spacecraft relative to the center of the constellation, $p^{L}_{A} \equiv p_{A} - p_{0}$.
 
@@
 
-We now turn to the transposition of~\eqref{eq:defresponse0} and~\eqref{eq:defresponseL} to the Fourier domain. Applying a pure delay as in~\eqref{eq:defresponse0} translates into
+We now turn to the transformation of~\eqref{eq:defresponse0} and~\eqref{eq:defresponseL} to the Fourier domain. Applying a pure delay as in~\eqref{eq:defresponse0} translates into
 \be\label{eq:G0}
@@
 
-For the constellation response~\eqref{eq:defresponseL}, we can separate the two delays by writing
+For the constellation response~\eqref{eq:defresponseL}, we can \jgb{separately treat the orbital delay $d_r$ and the constellation delay at spacecraft $s$, $d_{s,L}$}, by writing
 \begin{align}\label{eq:decomposeGslr}
@@
 \end{align}
-When also using the rigid approximation for the constellation, where the armlengths are fixed, a particular simplification occurs when combining these individual delays, thanks to the relation $p^{L}_{r} - p^{L}_{s} =  L n_{l}$:
+\jgb{Since we also assume the {\bf[right]} } rigid approximation for the constellation, where the armlengths are fixed, a particular simplification occurs when combining these individual delays, thanks to the relation $p^{L}_{r} - p^{L}_{s} =  L n_{l}$:
 \begin{align}\label{eq:GslrL}
@@
 
-Note that, if the corrections of Sec.~\ref{subsec:delays} are included for the constellation delays, the transfer function will not have this simple form anymore, as $\dot{d}$ will have a different velocity-dependent expression for the sending and receiving spacecraft.
-One must then handle separately $d_{s,L}$ and $d_{r}$ in~\eqref{eq:decomposeGslr}.
-
-The response~\eqref{eq:transfer0local}-\eqref{eq:delay0} takes a simple analytic form, but the phase contribution of this delay is significant on most of the frequency band and can be large for $f \gg f_{R}$.
-
-The response~\eqref{eq:GslrL}-\eqref{eq:transferLlocal} can be interpreted as the Fourier-domain translation of a discrete derivative taken on the waveform.
-The factor shows that the amplitude of the response is $\propto f$ in the low-frequency limit $f\ll f_{L}$, where the other factors are essentially unity.
-For $f\gtrsim f_{L}$, the $\sinc$ and the phase of the exponential give raise to additional structure in the response, including zero-crossings when the projected armlength corresponds to a wavelength.
-From~\eqref{eq:transferLlocal}, an expansion for small $f\ll f_{L}$ yields back a Fourier-domain analog of the low-frequency approximation of the response~\cite{Cutler97, RCP04}, which amounts to having two LIGO-type interferometers turned by $\pi/4$ and set in motion.
+Note that, if the corrections of Sec.~\ref{subsec:delays}, \jgb{for non-negligible $\dot d$}, are included for the constellation delays, the transfer function will not have this simple form anymore, as $\dot{d}$ will have a different velocity-dependent expression for the sending and receiving spacecraft.
+One must \jgb{separately handle $d_{s,L}$ and $d_{r}$ in~\eqref{eq:decomposeGslr} to compute the corrections.}
+
+The \jgb{orbital} response~\eqref{eq:transfer0local}-\eqref{eq:delay0} takes a simple analytic form, but the phase contribution of this delay is significant across most of the frequency band and can be large for $f \gg f_{R}$.
+
+The \jgb{constellation} response~\eqref{eq:GslrL}-\eqref{eq:transferLlocal} can be interpreted as the Fourier-domain translation of a discrete derivative taken on the waveform.
+The \jgb{leading factor in~\eqref{eq:GslrL}} shows that the amplitude of the response is proportional to $f$ in the low-frequency limit $f\ll f_{L}$, where the other factors are essentially unity.
+For $f\gtrsim f_{L}$, the $\sinc$ and the phase of the exponential generate additional structure in the response, including zero-crossings when the projected armlength \jgb{is an integer number of wavelengths}.
+From~\eqref{eq:transferLlocal}, an expansion for small $f\ll f_{L}$ yields back a Fourier-domain analog of the low-frequency approximation of the response~\cite{Cutler97, RCP04}, which \jgb{is equivalent to} having two LIGO-type interferometers turned by $\pi/4$ and set in motion.
 See Fig.~\ref{} and ~\ref{} \SM{[R]} for examples of LISA transfer functions.
 
-For the gravitational waveform itself, we will use the PhenomD model~\cite{Khan+15,Husa+15}, which provides Fourier-domain inspiral-merger-ringdown waveforms for aligned spins.
+\jgb{For analysis of the response we need a concrete set of gravitational waveforms.
+We will} use the PhenomD model~\cite{Khan+15,Husa+15}, which provides Fourier-domain inspiral-merger-ringdown waveforms for aligned spins.
 
@@
 The colored shaded areas indicate regions where $\epsilon_{\Psi} \geq 1$, where the perturbative formalism is expected to break down.
-In the gray area, $f_{\rm start}$ given in~\eqref{eq:fstartN} is lower than the lowest in-band frequency $f_{\rm min} = 10^{-5}\Hz$, so that the signal starts at $f_{\rm min}$ and $\epsilon_{\Psi}$ becomes independent of $\Delta t$.}
+In the gray area, $f_{\rm start}$ given in~\eqref{eq:fstartN} is lower than the lowest in-band frequency $f_{\rm min} = 10^{-5}\Hz$, so that the signal starts at $f_{\rm min}$ and $\epsilon_{\Psi}$ becomes independent of $\Delta t$.\jgb{\bf[Maybe not shade the ``grey area'' since it looks excluded, instead can call it ``the region to the right of the black line.'']} }
   \label{fig:lisafomPsiMcDeltat}
@@
 
-We will now use the approximate error measures $\epsilon$ introduced in~\eqref{eq:deffom} to estimate, for each type of correction (phase, amplitude, delay), the level of errors in the transfer function induced by ignoring this correction.
+Using the approximate error measures $\epsilon$ introduced in~\eqref{eq:deffom}, we will now estimate, for each type of correction (phase, amplitude, delay), the size of errors in the transfer function.
 
@@
 It is useful to separate the orbital response~\eqref{eq:transfer0local} and the constellation response~\eqref{eq:transferLlocal}, as the different baseline of the delays (orbital radius $R$ or armlength $L$) as well as the presence of a time-varying prefactor $F(t)$ both affect the result.
-When estimating the magnitude of the relevant derivatives of $G$, we have to take into account dimensionless delay factors of the type $2\pi f d$.
-
-We start with the orbital response, which takes the form of a pure delay $G_{0}(f, t) = e^{-2 i \pi f d_{0}(t)}$, and obtain
+When estimating the magnitude of the relevant derivatives of $G$, we must also take into account dimensionless delay factors of the type $2\pi f d$.
+
+We start with the orbital response, which takes the form of a pure delay $G_{0}(f, t) = e^{-2 i \pi f d_{0}(t)}$ \jgb{\bf[Why subscripted with zero not O or R?]}, and obtain
 \begin{subequations}
@@
 Since the one-arm constellation response~\eqref{eq:decomposeGslr}-\eqref{eq:transferLlocal} is analogous to a discrete time derivative of the signal, it is appropriate to keep explicit an overall factor $f$ reflecting this structure.
-Hence we write symbolically $G_{L}(f,t) \sim f F(t) e^{-2i \pi f d_{L}(t)}$, where $d_{L}$ represents a delay term and $F(t)$ represents the rest of the geometric factors in~\eqref{eq:transferLlocal} where we ignore the $f$-dependence.
+Hence we write symbolically $G_{L}(f,t) \sim f F(t) e^{-2i \pi f d_{L}(t)}$, where $d_{L}$ represents a delay term and $F(t)$ represents the rest of the geometric factors in~\eqref{eq:transferLlocal}, \jgb{for which we momentarily} ignore the $f$-dependence.
 This gives
@@
 
-An important point in the above is the presence of delay factors in the forms of powers of $2\pi f d$.
+\jgb{\bf[This paragraph is confusing.
+Can we point concretely to these factors somewhere, (e.g.
+in the expansion of the exponential.xxx)?
+Also it seems to say that we can't do what we just did.
+Should discuss.]} An important point in the above is the presence of delay factors in the forms of powers of $2\pi f d$.
 Thus, the magnitude of the corrections cannot be evaluated with the simple replacement $\partial_{t}^{n} \rightarrow \Omega_{0}^{n}$, and the frequency dependency of the error measures $\epsilon$ will depend on the frequency being above or below the transfer frequencies $f_{R}$ and $f_{L}$ defined in~\eqref{eq:transferfrequencies}.
@@
 \end{subequations}
-Note that if we consider that the frequency band of the instrument starts at a minimal frequency $f_{\rm min}$, for sufficiently high $\Mchirp$ this entry in the frequency band will mark the beginning of the signal, instead of $f_{\rm start}$ which becomes irrelevant.
+Note that if we \jgb{think of LISA as effectively insensitive below some minimal frequency $f_{\rm min}$, then for sufficiently high $\Mchirp$ this point of entry in the sensitive band will mark the beginning of the signal, obviating the relevance of $f_{\rm start}$}.
 For $f_{\rm min} = 10^{-5}\Hz$ and $\Delta t \leq 10\yr$, this is the case for $\Mchirp \geq 2.45\times 10^{6}\Msol $.
@@
 
-An important distinction needs to be made between what we will call merging binaries, systems close to coalescence that will merge during the LISA mission lifetime or a few years later, and slowly-chirping binaries still in the deep inspiral phase, of which we observe only a small snapshot in Fourier domain as they do not sweep to the end of the frequency band.
+\jgb{It is useful to distinguish} between what we will call merging binaries, systems close to coalescence that will merge during the LISA mission lifetime or a few years later, and slowly-chirping binaries still in the deep inspiral phase, of which we observe only a small snapshot in Fourier domain as they do not sweep to the end of the frequency band.
 The massive black hole binaries (MBH) that will be observed by LISA~\cite{LISA17} fall within the first category, with their merger in band, while the proposed population of stellar-origin black hole binaries (SOBH)~\cite{Sesana16}, with masses comparable to the LIGO/Virgo detections, will comprise both merging binaries, i.e.
@@
 The limit $\epsilon_{\Psi} = 1$ is used to single out areas where the perturbative treatment of Sec.~\ref{sec:formalism} is expected to break down, although the precise location of this boundary will vary, depending on the accuracy level required and on the orientation angles.
-We also single out the high-mass region where the start of the signal is set by its entry in the frequency band, taken to be $10^{-5}\Hz$, so that there is no dependence on the time to merger anymore.
-We find that, in the LIGO/Virgo mass range $10^{1}-10^{2} \Msol$, we reach the limit $\epsilon_{\Psi}=1$ first for the orbital response, for time-to-merger values within the expected observed range~\cite{Sesana16}.
+We also single out the high-mass region where the start of the signal is set by its entry \jgb{into the sensitive band, at $f_{\rm min}$ making the error measure independent of time to merger}.
+We find that, for binaries in the LIGO/Virgo mass range $10^{1}-10^{2} \Msol$, we reach the limit $\epsilon_{\Psi}=1$ first for the orbital response, for time-to-merger values within the expected observed range~\cite{Sesana16}.
 We will introduce in Sec~\ref{subsec:comblisa} an alternative approach to deal with those signals.
@@
 
-We now turn to the case of merging binaries, computing this time the three error measures $\epsilon_{\Psi}$, $\epsilon_{A1}$ and $\epsilon_{d}$.
+We now turn to the case of merging binaries.
+For these we compute the three error measures $\epsilon_{\Psi}$, $\epsilon_{A1}$ and $\epsilon_{d}$.
 To go beyond the crude analytical estimates built from~\eqref{eq:timescalesNfstart} and~\eqref{eq:estimatederivorb}-\eqref{eq:estimatederivconst}, we use numerical derivatives for the timescales~\eqref{eq:defTf}-\eqref{eq:defTA} and for the derivatives in~\eqref{eq:deffom}.
-We summarize the results in Fig.~\ref{fig:fomLISA}, both for the orbital delay and the constellation response, for an equal-mass system with total masses of $M=10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$ and a starting frequency corresponding to $\Delta t =10 \yr$ of observation.
+We summarize the results in Fig.~\ref{fig:fomLISA}, both for the orbital delay and the constellation response, considering equal-mass systems with total masses of $M=10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$ with a starting frequency corresponding to $\Delta t =10 \yr$ of observation.
 Since individual signals can show significant variation depending on the angular parameters, for the full computation with numerical derivatives we show a geometric average of $\epsilon$ over the sky position, inclination and polarization, together with $\pm 1\sigma$ geometric standard deviation.
@@
 
-For $\epsilon_{\Psi}$, we find as expected from~\eqref{eq:TfN} a steep rise towards lower frequencies, which shows that considering merging binaries with a limited time to coalescence is crucial here.
-Due to the different occurences of delay factors, we find a different behaviour between the orbital response, where for a fixed mass ratio the initial value of $\epsilon_{\Psi}$ grows towards lower masses, and the constellation response, where $\epsilon_{\Psi}$ grows towards higher masses.
-In both cases, the error measure remains below $\epsilon_{\Psi} \sim 0.1$, where the corrections are still well manageable by the perturbative formalism (as we will show in Sec.~\ref{subsec:errorsLISA}).
-
-For $\epsilon_{A1}$, we find that the structure in the waveform close to merger does have a visible but limited impact on the error measure.
+For $\epsilon_{\Psi}$, as expected from~\eqref{eq:TfN} we find a steep rise towards lower frequencies, which shows that considering merging binaries with a limited time to coalescence is crucial here.
+\jgb{We note differences in} behaviour between the orbital response, where for a fixed mass ratio the initial \jgb{(10yr before merger)} value of $\epsilon_{\Psi}$ grows towards lower masses, and the constellation response, where \jgb{initial} $\epsilon_{\Psi}$ grows towards higher masses.
+\jgb{Combining both parts of the response, for equal-mass mergers with $10^2<M<10^7$,} the error measure remains below $\epsilon_{\Psi} \sim 0.1$, where the corrections are still well manageable by the perturbative formalism (as we will show in Sec.~\ref{subsec:errorsLISA}).
+
+For $\epsilon_{A1}$, we find that the structure in the waveform close to merger does have a noticable but limited impact on the error measure.
 For the orbital response, the error measure remains small for all masses.

### 2018-04-05 docs/paper_method.tex
Small changes and edits to the paper.
- added y12 along y12L in the time-domain response
- default for bambi parameter eff changed from 0.1 to 0.5
- edits to the paper

@@
 For SOBH systems, however, $\eta$ will take a continuous set of values from roughly $10^{-4}$ to $1$ \SM{[check values]}.
-Although the systems for which $\eta$ is the largest are also the easiest to treat with the perturbative formalism of Sec~\ref{sec:formalism}, we propose here yet a third method for dealing with potentially troublesome systems.
+\SM{[add sentence: cover the intermediate ground]} Although the systems for which $\eta$ is the largest are also the easiest to treat with the perturbative formalism of Sec~\ref{sec:formalism}, we propose here yet a third method for dealing with potentially troublesome systems.
 We leave for the future a more detailed study of SOBH systems as a population, and the investigation of the precise boundaries and overlap areas of the three methods as well as their respective computational costs.
@@
 To assess the expected truncation error, we use a simple criterion based on the $L^{1}$-norm of the $\{c_{n}\}$ sequence.
-For a given target truncation error $\eta$, we define the truncation $N(f, \eta)$ as the smallest integer such that
+For a given target truncation error $\eta$, we define the truncation \SM{change notation to avoid confusion with $N$ from Fresnel stencil} $N(f, \eta)$ as the smallest integer such that
 \be\label{eq:criteriontruncationcomb}
@@
 The last angle $\gamma$ parametrizes the remaining freedom of rotation around this radiation axis.
-An important element of the construction of the precessing frame is the fixing of this third degree of freedom, for which we use the minimal rotation condition~\cite{Boyle+11}
+An important element of the construction of the precessing frame is the fixing of this third degree of freedom, for which we use the minimal rotation condition~\cite{Boyle+11} \SM{[explain what it means]}
 \be\label{eq:gammadot}
@@
 
-For the extension of the frame trajectory post-merger, we do not rely on the PhenomP procedure as it is merely effective.
+For the extension of the frame trajectory post-merger, we do not rely on the PhenomP procedure as it is merely effective \SM{[be clearer on the limitation here]}.
 Furthermore, we found that the ringdown attachment procedure, as implemented in SEOB, leads to frame trajectories that show oscillations in all the Euler angles when the spins are anti-aligned.
@@
 
+\SM{[write somewhere that we require mainly FD smoothness]}
+
 We now investigate the magnitude of higher-order corrections in the case of precessing binaries, using the error estimates $\epsilon_{\Psi 2}$, $\epsilon_{A1}$ and $\epsilon_{A 2}$ defined in~\eqref{eq:deffom}.
@@
 To obtain a qualitative picture of the timescales involved, here we will consider the most simple setting for precession in the post-Newtonian framework: at leading PN order and for a single-spin system, both the orbital angular momentum $\bm{L}$ and the spin vector $\bm{S}$ undergo what is called simple precession around the total angular momentum, $\bm{J}$ (see e.g.~\cite{Apostolatos+94, Kidder95}).
-By neglecting the slow evolution of the orbit under radiation reaction, ignoring higher-order corrections that are quadratic and higher in the spins, and orbit-averaging the precession equations (thus eliminating nutations in the precession of the orbital plane, which occur at the orbital timescale), the trajectories of $\bm{L}$ and $\bm{S}$ become cones around the direction of $\bm{J}$, with a constant precession velocity $\Omega_{\rm prec}$ that we will use to define the timescale of variations of the modulation functions $F$.
+By neglecting the slow evolution of the orbit under radiation reaction, ignoring higher-order corrections that are quadratic and higher in the spins, and orbit-averaging the precession equations (thus eliminating nutations in the precession of the orbital plane, which occur at the orbital timescale) \SM{[no nutation for single spin?]}, the trajectories of $\bm{L}$ and $\bm{S}$ become cones around the direction of $\bm{J}$, with a constant precession velocity $\Omega_{\rm prec}$ that we will use to define the timescale of variations of the modulation functions $F$.
 Note that the presence of two spins complicates the evolution of the system, but the picture of a precession cone for $\bm{\ell}$ remains approximately valid (see~\cite{} \SM{[C]} for a classification of the morphologies of precessing trajectories).
@@
 In both interpretations, either the truncated approximation of the Fourier series or  the trigonometric interpolation formulation, these coefficients are related to the coefficients of the inverse DFT.
-If we set $\omega \equiv e^{2i\pi/N}$ and define
+If we set $\omega \equiv e^{2i\pi/N}$ and define \SM{[check convention]}
 \be\label{eq:ykDFT}
@@
 The two cases $\epsilon\pm 1$ correspond simply to a complex conjugation of the coefficients $a_{N,k}^{\epsilon}$.
-Table~\ref{table:stencil} gives the resulting rational values for these coefficients for $N\leq 5$.
+Table~\ref{table:stencil} gives the resulting complex rational values for these coefficients for $N\leq 5$.
 

### 2018-04-05 LISAinference/LISAutils.c

@@ -337,7 +337,7 @@ Syntax: --PARAM-min\n\
     prior->logflat_massprior = 0;
 
     /* set default values for the run settings */
-    run->eff = 0.1;
+    run->eff = 0.5;
     run->tol = 0.5;
     run->consteff = 0;
     run->nlive = 1000;

### 2018-04-05 LISAsim/LISAgeometry.c

@@ -650,7 +650,7 @@ int EvaluateGABmode(
   const ResponseApproxtag responseapprox)  /* Tag to select possible low-f approximation level in FD response */
 {
   double phase = variant->ConstOmega*t + variant->ConstPhi0;
-  
+
   /* Precompute array of sine/cosine */
   for(int j=0; j<4; j++) {
     cosarray[j] = cos((j+1) * phase);
@@ -1044,7 +1044,7 @@ int RestoreInPlaceScaledFactorTDI(
 /*********************** Time-domain response ************************/
 
 /* Processing single mode in amp/phase form through orbital time delay */
-double hOTDAmpPhase(
+static double hOTDAmpPhase(
   const LISAconstellation *variant,    /* Description of LISA variant */
   double* amp,                             /* Output: amplitude */
   double* phase,                           /* Output: phase */
@@ -1073,9 +1073,9 @@ double hOTDAmpPhase(
   *phase = gsl_spline_eval(splinephase, t+delay, accelphase);
 }
 
-/* Functions evaluating yAB observables in time domain */
+/* Functions evaluating yAB observables in time domain - constellation response only */
 /* Note: includes both h22 and h2m2 contributions, assuming planar orbits so that h2-2 = h22* */
-double y12LTDfromh22AmpPhase(
+static double y12LTDfromh22AmpPhase(
   const LISAconstellation *variant,    /* Description of LISA variant */
   gsl_spline* splineamp,                   /* Input spline for h22 TD amp */
   gsl_spline* splinephase,                 /* Input spline for h22 TD phase */
@@ -1112,6 +1112,73 @@ double y12LTDfromh22AmpPhase(
   double factorc = (1./(1.-kn3)) * 0.5*n3Pn3cross;
   double firstdelay = -((kp1 + 1)*variant->ConstL)/C_SI;
   double seconddelay = -(kp2*variant->ConstL)/C_SI;
+
+  /* Values of Y22*h22 + Y2-2*h2-2 at 1 and 2 with delays, and hplus, hcross */
+  /* Note: includes both h22 and h2m2 contributions, assuming planar orbits so that h2-2 = h22* */
+  double A22at1 = gsl_spline_eval(splineamp, t+firstdelay, accelamp);
[…]

### 2018-04-05 LISAsim/LISAgeometry.h

@@ -187,17 +187,6 @@ int RestoreInPlaceScaledFactorTDI(
 
 /* Functions for the response in time domain */
 
-/* Processing single mode in amp/phase form through orbital time delay */
-double hOTDAmpPhase(
-  const LISAconstellation *variant,    /* Description of LISA variant */
-  double* amp,                             /* Output: amplitude */
-  double* phase,                           /* Output: phase */
-  gsl_spline* splineamp,                   /* Input spline for TD mode amplitude */
-  gsl_spline* splinephase,                 /* Input spline for TD mode phase */
-  gsl_interp_accel* accelamp,              /* Accelerator for amp spline */
-  gsl_interp_accel* accelphase,            /* Accelerator for phase spline */
-  const double t);                         /* Time */
-
 /* Basic yslr observables (including orbital delay) from hplus, hcross */
 double y12TD(
   const LISAconstellation *variant,    /* Description of LISA variant */

### 2018-05-25 docs/paper_method.tex
Edits to the paper.

@@
 \newcommand{\tfSPA}{t_{f}^{\rm SPA}}
+\newcommand{\TfSPA}{T_{f}^{\rm SPA}}
 
@@
 
+\SM{[style: below use texttt for models and codes (like \texttt{SpEC})]}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \jgb{If the effective width of of the kernel function $G(f,f’)$ is not quite negligible compared to the scale of significant variations with $f$ it can be useful to extend our approach beyond the leading-order approximation.
-With the waveform represented in phase-and-amplitude form~\eqref{eq:A9}, the elements of~\eqref{eq:FDkernel} may be Taylor expanded in the variable $f'$:
+With the waveform represented in phase-and-amplitude form~\eqref{eq:A9}, the elements of~\eqref{eq:FDkernel} may be Taylor expanded in the variable $f'$: \SM{[change notation, do not use $N$ as sum boundary for $\Psi$]}
 \begin{subequations}\label{eq:expandfprime}
@@
 \end{align}
-This leading-order result~\eqref{eq:transfer0local}-\eqref{eq:transferLlocal} is familiar, as it was derived in the past taking the point of view of monochromatic signals (see e.g.~\cite{} \SM{[C]}).
+This leading-order result~\eqref{eq:transfer0local}-\eqref{eq:transferLlocal} is familiar, as it was derived in the past taking the point of view of monochromatic signals (see e.g.~\cite{} \SM{[C]}) \SM{[rewrite that]}.
 
@@
 We find similar features as for the orbital response: including higher-corrections reduces the errors, but in the high-mass case the amplitude and delay corrections should be added together.
-We also find the same uncontrolled grow of errors at the very end of the signal in the high-mass case, in the region where the amplitude of the signal decays.
+We also find the same growth of errors at the very end of the signal in the high-mass case, in the region where the amplitude of the signal decays.
 
@@
 This treatment is referred to as the fast-slow decomposition, or heterodyning approach.
-If the gravitational wave signal extends only on the narrow frequency band $f \in [f_{*}, (1+\eta) f_{*}]$ with $\eta \ll 1$, then scaling out a carrier frequency of the signal by multpilying by $e^{-2 i \pi f_{*} t}$ will eliminate most of the time variability, allowing to process the signal through the time-domain response and to take a FFT with a Nyquist frequency shifted from $(1+\eta) f_{*}$ to $\eta f_{*}$, i.e.
+If the gravitational wave signal extends only on the narrow frequency band $f \in [f_{*}, (1+\eta) f_{*}]$ with $\eta \ll 1$, then scaling out a carrier frequency of the signal by multiplying by $e^{-2 i \pi f_{*} t}$ will eliminate most of the time variability, allowing to process the signal through the time-domain response and to take a FFT with a Nyquist frequency shifted from $(1+\eta) f_{*}$ to $\eta f_{*}$, i.e.
 with a much smaller number of samples.
@@
 
-Although straightforward and generic, this writing of the Fourier-domain response function will be in general more expansive to compute than the perturbative transfer function approach of Sec.~\ref{sec:formalism}.
+Although straightforward and generic, this writing of the Fourier-domain response function will be in general more expensive to compute than the perturbative transfer function approach of Sec.~\ref{sec:formalism}.
 
@@
 
+\SM{[replace $pp$ everywhere with $\perp\perp$]}
+
+\SM{[check error estimates with single-spin configurations $+$, $\perp$, $-$]}
+
+In this Section, we illustrate the application of the formalism described in~\ref{sec:formalism} to signals from precessing binaries, with modulations induced by the precession.
+We investigate how the separation of the relevant timescales evolves in the late inspiral and after the merger occurs.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In particular, one recovers qualitatively the hierarchy of mode amplitudes that prevails for non-precessing systems, with the modes $h_{22}$ and $h_{2,-2}$ being dominant.
-
-We define $(\alpha, \beta, \gamma)$ are the Euler angles relating the precessing frame to the inertial frame in the $(z,y,z)$ convention.
-The $z$ axis of the inertial frame is chosen to be the direction of $\bm{J}$, the total angular momentum, which is almost constant\footnote{Except in the cases of transitional precession~\cite{Apostolatos+94}, where for high mass ratios and large antialigned spins the orbital angular momentum and the spin can almost cancel each other.}.
-The first two angles, $\alpha$ and $\beta$, are the two spherical angles tracking the direction of dominant radiation reaction, which during the inspiral follows essentially the normal to the orbital plane.
+In the following, we will identify the $z$-axis of the co-precessing frame as the dominant eigenvector of the matrix representing the action of the angular momentum operator acting on the waveform modes, as proposed in~\cite{OShaughnessy+11} (see App.~\ref{app:wigner} for more details).
+
+We define $(\alpha, \beta, \gamma)$ are the Euler angles of the active rotation from the inertial frame to the precessing frame, in the $(z,y,z)$ convention.
+The first two angles, $\alpha$ and $\beta$, are the two spherical angles tracking the direction of the radiation axis, which during the inspiral follows essentially the normal to the orbital plane.
 The last angle $\gamma$ parametrizes the remaining freedom of rotation around this radiation axis.
-An important element of the construction of the precessing frame is the fixing of this third degree of freedom, for which we use the minimal rotation condition~\cite{Boyle+11} \SM{[explain what it means]}
+To fix this third degree of freedom, we use the minimal rotation condition~\cite{Boyle+11}, enforcing the absence of rotation of the precessing frame around the radiation axis.
+In terms of Euler angles, this condition translates to
 \be\label{eq:gammadot}
@@
 \ee
+A natural choice for the $z$ axis of the inertial frame is the direction of $\bm{J}$, the total angular momentum, which is almost constant\footnote{Except in the cases of transitional precession~\cite{Apostolatos+94}, where for high mass ratios and large antialigned spins the orbital angular momentum and the spin can almost cancel each other.}.
 Quite generically, and in particular in the case of simple precession~\cite{Apostolatos+94, Kidder95} but also in the model we will take for the post-merger precession, the radiation axis precesses on a cone around the direction of $\bm{J}$, with $\alpha$ increasing, while $\beta$, the opening angle of the precession cone, is slowly varying.
@@
 \end{subequations}
-where the coefficients $\calD^{\ell}_{mm'}$ are given by Wigner D-matrices~\cite{} \SM{[C]}, which take the expression
+where the coefficients $\calD^{\ell}_{mm'}$ are given by Wigner D-matrices~\cite{Wigner59} as
 \be\label{eq:defWignerD}
-	\calD^{\ell}_{mm'} (\alpha, \beta, \gamma) = e^{im \alpha} d^{\ell}_{mm'}(\beta) e^{im' \gamma}\,,
-\ee
-where the Wigner d-matrix $d^{\ell}_{mm'}(\beta)$ takes the form of a polynomial in $\cos (\beta/2)$, $\sin (\beta/2)$, and acts as an amplitude for the modulation function.
+	\calD^{\ell}_{mm'} (\alpha, \beta, \gamma) = e^{im \alpha} d^{\ell}_{mm'}(\beta) e^{im' \gamma}\,.
+\ee
+Here, the Wigner d-matrix $d^{\ell}_{mm'}(\beta)$ takes the form of a polynomial in $\cos (\beta/2)$, $\sin (\beta/2)$, and acts as an amplitude for the modulation function.
 We refer to App.~\ref{app:wigner} for explicit expressions.
+
+In the following, our objective will be to compute mode-by-mode transfer functions $\calT^{\ell}_{mm'}$, defined as
+\be\label{eq:defprectransfer}
+	\mathrm{FT} \left[ \calD^{\ell *}_{mm'} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm P} \right] (f) \equiv \calT^{\ell}_{mm'}(f) \tilde{h}_{\ell m'}^{\rm P} (f)
+\ee
+such that the complete Fourier-domain inertial-frame waveform will be given as the sum of these individual mode contributions,
+\be\label{eq:defprechIsum}
+	\tilde{h}_{\ell m}^{\rm I} (f) = \sum\limits_{m'=-\ell}^{\ell} \calT^{\ell}_{mm'}(f) \tilde{h}_{\ell m'}^{\rm P} (f) \,.
+\ee
 
@@
 This situation is quite generic, since large misalignments between $\bm{J}$ and $\bm{\ell}$ can only be reached with large spins and large mass ratios.
-From the explicit expressions~\eqref{} \SM{[R]} and~\eqref{} \SM{[R]}, one can see that $d^{\ell}_{mm'}(\beta)$ is greatly suppressed in the limit $\beta \rightarrow 0$ when increasing $|m-m'|$.
+From the explicit expression~\eqref{eq:defWignerdapp}, one can see that $d^{\ell}_{mm'}(\beta)$ is greatly suppressed in the limit $\beta \rightarrow 0$ when increasing $|m-m'|$.
 Intuitively, this means that, in the limit of a small misalignment between frames, the rotation produces mainly mode contributions with the same mode number.
-Similarly, when $\beta$ is constant or slowly varying, \eqref{eq:gammadot} gives $\gamma \simeq - \alpha \cos\beta$, and we have $\calD^{\ell *}_{mm'} (\alpha, \beta, \gamma) \propto e^{i(m' \cos\beta - m) \alpha}$ with $\cos\beta \simeq 1$ for small $\beta$.
+Additionally, when $\beta$ is constant or slowly varying, \eqref{eq:gammadot} gives $\gamma \simeq - \alpha \cos\beta$, and we have in that case
+\be\label{eq:wignerphasesimpleprec}
+	\calD^{\ell *}_{mm'} (\alpha, \beta, \gamma) \propto e^{i(m' \cos\beta - m) \alpha} \,,
+\ee
+with $\cos\beta \simeq 1$ for small $\beta$.
 This shows that the modulation for $m=m'$ has a suppressed phase, while increasing $|m-m'|$ increases the magnitude of the modulation phase.
@@
 Furthermore, rotations leave invariant the combined square amplitude $\calA_{\ell}$ for each $\ell$, as well as the total square amplitude:
-\be
+\be\label{eq:defsumamplitude}
 	\calA^{2} = \sum\limits_{\ell \geq 2}\calA_{\ell}^{2} = \sum\limits_{\ell \geq 2}\sum\limits_{m=-\ell}^{\ell} |h_{\ell m}|^{2} \,.
@@
 
-Several constructions for the precessing frame have been proposed~\cite{} \SM{[C]}.
-In the following, we will use a co-precessing frame based on the dominant eigenvector of the matrix representing the action of the angular momentum operator acting on the waveform modes, proposed in~\cite{OShaughnessy+11} (see App.~\ref{app:wigner} for more details \SM{[R]}).
-
-When approximating the precessing-frame waveform by a non-precessing one, one can use the assumption (exactly valid for non-precessing systems, see~\eqref{eq:symmetryhlminusm} \SM{[R]})
+When approximating the precessing-frame waveform by a non-precessing one, one can use the assumption (exactly valid for non-precessing systems, see~\eqref{eq:symmetryhlminusm})
 \be\label{eq:approxsymmetryhlminusm}
@@
 \ee
-Similarly, for Fourier-domain models, one can neglect either the negative or positive frequency band in the Fourier transform of precessing-frame modes (as in~\eqref{eq:zeronegativef})
+Similarly, in the Fourier domain, one can neglect either the negative or positive frequency band in the Fourier transform of precessing-frame modes (as in~\eqref{eq:zeronegativef})
 \begin{align}\label{eq:approxzeronegativef}
@@
 
-We will introduce mode-by-mode transfer functions $\calT^{\ell}_{mm'}$,
-\be\label{eq:defprectransfer}
-	\mathrm{FT} \left[ \calD^{\ell *}_{mm'} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm P} \right] (f) \equiv \calT^{\ell}_{mm'}(f) \tilde{h}_{\ell m'}^{\rm P} (f)
-\ee
-such that the complete Fourier-domain inertial-frame waveform will be given as the sum of these individual mode contributions,
-\be\label{eq:defprechIsum}
-	\tilde{h}_{\ell m}^{\rm I} (f) = \sum\limits_{m'=-\ell}^{\ell} \calT^{\ell}_{mm'}(f) \tilde{h}_{\ell m'}^{\rm P} (f) \,.
-\ee
-
-When using the assumptions~\eqref{eq:approxsymmetryhlminusm}-\eqref{eq:approxzeronegativef}, one can derive a symmetry relation in the transfer functions themselves.
+When using the approximations~\eqref{eq:approxsymmetryhlminusm}-\eqref{eq:approxzeronegativef}, one can derive a symmetry relation in the transfer functions themselves.
 From the explicit expression of the Wigner matrices, we have indeed (see App.~\ref{app:wigner})
@@
 \end{align}
-When including only the leading harmonic $h^{\rm P}_{22}$ in the precessing-frame waveform (as is done for instance in PhenomP~\cite{Hannam+13}), we have for $f<0$
+
+Further simplifications occur when including only the dominant harmonics $h^{\rm P}_{22}$, $h^{\rm P}_{2,-2}$ in the precessing-frame waveform (as is done for instance in PhenomP~\cite{Hannam+13}, as well as in our toy model~\ref{subsec:precmodel}), we have in this case for $f<0$
 \begin{subequations}
@@
 \end{subequations}
-In general, when including a superposition of modes with different values of $m'$ in the model (for instance, the modes $h^{\rm P}_{22}$ and $h^{\rm P}_{21}$), this symmetry exists only at the level of the mode contributions, and there is no relation such as~\eqref{eq:symmetryhIfor22only}  at the level of $\tilde{h}^{\rm I}_{\ell m}$, as the complete mode reconstruction includes factors $(-1)^{\ell+m+m'}$ in the sum.
-
-In the notation of our formalism of Sec.~\ref{sec:formalism}, 
-\be\label{eq:defmodulationprec}
-	s(t) = F(t) h(t) \,,
-\ee
-with $s$ being a mode contribution to $h^{\rm I}_{\ell m}$, $h$ one of the precessing-frame modes $h^{\rm P}_{\ell m'}$, and the modulation $F$ being a Wigner matrix $\calD^{\ell *}_{mm'}$.
-In the absence of delays, the convolution with a frequency-dependent kernel~\eqref{eq:FDkernel} becomes a simple convolution in the Fourier-domain,
-\be\label{eq:convolution}
-	\tilde{s} (f) = \int \ud f' \, \tilde{h}(f-f') \tilde{F} (f') \,.
-\ee
-Our objective is therefore to compute these convolution integrals to obtain the Fourier-domain transfer functions $\calT^{\ell}_{mm'}$.
-We note an important point about the structure of these comvolution integrals.
-In presence of the usual behaviour of precession around an almost constant $\bm{J}$, $\alpha$ is monotonously increasing and the modulation functions have the phase evolution $\calD^{\ell *}_{mm'} (\alpha, \beta, \gamma) \propto e^{i(m' \cos\beta - m) \alpha}$.
-For the 22 mode in the P-frame, $m'=2$, the coefficient of $\alpha$ is close to $0$ for small $\beta$ for $m=2$, and increasingly positive as we go down from $m=1$ to $m=-2$.
+With $m' = 2$ in the approximate phase given in Eq.~\eqref{eq:wignerphasesimpleprec}, the coefficient of $\alpha$ is close to $0$ for small $\beta$ for $m=2$, and increasingly positive when going down from $m=1$ to $m=-2$.
 In our Fourier convention~\eqref{eq:defFT}, the Fourier transforms $\tilde{F}$ thus have most of their support on negative frequencies.
-This means that the support of the convolution integral~\eqref{eq:convolution} extends to the right side for $\tilde{h}$, i.e.
+This means that the support of the convolution integral~\eqref{eq:FDkernel} extends to the right side for $\tilde{h}$, i.e.
 $f-f' > f$.
-This point will become important when considering the high-frequency part of the signal.
-
-\subsection{Previous approaches}
-\label{subsec:precpreviousapproaches}
-
-The above decomposition~\eqref{eq:wignerrot}, and simplifying assumptions~\eqref{eq:approxsymmetryhlminusm}-\eqref{eq:approxzeronegativef} for the precessing-frame waveform, have been widely used for the existing models of waveforms from precessing binaries.
-For the inspiral phase of the SEOBNRv3 model~\cite{Pan+13,BTB16}, the frame is assumed to follow the instantaneous orbital plane and the waveform in this frame is represented by a spin-aligned expression, with spin-aligned components that are left to vary according to the precessing dynamics.
-The ringdown phase of the signal is treated differently, the waveform being first rotated back to an inertial frame and then matched to a superposition of quasi-normal modes in this final frame.
-The PhenomP model~\cite{Hannam+13} uses the decomposition~\eqref{eq:wignerrot} above (although directly in the Fourier domain, as we will detail below) with a precessing-frame waveform given by a spin-aligned PhenomD waveform~\cite{Khan+15,Husa+15}.
-The SUA waveforms~\cite{KCY14,Chatziioannou+16,Chatziioannou+17}, focused on the inspiral phase, also use the decomposition~\eqref{eq:wignerrot} and a post-Newtonian non-precessing model for the modes $h_{\ell m}^{\rm P}$.
-
-To account for previous approaches to the problem of producing precessing waveforms directly in the Fourier domain, avoiding the use of a FFT \SM{[check acronym]}, it is useful to introduce schematically three levels of approximation (dropping the mode indices):
-\begin{itemize}
-	\item unstable SPA: $\calT(f) \tilde{h}^{\rm P}(f) = \mathrm{SPA}\left[ \calD^{*} h^{\rm P} \right](f)$,
-	\item SPA, $0^{\text{th}}$-order SUA, PhenomP: $\calT(f) = \calD^{*}(\tfSPA) $,
-	\item SUA: $\calT(f) = \frac{1}{2} \sum\limits_{k} a_{k} \calD^{*}(\tfSPA \pm k \Tf^{\rm SPA})$.
-\end{itemize}
-
-In the first option, one applies directly the SPA, as described in Sec.~\ref{subsec:SPA}, to the product of the modulation and the signal.
-This is known (see e.g.~\cite{KCY13}) to lead to possible pathologies, since the prefactor $1/\sqrt{\ddot{\phi}_{\rm prec}}$ can blow up due the precessing contributions to the phase, and is not used in waveform modelling.
-
-The second option corresponds to simply multiplying the Fourier-domain signal by the modulation function evaluated at the time $\tfSPA$ as defined in~\eqref{eq:deftfSPA}.
-In the SUA formalism of Ref.~\cite{KCY14}, this amounts to using a stencil reduced to a single point, i.e.
-using $a_{0,0} = 1$ in~\eqref{} \SM{[R]}.
-It is equivalent to the leading order of our formalism, as explained in Sec.~\ref{} \SM{[R]}, with the difference that we use a more general definition of $t_{f}$ (see~\eqref{eq:deftf}) that extends through the merger and ringdown.
-\SM{[Mention O'Shaugnessy single-spin FD waveforms]}.
-The PhenomP model~\cite{Hannam+13} implictly uses this level of approximation, although with the notable difference that a frequency-based expression of the modulation is used, allowing to cover the whole frequency band.
-The precessing-frame evolution is modelled using post-Newtonian expressions for the Euler angles $(\alpha^{\rm PN}, \beta^{\rm PN}, \gamma^{\rm PN}) (\omega)$, that are obtained as functions of the orbital frequency $\omega$ in the limit of a small opening angle of the precession cone and at the spin-orbit level~\cite{Hannam+13} (see also~\cite{BBF11, MBBB13}).
-Using the SPA-inspired correspondence~\eqref{} \SM{[R]} between the orbital and Fourier-domain frequency, one then writes
-\be\label{eq:precPhenomP}
-	\calT^{\ell}_{mm'}(f) = \calD^{\ell *}_{mm'}(\alpha^{\rm PN}, \beta^{\rm PN}, \gamma^{\rm PN}) \left( \omega=\pi f\right) \,.
-\ee
-When the SPA is valid for the underlying signal $h^{\rm P}_{\ell m}$, the condition $\omega = \pi f$ is equivalent by definition to $t=t^{\rm SPA}_{f}$, but writing things in terms of frequency gives an effective prescription that can cover the whole Fourier-domain frequency band, including the merger and ringdown.
-Note that this procedure amounts to using PN \SM{[check acronym]} results outside of their range of validity, since in the merger-ringdown phase both the PN perturbative treatment and the SPA approximation break down.
-In particular, seen as a function of time, $\omega^{\rm PN}(t)$ blows up at merger.
-In practice, one finds however that the frequency-based expressions~\eqref{eq:precPhenomP} are mildly varying as a function of frequency, and the agreement of the model with numerical relativity waveforms is good~\cite{} \SM{[C]}.
-
-In the SUA formalism proposed in~\cite{} \SM{[C]}, one extends the SPA to incorporate the leading-order correction during the inspiral phase.
-The transfer function is then computed by evaluating the modulation on a stencil of times centered around the time-of-frequency given by the SPA, following \SM{[say in words:tf-tfSPA]}
-\begin{widetext}
-\be
-	\calT(f) = \frac{1}{2} \sum\limits_{k=0}^{N} a_{N,k} \left[\calD^{\ell *}_{mm'}(\tfSPA + k \Tf^{\rm SPA}) + \calD^{\ell *}_{mm'}(\tfSPA - k \Tf^{\rm SPA}) \right] \,.
-\ee
-\end{widetext}
-The definition of $\Tf^{\rm SPA}$ was given in~\ref{} \SM{[R]}.
-As described in Sec.~\ref{} \SM{[R]}, this approach is equivalent to our formalism when keeping only the quadratic phase correction, but is a priori limited to the inspiral phase of the signal through the definitions of the times $t^{\rm SPA}_{f}$ and $T^{\rm SPA}_{f}$.
-For inspiral waveforms, Ref.~\cite{} \SM{[C]} showed that this treatment improves the accuracy with respect to the local approximation.
-
-The time-domain SEOB~\cite{} \SM{[C]} model includes the modes $h^{\rm P}_{2m}$ with $|m|=2$ and $|m|=1$ for the inspiral, and the ringdown constructed in the inertial frame contains a superposition of all modes $h^{\rm I}_{2 m}$ by construction.
-In the PhenomP~\cite{} \SM{[C]} model, only the dominant harmonics $h^{\rm P}_{22}$ and $h^{\rm P}_{2,-2}$ were included in the precessing-frame waveform.
-The SUA~\cite{} \SM{[C]} model can be constructed using any PN non-precessing model for an inspiral-only waveform, including higher harmonics.
-\SM{[Check what is implemented in LAL ?
-HM or not ?]}
+This point will become important when considering the high-frequency part of the signal in Sec.~\ref{subsec:trigopoly}.
 
@@
 
-\SM{[qualify how bad the model is with respect to full NR]}
-
-\SM{[explain shortcomings of pure SEOB in post-merger, problem of smoothness]}
-
-We now present our toy model for IMR waveforms of precessing binaries, which we will use in the following to illustrate the accuracy of the Fourier-domain transfer functions at various orders of approximation.
+\SM{[qualify how bad the model is with respect to full NR ?]}
+
+\SM{[explain shortcomings of pure SEOB in post-merger, problem of $\beta$ oscillating ?]}
+
+We wish to apply the formalism presented in Sec.~\ref{sec:formalism} to IMR waveforms of precessing binaries, exploring the separation of the timescales involved through the inspiral and merger, and assessing the relevance of the higher-order corrections summarized in~\ref{subsec:executivesummary}.
+To this end, we will use a simplified model for the precession, allowing a number of simplifications and idealizations.
+
+%To this end, we will use a simplified model for IMR waveforms of precessing binaries, which we will use in the following to .
 Since our purpose in this article is not to propose a full-fledged waveform model, but rather to investigate wether the formalism of Sec.~\ref{sec:formalism} is relevant to tackle the merger part of the signals, and wether its application reduces the errors in the reconstruction of transfer functions, we will allow for a number of idealizations and simplifications in this toy model.
 
-Although the frame decomposition and precessing-frame prescription is well understood for the inspiral, where the precessing frame roughly follows the orbital plane of the binary (see~\ref{} \SM{[R]}), we wish to investigate the merger and ringdown part of the signal as well, and need a way to extend this intuitive picture.
-
-\SM{[acronym QNM]}
-
-For this extension, one needs to define a precessing frame after merger, when the orbital plane has lost any meaning.
-This can be done naturally when using a frame prescription that is not tied to the dynamics but directly extracted from the gravitational waveform, such as~\eqref{} \SM{[R]}.
-Investigating the post-merger behaviour of this frame requires numerical relativity simulations of precessing binaries.
-Ref.~\cite{} \SM{[C]} proposed a qualitative model where the frame vector essentially keeps precessing around the final $\bm{J}$, but transitions to a faster precession rate.
-More precisely, they observe that the interpretation of the post-merger waveform as an excitation of a superposition of quasinormal modes suggests that the frame should evolve according to the difference of the leading QNM frequencies.
-In their simulations, they find indeed a precession rate around the final total angular momentum $\hat{\bm{J}}_{\rm final}$, asymptoting to the angular velocity
+We want to be able to investigate signals with a long inspiral phase, beyond the range covered by numerical relativity simulations, and we want to include the merger and ringdown phase.
+We will use the following three ingredients for this simplified model:
+\begin{itemize}
+	\item PhenomD~\cite{Husa+15, Khan+15} for the IMR Fourier-domain waveform in the precessing frame;
+	\item SEOBNRv3~\cite{Pan+13, BTB16} for the Euler angles during the inspiral;
+	\item an effective extension of the Euler angles post-merger based on~\cite{OShaughnessy+12} (see~\eqref{eq:OmegaframeQNM} below).
+\end{itemize}
+
+For the precessing-frame waveform, we make all the simplifying assumptions described above in Sec.~\ref{subsec:precdef}: we approximate it as a non-precessing waveform, further enforcing the approximations~\eqref{eq:approxsymmetryhlminusm} and~\eqref{eq:approxzeronegativef} and limiting our analysis to the dominant harmonics\footnote{See~\cite{London+17} for a recent extension of the model to include higher harmonics.} $h^{\rm P}_{22}$ and $h^{\rm P}_{2,-2}$.
+We will use the PhenomD waveform model~\cite{Husa+15, Khan+15}, an aligned-spin Fourier-domain model publicly available in the LIGO Algorithm Library (LAL) that covers the inpiral, merger and ringdown for binaries with generic spin magnitude.
+The amplitude and phase of the waveform are produced as piecewise analytical functions of the frequency.
+Using a Fourier-domain approximant that is smooth by construction avoids the Gibbs oscillations induced by the tapering of of finite-length time-domain waveforms, and will allow us to easily take Fourier-domain derivatives of the waveform\footnote{Note that the PhenomD amplitude and phase are smooth on three separate frequency bands (inspiral, intermediate, and ringdown) with junction conditions that are only of class $C^{1}$.
+To avoid spurious discontinuities, we introduce decaying corrective terms of the form $(f-f_{\rm join})^{2} e^{-\lambda(f-f_{\rm join})^{2}}$ on each side of $f_{\rm join}$, the junction frequency between any two given bands, resulting in functions of class $C^{2}$.}.
+
+For the frame trajectory in the inspiral phase, we use an SEOBNRv3 waveform~\cite{Pan+13, BTB16}, that incorporates all degrees of freedom of both spins.
+As the inspiral of the SEOBNRv3 waveform is constructed by using the instantaneous orbital plane of the dynamics as the precessing frame, the frame extraction procedure of~\cite{OShaughnessy+11} simply returns the normal to the orbital plane as the radiation axis.
+The frame thus presents oscillatory features of small amplitude at the orbital timescale, due to the nutation of the orbital plane, which we smooth out using a Gaussian filtering with width based on the orbital phase as in~\cite{Blackman+17a}.
+
+For the post-merger precession, the intuitive interpretation of the precessing frame as roughly following the orbital plane of the binary is lost.
+However, the prescription of~\cite{OShaughnessy+11} to extract the radiation axis is based entirely on the waveform itself, and can be applied to the ringdown as well.
+For $\ell=2$, analyzing numerical relativity waveforms, Ref.~\cite{OShaughnessy+12} described a qualitative model where the radiation axis essentially keeps precessing around the final angular momentum $\bm{J}$, but transitions to a faster precession rate, with an angular velocity determined by the quasinormal mode (QNM) frequencies of the remnant black hole as
 \be\label{eq:OmegaframeQNM}
@@
 \ee
-The behaviour of the opening angle of the precessing cone $\beta$, however, is harder to interpret.
-Although it usually shows some transition at merger (to a lower value), it does not show an exponential decay as the interpretation above would suggest.
-We found that overall this picture is at least qualitatively correct for the precessing waveforms	 currently available in the SXS catalog.
-\SM{[allowed to make such a vague statement about catalogue waveforms ?]}
-
-Note however that this model should be taken with caution, since the parameter space currently accessible to numerical simulations of precessing systems is limited in spin mass ratio.
-The behaviour of systems with a close-to-extremal remnant spin largely remains to be explored.
-In addition, systems with large mass ratios and large anti-aligned spins on the primary black hole can have a total angular momentum dominated by the primary spin instead of the orbital angular momentum (e.g.
-in the case of transitional precession~\cite{} \SM{[C]}).
-This can lead to the remnant spin being anti-aligned with respect to the precessing frame axis prior to merger.
-This could conceivably lead to a different structure in the excitation of the quasi-nomal modes, and to a qualitatively different frame behaviour.
-Assessing the limits of validity of this model would require a more extensive coverage of the parameter space by numerical relativity simulations and would go beyond the scope our present study.
-
-The basic ingredients of our model \SM{[apologize for the Frankenstein-ism !]} will be: \SM{[move most of all this to an appendix]}
-\begin{itemize}
-	\item PhenomD~\cite{} \SM{[C]} for the precessing-frame waveform;
-	\item SEOB for Euler angles during the inspiral;
-	\item an effective extension of the Euler angles post-merger based on~\eqref{eq:OmegaframeQNM}.
-\end{itemize}
-For the precessing-frame waveform, in keeping with the approximation~\eqref{} \SM{[R]} for the precessing-frame waveform we will use a slightly modified version of the PhenomD waveform model~\cite{} \SM{[C]}, a Fourier-domain model that covers the inpiral, merger and ringdown for binaries with generic aligned spins.
-The waveform is produced as piecewise analytical expressions\footnote{The PhenomD model is publicly available in the LIGO Algortitm Library (LAL).}.
-We limit ourselves to the dominant harmonics $h^{\rm P}_{22}$ and $h^{\rm P}_{2,-2}$ (see~\cite{} \SM{[C]} for a recent extension of the model to include higher harmonics).
-Using a Fourier-domain approximant that is smooth by construction avoids the problems arising from the tapering time-domain waveforms, as we will detail below, and will allow us to easily take Fourier-domain derivatives of the waveform.
-Since the PhenomD amplitude and phase are only of class $C^{1}$, to avoid spurious discontinuities we introduce decaying corrective terms of the form $(f-f_{\rm join})^{2} e^{-\lambda(f-f_{\rm join})^{2}}$ on each side of $f_{\rm join}$, the junction frequency between two bands.
-Here $\lambda$ is a large parameter ensuring that the modification remains local.
-The modificiations have very little effect overall but artificially enforce the continuity of second derivatives.
-
-For the frame trajectory in the inspiral phase, we extract the dominant-eigenvector frame as defined in~\eqref{} \SM{[R]} from an SEOB waveform.
-Since the SEOB waveform is constructed by using the instantaneous orbital plane of the dynamics as the precessing frame, the frame extraction procedure simply gives back the orbital plane (except very close to merger where the waveform starts to smoothly transition to the post-merger QNM superposition).
-The frame thus presents oscillatory features of small amplitude at the orbital timescale, due to the nutation of the orbital plane \cite{} \SM{[C]}.
-
-For the extension of the frame trajectory post-merger, we do not rely on the PhenomP procedure as it is merely effective \SM{[be clearer on the limitation here]}.
-Furthermore, we found that the ringdown attachment procedure, as implemented in SEOB, leads to frame trajectories that show oscillations in all the Euler angles when the spins are anti-aligned.
-These oscillations increase in amplitude as the spins go to larger misalignments, and are already present well beneath the regime of transitional precession.
-Since we checked that the qualitative behaviour~\eqref{eq:OmegaframeQNM} holds rather well for numerical relativity simulations in regions of the parameter space where SEOB starts to show these oscillations, and since the SEOB post-merger ringdown attachment is an effective representation, we chose to use an artificial post-merger extension of the frame, following
-\be\label{eq:alphaQNM}
-	\alpha_{\rm post-merger}(t) = \alpha(t_{\rm peak}) + (\omega_{220}^{\rm QNM} - \omega_{210}^{\rm QNM})(t-t_{\rm peak}) \,.
-\ee
-Since we do not have yet a quantitative model of the behaviour of the opening angle of the precessing cone $\beta$, we simply set it to a constant value post-merger, taken to be its value at merger.
-The third Euler angle $\gamma$ is extended using the minimal rotation condition~\eqref{} \SM{[R]}.
-For the QNM frequencies as a function of the final spin, we use values tabulated in Ref.~\cite{} \SM{[C]}.
-The final spin is taken to be the same as in the SEOB code~\cite{} \SM{[C]}, where the final-spin fit formula of Ref.~\cite{} \SM{[C]}, built for spin-aligned systems, is applied to the spin components \SM{[check that and complete]}.
-Our simple extension is an idealization, especially for the angle $\beta$, and is not to be considered quantitatively accurate.
-However, it allows us to capture what we believe to be the dominant post-merger behaviour, in the sense that the frame is set to rotate with a velocity of the right order of magnitude, which is crucial for us since we want to investigate how the separation between the orbital (or $h^{\rm P}_{\ell m}$) timescale and the precessing timescale is challenged post-merger.
-We show in Fig~\ref{} the frequencies $\omega_{220}^{\rm QNM}$, $\omega_{210}^{\rm QNM}$ and their difference $\Omega_{\rm frame}$.
-The behaviour for near-to-extremal spin of the remnant would require more exploration, as the QNM frequencies vary a lot when approaching extramality.
-\SM{[add plot of QNM frequency and frequency diff according to final spin]}.
-Fig.~\ref{fig:prectoymodel} presents a comparison of the frame trajectory at merger for the SEOB, NR waveforms and for our model.
-The example catalog waveform is SXS:BBH:0058~\cite{} \SM{[C]}, which presents a rather high mass ratio of $q=5$, with a single spin in-plane of $\chi_{1} = 0.5$.
-The NR and SEOB post-merger frames disagree mainly due to a different value of the final spin $\chi_{f}$\footnote{Work is underway to update the fit formulas used to compute $\chi_{f}$ in the SEOB code, which is expected to reduce these disagreement~\cite{}.
-\SM{[C]} \SM{[work in preparation by SEOB people, essentially already done]}}, which leads to a different $\Omega_{\rm frame}$.
-Given their different value of $\chi_{f}$, however, both agree with the qualitative prescription~\eqref{eq:OmegaframeQNM}.
-The oscillations developing after $t\sim 50 M$ occur in a regime where the overall waveform amplitude has decayed to small values, and are not important in that regard.
-Our model follows well the SEOB behaviour for $\alpha$ and $\gamma$, but departs more for $\beta$ as it ignores its variation at merger.
-
-To combine our various waveform ingredients, we first ensure that the PhenomD and SEOB waveforms are aligned in time, by performing an inverse Fourier transform of PhenomD and applying a time shift to both waveforms so that the peak amplitude occurs at $t=0$ (taken for SEOB to be the $\ell=2$ combined square amplitude~\eqref{}, and for PhenomD just the peak of the 22 mode for PhenomD).
-We apply to the resulting frame trajectory a Gaussian filtering following the same procedure as the one used for numerical-relativity surrogate of Ref.~\cite{} \SM{[]}: the scale of the filter is set to a fixed width in waveform phase \SM{[detail what we used exactly]}.
-This filtering ensures a continuity of the modulation function and their derivatives at the point of attachment.
-It also eliminates the nutation of the frame at the orbital timescale, which are generically less pronounced in the frame extracted following~\eqref{} \SM{[R]} from numerical relativity simulations than for the normal to the orbital plane.
-
-Note an important caveat of our approach: the application of the perturbative formalism presented in Sec.~\ref{sec:formalism} relies on the smoothness of both the time-domain modulation and the Fourier-domain signal.
-For the time-domain modulation, the assumption of the separation of timescales~\eqref{} \SM{[R]} relies on the identification of a characeteristic frequency $\Omega_{\rm prec}$, which is based on the time derivative of $F$.
-It is always possible to add and remove arbitrarily small oscillations in $F$ that would give arbitrarily large contributions to its derivatives, hence affecting any local computation of $\Omega_{\rm prec}$.
-One can consider that our approach is to be applied to such a slowly varying approximation of $F$, that can be obtained by filtering away rapidly-varying, small-amplitude features.
-However, we should keep in mind that any smoothing procedure could possibly eliminate  some information present in the original waveforms.
-
-For the smoothness of the signal in the Fourier domain, one has to keep in mind that, for waveforms computed starting with a time-domain waveform, Gibbs oscillations can be mitigated by properly tapering the waveform in time domain before performing the FFT, however taking Fourier-domain derivatives magnifies this oscillations, to the point that they would dominate among others the timescale $\Tf$ (unless one takes much longer waveforms to push this problem to lower frequencies).
-This is avoided in our case by working with a pure Fourier-domain PhenomD waveform that is smooth by construction, but would become important if one were to apply our formalism to the FFT of NR or SEOB waveforms.
-
-Another caveat is that for the precessing-frame waveform, the assumption~\eqref{eq:approxsymmetryhlminusm} is only approximate.
-Mode asymmetries are present in the late inspiral and merger phases of precessing binaries, as shown by numerical relativity simulations (see e.g.~\cite{} \SM{[C]}).
-These asymmetries play a role in the emission of linear momentum and in the resulting kick of the black hole remnant, and they also result in amplitude and phase variations on the orbital timescale.
-
-In the following, we will illustrate our formalism by applying it to three examples.
+This translates in general in a significant acceleration of the precession post-merger with respect to the inspiral (see Fig.~\ref{fig:precmodel}).
+For the opening angle of the precessing cone $\beta$, Ref.~\cite{OShaughnessy+12} proposes an exponential decay driven similarly by the difference between damping frequencies.
+We found by a non-exhaustive \SM{[allowed to say that ?
+the catalog is vast now]} inspection of precessing waveforms currently available in the SXS catalog (see e.g.~\cite{SXScatalog, Mroue+12, Mroue+13}) that this picture is at least qualitatively correct for $\ell=2$, with $\beta$ appearing to transition to a lower value rather than decaying all the way to zero.
+By contrast, we found that the current ringdown prescription in the SEOBNRv3 model can cause a qualitatively different behaviour, with oscillations in $\beta$.
+\SM{[motivate this with final spin estimate as well, if we update this ?]}
+%and its estimate for the spin of the final remnant~\cite{} would require to be updated (see e.g.~\cite{}) as shown in Fig.~\ref{fig:precmodel}.
+
+In our model, after reaching the time of peak amplitude as defined in~\eqref{eq:defsumamplitude}, we will use the prescription~\eqref{eq:OmegaframeQNM} for the post-merger frame precession and simply keep $\beta$ constant,
+\begin{subequations}
+\begin{align}\label{eq:eulerQNM}
+	\alpha_{\rm post-merger}(t) &= \alpha(t_{\rm peak}) + (\omega_{220}^{\rm QNM} - \omega_{210}^{\rm QNM})(t-t_{\rm peak}) \,, \\
+	\beta_{\rm post-merger}(t) &= \beta(t_{\rm peak}) \,, \\
+	\gamma_{\rm post-merger}(t) &= \gamma(t_{\rm peak}) - (\alpha(t) - \alpha(t_{\rm peak}) )\cos \beta \,,
+\end{align}
+\end{subequations}
+where we use the minimal-rotation condition~\eqref{eq:gammadot} for a constant $\beta$.
+For the QNM frequencies as a function of the final spin, we use values computed in Ref.~\cite{Berti+05}.
+The final spin is taken to be the same as the one computed internally to the SEOBNRv3 code~\cite{Pan+13, BTB16}, where the final-spin fit formula of Ref.~\cite{} \SM{[C]}, built for spin-aligned systems, is applied to the spin components projected on the orbital angular momentum $\bm{L}$ at merger.
+
+Fig.~\ref{fig:precmodel} presents a comparison of the frame trajectory at merger for the SEOBNRv3, NR waveforms and for our model~\eqref{eq:eulerQNM}.
+The example catalog waveform is SXS:BBH:0058, generated with SpEC~\cite{SXScatalog, SpEC, Mroue+12, Mroue+13}, with a mass ratio of $q=5$, and a single in-plane spin of $\chi_{1} = 0.5$.
+The NR and SEOB post-merger frames disagree mainly due to a different value of the final spin $\chi_{f}$\footnote{The fitting formulas used to compute $\chi_{f}$ in the next version of the SEOB code, which is expected to reduce or remove this disagreement~\cite{Ossokine+18}.
+\SM{[update this reference in prep.]}}, which leads to a different $\Omega_{\rm frame}$.
+Given their different value of $\chi_{f}$, however, both agree with the qualitative description~\eqref{eq:OmegaframeQNM}.
+The oscillations developing after $t\sim 50 M$ occur in a regime where the overall waveform amplitude has decayed to small values.
+Our simple model~\eqref{eq:eulerQNM} follows well the SEOBNRv3 behaviour for $\alpha$ and $\gamma$, but departs more for $\beta$ as we do not model its variation at merger.
+
+We would now like to underline a number of caveats and limitations of our simplified model.
+First, it is clear from Fig.~\ref{fig:precmodel} that the acceleration of the precession post-merger is going to be the most challenging feature for the separation of timescales at the basis of Sec.~\ref{sec:formalism}.
+While we checked the qualitative soundness of the decomposition outlined in Sec.~\ref{subsec:precdef} and of the post-merger precession~\eqref{eq:OmegaframeQNM} in some of the available numerical relativity waveforms~\cite{SXScatalog}, it is important to stress that there is no guarantee that another prescription for the precessing frame would not yield a better separation of timescales.
+Secondly, approximating the precessing-frame waveform by a spin-aligned one comes with known limitations, ignoring for instance mode asymmetries departing from~\eqref{eq:approxsymmetryhlminusm} as shown in~\cite{Boyle+14}.
+Thirdly, more exploration of the parameter space would be needed, as the QNM frequencies entering~\eqref{eq:OmegaframeQNM} vary rapidly for large $\chi_{f}$.
+Finally, we ensure the smoothness both of the Fourier-domain precessing-frame waveform and of the time-domain modulation, allowing us to take the derivatives required by the formalism of Sec.~\ref{sec:formalism}, which is an idealization.
+
+Leaving aside the question of a more in-depth investigation of the best representation of precessing waveforms in their post-merger phase, we will proceed to investigating the separation of timescales in three chosen examples.
 The parameters of these example cases are summarized in Table~\ref{table:precparams}.
-We consider a mass ratio of $q=4$, close-to-maximal spins of $\chi_{1} = \chi_{2} = 0.95$, and vary both spins misalignmnent angles to be $\pi/6$ (close to spin-aligned), $\pi/2$ (spins in the plane) and $5\pi/6$ (close to anti-aligned spins).
-Table~\ref{table:precparams} also shows the remnant spins that are internally computed in both the SEOB and the PhenomD/PhenomP codes, as well as the QNM frequencies and the resulting frame precession frequency $\Omega_{\rm frame}$.
-Note that the case $--$ shows a larger variation in the direction of $\bm{J}$, and the spin of the remnant is almost 0, due to the fact that the spin is large and antialigned, with a large mass ratio.
+We consider a mass ratio of $q=4$, close-to-maximal spins of $\chi_{1} = \chi_{2} = 0.95$, and vary both spins misalignmnent angles to be $\pi/6$ (case labeled $++$, close to aligned spins), $\pi/2$ ($\perp\perp$, spins in the plane) and $5\pi/6$ ($--$, close to anti-aligned spins).
+Table~\ref{table:precparams} also shows the remnant spins that are internally computed in both the SEOBNRv3 and the PhenomD/PhenomP codes, as well as the QNM frequencies and the resulting frame precession frequency $\Omega_{\rm frame}$.
+Note that the case $--$ shows a larger variation in the direction of $\bm{J}$, and the spin of the remnant is almost 0, due to the fact that the spin is large and antialigned with a large mass ratio.
 This example therefore serves the purpose of lying at the edge of validity of the picture of standard precession along a cone around an almost-fixed direction, and the standard hierarchy between mode contributions is not valid in this case.
@@
   \includegraphics[width=.98\linewidth]{plots/eulerSXS0058_py.pdf}
-  \caption{Evolution of Euler angles $(\alpha, \beta, \gamma)$ near merger for the example waveform SXS:BBH:0058~\cite{} \SM{[C]}.
+  \caption{Evolution of Euler angles $(\alpha, \beta, \gamma)$ near merger for the example SpEC waveform SXS:BBH:0058~\cite{SpEC, SXScatalog, Mroue+12, Mroue+13} \SM{[C]}.
 Full, dashed and dotted lines represent the NR, SEOB and extended SEOB waveforms respectively.
@@
 Add Euler angles produced by IFFT of PhenomP ?]}}
-  \label{fig:prectoymodel}
+  \label{fig:precmodel}
 \end{figure}
@@
 The table gives also the final spins magnitudes (for SEOB as well as PhenomD for comparison), angles between the initial and final angular momenta $\theta(\bm{J}_{i}, \bm{J}_{f})$, as well as QNM frequencies and frame rotation velocity used for the post-merger extension.
-\SM{[To be updated with longer waveforms, 10Hz 10Msol.]} \SM{[add misalignment angles $(\ell, \bm{J}_{i})$]}}\label{table:precparams}
+\SM{[To be updated with longer waveforms, 10Hz 10Msol.
+?]} \SM{[add misalignment angles $(\ell, \bm{J}_{i})$?]}}\label{table:precparams}
 %\begin{tabular}{C{.2\columnwidth}C{.24\columnwidth}C{.15\columnwidth}}
@@
 	\hline\hline
-	Case && ++ && $pp$ && $--$ \\
+	Case && ++ && $\perp\perp$ && $--$ \\
 	\hline
@@
 
-\subsection{Estimates for the magnitude of higher-order corrections}
+\subsection{Estimates for the separation of timescales}
 \label{subsec:sizecorrPrec}
 
-\SM{[Add full expressions for the $\epsilon$ quantities]}
+\begin{figure}
+  \centering
+  \includegraphics[width=.98\linewidth]{plots/postmergeromega_py.pdf}
+  \caption{Quasi-normal mode frequencies for the 0th overtone of the modes $22$ and $21$, as a function of the dimensionless spin of the final black hole $\chi_{f}$.
+We also show their difference $\Omega_{\rm frame}$ (see~\eqref{eq:OmegaframeQNM}), and the ratio $\Omega_{\rm frame} / \omega^{\rm QNM}_{220}$ characteristic of the separation of timescales between the the phase of the modulation~\eqref{eq:wignerphasesimpleprec} and the P-frame waveform phase.}
+  \label{fig:precmodel}
+\end{figure}
+
+%We show in Fig~\ref{} the frequencies $\omega_{220}^{\rm QNM}$, $\omega_{210}^{\rm QNM}$ and their difference $\Omega_{\rm frame}$.
+The behaviour for near-to-extremal spin of the remnant would require more exploration, as the QNM frequencies vary a lot when approaching extremality.
+
+\SM{[add plot of QNM frequency and frequency diff according to final spin ?]}.
 
@@
 
-We now investigate the magnitude of higher-order corrections in the case of precessing binaries, using the error estimates $\epsilon_{\Psi 2}$, $\epsilon_{A1}$ and $\epsilon_{A 2}$ defined in~\eqref{eq:deffom}.
-We would like to stress that these figures are to be understood as merely order-of-magnitude estimates, since they compute the error of each approximation by gauging the magnitude of the first corrective term in a Taylor series.
-In particular, they rely on computing derivatives of the time-domain modulation, which can be sensitive to oscillatory features.
-
-First, we would like to stress that the modulation functions $\calD^{\ell *}_{mm'}$. As shown in~\ref{app:Wigner}, their phase dependence is. \SM{[complete]}
+We now turn to the magnitude of higher-order corrections in the case of precessing binaries, as can be estimated by the quantities $\epsilon_{\Psi 2}$, $\epsilon_{A1}$ and $\epsilon_{A 2}$ introduced in~\eqref{eq:deffom}.
 
@@
 \ee
-For each mode contribution $\calD^{\ell *}_{mm'}$, we have
+For each mode contribution $\calD^{\ell *}_{mm'}$, we have \SM{[said before]}
 \be
@@
 which gives an additional geometric factors of $(m' \cos\beta - m)$.
-Under our assumptions of simple precession, ignoring radiation reaction, and using the leading-order expressions~\eqref{} \SM{[R]} for the Fourier-domain timescales, we finally obtain the following error estimates for individual modes:
+Under our assumptions of simple precession, ignoring radiation reaction, and using the leading-order expressions~\eqref{} \SM{[R]} for the Fourier-domain timescales, we finally obtain the following error estimates for individual modes \SM{[recall what $m$, $m'$ mean]}:
 \begin{subequations}
@@
 The overall $\nu$ scalings indicate a better separation of timescales for larger mass ratio systems.
-These expressions also show that, since the total angular momentum appear as a factor in $\Omega_{\rm prec}$, higher-order corrections will be larger for spin-aligned systems than for antialigned spins \SM{[language: antialigned or anti-aligned ?]}, an effect that becomes significant for large mass ratios.
+These expressions also show that, since the total angular momentum appear as a factor in $\Omega_{\rm prec}$, higher-order corrections will be larger for spin-aligned systems than for anti-aligned spins, an effect that becomes significant for large mass ratios.
 Note that when $\xi$ gets close to 0, it means that the system is in the transitional precession range, with the spin of the primary compensating the orbital angular momentum, and our analysis based on simple precession is not valid anymore.
 
+Note that these figures are to be understood as merely order-of-magnitude estimates, and that they rely on computing derivatives of the time-domain modulation, which can be sensitive to oscillatory features.
+
+\SM{[this is crude anyway]}
 The previous expressions for the error estimate only aim at providing a crude order-of-magnitude picture.
@@
 The results are shown in Fig.~\ref{fig:fomprec}, using the frequency at merger, shown by the vertical lines, to give an idea of the separation between the inspiral and post-merger phases.
-The first two cases, $++$ and $pp$, have \SM{[come back to this after having checked trajectories]} the usual behaviour of precession on a cone, with a larger opening angle for $pp$ due to the larger in-plane spin components.
-After merger, the case $++$ shows a much faster frame rotation than the $pp$ case, as shown by the values of $\Omega_{\rm frame}$ in Table~\ref{table:precexamples}, due to the larger remnant spin that incorporates the aligned spin components.
+The first two cases, $++$ and $\perp\perp$, have \SM{[come back to this after having checked trajectories]} the usual behaviour of precession on a cone, with a larger opening angle for $\perp\perp$ due to the larger in-plane spin components.
+After merger, the case $++$ shows a much faster frame rotation than the $\perp\perp$ case, as shown by the values of $\Omega_{\rm frame}$ in Table~\ref{table:precexamples}, due to the larger remnant spin that incorporates the aligned spin components.
 The case $--$ is qualitatively different, with a larger variation of $\bm{J}$ in direction, and a large and varying $\beta$.
@@
 The mode $m=2$, the most important to model once the amplitudes are refactored in, is mainly in the perturbative regime, but we see that the perturbative treatment can be challenged for modes $m\leq 0$ and also for $m=1$ in the case $++$, especially for the post-merger.
-The amplitude corrections are found to be in the perturbative regime for all cases during the inspiral, however both the $++$ and $pp$ cases show a sharp increase of $\epsilon_{A1}$, $\epsilon_{A2}$ post-merger.
+The amplitude corrections are found to be in the perturbative regime for all cases during the inspiral, however both the $++$ and $\perp\perp$ cases show a sharp increase of $\epsilon_{A1}$, $\epsilon_{A2}$ post-merger.
 The case $--$, with its small remnant spin and mild post-merger frame rotation, shows no such increase.
 
-Overall, the conclusion to be drawn of Fig.~\ref{} is that for the $++$ and $pp$ cases we expect the perturbative approach to be applicable only during the inspiral, with a possible breakdown for the post-merger phase, especially for the case $++$ with its fast post-merger frame rotation.
+Overall, the conclusion to be drawn of Fig.~\ref{} is that for the $++$ and $\perp\perp$ cases we expect the perturbative approach to be applicable only during the inspiral, with a possible breakdown for the post-merger phase, especially for the case $++$ with its fast post-merger frame rotation.
 The higher $|m-m'|$ modes are expected to be more challenging to compute accurately, including during the inspiral.
@@
 The vertical lines indicate the merger frequencies.
-\SM{[color for vertical lines at fpeak]} \SM{[add bottom panel with residual ?]} \SM{[check definition of A]}}
+\SM{[check definition of A]}}
   \label{fig:trigopoly}
@@
 \SM{[possibly move this to the formalism section ? but plots for our three examples]}
+
+\SM{[say that another possibility would be to go back to time-domain and FFT for the limited-length near-merger part of the signal]}
+
+\SM{[expand motivation: sharp feature TD $\rightarrow$ smooth in FD $\rightarrow$ hope to have a cheap trigo representation of FD merger range]}
 
@@
 
-Additionaly, we taper the amplitude to artificially flatten the amplitude at $f_{0}$, at the price of introducing a deviation from the amplitude of the original signal.
+Additionally, we taper the amplitude to artificially flatten the amplitude at $f_{0}$, at the price of introducing a deviation from the amplitude of the original signal.
 This tapering is built in practice as the discrete integral of a cosine window function on just the first two samples.
@@
 
-For the $pp$ case, we find a clearer picture.
+For the $\perp\perp$ case, we find a clearer picture.
 Errors are increasing to larger $|m-m'|$, higher-order corrections improve the reconstruction but are unsufficient for the merger-ringdown, and the convolution works for all modes on the frequency range where it is applied.
@@
 
-For the $pp$ case, both the perturbative corrections and the convolution make a clear improvement.
+For the $\perp\perp$ case, both the perturbative corrections and the convolution make a clear improvement.
 One can see clearly that the convolution becomes unimportant at low masses, as the inspiral dominates here.
@@
   \includegraphics[width=.98\linewidth]{plots/precerror_pp_py.pdf}
-  \caption{Amplitude of the inertial-frame modes $h^{I}_{2m}$ and errors in the Fourier-domain transfer fonctions $\calT^{2}_{m2}$ for the configuration $pp$.
+  \caption{Amplitude of the inertial-frame modes $h^{I}_{2m}$ and errors in the Fourier-domain transfer fonctions $\calT^{2}_{m2}$ for the configuration $\perp\perp$.
 The columns show the orders of approximation $\{\Psi:0,A:0\}$, $\{\Psi:2,A:2\}$, $\{\Psi:2,A:2,\text{conv}\}$.
@@
 \SM{[say how the problems with reconstruction notably for ++ are probably caused by the convolution extending to the right and not to the left]}
+
+\SM{[be honest: to get 1\% mismatch we need to combine high spin, high mass and high inclination]}
 
@@
 Computed here with $f_{\rm min} = 20\mathrm{Hz}$, for masses between 20 and 400 $\Msol$, using the standard aLIGO noise curve \SM{[reference for that curve]}.
-\SM{[Extend computation to fmin=10 and Mmin=10]} \SM{[relate unfaithfulness levels to future detectors in the text]} }
+\SM{[Extend computation to fmin=10 and Mmin=10]} \SM{[relate unfaithfulness levels to future detectors in the text]} \SM{[check definition of inclination: from $J$ or initial $L$ ?]} \SM{[legend is missing A:2]} \SM{[do we want to show A:2 at all ?]}}
   \label{fig:precunfaithfulness}
@@
 In this Appendix, we summarize our conventions for the Wigner matrices.
+
+\SM{[add O'Shaughnessy frame extraction]}
+\SM{[cite other prescriptions]} \SM{[C]}
+
+Several constructions for the radiation axis of the wa precessing frame have been proposed~\cite{} \SM{[C]}.
 
@@
 In this Appendix, we give explicit expressions for the stencil coefficients entering~\eqref{eq:stencilfresnel}.
-In this work we used only low order stencils with usually $N\leq 5$, and one can simply numerically invert of the system~\eqref{eq:stencilsystem}, as this operation has to be done only once.
-For future reference, we give here exact expressions, as the specific choice of samples at $\pm kT$ gives to the system~\eqref{eq:stencilsystem} the form of a Vandermonde system.
+In this work we used only low order stencils with $N\leq 20$, and one can trivially invert of the system~\eqref{eq:stencilsystem}, as this operation has to be done only once.
+One can also obtain closed-form expressions for these coefficients, thanks to the particular choice of samples at $\pm kT$ that gives to the system~\eqref{eq:stencilsystem} the form of a Vandermonde system.
 Defining the Vandermonde matrix $V(x_{0},\dots,x_{N})$ as $V_{ij} = (x_{j})^{i}$, setting $x_{j} = j^{2}$, and $b_{p} \equiv (-i\epsilon)^{p}(2p-1)!!$, the linear system~\eqref{eq:stencilsystem} becomes
 \be
-	b_{p} = \sum\limits_{k=0}^{N} V_{p,k} a_{N,k}^{\epsilon} \,.
+	b_{p} = \sum\limits_{k=0}^{N} V_{pk} a_{N,k}^{\epsilon} \,.
 \ee
@@
 %\ee
-The expression of the inverse of a Vandermonde matrix in terms of symmetric polynomials then allows us to write a closed-form (but rather impractical) expression for the $a_{N,k}^{\epsilon}$, for every finite order $N$:
+The expression of the inverse of a Vandermonde matrix in terms of symmetric polynomials then allows us to write the $a_{N,k}^{\epsilon}$, for every finite order $N$, as:
 \begin{align}
@@
 
+\section{Precession in current Fourier-domain waveform models}
+\label{subsec:precpreviousapproaches}
+
+\SM{[citation O'Shaugnessy]}
+
+In this Appendix, we give a more detailed overview of the treatment of precession in existing waveform models that generate signals directly in the Fourier domain.
+Our objective is not to give an exhaustive account of the various existing models, but rather to highlight how their treatment relates to ours.
+
+To describe previous approaches to the problem of producing precessing waveforms directly in the Fourier domain, avoiding the use of a time-domain generation followed by a FFT \SM{[check acronym]}, it is useful to introduce schematically four different approximations (dropping the mode indices):
+\begin{itemize}
+	\item unstable SPA: $\calT(f) \tilde{h}^{\rm P}(f) = \mathrm{SPA}\left[ \calD^{*} h^{\rm P} \right](f)$,
+	\item $0^{\text{th}}$-order SUA: $\calT(f) = \calD^{*}(\tfSPA) $,
+	\item Fourier-domain $0^{\text{th}}$-order SUA: $\calT(f) = \calD^{*}(\omega = \pi f ) $,
+	\item SUA: $\calT(f) = \frac{1}{2} \sum\limits_{k} a_{k} \calD^{*}(\tfSPA \pm k \Tf^{\rm SPA})$.
+\end{itemize}
+
+In the first option, one applies directly the SPA, as described in Sec.~\ref{subsec:SPA}, to the product of the modulation and the signal.
+This is known (see e.g.~\cite{KCY13}) to lead to possible pathologies, since the prefactor $1/\sqrt{\ddot{\phi}_{\rm prec}}$ can blow up due the precessing contributions to the phase of the inertial-frame waveform.
+
+The second option corresponds to simply multiplying the Fourier-domain signal by the modulation function evaluated at the time $\tfSPA$ as defined in~\eqref{eq:deftfSPA}.
+In the SUA formalism of Ref.~\cite{KCY14}, this treatment could be called the 0th order as it amounts to using a stencil reduced to a single point, i.e.
+using only one term with $a_{0,0} = 1$ in~\eqref{eq:stencilfresnel}.
+It is equivalent to the leading order of our formalism ($\{0, A:0, d:0 \}$ in the terminology of Sec.~\ref{subsec:executivesummary}), with the difference that we use a more general definition of $t_{f}$ (see~\eqref{eq:deftf}) that extends through the merger and ringdown.
+
+While keeping implictly the same level of approximation, one can also use frequency-based expressions for the modulation, as in Refs.~\cite{LOS13, Hannam+13}.
+The precessing-frame evolution is then modelled using post-Newtonian expressions for the Euler angles $(\alpha^{\rm PN}, \beta^{\rm PN}, \gamma^{\rm PN})$ as functions of the orbital frequency $\omega$.
+Using the SPA-inspired correspondence~\eqref{eq:deftfSPA} between the orbital and Fourier-domain frequency, one then writes
+\be\label{eq:precPhenomP}
+	\calT^{\ell}_{mm'}(f) = \calD^{\ell *}_{mm'}(\alpha^{\rm PN}, \beta^{\rm PN}, \gamma^{\rm PN}) \left( \omega=\pi f\right) \,.
+\ee
+When the SPA is valid for the underlying signal $h^{\rm P}_{\ell m}$, the condition $\omega(t) = \pi f$ is equivalent by definition to $t=t^{\rm SPA}_{f}$.
+In~\cite{LOS13}, only the inspiral phase is modelled, and only single-spin configurations, for which the system undergoes simple precession~\cite{Apostolatos+94, Kidder95}.
+The PhenomP model~\cite{Hannam+13} uses Euler angles computed in the limit of a small opening angle of the precession cone and at the spin-orbit level (see also~\cite{BBF11, MBBB13}).
+
+In the PhenomP model~\cite{Hannam+13}, the frequency-based transfer function~\eqref{eq:precPhenomP} is used as an effective prescription covering the whole Fourier-domain frequency band, including the merger and ringdown.
+Note that this procedure amounts to using PN \SM{[check acronym]} results outside of their range of validity, since in the merger-ringdown phase both the PN perturbative treatment and the SPA approximation break down.
+In particular, when computed as a function of time from the balance equation between emitted flux and orbital energy $\calF = -dE/dt$, the orbital $\omega^{\rm PN}(t)$ blows up and cannot be extended through merger.
+In practice, one finds however that the extended frequency-based expressions~\eqref{eq:precPhenomP} are mildly varying when extrapolating to higher frequencies, and this approach has been validated by comparisons to numerical relativity waveforms~\cite{Hannam+13}.
+\SM{[refer to PhenomPv3 in prep.]}
+
+In the SUA formalism proposed in~\cite{KCY13, KCY14}, one extends the SPA to incorporate the leading-order correction during the inspiral phase.
+The transfer function is then computed by evaluating the modulation on a stencil of times centered around the time-of-frequency given by the SPA, arriving at the formula~\eqref{eq:stencilfresnel} with the times $t = \tfSPA$ and $T = \TfSPA$ as defined in~\eqref{eq:deftfSPA} and~\eqref{eq:TfSPA}.
+As described in Sec.~\ref{sec:formalism}, our formalism reduces to the SUA when keeping only the quadratic phase correction, but the SUA is a priori limited to the inspiral phase of the signal through the definitions of the times $t^{\rm SPA}_{f}$ and $T^{\rm SPA}_{f}$.
+For inspiral waveforms, Ref.~\cite{KCY14} showed that this treatment improves the accuracy with respect to the 0th-order approximation.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-05-31 docs/paper_PE.tex
Start for Parameter estimation paper

\documentclass[aps,showpacs,twocolumn,prd,superscriptaddress,nofootinbib]{revtex4}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{color}
\usepackage{enumerate}
\usepackage{ulem}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand\ud{{\mathrm{d}}}
\newcommand\uD{{\mathrm{D}}}
\newcommand\calO{{\mathcal{O}}}
\newcommand\bfx{\mathbf{x}}
\newcommand{\ov}[1]{\overline{#1}}
\newcommand{\ph}[1]{\phantom{#1}}
\newcommand{\cte}{\mathrm{cte}}
\newcommand{\nn}{\nonumber}
\newcommand{\hatk}{\hat{k}}
\newcommand{\Hz}{\,\mathrm{Hz}}
\newcommand{\sinc}{\,\mathrm{sinc}}
\newcommand{\Msol}{M_{\odot}}

\begin{document}

\title{Bayesian methods for black hole merger parameter estimation with LISA.}

\author{John G. Baker}
\affiliation{Gravitational Astrophysics Laboratory, NASA Goddard Space Flight Center, 8800 Greenbelt Rd., Greenbelt, MD 20771, USA}
\author{Sylvain Marsat}
\affiliation{Department of Physics, University of Maryland, College Park, MD 20742, USA}
\affiliation{Gravitational Astrophysics Laboratory, NASA Goddard Space Flight Center, 8800 Greenbelt Rd., Greenbelt, MD 20771, USA}
\affiliation{Max Planck Institute for Gravitational Physics (Albert Einstein Institute), Am M\"uhlenberg 1, Potsdam-Golm, 14476, Germany}

\date{\today}

\begin{abstract}

[Abstract]

\end{abstract}

\pacs{
04.70.Bw, % classical black holes
04.80.Nn, % Gravitational wave detectors and experiments
95.30.Sf, % relativity and gravitation
95.55.Ym, % Gravitational radiation detectors
97.60.Lf  % black holes (astrophysics)
}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}

[Introduction]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Methodology}
\label{sec:intro}

\subsection{Fast frequency domain LISA response}
\label{sec:response}

-- Geometric setting, definitions, summarize the FD response

\subsection{Reduced order model for EOBNRv2 waveforms.}
\label{sec:waveforms}

-- Waveforms: EOBNRv2HMROM + PNext

\subsection{Fast likelihood computation}
\label{sec:likelihood}

-- Likelihood computation: fast overlaps

\subsection{Fisher matrix parameter estimation}
\label{sec:Fisher}

-- Fisher matrices computation: step self-tuning,...

\subsection{Bayesian sampling}
\label{sec:samplers}

-- Bayesian samplers: Multinest, PTMCMC

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Morphology of the signals}
\label{sec:morph}

- Morphology of the response

\subsection{Temporal development of SNR}
\label{sec:timeSNR}

-- SNR accumulation with time or frequency: inspiral/MRD balance across mass range

\subsection{LISAframe parameters}
\label{sec:LISAframe}

-- LISA frame angles

\subsection{The low-frequency limit}
\label{sec:low-freq}

-- make the connection to low-f response (2 interferometers)

\subsection{A simplified likelihood model}
\label{sec:simple-like}

-- simplified response and closed-form likelihood

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Illustrating examples}
\label{sec:examples}

- Demonstration/examples

\subsection{A non-degenerate case}
\label{sec:high-M-non-deg}

-- High-mass system, non-degenerate, 22 and HM + Fisher - stacked SNR contours
eg:
m1,    m2,    t0,D,     phi0,i,   lam,  bet, psi
8.89e5,1.11e5,0, 3.67e4,pi/3,pi/2,3pi/4,pi/3,pi/3

-- Performance: waveform and likelihood costs, number of sampler evaluations

\subsection{Stellar origin black holes}
\label{sec:SOBH}

-- SOBH system, SNR sufficient for clean analysis + Fisher - stacked SNR contours
-- Performance: waveform and likelihood costs, number of sampler evaluations

\subsection{Highlighting degeneracies}
\label{sec:degen}

- Highlighting degeneracies

-- High-mass degenerate example, 22 vs HM - stacked SNR contours
-- show degenerate waveforms
-- simplified response and closed-form likelihood, full degeneracies
-- (explore simplified response further: parameter maps, other off-the-shelf samplers ?)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}
\label{sec:discussion}

- Summary/conclusions

----------------
Other questions:
- special orientations/sky locations ?
- Fisher matrix qualification - explore parameter choices
- continuously increase SNR - include on these examples
- continuously accumulate information over time
- full exploration of parameter space for all of the above...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2018-06-02 docs/paper_method.tex
Edits to the paper.

@@
 
+\SM{[say that we reverse point of view: waveform is given FD, modulation TD, instead of waveform+modulation TD]}
+
+\SM{[say a word of how to gauge relevance of a given error: 1/SNR in direct error ? We won't give anything but waveform reconstruction errors...]}
+
 \SM{[style: below use texttt for models and codes (like \texttt{SpEC})]}
@@
 As a preliminary step, we recall here an approximation widely used for similar purposes in treating gravitational waves signals emitted by inspiraling binaries, the Stationary Phase Approximation (thereafter SPA, and also sometimes called the steepest-descent method).
-It applies in general to chirping signals, and we refer the reader to~\ref{} for details.
+It applies in general to chirping signals, and we refer the reader to~\cite{FC93, CF94} for details.
 One first writes the time-domain signal in amplitude and phase form as $h(t) = a(t) e^{-2i\varphi(t)}$, where for gravitational wave signals $\varphi$ will correspond to the orbital phase, with the orbital frequency being $\omega = \dot{\varphi}$.
 In order to keep close to the notation used in the gravitational wave literature, we introduced a factor of 2 in the phase of the wave, which is appropriate for the dominant 22 harmonic of the signal.
-The approximation then applies to signals verifying the conditions \SM{[say that the third is sometimes missing ?]} \SM{[add reference to Finn-Chernoff, Cutler-Flanagan where the third one is missing]}
+The approximation then applies to signals verifying the conditions\footnote{Note that the third condition is not always written explicitly, in particular not in Refs.~\cite{FC93, CF94}.
+It becomes important if we generalize $a$ to be an envelope function incorporating a modulation.} \SM{[say that the third is sometimes missing ?]} \SM{[add reference to Finn-Chernoff, Cutler-Flanagan where the third one is missing]}
 \be\label{eq:conditionsSPA}
@@
 Note that to be able to treat the amplitude as a constant in the integral above, we used the third condition in~\eqref{eq:conditionSPA}.
-The resulting complex Gaussian integral yields\footnote{Note that the expressions given here are valid for the dominant mode $h_{22}$ of the waveform.
-They can be readily generalized to other modes $h_{\ell m}$ with phase $e^{-im\varphi}$.}
+The resulting complex Gaussian integral yields\footnote{The expressions given here are valid for the dominant mode $h_{22}$ of the waveform.
+They can be generalized to other modes $h_{\ell m}$, $m\neq 0$ with phase $e^{-im\varphi}$ as follows: $\Psi$ acquires a factor $m/2$, $A$ a factor $\sqrt{2/m}$, and $v$ (see~\eqref{eq:SPAN}) has to be replaced by $(2/m)^{1/3} v$.}
 \begin{subequations}
@@
 \be\label{eq:a22N}
-	a_{22}^{\rm N} (t) = \frac{2 G M \nu x}{D c^{2}} \sqrt{\frac{16 \pi}{5}} \,,
-\ee
-where we set $x = (G M\omega/c^{3})^{2/3}$ and where $D$ is the luminosity distance to the observer.
+	a_{22}^{\rm N} (t) = \frac{2 G M \nu v^{2}}{D c^{2}} \sqrt{\frac{16 \pi}{5}} \,,
+\ee
+where we set $v = (G M\omega/c^{3})^{1/3}$ and where $D$ is the luminosity distance to the observer.
 For this Newtonian inspiral, applying the SPA gives for the $22$ mode $\tilde{h}(f) = A_{\rm N}(f)e^{-i\Psi_{\rm N}(f)}$ with \SM{[check that total mass convention is $M$ everywhere]}
@@
 \end{subequations}
-where we used the notation $v=(G M \pi f/c^{3})^{1/3}$ ($x=v^{2}$ with the SPA correspondence~\eqref{eq:deftfSPA}), and where $t_{0}, \phi_{0}$ are constants.
+where $v=(G M \pi f/c^{3})^{1/3}$ according to the SPA correspondence~\eqref{eq:deftfSPA}, and where $t_{0}, \phi_{0}$ are constants.
 It is also customary to rewrite the above relations in terms of the chirp mass $\Mchirp \equiv M\nu^{3/5}$, which is the only mass combination characterizing the signal at the leading PN order.
@@
 \ee
-Thus, when the SPA applies, $\Tf$ corresponds to} the radiation-reaction timescale: the shorter this timescale, the faster the binary chirps to higher frequencies on its quasi-circular inspiral.
+Thus, when the SPA applies, $\Tf$ corresponds to} the radiation-reaction timescale \SM{[make the connection to $Q$ used for SEOB comparisons in the Nagar group]}: the shorter this timescale, the faster the binary chirps to higher frequencies on its quasi-circular inspiral.
 
@@
 
-We can obtain useful estimates for these timescales from the leading order post-Newtonian expressions~\eqref{eq:SPAN}. The leading-order radiation-reaction and amplitude timescales are:
+We can obtain useful estimates for these timescales from the leading order post-Newtonian expressions~\eqref{eq:SPAN}, valid for the dominant harmonic $h_{22}$.
+The leading-order radiation-reaction and amplitude timescales are:
 \begin{subequations}\label{eq:timescalesN}
 \begin{align}
-	\Tf^{\rm N} &= \frac{1}{8} \sqrt{\frac{5}{3\nu}} \frac{G m}{c^{3}} v^{-11/2} \,, \label{eq:TfN}\\
+	\Tf^{\rm N} &= \frac{1}{8} \sqrt{\frac{5}{3\nu}} \frac{G M}{c^{3}} v^{-11/2} \,, \label{eq:TfN}\\
 	T_{A1}^{\rm N} &= \frac{7}{6} \frac{1}{2\pi f}\,, \quad T_{A2}^{\rm N} = \frac{\sqrt{91}}{6} \frac{1}{2\pi f} \,. \label{eq:TA1N-TA2N}
@@
 With a simple power-law amplitude as in~\eqref{eq:ASPAN}, all higher-order amplitude timescales are also simply proportional to $1/f$ and differ only by their numerical factor.
+For higher harmonics $h_{\ell m}$ with $m\neq 0$, as discussed in Sec.~\ref{subsec:SPA}, $T_{f}$ acquires a factor $m/2$, and $v$ has to be replaced by $(2/m)^{1/3} v$.
+The amplitude of higher harmonics starts at a higher PN order~\cite{BlanchetLiving}, differing from~\eqref{eq:a22N} by a different constant and an additional scaling $v^{\kappa_{\ell m}}$ with $\kappa_{\ell m} = \ell - 2 + (\ell + m \; \mathrm{mod} \; 2)$, which changes the constant in the amplitude timescale as $7/6 \rightarrow (7-2\kappa_{\ell m})/6$ in $T_{A1}^{\rm N}$ and $\sqrt{91}/6 \rightarrow \sqrt{(7-2\kappa_{\ell m})(13-2\kappa_{\ell m})}/6$ in $T_{A2}^{\rm N}$.
 
@@
 In the following, it will be convenient to introduce error estimates built from the Taylor-like series~\eqref{eq:resulttaylorPsi},~\eqref{eq:resulttaylorA} and~\eqref{eq:resulttaylordelay}.
-We simply define  this error estimates at a certain level of approximation as the magnitude of the first term ignored in the original Taylor series.
+We simply define these error estimates at a certain level of approximation as the magnitude of the first term ignored in the original Taylor series.
+To give these quantities a relative meaning, we divide by the leading term.
 Thus we define, with $G(f,t) = F(t) e^{-2i\pi f d(t)}$,
@@
 We investigate how the separation of the relevant timescales evolves in the late inspiral and after the merger occurs.
+We will also give in App.~\ref{app:precpreviousapproaches} a short overview of previous approaches to this problem of the Fourier-domain precession, and how they relate to our formalism.
 
@@
 \be
-	\mathrm{FT}[g^{*}](f) = \tilde{g}(-f)^{*} \,,
+	\mathrm{FT}[g^{*}](-f) = \tilde{g}(f)^{*} \,,
 \ee
@@
 \ee
-This means that such a model is required to cover only the positive frequency band and the values $m'>0$, the negative frequencies being deduced from
-\begin{align}
+This means that such a model is required to cover only the positive frequency band and the values $m'>0$, since
+\begin{align}\label{eq:hIlmposnegfreq}
 	\tilde{h}^{\rm I}_{\ell m}(f) &= \sum_{m'>0} \calT^{\ell}_{mm'}(f) \tilde{h}^{\rm P}_{\ell, m'}(f) \text{ for } f>0 \,, \nn\\
@@
 
-Further simplifications occur when including only the dominant harmonics $h^{\rm P}_{22}$, $h^{\rm P}_{2,-2}$ in the precessing-frame waveform (as is done for instance in PhenomP~\cite{Hannam+13}, as well as in our toy model~\ref{subsec:precmodel}), we have in this case for $f<0$
+Further simplifications occur when including only the dominant harmonics $h^{\rm P}_{22}$, $h^{\rm P}_{2,-2}$ in the precessing-frame waveform (as is done for instance in PhenomP~\cite{Hannam+13}, as well as in our toy model~\ref{subsec:precmodel}), we have in this case for $f<0$ \SM{[all $f$ ?]}
 \begin{subequations}
 \begin{align}
-	\calT^{2}_{m,-2}(f) &= (-1)^{m} \calT^{\ell *}_{-m,2}(-f) \,, \\
+	\calT^{2}_{m,-2}(f) &= (-1)^{m} \calT^{2 *}_{-m,2}(-f) \,, \\
 	\tilde{h}^{\rm I}_{2 m}(f) &= (-1)^{m} \tilde{h}^{\rm I *}_{2,-m}(-f) \,. \label{eq:symmetryhIfor22only}
@@
 where we use the minimal-rotation condition~\eqref{eq:gammadot} for a constant $\beta$.
-For the QNM frequencies as a function of the final spin, we use values computed in Ref.~\cite{Berti+05}.
-The final spin is taken to be the same as the one computed internally to the SEOBNRv3 code~\cite{Pan+13, BTB16}, where the final-spin fit formula of Ref.~\cite{} \SM{[C]}, built for spin-aligned systems, is applied to the spin components projected on the orbital angular momentum $\bm{L}$ at merger.
+To compute the QNM frequencies as a function of the final spin $\chi_{f}$, we use fits constructed in Ref.~\cite{Berti+05}, illustrated in Fig.~\ref{fig:QNM}.
+The final spin is taken to be the same as the one computed internally to the SEOBNRv3 code~\cite{Pan+13, BTB16}, where the final-spin fit formula of Ref.~\cite{BR09} \SM{[see if we can update the final-spin formula]}, built for spin-aligned systems, is applied to the spin components projected on the orbital angular momentum $\bm{L}$ at merger.
 
@@
 The example catalog waveform is SXS:BBH:0058, generated with SpEC~\cite{SXScatalog, SpEC, Mroue+12, Mroue+13}, with a mass ratio of $q=5$, and a single in-plane spin of $\chi_{1} = 0.5$.
-The NR and SEOB post-merger frames disagree mainly due to a different value of the final spin $\chi_{f}$\footnote{The fitting formulas used to compute $\chi_{f}$ in the next version of the SEOB code, which is expected to reduce or remove this disagreement~\cite{Ossokine+18}.
+The NR and SEOB post-merger frames disagree mainly due to a different value of the final spin $\chi_{f}$\footnote{To compute $\chi_{f}$, SEOBNRv3 uses fitting formulas from~\cite{BR09}.
+These formulas will be updated to more recent fits incorporating additional NR data (see e.g.~\cite{HBR16}) in the next version of the SEOB code~\cite{Ossokine+18}, which is expected to reduce or remove this disagreement.
 \SM{[update this reference in prep.]}}, which leads to a different $\Omega_{\rm frame}$.
@@
 Leaving aside the question of a more in-depth investigation of the best representation of precessing waveforms in their post-merger phase, we will proceed to investigating the separation of timescales in three chosen examples.
-The parameters of these example cases are summarized in Table~\ref{table:precparams}.
+The parameters of these example cases are summarized in Table~\ref{tab:precparams}.
 We consider a mass ratio of $q=4$, close-to-maximal spins of $\chi_{1} = \chi_{2} = 0.95$, and vary both spins misalignmnent angles to be $\pi/6$ (case labeled $++$, close to aligned spins), $\pi/2$ ($\perp\perp$, spins in the plane) and $5\pi/6$ ($--$, close to anti-aligned spins).
-Table~\ref{table:precparams} also shows the remnant spins that are internally computed in both the SEOBNRv3 and the PhenomD/PhenomP codes, as well as the QNM frequencies and the resulting frame precession frequency $\Omega_{\rm frame}$.
+Table~\ref{tab:precparams} also shows the remnant spins that are internally computed in both the SEOBNRv3 and the PhenomD/PhenomP codes, as well as the QNM frequencies and the resulting frame precession frequency $\Omega_{\rm frame}$.
 Note that the case $--$ shows a larger variation in the direction of $\bm{J}$, and the spin of the remnant is almost 0, due to the fact that the spin is large and antialigned with a large mass ratio.
@@
 \SM{[To be updated with longer waveforms, 10Hz 10Msol.
-?]} \SM{[add misalignment angles $(\ell, \bm{J}_{i})$?]}}\label{table:precparams}
+?]} \SM{[add misalignment angles $(\ell, \bm{J}_{i})$?]}}\label{tab:precparams}
 %\begin{tabular}{C{.2\columnwidth}C{.24\columnwidth}C{.15\columnwidth}}
-\begin{tabular}{ccccccc}\label{table:precexamples}
+\begin{tabular}{ccccccc}\label{tab:precexamples}
 	$f_{\rm min}$ & $ M_{\rm min} $ & $q$ & $\chi_{1}$ & $\chi_{2}$ & $ \phi_{1} $ & $ \phi_{2} $ \\
@@
 %\begin{table}[t]
-%\begin{ruledtabular}\caption{Parameters of the three example cases that we use to illustrate our formalism.}\label{table:precparams}
+%\begin{ruledtabular}\caption{Parameters of the three example cases that we use to illustrate our formalism.}\label{tab:precparams}
 %%\begin{tabular}{C{.2\columnwidth}C{.24\columnwidth}C{.15\columnwidth}}
@@
   \includegraphics[width=.98\linewidth]{plots/postmergeromega_py.pdf}
-  \caption{Quasi-normal mode frequencies for the 0th overtone of the modes $22$ and $21$, as a function of the dimensionless spin of the final black hole $\chi_{f}$.
-We also show their difference $\Omega_{\rm frame}$ (see~\eqref{eq:OmegaframeQNM}), and the ratio $\Omega_{\rm frame} / \omega^{\rm QNM}_{220}$ characteristic of the separation of timescales between the the phase of the modulation~\eqref{eq:wignerphasesimpleprec} and the P-frame waveform phase.}
-  \label{fig:precmodel}
+  \caption{\SM{[label y axis]} Quasi-normal mode frequencies for the 0th overtone of the modes $22$ and $21$, as a function of the dimensionless spin of the final black hole $\chi_{f}$, as well as their difference $\Omega_{\rm frame}$ (see~\eqref{eq:OmegaframeQNM}).
+The lower panel shows the  ratio $\Omega_{\rm frame} / \omega^{\rm QNM}_{220}$ characteristic of the separation of timescales between the phase of the modulation~\eqref{eq:wignerphasesimpleprec} and the P-frame waveform phase.}
+  \label{fig:QNM}
 \end{figure}
@@
 
-We now turn to the magnitude of higher-order corrections in the case of precessing binaries, as can be estimated by the quantities $\epsilon_{\Psi 2}$, $\epsilon_{A1}$ and $\epsilon_{A 2}$ introduced in~\eqref{eq:deffom}.
-
-For the inspiral phase of the dynamics, one can read the quantities $\epsilon_{\Psi 2}$, $\epsilon_{A1}$ and $\epsilon_{A 2}$ from the outcome of post-Newtonian computations.
-To obtain a qualitative picture of the timescales involved, here we will consider the most simple setting for precession in the post-Newtonian framework: at leading PN order and for a single-spin system, both the orbital angular momentum $\bm{L}$ and the spin vector $\bm{S}$ undergo what is called simple precession around the total angular momentum, $\bm{J}$ (see e.g.~\cite{Apostolatos+94, Kidder95}).
-By neglecting the slow evolution of the orbit under radiation reaction, ignoring higher-order corrections that are quadratic and higher in the spins, and orbit-averaging the precession equations (thus eliminating nutations in the precession of the orbital plane, which occur at the orbital timescale) \SM{[no nutation for single spin?]}, the trajectories of $\bm{L}$ and $\bm{S}$ become cones around the direction of $\bm{J}$, with a constant precession velocity $\Omega_{\rm prec}$ that we will use to define the timescale of variations of the modulation functions $F$.
-Note that the presence of two spins complicates the evolution of the system, but the picture of a precession cone for $\bm{\ell}$ remains approximately valid (see~\cite{} \SM{[C]} for a classification of the morphologies of precessing trajectories).
-
-In the following, we will use the notation \SM{[use M instead of m]} $m=m_{1}+m_{2}$, $\nu=m_{1}m_{2}/m^{2}$, $\delta = (m_{1}-m_{2})/m$, $|\bm{S}_{A}|=Gm_{A}^{2} \chi_{A}$ for $A=1,2$.
-We take the convention $m_{1} \geq m_{2}$.
-We define $\bm{\ell}$ as the unit vector normal to the orbital plane, and assume that only the more massive object has a spin, so that $\bm{S}_{2} = 0$.
-A factor $c$ is scaled out of the spins so that $S_{A} = G m_{A}^{2} \chi$, with $\chi$ the dimensionless spin between $0$ and $1$ for Kerr black holes.
-We will use the notation $v = (G m \omega/c^{3})^{1/3}$ with $\omega$ the orbital frequency, which is equivalent to the frequency-domain definition $v=(G \pi m f/c^{3})^{1/3}$ when the SPA applies.
-\SM{[Check redundancy with notations of previous sections]} The Newtonian order orbital angular momentum is $\bm{L} = L_{N} \bm{\ell}$, with
+We now turn to the separation of timescales and to the magnitude of the higher-order corrections derived in Sec.~\ref{sec:formalism} in the case of precessing binaries, as can be estimated by the quantities $\epsilon_{\Psi 2}$, $\epsilon_{A1}$ and $\epsilon_{A 2}$ introduced in~\eqref{eq:deffom}.
+
+In the inspiral phase, we can obtain a qualitative picture of the timescales involved by using well-known leading-order post-Newtonian results.
+To simplify things further, we will consider a single-spin system, and use orbit-averaged precession equations.
+In this configuration, both the orbital angular momentum $\bm{L}$ and the spin vector $\bm{S}$ undergo simple precession on a cone around the total angular momentum $\bm{J}$ (see e.g.~\cite{Apostolatos+94, Kidder95}), with an opening angle of the cone and a precession velocity $\Omega_{\rm prec}$ that vary only on the radiation reaction timescale.
+In the general case, the presence of two spins complicates the evolution of the system, but the picture of a precession cone for $\bm{L}$ remains approximately valid (see~\cite{Kesden+14} for a classification of generic precession trajectories).
+
+We will use the following notations: for $m_{1}$, $m_{2}$ the masses of the two bodies, we set $M=m_{1}+m_{2}$, $\nu=m_{1}m_{2}/M^{2}$, $\delta = (m_{1}-m_{2})/M$, and $|\bm{S}_{A}|=Gm_{A}^{2} \chi_{A}$ for $A=1,2$, with $\chi$ the dimensionless spin between $0$ and $1$ for Kerr black holes.
+We define $\bm{\ell}$ as the unit vector normal to the orbital plane.
+We take the convention $m_{1} \geq m_{2}$ and assume that only the more massive object has a spin, $\bm{S}_{2} = 0$.
+As in Sec.~\ref{subsec:SPA}, we will use the notation $v = (G M \omega/c^{3})^{1/3}$ with $\omega = \dot{\varphi}$ the orbital frequency, which translates to $v=(G \pi M f/c^{3})^{1/3}$ for the 22 mode when the SPA applies.
+At the Newtonian order, the orbital angular momentum is $\bm{L} = L_{N} \bm{\ell}$, with
 \be\label{eq:defLN}
-	L_{N} = \frac{G m^{2} \nu}{c v} \,.
-\ee
-Since we neglect radiation reaction, $\bm{J}$ is treated a constant that we use to set the z-axis so that $\bm{J} = J \bm{e}_{z}$, and we decompose the spin in its aligned and perpendicular components as $\bm{S}_{1} = S_{1}^{z} \bm{e}_{z} + \bm{S}_{1}^{\perp}$.
+	L_{N} = \frac{G M^{2} \nu}{c v} \,.
+\ee
+Since we neglect radiation reaction, $\bm{J} = \bm{L} + \bm{S}_{1}/c$ is treated a constant that we use to set the z-axis so that $\bm{J} = J \bm{e}_{z}$, and we decompose the spin in its aligned and perpendicular components as $\bm{S}_{1} = S_{1}^{z} \bm{e}_{z} + \bm{S}_{1}^{\perp}$.
 At leading order, the precession equations read
@@
 	\dot{\bm{S}}_{1} &= \bm{\Omega}_{1} \times \bm{S}_{1} \,, \nn\\
-	\dot{\bm{\ell}} &= - \frac{1}{c L_{N}}  \dot{\bm{S}}_{1}\,, \nn\\
-\end{align}
-with the spin precession velocity \cite{} \SM{[C]}
+	\dot{\bm{\ell}} &= - \frac{1}{c L_{N}}  \dot{\bm{S}}_{1}\,,
+\end{align}
+with the spin precession velocity~\cite{Kidder95}
 \be
-	\bm{\Omega}_{1} = \Omega_{1} \bm{\ell} = \frac{c^{3}}{G m} \left( \frac{3}{4} + \frac{\nu}{2} - \frac{3\delta}{4} \right) v^{5} \bm{\ell} \,.
+	\bm{\Omega}_{1} = \Omega_{1} \bm{\ell} = \frac{c^{3}}{G M} \left( \frac{3}{4} + \frac{\nu}{2} - \frac{3\delta}{4} \right) v^{5} \bm{\ell} \,.
 \ee
 which is formally a 1PN quantity.
-Decomposing the vectors in their in-plane component and projection on $z$, and using $\delta^{2} = 1-4\nu$ to make explicit the overall scaling in $\nu$, we obtain for the frame precession velocity (defined such that $\dot{\alpha} = \Omega_{\rm prec}$)
+Considering only the leading PN order, we ignore effects quadratic in the spin that would enter here at 1.5PN.
+Decomposing the vectors in their in-plane component and projection on $z$, and using $\delta^{2} = 1-4\nu$ to make explicit the overall scaling in $\nu$, we obtain for the frame precession velocity~\cite{Kidder95}
 \be
-	\Omega_{\rm prec} = \frac{c^{4}}{G^{2} m^{3}} \frac{7+\delta}{2(1+\delta)} v^{6} J\,.
-\ee
-For convenience, we will use the notation
-\begin{subequations}
-\begin{align}
-	\Lambda &\equiv \frac{7+\delta}{4(1+\delta)} \,, \\
-	\xi &\equiv 1 + \frac{v S_{1}^{z}}{G m^{2} \nu} \,.
-\end{align}
-\end{subequations}
-$\Lambda$ is a mass ratio-dependent factor chosen to be of order 1, varying from $7/4$ for equal masses to $1$ in the test-mass limit.
-The factor $\xi$ shows how the contribution of the spin to the total angular momentum affects the precession rate.
+	\dot{\alpha} \equiv \Omega_{\rm prec} = \frac{c^{4}}{G^{2} M^{3}} \frac{7+\delta}{2(1+\delta)} v^{6} J\,.
+\ee
+Separating the factors as
+\be\label{eq:defLambdaxi}
+	\Lambda \equiv \frac{7+\delta}{4(1+\delta)} \,, \quad	\xi \equiv 1 + \frac{v S_{1}^{z}}{G M^{2} \nu} \,,
+\ee
+we see that $\Lambda$ is a mass ratio-dependent factor chosen to be always of order 1, varying from $7/4$ for equal masses to $1$ in the test-mass limit, while the factor $\xi$ contains the contribution of the aligned component of the spin to the precession rate.
 With this notation,
 \be\label{eq:Omegaprec}
-	\Omega_{\rm prec} = \frac{2c^{3} \nu}{G m} \Lambda \xi v^{5} \,.
-\ee
-If precession effects are in general larger for larger mass ratios, as the precession cones widens, the overall scaling of $\nu$ in~\eqref{eq:Omegaprec} above ($\Lambda$ has a mild variation with masses, which is also decreasing with increasing mass ratio) shows that, as long as the orbital angular momentum still dominates the spin, high mass ratio systems have also a slower precession rate.
+	\Omega_{\rm prec} = \frac{2c^{3} \nu}{G M} \Lambda \xi v^{5} \,.
+\ee
+If precession effects are in general larger for larger mass ratios, as the precession cones widens, the overall scaling of $\nu$ in~\eqref{eq:Omegaprec} above shows that, as long as the orbital angular momentum still dominates the spin in $\xi$, increasing the mass ratio yields a slower precession rate.
+For high mass ratios and spins, the correction to $\xi$ in~\eqref{eq:defLambdaxi} starts to become important.
 
 We now turn to the Euler angles and modulation functions $\calD^{\ell *}_{mm'}(\alpha, \beta, \gamma)$, given explicitly in~\eqref{eq:defWignerD}.
-In the case of simple precession, the opening angle of the precession cone, $\beta$, evolves solely due to radiation reaction.
-If we treat it as a constant, the minimal rotation condition relates $\gamma$ to $\alpha$ by $\gamma  = -\alpha \cos \beta$ (up to a constant), and the only variable part of the modulation functions~\eqref{} \SM{[R]} are the phases.
+As explained in~\ref{subsec:precdef}, in the case of simple precession, the opening angle of the precession cone $\beta$ is essentially constant, so that the minimal rotation condition~\eqref{eq:gammadot} gives $\gamma  = -\alpha \cos \beta$ (up to a constant), and the only variable part of the precession modulation functions in~\eqref{eq:wignerrot} are the phases, according to~\eqref{eq:wignerphasesimpleprec}.
 The constant rate of rotation around $\bm{J}$ translates into $\dot{\alpha} = \Omega_{\rm prec}$.
-From the closure relation $\bm{J} = \bm{L} + \bm{S}_{1}$, we have
+From the closure relation $\bm{J} = \bm{L} + \bm{S}_{1}/c$, we have
 \be\label{eq:betaconst}
-	\cos \beta = \sqrt{1 - \left( \frac{S_{1}^{\perp}}{c L_{N}} \right)^{2}}\,.
-\ee
-For each mode contribution $\calD^{\ell *}_{mm'}$, we have \SM{[said before]}
-\be
-	\calD^{\ell *}_{mm'} (\alpha, \beta, \gamma) \propto e^{i(m' \cos\beta - m) \alpha}
-\ee
-which gives an additional geometric factors of $(m' \cos\beta - m)$.
-Under our assumptions of simple precession, ignoring radiation reaction, and using the leading-order expressions~\eqref{} \SM{[R]} for the Fourier-domain timescales, we finally obtain the following error estimates for individual modes \SM{[recall what $m$, $m'$ mean]}:
-\begin{subequations}
-\begin{align}
-	\epsilon_{\Psi 2} &= \frac{5\nu}{96} \Lambda^{2} \xi^{2} (m' \cos\beta - m)^{2} v^{-1} \,, \\
-	\epsilon_{A 1} &=  \frac{7\nu}{6} \Lambda \xi |m' \cos\beta - m| v^{2} \,, \\
-	\epsilon_{A 2} &= \frac{91 \nu^{2}}{72} \Lambda^{2} \xi^{2} (m' \cos\beta - m)^{2} v^{4} \,.
+	\cos \beta = \sqrt{1 - \left( \frac{vS_{1}^{\perp}}{G M^{2}\nu} \right)^{2}}\,.
+\ee
+For a given mode contribution $\calD^{\ell *}_{mm'}$, \eqref{eq:wignerphasesimpleprec} shows that we will have factors of $(m' \cos\beta - m)$ when taking derivatives.
+
+We have now everything we need to compute the error estimates~\eqref{eq:deffom} for the transfer function $\calT^{\ell}_{mm'}$~\eqref{eq:defprectransfer} with $m' \neq 0$.
+Given our restrictive  assumptions of orbit-averaged leading-order PN and single-spin simple precession, the result will only be a crude order-of-magnitude estimate.
+Using the Fourier-domain leading-order timescales generalized for modes $h^{\rm P}_{\ell m'}$ in the text below~\eqref{eq:timescalesN} yields: \SM{[check]}
+\begin{subequations}\label{eq:precfomPN}
+\begin{align}
+	\epsilon_{\Psi 2} &= \frac{m'}{2} \frac{5\nu}{96} \Lambda^{2} \xi^{2} (m' \cos\beta - m)^{2} v^{-1} \,, \\
+	\epsilon_{A 1} &=  \frac{(7 - 2\kappa_{\ell m'})\nu}{6} \Lambda \xi |m' \cos\beta - m| v^{2} \,, \\
+	\epsilon_{A 2} &= \frac{(7 - 2\kappa_{\ell m'}) (13 - 2\kappa_{\ell m'}) \nu^{2}}{72} \Lambda^{2} \xi^{2} (m' \cos\beta - m)^{2} v^{4} \,,
 \end{align}
 \end{subequations}
-As noted in Section~\ref{} \SM{[R]}, $\epsilon_{\Psi 2}$ has a factor $v^{-1}$, formally at $-0.5$PN order, which means that the relative size of this correction grows towards smaller frequencies, away from merger.
-Remember however that the opening angle of the precession cone, giving the overall normalization for precession effects in the waveform, also goes to 0 as $v$ in that limit.
-The amplitude error estimates $\epsilon_{A1}$, $\epsilon_{A2}$, by contrast, have the more usual behaviour of PN corrections (of order 1PN and 2PN respectively) growing towards merger.
-The geometric factors $(m' \cos\beta - m)$ show that higher $|m-m'|$ mode contributions are harder to model, with larger $\epsilon$, although they also have smaller amplitudes for small $\beta$.
-The overall $\nu$ scalings indicate a better separation of timescales for larger mass ratio systems.
-These expressions also show that, since the total angular momentum appear as a factor in $\Omega_{\rm prec}$, higher-order corrections will be larger for spin-aligned systems than for anti-aligned spins, an effect that becomes significant for large mass ratios.
-Note that when $\xi$ gets close to 0, it means that the system is in the transitional precession range, with the spin of the primary compensating the orbital angular momentum, and our analysis based on simple precession is not valid anymore.
-
-Note that these figures are to be understood as merely order-of-magnitude estimates, and that they rely on computing derivatives of the time-domain modulation, which can be sensitive to oscillatory features.
-
-\SM{[this is crude anyway]}
-The previous expressions for the error estimate only aim at providing a crude order-of-magnitude picture.
-We now present a more complete numerical computation of these quantities for our post-merger extended precession model presented in Sec.~\ref{subsec:precmodel}, for the three examples summarized in Table.~\ref{table:precexamples}.
-Here and in the following we present only the $h^{\rm I}_{2 m}$ mode contributions induced by $h^{\rm P}_{22}$, knowing that the ones induced by $h^{\rm P}_{2,-2}$ can be deduced from~\eqref{} \SM{[R]}.
-We compute the Fourier-domain based timescales $\Tf$, $T_{A1}$ and $T_{A2}$ from the analytic derivatives of the PhenomD phase and amplitude (modified to be of class $C^{2}$).
-The time derivatives of the time-domain modulation are computed from an interpolating cubic spline.
-The results are shown in Fig.~\ref{fig:fomprec}, using the frequency at merger, shown by the vertical lines, to give an idea of the separation between the inspiral and post-merger phases.
-The first two cases, $++$ and $\perp\perp$, have \SM{[come back to this after having checked trajectories]} the usual behaviour of precession on a cone, with a larger opening angle for $\perp\perp$ due to the larger in-plane spin components.
-After merger, the case $++$ shows a much faster frame rotation than the $\perp\perp$ case, as shown by the values of $\Omega_{\rm frame}$ in Table~\ref{table:precexamples}, due to the larger remnant spin that incorporates the aligned spin components.
-The case $--$ is qualitatively different, with a larger variation of $\bm{J}$ in direction, and a large and varying $\beta$.
-Since the $\epsilon$ quantities are the magnitude of the first or second term in a Taylor series, we will take $\epsilon \sim 1$ as an indication of the breakdown of a perturbative treatment.
-We recall an important point about Fig.~\ref{fig:fomprec}: since the error estimates are computed on a mode-by-mode basis and normalized by the transfer function for this mode, the amplitude hierarchy between mode contributions is erased here.
-What is plotted is an error estimate in relative sense, and one should keep in mind that higher $|m-m'|$ modes, which we will find harder to model accurately, can be very subdominant in the final waveform.
-
-The resulting magnitude of the error estimates in Fig.~\ref{fig:fomprec} is roughly in agreement with the PN-inspired computation~\ref{eq:} \SM{[R]} above for the inspiral \SM{[check if the PN is quantitatively ok -- overlay it on the plot ?]}.
-For the inspiral, the magnitude of the quadratic phase term $\epsilon_{\Psi 2}$ has a negative slope $v^{-1}=(Mf)^{-1/3}$ \SM{[check the slope on the figure]}, with a clear hierarchy that is due to the prefactor of $\alpha$ in the phase of the modulation, as explained above.
-This hierarchy is not respected for $--$.
-The mode $m=2$, the most important to model once the amplitudes are refactored in, is mainly in the perturbative regime, but we see that the perturbative treatment can be challenged for modes $m\leq 0$ and also for $m=1$ in the case $++$, especially for the post-merger.
+where $v = (G M \omega/c^{3})^{1/3}$ is related to the Fourier frequency $f$ by $v = (2/m')^{1/3} (GM \pi f/c^{3})^{1/3}$ for a mode $h^{\rm P}_{\ell m'}$.
+It is worth noting that $\epsilon_{\Psi 2}$ has an overall frequency scaling of $v^{-1}$, formally at $-0.5$PN order, which means that the relative size of this correction grows towards smaller frequencies, away from merger.
+Remember however that the quantities $\epsilon$ are meant to measure errors relatively to the leading order, and the opening angle of the precession cone, giving the overall normalization for precession effects in the waveform, also goes to 0 as $v$ in that limit.
+The amplitude error estimates $\epsilon_{A1}$, $\epsilon_{A2}$, by contrast, have the more usual behaviour of PN corrections growing towards merger.
+The geometric factors $(m' \cos\beta - m)$ show that mode contributions with a larger $|m-m'|$ are harder to model, again in a relative sense, but those contributions are however suppressed in amplitude (see discussion below~\eqref{eq:defprechIsum}).
+
+The overall $\nu$ scaling indicates a better separation of timescales for larger mass ratio systems.
+These expressions also show that, since the total angular momentum appears as a factor in $\Omega_{\rm prec}$, higher-order corrections will be larger for spin-aligned systems than for anti-aligned spins, an effect that becomes significant for large mass ratios.
+Note that the regime where $\xi$ gets close to 0 corresponds to the transitional precession range~\cite{Apostolatos+94}, with the spin of the primary compensating the orbital angular momentum, and our analysis based on simple precession is not valid anymore.
+
+We can somewhat complement this picture for the post-merger precession by considering the accelerated frame rotation rate described by the model~\eqref{eq:OmegaframeQNM} and illustrated in Fig.~\ref{fig:precmodel}.
+We show in Fig.~\ref{fig:QNM} the dependency of the QNM frequencies $\omega_{220}^{\rm QNM}$, $\omega_{210}^{\rm QNM}$ with the final spin of the remnant black hole $\chi_{f}$, together with the ratio $\Omega_{\rm frame} / \omega^{\rm QNM}_{220}$.
+This ratio is characteristic of the separation of timescales, in the ringdown regime, between the phase of the modulation~\eqref{eq:wignerphasesimpleprec} and the P-frame waveform phase, and increases monotonically with $\chi_{f}$, reaching $\sim 0.4$ for $\chi_{f} \rightarrow 1$.
+We cannot translate readily the time-domain separation of timescales in the ringdown regime to the Fourier-domain error measures defined in~\eqref{eq:deffom}, as the time-to-frequency correspondence~\eqref{eq:deftf} does not reach times beyond the merger, as shown by Fig.~\ref{fig:tf}.
+However, a faster frame rotation will yield a more extended Fourier transform of the modulation~\eqref{eq:defG}, and will be more challenging to accomodate with the formalism of Sec.~\ref{sec:formalism}.
+
+To go beyond the above order-of-magnitude picture, we now present a numerical computation of the error estimates~\eqref{eq:deffom} for our post-merger extended precession model presented in Sec.~\ref{subsec:precmodel}, for the three examples summarized in Table.~\ref{tab:precexamples}.
+Here and in the following we consider only the $h^{\rm I}_{2 m}$ mode contributions induced by $h^{\rm P}_{22}$ for positive frequencies, knowing that the ones induced by $h^{\rm P}_{2,-2}$ for negative frequencies can be deduced using~\eqref{eq:hIlmposnegfreq}.
+We use analytic derivatives of the PhenomD phase and amplitude for the Fourier-domain based timescales $\Tf$, $T_{A1}$ and $T_{A2}$, and numerical derivatives for the time-domain modulation.
+
+\SM{[Describe the difference in morphology of the precession in the three cases ?]}
+%The first two cases, $++$ and $\perp\perp$, have \SM{[come back to this after having checked trajectories]} the usual behaviour of precession on a cone, with a larger opening angle for $\perp\perp$ due to the larger in-plane spin components.
+After merger, the case $++$, with a lerger remnant spin, displays a much faster frame rotation $\Omega_{\rm frame}$ than the other cases, as shown by Table~\ref{tab:precexamples}.
+The case $--$ is qualitatively different, close to the transitional precession regime, with a large and varying $\beta$ and a significant change in the direction of $\bm{J}$.
+
+The results are shown in Fig.~\ref{fig:fomprec}, using the merger frequency and the ringdown frequency, shown by the vertical lines, to give an idea of the separation between the inspiral and post-merger phases.
+We can take $\epsilon \sim 1$ as an order-of-magnitude indication of the breakdown of a perturbative treatment.
+An important point about Fig.~\ref{fig:fomprec} is that the error estimates are relative for each mode, thus higher $|m-m'|$ modes, which are found to be the hardest to model precisely, can be very subdominant in the final waveform.
+
+The magnitude of the error estimates for $++$ and $\perp\perp$ is roughly in agreement with the PN-inspired computation~\eqref{eq:precfomPN} above for the inspiral part of the signal \SM{[check if the PN is quantitatively ok -- overlay it on the plot ?]}, with $\epsilon_{\Psi 2}$ showing a negative slope $v^{-1}=(Mf)^{-1/3}$, and with a hierarchy between modes due to the factors of $|m' \cos \beta - m|$.
+The case $--$ departs from the simple precession picture, and~\eqref{eq:precfomPN} does not apply.
+Before merger, we see that amplitude corrections are within the preturbative regime, contrarily to $\epsilon_{\Psi 2}$ that can exceed $1$ for subdominant modes in the $++$ and $\perp\perp$ cases.
 The amplitude corrections are found to be in the perturbative regime for all cases during the inspiral, however both the $++$ and $\perp\perp$ cases show a sharp increase of $\epsilon_{A1}$, $\epsilon_{A2}$ post-merger.
@@
 
-Overall, the conclusion to be drawn of Fig.~\ref{} is that for the $++$ and $\perp\perp$ cases we expect the perturbative approach to be applicable only during the inspiral, with a possible breakdown for the post-merger phase, especially for the case $++$ with its fast post-merger frame rotation.
-The higher $|m-m'|$ modes are expected to be more challenging to compute accurately, including during the inspiral.
+Overall, the conclusion to be drawn of Fig.~\ref{fig:fomprec} is that for the $++$ and $\perp\perp$ cases we expect the perturbative approach to be applicable only during the inspiral, with a possible breakdown for the post-merger phase, especially for the case $++$ with its fast post-merger frame rotation.
+The higher $|m-m'|$ modes are expected to be more challenging for the perturbative formalism, including during the inspiral.
 To make this picture quantitative, we need a full comparison of the signals processed at different orders of approximation against a numerical Fourier transform, which will be presented in Sec.~\ref{subsec:precerror} below.
@@
   \includegraphics[width=.98\linewidth]{plots/precfom_py.pdf}
-  \caption{Error estimates of the approximation as defined in~\eqref{eq:deffom} for the precession modulations. \SM{[complete caption]}}
+  \caption{Error estimates of the approximation as defined in~\eqref{eq:deffom} for the precession modulations, for the three cases listed in Table~\ref{tab:precexamples} \SM{[complete caption]}}
   \label{fig:fomprec}
@@
   \includegraphics[width=.98\linewidth]{plots/trigopoly_py.pdf}
-  \caption{Amplitude (left panel) and phase (right panel), compared to its trigonometric polynomial representation for the three cases listed in Table~\ref{table:precexamples}.
-The vertical lines indicate the merger frequencies.
-\SM{[check definition of A]}}
+  \caption{Amplitude (left panel) and phase (right panel), compared to its trigonometric polynomial representation for the three cases listed in Table~\ref{tab:precexamples}.
+The thick and thin vertical lines indicate the frequency of the merger and ringdown (the QNM frequency) respectively.
+\SM{[thin lines ?]} \SM{[check definition of A]}}
   \label{fig:trigopoly}
@@
 \subsection{Error control for the Fourier-domain modulation}
-\label{subsec:errorsPrec}
-
-We now assess the accuracy of the transfer function computation, applying the formalism of Secs.~\ref{} and \ref{} to our three examples listed in Table~\ref{table:precexamples} and comparing the result to a reference numerical computation.
+\label{subsec:precerror}
+
+We now assess the accuracy of the transfer function computation, applying the formalism of Secs.~\ref{} and \ref{} to our three examples listed in Table~\ref{tab:precexamples} and comparing the result to a reference numerical computation.
 
@@
 
+\SM{[discuss the shortcoming of our LISA treatment for precessing signals]}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{ruledtabular}\caption{Stencil coefficients $a_{N,k}^{\epsilon}$ entering the formula~\eqref{eq:stencilfresnel}, given here for $\epsilon=1$.
-The case $\epsilon=-1$ is obtained by complex conjugation.}\label{table:stencil}
+The case $\epsilon=-1$ is obtained by complex conjugation.}\label{tab:stencil}
 \begin{tabular}{c|cccccc}
@@
 The two cases $\epsilon\pm 1$ correspond simply to a complex conjugation of the coefficients $a_{N,k}^{\epsilon}$.
-Table~\ref{table:stencil} gives the resulting complex rational values for these coefficients for $N\leq 5$.
+Table~\ref{tab:stencil} gives the resulting complex rational values for these coefficients for $N\leq 5$.
 
 \section{Precession in current Fourier-domain waveform models}
-\label{subsec:precpreviousapproaches}
-
-\SM{[citation O'Shaugnessy]}
+\label{app:precpreviousapproaches}
 
@@
 In the PhenomP model~\cite{Hannam+13}, the frequency-based transfer function~\eqref{eq:precPhenomP} is used as an effective prescription covering the whole Fourier-domain frequency band, including the merger and ringdown.
-Note that this procedure amounts to using PN \SM{[check acronym]} results outside of their range of validity, since in the merger-ringdown phase both the PN perturbative treatment and the SPA approximation break down.
+Note that this procedure amounts to using PN results outside of their range of validity, since in the merger-ringdown phase both the PN perturbative treatment and the SPA approximation break down.
 In particular, when computed as a function of time from the balance equation between emitted flux and orbital energy $\calF = -dE/dt$, the orbital $\omega^{\rm PN}(t)$ blows up and cannot be extended through merger.

### 2018-06-02 docs/paper_PE.tex

@@
 eg:
+
+Case 12?
 m1,    m2,    t0,D,     phi0,i,   lam,  bet, psi
 8.89e5,1.11e5,0, 3.67e4,pi/3,pi/2,3pi/4,pi/3,pi/3
+
+Case 0
 

### 2018-06-05 docs/paper_method.tex
Edits of secs 5a,5b

@@
 
-We will now turn to the application of our formalism to the decomposition of signals from precessing binaries.
+\sout{We will next apply our formalism to the decomposition of signals from precessing binaries.}
 As we already discussed in Sec.~\ref{subsec:modulationPrec}, in the presence of spin components that are not aligned with the orbital angular momentum, an inspiraling binary system will undergo precession of its orbital plane~\cite{Apostolatos+94, Kidder95}.
@@
 As proposed by several authors~\cite{BCV03b, BCPTV05, Schmidt+10, OShaughnessy+11, Boyle+11}, if one performs the mode decomposition of the waveform~\eqref{eq:defmodes} not in a fixed inertial frame but rather in a time-dependent, rotating frame that follows the plane of the orbit, it is possible to restore much of the structure of a non-precessing waveform.
-In particular, one recovers qualitatively the hierarchy of mode amplitudes that prevails for non-precessing systems, with the modes $h_{22}$ and $h_{2,-2}$ being dominant.
+In particular, one recovers qualitatively the hierarchy of mode amplitudes that prevails for non-precessing systems, with the modes $h_{22}$ and $h_{2,-2}$ being dominant\jgb{, and with each harmonic mode characterized by a chirp with smoothly evolving amplitude and phase}.
 In the following, we will identify the $z$-axis of the co-precessing frame as the dominant eigenvector of the matrix representing the action of the angular momentum operator acting on the waveform modes, as proposed in~\cite{OShaughnessy+11} (see App.~\ref{app:wigner} for more details).
 
-We define $(\alpha, \beta, \gamma)$ are the Euler angles of the active rotation from the inertial frame to the precessing frame, in the $(z,y,z)$ convention.
+We define $(\alpha, \beta, \gamma)$ as the Euler angles of the active rotation from the inertial frame to the precessing frame, in the $(z,y,z)$ convention.
 The first two angles, $\alpha$ and $\beta$, are the two spherical angles tracking the direction of the radiation axis, which during the inspiral follows essentially the normal to the orbital plane.
@@
 \ee
-Here, the Wigner d-matrix $d^{\ell}_{mm'}(\beta)$ takes the form of a polynomial in $\cos (\beta/2)$, $\sin (\beta/2)$, and acts as an amplitude for the modulation function.
+Here, the \jgb{real-valued} Wigner d-matrix $d^{\ell}_{mm'}(\beta)$ takes the form of a polynomial in $\cos (\beta/2)$, $\sin (\beta/2)$, and acts as an amplitude for the modulation function.
 We refer to App.~\ref{app:wigner} for explicit expressions.
@@
 
-An important qualitative observation is that, for small to moderate opening angles, the modulation functions show a clear hierarchy in amplitude.
-This situation is quite generic, since large misalignments between $\bm{J}$ and $\bm{\ell}$ can only be reached with large spins and large mass ratios.
-From the explicit expression~\eqref{eq:defWignerdapp}, one can see that $d^{\ell}_{mm'}(\beta)$ is greatly suppressed in the limit $\beta \rightarrow 0$ when increasing $|m-m'|$.
+\jgb{An important qualitative observation is that the opening angle $\beta$ is typically small.
+Large misalignments between $\bm{J}$ and $\bm{\ell}$ can only be reached with large spins and large mass ratios.
+When $\beta$ is small,} from the explicit expression~\eqref{eq:defWignerdapp}, one can see that $d^{\ell}_{mm'}(\beta)$ is greatly suppressed in the limit $\beta \rightarrow 0$ when increasing $|m-m'|$.
 Intuitively, this means that, in the limit of a small misalignment between frames, the rotation produces mainly mode contributions with the same mode number.
@@
 
-When approximating the precessing-frame waveform by a non-precessing one, one can use the assumption (exactly valid for non-precessing systems, see~\eqref{eq:symmetryhlminusm})
+\jgb{The dominant features of the the precessing-frame waveform, are approximtely reflection symmetric about the orbital plane (exactly valid for non-precessing systems, see~\eqref{eq:symmetryhlminusm}), implying}
 \be\label{eq:approxsymmetryhlminusm}
@@
 
+\jgb{Some attention is required in specifying an model for the post merger precession.}
 For the post-merger precession, the intuitive interpretation of the precessing frame as roughly following the orbital plane of the binary is lost.
-However, the prescription of~\cite{OShaughnessy+11} to extract the radiation axis is based entirely on the waveform itself, and can be applied to the ringdown as well.
+However, the prescription of~\cite{OShaughnessy+11} to extract the radiation axis is based entirely on the waveform itself,\jgb{ [Should state the principle, minimizing m=0 mode?]} and can be applied to the ringdown as well.
 For $\ell=2$, analyzing numerical relativity waveforms, Ref.~\cite{OShaughnessy+12} described a qualitative model where the radiation axis essentially keeps precessing around the final angular momentum $\bm{J}$, but transitions to a faster precession rate, with an angular velocity determined by the quasinormal mode (QNM) frequencies of the remnant black hole as
@@
 \ee
-This translates in general in a significant acceleration of the precession post-merger with respect to the inspiral (see Fig.~\ref{fig:precmodel}).
+This \jgb{implies} a significant acceleration of the precession post-merger with respect to the inspiral (see Fig.~\ref{fig:precmodel}).
 For the opening angle of the precessing cone $\beta$, Ref.~\cite{OShaughnessy+12} proposes an exponential decay driven similarly by the difference between damping frequencies.
-We found by a non-exhaustive \SM{[allowed to say that ?
-the catalog is vast now]} inspection of precessing waveforms currently available in the SXS catalog (see e.g.~\cite{SXScatalog, Mroue+12, Mroue+13}) that this picture is at least qualitatively correct for $\ell=2$, with $\beta$ appearing to transition to a lower value rather than decaying all the way to zero.
+We found by \sout{a non-exhaustive \SM{[allowed to say that ?
+the catalog is vast now]}} inspection of \jgb{various} precessing waveforms currently available in the SXS catalog (see e.g.~\cite{SXScatalog, Mroue+12, Mroue+13}) that this picture is at least qualitatively correct for $\ell=2$, with $\beta$ appearing to transition to a lower value rather than decaying all the way to zero.
 By contrast, we found that the current ringdown prescription in the SEOBNRv3 model can cause a qualitatively different behaviour, with oscillations in $\beta$.
@@
 
-We would now like to underline a number of caveats and limitations of our simplified model.
+\jgb{In this paper, we will need this model only as an example for demonstrating our frequency-domain precession response, but this or a similar approach may be also useful in borader gravitational wave analysis applications.
+In that context,}
+we would now like to underline a number of caveats and limitations of our simplified model.
 First, it is clear from Fig.~\ref{fig:precmodel} that the acceleration of the precession post-merger is going to be the most challenging feature for the separation of timescales at the basis of Sec.~\ref{sec:formalism}.

### 2018-06-05 docs/paper_method.tex
Edits to the paper.

--- 2018-06-02 docs/paper_method.tex
@@
 \begin{widetext}
-\begin{subequations}
+\begin{subequations}\label{eq:summaryNAd}
 \begin{align}
@@
 In the context of precessing binaries, the delays are absent and we will go up to $k=2$ in~\eqref{eq:transferfinal}. The tranfer function at different orders of approximation will be:
-\begin{subequations}
+\begin{subequations}\label{eq:summaryNA}
 \begin{align}
@@
 \ee
+In the notations that we used in Sec.~\ref{sec:formalism}, in the absence delays, we wish to compute the convolutions
+\be\label{eq:precconvolution}
+	\tilde{s} (f) = \int df' \; \tilde{F}(f') \tilde{h} (f-f') \,,
+\ee
+with $\tilde{s}(f)$ being one mode contribution to $\tilde{h}_{\ell m}^{\rm I} (f)$ in~\eqref{eq:defprechIsum}, $\tilde{h}$ being one of the P-frame modes $\tilde{h}_{\ell m'}^{\rm P} (f)$, and the modulation $F(t)$ being one of the time-dependent Wigner matrices $\calD^{\ell *}_{mm'}$.
 
@@
 
-Further simplifications occur when including only the dominant harmonics $h^{\rm P}_{22}$, $h^{\rm P}_{2,-2}$ in the precessing-frame waveform (as is done for instance in PhenomP~\cite{Hannam+13}, as well as in our toy model~\ref{subsec:precmodel}), we have in this case for $f<0$ \SM{[all $f$ ?]}
-\begin{subequations}
-\begin{align}
-	\calT^{2}_{m,-2}(f) &= (-1)^{m} \calT^{2 *}_{-m,2}(-f) \,, \\
-	\tilde{h}^{\rm I}_{2 m}(f) &= (-1)^{m} \tilde{h}^{\rm I *}_{2,-m}(-f) \,. \label{eq:symmetryhIfor22only}
-\end{align}
-\end{subequations}
-With $m' = 2$ in the approximate phase given in Eq.~\eqref{eq:wignerphasesimpleprec}, the coefficient of $\alpha$ is close to $0$ for small $\beta$ for $m=2$, and increasingly positive when going down from $m=1$ to $m=-2$.
+Further simplifications occur when including only the dominant harmonics $h^{\rm P}_{22}$, $h^{\rm P}_{2,-2}$ in the precessing-frame waveform (as is done for instance in PhenomP~\cite{Hannam+13}, as well as in our toy model~\ref{subsec:precmodel}), we have in this case for $f<0$
+\be
+	\calT^{2}_{m,-2}(f) = (-1)^{m} \calT^{2 *}_{-m,2}(-f) \,,
+\ee
+which for both $f>0$ and $f<0$ translates into
+\be\label{eq:symmetryhIfor22only}
+	\tilde{h}^{\rm I}_{2 m}(f) = (-1)^{m} \tilde{h}^{\rm I *}_{2,-m}(-f) \,.
+\ee
+When reconstructing the polarizations $h_{+},h_{\times}$ according to~\eqref{eq:hpcfrommodes}, we have in this case
+\be
+	\tilde{h}_{+,\times} (f) = \sum_{m = -2}^{2} L^{+,\times}_{2 m} \tilde{h}^{\rm I}_{2 m}(f) \,,
+\ee
+where we defined
+\begin{align}
+	L^{+}_{\ell m} &\equiv \frac{1}{2} \left( {}_{-2}Y_{\ell m} + (-1)^{m} {}_{-2}Y_{\ell, -m}^{*} \right) \,, \nn\\
+	L^{\times}_{\ell m} &\equiv \frac{i}{2} \left( {}_{-2}Y_{\ell m} - (-1)^{m} {}_{-2}Y_{\ell, -m}^{*} \right) \,.
+\end{align}
+
+Finally, we note that with the restriction to $\ell = 2$ and $m' = 2$, in the approximate phase given in Eq.~\eqref{eq:wignerphasesimpleprec}, the coefficient of $\alpha$ is close to $0$ for small $\beta$ for $m=2$, and increasingly positive when going down from $m=1$ to $m=-2$.
 In our Fourier convention~\eqref{eq:defFT}, the Fourier transforms $\tilde{F}$ thus have most of their support on negative frequencies.
-This means that the support of the convolution integral~\eqref{eq:FDkernel} extends to the right side for $\tilde{h}$, i.e.
+This means that the support of the convolution integral~\eqref{eq:precconvolution} extends to the right side for $\tilde{h}$, i.e.
 $f-f' > f$.
 This point will become important when considering the high-frequency part of the signal in Sec.~\ref{subsec:trigopoly}.
+Note, however, that this statement might not be true anymore when including more modes in our model.
 
@@
 
+\SM{[Describe the difference in morphology of the precession in the three cases ?]}
 We can somewhat complement this picture for the post-merger precession by considering the accelerated frame rotation rate described by the model~\eqref{eq:OmegaframeQNM} and illustrated in Fig.~\ref{fig:precmodel}.
@@
 
-\SM{[Describe the difference in morphology of the precession in the three cases ?]}
 %The first two cases, $++$ and $\perp\perp$, have \SM{[come back to this after having checked trajectories]} the usual behaviour of precession on a cone, with a larger opening angle for $\perp\perp$ due to the larger in-plane spin components.
@@
 
-The magnitude of the error estimates for $++$ and $\perp\perp$ is roughly in agreement with the PN-inspired computation~\eqref{eq:precfomPN} above for the inspiral part of the signal \SM{[check if the PN is quantitatively ok -- overlay it on the plot ?]}, with $\epsilon_{\Psi 2}$ showing a negative slope $v^{-1}=(Mf)^{-1/3}$, and with a hierarchy between modes due to the factors of $|m' \cos \beta - m|$.
+The magnitude of the error estimates for $++$ and $\perp\perp$ is roughly in agreement with the PN-inspired computation~\eqref{eq:precfomPN} above for the inspiral part of the signal \SM{[check if the PN is quantitatively ok -- overlay it on the plot ?]} \SM{[Pb: scaling of FOM might be dominated here by SS oscillations, notably for 2nd der.
+?
+To check - check also geometric scaling from mode to mode]}, with $\epsilon_{\Psi 2}$ showing a negative slope $v^{-1}=(Mf)^{-1/3}$, and with a hierarchy between modes due to the factors of $|m' \cos \beta - m|$.
 The case $--$ departs from the simple precession picture, and~\eqref{eq:precfomPN} does not apply.
@@
 To make this picture quantitative, we need a full comparison of the signals processed at different orders of approximation against a numerical Fourier transform, which will be presented in Sec.~\ref{subsec:precerror} below.
-
-\SM{[Pb: scaling of FOM might be dominated here by SS oscillations, notably for 2nd der. ? To check - check also geometric scaling from mode to mode]}
 
@@
   \includegraphics[width=.98\linewidth]{plots/precfom_py.pdf}
-  \caption{Error estimates of the approximation as defined in~\eqref{eq:deffom} for the precession modulations, for the three cases listed in Table~\ref{tab:precexamples} \SM{[complete caption]}}
+  \caption{Error estimates of the approximation as defined in~\eqref{eq:deffom} for the precession modulations, for the three cases $++$, $\perp\perp$ and $--$ listed in Table~\ref{tab:precexamples}.
+The thick and thin vertical lines show, respectively, the merger frequency and the asymptotic ringdown frequency.
+The colors correspond to different values of $m$ in $\calT^{2}_{m 2}$ as defined in~\eqref{eq:defprectransfer}.
+We show the three error estimates $\epsilon_{\Psi 2}$, $\epsilon_{A1}$, $\epsilon_{A2}$ defined in~\eqref{eq:deffom}, and the range $\epsilon \gtrsim 1$ indicates a breakdown of the formalism of Sec.~\ref{sec:formalism}.
+\SM{[overlay PN simple precession ?]}}
   \label{fig:fomprec}
@@
 The thick and thin vertical lines indicate the frequency of the merger and ringdown (the QNM frequency) respectively.
-\SM{[thin lines ?]} \SM{[check definition of A]}}
+The continuous line show the target signal $\tilde{h}(f)$, while the dashed line shows the artificially symmetrized trigonometric polynomial~\eqref{eq:hsymtrigo}.
+The dots show the discrete samples entering~\eqref{eq:ykDFTprec}.
+The lower panels show the amplitude and phase residuals of~\eqref{eq:hsymtrigo} compared to the oringial $\tilde{h}(f)$.}
   \label{fig:trigopoly}
@@
 
-As will be shown by Fig.~\ref{} below, although including higher-order amplitude corrections does capture some effects near merger, it is not sufficiently accurate and seems to be too sensitive to the details of the evolution of the modulation $F(t)$ in the merger region.
-Motivated by this shortcoming of the Taylor-like expansion approach, here we investigate an alternative way of handling the merger-ringdown part of the signal.
-
-Thanks to the separation of timescales between the precessional timescale and the orbital timescale, for sufficiently high frequencies the convolution~\eqref{eq:convolution} will have a support that will be entirely comprised in the high-frequency part of the signal $\tilde{h}(f)$, which is rather featureless and slowly varying as a function of $f$.
-Taking advantage of this smoothness, we will adopt a trigonometric polynomial representation for $\tilde{h}(f)$.
-For frequencies high enough that the support of the convolution~\eqref{eq:convolution} does not extend beyond the range covered by this trigonometric representation, we will be able to write the result directly as a Discrete Fourier Transform (DFT) with a limited number of samples.
+As shown by Fig.~\ref{fig:fomprec}, the faster evolution of the modulation functions in the post-merger phase and the resulting weaker separation of timescales can be expected to be challenging for the perturbative formalism layed out in Sec.~\ref{sec:formalism}.
+Motivated by this forecasted shortcoming of the Taylor-like expansion approach, here we investigate an alternative way of handling the merger-ringdown part of the signal.
+
+In the correspondence between the time and Fourier domain, sharp features in the time domain map to extended features in the Fourier domain, and vice versa.
+The merger and ringdown part of the Fourier-domain signal extends over a wide range of frequencies, while it corresponds to a short interval of times, much shorter than the inspiral phase.
+One approach would be to separate the waveform between the inspiral and merger phase, build a Fourier-domain model for the inspiral part only, to which the formalism of Sec.~\ref{sec:formalism} could be applied, while the merger and ringdown part could be handled by a direct DFT \SM{[acronym]}, which would be a cheap operation on a limited range in time.
+Although this approach should be the most robust, here we will use an alternative, simpler to implement and keeping the setting of a Fourier-domain precessing-frame waveform combined with a time-domain precession modulation.
+
+As argued in Sec.~\ref{subsec:precdef}, the support of the convolution~\eqref{eq:precconvolution} will be mainly one-sided towards the high-frequency part of $\tilde{h}(f)$, which is featureless and slowly varying as a function of $f$.
+Taking advantage of this, we will adopt a one-sided trigonometric polynomial representation for $\tilde{h}(f)$.
+For frequencies high enough that the support of the convolution~\eqref{eq:precconvolution} does not extend beyond the range covered by this trigonometric representation, the result will be obtained directly as a DFT \SM{[acronym]} with a limited number of samples.
+The limitation to the high-frequency range is crucial here: the Fourier-domain amplitude and phase diverge as $f^{-7/6}$ and $f^{-5/3}$ respectively for $f \rightarrow 0$, and our procedure would require a much finer sampling to cover part of the inspiral, going back to being equivalent to an inverse DFT of the full signal if we were to cover all frequencies.
 
@@
 In practice, we define $f_{\rm knee}$ as the peak of $f^{2}A(f)$, representing the onset of the decay in amplitude, and roughly corresponding to $\omega_{22}^{\rm QNM}/\pi$.
-We set $f_{0}\equiv 2/3f_{\rm knee}$, which is close to the merger frequency, and $f_{\rm max}$ is chosen so that the amplitude is $10^{-4}$ of the amplitude at $f_{\rm knee}$.
+We set $f_{0}\equiv 2/3f_{\rm knee}$, which is close in practice to the merger frequency, and $f_{\rm max}$ is chosen so that the amplitude is $10^{-4}$ of the amplitude at $f_{\rm knee}$.
 We also eliminate a constant and a linear term in the phase by choosing another frequency central to the high-frequency range we want to represent, which we take to be $f_{p} = f_{\rm knee}^{2/3} f_{\rm max}^{1/3}$ with corresponding time $t_{p}\equiv \tf(f_{p})$.
@@
 
-Additionally, we taper the amplitude to artificially flatten the amplitude at $f_{0}$, at the price of introducing a deviation from the amplitude of the original signal.
-This tapering is built in practice as the discrete integral of a cosine window function on just the first two samples.
-We taper to 0 the last three samples before $f_{\rm max}$.
-To ensure good continuity properties, we perform an artificial symmetrization to a fictitious range $f\in [f_{0} - (f_{\rm max} - f_{0}), f_{0}]$, by imposing symmetric amplitudes and phases about $f_{0}$.
+Instead of tapering the signal to $0$ at $f_{0}$, to limit the deviation from the original signal we take advantage of the one-sidedness and only flatten the amplitude\footnote{In practice, this is implemented as the discrete integral of a cosine window function on just the first two samples.
+We also taper to 0 the last three samples before $f_{\rm max}$, and $0$-pad by a factor of 2.}, and artificially symmetrize the signal.
+To ensure continuity, this artificial symmetrization to a fictitious range $f\in [f_{0} - (f_{\rm max} - f_{0}), f_{0}]$ is done by imposing symmetric amplitudes and phases about $f_{0}$.
 Defining $\Delta f \equiv 2 (f_{\rm max} - f_{0})$, we write
@@
 
-Next, we build a a trigonometric polynomial representation of $\tilde{h}_{\rm sym }$.
-This construction is intimately related to the DFT, but we recall it for completeness.
-For a periodic function $F(x)$ defined on $x\in [x_{0}, x_{0} + \Delta x]$, and represented by $N$ samples $x_{j} = \ov{x} + j \Delta x/N$ (with $N$ chosen to satisfy the Nyquist criterion \SM{[historically should be called Shannon ?]}), we can build a trigonometric interpolant $P(x)$ as
-\be
-	P(x) = \sum\limits_{k=-M}^{+M} c_{k} e^{2i\pi k \frac{x-x_{0}}{\Delta x}}
-\ee
-that will satisfy the system $P(x_{j}) = F(x_{j})$ for $j=0,\dots, N-1$.
-Here we set $M=N/2$ (assuming N is even).
-The coefficients $c_{k}$ are related to the coefficients of the inverse DFT of the series for $F$.
-If we set $\omega \equiv e^{2i\pi/N}$ and define
-\be
-	y_{k} = \frac{1}{N} \sum\limits_{j=0}^{N-1} F(x_{j}) \omega^{jk} \,,
-\ee
-which is the expression of the inverse DFT in our sign convention~\eqref{eq:defFT}, the coefficients $c_{k}$ are given by \SM{[move generic trigopoly things to the appendix, as it is common with comb approach for LISA]}
-\begin{align}\label{eq:ckyk}
-	c_{k} &= y_{k} \text{ for } k=0,\dots, M-1 \,, \nn\\
-	c_{k} &= y_{k+N} \text{ for } k=-M+1,\dots, -1 \,, \nn\\
-	c_{M} &= c_{-M} = \frac{y_{M}}{2} \,,
-\end{align}
-where the condition $c_{M} = c_{-M}$ is an arbitrary condition enforced to match the number of degrees of freedom.
-In practice, we also 0-pad by a factor of 2 to high frequencies.
-Fig~\ref{fig:convolsymmetrized} shows both the original high-frequency part of the signal and its reconstruction as a trigonometric polynomial.
-
-Now, assuming that we wish to compute~\eqref{eq:convolution} for a frequency $f$ such that the extent of the convolution does not extend beyond $[f_{0}, 2f_{\rm max} - f_{0}]$, we can approximate
+Next, we build a a trigonometric polynomial representation of $\tilde{h}_{\rm sym}$, a construction intimately related to the DFT, that we recall in App.~\ref{app:notation}.
+Over the frequency range $f\in [f_{0}, 2f_{\rm max} - f_{0}]$, we can approximate
 \be\label{eq:hsymtrigo}
-	\tilde{h}_{\rm sym } (f) \simeq \sum\limits_{k=-M}^{+M} (-1)^{k} c_{k} e^{2i\pi k \frac{f-f_{0}}{\Delta f}} \,,
+	\tilde{h}(f) \simeq \tilde{h}_{\rm sym } (f) \simeq \sum\limits_{k=-M}^{+M} (-1)^{k} c_{k} e^{2i\pi k \frac{f-f_{0}}{\Delta f}} \,,
 \ee
@@
 The coefficients $c_{k}$ are built following the rules~\eqref{eq:ckyk} from the inverse DFT coefficients
-\be\label{eq:ykDFT}
+\be\label{eq:ykDFTprec}
 	y_{k} = \frac{1}{N} \sum\limits_{j=0}^{N-1} \tilde{h}_{\rm sym}\left( f_{0} + \frac{2j - N}{N} \Delta f \right) \omega^{jk} \,.
 \ee
-When inserting this representation~\eqref{eq:defhsym} and~\eqref{eq:hsymtrigo} into~\eqref{eq:convolution}, we obtain
+Taking the point of view of an interpolation problem, Fig~\ref{fig:trigopoly} shows the accuracy of this representation of the high-frequency part of the signal, by comparing the orginal $\tilde{h}(f)$ to its  trigonometric-polynomial representation~\eqref{eq:hsymtrigo}, for the three cases listed in Table~\ref{tab:precexamples}.
+We see that we can achieve a good agreement already for 32 samples (128 counting the symmetrization and 0-padding) \SM{[check this number]}.
+Errors in the phase $\Psi$ grow towards high frequencies because they are essentially errors in a relative sense, and amplitudes are decaying in this region.
+
+The crucial point in this approach is that the convolution integral in~\eqref{eq:precconvolution} has support mainly on $f'<0$, as discussed in Sec.~\ref{subsec:precdef}, which means that when we try to compute the transfer function $\calT(f)$ at a given frequency $f$, we only need the trigonometric-polynomial representation~\eqref{eq:hsymtrigo} to be accurate for frequencies $>f$ in Fig.~\ref{fig:trigopoly}.
+Thus, the trigonometric representation of $\tilde{h}(f)$ can be used to compute the convolution almost all the way down to $f_{0}$, effects of the tapering aside.
+However, this statement is tied to our restriction to the $22$-mode, and extending the method to a precessing-frame waveform that includes more modes $h^{\rm P}_{\ell m}$ will require care.
+
+When inserting this representation~\eqref{eq:defhsym} and~\eqref{eq:hsymtrigo} into~\eqref{eq:precconvolution}, we obtain
 \begin{align}\label{eq:resultdirectconvol}
@@
 We see that we are left with a DFT-like expression to compute from $N+1$ time samples of the modulation function $F$, centered around $t_{p}$.
-
-The crucial point in this approach is that the convolution integral in~\eqref{eq:convolution} has support mainly on $f'<0$, which means to the right of $f$ in Fig.~\ref{fig:trigopoly} as discussed in Sec.~\ref{subsec:} \SM{[R]}.
-This means that the trigonometric representation of $\tilde{h}(f)$ can be used to compute the convolution almost all the way down to $f_{0}$, effects of the tapering aside.
-
-In terms of performance, the implementation of~\eqref{eq:ykDFT} and~\eqref{eq:resultdirectconvol} above is expected to have a very reasonable cost.
-We obtained results with a good accuracy, as shown in Fig.~\ref{fig:trigopoly}, by including $M=64$ samples ($32$ useful samples before 0-padding, $N=128$ samples in total for the artificially symmetrized signal).
-The calculation described here is done only once for a given waveform and modulation function.
-The number of samples is of the same order of magnitude as the number of (logarithmically-spaced) frequency samples used to represent the Fourier-domain waveform with its amplitude and phase in the first place, thus the cost is comparable to the Taylor-like approach presented in Sec.~\ref{sec:quadphasedelay} \SM{[R]}.
-
-\SM{[update performance - plots were made with double number of points]}
-
-\SM{[explain what this all means translated in terms of sampling in the time-domain -- better relate this approach to simply IFFT]}
-
-\SM{[for plot illustrating reconstruction trigo of amplitude : what normalization to pick for A FD ?]}
+We see that, apart from the conditioning described above with the artificial symmetrization, this is analogous to a DFT of the product of the modulation with the time-domain signal obtained through an inverse DFT.
+
+In terms of computational performance, the implementation of this approach is expected to have a reasonable cost.
+In the following, we will use $M=64$ samples ($32$ useful samples before 0-padding, $N=128$ samples in total for the artificially symmetrized signal).
+The computation of~\eqref{eq:ykDFTprec} and~\eqref{eq:resultdirectconvol} amounts to two DFT operations, implemented through an FFT/IFFT, and is done only once for a given waveform and modulation function.
+The number of samples is of the same order of magnitude as the one required to represent the Fourier-domain waveform with an interpolating cubic spline for its amplitude and phase (a few hundreds for the full frequency band, see e.g~\cite{Puerrer14}), thus the cost should be comparable to the Taylor-like approach presented in Sec.~\ref{subsec:executivesummary}.
 
@@
 
-We now assess the accuracy of the transfer function computation, applying the formalism of Secs.~\ref{} and \ref{} to our three examples listed in Table~\ref{tab:precexamples} and comparing the result to a reference numerical computation.
-
-To numerically compute the transfer functions, we first have to perform an IFFT of our Fourier-domain P-frame waveform, to obtain $h^{\rm P}_{22}$ as a function of time.
-To mitigate the effect of the necessary tapering of the waveform when computing this numerical inverse Fourier transform, we apply a Planck-window tapering on the range $f\in 16-20 \Hz$, using a total mass of $M=20 \Msol$.
-\SM{[Update this to $f_{\rm min} = 10\Hz$]}.
+\SM{[update to $10 M_{\odot}$ and $10\mathrm{Hz}$ ?]}
+
+We now assess the accuracy of the transfer function computation, applying the formalism of Secs.~\ref{subsec:executivesummary} and~\ref{subsec:trigopoly} to the three examples listed in Table~\ref{tab:precexamples}, and comparing the result to a reference numerical computation.
+
+To numerically compute the reference transfer functions, we first have to perform an IFFT of our Fourier-domain P-frame waveform, to obtain $h^{\rm P}_{22}$ as a function of time.
+To mitigate the effect of the necessary tapering of the waveform when computing this numerical inverse Fourier transform, we apply a Planck-window tapering on the range $f\in 16-20 \Hz$ for a total mass of $M=20 \Msol$.
 In parallel, we generate an SEOB waveform with the appropriate length in time.
-Both are aligned to peak at $t=0$, we build the post-merger extended modulation functions as explained in Sec.~\ref{} \SM{[R]}, and we compute the Fourier-domain modes $h^{\rm I}_{2m}$ with an FFT.
-The transfer functions are then computed using~\eqref{eq:defprectransfer}.
-
-In the following, we will either present the errors obtained with the formalism of Sec.~\ref{sec:formalism} applied to the whole frequency band, including the merger and ringdown part of the signal, or cover the high-frequency range with the convolution treatment of Sec.~\ref{subsec:trigopoly}, with a Planckian window on $Mf\in [Mf_{0}, Mf_{\rm knee}]$ for a smooth transition between the two formalisms.
-
-The figures Fig.~\ref{fig:precerrors++}, Fig.~\ref{fig:precerrorspp} and Fig.~\ref{fig:precerrors--} display three successive approximations: $\{\Psi:0,A:0\}$, which is simply the leading-order transfer function~\eqref{eq:transferlocal}, ignoring all the corrections; $\{\Psi:2,A:2\}$, which incorporates both the phase corrections~\eqref{eq:Fresnelstencil}, using a stencil size $N=3$, and the amplitude corrections up to second order in~\eqref{eq:transferfinal}; and $\{\Psi:2,A:2,\text{conv}\}$, which is the same as the previous for the inspiral but uses the convolution formalism of Sec.~\ref{}.
-They show the Fourier-domain amplitude for each of the modes, both exact and from the reconstruction, the relative errors in amplitude, and the errors in phase.
-The errors are given in a relative sense, and the amplitude allows to visualize the mode hierarchy, and give an idea of how much the relative errors matter in an absolute sense.
-
-The $++$ case, as expected from the analysis of the previous section, is the most challenging, with large relative errors.
-Applying higher-order corrections improves the accuracy in the inspiral, but the most difficult modes $m=-1$ and $m=-2$ still show large errors.
-We also find that higher-order corrections do not reduce the errors in the merger-ringdown region, indicating a breakdown of the perturbative formalism as indicated by Fig.~\ref{fig:fomprec}.
-Using the convolution treatment improves the main modes $m=2$, $m=1$, but not the most challenging modes $m=-1$ and $m=-2$, which can be seen to depart from the perturbative treatment before the range covered by the convolution.
-Note however that, as we will show below by computing the full unfaithfulness these large errors for the higher modes do not affect much the full waveform.
-\SM{[explain how the convolution formalism fails - convolution support to the right, in low-amp region, not well controlled]}
-
-For the $\perp\perp$ case, we find a clearer picture.
+Both are aligned to peak at $t=0$, we build the post-merger extended modulation functions as explained in Sec.~\ref{subsec:precmodel}, compute the inertial-frame modes $h^{\rm I}_{2m}$ following~\eqref{eq:wignerrot} before computing their Fourier-domain counterparts $\tilde{h}^{\rm I}_{2m}$ with an FFT.
+The Fourier-domain transfer functions are then computed using~\eqref{eq:defprectransfer}.
+
+The figures Fig.~\ref{fig:precerrors++}, Fig.~\ref{fig:precerrorspp} and Fig.~\ref{fig:precerrors--} show the result for the cases $++$, $\perp\perp$ and $--$ of Table~\ref{tab:precexamples} at three successive approximations, following the notation of~\eqref{eq:summaryNA}: $\{N:0,A:0, \mathrm{No \; Conv.}\}$, which is simply the leading-order transfer function~\eqref{eq:transferlocal}, ignoring all the corrections; $\{N:3,A:2, \mathrm{No \; Conv.}\}$, which incorporates both the phase corrections~\eqref{eq:stencilfresnel}, using a stencil size $N=3$, and the amplitude corrections up to second order in~\eqref{eq:summaryNA}; and $\{\Psi:2,A:2,\text{conv}\}$, which is the same as the previous for the inspiral but uses the convolution formalism of Sec.~\ref{subsec:trigopoly} to cover the high frequency range, with a smooth transition in the shaded range.
+The panels show the Fourier-domain amplitude for each of the modes, both exact and from the reconstruction, the relative errors in amplitude, and the errors in phase.
+The errors here are relative to each mode, not to the dominant mode.
+Thus, the amplitude plots importantly allow to visualize the mode hierarchy, giving an idea of the impact of relative errors in the subdominant modes on the full waveform.
+
+The $++$ case, shown in Fig.~\ref{fig:precerrors++}, is the most challenging.
+As expected from the analysis of Sec.~\ref{subsec:sizecorrPrec}, the presence of strong aligned spins and a large spin of the remnant degrades the separation of timescales, and the reconstruction shows large relative errors, at least in subdominant mode contributions.
+Applying the higher-order corrections of Sec.~\ref{subsec:executivesummary} does improve the accuracy in the inspiral, but the most difficult modes $m=-1$ and $m=-2$ still show large errors.
+We also find that for this case higher-order corrections do not reduce the errors in the merger-ringdown region, consistently with the breakdown of the perturbative formalism indicated by Fig.~\ref{fig:fomprec}.
+Using the convolution treatment improves the main modes $m=2$, $m=1$, but not the most challenging modes $m=-1$ and $m=-2$, which can be seen to depart from the perturbative treatment before the range covered by the convolution and for which the convolution~\eqref{eq:precconvolution} extends to very high frequencies where our trigonometric polynomial representation of the signal is not accurate (see Fig.~\ref{fig:trigopoly}).
+Note however that, as will be shown below by unfaithfulness computations, these large relative errors for the subdominant modes do not affect much the full waveform.
+
+The $\perp\perp$ case is shown in Fig.~\ref{fig:precerrorspp}.
 Errors are increasing to larger $|m-m'|$, higher-order corrections improve the reconstruction but are unsufficient for the merger-ringdown, and the convolution works for all modes on the frequency range where it is applied.
@@
 
-For the $--$ case, the mode hierarchy is not respected, as the frame trajectory does not quite follow the picture of simple precession \SM{[mention angle Ji-Jf]}.
-The precession velocity is much milder, leading to smaller errors.
-Apart from some amplitude errors in the merger-ringdown region, we find that in this case even the leading-order treatment gives very reasonable results.
+In the case $--$, shown in Fig.~\ref{fig:precerrors--}, the mode hierarchy is not respected, as the frame trajectory does not quite follow the picture of simple precession, and the estimates~\eqref{eq:precfomPN} do not apply.
+The precession velocity is much milder, both in the inspiral and in the merger-ringdown range, thanks to a low remnant spin, and the separation of timescales is better, as shown in Fig.~\ref{fig:fomprec}.
+This leads to smaller errors, and, apart from some amplitude errors in the merger-ringdown region, we find that in this case even the leading-order treatment gives good results.
 
@@
 This unfaithfulness figure, although giving a simplified view of waveform inaccuracies, is commonly used to quantify disagreements between template families and to compare them to numerical relativity waveforms.
-For real signals $h_{1}$ and $h_{2}$, one introduces the usual noise-weighted scalar product
+For real signals $h_{1}$ and $h_{2}$, one introduces the usual noise-weighted scalar product~\cite{CF94}
 \be\label{eq:defoverlap}
@@
 \ee
-with $S_{n}(f)$ the noise power spectral density, for which we use the aLIGO ZDHP noise curve~\cite{} \SM{[R]}.
+with $S_{n}(f)$ the noise power spectral density, for which we use the aLIGO ZDHP noise curve~\cite{LIGOProspects13}.
+
+To define the unfaithfulness, different prescriptions are possible, taking into account or not th edetector response and optimizing over different sets of parameters.
+Here, we will use a simple prescription~\cite{Blackman+17a} applied directly at the level of $h_{+}$, $h_{\times}$, optimizing over time, phase, and polarization.
 
@@
 
-Overall, we find that all the corrections we investigated, perturbative in phase, perturbative in amplitude, and the convolution treatment, are relveant in lowering the unfaithfulness.
+Overall, we find that all the corrections we investigated, perturbative in phase, perturbative in amplitude, and the convolution treatment, are relevant in lowering the unfaithfulness.
 The only exception are second-order amplitude corrections that make little difference, and can even give worse results than first-order in some cases.
@@
 
-\SM{[discuss the shortcoming of our LISA treatment for precessing signals]}
+\SM{[discuss the extension of our LISA treatment for precessing signals]}
+
+\SM{[discuss other detectors, e.g. ET]}
+
+\SM{[we should investigate broader parameter space]}
 
@@
 \end{align}
+When both and~\eqref{eq:zeronegativef} apply, \eqref{eq:hpcfrommodes} becomes simpler. For $f>0$, we have then
+\be
+	\tilde{h}_{+,\times} (f) = \sum_{\ell \geq 2} \sum_{m = 1}^{\ell} K^{+,\times}_{\ell m} \tilde{h}_{\ell m}(f) \,,
+\ee
+where we set
+\begin{align}
+	K^{+}_{\ell m} &\equiv \frac{1}{2} \left( {}_{-2}Y_{\ell m} + (-1)^{\ell} {}_{-2}Y_{\ell, -m}^{*} \right) \,, \nn\\
+	K^{\times}_{\ell m} &\equiv \frac{i}{2} \left( {}_{-2}Y_{\ell m} - (-1)^{\ell} {}_{-2}Y_{\ell, -m}^{*} \right) \,.
+\end{align}
+the range $f<0$ being obtained readily, since $h_{+},h_{\times}$ are real quantities, from $\tilde{h}_{+,\times} (-f) = \tilde{h}_{+,\times}(f)^{*}$.
 
@@
 Note that, if we are considering real signals $s(t)$ (such as $h_{+}$, $h_{\times}$, or any detector output), we can focus  on the positive-frequencies range and reconstruct the negative frequencies as $\tilde{s}(-f) = \tilde{s}(f)^{*}$.
-For positive frequencies $f>0$, we will use the generic notation $h$ as
+For positive frequencies $f>0$, we will use the generic notation $h$ \SM{[not using the 'generic notation' anymore]} as
 \begin{align}\label{eq:notationhsignm}

### 2018-06-08 docs/paper_method.tex
Edits of the paper.

@@
 In our model, after reaching the time of peak amplitude as defined in~\eqref{eq:defsumamplitude}, we will use the prescription~\eqref{eq:OmegaframeQNM} for the post-merger frame precession and simply keep $\beta$ constant,
-\begin{subequations}
-\begin{align}\label{eq:eulerQNM}
+\begin{subequations}\label{eq:eulerQNM}
+\begin{align}
 	\alpha_{\rm post-merger}(t) &= \alpha(t_{\rm peak}) + (\omega_{220}^{\rm QNM} - \omega_{210}^{\rm QNM})(t-t_{\rm peak}) \,, \\
@@
 \jgb{In this paper, we will need this model only as an example for demonstrating our frequency-domain precession response, but this or a similar approach may be also useful in borader gravitational wave analysis applications.
-In that context,}
+In that context,} \SM{[make clear: goal is illustration]}
 we would now like to underline a number of caveats and limitations of our simplified model.
@@
 \SM{[Add amplitude/$\omega$ plots comparing NR/SEOB, and possibly PhenomD for the 22 mode ?
-Add Euler angles produced by IFFT of PhenomP ?]}}
+Add Euler angles produced by IFFT of PhenomP ?]} \SM{[remove SEOBext in the legend]} \SM{[confusion of dashed black lines]}}
   \label{fig:precmodel}
@@
 
-\subsection{Error control for the Fourier-domain modulation}
+\subsection{Error control for the Fourier-domain precession modulation}
 \label{subsec:precerror}
@@
 \SM{[update to $10 M_{\odot}$ and $10\mathrm{Hz}$ ?]}
+
+\SM{[the mismatch computation should be averaged over injected phase]}
 
@@
 
-The figures Fig.~\ref{fig:precerrors++}, Fig.~\ref{fig:precerrorspp} and Fig.~\ref{fig:precerrors--} show the result for the cases $++$, $\perp\perp$ and $--$ of Table~\ref{tab:precexamples} at three successive approximations, following the notation of~\eqref{eq:summaryNA}: $\{N:0,A:0, \mathrm{No \; Conv.}\}$, which is simply the leading-order transfer function~\eqref{eq:transferlocal}, ignoring all the corrections; $\{N:3,A:2, \mathrm{No \; Conv.}\}$, which incorporates both the phase corrections~\eqref{eq:stencilfresnel}, using a stencil size $N=3$, and the amplitude corrections up to second order in~\eqref{eq:summaryNA}; and $\{\Psi:2,A:2,\text{conv}\}$, which is the same as the previous for the inspiral but uses the convolution formalism of Sec.~\ref{subsec:trigopoly} to cover the high frequency range, with a smooth transition in the shaded range.
+The figures Fig.~\ref{fig:precerrors++}, Fig.~\ref{fig:precerrorspp} and Fig.~\ref{fig:precerrors--} show the result for the cases $++$, $\perp\perp$ and $--$ of Table~\ref{tab:precexamples} at three successive approximations, following the notation of~\eqref{eq:summaryNA}: $\{N:0,A:0, \mathrm{No \; Conv.}\}$, which is simply the leading-order transfer function~\eqref{eq:transferlocal}, ignoring all the corrections; $\{N:3,A:2, \mathrm{No \; Conv.}\}$, which incorporates both the phase corrections~\eqref{eq:stencilfresnel}, using a stencil size $N=3$, and the amplitude corrections up to second order in~\eqref{eq:summaryNA}; and $\{N:3,A:2,\text{conv}\}$, which is the same as the previous setting for the inspiral but uses the convolution formalism of Sec.~\ref{subsec:trigopoly} to cover the high frequency range, with a smooth transition in the shaded range.
 The panels show the Fourier-domain amplitude for each of the modes, both exact and from the reconstruction, the relative errors in amplitude, and the errors in phase.
@@
 This unfaithfulness figure, although giving a simplified view of waveform inaccuracies, is commonly used to quantify disagreements between template families and to compare them to numerical relativity waveforms.
-For real signals $h_{1}$ and $h_{2}$, one introduces the usual noise-weighted scalar product~\cite{CF94}
+To define the unfaithfulness, different prescriptions are possible, taking into account or not the detector orientation and optimizing over different sets of parameters.
+
+Here, we will use directly the wave polarizations $h_{+}$, $h_{\times}$, optimizing over time, phase, and polarization angle.
+For real or complex signals $a$, $b$, one introduces the usual Hermitian noise-weighted scalar product~\cite{CF94}
 \be\label{eq:defoverlap}
-	\left( h_{1} | h_{2} \right) = 4\text{Re} \int_{0}^{+\infty} \ud f \frac{\tilde{h}_{1}(f) \tilde{h}_{2}^{*}(f)}{S_{n}(f)} \,,
-\ee
-with $S_{n}(f)$ the noise power spectral density, for which we use the aLIGO ZDHP noise curve~\cite{LIGOProspects13}.
-
-To define the unfaithfulness, different prescriptions are possible, taking into account or not th edetector response and optimizing over different sets of parameters.
-Here, we will use a simple prescription~\cite{Blackman+17a} applied directly at the level of $h_{+}$, $h_{\times}$, optimizing over time, phase, and polarization.
-
-For spin-aligned waveforms, the unfaithfulness is customarily defined as the normalized overlap~\eqref{eq:defoverlap} above optimized over the time at coalescence and global phase of the signal.
-In the precessing case, different prescriptions coexist, differing by the extrinsic parameters that are optimized over.
-Here we adopt the same prescription that was used in the construction of a numerical relativity surrogate model for precessing waveforms~\cite{} \SM{[R]}.
-We refer to the Appendix.~\SM{[R]} of~\cite{} for more details \SM{[complete a bit here]}.
-
-Here we take our illustrative model of Sec.~\ref{} \SM{[R]} as the reference waveform, thus ignoring the inaccuracy introduced by our approximate treatment of post-merger precession and focusing only on the passage from time-domain to Fourier-domain.
-We show the unfaithfulness obtained for 3 inclination angles between the line-of-sight and the direction of the final $\bm{J}$, $0$ (face-on), $\pi/3$ and $\pi/2$ (edge-on), using different approximations (note that we consider more options than for the preceding error plots).
+	\left( a | b \right) = 2 \int \ud f \frac{\tilde{a}_{1}(f) \tilde{b}_{2}(f)^{*}}{S_{n}(f)} \,,
+\ee
+where the integral extends over all frequencies (in practice, two intervals $[-f_{\rm max}, -f_{\rm min}]$ and $[f_{\rm min}, f_{\rm max}]$).
+In the above, $S_{n}(f)$ the noise power spectral density, for which we use the aLIGO ZDHP noise curve~\cite{LIGOProspects13}.
+We use the notation $h \equiv h_{+} - i h_{\times}$ for the complex strain, and write $h[\delta t, \delta \Phi, \delta \psi]$ to represent a signal where we shifted the time of coalescence, the phase of the line of sight to the observer and the polarization angle.
+The dependencies in time and polarization are given by~\eqref{eq:shifttime} and $h[\delta \psi] = e^{2 i \delta \psi} h$, while the phase $\delta \Phi$ affects differently each waveform mode in~\eqref{eq:defmodes}.
+In the norm $(h|h)$, only the dependence on $\delta\Phi$ remains.
+For two signals $h_{1}$ and $h_{2}$, the mismatch between them is then defined as \SM{[check]}
+\be
+	\mathrm{MM} \equiv \mathrm{Min}_{\delta t, \delta \varphi, \delta \psi} \left( 1 - \frac{ \mathrm{Re} \left[ (h_{1}[\delta t, \delta \Phi, \delta \psi] | h_{2}) \right]}{\sqrt{(h_{1}|h_{1})[\delta\Phi]}\sqrt{(h_{2}|h_{2})}} \right)
+\ee
+where an optimization over a shift in time, phase and polarization is performed.
+This definition is the same as in~\cite{Blackman+17a} (see App.~D there), except that we added the optimization over the phase.
+The target unfaithfulness for waveform models depends on the application.
+For ground-based detectors like advanced LIGO and Virgo, it is often set to 1\% or 3\%, while detectors of the next generations, including LISA, will require a lower unfaithfulness for their higher SNR \SM{[acronym]} signals.
+
+In this unfaithfulness computation, we take our illustrative model of Sec.~\ref{subsec:precmodel} as the reference waveform, thus ignoring all limitations of our model for precession and focusing only on the passage from time-domain to Fourier-domain for a given waveform.
+We show in Fig.~\ref{fig:precunfaithfulness} the unfaithfulness obtained for three inclination angles between the line-of-sight and the direction of the final angular momentum $\bm{J}$, $0$ (face-on), $\pi/3$ and $\pi/2$ (edge-on).
+Again, we use different approximation levels in~\eqref{eq:summaryNA}, indicated by the color, including or not the treatment of Sec.~\ref{subsec:trigopoly} for the high frequencies as indicated by a continuous or dashed line.
 The total mass ranges from $20\Msol$ (where the inspiral dominates) \SM{[to be updated to $10 \Msol$]} to $400\Msol$ (where the merger-ringdown phase dominates).
-The symbol $\Psi:n$ with $n=0$ or $n=2$ means that the phase corrections~\eqref{eq:Fresnelstencil} are either ignored or included with a stencil size $N=3$.
-The symbol $A:n$ with $n=0,1,2$ indicates the order at which the amplitude corrections in~\eqref{eq:transferfinal} are included.
-Additionally, for all these approximation orders we either apply the perturbative formalism of Sec.~\ref{sec:formalism} to the whole frequency band (symbol ``no conv''), or treat the high-frequency part with the convolution formalism of Sec.~\ref{subsec:trigopoly} (symbol ``conv'').
-
-Note that the order of approximation $\{\Psi:0,A:0\}$ corresponds to the leading-order formula~\eqref{eq:transferlocal}, which is close (but not identical; see the discussion in Sec.~\ref{} \SM{[R]}) to the PhenomP~\cite{} \SM{[C]} treatment.
-The order of approximation $\{\Psi:2,A:0\}$ is equivalent to formalism of Ref.~\cite{KCY14} with a stencil order of $N=3$, except that the latter formalism is based on an SPA representation of the P-frame waveform, which we bypass by keeping to the Fourier-domain, as discussed above in Sec.~\ref{} \SM{[R]}.
-
-For the $++$ case, the face-on computation (where only the $h^{\rm I}_{22}$ mode contributes) shows good agreement even at leading order.
-By contrast, with $\pi/3$ or $\pi/2$ inclination the mismatch reaches $10^{-2}$ for high masses, even with the perturbative corrections.
-Using the convolution limits the mismatch, and the inclusion of amplitude corrections makes a clear improvement.
-
-For the $\perp\perp$ case, both the perturbative corrections and the convolution make a clear improvement.
-One can see clearly that the convolution becomes unimportant at low masses, as the inspiral dominates here.
-
-The $--$ case does not show the same behaviour with inclination as the other ones, as we measure the inclination angle with respect to the final direction of $\bm{J}$, which moves during the inspiral.
-The mismatches are small in this case.
-
-Overall, we find that all the corrections we investigated, perturbative in phase, perturbative in amplitude, and the convolution treatment, are relevant in lowering the unfaithfulness.
-The only exception are second-order amplitude corrections that make little difference, and can even give worse results than first-order in some cases.
+
+As discussed in App.~\ref{app:precpreviousapproaches}, the approximation $\{N:0,A:0\}$ corresponds to the leading-order formula~\eqref{eq:transferlocal}, which is close to the treatment of PhenomP~\cite{Hannam+13}, with the difference that this model uses effective Fourier-domain Euler angles instead of a time-domain modulation through merger and ringdown.
+The order of approximation $\{N:3,A:0\}$ is equivalent to formalism of Ref.~\cite{KCY14} with a stencil order of $N=3$, except that the latter formalism is based on an SPA representation of the P-frame waveform and limited to the inspiral, while our Fourier-domain approach is an attempt at covering the whole frequency band.
+
+Fig.~\ref{fig:precunfaithfulness} shows that, tor the $++$ case, the face-on case shows good agreement even at leading order, as the effects of precession on the full waveform are suppressed for a zero inclination.
+By contrast, with $\pi/3$ or $\pi/2$ inclination the mismatch can reach $10^{-2}$ for high masses above $M \simeq 100 \Msol$, even with the perturbative corrections.
+Using the convolution treatment of the merger and ringdown limits the mismatch, and including both phase (going from blue to red) and amplitude corrections (going from red to yellow and green) makes a clear improvement.
+Similarly, for the case $\perp\perp$, both the perturbative corrections and the convolution at high frequencies reduce the unfaithfulness.
+In both cases, one can see that, as expected, the high-frequency convolution becomes unimportant at low masses, where the inspiral dominates.
+The $--$ case does not show the same dependency on inclination as the other ones, as we measure the inclination angle with respect to the final direction of $\bm{J}$, and this case is close to transitional precession and departs from the picture of simple precession on a cone.
+As expected from the error estimates and the transfer function errors shown in Figs.~\ref{fig:fomprec} and~\ref{fig:precerrors--}, the milder precession velocity means that the mismatches are small in this case already for the leading-order treatment.
+We note, however, that in all cases increasing the order of the amplitude corrections from $A:1$ to $A:2$ in~\eqref{eq:summaryNA} does not yield an improvement, and gives even slightly worse errors in some cases.
+This points to a limitation of our analysis.
+Higher-order time derivatives are harder to extract numerically, and this might be an indication that, even in our smooth model of Sec.~\ref{subsec:precmodel}, we should limit the formalism to first derivatives.
+
+Overall, we found that, although the configurations listed in Table~\ref{tab:precexamples} are strongly spinning ($\chi = 0.95$) with a quite high mass ratio ($q=4$), which both enhance precession effects, the unfaithfulness due to the leading-order Fourier-domain treatment of the precession remains mainly below 1\%, except at larger masses.
+This indicates that this treatment (used in particular in PhenomP~\cite{Hannam+13}, see the discussion of App.~\ref{app:precpreviousapproaches}) should be good enough for current LIGO applications.
+All types of corrections we investigated, either perturbative in phase and amplitude from Sec.~\ref{subsec:executivesummary}, or generated with the convolution treatment of Sec.~\ref{subsec:trigopoly}, consistently lower the unfaithfulness below the 1\% level.
+One exception are second-order amplitude corrections, which we found to make little or no difference.
 Using all the tools at our hands, we are able to keep the unfaithfulness of our three example waveforms below $2.10^{-3}$ for all masses and inclinations.
@@
   \includegraphics[width=.98\linewidth]{plots/precerror_++_py.pdf}
-  \caption{Amplitude of the inertial-frame modes $h^{I}_{2m}$ and errors in the Fourier-domain transfer fonctions $\calT^{2}_{m2}$ for the configuration $++$.
+  \caption{Fourier-domain amplitudes of the modes $\tilde{h}^{\rm I}_{\ell m}$ (top panels), reconstruction errors in the transfer fonctions $\calT^{2}_{m2}$ for the amplitudes (middle panels) and for the phases (bottom panels),  for the configuration $++$ of Table~\ref{tab:precexamples}, with spins almost aligned.
+The thick and thin vertical lines show the merger and asymptotic ringdown frequencies respectively.
 The columns show the orders of approximation $\{\Psi:0,A:0\}$, $\{\Psi:2,A:2\}$, $\{\Psi:2,A:2,\text{conv}\}$.
-\SM{[Compact the columns to share the x axis, add vertical line for merger frequency and shaded area for stitching pert/conv]} \SM{[check: where is the phase of the orange, mode 20 ?]} \SM{[add vertical lines for merger and RD freqs]} \SM{[scaling of the amplitude ?]}}
+In the right column, the region of transition between the perturbative treatment of Sec.~\ref{subsec:executivesummary} and the high-frequency treatment of Sec.~\ref{subsec:trigopoly} is shaded in grey.
+Note that the errors shown here are relative to a given mode, not to the leading mode, and the errors are largest in this sense for the most subdominant modes; the top panels illustrate the mode hierarchy.}
   \label{fig:precerrors++}
@@
   \includegraphics[width=.98\linewidth]{plots/precerror_pp_py.pdf}
-  \caption{Amplitude of the inertial-frame modes $h^{I}_{2m}$ and errors in the Fourier-domain transfer fonctions $\calT^{2}_{m2}$ for the configuration $\perp\perp$.
-The columns show the orders of approximation $\{\Psi:0,A:0\}$, $\{\Psi:2,A:2\}$, $\{\Psi:2,A:2,\text{conv}\}$.
-\SM{[Compact the columns to share the x axis, add vertical line for merger frequency and shaded area for stitching pert/conv]} \SM{[add vertical lines for merger and RD freqs]}}
+  \caption{Same as Fig.~\ref{fig:precerrors++}, but for the configuration $\perp\perp$ of Table~\ref{tab:precexamples}, with both spins in the orbital plane.}
   \label{fig:precerrorspp}
@@
   \includegraphics[width=.98\linewidth]{plots/precerror_--_py.pdf}
-  \caption{Amplitude of the inertial-frame modes $h^{I}_{2m}$ and errors in the Fourier-domain transfer fonctions $\calT^{2}_{m2}$ for the configuration $--$.
-The columns show the orders of approximation $\{\Psi:0,A:0\}$, $\{\Psi:2,A:2\}$, $\{\Psi:2,A:2,\text{conv}\}$.
-\SM{[Compact the columns to share the x axis, add vertical line for merger frequency and shaded area for stitching pert/conv.
-Check the post-merger frame trajectory for - -.]} \SM{[add vertical lines for merger and RD freqs]}}
+  \caption{Same as Fig.~\ref{fig:precerrors++}, but for the configuration $--$ of Table~\ref{tab:precexamples}, with almost anti-aligned spins.}
   \label{fig:precerrors--}
@@
 %\end{figure*}
-
-\SM{[say how the problems with reconstruction notably for ++ are probably caused by the convolution extending to the right and not to the left]}
-
-\SM{[be honest: to get 1\% mismatch we need to combine high spin, high mass and high inclination]}
 
@@
   \caption{Unfaithfulness.
-Computed here with $f_{\rm min} = 20\mathrm{Hz}$, for masses between 20 and 400 $\Msol$, using the standard aLIGO noise curve \SM{[reference for that curve]}.
+Computed here with $f_{\rm min} = 20\mathrm{Hz}$, for masses between 20 and 400 $\Msol$, using the aLIGO ZDHP noise curve~\cite{LIGOProspects13}.
+The inclination is defined here as tha angle between the line of sight and the final $\bm{J}$.
+\SM{[reference for that curve]}.
 \SM{[Extend computation to fmin=10 and Mmin=10]} \SM{[relate unfaithfulness levels to future detectors in the text]} \SM{[check definition of inclination: from $J$ or initial $L$ ?]} \SM{[legend is missing A:2]} \SM{[do we want to show A:2 at all ?]}}

### 2018-06-14 docs/paper_method.tex

@@
 	\tilde{h}_{\rm SPA}(f) &= A_{\rm SPA}(f) e^{-i\Psi_{\rm SPA}(f)} \,, \\
-	A_{\rm SPA}(f) &= a(\tfSPA) \sqrt{\frac{\pi}{\dot{\omega}(\tfSPA)}} \,, \\
+	A_{\rm SPA}(f) &= a(\tfSPA) \sqrt{\frac{\pi}{\dot{\omega}(\tfSPA)}} \,, \label{eq:ASPA} \\
 	\Psi_{\rm SPA}(f) &= 2\varphi(\tfSPA) - 2\pi f \tfSPA + \frac{\pi}{4} \,. \label{eq:PsiSPA}
@@
 Indeed, using~\eqref{eq:deftfSPA} one can verify that the derivative of the SPA phase $\Psi_{\rm SPA}$~\eqref{eq:PsiSPA} with respect to $f$ yields back $\tfSPA$, as
-\be
+\be\label{eq:tfSPA}
 	\tfSPA = -\frac{1}{2\pi} \frac{\ud \Psi_{\rm SPA}}{\ud f} \,.
@@
 For high mass ratios and spins, the correction to $\xi$ in~\eqref{eq:defLambdaxi} starts to become important.
+Here, $\Omega_{\rm prec}$ evolves only on the radiation-reaction timescale through its dependence in $v$; this is a consequence of our simple-precession assumption, as $\Omega_{\rm prec}$ varies on the precession timescale in the generic case (see e.g.~\cite{Chatziioannou+17}).
 
@@
 The dots show the discrete samples entering~\eqref{eq:ykDFTprec}.
-The lower panels show the amplitude and phase residuals of~\eqref{eq:hsymtrigo} compared to the oringial $\tilde{h}(f)$.}
+The lower panels show the amplitude and phase residuals of~\eqref{eq:hsymtrigo} compared to the original $\tilde{h}(f)$.}
   \label{fig:trigopoly}
@@
 For two signals $h_{1}$ and $h_{2}$, the mismatch between them is then defined as \SM{[check]}
-\be
+\be\label{eq:defMM}
 	\mathrm{MM} \equiv \mathrm{Min}_{\delta t, \delta \varphi, \delta \psi} \left( 1 - \frac{ \mathrm{Re} \left[ (h_{1}[\delta t, \delta \Phi, \delta \psi] | h_{2}) \right]}{\sqrt{(h_{1}|h_{1})[\delta\Phi]}\sqrt{(h_{2}|h_{2})}} \right)
@@
   \includegraphics[width=.98\linewidth]{plots/precMM_py.pdf}
-  \caption{Unfaithfulness.
-Computed here with $f_{\rm min} = 20\mathrm{Hz}$, for masses between 20 and 400 $\Msol$, using the aLIGO ZDHP noise curve~\cite{LIGOProspects13}.
-The inclination is defined here as tha angle between the line of sight and the final $\bm{J}$.
-\SM{[reference for that curve]}.
-\SM{[Extend computation to fmin=10 and Mmin=10]} \SM{[relate unfaithfulness levels to future detectors in the text]} \SM{[check definition of inclination: from $J$ or initial $L$ ?]} \SM{[legend is missing A:2]} \SM{[do we want to show A:2 at all ?]}}
+  \caption{Unfaithfulness at various orders of approximation and a numerical FFT, for the three precessing systems of Table~\ref{tab:precexamples}.
+The color indicates the order of approximation of the perturbative treatment of Sec.~\ref{subsec:executivesummary}.
+Dashed lines are obtained applying the treatment of Sec.~\ref{subsec:executivesummary} to the whole frequency band, while full lines are obtained using Sec.~\ref{subsec:trigopoly} for the high frequency range.
+The computation is done using~\eqref{eq:defMM} with $f_{\rm min} = 20\mathrm{Hz}$, for masses between 20 and 400 $\Msol$, and with the aLIGO ZDHP noise curve~\cite{LIGOProspects13}.
+The inclination is defined here as the angle between the line of sight and the final $\bm{J}$.
+\SM{[Extend computation to fmin=10 and Mmin=10 ?]} \SM{[re-check definition of inclination: from $J$ or initial $L$ ?]} }
   \label{fig:precunfaithfulness}
@@
 
-[]
+We presented a formalism to process
+
+\SM{[we should investigate broader parameter space]}
+
+\SM{[move to better performance metrics than just waveform error for LISA]}
+
+\SM{[discuss EMRI]}
 
@@
 \SM{[discuss other detectors, e.g. ET]}
-
-\SM{[we should investigate broader parameter space]}
 
@@
 
+\section{Precessing waveform and the LISA Fourier-domain response}
+\label{app:precLISA}
+
+In this Appendix, we present an argument for the applicability of the treatment of Section~\ref{sec:formalism} to a Fourier-domain waveform that already contains precession effects, thus combining our investigations of Secs.~\ref{sec:LISA} and~\ref{sec:precession}.
+Let us first note that one possible approach would be to incorporate both the precession modulation and the LISA modulation and delay in a single $G(f,t)$ kernel.
+This is what is done notably in~\cite{Klein+15}, with inspiral-only precessing waveforms (and a low-frequency approximation for the response).
+In that approach, comparing the results of Secs.~\ref{sec:LISA} and Sec.~\ref{sec:precession}, it appears that the precession is more challenging to model than the LISA response -- thus the accuracy of the final result should depend mainly on the precession treatment.
+In the following, we take a different route and investigate wether we can directly process Fourier-domain precessing mode contributions of the form $\calT^{\ell}_{m m'}(f) \tilde{h}_{\ell m'}(f)$ through the LISA response.
+We will make simplifying assumptions, assuming simple precession and keeping to order-of-magnitude estimates only.
+We leave for future work a more thorough investigation of precessing waveforms for LISA beyond simple precession and for IMR waveforms.
+
+For simplicity, we assume that the precession transfer function has been computed using the leading-order approximation, as including corrections should not change the timescales involved.
+Thus, we consider the signal
+\be\label{eq:htildeasaprecmodecontrib}
+	\tilde{h} (f) \equiv \calT^{\ell}_{m m'}(f) \tilde{h}_{\ell m'}(f) = \calD^{\ell *}_{m m'}(\alpha, \beta, \gamma)(\tf) \tilde{h}_{\ell m'}(f)
+\ee
+to be processed through the LISA response.
+If we assume simple precession, we can use the results of Sec.~\ref{subsec:sizecorrPrec} for the behaviour of $(\alpha, \beta, \gamma)$ as a function of $t$.
+Restricting ourselves to the inspiral part, we will also use the SPA expressions~\eqref{eq:tfSPA} and~\eqref{eq:TfSPA} for the times $\tf$ and $\Tf$.
+According to~\eqref{eq:wignerphasesimpleprec}, the mode contribution~\eqref{eq:htildeasaprecmodecontrib} above acquires a Fourier-domain phase $\Psi \rightarrow \Psi + \Phi_{\rm prec}$ with
+\be\label{eq:additionalPhiprec}
+	\Phi_{\rm prec} = -(m' \cos \beta(\tf) - m) \alpha(\tf) \,
+\ee
+with $\beta(t)$ varying on the radiation-reaction timescale and $\alpha(t)$ varying on the precession timescale with $\dot{\alpha} = \Omega_{\rm prec}$.
+The Fourier-domain amplitude becomes $A\rightarrow A \times d^{\ell}_{mm'} (\beta(\tf))$.
+If $\beta$ varies on the radiation-reaction timescale, this extra factor can be absorbed in a redefinition of the $a(\tf)$ amplitude factor in~\eqref{eq:ASPA}, so that it won't affect the separation of timescales.
+
+Using the derivatives
+\be
+	\frac{\ud \tf}{\ud f} = 2\pi \Tf^{2}
+\ee
+and
+\be
+	\frac{\ud \Tf}{\ud f} = -2\pi \Tf^{5} \ddot{\omega}(\tf) = - 2\pi \frac{11}{12} \frac{\Tf}{\pi f} \,,
+\ee
+where we have used $\ddot{\omega} = 11/3 (\dot{\omega})^{2}/\omega$ following~\eqref{eq:omegaphiN}, the extra Fourier-domain phase contribution~\eqref{eq:additionalPhiprec} cause new contributions in~\eqref{eq:deftf} and~\eqref{eq:defTf} as \SM{[check]}
+\begin{subequations}
+\begin{align}
+	-\frac{1}{2\pi} \frac{\ud \Phi_{\rm prec}}{\ud f} &\sim (m' \cos \beta - m) \Tf^{2} \Omega_{\rm prec} \,, \label{eq:additionaltf}\\
+	-\frac{1}{4\pi^{2}} \frac{\ud^{2} \Phi_{\rm prec}}{\ud f^{2}} &\sim (m' \cos \beta - m) \Tf^{2} \left[ \frac{11}{12} \frac{\Omega_{\rm prec}}{\pi f} - \Tf^{2} \dot{\Omega}_{\rm prec} \right] \,, \label{eq:additionalTf}
+\end{align}
+\end{subequations}
+where we ignored $\dot{\beta}$. The new term~\eqref{eq:additionaltf} means that, when processing the signal through the LISA response, we would attribute a different $\tf$ to each mode contribution.
+
+The new term~\eqref{eq:additionalTf} will enter $\epsilon_{\Psi 2}$ as defined in~\eqref{eq:deffom}.
+Comparing to $\Tf^{2}$, and ignoring the prefactor depending on the mode number (of order a few), we have two terms in the bracket.
+The first term is $\Omega_{\rm prec} / \omega$, which is a quantity of order 1PN, and should remain well below 1.
+The second term is $\dot{\Omega}_{\rm prec} / 2\dot{\omega}$.
+Under the simple precession assumption, \eqref{eq:Omegaprec} gives $\dot{\Omega}_{\rm prec} \sim \dot{\omega}/\omega \Omega_{\rm prec}$, and this term suppressed like the first term.
+Thus, in this case the contribution to $\epsilon_{\Psi 2}$ should be subdominant, the findings of Sec.~\ref{sec:LISA} for non-spinning waveforms should apply.
+
+In the generic case of double-spin precession, however, the above argument does not apply anymore as $\Omega_{\rm prec}$ varies itself on the precessional timescale and $\Omega_{\rm prec} / \dot{\omega}$ is not small in general.
+In fact, the cases where $\dot{\Omega}_{\rm prec} \sim \dot{\omega}$, with possible cancellations, are the ``catastrophes'' preventing the application of the SPA to the full time-domain precessing signal~\cite{KCY13}.
+We leave to future work the question of wether this situation could be challenging for the perturbative formalism, after taking into account the appropriate factors for the error estimates as in Sec.~\ref{subsec:lisafom}.
+
+Note that, if we were instead to incorporate the full precession transfer function $\calT^{\ell}_{mm'}$ as an envelope function multiplying $A(f)$, then applying the criterion~\eqref{} would show that, even in the case of simple precession, the perturbative analysis would be challenged at both ends of the mass spectrum, for low-mass and high-mass systems.
+
+\SM{[say: to update with double precession, in particular $\ddot{\Phi}_{\rm prec}$]}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-06-14 docs/paper_method.tex
Edits to the paper.

@@
 By contrast, for the constellation response, to our knowledge the coefficients $c_{n}^{L}$ do not admit such a simple close-form expression, and they must be computed numerically using the integrand given by~\eqref{eq:GslrL}.
-Truncating~\eqref{eq:Gdiscretefourier} to a finite order $N$, this computation reduces to a DFT \SM{[acronym]}, and the $c_{n}$ coefficients are given by~\eqref{eq:ykDFT} and~\eqref{eq:ckyk}.
+Truncating~\eqref{eq:Gdiscretefourier} to a finite order $N$, this computation reduces to an FFT and the $c_{n}$ coefficients are given by~\eqref{eq:ykDFT} and~\eqref{eq:ckyk}.
 
@@
 The truncation order $N(f,\eta)$ is frequency-dependent, and also depends on the orientation angles.
-For $c_{n}^{L}$, it also determines the number of function evaluations and the cost of the DFT.
+For $c_{n}^{L}$, it also determines the number of function evaluations and the cost of the FFT.
 For $c_{n}^{0}$, the expression~\eqref{eq:cn0} can provide an asymptotic bound on the width of the comb.
@@
 The merger and ringdown part of the Fourier-domain signal extends over a wide range of frequencies, while it corresponds to a short interval of times, much shorter than the inspiral phase.
-One approach would be to separate the waveform between the inspiral and merger phase, build a Fourier-domain model for the inspiral part only, to which the formalism of Sec.~\ref{sec:formalism} could be applied, while the merger and ringdown part could be handled by a direct DFT \SM{[acronym]}, which would be a cheap operation on a limited range in time.
-Although this approach should be the most robust, here we will use an alternative, simpler to implement and keeping the setting of a Fourier-domain precessing-frame waveform combined with a time-domain precession modulation.
+One approach would be to separate the time-domain waveform between the inspiral and merger phase, build a Fourier-domain model for the inspiral part only, to which the formalism of Sec.~\ref{sec:formalism} could be applied, while the merger and ringdown part could be handled by a direct FFT, which would be a cheap operation on a limited range in time.
+Although this approach should be straightforward, here we will use an alternative method, simpler to implement as it allows us to keep the setting of a Fourier-domain precessing-frame waveform combined with a time-domain precession modulation.
 
@@
 Taking advantage of this, we will adopt a one-sided trigonometric polynomial representation for $\tilde{h}(f)$.
-For frequencies high enough that the support of the convolution~\eqref{eq:precconvolution} does not extend beyond the range covered by this trigonometric representation, the result will be obtained directly as a DFT \SM{[acronym]} with a limited number of samples.
-The limitation to the high-frequency range is crucial here: the Fourier-domain amplitude and phase diverge as $f^{-7/6}$ and $f^{-5/3}$ respectively for $f \rightarrow 0$, and our procedure would require a much finer sampling to cover part of the inspiral, going back to being equivalent to an inverse DFT of the full signal if we were to cover all frequencies.
+For frequencies high enough that the support of the convolution~\eqref{eq:precconvolution} does not extend beyond the range covered by this trigonometric representation, the result will be obtained directly as an FFT/IFFT with a limited number of samples.
+The limitation to the high-frequency range is crucial here: the Fourier-domain amplitude and phase diverge as $f^{-7/6}$ and $f^{-5/3}$ respectively for $f \rightarrow 0$, and our procedure would require a much finer sampling to cover part of the inspiral, going back to being equivalent to an IFFT of the full signal if we were to cover all frequencies.
 
@@
 
-Next, we build a a trigonometric polynomial representation of $\tilde{h}_{\rm sym}$, a construction intimately related to the DFT, that we recall in App.~\ref{app:notation}.
+Next, we build a a trigonometric polynomial representation of $\tilde{h}_{\rm sym}$, a construction intimately related to the FFT, that we recall in App.~\ref{app:notation}.
 Over the frequency range $f\in [f_{0}, 2f_{\rm max} - f_{0}]$, we can approximate
@@
 \ee
-where the factor $(-1)^{k}$ comes from the fact that $f_{0}$ is here at the center of the interval.
-The coefficients $c_{k}$ are built following the rules~\eqref{eq:ckyk} from the inverse DFT coefficients
+where the factor $(-1)^{k}$ comes from the fact that $f_{0}$ is here at the center of the interval. The coefficients $c_{k}$ are built following the rules~\eqref{eq:ckyk} from the IFFT coefficients
 \be\label{eq:ykDFTprec}
@@
 where we defined the time sampling $\delta t \equiv 1/\Delta f$.
-We see that we are left with a DFT-like expression to compute from $N+1$ time samples of the modulation function $F$, centered around $t_{p}$.
-We see that, apart from the conditioning described above with the artificial symmetrization, this is analogous to a DFT of the product of the modulation with the time-domain signal obtained through an inverse DFT.
+We see that we are left with an FFT-like expression to compute from $N+1$ time samples of the modulation function $F$, centered around $t_{p}$.
+We see that, apart from the conditioning described above with the artificial symmetrization, this is analogous to an FFT of the product of the modulation with the time-domain signal obtained through an IFFT.
 
@@
 In the following, we will use $M=64$ samples ($32$ useful samples before 0-padding, $N=128$ samples in total for the artificially symmetrized signal).
-The computation of~\eqref{eq:ykDFTprec} and~\eqref{eq:resultdirectconvol} amounts to two DFT operations, implemented through an FFT/IFFT, and is done only once for a given waveform and modulation function.
+The computation of~\eqref{eq:ykDFTprec} and~\eqref{eq:resultdirectconvol} amounts to two FFT/IFFT operations, and is done only once for a given waveform and modulation function.
 The number of samples is of the same order of magnitude as the one required to represent the Fourier-domain waveform with an interpolating cubic spline for its amplitude and phase (a few hundreds for the full frequency band, see e.g~\cite{Puerrer14}), thus the cost should be comparable to the Taylor-like approach presented in Sec.~\ref{subsec:executivesummary}.
@@
 
-We presented a formalism to process
-
-\SM{[we should investigate broader parameter space]}
-
-\SM{[move to better performance metrics than just waveform error for LISA]}
-
-\SM{[discuss EMRI]}
-
-\SM{[discuss the extension of our LISA treatment for precessing signals]}
-
-\SM{[discuss other detectors, e.g. ET]}
+We presented a formalism to process gravitational-wave signals described in the Fourier domain through a time-domain modulation and delay, of the type created by the motion and response of a LISA-type detector and by the precessing motion of spinning compact binaries.
+The natural leading-order of this perturbative formalism simply amounts to building a time-to-frequency correspondence~\eqref{eq:deftf}, generalizing the SPA correspondence~\eqref{eq:deftfSPA} through merger and ringdown.
+
+This formalism is based on a the separation of timescales between the modulation and the signal.
+By a Fourier-domain expansion of the integrand of convolution integrals, corrections beyond the leading order are straightforwardly generated.
+The most important are quadratic phase corrections, for which the formalism yields a natural extension of the results of~\cite{KCY14} to IMR signals, thanks to the generalization of the radiation reaction timescale with a Fourier-domain expression~\eqref{eq:defTf}.
+Other corrections include amplitude and delay terms.
+We provided order-of-magnitude error estimates~\eqref{eq:deffom} allowing a simple diagnosis of the relevance of these various corrections, forecasting where the perturbative formalism will fail due to an insufficient separation of timescales.
+
+We applied our approach to the case of non-precessing comparable-mass binary systems seen by LISA, investigating examples of supermassive black hole mergers as well as inspirals of stellar-mass black holes.
+The population of stellar-mass binary black hole signals seen in LISA will include chirping systems, merging within years of the LISA observations, as well as slowly-chirping systems hundreds or thousands of years away from merger.
+We found that our treatment is well applicable for both high-mass and chirping low-mass systems.
+The leading-order treatment performs quite well, with at most a localized $\sim10\%$ in amplitude or phase error, while the corrections we derived can reduce the errors in the transfer functions to much lower levels if needed.
+Quadratic phase corrections at the start of the waveform and delay corrections at the high end of the frequency band are the most relevant.
+On the other hand, the morphology of the slowest-chirping stellar-mass signals is close to quasi-monochromatic galactic binaries, and they cannot be handled by this perturbative formalism.
+We developed a separate method, using discrete Fourier coefficients, to compute the instrument response in this case.
+
+To explore the application of the formalism to precessing binaries, we built a toy model for the precession of a two-spins system including a post-merger precessing-frame trajectory.
+For the high spin, moderate mass ratio examples we explored, we found that subdominant precessing mode contributions can be challenging for the formalism and have large relative errors, while they are also suppressed in amplitude in the complete waveform.
+For the inspiral, the most relevant corrections are the quadratic in phase ones that were already derived in~\cite{KCY14}.
+Our formalism extends to the merger and ringdown, where our model for the post-merger precession gives a weaker separation of timescales.
+We developed an alternative method to handle the high-frequency, late-time part of the signal, with a direct convolution of the Fourier-domain signal.
+We showed that although the leading-order treatment is acceptable for our examples with an unfaithfulness at or below 1\% for typical LIGO masses, including in the inspiral all the corrections we derived, as well as using the convolution treatment for the merger, both help lowering the unfaithfulness down to the $10^{-4}$-$10^{-3}$ range.
+
+Our investigation of the precession provides tools that will be potentially useful for the next generations of Fourier-domain waveform models for IMR precessing systems.
+Our treatment of the Fourier-domain LISA instrumental response will allow us to leverage modern, fast Fourier-domain codes producing IMR waveforms to accelerate computationally intensive data analysis applications, most notably Bayesian source parameter estimation on synthetic data, that have become urgent with the selection of LISA as a mission by ESA.
+The LISA response described in Sec.~\ref{sec:LISA} has been implemented in the LISA Data Challenge~\cite{} software suite.
+\SM{[cite LDC, how ?]}
+
+Our results call for straightforward extensions.
+First, we only considered the leading harmonic of non-spinning waveforms in the LISA response.
+Higher harmonics and aligned spin components are not expected to matter, but precession will leave a definite imprint on the Fourier-domain signal.
+One option in this case would be to handle the precession and response modulation togeter, as in~\cite{Klein+15}.
+However, we present in App.~\ref{app:precLISA} an argument showing that our formalism should apply straightforwardly to waveforms with simple precession.
+We leave the general case for future study.
+
+A natural extension of our formalism for the LISA response would be the third-generation ground-based detectors, like for instance the Einstein Telescope.
+For such detectors, signals are sufficiently long for the rotation of the Earth to matter, and their arms are long enough that the long wavelength approximation is not valid at high frequencies and that the response becomes frequency-dependent~\cite{Essick+17}.
+A preliminary order-of-magnitude analysis shows that the leading-order treatment should be applicable \SM{[check]}, but this potential application is to be studied in more details.
+
+Finally, we restricted ourselves to comparable-mass systems, while Extreme Mass Ratio Inspirals (EMRI) will be very interesting signals for LISA.
+Those signals will show typically strong precession and also a large eccentricity, so that we would need to apply our treatment separately to each of the numerous individual harmonics produced (see e.g.~\cite{Chua+17}).
 
@@
 
-In our convention, the link between the Discrete Fourier Transform (DFT) \SM{[check acronym]} and the trigonometric polynomial representation of a function goes as follows.
+Throughout this paper we will refer to Discrete Fourier Transform (DFT) with the acronyms FFT/IFFT for the Fast Fourier Transform and its inverse.
+In our convention~\eqref{eq:defFT}, the link between the DFT and the trigonometric polynomial representation of a function goes as follows.
 For a periodic function $F(x)$ defined on $x\in [x_{0}, x_{0} + \Delta x]$, and represented by $N$ samples $x_{j} = \ov{x} + j \Delta x/N$, $j=0,\dots,N-1$, with $N$ large enough to satisfy, at least approximately, the Nyquist criterion, we can build a trigonometric interpolant $P(x)$ as
@@
 This trigonometric polynomial representation is equivalent to truncating the formal Fourier series, representing the full signal, to a finite order $M$.
-The coefficients $c_{k}$, in the full series, are defined as integrals over the interval of $F$ weighted by the appropriate exponential (as in~\eqref{eq:defcn}).
-In both interpretations, either the truncated approximation of the Fourier series or  the trigonometric interpolation formulation, these coefficients are related to the coefficients of the inverse DFT.
+The coefficients $c_{k}$, in the full series, are defined as
+\be\label{eq:defcnApp}
+	c_{n}(F) = \frac{1}{\Delta x} \int_{0}^{\Delta x} \ud x \, e^{\frac{2 i \pi n x }{ \Delta x}} F(x) \,.
+\ee
+%integrals over the interval of $F$ weighted by the appropriate exponential (as in~\eqref{eq:defcn}).
+In both interpretations, either the truncated approximation of the Fourier series or  the trigonometric interpolation formulation, these coefficients are related to the coefficients of the IFFT.
 If we set $\omega \equiv e^{2i\pi/N}$ and define \SM{[check convention]}
@@
 \ee
-which is the expression of the inverse DFT in our sign convention~\eqref{eq:defFT}, the coefficients $c_{k}$ are given by
+which is the expression of the IFFT in our sign convention~\eqref{eq:defFT}, the coefficients $c_{k}$ are given by
 \begin{align}\label{eq:ckyk}
@@
 
-Note that, if we were instead to incorporate the full precession transfer function $\calT^{\ell}_{mm'}$ as an envelope function multiplying $A(f)$, then applying the criterion~\eqref{} would show that, even in the case of simple precession, the perturbative analysis would be challenged at both ends of the mass spectrum, for low-mass and high-mass systems.
-
-\SM{[say: to update with double precession, in particular $\ddot{\Phi}_{\rm prec}$]}
+Note that, if we were instead to incorporate the full precession transfer function $\calT^{\ell}_{mm'}$ as an envelope function multiplying $A(f)$, then applying the criterion~\eqref{eq:deffom} for $\epsilon_{A1}$ would show that, even in the case of simple precession, the perturbative analysis would be challenged at both ends of the mass spectrum, for low-mass and high-mass systems.
 

### 2018-06-15 docs/paper_method.tex
Some mostly minor edits in sec V.  Search for "CHECK".

--- 2018-06-14 docs/paper_method.tex
@@
 \ee
-In the notations that we used in Sec.~\ref{sec:formalism}, in the absence delays, we wish to compute the convolutions
+In the notation that we used in Sec.~\ref{sec:formalism}, in the absence delays, we wish to compute the convolutions
 \be\label{eq:precconvolution}
@@
 
-We will use the following notations: for $m_{1}$, $m_{2}$ the masses of the two bodies, we set $M=m_{1}+m_{2}$, $\nu=m_{1}m_{2}/M^{2}$, $\delta = (m_{1}-m_{2})/M$, and $|\bm{S}_{A}|=Gm_{A}^{2} \chi_{A}$ for $A=1,2$, with $\chi$ the dimensionless spin between $0$ and $1$ for Kerr black holes.
+We will use the following notation: for $m_{1}$, $m_{2}$ the masses of the two bodies, we set $M=m_{1}+m_{2}$, $\nu=m_{1}m_{2}/M^{2}$, $\delta = (m_{1}-m_{2})/M$, and $|\bm{S}_{A}|=Gm_{A}^{2} \chi_{A}$ for $A=1,2$, with $\chi$ the dimensionless spin between $0$ and $1$ for Kerr black holes.
 We define $\bm{\ell}$ as the unit vector normal to the orbital plane.
@@
 The overall $\nu$ scaling indicates a better separation of timescales for larger mass ratio systems.
-These expressions also show that, since the total angular momentum appears as a factor in $\Omega_{\rm prec}$, higher-order corrections will be larger for spin-aligned systems than for anti-aligned spins, an effect that becomes significant for large mass ratios.
+These expressions also show that, since the total angular momentum appears as a factor in $\Omega_{\rm prec}$, higher-order corrections will be larger for \jgb{roughly} spin-aligned systems than for anti-aligned spins, an effect that becomes significant for large mass ratios.
 Note that the regime where $\xi$ gets close to 0 corresponds to the transitional precession range~\cite{Apostolatos+94}, with the spin of the primary compensating the orbital angular momentum, and our analysis based on simple precession is not valid anymore.
@@
 
-Instead of tapering the signal to $0$ at $f_{0}$, to limit the deviation from the original signal we take advantage of the one-sidedness and only flatten the amplitude\footnote{In practice, this is implemented as the discrete integral of a cosine window function on just the first two samples.
+Instead of tapering the signal to $0$ \jgb{below [CHECK]} $f_{0}$, to limit the deviation from the original signal we take advantage of the one-sidedness and only flatten the amplitude\footnote{In practice, this is implemented as the discrete integral of a cosine window function on just the first two samples.
 We also taper to 0 the last three samples before $f_{\rm max}$, and $0$-pad by a factor of 2.}, and artificially symmetrize the signal.
@@
 The thick and thin vertical lines show the merger and asymptotic ringdown frequencies respectively.
+\jgb{For the solid curves in the top panel, precession modulations were computed in the time domain, while the dashed curves show the results of the Fourier-domain reconstruction techniques presented here.[CHECK]}
 The columns show the orders of approximation $\{\Psi:0,A:0\}$, $\{\Psi:2,A:2\}$, $\{\Psi:2,A:2,\text{conv}\}$.
@@
   \includegraphics[width=.98\linewidth]{plots/precMM_py.pdf}
-  \caption{Unfaithfulness at various orders of approximation and a numerical FFT, for the three precessing systems of Table~\ref{tab:precexamples}.
+  \caption{Unfaithfulness at various orders of approximation \jgb{as compared with a time-domain treatment using a full-domain[CHECK]} a numerical FFT, for the three precessing systems of Table~\ref{tab:precexamples}.
 The color indicates the order of approximation of the perturbative treatment of Sec.~\ref{subsec:executivesummary}.

### 2018-06-15 docs/paper_method.tex
Edits in the summary.

@@
 
-We presented a formalism to process gravitational-wave signals described in the Fourier domain through a time-domain modulation and delay, of the type created by the motion and response of a LISA-type detector and by the precessing motion of spinning compact binaries.
+We presented a formalism to process gravitational-wave signals described in the Fourier domain through a time-domain modulation and delay of the type created by the motion and response of a LISA-type detector and by the precessing motion of spinning compact binaries.
 The natural leading-order of this perturbative formalism simply amounts to building a time-to-frequency correspondence~\eqref{eq:deftf}, generalizing the SPA correspondence~\eqref{eq:deftfSPA} through merger and ringdown.
@@
 To explore the application of the formalism to precessing binaries, we built a toy model for the precession of a two-spins system including a post-merger precessing-frame trajectory.
-For the high spin, moderate mass ratio examples we explored, we found that subdominant precessing mode contributions can be challenging for the formalism and have large relative errors, while they are also suppressed in amplitude in the complete waveform.
+For the high spin, moderate mass ratio examples we explored, we found that subdominant precessing mode contributions can be challenging for the formalism and have large relative errors, \jgb{although they are suppressed} in amplitude in the complete waveform.
 For the inspiral, the most relevant corrections are the quadratic in phase ones that were already derived in~\cite{KCY14}.
@@
 
-Our results call for straightforward extensions.
+Our results \jgb{suggest a few} straightforward extensions.
 First, we only considered the leading harmonic of non-spinning waveforms in the LISA response.
-Higher harmonics and aligned spin components are not expected to matter, but precession will leave a definite imprint on the Fourier-domain signal.
-One option in this case would be to handle the precession and response modulation togeter, as in~\cite{Klein+15}.
+\jgb{The same approach is also well-suited to treating higher harmonics and aligned spin signal}, but precession will leave a definite imprint on the Fourier-domain signal.
+One option in this case would be to handle the precession and response modulation together, as in~\cite{Klein+15}.
 However, we present in App.~\ref{app:precLISA} an argument showing that our formalism should apply straightforwardly to waveforms with simple precession.
@@
 
-A natural extension of our formalism for the LISA response would be the third-generation ground-based detectors, like for instance the Einstein Telescope.
-For such detectors, signals are sufficiently long for the rotation of the Earth to matter, and their arms are long enough that the long wavelength approximation is not valid at high frequencies and that the response becomes frequency-dependent~\cite{Essick+17}.
-A preliminary order-of-magnitude analysis shows that the leading-order treatment should be applicable \SM{[check]}, but this potential application is to be studied in more details.
-
-Finally, we restricted ourselves to comparable-mass systems, while Extreme Mass Ratio Inspirals (EMRI) will be very interesting signals for LISA.
-Those signals will show typically strong precession and also a large eccentricity, so that we would need to apply our treatment separately to each of the numerous individual harmonics produced (see e.g.~\cite{Chua+17}).
+\jgb{Another natural application of our formalism developed for the LISA response is to third-generation ground-based detectors.
+For detectors like the Einstein Telescope,} signals are sufficiently long for the rotation of the Earth to matter, and their arms are long enough that the long wavelength approximation is not valid at high frequencies and that the response becomes frequency-dependent~\cite{Essick+17}.
+A preliminary order-of-magnitude analysis shows that the leading-order treatment should be applicable \SM{[check]}, \jgb{but more investigation is needed.}
+
+\jgb{Finally, while we restricted ourselves to comparable-mass binary systems, Extreme Mass Ratio Inspirals (EMRI) will also be very important signals for LISA.
+Those signals generally include not only strong precession but also large eccentricity.
+An effective frequency-domain LISA response may be possible by apply our treatment separately to each of the numerous individual harmonics produced (see e.g.~\cite{Chua+17}).}
 
@@
 
-[Acknowledgments]
+\jgb{This work was supported by a NASA Astrophysics Theory Program grant 11-ATP-046.}
 

### 2018-06-15 docs/paper_method.tex
Edits to the paper.

--- 2018-06-15 docs/paper_method.tex
@@
 
+\SM{[check citation of Cornish-Rubbo-Poujade early work]}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \includegraphics[width=.98\linewidth]{plots/tf_py.pdf}
-  \caption{Time-to-frequency correspondence $t_{f}$, as defined directly from the phase of the Fourier-domain signal as in~\eqref{eq:deftf}, in geometric units.
+  \caption{Time-to-frequency correspondence $t_{f}$, as defined directly from the phase of the Fourier-domain signal in~\eqref{eq:deftf}, in geometric units.
 We show only the high-frequency part of the signal, corresponding to the merger region.
-The PhenomD waveforms have been aligned such that the time-domain amplitude (obtained by an IFFT \SM{[check acronym]}) peaks at $t=0$.
+The \texttt{PhenomD} waveforms have been aligned such that the time-domain amplitude (obtained by an IFFT) peaks at $t=0$.
 The two aligned spins are equal, with components of $0.95$ ($++$, blue), $0$ ($00$, red), and $-0.95$ ($--$, yellow), for mass ratios $q=1$ (full line) and $q=8$ (dashed).
-The vertical lines shows the time-domain instantaneous frequency at the peak $\omega^{22}_{\rm peak}/(2\pi)$, again measured from an IFFT, and the fact that the curves $t_{f}$ do not pass exactly by the crossing with the $t=0$ horizontal line reflects the fact that the link between time-domain frequency and Fourier frequency $f$ is only approximate at merger.
-Note that $t_{f}$ increases to at most $\sim 30M$ after the peak, and is not monotonous.
-\SM{[add color to the vertical lines]} \SM{[add dots to highlights the crossings]}}
+The vertical lines shows the time-domain instantaneous frequency at the peak $\omega^{22}_{\rm peak}/(2\pi)$, and the fact that the curves $t_{f}$ do not pass exactly by the crossing with the $t=0$ horizontal line reflects the fact that the link between time-domain frequency and Fourier frequency $f$ is only approximate at merger.
+Note that $t_{f}$ increases to at most $\sim 30M$ after the peak, and is not monotonous.}
   \label{fig:tf}
@@
 \end{align}
-\jgb{which we can think of as a local evaluation of the kernel function $G(f,t)$ at a frequency-dependent effective time}
+which we can think of as a local evaluation of the kernel function $G(f,t)$ at a frequency-dependent effective time
 \be\label{eq:deftf}
@@
 In particular, one should note that $\tf$ is not monotonically increasing with frequency anymore after reaching in the high-frequency part of the waveform, corresponding to the ringdown.
-\jgb{As long as the Fourier phase is differentiable $t_{f}$ is a well-defined function of $f$.
-While its non-monotonicity would forbid an unambiguous definition of a reciprocal frequency-of-time function $f(t)$, like that in the SPA, no such function will be needed in our treatment.}
-
-\jgb{With \eqref{eq:leadingorderresponse} we have brought the modulated and delayed signal in to the form~\eqref{eq:deftransfer} with transfer function}
+As long as the Fourier phase is differentiable $t_{f}$ is a well-defined function of $f$.
+While its non-monotonicity would forbid an unambiguous definition of a reciprocal frequency-of-time function $f(t)$, like that in the SPA, no such function will be needed in our treatment.
+
+With \eqref{eq:leadingorderresponse} we have brought the modulated and delayed signal in to the form~\eqref{eq:deftransfer} with transfer function
 \be\label{eq:transferlocal}
@@
 
-\jgb{If the effective width of of the kernel function $G(f,f’)$ is not quite negligible compared to the scale of significant variations with $f$ it can be useful to extend our approach beyond the leading-order approximation.
+\jgb{If the effective width of the kernel function $G(f,f’)$ is not quite negligible compared to the scale of significant variations with $f$ it can be useful to extend our approach beyond the leading-order approximation.
 With the waveform represented in phase-and-amplitude form~\eqref{eq:A9}, the elements of~\eqref{eq:FDkernel} may be Taylor expanded in the variable $f'$: \SM{[change notation, do not use $N$ as sum boundary for $\Psi$]}
@@
 The time-domain frequency at merger is represented by the vertical line.
-The derivatives from which these timescales are built encounter zero-crossings and change sign in the high-frequency range.
-\SM{[discuss discontinuity of third derivative]}\SM{[add vertical line for RD, make merger thick]}}
+The derivatives from which these timescales are built encounter zero-crossings and change sign in the high-frequency range.}
   \label{fig:TfTA}
@@
 \ee
-\jgb{This expression provides a summary of the perturbative treatment we will apply in later sections, with the additional understanding that the Frensel transform is treated approximately using \eqref{eq:fresnelstencil}.[Check this]}
+\jgb{This expression provides a summary of the perturbative treatment we will apply in later sections, with the additional understanding that the Fresnel transform is treated approximately using \eqref{eq:fresnelstencil}.[Check this]}
 \end{widetext}
@@
 
-For the constellation response~\eqref{eq:defresponseL}, we can \jgb{separately treat the orbital delay $d_r$ and the constellation delay at spacecraft $s$, $d_{s,L}$}, by writing
+For the constellation part of response~\eqref{eq:defresponseL}, treated separately from the delay~\eqref{eq:delay0}, we write
 \begin{align}\label{eq:decomposeGslr}
@@
 \end{align}
-\jgb{Since we also assume the {\bf[right]} } rigid approximation for the constellation, where the armlengths are fixed, a particular simplification occurs when combining these individual delays, thanks to the relation $p^{L}_{r} - p^{L}_{s} =  L n_{l}$:
+The superscript $L$ indicates that the orbital delay~\eqref{eq:delay0} is not included.
+Since we also assume the rigid approximation for the constellation, where the armlengths are fixed, a particular simplification occurs when combining these individual delays, thanks to the relation $p^{L}_{r} - p^{L}_{s} =  L n_{l}$:
 \begin{align}\label{eq:GslrL}
@@
 This leading-order result~\eqref{eq:transfer0local}-\eqref{eq:transferLlocal} is familiar, as it was derived in the past taking the point of view of monochromatic signals (see e.g.~\cite{} \SM{[C]}) \SM{[rewrite that]}.
+For plotting purposes, we will define
+\be\label{eq:transferLenvelope}
+	\overline{\calT}_{slr}^{L} (f) = \frac{i \pi f L}{2} n_{l} \cdot P_{\ell m} \cdot n_{l} (\tf)
+\ee
+which will serve as an estimate for the enveloppe function of the response, devoid of the zero-crossings at high frequencies of the $\sinc$ term in~\eqref{eq:GslrL}.
 
@@
   \includegraphics[width=.9\linewidth]{plots/lisafom_Psi_McDeltat.pdf}
-  \caption{\SM{[To be checked and redone]} Contour levels for the analytical estimate of the error measure $\epsilon_{\Psi}$ at the starting frequency, as a function of chirp mass $\Mchirp$ and time to coalescence $\Delta t$.
+  \caption{\SM{[To be checked and redone in matplotlib]} \SM{[remove grey shading]} \SM{[add labels to contours]} Contour levels for the analytical estimate of the error measure $\epsilon_{\Psi 2}$ at the starting frequency, as a function of chirp mass $\Mchirp$ and time to coalescence $\Delta t$.
 Blue corresponds to the orbital response and red to the constellation response.
-The colored shaded areas indicate regions where $\epsilon_{\Psi} \geq 1$, where the perturbative formalism is expected to break down.
-In the gray area, $f_{\rm start}$ given in~\eqref{eq:fstartN} is lower than the lowest in-band frequency $f_{\rm min} = 10^{-5}\Hz$, so that the signal starts at $f_{\rm min}$ and $\epsilon_{\Psi}$ becomes independent of $\Delta t$.\jgb{\bf[Maybe not shade the ``grey area'' since it looks excluded, instead can call it ``the region to the right of the black line.'']} }
+The colored shaded areas indicate regions where $\epsilon_{\Psi 2} \geq 1$, where the perturbative formalism is expected to break down.
+In the region to the right of the black line, $f_{\rm start}$ given in~\eqref{eq:fstartN} is lower than the lowest in-band frequency $f_{\rm min} = 10^{-5}\Hz$, so that the signal starts at $f_{\rm min}$ and $\epsilon_{\Psi 2}$ becomes independent of $\Delta t$.}
   \label{fig:lisafomPsiMcDeltat}
@@
 The top row corresponds to the orbital delay part of the response~\eqref{eq:defresponse0}, and the lower row shows the LISA-centered constellation response~\eqref{eq:defresponseL}.
-The error measures $\epsilon_{\Psi}$ for the phase corrections, $\epsilon_{A1}$ for the amplitude, $\epsilon_{d}$ for the delay are shown from left to right.
+The error measures $\epsilon_{\Psi 2}$ for the phase corrections, $\epsilon_{A1}$ for the amplitude, $\epsilon_{d}$ for the delay are shown from left to right.
 The central line and interval are the mean and $1\sigma$ standard deviation of the logarithm of $\epsilon$ computed with numerical derivatives over 400 random values for the position in the sky, inclination and polarization.
@@
 The error measure $\epsilon_{d}$ does not depend on signal-dependent timescales and can be directly read off the estimates above, with $\epsilon_{d} = |\partial_{tf}G/(2\pi G)|$.
-For the error measures $\epsilon_{\Psi}$ and $\epsilon_{A1}$, we can combine the above derivatives with the Newtonian timescales given in~\eqref{eq:timescalesN} for the inspiral phase of the signal.
+For the error measures $\epsilon_{\Psi 2}$ and $\epsilon_{A1}$, we can combine the above derivatives with the Newtonian timescales given in~\eqref{eq:timescalesN} for the inspiral phase of the signal.
 
@@
 We turn first to the slowly-chirping binaries.
-For these systems, we will focus on $\epsilon_{\Psi}$, which will be the most important error measure.
-Fig.~\ref{fig:lisafomPsiMcDeltat} shows contour levels for the value of our simple analytical estimate for $\epsilon_{\Psi}$ at the beginning of the observations, as a function of the chirp mass and time to merger, for both the orbital and the constellation response.
-The limit $\epsilon_{\Psi} = 1$ is used to single out areas where the perturbative treatment of Sec.~\ref{sec:formalism} is expected to break down, although the precise location of this boundary will vary, depending on the accuracy level required and on the orientation angles.
+For these systems, we will focus on $\epsilon_{\Psi 2}$, which will be the most important error measure.
+Fig.~\ref{fig:lisafomPsiMcDeltat} shows contour levels for the value of our simple analytical estimate for $\epsilon_{\Psi 2}$ at the beginning of the observations, as a function of the chirp mass and time to merger, for both the orbital and the constellation response.
+The limit $\epsilon_{\Psi 2} = 1$ is used to single out areas where the perturbative treatment of Sec.~\ref{sec:formalism} is expected to break down, although the precise location of this boundary will vary, depending on the accuracy level required and on the orientation angles.
 We also single out the high-mass region where the start of the signal is set by its entry \jgb{into the sensitive band, at $f_{\rm min}$ making the error measure independent of time to merger}.
-We find that, for binaries in the LIGO/Virgo mass range $10^{1}-10^{2} \Msol$, we reach the limit $\epsilon_{\Psi}=1$ first for the orbital response, for time-to-merger values within the expected observed range~\cite{Sesana16}.
+We find that, for binaries in the LIGO/Virgo mass range $10^{1}-10^{2} \Msol$, we reach the limit $\epsilon_{\Psi 2}=1$ first for the orbital response, for time-to-merger values within the expected observed range~\cite{Sesana16}.
 We will introduce in Sec~\ref{subsec:comblisa} an alternative approach to deal with those signals.
@@
 We now turn to the case of merging binaries.
-For these we compute the three error measures $\epsilon_{\Psi}$, $\epsilon_{A1}$ and $\epsilon_{d}$.
+For these we compute the three error measures $\epsilon_{\Psi 2}$, $\epsilon_{A1}$ and $\epsilon_{d}$.
 To go beyond the crude analytical estimates built from~\eqref{eq:timescalesNfstart} and~\eqref{eq:estimatederivorb}-\eqref{eq:estimatederivconst}, we use numerical derivatives for the timescales~\eqref{eq:defTf}-\eqref{eq:defTA} and for the derivatives in~\eqref{eq:deffom}.
@@
 
-For $\epsilon_{\Psi}$, as expected from~\eqref{eq:TfN} we find a steep rise towards lower frequencies, which shows that considering merging binaries with a limited time to coalescence is crucial here.
-\jgb{We note differences in} behaviour between the orbital response, where for a fixed mass ratio the initial \jgb{(10yr before merger)} value of $\epsilon_{\Psi}$ grows towards lower masses, and the constellation response, where \jgb{initial} $\epsilon_{\Psi}$ grows towards higher masses.
-\jgb{Combining both parts of the response, for equal-mass mergers with $10^2<M<10^7$,} the error measure remains below $\epsilon_{\Psi} \sim 0.1$, where the corrections are still well manageable by the perturbative formalism (as we will show in Sec.~\ref{subsec:errorsLISA}).
+For $\epsilon_{\Psi 2}$, as expected from~\eqref{eq:TfN} we find a steep rise towards lower frequencies, which shows that considering merging binaries with a limited time to coalescence is crucial here.
+\jgb{We note differences in} behaviour between the orbital response, where for a fixed mass ratio the initial \jgb{(10yr before merger)} value of $\epsilon_{\Psi 2}$ grows towards lower masses, and the constellation response, where \jgb{initial} $\epsilon_{\Psi 2}$ grows towards higher masses.
+\jgb{Combining both parts of the response, for equal-mass mergers with $10^2<M<10^7$,} the error measure remains below $\epsilon_{\Psi 2} \sim 0.1$, where the corrections are still well manageable by the perturbative formalism (as we will show in Sec.~\ref{subsec:errorsLISA}).
 
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e7dO_py.pdf}
-  \caption{Error in the transfer function for the orbital delay $d_{0}$, for $M=10^{7} \Msol$.}
+  \caption{Transfer function and residual for the orbital delay $d_{0}$~\eqref{eq:delay0}, for an equal-mass system with $M=10^{7} \Msol$.
+The first and second panels show the amplitude and phase of the transfer function.
+The bottom panel shows the relative modulus of the residual of the perturbative treatment~\eqref{eq:summaryNAd}, with the color indicating the order of the approximation, compared to a numerical FFT.
+The starting frequency is set by the LISA sensitivity band at $10^{-5}\mathrm{Hz}$, corresponding to $3.8\mathrm{yr}$.
+The thick and thin vertical lines indicate the merger and ringdown frequencies.}
   \label{fig:LISAerrorM1e7orb}
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e2dO_py.pdf}
-  \caption{Transfer function and reconstruction error for the orbital delay $d_{0}$, for $M=10^{2} \Msol$.}
+  \caption{Transfer function and residual for the orbital delay $d_{0}$~\eqref{eq:delay0}, for an equal-mass system with $M=10^{2} \Msol$.
+The panels are as in Fig.~\ref{fig:LISAerrorM1e7orb}.
+The starting frequency corresponds to four years \SM{[check]} before merger.
+The ending frequency is set by the LISA sensitivity band at $1\mathrm{Hz}$, and the merger is out of band.
+The jump in the residuals for $\{N:3|A:1|d:1\}$ is due to the fact that we split the band in two, with two different sampling rates, for the FFT; it shows that, at this level, the details of the conditioning for the FFT start to cause numerical errors in the reference transfer function.}
   \label{fig:LISAerrorM1e2orb}
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e7y12L_py.pdf}
-  \caption{Error in the transfer function for the basic observable $y_{12}$, for $M=10^{7} \Msol$.}
+  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:GslrL}, for an equal-mass system with $M=10^{7} \Msol$.
+The top panel shows the real and imaginary parts of the transfer function, rescaled by the overall scaling $\pi f L$ of the response at low frequencies, while the bottom panel shows the relative residuals compared to an FFT.
+The frequencies are the same as in Fig.~\ref{fig:LISAerrorM1e7orb}.}
   \label{fig:LISAerrorM1e7const}
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e2y12L_py.pdf}
-  \caption{Transfer function and reconstruction error for the basic observable $y_{12}$, for $M=10^{2} \Msol$. \SM{[check $\overline{\calT}$ is defined]}}
+  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:GslrL}, for an equal-mass system with $M=10^{2} \Msol$.
+The top panel shows the real and imaginary parts of the transfer function, without the rescaling of Fig.~\ref{fig:LISAerrorM1e7const}.
+The bottom panels shows the residuals normalized by~\eqref{eq:transferLenvelope} to avoid zero-crossings.
+The frequencies are the same as in Fig.~\ref{fig:LISAerrorM1e2orb}, and the discontinuity in the residuals for $\{N:3|A:1|d:1\}$ is again due to the conditioning of the FFT.}
   \label{fig:LISAerrorM1e2const}
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi1d0corr_py.pdf}
-  \caption{Error in the transfer function for the orbital delay $d_{0}$, for a SOBH with $\epsilon_{\Psi}\sim 1$.}
-  \label{fig:}
+  \caption{Transfer function and residual for the orbital delay $d_{0}$~\eqref{eq:delay0}, for an SOBH with $\epsilon_{\Psi 2}\sim 1$.
+The frequency band corresponds to 4 years of observation for this system with $M=50\Msol$, $\Delta t = 200\yr$ away from merger.
+The top panels show the transfer function amplitude and phase at various orders of approximation in~\eqref{eq:summaryNAd}, with the numerical result of the FFT in black and the comb result~\eqref{eq:transferdiscreteconvolution} in green.
+The bottom panel shows the relative residuals with respect ot the FFT.}
+  \label{fig:LISAerrorSOBHepsPsi1d0}
 \end{figure}
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi10d0corr_py.pdf}
-  \caption{Error in the transfer function for the orbital delay $d_{0}$, for a SOBH with $\epsilon_{\Psi}\sim 10$.}
-  \label{fig:}
+  \caption{Same as Fig.~\ref{fig:LISAerrorSOBHepsPsi1d0}, but for an SOBH with $\epsilon_{\Psi 2}\sim 10$.
+The frequency band corresponds to 4 years of observation for this system with $M=15\Msol$, $\Delta t = 1500\yr$ away from merger.
+The perturbative treatment breaks down in this case, with errors of order 1.}
+  \label{fig:LISAerrorSOBHepsPsi10d0}
 \end{figure}
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi1y12L_py.pdf}
-  \caption{Error in the transfer function for the constellation response $y_{132}$, for a SOBH with $\epsilon_{\Psi}\sim 1$.}
+  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:GslrL}, for the same SOBH as in Fig.~\ref{fig:LISAerrorSOBHepsPsi1d0}, with $\epsilon_{\Psi 2}\sim 1$.
+The top panel shows the real and imaginary part of the transfer function.}
   \label{fig:}
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi10y12L_py.pdf}
-  \caption{Error in the transfer function for the constellation response $y_{132}$, for a SOBH with $\epsilon_{\Psi}\sim 10$.}
+  \caption{Same as }
   \label{fig:}
@@
 As explained in Sec.~\ref{subsec:lisafom}, slowly-chirping binaries can be problematic for the perturbative formalism of Sec.~\ref{sec:formalism}.
-Indeed, when far enough from merger, SOBH systems~\cite{Sesana16} can have error estimates $\epsilon_{\Psi}$ reaching and exceeding $1$, as shown in Fig.~\ref{fig:lisafomPsiMcDeltat}.
+Indeed, when far enough from merger, SOBH systems~\cite{Sesana16} can have error estimates $\epsilon_{\Psi 2}$ reaching and exceeding $1$, as shown in Fig.~\ref{fig:lisafomPsiMcDeltat}.
 Here we investigate these sources in more details, and propose an alternative treatment for their instrument response.
@@
 
-\sout{We will next apply our formalism to the decomposition of signals from precessing binaries.}
 As we already discussed in Sec.~\ref{subsec:modulationPrec}, in the presence of spin components that are not aligned with the orbital angular momentum, an inspiraling binary system will undergo precession of its orbital plane~\cite{Apostolatos+94, Kidder95}.
@@
 \ee
-In the notation that we used in Sec.~\ref{sec:formalism}, in the absence delays, we wish to compute the convolutions
+In the notations that we used in Sec.~\ref{sec:formalism}, in the absence delays, we wish to compute the convolutions
 \be\label{eq:precconvolution}
@@
 
-We will use the following notation: for $m_{1}$, $m_{2}$ the masses of the two bodies, we set $M=m_{1}+m_{2}$, $\nu=m_{1}m_{2}/M^{2}$, $\delta = (m_{1}-m_{2})/M$, and $|\bm{S}_{A}|=Gm_{A}^{2} \chi_{A}$ for $A=1,2$, with $\chi$ the dimensionless spin between $0$ and $1$ for Kerr black holes.
+We will use the following notations: for $m_{1}$, $m_{2}$ the masses of the two bodies, we set $M=m_{1}+m_{2}$, $\nu=m_{1}m_{2}/M^{2}$, $\delta = (m_{1}-m_{2})/M$, and $|\bm{S}_{A}|=Gm_{A}^{2} \chi_{A}$ for $A=1,2$, with $\chi$ the dimensionless spin between $0$ and $1$ for Kerr black holes.
 We define $\bm{\ell}$ as the unit vector normal to the orbital plane.
@@
 The overall $\nu$ scaling indicates a better separation of timescales for larger mass ratio systems.
-These expressions also show that, since the total angular momentum appears as a factor in $\Omega_{\rm prec}$, higher-order corrections will be larger for \jgb{roughly} spin-aligned systems than for anti-aligned spins, an effect that becomes significant for large mass ratios.
+These expressions also show that, since the total angular momentum appears as a factor in $\Omega_{\rm prec}$, higher-order corrections will be larger for spin-aligned systems than for anti-aligned spins, an effect that becomes significant for large mass ratios.
 Note that the regime where $\xi$ gets close to 0 corresponds to the transitional precession range~\cite{Apostolatos+94}, with the spin of the primary compensating the orbital angular momentum, and our analysis based on simple precession is not valid anymore.
@@
 
-Instead of tapering the signal to $0$ \jgb{below [CHECK]} $f_{0}$, to limit the deviation from the original signal we take advantage of the one-sidedness and only flatten the amplitude\footnote{In practice, this is implemented as the discrete integral of a cosine window function on just the first two samples.
+Instead of tapering the signal to $0$ at $f_{0}$, to limit the deviation from the original signal we take advantage of the one-sidedness and only flatten the amplitude\footnote{In practice, this is implemented as the discrete integral of a cosine window function on just the first two samples.
 We also taper to 0 the last three samples before $f_{\rm max}$, and $0$-pad by a factor of 2.}, and artificially symmetrize the signal.
@@
 The thick and thin vertical lines show the merger and asymptotic ringdown frequencies respectively.
-\jgb{For the solid curves in the top panel, precession modulations were computed in the time domain, while the dashed curves show the results of the Fourier-domain reconstruction techniques presented here.[CHECK]}
 The columns show the orders of approximation $\{\Psi:0,A:0\}$, $\{\Psi:2,A:2\}$, $\{\Psi:2,A:2,\text{conv}\}$.
@@
   \includegraphics[width=.98\linewidth]{plots/precMM_py.pdf}
-  \caption{Unfaithfulness at various orders of approximation \jgb{as compared with a time-domain treatment using a full-domain[CHECK]} a numerical FFT, for the three precessing systems of Table~\ref{tab:precexamples}.
+  \caption{Unfaithfulness at various orders of approximation and a numerical FFT, for the three precessing systems of Table~\ref{tab:precexamples}.
 The color indicates the order of approximation of the perturbative treatment of Sec.~\ref{subsec:executivesummary}.

### 2018-06-15 docs/paper_method.tex
Edits to the paper.

@@
 
-\SM{[style: below use texttt for models and codes (like \texttt{SpEC})]}
-
 \SM{[check citation of Cornish-Rubbo-Poujade early work]}
@@
 We show only the high-frequency part of the signal, corresponding to the merger region.
-The \texttt{PhenomD} waveforms have been aligned such that the time-domain amplitude (obtained by an IFFT) peaks at $t=0$.
+The PhenomD waveforms have been aligned such that the time-domain amplitude (obtained by an IFFT) peaks at $t=0$.
 The two aligned spins are equal, with components of $0.95$ ($++$, blue), $0$ ($00$, red), and $-0.95$ ($--$, yellow), for mass ratios $q=1$ (full line) and $q=8$ (dashed).
@@
 
-Having at hand estimates for the relevance of higher-order corrections in the LISA response, we now demonstrate that including these higher-order corrections does indeed reduce the reconstruction error of the Fourier-domain transfer function.
-We will consider an equal-mass and non-spinning system, with only two values for the total mass, $M=10^{7} \Msol$ and $M=10^{2} \Msol$.
-For this example, we will also focus on the orbital response and on a single-arm constellation response $y_{132}$, as TDI combinations will appear (within the approximations listed in Sec.~\ref{subsec:modelLISA}) as linear combinations of these basic observables.
+Having at hand estimates for the relevance of higher-order corrections in the LISA response, we assess the performance of the leading-order treatment and we demonstrate that including higher-order corrections consistently reduces the reconstruction error of the Fourier-domain transfer function.
+We will consider two equal-mass and non-spinning systems for this illustration.
+The first will be a high-mass system with $M=10^{7} \Msol$, entering the LISA band at $f=10^{-5}\Hz$ and inspiralling for $\Delta t = 3.8 \yr$ before merging.
+The second will be a light system with $M=10^{2} \Msol$, close to the mass range of SOBHs~\cite{Sesana16}, starting at $\Delta t = 10\yr$ before merger and exiting the LISA band at $f=1\Hz$.
+We will discuss SOBH systems farther away from merger in the next section.
+We will also focus on the orbital response and on the single-arm constellation response $y_{132}$, as TDI combinations will appear (within the approximations listed in Sec.~\ref{subsec:modelLISA}) as linear combinations of these basic observables.
 The orientation angles chosen for these examples are $[\lambda,\beta,\iota,\psi] = [\pi/4,\pi/3,\pi/3,0]$.
@@
 On one hand, we perform a numerical inverse Fourier transform (IFFT) of the Fourier-domain PhenomD waveform, process the signal through the time-domain response~\eqref{eq:defresponse0} or~\eqref{eq:defresponseL}, and perform a numerical Fourier transform (FFT).
-This gives us the target Fourier-domain waveform, considered to be exact but for possible numerical artifacts (most notably, oscillations induced by the necessary tapering of the signal).
+This gives us the target Fourier-domain waveform, considered to be exact but for possible numerical artifacts\footnote{Numerical Fourier transforms include most notably oscillations induced by the necessary tapering of the signal.
+In the low-mass case, in practice we stitch together two frequency bands with different sampling rates, which leaves some visible residual in Figs.~\ref{fig:LISAerrorM1e2orb} and~\ref{fig:LISAerrorM1e2const}.}.
 On the other hand, we process the Fourier-domain signal through the response summarized in Sec.~\ref{subsec:executivesummary}, including various higher-order corrections.
@@
 
-We show the results for the orbital response in Figs.~\ref{fig:LISAerrorM1e7orb} and~\ref{fig:LISAerrorM1e2orb} in amplitude and phase form.
+The results for the orbital response~\eqref{eq:defresponse0} are shown in Figs.~\ref{fig:LISAerrorM1e7orb} and~\ref{fig:LISAerrorM1e2orb} in amplitude and phase form.
 In both cases, in accordance with~\eqref{eq:transfer0local}, the transfer function is essentially a phase, and contains many more cycles in the low-mass case in the higher frequency band.
-We find that including higher-order corrections does reduce the reconstruction errors, but with the distinctive feature that including only one of the amplitude and delay corrections can make the error actually worse, while including both does improve the accuracy.
-In the high-mass case, we also find for all orders of approximation an increase in the errors at the very end.
-This is likely to be due to the fact that the errors in the transfer function are to be interpreted in the relative sense, and the amplitude of the signal is decaying rapidly \SM{[ref to amplitude plot somewhere]} in this post-merger region.
-Although we did not identify the precise cause of this rise in errors, one possible explanation would be tapering-induced oscillations in the numerical Fourier transforms.
-
-For the constellation response, we show the results in Figs.~\ref{fig:LISAerrorM1e7const} and~\ref{fig:LISAerrorM1e2const}, using directly the real and imaginary part of the transfer function.
-We find similar features as for the orbital response: including higher-corrections reduces the errors, but in the high-mass case the amplitude and delay corrections should be added together.
-We also find the same growth of errors at the very end of the signal in the high-mass case, in the region where the amplitude of the signal decays.
+The leading-order treatment is accurate at $10^{-3}$ in the high-mass case, while in the low-mass case errors reach $10\%$ at the start of observations and a few percent at high frequencies.
+In both cases, including higher-order corrections does reduce the reconstruction errors, with the distinctive feature that including only one of the amplitude and delay corrections can make the error actually worse, while including both does improve the accuracy down to $10^{-5}$ or better.
+
+For the constellation response~\eqref{eq:defresponseL}, we show the results in Figs.~\ref{fig:LISAerrorM1e7const} and~\ref{fig:LISAerrorM1e2const}, displaying the real and imaginary part of the transfer function.
+In the high-mass case, we rescale the response by the overall scaling $\pi f L$ in~\eqref{eq:GslrL}.
+The leading-order treatment reaches an inaccuracy of $\sim 10\%$ at the lowest frequencies of the high-mass case, while it is better than $10^{-3}$ for the low-mass case.
+Similarly to the orbital response, including all higher-corrections reduces the errors to better than $10^{-5}$, with the only exception of the high-mass case showing errors rising at the low-frequency end and beyond the ringdown frequency.
 
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e7dO_py.pdf}
-  \caption{Transfer function and residual for the orbital delay $d_{0}$~\eqref{eq:delay0}, for an equal-mass system with $M=10^{7} \Msol$.
+  \caption{Transfer function and residual for the orbital response~\eqref{eq:defresponse0}, for an equal-mass system with $M=10^{7} \Msol$.
 The first and second panels show the amplitude and phase of the transfer function.
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e2dO_py.pdf}
-  \caption{Transfer function and residual for the orbital delay $d_{0}$~\eqref{eq:delay0}, for an equal-mass system with $M=10^{2} \Msol$.
+  \caption{Transfer function and residual for the orbital response~\eqref{eq:defresponse0}, for an equal-mass system with $M=10^{2} \Msol$.
 The panels are as in Fig.~\ref{fig:LISAerrorM1e7orb}.
-The starting frequency corresponds to four years \SM{[check]} before merger.
+The starting frequency corresponds to 10 years before merger.
 The ending frequency is set by the LISA sensitivity band at $1\mathrm{Hz}$, and the merger is out of band.
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e7y12L_py.pdf}
-  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:GslrL}, for an equal-mass system with $M=10^{7} \Msol$.
+  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:defresponseL}, for an equal-mass system with $M=10^{7} \Msol$.
 The top panel shows the real and imaginary parts of the transfer function, rescaled by the overall scaling $\pi f L$ of the response at low frequencies, while the bottom panel shows the relative residuals compared to an FFT.
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e2y12L_py.pdf}
-  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:GslrL}, for an equal-mass system with $M=10^{2} \Msol$.
+  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:defresponseL}, for an equal-mass system with $M=10^{2} \Msol$.
 The top panel shows the real and imaginary parts of the transfer function, without the rescaling of Fig.~\ref{fig:LISAerrorM1e7const}.
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi1d0corr_py.pdf}
-  \caption{Transfer function and residual for the orbital delay $d_{0}$~\eqref{eq:delay0}, for an SOBH with $\epsilon_{\Psi 2}\sim 1$.
+  \caption{Transfer function and residual for the orbital delay $d_{0}$~\eqref{eq:delay0}, for an SOBH with $\epsilon_{\Psi 2}\sim 1$ for the orbital response.
 The frequency band corresponds to 4 years of observation for this system with $M=50\Msol$, $\Delta t = 200\yr$ away from merger.
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi10d0corr_py.pdf}
-  \caption{Same as Fig.~\ref{fig:LISAerrorSOBHepsPsi1d0}, but for an SOBH with $\epsilon_{\Psi 2}\sim 10$.
+  \caption{Same as Fig.~\ref{fig:LISAerrorSOBHepsPsi1d0}, but for an SOBH with $\epsilon_{\Psi 2}\sim 7$ for the orbital response.
 The frequency band corresponds to 4 years of observation for this system with $M=15\Msol$, $\Delta t = 1500\yr$ away from merger.
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi1y12L_py.pdf}
-  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:GslrL}, for the same SOBH as in Fig.~\ref{fig:LISAerrorSOBHepsPsi1d0}, with $\epsilon_{\Psi 2}\sim 1$.
+  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:GslrL}, for the same SOBH as in Fig.~\ref{fig:LISAerrorSOBHepsPsi1d0}.
+Here $\epsilon_{\Psi 2}\sim 0.01$ for the constellation part of the response.
 The top panel shows the real and imaginary part of the transfer function.}
@@
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi10y12L_py.pdf}
-  \caption{Same as }
+  \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:GslrL}, for the same SOBH as in Fig.~\ref{fig:LISAerrorSOBHepsPsi10d0}.
+Here $\epsilon_{\Psi 2}\sim 0.1$ for the constellation part of the response.
+The top panel shows the real and imaginary part of the transfer function.}
   \label{fig:}

### 2018-06-17 docs/paper_method.tex
Edits to the paper.

@@
 From~\eqref{eq:transferLlocal}, an expansion for small $f\ll f_{L}$ yields back a Fourier-domain analog of the low-frequency approximation of the response~\cite{Cutler97, RCP04}, which \jgb{is equivalent to} having two LIGO-type interferometers turned by $\pi/4$ and set in motion.
-See Fig.~\ref{} and ~\ref{} \SM{[R]} for examples of LISA transfer functions.
-
-\jgb{For analysis of the response we need a concrete set of gravitational waveforms.
-We will} use the PhenomD model~\cite{Khan+15,Husa+15}, which provides Fourier-domain inspiral-merger-ringdown waveforms for aligned spins.
+
+For analysis of the response we need a concrete set of gravitational waveforms.
+We will use the PhenomD model~\cite{Khan+15,Husa+15}, which provides Fourier-domain inspiral-merger-ringdown waveforms for aligned spins.
 
@@
 To obtain an order-of-magnitude estimate for the error measures $\epsilon$~\eqref{eq:deffom}, we will use the Newtonian-order expressions~\eqref{eq:timescalesN} for the signal-dependent timescales $\Tf$ and $T_{A1}$.
-Higher-order amplitude terms beyond the first one in~\eqref{eq:resulttaylorA}, as well as phase terms beyond the second derivative in~\eqref{eq:expandPsi}, will turn out to be completely negligible in the LISA case and we will ignore them in the following.
+Higher-order amplitude terms beyond the first one in~\eqref{eq:resulttaylorA}, as well as phase terms beyond the second derivative in~\eqref{eq:expandPsi}, will turn out to be  negligible in the LISA case and we will ignore them in the following.
 It is useful to separate the orbital response~\eqref{eq:transfer0local} and the constellation response~\eqref{eq:transferLlocal}, as the different baseline of the delays (orbital radius $R$ or armlength $L$) as well as the presence of a time-varying prefactor $F(t)$ both affect the result.
@@
 
-We start with the orbital response, which takes the form of a pure delay $G_{0}(f, t) = e^{-2 i \pi f d_{0}(t)}$ \jgb{\bf[Why subscripted with zero not O or R?]}, and obtain
+We start with the orbital response, which takes the form of a pure delay $G_{0}(f, t) = e^{-2 i \pi f d_{0}(t)}$, and obtain
 \begin{subequations}
@@
 Since the one-arm constellation response~\eqref{eq:decomposeGslr}-\eqref{eq:transferLlocal} is analogous to a discrete time derivative of the signal, it is appropriate to keep explicit an overall factor $f$ reflecting this structure.
-Hence we write symbolically $G_{L}(f,t) \sim f F(t) e^{-2i \pi f d_{L}(t)}$, where $d_{L}$ represents a delay term and $F(t)$ represents the rest of the geometric factors in~\eqref{eq:transferLlocal}, \jgb{for which we momentarily} ignore the $f$-dependence.
+Hence we write symbolically $G_{L}(f,t) \sim f F(t) e^{-2i \pi f d_{L}(t)}$, where $d_{L}$ represents a delay term and $F(t)$ represents the rest of the geometric factors in~\eqref{eq:transferLlocal}, for which we momentarily ignore the $f$-dependence.
 This gives
@@
 
-\jgb{\bf[This paragraph is confusing.
-Can we point concretely to these factors somewhere, (e.g.
-in the expansion of the exponential.xxx)?
-Also it seems to say that we can't do what we just did.
-Should discuss.]} An important point in the above is the presence of delay factors in the forms of powers of $2\pi f d$.
-Thus, the magnitude of the corrections cannot be evaluated with the simple replacement $\partial_{t}^{n} \rightarrow \Omega_{0}^{n}$, and the frequency dependency of the error measures $\epsilon$ will depend on the frequency being above or below the transfer frequencies $f_{R}$ and $f_{L}$ defined in~\eqref{eq:transferfrequencies}.
+An important point in~\eqref{eq:estimatederivorb} and~\eqref{eq:estimatederivconst} above is the presence of delay factors in the forms of powers of $2\pi f d$, so that we cannot use the simple replacement $\partial_{t}^{n}G \rightarrow \Omega_{0}^{n}G$.
+Thus, the suitability of the leading-order treatment cannot be estimated by a mere separation between the signal timescales and the annual timescale for the response, but will also depend on the frequency being above or below the transfer frequencies $f_{R}$ and $f_{L}$ defined in~\eqref{eq:transferfrequencies}.
 The error measure $\epsilon_{d}$ does not depend on signal-dependent timescales and can be directly read off the estimates above, with $\epsilon_{d} = |\partial_{tf}G/(2\pi G)|$.
-For the error measures $\epsilon_{\Psi 2}$ and $\epsilon_{A1}$, we can combine the above derivatives with the Newtonian timescales given in~\eqref{eq:timescalesN} for the inspiral phase of the signal.
+For the error measures $\epsilon_{\Psi 2}$ and $\epsilon_{A1}$, we will combine the above derivatives with the Newtonian timescales given in~\eqref{eq:timescalesN} for the inspiral phase of the signal.
 
@@
 \end{subequations}
-Note that if we \jgb{think of LISA as effectively insensitive below some minimal frequency $f_{\rm min}$, then for sufficiently high $\Mchirp$ this point of entry in the sensitive band will mark the beginning of the signal, obviating the relevance of $f_{\rm start}$}.
+Note that if we think of LISA as effectively insensitive below some minimal frequency $f_{\rm min}$, then for sufficiently high $\Mchirp$ this point of entry in the sensitive band will mark the beginning of the signal, obviating the relevance of $f_{\rm start}$.
 For $f_{\rm min} = 10^{-5}\Hz$ and $\Delta t \leq 10\yr$, this is the case for $\Mchirp \geq 2.45\times 10^{6}\Msol $.
@@
 
-\jgb{It is useful to distinguish} between what we will call merging binaries, systems close to coalescence that will merge during the LISA mission lifetime or a few years later, and slowly-chirping binaries still in the deep inspiral phase, of which we observe only a small snapshot in Fourier domain as they do not sweep to the end of the frequency band.
+It is useful to distinguish between what we will call merging binaries, systems close to coalescence that will merge during the LISA mission lifetime or a few years later, and slowly-chirping binaries still in the deep inspiral phase, of which we observe only a small snapshot in Fourier domain as they do not sweep to the end of the frequency band.
 The massive black hole binaries (MBH) that will be observed by LISA~\cite{LISA17} fall within the first category, with their merger in band, while the proposed population of stellar-origin black hole binaries (SOBH)~\cite{Sesana16}, with masses comparable to the LIGO/Virgo detections, will comprise both merging binaries, i.e.
 exiting the LISA band towards larger frequencies during observations, and slowly-chirping binaries hundereds or thousands of years away from merger.
-Galactic binaries (GB) fall entirely within the second category, and are very close to monochromatic over the lifetime of LISA~\cite{LISA17} \SM{[C: complete]}.
 
@@
 The limit $\epsilon_{\Psi 2} = 1$ is used to single out areas where the perturbative treatment of Sec.~\ref{sec:formalism} is expected to break down, although the precise location of this boundary will vary, depending on the accuracy level required and on the orientation angles.
-We also single out the high-mass region where the start of the signal is set by its entry \jgb{into the sensitive band, at $f_{\rm min}$ making the error measure independent of time to merger}.
+We also single out the high-mass region where the start of the signal is set by its entry into the sensitive band at $f_{\rm min}$, making the error measure independent of the time to merger.
 We find that, for binaries in the LIGO/Virgo mass range $10^{1}-10^{2} \Msol$, we reach the limit $\epsilon_{\Psi 2}=1$ first for the orbital response, for time-to-merger values within the expected observed range~\cite{Sesana16}.
 We will introduce in Sec~\ref{subsec:comblisa} an alternative approach to deal with those signals.
-For completeness, we also show in Fig.~\ref{fig:lisafomPsiMcDeltat} the result for intermediate and massive black hole systems, although we expect to observe mainly merging systems for this mass range.
+For completeness, we also show in Fig.~\ref{fig:lisafomPsiMcDeltat} the result for intermediate and massive black hole systems, although we expect to observe merging systems for this mass range.
 
@@
 To go beyond the crude analytical estimates built from~\eqref{eq:timescalesNfstart} and~\eqref{eq:estimatederivorb}-\eqref{eq:estimatederivconst}, we use numerical derivatives for the timescales~\eqref{eq:defTf}-\eqref{eq:defTA} and for the derivatives in~\eqref{eq:deffom}.
-We summarize the results in Fig.~\ref{fig:fomLISA}, both for the orbital delay and the constellation response, considering equal-mass systems with total masses of $M=10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$ with a starting frequency corresponding to $\Delta t =10 \yr$ of observation.
-Since individual signals can show significant variation depending on the angular parameters, for the full computation with numerical derivatives we show a geometric average of $\epsilon$ over the sky position, inclination and polarization, together with $\pm 1\sigma$ geometric standard deviation.
+We summarize the results in Fig.~\ref{fig:fomLISA}, both for the orbital delay and the constellation response, considering equal-mass non-spinning systems with total masses of $M=10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$ with a starting frequency corresponding to $\Delta t =10 \yr$ of observation.
+Since individual signals can show significant variation depending on the orientation parameters, for the full computation with numerical derivatives we show a geometric average of $\epsilon$ over the sky position, inclination and polarization, together with $\pm 1\sigma$ geometric standard deviation.
 Here and in the following, for completeness we show results on the frequency band from $10^{-5}\Hz$ to $1\Hz$, while the signals are expected to have little signal-to-noise ratio outside of the band from $10^{-4}\Hz$ to $10^{-1}\Hz$, which is sometimes taken as a reference~\cite{LISA17}.
@@
 For $\epsilon_{\Psi 2}$, as expected from~\eqref{eq:TfN} we find a steep rise towards lower frequencies, which shows that considering merging binaries with a limited time to coalescence is crucial here.
-\jgb{We note differences in} behaviour between the orbital response, where for a fixed mass ratio the initial \jgb{(10yr before merger)} value of $\epsilon_{\Psi 2}$ grows towards lower masses, and the constellation response, where \jgb{initial} $\epsilon_{\Psi 2}$ grows towards higher masses.
-\jgb{Combining both parts of the response, for equal-mass mergers with $10^2<M<10^7$,} the error measure remains below $\epsilon_{\Psi 2} \sim 0.1$, where the corrections are still well manageable by the perturbative formalism (as we will show in Sec.~\ref{subsec:errorsLISA}).
+We note differences in behaviour between the orbital response, where for a fixed mass ratio the initial (10yr before merger) value of $\epsilon_{\Psi 2}$ grows towards lower masses, and the constellation response, where the initial $\epsilon_{\Psi 2}$ grows towards higher masses.
+Combining both parts of the response, for mergers with $10^2<M<10^7$, the error measure remains below $\epsilon_{\Psi 2} \sim 0.1$, where the corrections are still well manageable by the perturbative formalism (as we will show in Sec.~\ref{subsec:errorsLISA}).
 
@@
 For the orbital response, $\epsilon_{d_{0}}$ grows for higher frequencies, reaching the $\sim 0.1$ at around $1\Hz$.
-For the constellation response, $\epsilon_{d_{c}}$ is mainly important at lower frequencies, reaching $\sim 0.01$ at around $10^{-5}\Hz$.
-
-Overall, we find that for merging binaries the perturbative approach presented in Sec.~\ref{sec:formalism} will be applicable, as the smallness of the annual frequency $f_{0} \simeq 3.1\times10^{-8}\mathrm{Hz}$, compared to the timescales relevant for the signal, ensures a good separation of timescales on the whole LISA frequency band.
+For the constellation response, $\epsilon_{d_{L}}$ is mainly important at lower frequencies, reaching $\sim 0.01$ at around $10^{-5}\Hz$.
+
+Overall, we find that for merging binaries the perturbative approach presented in Sec.~\ref{sec:formalism} will be applicable on the whole LISA frequency band.
 The most relevant higher-order corrections are expected to be the phase corrections close to the starting frequency of the signal, and the delay correction at high frequencies for the orbital response.
@@
 \label{subsec:errorsLISA}
-
-\SM{[We have to say why we take a non-spinning, equal-mass system - say why the method should also work for precessing signals]}
-
-\SM{[Write explicitly what the orders A1, d1 mean]}
-
-\SM{[argument why we do not compute mismatches]}
-
-\SM{[stress that binaries very far from merger could have arbitrarily large Psi2 corrections]}
-
-\SM{[say that we take a large number of samples to eliminate spline interpolation errors]}
 
@@
 For $c_{n}^{0}$, the expression~\eqref{eq:cn0} can provide an asymptotic bound on the width of the comb.
-For large values of $n$ we have indeed the equivalent $| J_{n}[z] | \sim (e z/2n)^{n}/\sqrt{2\pi n} \; \left[ n\rightarrow + \infty \right]$~\cite{} \SM{[C: dlmf?]}, so that a conservative estimate for the truncation order (almost independent of $\eta$) is given by $N \geq  |e \pi f R \cos \beta|$.
+For large values of $n$ we have indeed the equivalent $| J_{n}[z] | \sim (e z/2n)^{n}/\sqrt{2\pi n} \; \left[ n\rightarrow + \infty \right]$~\cite{DLMF} \SM{[C: dlmf?]}, so that a conservative estimate for the truncation order (almost independent of $\eta$) is given by $N \geq  |e \pi f R \cos \beta|$.
 Fig.~\ref{fig:lisacombextent} shows numerical computations for $N(f,\eta)$, averaged over orientation angles, for both the constellation and the orbital response and for the truncation levels  $\eta=10^{-12}$, $10^{-6}$, $10^{-3}$.
@@
   \includegraphics[width=.98\linewidth]{plots/eulerSXS0058_py.pdf}
-  \caption{Evolution of Euler angles $(\alpha, \beta, \gamma)$ near merger for the example SpEC waveform SXS:BBH:0058~\cite{SpEC, SXScatalog, Mroue+12, Mroue+13} \SM{[C]}.
+  \caption{Evolution of Euler angles $(\alpha, \gamma)$ (top panel) and $\beta$ (bottom panel) near merger for the example SpEC waveform SXS:BBH:0058~\cite{SpEC, SXScatalog, Mroue+12, Mroue+13}.
 Full, dashed and dotted lines represent the NR, SEOB and extended SEOB waveforms respectively.
@@
 \SM{[Add amplitude/$\omega$ plots comparing NR/SEOB, and possibly PhenomD for the 22 mode ?
-Add Euler angles produced by IFFT of PhenomP ?]} \SM{[remove SEOBext in the legend]} \SM{[confusion of dashed black lines]}}
+Add Euler angles produced by IFFT of PhenomP ?]}}
   \label{fig:precmodel}
@@
 
-\SM{[add plot of QNM frequency and frequency diff according to final spin ?]}.
-
 \SM{[justify somewhere smallness of cubic-in-phase term]}
@@
 The results are shown in Fig.~\ref{fig:fomprec}, using the merger frequency and the ringdown frequency, shown by the vertical lines, to give an idea of the separation between the inspiral and post-merger phases.
+The leading-order PN prediction for single-spin simple precession~\eqref{eq:precfomPN} is overlayed for the cases $++$ and $\perp\perp$, but not for the case $--$ as it departs significantly from simple precession.
 We can take $\epsilon \sim 1$ as an order-of-magnitude indication of the breakdown of a perturbative treatment.
@@
 
-The magnitude of the error estimates for $++$ and $\perp\perp$ is roughly in agreement with the PN-inspired computation~\eqref{eq:precfomPN} above for the inspiral part of the signal \SM{[check if the PN is quantitatively ok -- overlay it on the plot ?]} \SM{[Pb: scaling of FOM might be dominated here by SS oscillations, notably for 2nd der.
-?
-To check - check also geometric scaling from mode to mode]}, with $\epsilon_{\Psi 2}$ showing a negative slope $v^{-1}=(Mf)^{-1/3}$, and with a hierarchy between modes due to the factors of $|m' \cos \beta - m|$.
+The magnitude of the error estimates for $++$ and $\perp\perp$ is roughly in agreement with the PN-inspired computation~\eqref{eq:precfomPN} above for the inspiral part of the signal, with $\epsilon_{\Psi 2}$ showing a negative slope $v^{-1}=(Mf)^{-1/3}$, and with a hierarchy between modes due to the factors of $|m' \cos \beta - m|$.
+The PN estimates are missing oscillations on the precession timescale due to the double-spin precession.
 The case $--$ departs from the simple precession picture, and~\eqref{eq:precfomPN} does not apply.
@@
 The colors correspond to different values of $m$ in $\calT^{2}_{m 2}$ as defined in~\eqref{eq:defprectransfer}.
-We show the three error estimates $\epsilon_{\Psi 2}$, $\epsilon_{A1}$, $\epsilon_{A2}$ defined in~\eqref{eq:deffom}, and the range $\epsilon \gtrsim 1$ indicates a breakdown of the formalism of Sec.~\ref{sec:formalism}.
-\SM{[overlay PN simple precession ?]}}
+The full lines are computed with numerical derivatives for the full waveform and modulation, while in the cases $++$ and $\perp\perp$ the dashed lines show the leading-order PN estimates~\eqref{eq:precfomPN} for single-spin simple precession.
+We show the three error estimates $\epsilon_{\Psi 2}$, $\epsilon_{A1}$, $\epsilon_{A2}$ defined in~\eqref{eq:deffom}, and the range $\epsilon \gtrsim 1$ indicates a breakdown of the formalism of Sec.~\ref{sec:formalism}.}
   \label{fig:fomprec}
@@
 
-\SM{[possibly move this to the formalism section ? but plots for our three examples]}
-
-\SM{[say that another possibility would be to go back to time-domain and FFT for the limited-length near-merger part of the signal]}
-
-\SM{[expand motivation: sharp feature TD $\rightarrow$ smooth in FD $\rightarrow$ hope to have a cheap trigo representation of FD merger range]}
-
 As shown by Fig.~\ref{fig:fomprec}, the faster evolution of the modulation functions in the post-merger phase and the resulting weaker separation of timescales can be expected to be challenging for the perturbative formalism layed out in Sec.~\ref{sec:formalism}.
@@
 In the correspondence between the time and Fourier domain, sharp features in the time domain map to extended features in the Fourier domain, and vice versa.
-The merger and ringdown part of the Fourier-domain signal extends over a wide range of frequencies, while it corresponds to a short interval of times, much shorter than the inspiral phase.
+The merger and ringdown part of the Fourier-domain signal extends over a wide range of frequencies, while it corresponds to a short interval of times, much shorter than the inspiral part.
 One approach would be to separate the time-domain waveform between the inspiral and merger phase, build a Fourier-domain model for the inspiral part only, to which the formalism of Sec.~\ref{sec:formalism} could be applied, while the merger and ringdown part could be handled by a direct FFT, which would be a cheap operation on a limited range in time.
@@
 Our treatment of the Fourier-domain LISA instrumental response will allow us to leverage modern, fast Fourier-domain codes producing IMR waveforms to accelerate computationally intensive data analysis applications, most notably Bayesian source parameter estimation on synthetic data, that have become urgent with the selection of LISA as a mission by ESA.
-The LISA response described in Sec.~\ref{sec:LISA} has been implemented in the LISA Data Challenge~\cite{} software suite.
-\SM{[cite LDC, how ?]}
+The LISA response described in Sec.~\ref{sec:LISA} has been implemented in the LISA Data Challenge~\cite{LDC} software suite.
 
@@
 
-\jgb{This work was supported by a NASA Astrophysics Theory Program grant 11-ATP-046.}
+We are indebted to Alejandro Boh\'{e}, Alessandra Buonanno, Scott Field, Philip Graff, Sebastian Khan, Antoine Klein, Serguei Ossokine, Michael P\"{u}rrer for useful discussions and comments.
+We used the software GWFrames~\cite{Boyle+14} in our study.
+This work was supported by a NASA Astrophysics Theory Program grant 11-ATP-046.
 
@@
 
-In the following, we will drop the mode indices and the sum over modes, and we will denote generically by $h$ an individual mode of the waveform\footnote{This notation for a generic individual mode is not to be confused with the often encountered notation $h\equiv h_{+} - i h_{\times}$.}.
+%In the following, we will drop the mode indices and the sum over modes, and we will denote generically by $h$ an individual mode of the waveform\footnote{This notation for a generic individual mode is not to be confused with the often encountered notation $h\equiv h_{+} - i h_{\times}$.}.
 Note that, if we are considering real signals $s(t)$ (such as $h_{+}$, $h_{\times}$, or any detector output), we can focus  on the positive-frequencies range and reconstruct the negative frequencies as $\tilde{s}(-f) = \tilde{s}(f)^{*}$.
 For positive frequencies $f>0$, we will use the generic notation $h$ \SM{[not using the 'generic notation' anymore]} as
-\begin{align}\label{eq:notationhsignm}
-	h(t) &\equiv h_{\ell m}(t) \,, \quad \tilde{h}(f) \equiv \tilde{h}_{\ell m}(f) \text{ for } m>0 \,, \nn\\
-	h(t) &\equiv h_{\ell m}^{*}(t) \,, \quad \tilde{h}(f) \equiv \tilde{h}_{\ell m}(-f)^{*} \text{ for } m<0 \,.
-\end{align}
-We will further decompose $\tilde{h}$ into a Fourier-domain amplitude $A$ and a phase $\Psi$ according to
+%\begin{align}\label{eq:notationhsignm}
+%	h(t) &\equiv h_{\ell m}(t) \,, \quad \tilde{h}(f) \equiv \tilde{h}_{\ell m}(f) \text{ for } m>0 \,, \nn\\
+%	h(t) &\equiv h_{\ell m}^{*}(t) \,, \quad \tilde{h}(f) \equiv \tilde{h}_{\ell m}(-f)^{*} \text{ for } m<0 \,.
+%\end{align}
+Dropping the mode indices $\ell$, $m$, we will decompose a given mode\footnote{Not to be confused with the commonly used notation $h = h_{+} - i h_{\times}$ for the complex strain.} $\tilde{h}_{\ell m} = \tilde{h}$ into a Fourier-domain amplitude $A$ and a phase $\Psi$ according to
 \be\label{eq:defAPsi}
@@
 \ee
-
-Since we intend to work in the Fourier domain, we will use PhenomD waveforms~\cite{Khan+15}.
+%Since we intend to work in the Fourier domain, we will use PhenomD waveforms~\cite{Khan+15}.
 These phenomenological waveforms provide analytic fits for both the Fourier-domain amplitude and phase of the 22 mode, with fit coefficients that were calibrated on EOB/NR hybrids \SM{[describe smoothing across intervals ?]}.
@@
 \SM{[remove that and point to appendix on interpolation errors]}
-
 %\begin{figure*}
@@
 %\end{figure*}
-
-We recall that, with our sign convention, the effect of a positive shift in time $h' (t) = h(t- \Delta t)$ is
-\be
-	\tilde{h}'(f) = e^{2i\pi f \Delta t} \tilde{h}(f) \,.
-\ee
-If we also allow for a global shift in phase of the signal, which corresponds to a rotation of the unit vector pointing towards the observer in the frame of the gravitational waves source, a constant term can also be added (note that the effect of such a rotation gives different phase changes $e^{\pm i m \delta \phi}$ for the different modes of the radiation).
+%We recall that, with our sign convention, the effect of a positive shift in time $h' (t) = h(t- \Delta t)$ is
+%\be
+%	\tilde{h}'(f) = e^{2i\pi f \Delta t} \tilde{h}(f) \,.
+%\ee
+%If we also allow for a global shift in phase of the signal, which corresponds to a rotation of the unit vector pointing towards the observer in the frame of the gravitational waves source, a constant term can also be added (note that the effect of such a rotation gives different phase changes $e^{\pm i m \delta \phi}$ for the different modes of the radiation).
 We see that both a constant and a linear term in the Fourier-domain phase can be adjusted by changing extrinsic parameters.
@@
 
-To describe previous approaches to the problem of producing precessing waveforms directly in the Fourier domain, avoiding the use of a time-domain generation followed by a FFT \SM{[check acronym]}, it is useful to introduce schematically four different approximations (dropping the mode indices):
+To describe previous approaches to the problem of producing precessing waveforms directly in the Fourier domain, avoiding the use of a time-domain generation followed by an FFT, it is useful to introduce schematically four different approximations (dropping the mode indices):
 \begin{itemize}
@@
 In practice, one finds however that the extended frequency-based expressions~\eqref{eq:precPhenomP} are mildly varying when extrapolating to higher frequencies, and this approach has been validated by comparisons to numerical relativity waveforms~\cite{Hannam+13}.
-\SM{[refer to PhenomPv3 in prep.]}
 

### 2018-06-17 docs/paper_method.tex
Edits to the paper.

@@
 \usepackage{tabularx}
+\usepackage{braket}
 
@@
 Blue corresponds to the orbital delay comb coefficients $c_{n}^{0}$ given by~\eqref{eq:cn0}, and red to the constellation comb $c_{n}^{L}$ for the response~\eqref{eq:GslrL}.
-The result is shown after averaging over 100 random orientation angles.}
+The result is shown after averaging over 100 random orientation angles.
+Most of the slowly-chirping systems considered here will have $f \lesssim 10^{-2}\Hz$.}
   \label{fig:lisacombextent}
@@
 The top panel shows the real and imaginary part of the transfer function.}
-  \label{fig:}
+  \label{fig:LISAerrorSOBHepsPsi1const}
 \end{figure}
@@
 The top panel shows the real and imaginary part of the transfer function.}
-  \label{fig:}
+  \label{fig:LISAerrorSOBHepsPsi10const}
 \end{figure}
@@
 
-First, we should mention that these signals share similarities with the galactic binaries that will provide numerous quasi-monochromatic signals in the LISA band~\cite{LISA17}.
-They are far from merger, slowly chirping, and span only a narrow frequency band over the course of the LISA mission lifetime.
-For the galactic binaries, an accurate and very efficient numerical treatment of the response has been proposed in~\cite{} \SM{[C]} and widely used in applications (see e.g.~\cite{} \SM{[C]}).
+First, we should mention that these signals share similarities with the galactic binaries that will provide numerous quasi-monochromatic signals in the LISA band~\cite{Nelemans+01, LISA17}.
+They are far from merger, slowly chirping, and span only a narrow frequency band over the course of the LISA mission lifetime, although being less monochromatic that galactic binaries.
+For the galactic binaries, an accurate and efficient numerical treatment of the response has been proposed~\cite{CL07} and widely used in applications.
 This treatment is referred to as the fast-slow decomposition, or heterodyning approach.
@@
 The efficiency of this approach is contingent to the smallness of $\eta$.
-For galactic binaries, this quantity is typically \SM{[complete]}.
-For SOBH systems, however, $\eta$ will take a continuous set of values from roughly $10^{-4}$ to $1$ \SM{[check values]}.
-\SM{[add sentence: cover the intermediate ground]} Although the systems for which $\eta$ is the largest are also the easiest to treat with the perturbative formalism of Sec~\ref{sec:formalism}, we propose here yet a third method for dealing with potentially troublesome systems.
-We leave for the future a more detailed study of SOBH systems as a population, and the investigation of the precise boundaries and overlap areas of the three methods as well as their respective computational costs.
-
-The third approach we propose here exploits the fact that, in the LISA case, the modulations and delays entering~\eqref{eq:yslr} are periodic, with a period of one year and a frequency $f_{0} = \Omega_{0}/2\pi = 1/\mathrm{yr} \simeq 3.2\times 10^{-8}\mathrm{Hz}$.
+Galactic binaries are extremely close to monochromatic~\cite{Nelemans+01}, and this quantity can be as low as $10^{-6}-10^{-7}$.
+For SOBH systems, however, $\eta$ will take a continuous set of values from roughly $10^{-4}$ to $5$.
+Although the systems for which $\eta$ is the largest are also the easiest to treat with the perturbative formalism of Sec~\ref{sec:formalism}, we propose here yet a third method, based on a discrete Fourier comb, making sure we can cover the intermediate ground of slowly-chirping systems with a large $\eta$.
+We leave for the future a more detailed study of SOBH systems as a population, and the investigation of the precise boundaries and overlap areas of the three methods (heterodyning, Fourier comb, perturbative) as well as their respective computational costs.
+
+The Fourier comb approach we propose here exploits the fact that, in the LISA case, the modulations and delays entering~\eqref{eq:yslr} are periodic, with a period of one year and a frequency $f_{0} = \Omega_{0}/2\pi = 1/\mathrm{yr} \simeq 3.2\times 10^{-8}\mathrm{Hz}$.
 For any given frequency $f$, $G(f,t)$ is periodic in time, so that~\eqref{eq:defG} becomes a discrete Fourier series
@@
 \ee
-
+For slowly-chirping systems, the orbital-delay part of the response gives a transfer function that has significant structure, by contrast with fast-chirping systems where it reduces essentially to an extra phase contribution.
+Thus, in practice this approach is to be applied to $G(f,t)$ representing the full response, orbital delay and constellation modulation.
+
+For illustration purposes, however, we will keep the orbital and constellation response separated in the following.
 In the case of the orbital delay, with $G_{0}(f,t)$ given by~\eqref{eq:G0} and~\eqref{eq:delay0}, the particularly simple expression of the delay gives an analytic expression for the coefficients $c_{n}^{0}$ in terms of Bessel functions of the first kind, as
@@
 \ee
-By contrast, for the constellation response, to our knowledge the coefficients $c_{n}^{L}$ do not admit such a simple close-form expression, and they must be computed numerically using the integrand given by~\eqref{eq:GslrL}.
+By contrast, for the constellation response (or for the full response), to our knowledge the coefficients $c_{n}^{L}$ do not admit such a simple close-form expression, and they must be computed numerically using the integrand given by~\eqref{eq:GslrL}.
 Truncating~\eqref{eq:Gdiscretefourier} to a finite order $N$, this computation reduces to an FFT and the $c_{n}$ coefficients are given by~\eqref{eq:ykDFT} and~\eqref{eq:ckyk}.
@@
 Thus, computing the Fourier-domain response now requires to convolve the signal with a discrete comb with frequency-dependent coefficients $c_{n}$.
-In practice, this sum is to be truncated at $|n|\leq N$, for some finite order $N$ determining the accuracy of the approximation.
+In practice, this sum is to be truncated at $|n|\leq N_{\rm comb}$, for some finite order $N_{\rm comb}$ determining the accuracy of the approximation.
 To assess the expected truncation error, we use a simple criterion based on the $L^{1}$-norm of the $\{c_{n}\}$ sequence.
-For a given target truncation error $\eta$, we define the truncation \SM{change notation to avoid confusion with $N$ from Fresnel stencil} $N(f, \eta)$ as the smallest integer such that
+For a given target truncation error $\eta$, we define the truncation $N_{\rm comb}(f, \eta)$ as the smallest integer such that
 \be\label{eq:criteriontruncationcomb}
@@
 \ee
-The truncation order $N(f,\eta)$ is frequency-dependent, and also depends on the orientation angles.
-For $c_{n}^{L}$, it also determines the number of function evaluations and the cost of the FFT.
+The truncation order $N_{\rm comb}(f,\eta)$ is frequency-dependent, and also depends on the orientation angles.
 For $c_{n}^{0}$, the expression~\eqref{eq:cn0} can provide an asymptotic bound on the width of the comb.
-For large values of $n$ we have indeed the equivalent $| J_{n}[z] | \sim (e z/2n)^{n}/\sqrt{2\pi n} \; \left[ n\rightarrow + \infty \right]$~\cite{DLMF} \SM{[C: dlmf?]}, so that a conservative estimate for the truncation order (almost independent of $\eta$) is given by $N \geq  |e \pi f R \cos \beta|$.
-Fig.~\ref{fig:lisacombextent} shows numerical computations for $N(f,\eta)$, averaged over orientation angles, for both the constellation and the orbital response and for the truncation levels  $\eta=10^{-12}$, $10^{-6}$, $10^{-3}$.
+For large values of $n$ we have indeed the equivalent $| J_{n}[z] | \sim (e z/2n)^{n}/\sqrt{2\pi n} \; \left[ n\rightarrow + \infty \right]$ (see (10.19) in~\cite{DLMF}), so that a conservative estimate for the truncation order (almost independent of $\eta$) is given by $N_{\rm comb} \geq  |e \pi f R \cos \beta|$.
+Fig.~\ref{fig:lisacombextent} shows numerical computations for $N_{\rm comb}(f,\eta)$, averaged over orientation angles, for both the constellation and the orbital response and for the truncation levels  $\eta=10^{-12}$, $10^{-6}$, $10^{-3}$.
 We see that, especially at high frequencies, the orbital response requires more coefficients than the constellation response due to the longer baseline of the delay.
-At the frequency $f=0.1\Hz$, we have for the orbital response $N \simeq 200$.
+The majority of systems for which we wish to apply the comb method will have frequencies $\lesssim 10^{-2}\Hz$, where for the orbital response $N_{\rm comb} \simeq 40$.
 
 An important point is that, since the signals from slowly-chirping binaries extend on a narrow frequency band, for both responses the frequency-dependence of $c_{n}^{0}(f)$, $c_{n}^{L}(f)$ will be very mild, so that the coefficients can be computed at two or three frequencies and interpolated in-between.
-The computational cost of this approach is set by the truncation order $N$, and the number of intermediate frequencies where the transfer function resulting from the sum~\eqref{eq:transferdiscreteconvolution} must be computed for later interpolation, with the necessary waveform evaluations \SM{[possibly refer to appendix on interpolation errors here]}.
-We leave for future work a proper assessment of the computational performance of this method compared to the others \SM{[say a bit more ?]}.
-
-Although straightforward and generic, this writing of the Fourier-domain response function will be in general more expensive to compute than the perturbative transfer function approach of Sec.~\ref{sec:formalism}.
-
-Note that for slowly-chirping binaries, the large features in the orbital transfer functions (see Fig.~\ref{} \SM{[R]} and Fig.~\ref{} \SM{[R]}) start to dominate signal-dependent derivatives.
-For the clarity of the presentation, we still do two separate comparisons for the orbital response~\eqref{} \SM{[R]} and the constellation response~\eqref{} \SM{[R]}, while the actual treatment would require to process the signal through the two steps of the response at once, instead of sequentially.
-This would result in replacing the discrete comb~\eqref{eq:defcn} by a double sum, however the extent of the constellation $c_{n}$ is rather limited in $n$.
+The computational cost of this approach is then set by the convolution~\eqref{eq:transferdiscreteconvolution} that must be evaluated at as many frequencies as necessary to interpolate the transfer functions as functions of frequency (for examples, see Figs.~\ref{fig:LISAerrorSOBHepsPsi1d0}, \ref{fig:LISAerrorSOBHepsPsi10d0}, \ref{fig:LISAerrorSOBHepsPsi1const} and~\ref{fig:LISAerrorSOBHepsPsi10const}).
+In general, this Fourier comb approach will be more expensive than the perturbative response of Sec.~\ref{subsec:executivesummary}.
+We leave for future work a proper assessment of the computational performance of this method compared to the others.
+
+We now illustrate this approach by considering two examples of equal-mass SOBH systems.
+The first has $M = 50\Msol$ and is $\Delta t = 200\yr$ away from merger, and for the orbital response $\epsilon_{\Psi 2} \simeq 1$.
+The second has $M = 15\Msol$ and is $\Delta t = 1500\yr$ away from merger, and for the orbital response $\epsilon_{\Psi 2} \simeq 7$.
+The phase error measures for the constellation part are smaller, $\epsilon_{\Psi_{2}} \simeq 0.1$ and $0.01$ respectively.
+In keeping with the previous section, in our presentation we separate the orbital and the constellation response, keeping in mind that in practice the full response is to be handled in one step.
+
+The transfer functions for the orbital response and residual errors are shown for the two systems in Figs.~\ref{fig:LISAerrorSOBHepsPsi1d0} and~\ref{fig:LISAerrorSOBHepsPsi10d0}.
+We see that, contrarily to Sec.~\ref{subsec:errorsLISA}, the transfer function starts to develop more structure that just a phase contribution.
+The case $M=50\Msol$ with $\epsilon_{\Psi 2} \simeq 1$ shows that the leading-order treatment leads to relative errors of order $10\%$, while increasing $N$ to 5 or 20 brings the errors back to $\sim 10^{-4}$, with a marginal improvement from further amplitude and delay corrections.
+The case $M=50\Msol$ with $\epsilon_{\Psi 2} \simeq 7$, by contrast, is clearly outside the range of applicability of the perturbative formalism, as orders of approximation give errors of order $100\%$.
+In both cases, the comb treatment performs well, at better than $10^{-4}$.
+
+For the constellation response, the transfer functions and residual errors are shown for the two systems in Figs.~\ref{fig:LISAerrorSOBHepsPsi1const} and~\ref{fig:LISAerrorSOBHepsPsi10const}.
+Here, with $\epsilon_{\Psi 2} \sim 0.01$ and $0.1$, both systems are within reach of the perturbative formalism, although the case  $\epsilon_{\Psi 2} \sim 0.1$ shows errors of $10\%$ at leading order and requires a rather large $N=20$ to reach errors below $10^{-3}$.
+The comb treatment yields again errors below $10^{-4}$ in both cases.
+
+Thus, we have shown that some slowly-chirping systems will be out of reach of the perturbative treatment of Sec.~\ref{subsec:executivesummary}, and we have demonstrated that the Fourier comb approach presented here could be applied to these systems with a good accuracy.
 
@@
 The NR and SEOB post-merger frames disagree mainly due to a different value of the final spin $\chi_{f}$\footnote{To compute $\chi_{f}$, SEOBNRv3 uses fitting formulas from~\cite{BR09}.
-These formulas will be updated to more recent fits incorporating additional NR data (see e.g.~\cite{HBR16}) in the next version of the SEOB code~\cite{Ossokine+18}, which is expected to reduce or remove this disagreement.
-\SM{[update this reference in prep.]}}, which leads to a different $\Omega_{\rm frame}$.
+Using updated formulas incorporating additional NR data (see e.g.~\cite{HBR16}) would reduce or remove this disagreement.}, which leads to a different $\Omega_{\rm frame}$.
 Given their different value of $\chi_{f}$, however, both agree with the qualitative description~\eqref{eq:OmegaframeQNM}.
@@
 
-\jgb{In this paper, we will need this model only as an example for demonstrating our frequency-domain precession response, but this or a similar approach may be also useful in borader gravitational wave analysis applications.
-In that context,} \SM{[make clear: goal is illustration]}
-we would now like to underline a number of caveats and limitations of our simplified model.
+In this paper, we will need this model only as an example for demonstrating our frequency-domain precession response, and we only ask that it represents the qualitative features of a precessing waveform.
+In that context, we would like to underline a number of caveats and limitations of our simplified model.
 First, it is clear from Fig.~\ref{fig:precmodel} that the acceleration of the precession post-merger is going to be the most challenging feature for the separation of timescales at the basis of Sec.~\ref{sec:formalism}.
@@
 This example therefore serves the purpose of lying at the edge of validity of the picture of standard precession along a cone around an almost-fixed direction, and the standard hierarchy between mode contributions is not valid in this case.
-\SM{[soundness of trajectories, particularly in this case, are to be checked again]}
+\SM{[soundness of trajectories, particularly in this case, are to be checked again -- discrepancy for $--$ between Jf from dynamics and chif used in SEOB internals]}
 
@@
 %\end{figure*}
-
-\SM{[Check the mismatch for $--$ between Jf from dynamics and chif used in SEOB internals to do the ringdown attachment - check that the frame trajectory makes sense by comparing to orbital plane.]}
 
@@
 
-\SM{[We could add plots of the 3 frame trajectories, and of the FD transfer functions in Amp/Phase form.
-But the trajectories are not very informative except for - - (where our model might not be very solid) and for showing oscillations in $\beta$.
-The reconstruction errors plots already show the amplitude of the modes themselves, giving an idea of the hierarchy; what is the missing is the phase in Tlmmp, which is not very informative.]}
-
 %\begin{figure*}
@@
   \includegraphics[width=.98\linewidth]{plots/postmergeromega_py.pdf}
-  \caption{\SM{[label y axis]} Quasi-normal mode frequencies for the 0th overtone of the modes $22$ and $21$, as a function of the dimensionless spin of the final black hole $\chi_{f}$, as well as their difference $\Omega_{\rm frame}$ (see~\eqref{eq:OmegaframeQNM}).
+  \caption{Quasi-normal mode frequencies for the 0th overtone of the modes $22$ and $21$, as a function of the dimensionless spin of the final black hole $\chi_{f}$, as well as their difference $\Omega_{\rm frame}$ (see~\eqref{eq:OmegaframeQNM}).
 The lower panel shows the  ratio $\Omega_{\rm frame} / \omega^{\rm QNM}_{220}$ characteristic of the separation of timescales between the phase of the modulation~\eqref{eq:wignerphasesimpleprec} and the P-frame waveform phase.}
@@
 \end{figure}
-
-%We show in Fig~\ref{} the frequencies $\omega_{220}^{\rm QNM}$, $\omega_{210}^{\rm QNM}$ and their difference $\Omega_{\rm frame}$.
-The behaviour for near-to-extremal spin of the remnant would require more exploration, as the QNM frequencies vary a lot when approaching extremality.
-
-\SM{[justify somewhere smallness of cubic-in-phase term]}
-
-\SM{[write somewhere that we require mainly FD smoothness]}
 
@@
 
-\SM{[Describe the difference in morphology of the precession in the three cases ?]}
 We can somewhat complement this picture for the post-merger precession by considering the accelerated frame rotation rate described by the model~\eqref{eq:OmegaframeQNM} and illustrated in Fig.~\ref{fig:precmodel}.
@@
 We are indebted to Alejandro Boh\'{e}, Alessandra Buonanno, Scott Field, Philip Graff, Sebastian Khan, Antoine Klein, Serguei Ossokine, Michael P\"{u}rrer for useful discussions and comments.
-We used the software GWFrames~\cite{Boyle+14} in our study.
+We used the software GWFrames~\cite{Boyle13, Boyle+14} in our study.
 This work was supported by a NASA Astrophysics Theory Program grant 11-ATP-046.
@@
 
-In this Appendix, we summarize our conventions for the Wigner matrices.
-
-\SM{[add O'Shaughnessy frame extraction]}
-\SM{[cite other prescriptions]} \SM{[C]}
-
-Several constructions for the radiation axis of the wa precessing frame have been proposed~\cite{} \SM{[C]}.
-
-\SM{[This differs from G. Faye's definition by a transposition $m\leftrightarrow m'$]}
+In this Appendix, we summarize our conventions for the Wigner matrices and give a brief description of the construction of a precessing-frame directly from the waveform.
+
+If the $h_{\ell m}^{\rm I}$ are the spin-weighted spherical harmonics~\eqref{eq:defmodes} of the waveform in a fixed inertial frame (I), and if the $h_{\ell m}^{\rm P}$ are the modes of the waveform in a time-dependent precessing frame (P) constructed from the inertial frame by an active rotation with Euler angles $(\alpha, \beta, \gamma)$ (in the convention $(z,y,z)$), then the modes are related by
+\begin{subequations}
+\label{eq:wignerrotApp}
+\begin{align}
+	h_{\ell m}^{\rm I} = \sum\limits_{m=-\ell}^{\ell} \calD^{\ell *}_{mm'} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm P} \,, \\
+	h_{\ell m}^{\rm P} = \sum\limits_{m=-\ell}^{\ell} \calD^{\ell }_{m'm} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm I} \,.
+\end{align}
+\end{subequations}
+Here we introduced Wigner $\calD$-matrices
 \be\label{eq:defWignerDapp}
@@
 \ee
-\SM{[check]}
+with the real-valued Wigner $d$-matrix reading
 \begin{widetext}
@@
 where the boundaries of the sum, $k_{\rm min} = \mathrm{max}(0, m-m')$ and $k_{\rm max} = \mathrm{min}(\ell+m, \ell-m')$, can also be read by enforcing that the arguments of the factorials must be non-negative.
-
-\SM{[add here explicitly the dominant-eigenvector construction of the P-frame]}
+Note that our convention differs from the convention of~\cite{ABFO09} by a transposition,
+\be
+	\calD^{\ell}_{m m'} (\alpha, \beta, \gamma) = D^{\ell \, \mathrm{ABFO}}_{m' m} (\alpha, \beta, \gamma) \,.
+\ee
+In Sec.~\ref{sec:precession}, we considered only $\ell = 2$ and we restricted ourselves to the contributions of the precessing-frame mode $h_{22}^{\rm P}$.
+In that case, the relevant explicit expression for the $d$-matrix are
+\begin{subequations}
+\begin{align}
+	d^{2}_{22} (\beta) &= \cos^{4} \frac{\beta}{2}\,, \\
+	d^{2}_{12} (\beta) &= 2 \cos^{3} \frac{\beta}{2} \sin \frac{\beta}{2}\,, \\
+	d^{2}_{02} (\beta) &= \sqrt{6} \cos^{2} \frac{\beta}{2} \sin^{2} \frac{\beta}{2} \,, \\
+	d^{2}_{-12} (\beta) &= 2 \cos \frac{\beta}{2} \sin^{3} \frac{\beta}{2} \,, \\
+	d^{2}_{-22} (\beta) &= \sin^{4} \frac{\beta}{2} \,.
+\end{align}
+\end{subequations}
+
+Different prescriptions have been proposed to construct a precessing frame.
+Intuitively, the frame follows the plane of the orbit as the spins force the latter to precess, so one could define the precessing frame from the trajectory during the inspiral.
+This is in fact what is done for SEOBNRv3~\cite{Pan+13}, where a non-precessing waveform is generated in the orbital plane before being rotated back to an inertial frame.
+Alternatively, the precessing frame can be directly read off the waveform, an approach that is less coordinate-dependent and that can readily be extended through merger where there is no notion of an orbital plane.
+Proposals in this sense include a maximization of the modes $22$ and $2,-2$~\cite{Schmidt+10}, and the construction of an angular velocity for the waveform~\cite{Boyle13}.
+
+Here, we use the dominant eigenvector prescription of~\cite{OShaughnessy+11}, to which we refer (as well as~\cite{Boyle13}) for more details. One defines the matrix
+\be
+	\langle \bm{L} \bm{L} \rangle_{ab} \equiv \frac{\sum_{\ell, m, m'} h^{*}_{\ell m'} \bra{\ell, m'} L_{a} L_{b} \ket{\ell, m} h_{\ell m} }{ \sum_{\ell, m, m'} h^{*}_{\ell m'} h_{\ell m} } \,,
+\ee
+where the $\ket{\ell, m}$ are a notation for the spin-weighted spherical harmonics ${}_{-2}Y_{\ell m}$, the indices $a,b$ are ordinary spatial incides, and $L_{a}$ stands for the angular-momentum operator, acting on the bras and kets $\ket{\ell, m}$ in the usual fashion.
+Defining $L_{\pm} = L_{x} \pm i L_{y}$,
+\begin{subequations}
+\begin{align}
+	L_{+} \ket{\ell, m} &= \begin{cases}\sqrt{\ell (\ell+1) - m (m+1)} \ket{\ell, m+1} \;\text{if} \; m<\ell, \\ \;\; 0 \;\;\text{otherwise} \end{cases} \\
+	L_{-} \ket{\ell, m} &= \begin{cases}\sqrt{\ell (\ell+1) - m (m-1)} \ket{\ell, m-1} \; \text{if} \; m>-\ell, \\ \;\; 0 \;\;\text{otherwise} \end{cases} \\
+	L_{z} \ket{\ell, m} &= m \ket{\ell, m} \,.
+\end{align}
+\end{subequations}
+Note that in the original proposal~\cite{OShaughnessy+11}, $\psi_{4}$ was used instead of the strain.
+The $z$-axis of the precessing frame is then chosen to be the dominant eigenvetor of the matrix $\langle \bm{L} \bm{L} \rangle_{ab}$.
+The construction of the precessing frame is then supplemented by the minimal rotation condition~\cite{Boyle+11}, see~\eqref{eq:gammadot}.
 

### 2018-06-18 docs/paper_method.tex
Edits to the paper.

@@
 We further present an alternative direct convolution approach to accurately represent these post-merger features.
+\SM{[update to include SOBHs]}
 
@@
 
-\jgb{With the unprecedented recent gravitational-wave detections of coalescencing binary black holes and binary neutron stars, announced by the LIGO-Virgo collaboration~\cite{LIGO-theevent-2016,LIGO-christmasevent-2016,LIGO-O1BBH-2016}[update refs], gravitational-wave astronomy has entered its observational era.
-As LIGO prepares for even more sensitive observation runs, and with the recent expansion of the ground-based detectors network with Virgo~\cite{Virgo} ( and eventually also KAGRA~\cite{KAGRA} and LIGO-India~\cite{INDIGO}), observations of such compact object coalescences are expected at ever-increasing rate, with binary black holes likely to be the dominant sources~\cite{} \SM{[C]}.}
-
-  \jgb{Moreover, the European Space Agency has recently selected the Laser Interferometer Space Antenna (LISA)\cite{LISA17} to realize the ``Gravitational Universe'' science theme\cite{GravitationalUniverse2013} as the 3rd large space mission of its Cosmic Vision program, with a tentative launch around 2034.
+With the unprecedented recent gravitational-wave detections of coalescencing binary black holes and binary neutron stars, announced by the LIGO-Virgo collaboration~\cite{LIGO-theevent-2016, LIGO-christmasevent-2016, LIGO-O1BBH-2016, LIGO-BNSevent-2017}, gravitational-wave astronomy has entered its observational era.
+As LIGO prepares for even more sensitive observation runs, and with the recent expansion of the ground-based detectors network with Virgo~\cite{Virgo} (and eventually also KAGRA~\cite{KAGRA} and LIGO-India~\cite{INDIGO}), observations of such compact object coalescences are expected at ever-increasing rate.
+
+Moreover, the European Space Agency has recently selected the Laser Interferometer Space Antenna (LISA)\cite{LISA17} to realize the ``Gravitational Universe'' science theme\cite{elisa13} as the 3rd large space mission of its Cosmic Vision program, with a tentative launch around 2034.
 LISA will be able to detect and characterize, among several important gravitational wave source targets, comparable-mass binary black hole coalescences from cosmological distances over a wide range of masses.
-These will range from high-redshift observations of supermassive black hole binaries with $M\sim 10^{7} \Msol$ down to the observation of LIGO-type sources with $M\sim 10^{2} \Msol$~\cite{Sesana16}.}
+These will range from high-redshift observations of supermassive black hole binaries with $M\sim 10^{7} \Msol$ down to the observation of LIGO-type sources with $M\sim 10^{2} \Msol$~\cite{Sesana16}.
 
@@
 Bayesian analysis for parameter estimation of gravitational-wave signals, as was performed for the LIGO detections~\cite{LIGO-theeventPE-2016,LIGO-O1BBH-2016}, may require millions of evaluations of the likelihood function to sample the posterior probability distribution.
-\SM{[objective: accuracy, not necessary necessary for LIGO now but in the future yes]} The greater sensitivity of LISA and other future instruments will require further increases in the accuracy and computational efficiency of signal templates.
-
-  %This motivates an important and ongoing effort to improve gravitational wave templates.
-Two important focus areas for template improvements are need to cover all phases of the evolution of the system, including the inspiral, merger and ringdown (IMR), which became possible thanks to the breakthrough of numerical relativity~\cite{Pretorius05,Baker+06,Campanelli+06}, and the inclusion of the effects of the spins of the compact objects, that can lead to a precession of the orbital plane~\cite{ACST94,K95}.
-Besides accuracy, the computational efficiency of template generation is crucial.
+The greater sensitivity of LISA and other future instruments will require further increases in the accuracy and computational efficiency of signal templates.
 
@@
 
-\jgb{To motivate our signal treatment we consider challenges of both the LISA response and GW precession applications in terms of a more general problem which covers both domains.
-In each case must determine the Fourier transform a signal $\tilde s(f)$ which results from processing some incident waveform $h(t)$ through some combination of modulations and delays.}
-
-In the general problem, we apply a time-varying delay $d(t)$ to the waveform followed by a multiplicative modulation function $F(t)$,
+In our presentation, we will consider a formal signal processing problem encompassing the challenges posed by both the LISA response and the precession modulations.
+Given a signal $h(t)$, we apply a time-varying delay $d(t)$ to the waveform followed by a multiplicative modulation function $F(t)$,
 \be
@@
 \ee
-We then seek an \textit{efficient} way to compute the Fourier transform $\tilde{s}(f)$, expressed by means of a Fourier-domain transfer function $\calT$ such that
+We then seek an efficient way to compute the Fourier transform $\tilde{s}(f)$, expressed by means of a Fourier-domain transfer function $\calT$ such that
 \be\label{eq:deftransfer}
@@
 
-As is well known, GW signals decomposed in spin-weighted spherical harmonic components are smoothly varying functions of frequency in amplitude/phase form\cite{smoothAmpPhaseRefs}.
+As is well known, GW signals decomposed in spin-weighted spherical harmonic components are smoothly varying functions of frequency in amplitude/phase form.
 An important consequence is that these signal components can be accurately represented in terms of a relatively coarsely sampled frequency grid.
@@
 
-In the case of precessing binaries, the delays will be absent and the modulation functions will be the time-dependent Wigner coefficients applied to rotate the waveform from an \jgb{precessing orbital frame, in which waveform modes exhibit smooth amplitude and phase variation}, to an inertial frame where the observations take place.
+In the case of precessing binaries, the delays will be absent and the modulation functions will be the time-dependent Wigner coefficients applied to rotate the waveform from a precessing orbital frame, in which waveform modes exhibit smooth amplitude and phase variation, to an inertial frame where the observations take place.
 In the case of a LISA-type detector, the signal $h(t)$ will simply be the waveform in a fixed heliocentric frame, and the delays will come from the motion of each detector against the wave front, while the modulation will represent the time-variation in the detector orientation.
@@
 Furthermore, a natural quantitative criterion for the applicability of this idea is given by the comparison of the radiation-reaction timescale of the signal with the modulation timescale.
-In other words, the change in frequency of the signal over a characeteristic time of the modulation should be small for the separation of timescales to work.
+In other words, the change in frequency of the signal over a characteristic time of the modulation should be large for the separation of timescales to work.
+However, as we will see below, for this problem the dimensional analysis falls short of the full picture: the separation of timescales can be affected by frequency-dependent dimensionless factors in presence of delays.
 
@@
 We will use a rigid instantaneous model, approximating the geometry of the constellation by a moving equilateral triangle (neglecting the flexing of the arms induced by corrections in the orbits), and evaluating all geometric factors in~\eqref{eq:yslr} at a single time $t$ (neglecting point-ahead corrections).
-The additional delay $L$ (we use $c=1$) in the first term represents the light propagation time along the arm between spacecraft $s$ and $l$.
+The additional delay $L$ (we use $c=1$) in the first term represents the light propagation time along the arm between spacecraft $s$ and $r$.
 
@@
 Since in our rigid approximation these delays will take a simple form in Fourier domain, we will focus in this paper on the $y_{slr}$ observables.
-Sec.~\ref{subsec:modelLISA} provides more details on the response of LISA-type detectors, on the TDI observables, and on the approximations that enter the derivation of the basic response~\eqref{eq:yslr} above.
+Sec.~\ref{subsec:modelLISA} provides more details on the response of LISA-type detectors and on the approximations that enter the derivation of the basic response~\eqref{eq:yslr} above.
 
@@
 When neglecting small non-periodic orbital perturbations (such as those caused by the influence of the Earth and of the other planets on the orbits), both $F(t)$ and $d(t)$ are periodic.
-It will also be useful to separate the delays in two types of terms: the first, $d_{0} = -\hatk\cdot p_{0}$, relates the waveform at the SSB to the waveform at the center of the constellation, whereas the second, $d_{c}$, represents the various delays between the spacecraft of the constellation.
-\jgb{Concretely, we assume the basic reference design parameters of the LISA mission\cite{LISA17} recently selected by the European Space Agency(ESA), with orbit radius $R=1\,\mathrm{au}$ and armlength $L=2.5\times10^{6}\mathrm{km}$, and with $d_{0} \sim R/c \simeq 500\mathrm{s}$ and $d_{c} \sim L/c \simeq 8\mathrm{s}$.}
+It will also be useful to separate the delays in two types of terms: the first, $d_{0} = -\hatk\cdot p_{0}$, relates the waveform at the SSB to the waveform at the center of the constellation, whereas the second, $d_{L}$, represents the various delays between the spacecraft of the constellation.
+Concretely, we assume the basic reference design parameters of the LISA mission\cite{LISA17} recently selected by the European Space Agency(ESA), with orbit radius $R=1\,\mathrm{au}$ and armlength $L=2.5\times10^{6}\mathrm{km}$, and with $d_{0} \sim R/c \simeq 500\mathrm{s}$ and $d_{L} \sim L/c \simeq 8\mathrm{s}$.
 
@@
 
-For spinning compact objects, \jgb{angular momentum interactions typically} lead to the precession of the orbit~\cite{Apostolatos+94, Kidder95}, which can have large effects on the waveform.
+For spinning compact objects, angular momentum interactions typically lead to the precession of the orbit~\cite{Apostolatos+94, Kidder95}, which can have large effects on the waveform.
 In particular, it breaks the planar symmetry of the gravitational wave emission, causing modulations that are especially important for systems that are observed edge-on.
@@
 \ee
-where the $\calD^{\ell}_{mm'}$ are Wigner matrices~\cite{} \SM{[C]}. Notice that there is no mixing of the modes with different values of $\ell$.
+where the $\calD^{\ell}_{mm'}$ are Wigner matrices (see App.~\ref{app:wigner}).
 
@@
 where the modulation function $F(t)$ is given by a Wigner matrix and depends on time through the Euler angles $(\alpha, \beta, \gamma)(t)$.
-We will discuss more in Sec.~\ref{subsec:modelPrec} \SM{[R]} the limitations of the assumptions used to perform the decomposition above.
 
 In Eq.~\eqref{eq:defmodulationprec} above, the modulation function $F$ has time variations on the precessional timescale, which evolves throughout the inspiral.
-The separation between the precessional timescale and the orbital timescale will be crucial for our analysis.
-In the limit of low frequencies, we will see in Sec.~\ref{} \SM{[R]} below that, although these timescales become more and more separated, the decrease in the chirping rate gives raise to a corrective contribution that does not vanish in this limit.
-In the other limit, as the system gets close to merger, the separation of timescales becomes weaker, and we will explore in Sec.~\ref{} \SM{[R]} the application of our formalism to the merger and ringdown phase.
+In the limit of low frequencies, we will see in Sec.~\ref{sec:precession} below that, although the precession and orbital timescales become more and more separated, the decrease in the chirping rate gives raise to a corrective contribution that does not vanish in this limit.
+We will also explore  the application of our formalism to a precessing-frame decomposition of the waveform extended through the merger and ringdown phase.
 
@@
 The approximation then applies to signals verifying the conditions\footnote{Note that the third condition is not always written explicitly, in particular not in Refs.~\cite{FC93, CF94}.
-It becomes important if we generalize $a$ to be an envelope function incorporating a modulation.} \SM{[say that the third is sometimes missing ?]} \SM{[add reference to Finn-Chernoff, Cutler-Flanagan where the third one is missing]}
+It becomes important if we generalize $a$ to be an envelope function incorporating a modulation.}
 \be\label{eq:conditionsSPA}
@@
 The usual approach is simply to replace any time dependencies appearing the transfer function using \eqref{eq:deftfSPA}.
-\jgb{We will reference the SPA treatment, as a familiar touchstone, as we develop a more general formalism.}
+We will reference the SPA treatment, as a familiar touchstone, as we develop a more general formalism.
 
@@
 
-A direct computation of the generalized convolution~\eqref{eq:FDkernel} using~\eqref{eq:defG} will generally be computationally demanding, but we can exploit the separation of timescales in the problem to seek an accurate but efficient approximation.
-\jgb{We will explore approximations which allow the integral in~\eqref{eq:FDkernel} to be computed, so that the signal may be expressed the form of~\eqref{eq:deftransfer}, with the modulation and delay reduced to a multiplicative transfer function $\mathcal T(f)$.}
-
-\jgb{For example, under appropriate conditions with slowly varying modulations and delays, we can expect that} the Fourier transform $\tilde{G}(f,f')$ should have a compact support, limited to $f' \in [-f_{\rm max}, f_{\rm max}]$ with $f_{\rm max}$ a maximal frequency for the modulation, roughly the inverse of its characteristic timescale.
+A direct computation of the generalized convolution~\eqref{eq:FDkernel} using~\eqref{eq:defG} will generally be computationally demanding, but we can exploit the separation of timescales in the problem to seek an accurate but efficient approximation and to compute the transfer function $\calT(f)$ as in~\eqref{eq:deftransfer}.
+
+For example, under appropriate conditions with slowly varying modulations and delays, we can expect that the Fourier transform $\tilde{G}(f,f')$ should have a compact support, limited to $f' \in [-f_{\rm max}, f_{\rm max}]$ with $f_{\rm max}$ a maximal frequency for the modulation, roughly the inverse of its characteristic timescale.
 This makes the convolution integral~\eqref{eq:FDkernel} localized in frequency, as the waveform is to be evaluated only at frequencies close to $f$.
-We can then approximate $\tilde{h}(f-f')$ by $\tilde{h}(f)$, with a simple Taylor expansion for a perturbative treatment of the difference between the two.
+We can then approximate $\tilde{h}(f-f')$ by $\tilde{h}(f)$, with a simple Taylor expansion of their difference.
 Doing so, we will recover at leading order a locality in time: the response is approximately reduced to an evaluation of the modulation and delay at a representative signal-dependent time $\tf$, which will be equivalent to $\tfSPA$ for inspiral signals.
@@
 Depending on the frequency $f$, this can lead $G(f,t)$ to have faster variations than its nominal timescale ($1\mathrm{yr}$ for LISA).
-As we will see in Sec.~\ref{} \SM{[R]}, \jgb{to derive useful approximations we will not only need to compare modulation timescales, but also to constrain dimensionless factors of the form $2\pi f d$.}
-
-\jgb{In the next subsections we will examine the approximations necessary to allow a local-in-frequency response treatment, and consider low-order corrections which can improve accuracy in cases where the approriate conditions are more marginally realized.}
+Thus, to assess the validity of our approximations we cannot limit ourselves to comparing the dimensionful timescales at play in the problem.
+As we will see in Sec.~\ref{subsec:lisafom}, we must also include in the analysis dimensionless factors of the form $2\pi f d$.
 
@@
 
-\jgb{If the effective width of the kernel function $G(f,f’)$ is not quite negligible compared to the scale of significant variations with $f$ it can be useful to extend our approach beyond the leading-order approximation.
-With the waveform represented in phase-and-amplitude form~\eqref{eq:A9}, the elements of~\eqref{eq:FDkernel} may be Taylor expanded in the variable $f'$: \SM{[change notation, do not use $N$ as sum boundary for $\Psi$]}
+If the width of the kernel function $G(f-f',f')$ is not quite negligible compared to the scale of significant variations of $\tilde{h}(f)$ with $f$, it can be useful to extend our approach beyond the leading-order approximation.
+With the waveform represented in the amplitude and phase form~\eqref{eq:defAPsi}, the elements of~\eqref{eq:FDkernel} may be formally Taylor-expanded in the variable $f'$:
 \begin{subequations}\label{eq:expandfprime}
 \begin{align}
-	\Psi(f-f') &= \Psi(f) + 2\pi f' \tf + \sum\limits_{p\geq 2}^{N} \frac{(-1)^{p}}{p!} {f'}^{p} \frac{\ud^{p} \Psi}{\ud f^{p}} \,, \label{eq:expandPsi}\\
+	\Psi(f-f') &= \Psi(f) + 2\pi f' \tf + \sum\limits_{p\geq 2} \frac{(-1)^{p}}{p!} {f'}^{p} \frac{\ud^{p} \Psi}{\ud f^{p}} \,, \label{eq:expandPsi}\\
 	A(f-f') &= A(f)+A(f) \sum\limits_{q\geq 1} \frac{(-1)^{q}}{q!} {f'}^{q} \frac{1}{A}\frac{\ud^{q} A}{\ud f^{q}} \,, \label{eq:expandA}\\
@@
 \end{subequations}
-using the definition of $t_{f}$ introduced in~\eqref{eq:deftf}.}
-
-\jgb{The leading order transfer function~\eqref{eq:transferlocal} is obtained by leaving off all the terms in the sums.
-In the following, we will consider the resulting transfer functions when keeping some of the next few terms in each of these expansions.}
+using the definition of $t_{f}$ introduced in~\eqref{eq:deftf}.
+
+The leading order transfer function~\eqref{eq:transferlocal} is obtained by leaving off all the terms in the sums.
+In the following, we will consider the resulting transfer functions when keeping some of the next few terms in each of these expansions.
 Notice that we expand $\tilde{G}(f-f',f')$ in $f'$ only in its first argument, and that we can commute the $f$-derivatives of $G$ with the Fourier transform operation.
 
-\jgb{In some cases, we will also expand the exponential of the phase
+We can also formally expand the exponential of the phase corrections $\delta \Psi$ beyond the first two terms in~\eqref{eq:expandPsi} as
 \be
-\exp\left[ -i\Psi(f) \right] \simeq 1+\sum_{j\geq1}\frac{(-i\Psi(f))^j}{j!}
+	\exp\left[ -i\delta\Psi \right] = \sum_{j \geq 0}\frac{(-i\delta\Psi)^j}{j!} \,,
 \label{eq:expandexp}
 \ee
-to obtain a pure $f'$-expansion. The resulting power series in $f'$ can then be recast as a temporal Taylor series}, by applying the formal derivative rule
+to obtain a pure $f'$-expansion. The resulting power series in $f'$ can then be recast as a temporal Taylor series, by applying the formal derivative rule
 \allowdisplaybreaks
@@
 Instead, it will be more instructive to separately consider the different expansions in~\eqref{eq:expandfprime}.
-Later, in Sec.~\ref{subsec:executivesummary} will we come back to combining our results together.
+Later, in Sec.~\ref{subsec:executivesummary} we will come back to combining our results together.
 
@@
 We will find that the third and higher derivatives of the phase are always negligible for our purposes, and we will ignore them.
-Keeping only the first term of the \jgb{sum in~\eqref{eq:expandPsi}, corresponding to the second derivative of $\Psi$, with just the leading terms from \eqref{eq:expandA} and \eqref{eq:expandG}, and expanding the phase exponential \eqref{eq:expandexp} so that the result can be cast as a Taylor series in time} , we obtain straightforwardly
+Keeping only the first term of the sum in~\eqref{eq:expandPsi}, corresponding to the second derivative of $\Psi$, with just the leading terms from \eqref{eq:expandA} and \eqref{eq:expandG}, and expanding the phase exponential \eqref{eq:expandexp} so that the result can be cast as a Taylor series in time, we obtain straightforwardly
 \be\label{eq:resulttaylorPsi}
@@
 
-Similarly, applying the expansion of the Fourier-domain amplitude~\eqref{eq:expandA} while preserving only the leading order terms in \eqref{eq:expandPhi} and \eqref{eq:expandG} gives
+Similarly, applying the expansion of the Fourier-domain amplitude~\eqref{eq:expandA} while preserving only the leading order terms in \eqref{eq:expandPsi} and \eqref{eq:expandG} gives
 \be\label{eq:resulttaylorA}
@@
 \ee
-\jgb{[It seems like it might be better to label this version of the transfer function as ``kernel'' rather than ``delay'' since $G$ here is general. Do we ultimately use this at all?]}
 which interestingly looks like a Taylor expansion of $G$, but this time with joint derivatives in frequency and time.
@@
 
-As will be explained in Sec.~\ref{subsec:executivesummary}, we will also use these formal Taylor expansions to build error measures (constructed as the magnitude of the first term ignored in the series, see~\eqref{eq:deffom}) designed to estimate which of these three types of corrections are important to take into account.
+As will be explained in Sec.~\ref{subsec:executivesummary}, we will also use these formal Taylor expansions to build error measures (constructed as the magnitude of the first term ignored in the series, see~\eqref{eq:deffom}) designed to estimate if these three types of corrections are important to take into account.
 
@@
 
-\jgb{When considering the impact on the transfer function from including the next higher-order phase term, that is the difference between~\eqref{eq:resulttaylorPsi} and~\eqref{eq:transferlocal}, it is natural to define a new timescale, as a function of frequency,
+When considering the impact on the transfer function from including the next higher-order phase term, that is the difference between~\eqref{eq:resulttaylorPsi} and~\eqref{eq:transferlocal}, it is natural to define a new timescale, as a function of frequency,
 \be\label{eq:defTf}
@@
 \ee
-If the correction is small, this timescale should be small compared to the time-scale of variations in the modulation and delay encoded in $G(f,t)$.}
+For the correction to be small, this timescale should be small compared to the time-scale of variations in the kernel $G(f,t)$ encoding the modulation and delay.
 We can use this notation to rewrite~\eqref{eq:resulttaylorPsi} as
@@
 
-We can obtain a straighforward physical interpretation of this timescale $\Tf$ by considering inspiral signals for which the SPA is valid.
-\jgb{In that case
+We can obtain a straighforward physical interpretation of this timescale $\Tf$ by considering inspiral signals for which the SPA is valid. In that case
 \be
@@
 \ee
-Thus, when the SPA applies, $\Tf$ corresponds to} the radiation-reaction timescale \SM{[make the connection to $Q$ used for SEOB comparisons in the Nagar group]}: the shorter this timescale, the faster the binary chirps to higher frequencies on its quasi-circular inspiral.
+Thus, when the SPA applies, $\Tf$ corresponds to the radiation-reaction timescale: the shorter this timescale, the faster the binary chirps to higher frequencies on its quasi-circular inspiral.
 
@@
 
-\jgb{Next consider the impact on the transfer function~\eqref{eq:resulttaylorA} from amplitude corrections beyond leading order.}  This series also leads to the natural introduction of a set of another set of timescales related to the successive derivatives of the amplitude.
+Next, we consider the impact on the transfer function~\eqref{eq:resulttaylorA} from amplitude corrections beyond leading order.
+This series also leads to the natural introduction of a set of another set of timescales related to the successive derivatives of the amplitude.
 In an analogous manner to the definition~\eqref{eq:defTf} of the timescale $\Tf$, we can define
@@
 
-\jgb{By contrast, the impact of higher-order terms in the kernel function on the transfer function}~\eqref{eq:resulttaylordelay} is signal-independent.
+By contrast, the impact of higher-order terms in the kernel function on the transfer function~\eqref{eq:resulttaylordelay} is signal-independent.
 It does not lead to the introduction of new timescales since the coupled time and frequency derivatives are dimensionless.
-We will see however in Sec.~\ref{subsec:lisafom} that treating delays \jgb{does require limiting} dimensionless factors of the type $2\pi f d$.
+We will see however in Sec.~\ref{subsec:lisafom} that treating delays does require taking into account dimensionless factors of the type $2\pi f d$.
 
@@
 For higher harmonics $h_{\ell m}$ with $m\neq 0$, as discussed in Sec.~\ref{subsec:SPA}, $T_{f}$ acquires a factor $m/2$, and $v$ has to be replaced by $(2/m)^{1/3} v$.
-The amplitude of higher harmonics starts at a higher PN order~\cite{BlanchetLiving}, differing from~\eqref{eq:a22N} by a different constant and an additional scaling $v^{\kappa_{\ell m}}$ with $\kappa_{\ell m} = \ell - 2 + (\ell + m \; \mathrm{mod} \; 2)$, which changes the constant in the amplitude timescale as $7/6 \rightarrow (7-2\kappa_{\ell m})/6$ in $T_{A1}^{\rm N}$ and $\sqrt{91}/6 \rightarrow \sqrt{(7-2\kappa_{\ell m})(13-2\kappa_{\ell m})}/6$ in $T_{A2}^{\rm N}$.
+The amplitude of higher harmonics starts at a higher PN order~\cite{BlanchetLiving}, differing from~\eqref{eq:a22N} by a different constant and an additional scaling $v^{\kappa_{\ell m}}$ with $\kappa_{\ell m} = \ell - 2 + (\ell + m \; \mathrm{mod} \; 2)$, which changes the constant in the amplitude timescale as $7/6 \rightarrow (7-2\kappa_{\ell m})/6$ in $T_{A1}^{\rm N}$ and $\sqrt{91}/6 \rightarrow \sqrt{(7-2\kappa_{\ell m})(13-2\kappa_{\ell m})}/6$ in $T_{A2}^{\rm N}$ \SM{[check]}.
 
@@
 
-If we restrict to the case of a pure undelayed modulation, $d=0$ and $G(f,t) = F(t)$, we can relate our result~\eqref{eq:resultdffPsiTf} \jgb{for the transfer function including up to quadratic phase terms, with treatment} of Ref.~\cite{KCY14}, where the authors extend the SPA in a formalism called the Shifted Uniform Asymptotic expansion (SUA) \SM{[look for other reference beyond KCY]}.
-%The derivation includes an expansion of the modulation in Bessel functions, a study of the displacement of the stationary phase point induced by the precession, and a resummation of the result in the form of time derivatives.
+If we restrict to the case of a pure undelayed modulation, $d=0$ and $G(f,t) = F(t)$, we can relate our result~\eqref{eq:resultdffPsiTf} for the transfer function including up to quadratic phase terms, with the treatment of Ref.~\cite{KCY13, KCY14}, where the authors extend the SPA in a formalism called the Shifted Uniform Asymptotic expansion (SUA).
+
 The main intermediate result of~\cite{KCY14}, their equation~(34), reads exactly like~\eqref{eq:resultdffPsiTf} with $\epsilon=1$ and with the identifications $\tilde{H}_{corr}(f)\rightarrow \calT(f)$ for the transfer function, $T\rightarrow \Tf$ for the radiation-reaction timescale and $e^{-i\delta\phi} \rightarrow F$ for the modulation function (which is restricted in their framework to a phase, amplitudes being treated jointly with the amplitude of the signal).
@@
 \ee
+In practice, $N$ will be most often less than $5$, although we will consider in one case $N=10$ and $20$.
 In~\cite{KCY14}, only the case $\epsilon=1$ was considered.
@@
 Here, one simply evaluates the original smooth modulation function at shifted times.
-
-\jgb{In practice, our quadratic-in-phase treatment will apply this formalism with [something about the range of N?]. }
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Relation to the Fresnel transform}
+In the following, we will adopt this implementation for our quadratic-in-phase treatment.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Quadratic phase corrections as an integral transform}
 \label{subsec:fresneltransform}
 
-We can give an alternative interpretation of the \jgb{quadratic-phase} expansion~\eqref{eq:resultdffPsiTf} and of Sec.~\ref{subsec:resumquadphase} in terms of the Fresnel transform.
+We can give an alternative interpretation of the quadratic-phase expansion~\eqref{eq:resultdffPsiTf} and of Sec.~\ref{subsec:resumquadphase} in terms of an integral transform.
 For simplicity of notation, here we keep to the case of a pure modulation $F$ with no delays, $d=0$.
@@
 To obtain~\eqref{eq:resultdffPsiTf}, we expanded the phase exponential using~\eqref{eq:expandexp}.
-\jgb{Without this expansion, though}, we have
+If instead we do not expand this factor, we have
 \be\label{eq:integralquadphase}
@@
 where we recall that $\epsilon = -\mathrm{sgn}(\ud ^{2} \Psi/\ud f^{2})$.
-The integral over $f'$ is a simple complex Gaussian integral \jgb{which can be carried out explicitly.
-The result is a complex Gaussian integral over time which we recognize as a Fresnel transform~\cite{} \SM{[C]} of the function $F$.
-For the Fresnel transform, we introduce the notation}
+The integral over $f'$ is a simple complex Gaussian integral which can be carried out explicitly.
+The result is a complex Gaussian integral over time which is analogous to a Fresnel transform of the function $F$.
+For this Fresnel transform, we introduce the notation
 \be\label{eq:defFresnel}
@@
 
-The Fresnel transform~\cite{} \SM{[C]} is an integral transform that shares some properties with the Fourier transform.
-Contrarily to the latter, it is localized in the sense that the part of the integral that is centered around $t_{0}$ contributes predominantly, due to the cancelling oscillations far from $t_{0}$.
+The Fresnel transform~\eqref{eq:defFresnel} is localized, in the sense that the part of the integral that is centered around $t_{0}$ contributes predominantly, due to the cancelling oscillations far from $t_{0}$.
 The parameter $\tau$ determines how local the transform is.
@@
 Since we have seen that $\Tf$ can be interpreted as the radiation reaction timescale in the SPA regime, this means that a faster-chirping signal ($\Tf$ small) will have a Fresnel transform that is more focused, whereas a slower-chirping signal ($\Tf$ large) will have a Fresnel transform that is more extended.
-\jgb{The Fresnel width must then be compared} with how fast the function $F(t)$ in the integrand is varying.
-In Sec.~\ref{subsec:executivesummary}, we will build an estimate for the magnitude of these phase corrections by comparing of the radiation-reaction timescale to the timescale of variation of the modulation.
+The Fresnel width must then be compared with how fast the function $F(t)$ in the integrand is varying.
+In Sec.~\ref{subsec:executivesummary}, we will build an estimate for the magnitude of these phase corrections by comparing the radiation-reaction timescale to the timescale of variation of the modulation.
 
@@
 
-Note however that this quadrature rule does not fall into the class of Gaussian quadratures, due to the complex character of the kernel in~\eqref{eq:defFresnel}.
+%Note however that this quadrature rule does not fall into the class of Gaussian quadratures, due to the complex character of the kernel in~\eqref{eq:defFresnel}.
 The induced inner product is not positive definite and some key properties are lost, such as the construction of the optimal Gaussian nodes from the zeros of the orthonormalized polynomials.
 For more details on quadrature rules when applied to integrals with oscillatory kernels, we refer to~\cite{} \SM{[C]}.
+
 Note that the choice of a stencil with quadrature nodes $\tf \pm k\Tf$, even if natural, is by no means unique, and other choices would have led to different stencils.
 The formulation of the result~\eqref{eq:resultdffPsiTf} as a Fresnel transform~\eqref{eq:resultFresnel} opens the way for future investigations of different numerical approaches to the problem.
-
-\jgb{Going forward, the Fresnel transform provides a convenient way to non-perturbatively express the effect of the quadratic phase term on the transfer function, but practical compuations will rely on the approximations $\calF_{T, \epsilon}^{N}[F] (t)$ in \eqref{eq:stencilfresnel}.[Check that I have this right.
-]}
+For practical computations, in this work we will rely on the approximations $\calF_{T, \epsilon}^{N}[F] (t)$ in \eqref{eq:stencilfresnel}.
 
@@
 In the case of a LISA-type detector response, the presence of the delay $d(t)$ is responsible for the frequency-dependence of the kernel $G$ introduced in~\eqref{eq:defG}.
-\jgb{In practice we will not use the Taylor expansion of the kernel function in~\eqref{eq:resulttaylordelay}, and will instead expand the delay itself as a slowly varying constant.}
-
-\jgb{As a first step, we consider the transfer function when we keep only the leading order waveform~\eqref{eq:leadingorderwf}, leaving off additional amplitude and phase terms in~\eqref{eq:resulttaylorA} and~\eqref{eq:resulttaylorPsi}, but treat the kernel function nonperturbatively.} The response then depends on the signal only by the time-to-frequency correspondence and reads
+As an improvement over the Taylor expansion of the kernel function in~\eqref{eq:resulttaylordelay}, in this section we will propose a different approach based on a change of variable.
+
+As a first step, we consider the transfer function when we keep only the leading order of the expansion~\eqref{eq:leadingorderwf}, leaving off additional amplitude and phase terms in~\eqref{eq:resulttaylorA} and~\eqref{eq:resulttaylorPsi}, but treating the kernel function nonperturbatively.
+The response then depends on the signal only by the time-to-frequency correspondence and reads
 \begin{equation}
@@
 This motivates a change of variable to the delayed time function $t_{d}:t \mapsto t+d(t)$.
-\jgb{Assuming the delay does not vary too quickly, we can} also define the reciprocal function $t_{d}^{-1}$, and a modified time-to-frequency correspondence $\tfd = t_{d}^{-1}(\tf)$, defined implicitly by
+Assuming the delay does not vary too quickly, we can also define the reciprocal function $t_{d}^{-1}$, and a modified time-to-frequency correspondence $\tfd = t_{d}^{-1}(\tf)$, defined implicitly by
 \be
@@
 \end{align}
-\jgb{The difference with~\eqref{eq:transferlocal}, the extra factor comprising the denominator, shows the effect of using here the non-perturbative kernel function rather than just the leading order term in~\eqref{eq:expandG}.}
-
-For the proposed LISA configuration~\cite{LISA17}, we have for the orbital delays the scaling $d_{0}\sim R/c \simeq 500s$, and for the constellation delays the scaling $d_{c}\sim L/c \simeq 8s$ (ignoring the dependence on angular factors).
+This differs from~\eqref{eq:transferlocal} both by the replacement $\tf \rightarrow t_{f}^{d}$ and by the extra denominator.
+
+For the proposed LISA configuration~\cite{LISA17}, we have for the orbital delays the scaling $d_{0}\sim R/c \simeq 500s$, and for the constellation delays the scaling $d_{L}\sim L/c \simeq 8s$ (ignoring the dependence on angular factors).
 Since the motion of the constellation is anually periodic, with a frequency $\Omega_{0} \simeq 2 \times 10^{-7}\mathrm{rad}.s^{-1}$, we have $\dot{d}_{0} \sim \Omega_{0} R/c \simeq 10^{-4}$ and $\dot{d}_{c} \sim \Omega_{0} L/c \simeq 1.7 \times 10^{-6}$.
@@
 \end{align}
-Now, the most relevant correction in~\eqref{eq:delaycorrleading} comes from the phase factor at high frequencies, where the factors $2\pi f d_{0}$ and $2\pi f d_{c}$ give a magnification factors reaching respectively $3.10^{3}$ and $10^{2}$ at $1\Hz$.
+Now, the most relevant correction in~\eqref{eq:delaycorrleading} comes from the phase factor at high frequencies, where the factors $2\pi f d_{0}$ and $2\pi f d_{L}$ give a magnification factors reaching respectively $3.10^{3}$ and $10^{2}$ at $1\Hz$.
 Ignoring the other corrections, we thus arrive at the following form for the dominant delay correction in the transfer function:
@@
 \ee
-\jgb{This expression provides a summary of the perturbative treatment we will apply in later sections, with the additional understanding that the Fresnel transform is treated approximately using \eqref{eq:fresnelstencil}.[Check this]}
 \end{widetext}
@@
 We will investigate several orders of approximation, combining corrections from the phase, amplitude and delays, and explore which ones are relevant for a given level of accuracy.
-We will use symbols of the form $\{N,A,d\}$ to indicate the order of the stencil in~\eqref{eq:stencilfresnel}, the maximal order of the amplitude correction included, and the inclusion or not of the delay corrections.
-\jgb{[The meaning of this notation is unclear.
-Some more work may be needed in this subsection.]}
+We will use symbols of the form $\{N,A,d\}$ to indicate the order of the stencil in~\eqref{eq:stencilfresnel}, the maximal order of the amplitude correction included, and the inclusion or not of the delay corrections at first order in $\dot{d}$ as derived in Sec.~\ref{subsec:delays}.
 
@@
 We begin by detailing the model that we use for the response of a LISA-like detector, together with the assumptions used and their limitations.
-Since our aim is to assess the accuracy of our direct Fourier-domain treatment of the response, we can focus on the gravitational-wave contribution to the basic single-linkobservables.
-We therefore use a somewhat simplified model for the time-domain response, ignoring corrections that are crucial from the point of view of noise cancellations, keeping in mind that the response model can be enriched later without affecting the conclusions of the present analysis.
-
-The frequency-shift response for a single link~\eqref{eq:yslr} was derived in Ref.~\cite{EW75} (see also~\cite{RCP04, Finn08, Cornish09} for more modern derivations).
+Since our aim is to assess the accuracy of our direct Fourier-domain treatment of the response, we can focus on the gravitational-wave contribution to the basic single-link observables.
+We therefore use a somewhat simplified model for the time-domain response~\cite{Krolak+04}, ignoring corrections that are crucial from the point of view of noise cancellations, keeping in mind that the response model can be enriched later without affecting the conclusions of the present analysis.
+
+The frequency-shift response for a single link~\eqref{eq:yslr} was derived in Ref.~\cite{EW75} (see also~\cite{CR02, RCP04, Finn08, Cornish09}).
 Several assumptions enter the result as written in~\eqref{eq:yslr}: (i) effects of the order $v/c$ are neglected, including for instance the special relativistic Doppler effect created by the relative speeds of the spacecraft on their orbits (ii) the propagation is assumed to take place in a flat spacetime, perturbed only by the gravitational wave; thus the gravitational redshift as well as the deflection of light created by the gravitational potential of the Sun is ignored (iii) all geometric factors are evaluated at a single time, whereas one should consider the beam as propagating from the position of the first spacecraft at the time of emission to the position of the second scapecraft at the time of reception, leading to a pointing-ahead effect.
@@
 These simplified orbits neglect (iv) effects of order $e^{2}$ from the eccentricity of the individual Keplerian orbits, (v) the effect of gravitational perturbations coming from other celestial bodies, such as the Earth, the quadrupole of the Sun, and the other planets.
-
-Note that although we can neglect all the effects (i)-(v) for our present study, keeping track of these corrections is crucial for the purpose of laser noise cancellations, and led to the development of new generations of TDI observables~\cite{} \SM{[C]}.
-This simplified response has been abundantly used in previous studies (see e.g.~\cite{} \SM{[C]}).
+Note that although we can neglect all the effects (i)-(v) for our present study, keeping track of these corrections is crucial for the purpose of laser noise cancellations, and led to the development of new generations of TDI observables~\cite{Tintoliving}.
 
@@
 \end{subequations}
-The LISA response behaves qualitatively differently on the frequency bands separated by these transfer frequencies. \jgb{[\bf I don't understand the meaning or intent of the previous sentence]}
-
-\jgb{The first stage of the response, the orbital delay, consists simply in applying the varying time delay to bring the wavefront sampling point from the SSB reference to the center} of the LISA triangular constellation, common to all $y_{slr}$ observables.
+The LISA response will behave qualitatively differently on the three frequency bands $f \lesssim f_{R}$, $f_{R} \lesssim f \lesssim f_{L}$ and $f_{L} \lesssim f$.
+
+The first stage of the response, the orbital delay, consists simply in applying the varying time delay to bring the wavefront sampling point from the SSB reference to the center of the LISA triangular constellation, common to all $y_{slr}$ observables.
 For $h^{\rm TT}$ the transverse-traceless gravitational waveform in matrix form, we write this orbital time delay as
@@
 with $p_{0}$ the position of the constellation center, which follows the Earth orbit around the Sun.
-The second \jgb{stage of the response calculation comprises the remaining, constellation-centered response.
-For the single-link contribution to the response, for the laser link from spacecraft $s$ to spacecraft $r$ along a path in direction $n_l=$ [{\bf Check that I got the notation right}], we write}
+The second stage of the response calculation comprises the remaining, constellation-centered response.
+For the single-link contribution to the response, for the laser link from spacecraft $s$ to spacecraft $r$ along a path in direction $n_l$, we write
 \begin{align}\label{eq:defresponseL}
@@
 We focus only on positive frequencies.
-Assuming that the approximation~\eqref{eq:zeronegativef} applies, we consider a single mode contribution, either $h=h_{\ell m}$ for $m>0$ or $h=h_{\ell m}^{*}$ for $m<0$ (following~\eqref{eq:notationhsignm}).
-For each mode we define a complex matrix $P_{\ell m}$ incorporating the spin-weighted spherical harmonic constant factor as
+Assuming that the approximation~\eqref{eq:zeronegativef} applies, we consider a single mode contribution, $h=h_{\ell m}$ with $m>0$.
+For each given mode we define a complex matrix $P_{\ell m}$ incorporating the spin-weighted spherical harmonic constant factor as
 \be
@@
 \end{align}
-This leading-order result~\eqref{eq:transfer0local}-\eqref{eq:transferLlocal} is familiar, as it was derived in the past taking the point of view of monochromatic signals (see e.g.~\cite{} \SM{[C]}) \SM{[rewrite that]}.
-For plotting purposes, we will define
+For plotting purposes, we will also define
 \be\label{eq:transferLenvelope}
@@
 
-Note that, if the corrections of Sec.~\ref{subsec:delays}, \jgb{for non-negligible $\dot d$}, are included for the constellation delays, the transfer function will not have this simple form anymore, as $\dot{d}$ will have a different velocity-dependent expression for the sending and receiving spacecraft.
-One must \jgb{separately handle $d_{s,L}$ and $d_{r}$ in~\eqref{eq:decomposeGslr} to compute the corrections.}
-
-The \jgb{orbital} response~\eqref{eq:transfer0local}-\eqref{eq:delay0} takes a simple analytic form, but the phase contribution of this delay is significant across most of the frequency band and can be large for $f \gg f_{R}$.
-
-The \jgb{constellation} response~\eqref{eq:GslrL}-\eqref{eq:transferLlocal} can be interpreted as the Fourier-domain translation of a discrete derivative taken on the waveform.
-The \jgb{leading factor in~\eqref{eq:GslrL}} shows that the amplitude of the response is proportional to $f$ in the low-frequency limit $f\ll f_{L}$, where the other factors are essentially unity.
-For $f\gtrsim f_{L}$, the $\sinc$ and the phase of the exponential generate additional structure in the response, including zero-crossings when the projected armlength \jgb{is an integer number of wavelengths}.
-From~\eqref{eq:transferLlocal}, an expansion for small $f\ll f_{L}$ yields back a Fourier-domain analog of the low-frequency approximation of the response~\cite{Cutler97, RCP04}, which \jgb{is equivalent to} having two LIGO-type interferometers turned by $\pi/4$ and set in motion.
+Note that, if the corrections of Sec.~\ref{subsec:delays}, for non-negligible $\dot d$, are included for the constellation delays, the transfer function will not have this simple form anymore, as $\dot{d}$ will have a different velocity-dependent expression for the sending and receiving spacecraft.
+One must then separately handle $d_{s,L}$ and $d_{r}$ in~\eqref{eq:decomposeGslr} to compute the corrections.
+
+The orbital response~\eqref{eq:transfer0local}-\eqref{eq:delay0} takes a simple analytic form, but the phase contribution of this delay is significant across most of the frequency band and can be large for $f \gg f_{R}$.
+
+The constellation response~\eqref{eq:GslrL}-\eqref{eq:transferLlocal} can be interpreted as the Fourier-domain translation of a discrete derivative taken on the waveform.
+The leading factor in~\eqref{eq:GslrL} shows that the amplitude of the response is proportional to $f$ in the low-frequency limit $f\ll f_{L}$, where the other factors are essentially unity.
+For $f\gtrsim f_{L}$, the $\sinc$ and the phase of the exponential generate additional structure in the response, including zero-crossings when the projected armlength is an integer number of wavelengths.
+From~\eqref{eq:transferLlocal}, an expansion for small $f\ll f_{L}$ yields back a Fourier-domain analog of the low-frequency approximation of the response~\cite{Cutler97, RCP04}, which is equivalent to having two LIGO-type interferometers turned by $\pi/4$ and set in motion.
 
@@
 \begin{align}
-	\frac{1}{G_{L}} \partial_{t} G_{L} &\sim -2i\pi f \dot{d}_{c} + \frac{\dot{F}}{F}\,,\\
-	\frac{1}{G_{L}} \partial_{tt} G_{L} &\sim -2i\pi f \ddot{d}_{c} - 4\pi^{2} f^{2} \dot{d}_{c}^{2} - 2i\pi f \dot{d}_{c} \frac{\dot{F}}{F} - \frac{\dot{F}^{2}}{F^{2}} + \frac{\ddot{F}}{F} \,,\\
-	\frac{1}{G_{L}} \partial_{tf} G_{L} &\sim -4 i \pi \dot{d}_{c} - 2i\pi d_{c} \frac{\dot{F}}{F} - 4\pi^{2} f d_{c} \dot{d}_{c} + \frac{1}{f}\frac{\dot{F}}{F}\,.
+	\frac{1}{G_{L}} \partial_{t} G_{L} &\sim -2i\pi f \dot{d}_{L} + \frac{\dot{F}}{F}\,,\\
+	\frac{1}{G_{L}} \partial_{tt} G_{L} &\sim -2i\pi f \ddot{d}_{L} - 4\pi^{2} f^{2} \dot{d}_{L}^{2} - 2i\pi f \dot{d}_{L} \frac{\dot{F}}{F} - \frac{\dot{F}^{2}}{F^{2}} + \frac{\ddot{F}}{F} \,,\\
+	\frac{1}{G_{L}} \partial_{tf} G_{L} &\sim -4 i \pi \dot{d}_{L} - 2i\pi d_{L} \frac{\dot{F}}{F} - 4\pi^{2} f d_{L} \dot{d}_{L} + \frac{1}{f}\frac{\dot{F}}{F}\,.
 \end{align}
@@
 As proposed by several authors~\cite{BCV03b, BCPTV05, Schmidt+10, OShaughnessy+11, Boyle+11}, if one performs the mode decomposition of the waveform~\eqref{eq:defmodes} not in a fixed inertial frame but rather in a time-dependent, rotating frame that follows the plane of the orbit, it is possible to restore much of the structure of a non-precessing waveform.
-In particular, one recovers qualitatively the hierarchy of mode amplitudes that prevails for non-precessing systems, with the modes $h_{22}$ and $h_{2,-2}$ being dominant\jgb{, and with each harmonic mode characterized by a chirp with smoothly evolving amplitude and phase}.
+In particular, one recovers qualitatively the hierarchy of mode amplitudes that prevails for non-precessing systems, with the modes $h_{22}$ and $h_{2,-2}$ being dominant, and with each harmonic mode characterized by a chirp with smoothly evolving amplitude and phase.
 In the following, we will identify the $z$-axis of the co-precessing frame as the dominant eigenvector of the matrix representing the action of the angular momentum operator acting on the waveform modes, as proposed in~\cite{OShaughnessy+11} (see App.~\ref{app:wigner} for more details).
@@
 	h_{\ell m}^{\rm I} = \sum\limits_{m=-\ell}^{\ell} \calD^{\ell *}_{mm'} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm P} \,, \\
-	h_{\ell m}^{\rm P} = \sum\limits_{m=-\ell}^{\ell} \calD^{\ell }_{m'm} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm I} \,,
+	h_{\ell m}^{\rm P} = \sum\limits_{m=-\ell}^{\ell} \calD^{\ell }_{m'm} (\alpha,\beta,\gamma) h_{\ell m'}^{\rm I} \,.
 \end{align}
 \end{subequations}
-where the coefficients $\calD^{\ell}_{mm'}$ are given by Wigner D-matrices~\cite{Wigner59} as
+Notice that there is no mixing of the modes with different values of $\ell$. Here the coefficients $\calD^{\ell}_{mm'}$ are given by Wigner D-matrices~\cite{Wigner59} as
 \be\label{eq:defWignerD}
@@
 \ee
-Here, the \jgb{real-valued} Wigner d-matrix $d^{\ell}_{mm'}(\beta)$ takes the form of a polynomial in $\cos (\beta/2)$, $\sin (\beta/2)$, and acts as an amplitude for the modulation function.
+Here, the real-valued Wigner d-matrix $d^{\ell}_{mm'}(\beta)$ takes the form of a polynomial in $\cos (\beta/2)$, $\sin (\beta/2)$, and acts as an amplitude for the modulation function.
 We refer to App.~\ref{app:wigner} for explicit expressions.
@@
 
-\jgb{An important qualitative observation is that the opening angle $\beta$ is typically small.
+An important qualitative observation is that the opening angle $\beta$ is typically small.
 Large misalignments between $\bm{J}$ and $\bm{\ell}$ can only be reached with large spins and large mass ratios.
-When $\beta$ is small,} from the explicit expression~\eqref{eq:defWignerdapp}, one can see that $d^{\ell}_{mm'}(\beta)$ is greatly suppressed in the limit $\beta \rightarrow 0$ when increasing $|m-m'|$.
+When $\beta$ is small, from the explicit expression~\eqref{eq:defWignerdapp}, one can see that $d^{\ell}_{mm'}(\beta)$ is greatly suppressed in the limit $\beta \rightarrow 0$ when increasing $|m-m'|$.
 Intuitively, this means that, in the limit of a small misalignment between frames, the rotation produces mainly mode contributions with the same mode number.
@@
 
-\jgb{The dominant features of the the precessing-frame waveform, are approximtely reflection symmetric about the orbital plane (exactly valid for non-precessing systems, see~\eqref{eq:symmetryhlminusm}), implying}
+The dominant features of the the precessing-frame waveform, are approximtely reflection symmetric about the orbital plane (exactly valid for non-precessing systems, see~\eqref{eq:symmetryhlminusm}), implying
 \be\label{eq:approxsymmetryhlminusm}
@@
 
-\SM{[qualify how bad the model is with respect to full NR ?]}
-
-\SM{[explain shortcomings of pure SEOB in post-merger, problem of $\beta$ oscillating ?]}
-
 We wish to apply the formalism presented in Sec.~\ref{sec:formalism} to IMR waveforms of precessing binaries, exploring the separation of the timescales involved through the inspiral and merger, and assessing the relevance of the higher-order corrections summarized in~\ref{subsec:executivesummary}.
 To this end, we will use a simplified model for the precession, allowing a number of simplifications and idealizations.
-
-%To this end, we will use a simplified model for IMR waveforms of precessing binaries, which we will use in the following to .
-Since our purpose in this article is not to propose a full-fledged waveform model, but rather to investigate wether the formalism of Sec.~\ref{sec:formalism} is relevant to tackle the merger part of the signals, and wether its application reduces the errors in the reconstruction of transfer functions, we will allow for a number of idealizations and simplifications in this toy model.
 
@@
 
-\jgb{Some attention is required in specifying an model for the post merger precession.}
-For the post-merger precession, the intuitive interpretation of the precessing frame as roughly following the orbital plane of the binary is lost.
-However, the prescription of~\cite{OShaughnessy+11} to extract the radiation axis is based entirely on the waveform itself,\jgb{ [Should state the principle, minimizing m=0 mode?]} and can be applied to the ringdown as well.
+An extension is required in specifying a model for the precession post-merger.
+After the merger occurs, the intuitive interpretation of the precessing frame as roughly following the orbital plane of the binary is lost.
+However, the prescription of~\cite{OShaughnessy+11} to extract the radiation axis is based entirely on the waveform itself, and can be applied to the ringdown as well.
 For $\ell=2$, analyzing numerical relativity waveforms, Ref.~\cite{OShaughnessy+12} described a qualitative model where the radiation axis essentially keeps precessing around the final angular momentum $\bm{J}$, but transitions to a faster precession rate, with an angular velocity determined by the quasinormal mode (QNM) frequencies of the remnant black hole as
@@
 \ee
-This \jgb{implies} a significant acceleration of the precession post-merger with respect to the inspiral (see Fig.~\ref{fig:precmodel}).
+This implies a significant acceleration of the precession post-merger with respect to the inspiral (see Fig.~\ref{fig:precmodel}).
 For the opening angle of the precessing cone $\beta$, Ref.~\cite{OShaughnessy+12} proposes an exponential decay driven similarly by the difference between damping frequencies.
-We found by \sout{a non-exhaustive \SM{[allowed to say that ?
-the catalog is vast now]}} inspection of \jgb{various} precessing waveforms currently available in the SXS catalog (see e.g.~\cite{SXScatalog, Mroue+12, Mroue+13}) that this picture is at least qualitatively correct for $\ell=2$, with $\beta$ appearing to transition to a lower value rather than decaying all the way to zero.
+We found by inspection of various precessing waveforms currently available in the SXS catalog (see e.g.~\cite{SXScatalog, Mroue+12, Mroue+13}) that this picture is at least qualitatively correct for $\ell=2$, with $\beta$ appearing to transition to a lower value rather than decaying all the way to zero.
 By contrast, we found that the current ringdown prescription in the SEOBNRv3 model can cause a qualitatively different behaviour, with oscillations in $\beta$.
@@
 The difference between NR and SEOB there is due to the value of the final spin, with $\chi_{f}^{\rm NR} = 0.54$ and $\chi_{f}^{\rm SEOB} = 0.42$, which leads to a different $\Omega_{\rm frame}$.
-In the right panel, $\beta$ is asymptotically constant in our toy model.
-\SM{[Add amplitude/$\omega$ plots comparing NR/SEOB, and possibly PhenomD for the 22 mode ?
-Add Euler angles produced by IFFT of PhenomP ?]}}
+In the right panel, $\beta$ is asymptotically constant in our toy model.}
   \label{fig:precmodel}
@@
 The overall $\nu$ scaling indicates a better separation of timescales for larger mass ratio systems.
-These expressions also show that, since the total angular momentum appears as a factor in $\Omega_{\rm prec}$, higher-order corrections will be larger for \jgb{roughly} spin-aligned systems than for anti-aligned spins, an effect that becomes significant for large mass ratios.
+These expressions also show that, since the total angular momentum appears as a factor in $\Omega_{\rm prec}$, higher-order corrections will be larger for roughly spin-aligned systems than for anti-aligned spins, an effect that becomes significant for large mass ratios.
 Note that the regime where $\xi$ gets close to 0 corresponds to the transitional precession range~\cite{Apostolatos+94}, with the spin of the primary compensating the orbital angular momentum, and our analysis based on simple precession is not valid anymore.
@@
 
-Instead of tapering the signal to $0$ \jgb{below [CHECK]} $f_{0}$, to limit the deviation from the original signal we take advantage of the one-sidedness and only flatten the amplitude\footnote{In practice, this is implemented as the discrete integral of a cosine window function on just the first two samples.
+Instead of tapering the signal to $0$ below $f_{0}$, to limit the deviation from the original signal we take advantage of the one-sidedness and only flatten the amplitude\footnote{In practice, this is implemented as the discrete integral of a cosine window function on just the first two samples.
 We also taper to 0 the last three samples before $f_{\rm max}$, and $0$-pad by a factor of 2.}, and artificially symmetrize the signal.
@@
 The thick and thin vertical lines show the merger and asymptotic ringdown frequencies respectively.
-\jgb{For the solid curves in the top panel, precession modulations were computed in the time domain, while the dashed curves show the results of the Fourier-domain reconstruction techniques presented here.[CHECK]}
+For the solid curves in the top panel, transfer functions were computed from an FFT of the full time-domain precessing signal, while the dashed curves show the results of the Fourier-domain reconstruction techniques presented here.
 The columns show the orders of approximation $\{\Psi:0,A:0\}$, $\{\Psi:2,A:2\}$, $\{\Psi:2,A:2,\text{conv}\}$.
@@
   \includegraphics[width=.98\linewidth]{plots/precMM_py.pdf}
-  \caption{Unfaithfulness at various orders of approximation \jgb{as compared with a time-domain treatment using a full-domain[CHECK]} a numerical FFT, for the three precessing systems of Table~\ref{tab:precexamples}.
+  \caption{Unfaithfulness at various orders of approximation as compared with an FFT of the full time-domain precessing signal, for the three precessing systems of Table~\ref{tab:precexamples}.
 The color indicates the order of approximation of the perturbative treatment of Sec.~\ref{subsec:executivesummary}.
@@
 To explore the application of the formalism to precessing binaries, we built a toy model for the precession of a two-spins system including a post-merger precessing-frame trajectory.
-For the high spin, moderate mass ratio examples we explored, we found that subdominant precessing mode contributions can be challenging for the formalism and have large relative errors, \jgb{although they are suppressed} in amplitude in the complete waveform.
+For the high spin, moderate mass ratio examples we explored, we found that subdominant precessing mode contributions can be challenging for the formalism and have large relative errors, although they are suppressed in amplitude in the complete waveform.
 For the inspiral, the most relevant corrections are the quadratic in phase ones that were already derived in~\cite{KCY14}.
@@
 
-Our results \jgb{suggest a few} straightforward extensions.
+Our results suggest a few straightforward extensions.
 First, we only considered the leading harmonic of non-spinning waveforms in the LISA response.
-\jgb{The same approach is also well-suited to treating higher harmonics and aligned spin signal}, but precession will leave a definite imprint on the Fourier-domain signal.
+The same approach is also well-suited to treating higher harmonics and aligned spin signal, but precession will leave a definite imprint on the Fourier-domain signal.
 One option in this case would be to handle the precession and response modulation together, as in~\cite{Klein+15}.
@@
 
-\jgb{Another natural application of our formalism developed for the LISA response is to third-generation ground-based detectors.
-For detectors like the Einstein Telescope,} signals are sufficiently long for the rotation of the Earth to matter, and their arms are long enough that the long wavelength approximation is not valid at high frequencies and that the response becomes frequency-dependent~\cite{Essick+17}.
-A preliminary order-of-magnitude analysis shows that the leading-order treatment should be applicable \SM{[check]}, \jgb{but more investigation is needed.}
-
-\jgb{Finally, while we restricted ourselves to comparable-mass binary systems, Extreme Mass Ratio Inspirals (EMRI) will also be very important signals for LISA.
+Another natural application of our formalism developed for the LISA response is to third-generation ground-based detectors.
+For detectors like the Einstein Telescope, signals are sufficiently long for the rotation of the Earth to matter, and their arms are long enough that the long wavelength approximation is not valid at high frequencies and that the response becomes frequency-dependent~\cite{Essick+17}.
+A preliminary order-of-magnitude analysis shows that the leading-order treatment should be applicable \SM{[check]}, but more investigation is needed.
+
+Finally, while we restricted ourselves to comparable-mass binary systems, Extreme Mass Ratio Inspirals (EMRI) will also be very important signals for LISA.
 Those signals generally include not only strong precession but also large eccentricity.
-An effective frequency-domain LISA response may be possible by apply our treatment separately to each of the numerous individual harmonics produced (see e.g.~\cite{Chua+17}).}
+An effective frequency-domain LISA response may be possible by apply our treatment separately to each of the numerous individual harmonics produced (see e.g.~\cite{Chua+17}).
 
@@
 
-\SM{[say a word of resampling needed at low and high f for LISA, and in general interpolation errors, appendix or refer to paper PE ?]}
-
-\SM{[clean up part on discrete Fourier coeffs]}
+\SM{[check part on discrete Fourier coeffs]}
 
@@
 An additional approximation often used in waveform models consists in considering that the Fourier transforms $\tilde{h}_{\ell m}$ have support only on one side of the spectrum, either for positive of for negative frequencies depending on the sign of $m$.
-This holds in particular within the stationary phase approximation, as explained below \SM{[change here]}.
+This holds in particular within the stationary phase approximation (see Sec.~\ref{subsec:SPA}).
 With our sign convention~\eqref{eq:defFT}, this approximation reads

### 2018-06-18 docs/paper_method.tex
Edits to the paper.

@@
 In both cases, being able to compute transfer functions for these effects directly in the Fourier domain may enable performance improvements for data analysis applications by using fast frequency-domain waveforms.
-Our results generalize previous approaches based on the stationary phase approximation, extending them by including delays and higher-order corrections while being applicable to a broader class of signals including full inspiral-merger-ringdown waveforms.
-In the LISA case, we find that higher-order corrections remain small except at the two extremes of the detector's frequency band, where we show that including them reduces the errors.
+Our results generalize previous approaches based on the stationary phase approximation for inspiral signals, extending them by including delays and computing corrections beyond the leading order, while being applicable to the broader class of signals of inspiral-merger-ringdown waveforms.
+In the LISA case, we find that a leading-order treatment is accurate for high-mass and low-mass signals that are chirping enough, with reconstruction errors that can be reduced by consistently including the corrections we derived.
+By contrast, low-mass binary black holes, if far away from merger and slowly-chirping, cannot be handled by our perturbative formalism and we develop another approach for these systems.
 In the case of precessing binaries, we perform a limited exploration of the merger-ringdown range based on a toy model for the precession.
-We find that, depending on the post-merger frame velocity, higher-order corrections can become quantitatively large to the point of challenging our perturbative formalism, while affecting mainly the amplitude and giving rise to limited unfaithfulness.
-We further present an alternative direct convolution approach to accurately represent these post-merger features.
-\SM{[update to include SOBHs]}
+We find that, depending on the post-merger frame velocity, deviations from leading order can become qualitatively large to the point of challenging our perturbative formalism, while affecting mainly subdominant mode contributions and giving rise to a limited unfaithfulness in the full waveform.
+Including higher-order corrections consistently reduces the unfaithfulness, and we further develop an alternative approach to accurately represent post-merger features.
 
@@
 
-\jgb{The GW community is making progress assembling higher-fidelity tools for these coming challenges.
-State-of-the-art IMR templates combine information from the perturbative results of post-Newtonian theory covering the inspiral (see e.g.~\cite{BlanchetLiving}) and from numerical relativity simulations covering the end of the inspiral and the merger-ringdown phase (see e.g.~\cite{Pfeiffer12}).
-Approaches to template construction include phenomenological templates postulating an analytic ansatz for the Fourier-domain amplitude and phase~\cite{Husa+15, Khan+15,Hannam+13}, and the more physically motived Effective-One-Body (EOB) approach~\cite{BD99,Taracchini+13, Pan+13, Bohe+16}.
-Where needed, Reduced Order Models (ROM), also called surrogate models, have been developed to considerably speed up waveform generation, without losing accuracy~\cite{Field+13, Puerrer14, Bohe+16}.
-These tools provide efficient non-precessing IMR Fourier-domain waveforms.
-Importantly for this work, the resulting waveforms can be represented by an amplitude and phase for each mode, with only a few hundred samples~\cite{Puerrer14}.[JB:I cut this down significantly.]} Our challenge will be to accurately complete the representation of instrumental signals including precession effects and the LISA instrumental response and precession effects without sacrificing the computational efficiency enabled by these fast waveform templates.
+The GW community is making progress assembling higher-fidelity tools for these coming challenges.
+State-of-the-art IMR templates combine information from the perturbative results of post-Newtonian (PN) theory covering the inspiral (see e.g.~\cite{BlanchetLiving}) and from numerical relativity (NR) simulations covering the end of the inspiral and the merger-ringdown phase (see e.g.~\cite{Pfeiffer12}).
+Approaches to template construction include phenomenological templates postulating an analytic ansatz for the Fourier-domain amplitude and phase~\cite{Husa+15, Khan+15,Hannam+13}, and the Effective-One-Body (EOB) approach~\cite{BD99,Taracchini+13, Pan+13, Bohe+16} incorporating PN and NR information.
+Where needed, Reduced Order Models (ROM), also called surrogate models, have been developed to considerably speed up waveform generation, without losing accuracy~\cite{Field+13, Puerrer14, Blackman+17a}.
+Put together, these tools provide efficient non-precessing IMR Fourier-domain waveforms, while recent progress has been made for precessing systems as well~\cite{Hannam+13, Chatziioannou+17, Blackman+17b}.
+Importantly for this work, the resulting waveforms can be represented by an amplitude and phase for each mode, with only a few hundred samples~\cite{Puerrer14}.
+
+%Our challenge will be to accurately complete the representation of instrumental signals including precession effects and the LISA instrumental response and precession effects without sacrificing the computational efficiency enabled by these fast waveform templates.
+\SM{[rewrite that]}
 
 A complete representation of spin effects in fast IMR templates still remains a frontier of gravitational wave signal modelling.
-In presence of misaligned spins, the system will endure precession of the orbital plane as it evolves, leading to modulations of the signal as seen by the observer~\cite{ACST94}.
-%Precessing EOB templates incorporate all six parameters of the two spins~\cite{Pan+13, BTB16} but still require a costly time-domain integration; while precessing phenomenological templates~\cite{Hannam+13} decrease the dimensionnality by using effective spin parameters representing only the dominant effects of the spins.
-\jgb{[JB:I cut this down too.]}
-
-A promising approach to modeling the effect of precession on the emitted waveform, it has been proposed~\cite{BCV03b, BCPTV05, Schmidt+10, OShaughnessy+11, Boyle+11} to decompose precessing waveforms by performing a time-dependent rotation, following the precession of the orbital plane.
+In presence of misaligned spins, the system will endure precession of the orbital plane as it evolves, leading to modulations of the signal as seen by the observer~\cite{Apostolatos+94, Kidder95}.
+The presence of six degrees of freedom for the spins increases the dimensionality of the problem.
+
+A promising approach to modeling the effect of precession on the emitted waveform, as proposed in~\cite{BCV03b, BCPTV05, Schmidt+10, OShaughnessy+11, Boyle+11}, is to decompose precessing waveforms by performing a time-dependent rotation, following the precession of the orbital plane.
 The resulting waveform in the rotated frame can then be modelled by a non-precessing waveform, an approximation which is used both in the construction of the inspiral part of precessing EOB waveforms~\cite{Pan+13} and in the construction of precessing phenomenological waveforms~\cite{Hannam+13}.
-To follow this modelling approach and efficiently create Fourier-domain waveforms, one needs to understand how to translate the time-domain modulations created by the frame rotation into a Fourier-domain transfer function, which is our first motivation.
-
-Beyond a fast representation of the incident waveform, observational analyses also require transforming the signals through some instrumental response.
-For short duration mergers such as LIGO  has detected, this can be treated by a simple multiplier.
+To follow this modelling approach and efficiently create Fourier-domain waveforms, one needs to understand how to translate the time-domain modulations created by the frame rotation into a Fourier-domain transfer function.
+
+Beyond a fast representation of the incident gravitational wave, observational analyses also require transforming the signals through some instrumental response.
+For short duration mergers such as LIGO  has detected, this can be treated by a simple multiplier and a fixed timeshift between detectors.
 For future instruments though, the instrumental response will be more complicated.
@@
 
-\jgb{Whereas LIGO and Virgo are typically sensitive to chirping binaries for a minute or less, LISA signals may accumulating over months or years.
+Whereas LIGO and Virgo are typically sensitive to chirping binaries for a minute or less, LISA signals may accumulating over months or years.
 The response of a LISA-type instrument is thus time-dependent~\cite{Cutler97}.
@@
 These effects then convey information about the localization of the gravitational-wave source in the sky.
-Direct implementation of the detector response is straightforward\cite{SynthLISA,LISACode,LISASim} but at prohibitive computational cost for parameter-estimation analyses.
-To leverage the performance of state-of-the-art Fourier-domain IMR templates~\cite{BTB16,Khan+15}, we must efficiently process the signals through the time-dependent response of the dectector while staying in the Fourier domain.}
+Direct time-domain implementation of the detector response is straightforward~\cite{Vallisneri04, Petiteau+08, CR02, RCP04}, but at a prohibitive computational cost for parameter-estimation analyses.
+To leverage the performance of state-of-the-art Fourier-domain IMR templates~\cite{BTB16,Khan+15}, we must efficiently process the signals through the time-dependent response of the dectector while staying in the Fourier domain.
 
@@
 
-\jgb{In the case of precessing binaries, applying the SPA directly to the modulated signal is also prone to pathologies.
-In Ref.~\cite{KCY13,KCY14}, a formalism (called shifted uniform asymptotics or SUA) was introduced to go beyond the SPA and compute more accurately the modulation in the Fourier domain; however, this formalism still relies on the SPA for the underlying precessing-frame signal, and is as such limited to inspiraling signals.
-\textbf{[JB: Did I capture your thoughts correctly in the following.]}
-The simplified treatment of the precession response in the phenomenological waveforms in Ref.~\cite{Hannam+13} takes another approach, treating the precession modulation in the frequency domain by directly associating the Fourier frequency with (a multiple of) the post-Newtonain orbital frequency.
-As we will explain, this corresponds to the zeroth-order approximation of the SUA.}
-
-In the case of the LISA response, the SPA provides a natural map from time-domain (for the orbit) to Fourier-domain (for the signal)~\cite{Cutler97}.
-Ref.~\cite{Klein+15} included the orbital motion of the detector in the SUA treatment.
-Consistently extending these previous approaches to the merger-ringdown part of the signals, including the delays in the LISA case, is part of the objectives of this paper.
+In the case of precessing binaries, applying the SPA directly to the modulated signal is also prone to pathologies.
+In Refs.~\cite{KCY13,KCY14}, a formalism (called shifted uniform asymptotics or SUA) was introduced to go beyond the SPA and compute more accurately the modulation in the Fourier domain; however, this formalism still relies on the SPA for the underlying precessing-frame signal, and is as such limited to inspiraling signals.
+The simplified treatment of the precession response in the phenomenological waveforms of Ref.~\cite{Hannam+13} takes another approach, treating the precession modulation in the frequency domain by directly associating the Fourier frequency with the post-Newtonain orbital frequency.
+As we will explain, this corresponds to the zeroth-order approximation of the SUA.
+
+In the case of the LISA response, the SPA provides a natural map from time-domain (for the orbit) to Fourier-domain (for the signal)~\cite{Cutler97}, and was used by many previous studies with inspiral waveforms.
+Ref.~\cite{Klein+15} included the orbital motion of the detector in the SUA treatment of precessing inspiral signals, within a low-frequency approximation of the constellation response.
+Consistently extending these previous approaches to the merger-ringdown part of the signals, while including the delays in the full LISA response at all frequencies, is part of the objectives of this paper.
 
@@
 Second, there is often no clear way to improve the accuracy of these methods beyond the intuitive leading order treatment in order to meet the high-accuracy needs of future detectors.
-Our approach exploits separation of time-scales approximations, based on a general treatment directly in the Fourier domain of slowly varying delays and amplitude modulations for chirping waveforms with slowly varying amplitude and phase.
+Our approach exploits separation of time-scales approximations, based on a general treatment directly in the Fourier domain of slowly varying delays and amplitude modulations for chirping waveforms.
 
 The plan of the paper is as follows.
-In Sec.~\ref{sec:motivation}, we provide a general presentation of problem of Fourier-domain modulation and introduce the relevant timescales for both the response of LISA-type detectors and the modulation of precessing signals.
+In Sec.~\ref{sec:motivation}, we provide a general presentation of the problem of Fourier-domain modulation and introduce the relevant timescales for both the response of LISA-type detectors and the modulation of precessing signals.
 In Sec.~\ref{sec:formalism}, we present our general formalism, give its leading order approximation as well as higher-order corrections, introduce new timescales based on the Fourier-domain signal, refine the previous results for both the quadratic-in-phase corrections and the treatment of the delays, and make the connection to previous approaches of the problem.
@@
 We discuss and summarize our results in Sec.~\ref{sec:discussion}.
-
-\SM{[say that we reverse point of view: waveform is given FD, modulation TD, instead of waveform+modulation TD]}
-
-\SM{[say a word of how to gauge relevance of a given error: 1/SNR in direct error ? We won't give anything but waveform reconstruction errors...]}
-
-\SM{[check citation of Cornish-Rubbo-Poujade early work]}
 
@@
 where we set $v = (G M\omega/c^{3})^{1/3}$ and where $D$ is the luminosity distance to the observer.
-For this Newtonian inspiral, applying the SPA gives for the $22$ mode $\tilde{h}(f) = A_{\rm N}(f)e^{-i\Psi_{\rm N}(f)}$ with \SM{[check that total mass convention is $M$ everywhere]}
+For this Newtonian inspiral, applying the SPA gives for the $22$ mode $\tilde{h}(f) = A_{\rm N}(f)e^{-i\Psi_{\rm N}(f)}$ with
 \begin{subequations}\label{eq:SPAN}
@@
 
-\SM{[explain somewhere the fact that the modulation has support for negative f, convolution has support to larger f - pretty important for direct convolution]}
-
-\SM{[cite GWFrames]}
-
-\SM{[replace $pp$ everywhere with $\perp\perp$]}
-
-\SM{[check error estimates with single-spin configurations $+$, $\perp$, $-$]}
-
 In this Section, we illustrate the application of the formalism described in~\ref{sec:formalism} to signals from precessing binaries, with modulations induced by the precession.
@@
 By contrast, we found that the current ringdown prescription in the SEOBNRv3 model can cause a qualitatively different behaviour, with oscillations in $\beta$.
-\SM{[motivate this with final spin estimate as well, if we update this ?]}
-%and its estimate for the spin of the final remnant~\cite{} would require to be updated (see e.g.~\cite{}) as shown in Fig.~\ref{fig:precmodel}.
 
@@
 The target unfaithfulness for waveform models depends on the application.
-For ground-based detectors like advanced LIGO and Virgo, it is often set to 1\% or 3\%, while detectors of the next generations, including LISA, will require a lower unfaithfulness for their higher SNR \SM{[acronym]} signals.
+For ground-based detectors like advanced LIGO and Virgo, it is often set to 1\% or 3\%, while detectors of the next generations, including LISA, will require a lower unfaithfulness due to their higher signal-to-noise ratios.
 

### 2018-06-18 docs/paper_method.tex
Edits to the paper.

@@
 We will investigate several orders of approximation, combining corrections from the phase, amplitude and delays, and explore which ones are relevant for a given level of accuracy.
-We will use symbols of the form $\{N,A,d\}$ to indicate the order of the stencil in~\eqref{eq:stencilfresnel}, the maximal order of the amplitude correction included, and the inclusion or not of the delay corrections at first order in $\dot{d}$ as derived in Sec.~\ref{subsec:delays}.
+We will use symbols of the form $\{N | A | d\}$ to indicate the order of the stencil in~\eqref{eq:stencilfresnel}, the maximal order of the amplitude correction included, and the inclusion or not of the delay corrections at first order in $\dot{d}$ as derived in Sec.~\ref{subsec:delays}.
 
@@
 \begin{align}
-	\{N,A:0,d:0\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ F e^{-2i\pi f d} \right] (t_{f}) \,, \\
-	\{N,A:1,d:0\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ \left(F - i T_{A1} \left( \dot{F} - 2i\pi f \dot{d}\right) \right) e^{-2i\pi f d} \right] (t_{f}) \,, \\
-	\{N,A:0,d:1\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ \frac{F - d\dot{F}}{1+\dot{d}} e^{-2i\pi f d (1-\dot{d})} \right] (t_{f})\,, \\
-	\{N,A:1,d:1\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ \frac{1}{1+\dot{d}}\left(  F - d\dot{F} - i T_{A1} \left( \dot{F} - 2i\pi f \dot{d} \right) \right) e^{-2i\pi f d (1 - \dot{d})} \right] (t_{f})\,.
+	\{N | A:0 | d:0\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ F e^{-2i\pi f d} \right] (t_{f}) \,, \\
+	\{N | A:1 | d:0\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ \left(F - i T_{A1} \left( \dot{F} - 2i\pi f \dot{d}\right) \right) e^{-2i\pi f d} \right] (t_{f}) \,, \\
+	\{N | A:0 | d:1\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ \frac{F - d\dot{F}}{1+\dot{d}} e^{-2i\pi f d (1-\dot{d})} \right] (t_{f})\,, \\
+	\{N | A:1 | d:1\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ \frac{1}{1+\dot{d}}\left(  F - d\dot{F} - i T_{A1} \left( \dot{F} - 2i\pi f \dot{d} \right) \right) e^{-2i\pi f d (1 - \dot{d})} \right] (t_{f})\,.
 \end{align}
@@
 \begin{align}
-	\{N,A:0\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ F \right] (t_{f}) \,, \\
-	\{N,A:1\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ F - i T_{A1} \dot{F} \right] (t_{f}) \,, \\
-	\{N,A:2\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ F - i T_{A1} \dot{F} - \frac{1}{2} (T_{A2})^{2} \ddot{F} \right] (t_{f}) \,.
+	\{N | A:0\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ F \right] (t_{f}) \,, \\
+	\{N | A:1\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ F - i T_{A1} \dot{F} \right] (t_{f}) \,, \\
+	\{N | A:2\}&: \; \calT(f) = \calF^{N}_{\Tf, \epsilon} \left[ F - i T_{A1} \dot{F} - \frac{1}{2} (T_{A2})^{2} \ddot{F} \right] (t_{f}) \,.
 \end{align}
@@
 We will use the PhenomD model~\cite{Khan+15,Husa+15}, which provides Fourier-domain inspiral-merger-ringdown waveforms for aligned spins.
+We refer to App.~\ref{app:precLISA} for a brief discussion of the prospects for applying our formalism for the LISA response to precessing Fourier-domain waveforms.
 
@@
 
-The figures Fig.~\ref{fig:precerrors++}, Fig.~\ref{fig:precerrorspp} and Fig.~\ref{fig:precerrors--} show the result for the cases $++$, $\perp\perp$ and $--$ of Table~\ref{tab:precexamples} at three successive approximations, following the notation of~\eqref{eq:summaryNA}: $\{N:0,A:0, \mathrm{No \; Conv.}\}$, which is simply the leading-order transfer function~\eqref{eq:transferlocal}, ignoring all the corrections; $\{N:3,A:2, \mathrm{No \; Conv.}\}$, which incorporates both the phase corrections~\eqref{eq:stencilfresnel}, using a stencil size $N=3$, and the amplitude corrections up to second order in~\eqref{eq:summaryNA}; and $\{N:3,A:2,\text{conv}\}$, which is the same as the previous setting for the inspiral but uses the convolution formalism of Sec.~\ref{subsec:trigopoly} to cover the high frequency range, with a smooth transition in the shaded range.
+The figures Fig.~\ref{fig:precerrors++}, Fig.~\ref{fig:precerrorspp} and Fig.~\ref{fig:precerrors--} show the result for the cases $++$, $\perp\perp$ and $--$ of Table~\ref{tab:precexamples} at three successive approximations, following the notation of~\eqref{eq:summaryNA}: $\{N:0 | A:0 | \mathrm{No \; Conv.}\}$, which is simply the leading-order transfer function~\eqref{eq:transferlocal}, ignoring all the corrections; $\{N:3 | A:2 | \mathrm{No \; Conv.}\}$, which incorporates both the phase corrections~\eqref{eq:stencilfresnel}, using a stencil size $N=3$, and the amplitude corrections up to second order in~\eqref{eq:summaryNA}; and $\{N:3 | A:2 | \text{conv}\}$, which is the same as the previous setting for the inspiral but uses the convolution formalism of Sec.~\ref{subsec:trigopoly} to cover the high frequency range, with a smooth transition in the shaded range.
 The panels show the Fourier-domain amplitude for each of the modes, both exact and from the reconstruction, the relative errors in amplitude, and the errors in phase.
@@
 
-As discussed in App.~\ref{app:precpreviousapproaches}, the approximation $\{N:0,A:0\}$ corresponds to the leading-order formula~\eqref{eq:transferlocal}, which is close to the treatment of PhenomP~\cite{Hannam+13}, with the difference that this model uses effective Fourier-domain Euler angles instead of a time-domain modulation through merger and ringdown.
-The order of approximation $\{N:3,A:0\}$ is equivalent to formalism of Ref.~\cite{KCY14} with a stencil order of $N=3$, except that the latter formalism is based on an SPA representation of the P-frame waveform and limited to the inspiral, while our Fourier-domain approach is an attempt at covering the whole frequency band.
+As discussed in App.~\ref{app:precpreviousapproaches}, the approximation $\{N:0 | A:0\}$ corresponds to the leading-order formula~\eqref{eq:transferlocal}, which is close to the treatment of PhenomP~\cite{Hannam+13}, with the difference that this model uses effective Fourier-domain Euler angles instead of a time-domain modulation through merger and ringdown.
+The order of approximation $\{N:3 | A:0\}$ is equivalent to formalism of Ref.~\cite{KCY14} with a stencil order of $N=3$, except that the latter formalism is based on an SPA representation of the P-frame waveform and limited to the inspiral, while our Fourier-domain approach is an attempt at covering the whole frequency band.
 
@@
 For the solid curves in the top panel, transfer functions were computed from an FFT of the full time-domain precessing signal, while the dashed curves show the results of the Fourier-domain reconstruction techniques presented here.
-The columns show the orders of approximation $\{\Psi:0,A:0\}$, $\{\Psi:2,A:2\}$, $\{\Psi:2,A:2,\text{conv}\}$.
+The columns show the orders of approximation $\{N:0 | A:0 | \text{No Conv.}\}$, $\{N:3 | A:2 | \text{No Conv.}\}$, $\{N:3 | A:2 | \text{Conv.}\}$.
 In the right column, the region of transition between the perturbative treatment of Sec.~\ref{subsec:executivesummary} and the high-frequency treatment of Sec.~\ref{subsec:trigopoly} is shaded in grey.
@@
 using only one term with $a_{0,0} = 1$ in~\eqref{eq:stencilfresnel}.
-It is equivalent to the leading order of our formalism ($\{0, A:0, d:0 \}$ in the terminology of Sec.~\ref{subsec:executivesummary}), with the difference that we use a more general definition of $t_{f}$ (see~\eqref{eq:deftf}) that extends through the merger and ringdown.
+It is equivalent to the leading order of our formalism ($\{N:0 | A:0 | d:0 \}$ in the terminology of Sec.~\ref{subsec:executivesummary}), with the difference that we use a more general definition of $t_{f}$ (see~\eqref{eq:deftf}) that extends through the merger and ringdown.
 

### 2018-06-18 docs/paper_method.tex
Minor edits in appendicies.

@@
 
-In this Appendix, we summarize our conventions for the Wigner matrices and give a brief description of the construction of a precessing-frame directly from the waveform.
+In this appendix, we summarize our conventions for the Wigner matrices and give a brief description of the construction of a precessing-frame directly from the waveform.
 
@@
 In Sec.~\ref{sec:precession}, we considered only $\ell = 2$ and we restricted ourselves to the contributions of the precessing-frame mode $h_{22}^{\rm P}$.
-In that case, the relevant explicit expression for the $d$-matrix are
+In that case, the relevant explicit \jgb{expressions for the $d$-matrix elements} are
 \begin{subequations}
@@
 
-\section{Explicit expression for the stencil coefficients}
+\section{Explicit expression for the \jgb{Fresnel} stencil coefficients}
 \label{app:stencil}
@@
 
-In this Appendix, we give explicit expressions for the stencil coefficients entering~\eqref{eq:stencilfresnel}.
+In this appendix, we give explicit expressions for the stencil coefficients entering~\eqref{eq:stencilfresnel} \jgb{which approximates a Fresnel transform, as discussed in Sec.~\ref{subsec:fresneltransform}}.
 In this work we used only low order stencils with $N\leq 20$, and one can trivially invert of the system~\eqref{eq:stencilsystem}, as this operation has to be done only once.
@@
 
-In this Appendix, we give a more detailed overview of the treatment of precession in existing waveform models that generate signals directly in the Fourier domain.
+In this appendix, we give a more detailed overview of the treatment of precession in existing waveform models that generate signals directly in the Fourier domain.
 Our objective is not to give an exhaustive account of the various existing models, but rather to highlight how their treatment relates to ours.
@@
 
-In this Appendix, we present an argument for the applicability of the treatment of Section~\ref{sec:formalism} to a Fourier-domain waveform that already contains precession effects, thus combining our investigations of Secs.~\ref{sec:LISA} and~\ref{sec:precession}.
+In this appendix, we present an argument for the applicability of the treatment of Section~\ref{sec:formalism} to a Fourier-domain waveform that already contains precession effects, thus combining our investigations of Secs.~\ref{sec:LISA} and~\ref{sec:precession}.
 Let us first note that one possible approach would be to incorporate both the precession modulation and the LISA modulation and delay in a single $G(f,t)$ kernel.

### 2018-06-18 docs/paper_method.tex
Tiny addition to abstract.

@@
 
-We present a Fourier-domain based approach modulations and delays of gravitational wave signals, a problem which arises in two different contexts.
+We present a Fourier-domain based approach modulations and delays of gravitational wave signals, a problem which arises in \jgb{at least} two different contexts.
 For space-based detectors like LISA, the orbital motion of the detector introduces a time-dependency in the response of the detector, consisting of both a modulation and a varying delay.

### 2018-06-28 docs/paper_method.tex
First submitted version of the paper.

--- 2018-06-18 docs/paper_method.tex
@@
 \usepackage{enumerate}
-\usepackage{ulem}
 \usepackage{tabularx}
@@
 \newcommand{\nn}{\nonumber}
-%\newcommand{\hatk}{\hat{k}}
 \newcommand{\hatk}{k}
@@
 
-We present a Fourier-domain based approach modulations and delays of gravitational wave signals, a problem which arises in two different contexts.
+We present a Fourier-domain approach to modulations and delays of gravitational wave signals, a problem which arises in two different contexts.
 For space-based detectors like LISA, the orbital motion of the detector introduces a time-dependency in the response of the detector, consisting of both a modulation and a varying delay.
 In the context of signals from precessing spinning binary systems, a useful tool for building models of the waveform consists in representing the signal as a time-dependent rotation of a quasi-non-precessing waveform.
-In both cases, being able to compute transfer functions for these effects directly in the Fourier domain may enable performance improvements for data analysis applications by using fast frequency-domain waveforms.
-Our results generalize previous approaches based on the stationary phase approximation for inspiral signals, extending them by including delays and computing corrections beyond the leading order, while being applicable to the broader class of signals of inspiral-merger-ringdown waveforms.
-In the LISA case, we find that a leading-order treatment is accurate for high-mass and low-mass signals that are chirping enough, with reconstruction errors that can be reduced by consistently including the corrections we derived.
-By contrast, low-mass binary black holes, if far away from merger and slowly-chirping, cannot be handled by our perturbative formalism and we develop another approach for these systems.
-In the case of precessing binaries, we perform a limited exploration of the merger-ringdown range based on a toy model for the precession.
-We find that, depending on the post-merger frame velocity, deviations from leading order can become qualitatively large to the point of challenging our perturbative formalism, while affecting mainly subdominant mode contributions and giving rise to a limited unfaithfulness in the full waveform.
+In both cases, being able to compute transfer functions for these effects directly in the Fourier domain may enable performance gains for data analysis applications by using fast frequency-domain waveforms.
+Our results generalize previous approaches based on the stationary phase approximation for inspiral signals, extending them by including delays and computing corrections beyond the leading order, while being applicable to the broader class of inspiral-merger-ringdown signals.
+In the LISA case, we find that a leading-order treatment is accurate for high-mass and low-mass signals that are chirping fast enough, with errors consistently reduced by the corrections we derived.
+By contrast, low-mass binary black holes, if far away from merger and slowly-chirping, cannot be handled by this formalism and we develop another approach for these systems.
+In the case of precessing binaries, we explore the merger-ringdown range for a handful of cases, using a simple model for the post-merger precession.
+We find that deviations from leading order can give large fractional errors, while affecting mainly subdominant modes and giving rise to a limited unfaithfulness in the full waveform.
 Including higher-order corrections consistently reduces the unfaithfulness, and we further develop an alternative approach to accurately represent post-merger features.
@@
 With the unprecedented recent gravitational-wave detections of coalescencing binary black holes and binary neutron stars, announced by the LIGO-Virgo collaboration~\cite{LIGO-theevent-2016, LIGO-christmasevent-2016, LIGO-O1BBH-2016, LIGO-BNSevent-2017}, gravitational-wave astronomy has entered its observational era.
-As LIGO prepares for even more sensitive observation runs, and with the recent expansion of the ground-based detectors network with Virgo~\cite{Virgo} (and eventually also KAGRA~\cite{KAGRA} and LIGO-India~\cite{INDIGO}), observations of such compact object coalescences are expected at ever-increasing rate.
-
-Moreover, the European Space Agency has recently selected the Laser Interferometer Space Antenna (LISA)\cite{LISA17} to realize the ``Gravitational Universe'' science theme\cite{elisa13} as the 3rd large space mission of its Cosmic Vision program, with a tentative launch around 2034.
+As LIGO prepares for even more sensitive observation runs, and with the recent expansion of the ground-based detectors network with Virgo~\cite{Virgo} (and eventually also KAGRA~\cite{KAGRA} and LIGO-India~\cite{INDIGO}), observations of such compact object coalescences are expected at an ever-increasing rate.
+
+Moreover, the European Space Agency has recently selected the Laser Interferometer Space Antenna (LISA)~\cite{LISA17} to realize the ``Gravitational Universe'' science theme~\cite{elisa13} as the 3rd large space mission of its Cosmic Vision program, with a tentative launch around 2034.
+A technology demonstrator, LISA Pathfinder, has tested with great success some of the key mission technologies~\cite{LISAPathfinder2016, LISAPathfinder2018}.
 LISA will be able to detect and characterize, among several important gravitational wave source targets, comparable-mass binary black hole coalescences from cosmological distances over a wide range of masses.
-These will range from high-redshift observations of supermassive black hole binaries with $M\sim 10^{7} \Msol$ down to the observation of LIGO-type sources with $M\sim 10^{2} \Msol$~\cite{Sesana16}.
+These will range from high-redshift observations of supermassive black hole binaries with $M\sim 10^{7} \Msol$ down to the observation of LIGO-type sources with $M\sim 10-100 \Msol$~\cite{Sesana16}.
 
@@
 
-%Our challenge will be to accurately complete the representation of instrumental signals including precession effects and the LISA instrumental response and precession effects without sacrificing the computational efficiency enabled by these fast waveform templates.
-\SM{[rewrite that]}
-
-A complete representation of spin effects in fast IMR templates still remains a frontier of gravitational wave signal modelling.
+A complete representation of spin effects across parameter space in fast IMR templates still remains a frontier of gravitational wave signal modelling.
 In presence of misaligned spins, the system will endure precession of the orbital plane as it evolves, leading to modulations of the signal as seen by the observer~\cite{Apostolatos+94, Kidder95}.
@@
 
-%Prospective parameter recovery studies for LISA observations of comparable-mass binary black hole coalescences aim at characterizing the scientific performance of the future experiment.
-Although LISA~\cite{LISA17} is still at least a decade from launch, the ESA L3 mission is rapidly progressing into its early stages where mission design choices must be informed by science impact analysis.
-However, the recent progresses in waveform modelling, most notably the availability of fast and accurate IMR waveforms, have not yet made their way to the field of LISA data analysis studies.
-In the past, many such studies used post-Newtonian, inspiral-only models, and were limited to the cheaper Fisher-matrix formalism instead of a full Bayesian analysis (among numerous works, see e.g.
-\cite{Cutler97,LISAPE09}).
-Refs.~\cite{McWilliams+09,McWilliams+10,McWilliams+11} explored the effects of including the merger and ringdown, keeping the mass parameters fixed.
-However, the Fisher matrix analysis, although considerably cheaper to perform, is valid only in the limit of a high signal to noise ratio (SNR).
-Several implementations of Bayesian analyses were developed~\cite{CP06,Wickham+06,Babak+08,GP09,Feroz+09,Petiteau+10,PC13} (see also~\cite{MLDC4} and references therein), with Ref.~\cite{PC15} proposing a direct comparison of Fisher and Bayes analyses, but they were limited to inspiral-only waveforms.
-One exception was Ref.~\cite{Babak+08} where IMR signals were used in a Bayesian study, but there the masses of the black holes were kept fixed.
-In the recent Ref.~\cite{Klein+15}, the limitation to inspiral-only precessing post-Newtonian waveforms was compensated by an SNR reweighting procedure.
-Thus, due mainly to their high computational cost, full-fledged Bayesian parameter estimation simulations for IMR signals are still missing.
-\jgb{[Maybe there is too much detail in this paragraph?]}
-
-%Prospective parameter recovery studies for LISA observations of comparable-mass binary black hole coalescences aim at characterizing the scientific performance of the future experiment.
-Although LISA~\cite{LISA17} is still at least a decade from launch, the ESA L3 mission is rapidly progressing into its early stages where mission design choice must be informed by science impact analysis.
-However, the recent progresses in waveform modelling, most notably the availability of cheap IMR waveforms, have not yet made their way to the field of LISA data analysis studies.
-In the past, such studies used mainly post-Newtonian, inspiral-only models, and were limited to the Fisher-matrix formalism instead of full Bayesian analysis (among numerous works, see e.g.
-\cite{Cutler97,LISAPE09}).
-Refs.~\cite{McWilliams+09,McWilliams+10,McWilliams+11} explored the effects of including the merger and ringdown, keeping the mass parameters fixed.
-However, the Fisher matrix analysis, although considerably cheaper to perform, is valid only in the limit of a high signal to noise ratio (SNR) and has known shortcomings~\cite{Vallisneri08}.
-Several implementations of Bayesian analyses were developed~\cite{CP06,Wickham+06,Babak+08,GP09,Feroz+09,Petiteau+10,PC13} (see also~\cite{MLDC4} and references therein), with Ref.~\cite{PC15} proposing a direct comparison of Fisher and Bayes analyses, but they were limited to inspiral-only waveforms.
-One exception was Ref.~\cite{Babak+08} where IMR signals were used in a Bayesian study, but there the masses of the black holes were kept fixed.
-In the recent Ref.~\cite{Klein+15}, the limitation to inspiral-only precessing post-Newtonian waveforms was compensated by an SNR reweighting procedure.
-Thus, due mainly to their high computational cost, full-fledged Bayesian parameter estimation simulations for IMR signals are still missing.
-\jgb{[Maybe there is too much detail in this paragraph?]}
-
-Whereas LIGO and Virgo are typically sensitive to chirping binaries for a minute or less, LISA signals may accumulating over months or years.
+Whereas LIGO and Virgo are typically sensitive to chirping binaries for a minute or less, LISA signals may accumulate over months or years.
 The response of a LISA-type instrument is thus time-dependent~\cite{Cutler97}.
@@
 These effects then convey information about the localization of the gravitational-wave source in the sky.
-Direct time-domain implementation of the detector response is straightforward~\cite{Vallisneri04, Petiteau+08, CR02, RCP04}, but at a prohibitive computational cost for parameter-estimation analyses.
+Direct time-domain implementation of the detector response is straightforward~\cite{Vallisneri04, Petiteau+08, CR02, RCP04}, but at a high computational cost for parameter-estimation analyses.
 To leverage the performance of state-of-the-art Fourier-domain IMR templates~\cite{BTB16,Khan+15}, we must efficiently process the signals through the time-dependent response of the dectector while staying in the Fourier domain.
@@
 
-In the case of precessing binaries, applying the SPA directly to the modulated signal is also prone to pathologies.
+In the case of precessing binaries, applying the SPA directly to the modulated signal is prone to pathologies.
 In Refs.~\cite{KCY13,KCY14}, a formalism (called shifted uniform asymptotics or SUA) was introduced to go beyond the SPA and compute more accurately the modulation in the Fourier domain; however, this formalism still relies on the SPA for the underlying precessing-frame signal, and is as such limited to inspiraling signals.
@@
 Ref.~\cite{Klein+15} included the orbital motion of the detector in the SUA treatment of precessing inspiral signals, within a low-frequency approximation of the constellation response.
-Consistently extending these previous approaches to the merger-ringdown part of the signals, while including the delays in the full LISA response at all frequencies, is part of the objectives of this paper.
+Consistently extending these previous approaches to the merger-ringdown part of the signals, while including the delays in the full LISA response at all frequencies, assessing and understanding the errors made along the way, is part of the objectives of this paper.
 
@@
 In Sec.~\ref{sec:motivation}, we provide a general presentation of the problem of Fourier-domain modulation and introduce the relevant timescales for both the response of LISA-type detectors and the modulation of precessing signals.
-In Sec.~\ref{sec:formalism}, we present our general formalism, give its leading order approximation as well as higher-order corrections, introduce new timescales based on the Fourier-domain signal, refine the previous results for both the quadratic-in-phase corrections and the treatment of the delays, and make the connection to previous approaches of the problem.
+In Sec.~\ref{sec:formalism}, we present our general formalism, give its leading order approximation as well as higher-order corrections, introduce new timescales based on the Fourier-domain signal, and refine the previous results for both the quadratic-in-phase corrections and the treatment of the delays.
 We then apply our formalism to the response of the LISA detector in Sec.~\ref{sec:LISA}, and to the case of signals from precessing binaries in Sec~\ref{sec:precession}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\subsection{Basic considerations}
-%\label{subsec:basic}
 
@@
 An important consequence is that these signal components can be accurately represented in terms of a relatively coarsely sampled frequency grid.
-This property will also extend to the transfer functions, allowing us to keep a compact representation of the full signals.
+This property will also extend to the transfer functions, allowing one to keep a compact representation of the full signals.
 (See Appendix~\ref{app:notation} for details of our notation and conventions for Fourier transforms and spherical harmonics.)
@@
 \end{align}
-Note that to be able to treat the amplitude as a constant in the integral above, we used the third condition in~\eqref{eq:conditionSPA}.
+Note that to be able to treat the amplitude as a constant in the integral above, we used the third condition in~\eqref{eq:conditionsSPA}.
 The resulting complex Gaussian integral yields\footnote{The expressions given here are valid for the dominant mode $h_{22}$ of the waveform.
-They can be generalized to other modes $h_{\ell m}$, $m\neq 0$ with phase $e^{-im\varphi}$ as follows: $\Psi$ acquires a factor $m/2$, $A$ a factor $\sqrt{2/m}$, and $v$ (see~\eqref{eq:SPAN}) has to be replaced by $(2/m)^{1/3} v$.}
+They can be generalized to other modes $h_{\ell m}$, $m\neq 0$ with phase $e^{-im\varphi}$ as follows: $t_{f}^{\rm SPA}$ is now such that $m\omega = 2\pi f$, $A$ acquires a factor $\sqrt{2/m}$, in $\Psi$ the term $2\varphi$ becomes $m\varphi$.}
 \begin{subequations}
@@
 \end{subequations}
-where $v=(G M \pi f/c^{3})^{1/3}$ according to the SPA correspondence~\eqref{eq:deftfSPA}, and where $t_{0}, \phi_{0}$ are constants.
+where $v=(G M \pi f/c^{3})^{1/3}$ according to the SPA correspondence~\eqref{eq:deftfSPA}, and where $t_{0}, \phi_{0}$ are constants\footnote{For modes $h_{\ell m}$ with $m\neq 0$, the phase $\Psi$ acquires a factor $m/2$ and has to be evaluated at $(2/m)^{1/3} v$.}.
 It is also customary to rewrite the above relations in terms of the chirp mass $\Mchirp \equiv M\nu^{3/5}$, which is the only mass combination characterizing the signal at the leading PN order.
@@
 Near and after the merger, the SPA treatment is not applicable.
-%, as can be seen by the breakdown of the relation~\eqref{eq:deftfSPA} as $\omega$ asymptotes the quasi-normal mode frequency in the ringdown.
-Because of this limitation, we aim at developing a formalism that would not rely on this approximation.
 Nonetheless, the SPA has been a useful workhorse in many gravitational analyses involving frequency domain transfer functions of the form~\eqref{eq:deftransfer}.
@@
 \ee
-where here and in the rest of the paper, such derivatives with respect to the frequency will always be evaluated at $f$. Plugging this relation into~\eqref{eq:FDkernel}, we obtain
+Plugging this relation into~\eqref{eq:FDkernel}, we obtain
 \begin{align}
@@
 \ee
-However, our defintion~\eqref{eq:deftf} refers only to the Fourier-domain waveform.
-We do not need to relate the frequency $f$ to a time-domain frequency like the orbital frequency $\omega$, and the defintion is independent of the SPA being valid or not for the underlying signal $\tilde{h}$.
-
-The main advantage of our time-of-frequency function~\eqref{eq:deftf} is that it extends naturally to the merger-ringdown part of the signals.
+However, $\tf$ refers only to the Fourier-domain waveform.
+We do not need to relate the frequency $f$ to a time-domain frequency like the orbital frequency $\omega$, and the definition is independent of the SPA being valid or not for the underlying signal $\tilde{h}$.
+
+The main advantage of the time-of-frequency function~\eqref{eq:deftf} is that it extends naturally to the merger-ringdown part of the signals.
+As such, it is used in the PhenomD and PhenomHM waveform models~\cite{Khan+15, London+17}.
 Fig.~\ref{fig:tf} shows the behaviour of the time function $\tf$ around merger for six example waveforms, for mass ratios $q=1$ and $q=8$ and for aligned spin components $\chi=0.95,0.,-0.95$.
@@
   \centering
-  %\includegraphics[width=.99\linewidth]{plots/TfTA.pdf}
   \includegraphics[width=.98\linewidth]{plots/TfTA_py.pdf}
@@
 The hierarchy of these timescales is roughly the same for higher mass ratio and higher spin systems.
-The time-domain frequency at merger is represented by the vertical line.
+The time-domain frequency at merger and the ringdown frequency are represented by the thick and thin vertical lines.
 The derivatives from which these timescales are built encounter zero-crossings and change sign in the high-frequency range.}
@@
 
-However, in the same way that our definition~\eqref{eq:deftf} for the time-of-frequency function $t_{f}$ generalizes the SPA definition~\eqref{eq:deftfSPA}, the definition~\eqref{eq:defTf} only refers to the phase of the Fourier-domain signal and does not require introducing a time-domain frequency like $\omega$.
+However, in the same way that the definition~\eqref{eq:deftf} for the time-of-frequency function $t_{f}$ generalizes the SPA definition~\eqref{eq:deftfSPA}, the definition~\eqref{eq:defTf} only refers to the phase of the Fourier-domain signal and does not require introducing a time-domain frequency like $\omega$.
 This defintion thus extends naturally to the merger-ringdown part of the signal.
@@
 With a simple power-law amplitude as in~\eqref{eq:ASPAN}, all higher-order amplitude timescales are also simply proportional to $1/f$ and differ only by their numerical factor.
-For higher harmonics $h_{\ell m}$ with $m\neq 0$, as discussed in Sec.~\ref{subsec:SPA}, $T_{f}$ acquires a factor $m/2$, and $v$ has to be replaced by $(2/m)^{1/3} v$.
-The amplitude of higher harmonics starts at a higher PN order~\cite{BlanchetLiving}, differing from~\eqref{eq:a22N} by a different constant and an additional scaling $v^{\kappa_{\ell m}}$ with $\kappa_{\ell m} = \ell - 2 + (\ell + m \; \mathrm{mod} \; 2)$, which changes the constant in the amplitude timescale as $7/6 \rightarrow (7-2\kappa_{\ell m})/6$ in $T_{A1}^{\rm N}$ and $\sqrt{91}/6 \rightarrow \sqrt{(7-2\kappa_{\ell m})(13-2\kappa_{\ell m})}/6$ in $T_{A2}^{\rm N}$ \SM{[check]}.
+For higher harmonics $h_{\ell m}$ with $m\neq 0$, as discussed in Sec.~\ref{subsec:SPA}, $T_{f}$ acquires a factor $(m/2)^{4/3}$.
+The amplitude of higher harmonics starts at a higher PN order~\cite{BlanchetLiving}, differing from~\eqref{eq:a22N} by a different constant and an additional scaling $v^{\kappa_{\ell m}}$ with $\kappa_{\ell m} = \ell - 2 + (\ell + m \; \mathrm{mod} \; 2)$, which changes the constant in the amplitude timescale as $7/6 \rightarrow (7-2\kappa_{\ell m})/6$ in $T_{A1}^{\rm N}$ and $\sqrt{91}/6 \rightarrow \sqrt{(7-2\kappa_{\ell m})(13-2\kappa_{\ell m})}/6$ in $T_{A2}^{\rm N}$.
 
@@
 In practice, $N$ will be most often less than $5$, although we will consider in one case $N=10$ and $20$.
-In~\cite{KCY14}, only the case $\epsilon=1$ was considered.
+In~\cite{KCY14}, only the case $\epsilon=1$ was needed, as inspiral signals were considered.
 The two solutions for the stencil coefficients in the two cases $\epsilon = \pm 1$ are simply related by a complex conjugation.
@@
 
-%Note however that this quadrature rule does not fall into the class of Gaussian quadratures, due to the complex character of the kernel in~\eqref{eq:defFresnel}.
-The induced inner product is not positive definite and some key properties are lost, such as the construction of the optimal Gaussian nodes from the zeros of the orthonormalized polynomials.
-For more details on quadrature rules when applied to integrals with oscillatory kernels, we refer to~\cite{} \SM{[C]}.
-
 Note that the choice of a stencil with quadrature nodes $\tf \pm k\Tf$, even if natural, is by no means unique, and other choices would have led to different stencils.
 The formulation of the result~\eqref{eq:resultdffPsiTf} as a Fresnel transform~\eqref{eq:resultFresnel} opens the way for future investigations of different numerical approaches to the problem.
-For practical computations, in this work we will rely on the approximations $\calF_{T, \epsilon}^{N}[F] (t)$ in \eqref{eq:stencilfresnel}.
+For practical computations, in this work we will rely on the approximations $\calF_{T, \epsilon}^{N}[F] (t)$ in \eqref{eq:stencilfresnel}, introduced by~\cite{KCY14}.
 
@@
 For the proposed LISA configuration~\cite{LISA17}, we have for the orbital delays the scaling $d_{0}\sim R/c \simeq 500s$, and for the constellation delays the scaling $d_{L}\sim L/c \simeq 8s$ (ignoring the dependence on angular factors).
-Since the motion of the constellation is anually periodic, with a frequency $\Omega_{0} \simeq 2 \times 10^{-7}\mathrm{rad}.s^{-1}$, we have $\dot{d}_{0} \sim \Omega_{0} R/c \simeq 10^{-4}$ and $\dot{d}_{c} \sim \Omega_{0} L/c \simeq 1.7 \times 10^{-6}$.
+Since the motion of the constellation is anually periodic, with a frequency $\Omega_{0} \simeq 2 \times 10^{-7}\mathrm{rad}.s^{-1}$, we have $\dot{d}_{0} \sim \Omega_{0} R/c \simeq 10^{-4}$ and $\dot{d}_{L} \sim \Omega_{0} L/c \simeq 1.7 \times 10^{-6}$.
 The smallness of the dimensionless quantity $\dot{d} \ll 1$ (and of its subsequent derivatives) will allow us to treat it perturbatively with a very good approximation, and shows also that the function $t_{d}$ is univalued and that there is no ambiguity in defining the reciprocal $t_{d}^{-1}$.
@@
 \end{align}
-Now, the most relevant correction in~\eqref{eq:delaycorrleading} comes from the phase factor at high frequencies, where the factors $2\pi f d_{0}$ and $2\pi f d_{L}$ give a magnification factors reaching respectively $3.10^{3}$ and $10^{2}$ at $1\Hz$.
+Now, the most relevant correction in~\eqref{eq:delaycorrleading} comes from the phase factor at high frequencies, where the factors $2\pi f d_{0}$ and $2\pi f d_{L}$ give a magnification reaching respectively $3.10^{3}$ and $10^{2}$ at $1\Hz$.
 Ignoring the other corrections, we thus arrive at the following form for the dominant delay correction in the transfer function:
@@
 
-Next, we consider the case where the quadratic phase correction is kept as well, as in Sec.~\ref{subsec:fresneltransform}.
-Keeping as before only the first-order terms in $\dot{d}$ (and neglecting its higher derivatives), we can write
+Next, we consider the case where the quadratic phase correction is kept as well, as in Sec.~\ref{subsec:fresneltransform}, and where we keep all the first-order terms in $\dot{d}$ (neglecting its higher derivatives).
+We can write
 \begin{widetext}
@@
 Finally, when considering amplitude corrections as well, as in~\eqref{eq:resulttaylorA}, additional powers of $f'$ can be translated as time derivatives with respect to the variable $\tau$ after performing the change of variables.
-This produces the final result:
+This produces the result:
 \be\label{eq:transferfinal}
@@
 In this Section, we apply the formalism of Sec.~\ref{sec:formalism} to the Fourier-domain response of a LISA-like detector, and assess the accuracy of our approach at various levels of approximation.
-We also propose an alternative approach for slowly-chipring signals.
+We also propose an alternative approach for slowly-chirping signals.
 
@@
 Since our aim is to assess the accuracy of our direct Fourier-domain treatment of the response, we can focus on the gravitational-wave contribution to the basic single-link observables.
-We therefore use a somewhat simplified model for the time-domain response~\cite{Krolak+04}, ignoring corrections that are crucial from the point of view of noise cancellations, keeping in mind that the response model can be enriched later without affecting the conclusions of the present analysis.
+We therefore use a somewhat simplified model for the time-domain response~\cite{Krolak+04}, ignoring corrections that would be crucial from the point of view of noise cancellations, keeping in mind that the response model can be enriched later without affecting the conclusions of the present analysis.
 
 The frequency-shift response for a single link~\eqref{eq:yslr} was derived in Ref.~\cite{EW75} (see also~\cite{CR02, RCP04, Finn08, Cornish09}).
-Several assumptions enter the result as written in~\eqref{eq:yslr}: (i) effects of the order $v/c$ are neglected, including for instance the special relativistic Doppler effect created by the relative speeds of the spacecraft on their orbits (ii) the propagation is assumed to take place in a flat spacetime, perturbed only by the gravitational wave; thus the gravitational redshift as well as the deflection of light created by the gravitational potential of the Sun is ignored (iii) all geometric factors are evaluated at a single time, whereas one should consider the beam as propagating from the position of the first spacecraft at the time of emission to the position of the second scapecraft at the time of reception, leading to a pointing-ahead effect.
+Several assumptions enter the result as written in~\eqref{eq:yslr}: (i) effects of the order $v/c$ are neglected, including for instance the special relativistic Doppler effect created by the relative speeds of the spacecraft on their orbits (ii) the propagation is assumed to take place in a flat spacetime, perturbed only by the gravitational wave; thus the gravitational redshift as well as the deflection of light created by the gravitational potential of the Sun is ignored (iii) all geometric factors are evaluated at a single time, whereas one should consider the beam as propagating from the position of the first spacecraft at the time of emission to the position of the second scapecraft at the time of reception, leading to a point-ahead effect.
 
@@
 \end{subequations}
-The LISA response will behave qualitatively differently on the three frequency bands $f \lesssim f_{R}$, $f_{R} \lesssim f \lesssim f_{L}$ and $f_{L} \lesssim f$.
+The LISA response will behave qualitatively differently on the three frequency bands $f \leq f_{R}$, $f_{R} \leq f \leq f_{L}$ and $f_{L} \leq f$.
 
@@
 	F_{slr}^{L}(t) &= \frac{1}{2} \frac{1}{1 - \hatk\cdot n_{l}(t)} n_{l}(t) \cdot P_{\ell m} \cdot n_{l} (t) \,,\nn\\
-	d_{s,L}(t) &= -L - k\cdot p_{s}^{L}(t) \,, \quad d_{r}(t) = - k\cdot p_{r}^{L}(t) \,,\nn\\
-	G_{slr}^{L}(f,t) &=  F_{slr}^{L}(t) \left( e^{-2i\pi f d_{s,L}(t)} - e^{-2i\pi f d_{r}(t)} \right) \,.
+	d_{s}(t) &= - k\cdot p_{s}^{L}(t) \,, \quad d_{r}(t) = - k\cdot p_{r}^{L}(t) \,,\nn\\
+	G_{slr}^{L}(f,t) &=  F_{slr}^{L}(t) \left( e^{-2i\pi f (d_{s,L}(t) + L)} - e^{-2i\pi f d_{r}(t)} \right) \,.
 \end{align}
@@
 \end{align}
-with all time-dependent vectors evaluated at $t$. In the local approximation~\eqref{eq:transferlocal}, the Fourier-domain transfer function then reads
+with all time-dependent vectors evaluated at $t$.
+This expression is well known as describing the frequency-dependency in the LISA response~\cite{Larson+99, Cornish01, CR02, RCP04}.
+In the local approximation~\eqref{eq:transferlocal}, the Fourier-domain transfer function then reads
 \begin{align}\label{eq:transferLlocal}
@@
 Note that, if the corrections of Sec.~\ref{subsec:delays}, for non-negligible $\dot d$, are included for the constellation delays, the transfer function will not have this simple form anymore, as $\dot{d}$ will have a different velocity-dependent expression for the sending and receiving spacecraft.
-One must then separately handle $d_{s,L}$ and $d_{r}$ in~\eqref{eq:decomposeGslr} to compute the corrections.
+One must then separately handle $d_{s}$ and $d_{r}$ in~\eqref{eq:decomposeGslr} to compute the corrections.
 
@@
   \centering
-  \includegraphics[width=.9\linewidth]{plots/lisafom_Psi_McDeltat.pdf}
-  \caption{\SM{[To be checked and redone in matplotlib]} \SM{[remove grey shading]} \SM{[add labels to contours]} Contour levels for the analytical estimate of the error measure $\epsilon_{\Psi 2}$ at the starting frequency, as a function of chirp mass $\Mchirp$ and time to coalescence $\Delta t$.
+  \includegraphics[width=.99\linewidth]{plots/lisafom_Psi_McDeltat_py.pdf}
+  \caption{Contour levels for the analytical estimate of the error measure $\epsilon_{\Psi 2}$ at the starting frequency, as a function of chirp mass $\Mchirp$ and time to coalescence $\Delta t$.
 Blue corresponds to the orbital response and red to the constellation response.
@@
   \centering
-%  \includegraphics[width=.32\linewidth]{plots/lisafom_Psiorb.pdf}
-%  \includegraphics[width=.32\linewidth]{plots/lisafom_A1orb.pdf}
-%  \includegraphics[width=.32\linewidth]{plots/lisafom_dorb.pdf}
-%  %
-%  \includegraphics[width=.32\linewidth]{plots/lisafom_Psiconst.pdf}
-%  \includegraphics[width=.32\linewidth]{plots/lisafom_A1const.pdf}
-%  \includegraphics[width=.32\linewidth]{plots/lisafom_dconst.pdf}
   \includegraphics[width=.98\linewidth]{plots/lisafom_py.pdf}
@@
 To obtain an order-of-magnitude estimate for the error measures $\epsilon$~\eqref{eq:deffom}, we will use the Newtonian-order expressions~\eqref{eq:timescalesN} for the signal-dependent timescales $\Tf$ and $T_{A1}$.
-Higher-order amplitude terms beyond the first one in~\eqref{eq:resulttaylorA}, as well as phase terms beyond the second derivative in~\eqref{eq:expandPsi}, will turn out to be  negligible in the LISA case and we will ignore them in the following.
+Higher-order amplitude terms beyond the first one in~\eqref{eq:resulttaylorA}, as well as phase terms beyond the second derivative in~\eqref{eq:expandPsi}, will be negligible and we will ignore them in the following.
 It is useful to separate the orbital response~\eqref{eq:transfer0local} and the constellation response~\eqref{eq:transferLlocal}, as the different baseline of the delays (orbital radius $R$ or armlength $L$) as well as the presence of a time-varying prefactor $F(t)$ both affect the result.
@@
 These scalings are only approximate as different orientation angles can lead to significant variations.
-To represent the average of the geometric projection factor of the gravitational wave propagation vector on the plane of the orbit, we will assume $d_{0} \sim R_{\rm eff} \equiv 2R/\pi $.
+To represent the average of the geometric projection factor of the gravitational wave propagation vector on the plane of the orbit, we will take $d_{0} \sim R/2 $.
 For the constellation delays, we simply take $d_{L} \sim L$ as there projection effects can lead to variations in both ways.
@@
 \begin{align}
-	\left| \frac{1}{G_{0}} \partial_{t} G_{0} \right| &\sim 2 \pi f \Omega_{0} R_{\rm eff}\,,\\
-	\left| \frac{1}{G_{0}} \partial_{tt} G_{0} \right| &\sim \text{max} \left[ 2 \pi f \Omega_{0}^{2} R_{\rm eff} , 4\pi^{2} f^{2} \Omega_{0}^{2} R_{\rm eff}^{2}\right] \,,\\
-	\left| \frac{1}{G_{0}} \partial_{tf} G_{0} \right| &\sim \text{max} \left[ 2 \pi  \Omega_{0} R_{\rm eff}, 4 \pi^{2} f \Omega_{0} R_{\rm eff}^{2} \right] \,,
+	\left| \frac{1}{G_{0}} \partial_{t} G_{0} \right| &\sim \pi f \Omega_{0} R\,,\\
+	\left| \frac{1}{G_{0}} \partial_{tt} G_{0} \right| &\sim \text{max} \left[ \pi f \Omega_{0}^{2} R, \pi^{2} f^{2} \Omega_{0}^{2} R^{2}\right] \,,\\
+	\left| \frac{1}{G_{0}} \partial_{tf} G_{0} \right| &\sim \text{max} \left[ \pi  \Omega_{0} R, \pi^{2} f \Omega_{0} R^{2} \right] \,,
 \end{align}
@@
 The massive black hole binaries (MBH) that will be observed by LISA~\cite{LISA17} fall within the first category, with their merger in band, while the proposed population of stellar-origin black hole binaries (SOBH)~\cite{Sesana16}, with masses comparable to the LIGO/Virgo detections, will comprise both merging binaries, i.e.
-exiting the LISA band towards larger frequencies during observations, and slowly-chirping binaries hundereds or thousands of years away from merger.
+exiting the LISA band towards larger frequencies during observations, and slowly-chirping binaries hundreds or thousands of years away from merger.
 
@@
 The $\epsilon_{d}$ is signal-independent, and follows a quite different behaviour for the orbital and constellation responses.
-For the orbital response, $\epsilon_{d_{0}}$ grows for higher frequencies, reaching the $\sim 0.1$ at around $1\Hz$.
-For the constellation response, $\epsilon_{d_{L}}$ is mainly important at lower frequencies, reaching $\sim 0.01$ at around $10^{-5}\Hz$.
+For the orbital response, $\epsilon_{d}^{0}$ grows for higher frequencies, reaching $\sim 0.1$ at around $1\Hz$.
+For the constellation response, $\epsilon_{d}^{L}$ is mainly important at lower frequencies, reaching $\sim 0.01$ at around $10^{-5}\Hz$.
 
@@
 We will also focus on the orbital response and on the single-arm constellation response $y_{132}$, as TDI combinations will appear (within the approximations listed in Sec.~\ref{subsec:modelLISA}) as linear combinations of these basic observables.
-The orientation angles chosen for these examples are $[\lambda,\beta,\iota,\psi] = [\pi/4,\pi/3,\pi/3,0]$.
+The orientation angles chosen for these examples are $[\lambda,\beta,\iota,\psi] = [\pi/4,\pi/3,\pi/3,0]$ for the ecliptic longitude, latitude, inclination and polarization respectively.
 
@@
   \centering
-%  \includegraphics[width=.8\linewidth]{plots/LISAtransferM1e7dOamp.pdf}
-%  \hspace{0.1cm}
-%  \includegraphics[width=.8\linewidth]{plots/LISAtransferM1e7dOphase.pdf}
-%  \hspace{0.1cm}
-%  \includegraphics[width=.8\linewidth]{plots/LISAerrorM1e7dO.pdf}
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e7dO_py.pdf}
@@
   \centering
-%  \includegraphics[width=.8\linewidth]{plots/LISAtransferM1e2dOamp.pdf}
-%  \hspace{0.1cm}
-%  \includegraphics[width=.8\linewidth]{plots/LISAtransferM1e2dOphase.pdf}
-%  \hspace{0.1cm}
-%  \includegraphics[width=.8\linewidth]{plots/LISAerrorM1e2dO.pdf}
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e2dO_py.pdf}
@@
   \centering
-  %\includegraphics[width=.98\linewidth]{plots/LISAtransferM1e7y12LReIm.pdf}
-  %\includegraphics[width=.98\linewidth]{plots/LISAerrorM1e7y12L.pdf}
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e7y12L_py.pdf}
@@
   \centering
-  %\includegraphics[width=.98\linewidth]{plots/LISAtransferM1e2y12LReIm.pdf}
-  %
-  %\includegraphics[width=.98\linewidth]{plots/LISAerrorM1e2y12L.pdf}
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e2y12L_py.pdf}
@@
   \centering
-%  \includegraphics[width=.8\linewidth]{plots/LISAtransferSOBHepsPsi1d0ampcorr.pdf}
-%  \hspace{0.1cm}
-%  \includegraphics[width=.8\linewidth]{plots/LISAtransferSOBHepsPsi1d0phasecorr.pdf}
-%  \hspace{0.1cm}
-%  \includegraphics[width=.8\linewidth]{plots/LISAerrorSOBHepsPsi1d0.pdf}
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi1d0corr_py.pdf}
-  \caption{Transfer function and residual for the orbital delay $d_{0}$~\eqref{eq:delay0}, for an SOBH with $\epsilon_{\Psi 2}\sim 1$ for the orbital response.
+  \caption{Transfer function and residual for the orbital delay $d_{0}$~\eqref{eq:delay0}, for an SOBH with $\epsilon_{\Psi 2}\sim 0.6$ for the orbital response.
 The frequency band corresponds to 4 years of observation for this system with $M=50\Msol$, $\Delta t = 200\yr$ away from merger.
@@
   \centering
-%  \includegraphics[width=.8\linewidth]{plots/LISAtransferSOBHepsPsi10d0ampcorr.pdf}
-%  \hspace{0.1cm}
-%  \includegraphics[width=.8\linewidth]{plots/LISAtransferSOBHepsPsi10d0phasecorr.pdf}
-%  \hspace{0.1cm}
-%  \includegraphics[width=.8\linewidth]{plots/LISAerrorSOBHepsPsi10d0.pdf}
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi10d0corr_py.pdf}
-  \caption{Same as Fig.~\ref{fig:LISAerrorSOBHepsPsi1d0}, but for an SOBH with $\epsilon_{\Psi 2}\sim 7$ for the orbital response.
+  \caption{Same as Fig.~\ref{fig:LISAerrorSOBHepsPsi1d0}, but for an SOBH with $\epsilon_{\Psi 2}\sim 4.5$ for the orbital response.
 The frequency band corresponds to 4 years of observation for this system with $M=15\Msol$, $\Delta t = 1500\yr$ away from merger.
@@
   \centering
-%  \includegraphics[width=.8\linewidth]{plots/LISAtransferSOBHepsPsi1y12LReIm.pdf}
-%  \hspace{0.1cm}
-%  \includegraphics[width=.8\linewidth]{plots/LISAerrorSOBHepsPsi1y12L.pdf}
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi1y12L_py.pdf}
@@
   \centering
-%  \includegraphics[width=.8\linewidth]{plots/LISAtransferSOBHepsPsi10y12LReIm.pdf}
-%  \hspace{0.1cm}
-%  \includegraphics[width=.8\linewidth]{plots/LISAerrorSOBHepsPsi10y12L.pdf}
   \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi10y12L_py.pdf}
@@
 They are far from merger, slowly chirping, and span only a narrow frequency band over the course of the LISA mission lifetime, although being less monochromatic that galactic binaries.
-For the galactic binaries, an accurate and efficient numerical treatment of the response has been proposed~\cite{CL07} and widely used in applications.
+For the galactic binaries, an accurate and efficient numerical treatment of the response has been proposed and widely used in applications~\cite{CL07}.
 This treatment is referred to as the fast-slow decomposition, or heterodyning approach.
@@
 We now illustrate this approach by considering two examples of equal-mass SOBH systems.
-The first has $M = 50\Msol$ and is $\Delta t = 200\yr$ away from merger, and for the orbital response $\epsilon_{\Psi 2} \simeq 1$.
-The second has $M = 15\Msol$ and is $\Delta t = 1500\yr$ away from merger, and for the orbital response $\epsilon_{\Psi 2} \simeq 7$.
-The phase error measures for the constellation part are smaller, $\epsilon_{\Psi_{2}} \simeq 0.1$ and $0.01$ respectively.
+The first has $M = 50\Msol$ and is $\Delta t = 200\yr$ away from merger, and for the orbital response $\epsilon_{\Psi 2} \simeq 0.6$.
+The second has $M = 15\Msol$ and is $\Delta t = 1500\yr$ away from merger, and for the orbital response $\epsilon_{\Psi 2} \simeq 4.5$.
+The phase error measures for the constellation response are smaller, $\epsilon_{\Psi_{2}} \simeq 0.1$ and $0.01$ respectively.
 In keeping with the previous section, in our presentation we separate the orbital and the constellation response, keeping in mind that in practice the full response is to be handled in one step.
@@
 We see that, contrarily to Sec.~\ref{subsec:errorsLISA}, the transfer function starts to develop more structure that just a phase contribution.
-The case $M=50\Msol$ with $\epsilon_{\Psi 2} \simeq 1$ shows that the leading-order treatment leads to relative errors of order $10\%$, while increasing $N$ to 5 or 20 brings the errors back to $\sim 10^{-4}$, with a marginal improvement from further amplitude and delay corrections.
-The case $M=50\Msol$ with $\epsilon_{\Psi 2} \simeq 7$, by contrast, is clearly outside the range of applicability of the perturbative formalism, as orders of approximation give errors of order $100\%$.
+The case $M=50\Msol$ with $\epsilon_{\Psi 2} \simeq 0.6$ shows that the leading-order treatment leads to relative errors of order $10\%$, while increasing $N$ to 5 or 20 brings the errors back to $\sim 10^{-4}$, with a marginal improvement from further amplitude and delay corrections.
+The case $M=50\Msol$ with $\epsilon_{\Psi 2} \simeq 4.5$, by contrast, is clearly outside the range of applicability of the perturbative formalism, as all orders of approximation give errors of order $100\%$.
 In both cases, the comb treatment performs well, at better than $10^{-4}$.
@@
 
-The dominant features of the the precessing-frame waveform, are approximtely reflection symmetric about the orbital plane (exactly valid for non-precessing systems, see~\eqref{eq:symmetryhlminusm}), implying
+The dominant features of the the precessing-frame waveform are approximately reflection symmetric about the orbital plane (exactly valid for non-precessing systems, see~\eqref{eq:symmetryhlminusm}), implying
 \be\label{eq:approxsymmetryhlminusm}
@@
 This point will become important when considering the high-frequency part of the signal in Sec.~\ref{subsec:trigopoly}.
-Note, however, that this statement might not be true anymore when including more modes in our model.
+Note, however, that this statement might not be true anymore when including more modes in our model, in particular modes $\ell \neq m$.
 
@@
 
-We want to be able to investigate signals with a long inspiral phase, beyond the range covered by numerical relativity simulations, and we want to include the merger and ringdown phase.
+We want to be able to investigate generic-spin signals with a long inspiral phase, beyond the range covered by numerical relativity simulations, and we want to include the merger and ringdown phase.
 We will use the following three ingredients for this simplified model:
@@
 
-For the precessing-frame waveform, we make all the simplifying assumptions described above in Sec.~\ref{subsec:precdef}: we approximate it as a non-precessing waveform, further enforcing the approximations~\eqref{eq:approxsymmetryhlminusm} and~\eqref{eq:approxzeronegativef} and limiting our analysis to the dominant harmonics\footnote{See~\cite{London+17} for a recent extension of the model to include higher harmonics.} $h^{\rm P}_{22}$ and $h^{\rm P}_{2,-2}$.
-We will use the PhenomD waveform model~\cite{Husa+15, Khan+15}, an aligned-spin Fourier-domain model publicly available in the LIGO Algorithm Library (LAL) that covers the inpiral, merger and ringdown for binaries with generic spin magnitude.
+For the precessing-frame waveform, we make all the simplifying assumptions described above in Sec.~\ref{subsec:precdef}: we approximate it as a non-precessing waveform, further enforcing the approximations~\eqref{eq:approxsymmetryhlminusm} and~\eqref{eq:approxzeronegativef}.
+Limiting our analysis to the dominant harmonics $h^{\rm P}_{22}$ and $h^{\rm P}_{2,-2}$, we will use the PhenomD waveform model~\cite{Husa+15, Khan+15}, an aligned-spin Fourier-domain model publicly available in the LIGO Algorithm Library (LAL) that covers the inpiral, merger and ringdown for binaries with generic spin magnitude\footnote{See~\cite{London+17} for a recent extension of the model to include higher harmonics.}.
 The amplitude and phase of the waveform are produced as piecewise analytical functions of the frequency.
@@
 To compute the QNM frequencies as a function of the final spin $\chi_{f}$, we use fits constructed in Ref.~\cite{Berti+05}, illustrated in Fig.~\ref{fig:QNM}.
-The final spin is taken to be the same as the one computed internally to the SEOBNRv3 code~\cite{Pan+13, BTB16}, where the final-spin fit formula of Ref.~\cite{BR09} \SM{[see if we can update the final-spin formula]}, built for spin-aligned systems, is applied to the spin components projected on the orbital angular momentum $\bm{L}$ at merger.
+The final spin is taken to be the same as the one computed internally to the SEOBNRv3 code~\cite{Pan+13, BTB16}, where the final-spin fit formula of Ref.~\cite{BR09}, built for spin-aligned systems, is applied to the spin components projected on the orbital angular momentum $\bm{L}$ at merger.
 
@@
 This example therefore serves the purpose of lying at the edge of validity of the picture of standard precession along a cone around an almost-fixed direction, and the standard hierarchy between mode contributions is not valid in this case.
-\SM{[soundness of trajectories, particularly in this case, are to be checked again -- discrepancy for $--$ between Jf from dynamics and chif used in SEOB internals]}
 
@@
   \centering
-%  \includegraphics[width=.45\linewidth]{plots/eulerSXS0058alphagamma.pdf}
-%  \hspace{0.2cm}
-%  \includegraphics[width=.45\linewidth]{plots/eulerSXS0058beta.pdf}
   \includegraphics[width=.98\linewidth]{plots/eulerSXS0058_py.pdf}
@@
 
-%\begin{figure*}
-%  \centering
-%  \includegraphics[width=.48\linewidth]{plots/prectoymodelalphagamma.pdf}
-%  \hspace{0.2cm}
-%  \includegraphics[width=.48\linewidth]{plots/prectoymodelbeta.pdf}
-%  \caption{Evolution of Euler angles $(\alpha, \beta, \gamma)$ for our two toy models.
-Dashed lines represent Case I and solid lines Case II.
-The vertical line indicates the time of merger.
-In the left panel the dashed black lines indicate the enforced asymptotic behaviour in Case II, where the frame rotates with constant angular velocity $\omega_{220} -\omega_{210}$.
-In the right panel, $\beta$ is asymptotically constant.}
-%  \label{fig:prectoymodel}
-%\end{figure*}
-
-%In Eq.~\eqref{eq:defmodulationprec} above, the modulation function $F$ evolves on the precessional timescale, which evolves throughout the inspiral.
-The separation between the precessional timescale and the orbital timescale will be crucial for our analysis.
-In the limit of low frequencies, we will see in Sec.~\ref{} below that, although these timescales become more and more separated, the decrease in the chirping rate gives raise to a corrective contribution that does not vanish in this limit.
-In the other limit, as the system gets close to merger, the separation of timescales becomes weaker, and we will explore in Sec.~\ref{} the application of our formalism to the merger and ringdown phase.
-
 \begin{table}[t]
 \begin{ruledtabular}\caption{Parameters of of the three example cases that we use to illustrate our formalism.
-The three cases differ by their spin alignment angles $\theta_{A} = \bm{\chi}_{A} \cdot \hat{\bm{L}}_{i}$, the spins being in the orbital plane, almost aligned and almost anti-aligned.
-The table gives also the final spins magnitudes (for SEOB as well as PhenomD for comparison), angles between the initial and final angular momenta $\theta(\bm{J}_{i}, \bm{J}_{f})$, as well as QNM frequencies and frame rotation velocity used for the post-merger extension.
-\SM{[To be updated with longer waveforms, 10Hz 10Msol.
-?]} \SM{[add misalignment angles $(\ell, \bm{J}_{i})$?]}}\label{tab:precparams}
-%\begin{tabular}{C{.2\columnwidth}C{.24\columnwidth}C{.15\columnwidth}}
+The three cases differ by their spin alignment angles $\theta_{A} = \bm{\chi}_{A} \cdot \hat{\bm{L}}_{i}$, the spins being almost aligned, in the orbital plane and almost anti-aligned.
+The table gives also the final spins magnitudes (for SEOB as well as PhenomD for comparison), angles between the initial and final angular momenta $\theta(\bm{J}_{i}, \bm{J}_{f})$, as well as QNM frequencies and frame rotation velocity used for the post-merger extension.}\label{tab:precparams}
 \begin{tabular}{ccccccc}\label{tab:precexamples}
@@
 
-%\begin{table}[t]
-%\begin{ruledtabular}\caption{Parameters of the three example cases that we use to illustrate our formalism.}\label{tab:precparams}
-%%\begin{tabular}{C{.2\columnwidth}C{.24\columnwidth}C{.15\columnwidth}}
-%\begin{tabular}{cccc}
-%	Case & pp & $++$ & $--$ \\
-%	\hline
-%	$q$ & $4.0$ & $4.0$ & $4.0$ \\
-%	$\chi_{1}$ & $0.95$ & $0.95$ & $0.95$ \\
-%	$\chi_{2}$ & $0.95$ & $0.95$ & $0.95$ \\
-%	$\theta_{1}$ & $\pi/2$ & $\pi/6$ & $5\pi/6$ \\
-%	$\theta_{2}$ & $\pi/2$ & $\pi/6$ & $5\pi/6$ \\
-%	$\phi_{2}$ & $0.0$ & $0.0$ & $0.0$ \\
-%	$\phi_{2}$ & $\pi/2$ & $\pi/2$ & $\pi/2$
-%\end{tabular}
-%\end{ruledtabular}
-%\end{table}
-
-%\begin{figure*}
-%  \centering
-%  \includegraphics[width=.48\linewidth]{plots/prectransferAcaseI.pdf}
-%  \hspace{0.2cm}
-%  \includegraphics[width=.48\linewidth]{plots/prectransferPsicaseI.pdf}
-%  %
-%  \includegraphics[width=.48\linewidth]{plots/prectransferAcaseII.pdf}
-%  \hspace{0.2cm}
-%  \includegraphics[width=.48\linewidth]{plots/prectransferPsicaseII.pdf}  
-%  \caption{Amplitude and phase of the Fourier-domain transfer fonctions $\calT_{22,2m'}$ as defined in~\eqref{eq:defTlmlmp} for individual mode contributions.
-The left panels show the amplitude and the right panels the phase.}
-%  \label{fig:prectransfer}
-%\end{figure*}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In the inspiral phase, we can obtain a qualitative picture of the timescales involved by using well-known leading-order post-Newtonian results.
-To simplify things further, we will consider a single-spin system, and use orbit-averaged precession equations.
+To simplify things further, we will consider a single-spin system, and use orbit-averaging to ignore nutation features in the normal to the orbital plane.
 In this configuration, both the orbital angular momentum $\bm{L}$ and the spin vector $\bm{S}$ undergo simple precession on a cone around the total angular momentum $\bm{J}$ (see e.g.~\cite{Apostolatos+94, Kidder95}), with an opening angle of the cone and a precession velocity $\Omega_{\rm prec}$ that vary only on the radiation reaction timescale.
@@
 \ee
-Since we neglect radiation reaction, $\bm{J} = \bm{L} + \bm{S}_{1}/c$ is treated a constant that we use to set the z-axis so that $\bm{J} = J \bm{e}_{z}$, and we decompose the spin in its aligned and perpendicular components as $\bm{S}_{1} = S_{1}^{z} \bm{e}_{z} + \bm{S}_{1}^{\perp}$.
+Since we neglect radiation reaction, $\bm{J} = \bm{L} + \bm{S}_{1}/c$ is treated as a constant that we use to set the z-axis so that $\bm{J} = J \bm{e}_{z}$, and we decompose the spin in its aligned and perpendicular components as $\bm{S}_{1} = S_{1}^{z} \bm{e}_{z} + \bm{S}_{1}^{\perp}$.
 At leading order, the precession equations read
@@
 Given our restrictive  assumptions of orbit-averaged leading-order PN and single-spin simple precession, the result will only be a crude order-of-magnitude estimate.
-Using the Fourier-domain leading-order timescales generalized for modes $h^{\rm P}_{\ell m'}$ in the text below~\eqref{eq:timescalesN} yields: \SM{[check]}
+Using the Fourier-domain leading-order timescales generalized for modes $h^{\rm P}_{\ell m'}$ in the text below~\eqref{eq:timescalesN} yields:
 \begin{subequations}\label{eq:precfomPN}
 \begin{align}
-	\epsilon_{\Psi 2} &= \frac{m'}{2} \frac{5\nu}{96} \Lambda^{2} \xi^{2} (m' \cos\beta - m)^{2} v^{-1} \,, \\
-	\epsilon_{A 1} &=  \frac{(7 - 2\kappa_{\ell m'})\nu}{6} \Lambda \xi |m' \cos\beta - m| v^{2} \,, \\
-	\epsilon_{A 2} &= \frac{(7 - 2\kappa_{\ell m'}) (13 - 2\kappa_{\ell m'}) \nu^{2}}{72} \Lambda^{2} \xi^{2} (m' \cos\beta - m)^{2} v^{4} \,,
-\end{align}
-\end{subequations}
-where $v = (G M \omega/c^{3})^{1/3}$ is related to the Fourier frequency $f$ by $v = (2/m')^{1/3} (GM \pi f/c^{3})^{1/3}$ for a mode $h^{\rm P}_{\ell m'}$.
+	\epsilon_{\Psi 2} &= \left(\frac{2}{m'}\right)^{\frac{2}{3}} \frac{5\nu}{96} \Lambda^{2} \xi^{2} (m' \cos\beta - m)^{2} v^{-1} \,, \\
+	\epsilon_{A 1} &= \left(\frac{2}{m'}\right)^{\frac{5}{3}} \frac{(7 - 2\kappa_{\ell m'})\nu}{6} \Lambda \xi |m' \cos\beta - m| v^{2} \,, \\
+	\epsilon_{A 2} &= \left(\frac{2}{m'}\right)^{\frac{10}{3}} \frac{(7 - 2\kappa_{\ell m'}) (13 - 2\kappa_{\ell m'}) \nu^{2}}{72} \nn\\
+	& \quad\quad\quad \cdot \Lambda^{2} \xi^{2} (m' \cos\beta - m)^{2} v^{4} \,,
+\end{align}
+\end{subequations}
+where $v = (GM \pi f/c^{3})^{1/3}$.
 It is worth noting that $\epsilon_{\Psi 2}$ has an overall frequency scaling of $v^{-1}$, formally at $-0.5$PN order, which means that the relative size of this correction grows towards smaller frequencies, away from merger.
-Remember however that the quantities $\epsilon$ are meant to measure errors relatively to the leading order, and the opening angle of the precession cone, giving the overall normalization for precession effects in the waveform, also goes to 0 as $v$ in that limit.
+Remember however that the quantities $\epsilon$ are meant to fractional errors, and the opening angle of the precession cone, giving the overall normalization for precession effects in the waveform, also goes to 0 as $v$ in that limit.
 The amplitude error estimates $\epsilon_{A1}$, $\epsilon_{A2}$, by contrast, have the more usual behaviour of PN corrections growing towards merger.
@@
 
-The overall $\nu$ scaling indicates a better separation of timescales for larger mass ratio systems.
+The overall $\nu$ scaling indicates a better separation of timescales when increasing the mass ratio away from equal mass.
 These expressions also show that, since the total angular momentum appears as a factor in $\Omega_{\rm prec}$, higher-order corrections will be larger for roughly spin-aligned systems than for anti-aligned spins, an effect that becomes significant for large mass ratios.
@@
 
-%The first two cases, $++$ and $\perp\perp$, have \SM{[come back to this after having checked trajectories]} the usual behaviour of precession on a cone, with a larger opening angle for $\perp\perp$ due to the larger in-plane spin components.
-After merger, the case $++$, with a lerger remnant spin, displays a much faster frame rotation $\Omega_{\rm frame}$ than the other cases, as shown by Table~\ref{tab:precexamples}.
-The case $--$ is qualitatively different, close to the transitional precession regime, with a large and varying $\beta$ and a significant change in the direction of $\bm{J}$.
-
 The results are shown in Fig.~\ref{fig:fomprec}, using the merger frequency and the ringdown frequency, shown by the vertical lines, to give an idea of the separation between the inspiral and post-merger phases.
@@
 The case $--$ departs from the simple precession picture, and~\eqref{eq:precfomPN} does not apply.
-Before merger, we see that amplitude corrections are within the preturbative regime, contrarily to $\epsilon_{\Psi 2}$ that can exceed $1$ for subdominant modes in the $++$ and $\perp\perp$ cases.
+Before merger, we see that amplitude corrections are within the perturbative regime, contrarily to $\epsilon_{\Psi 2}$ that can exceed $1$ for subdominant modes in the $++$ and $\perp\perp$ cases.
 The amplitude corrections are found to be in the perturbative regime for all cases during the inspiral, however both the $++$ and $\perp\perp$ cases show a sharp increase of $\epsilon_{A1}$, $\epsilon_{A2}$ post-merger.
@@
   \centering
-%  \includegraphics[width=.32\linewidth]{plots/fom_pp_Psi2.pdf}
-%  \hspace{0cm}
-%  \includegraphics[width=.32\linewidth]{plots/fom_pp_A1.pdf}
-%  \hspace{0cm}
-%  \includegraphics[width=.32\linewidth]{plots/fom_pp_A2.pdf}
-%  %
-%  \includegraphics[width=.32\linewidth]{plots/fom_++_Psi2.pdf}
-%  \hspace{0cm}
-%  \includegraphics[width=.32\linewidth]{plots/fom_++_A1.pdf}
-%  \hspace{0cm}
-%  \includegraphics[width=.32\linewidth]{plots/fom_++_A2.pdf}
-%  %
-%  \includegraphics[width=.32\linewidth]{plots/fom_--_Psi2.pdf}
-%  \hspace{0cm}
-%  \includegraphics[width=.32\linewidth]{plots/fom_--_A1.pdf}
-%  \hspace{0cm}
-%  \includegraphics[width=.32\linewidth]{plots/fom_--_A2.pdf}
   \includegraphics[width=.98\linewidth]{plots/precfom_py.pdf}
@@
 
-%\begin{figure*}
-%  \centering
-%  \includegraphics[width=.48\linewidth]{plots/precfom22caseI.pdf}
-%  \hspace{0.2cm}
-%  \includegraphics[width=.48\linewidth]{plots/precfom21caseI.pdf}
-%  %
-%  \includegraphics[width=.48\linewidth]{plots/precfom22caseII.pdf}
-%  \hspace{0.2cm}
-%  \includegraphics[width=.48\linewidth]{plots/precfom21caseII.pdf}
-%  \caption{Figures of merit of the approximation as defined in~\eqref{eq:deffom} for the precession modulations.
-The precessing-frame waveform is taken to be the $22$ mode of an equal-mass, non-spinning system.
-The two upper panels show the case where the frame rotation is freezed at merger (Case I), while the two lower panels show the case where it rotates after merger (Case II) (see Sec.~\ref{subsec:precmodel} for their definitions).
-The left panels show the figures of merit for $F_{22, 22}$, and the right panels for $F_{22, 21}$.
-The vertical line represents the frequency at merger.
-The amplitude corrections become non-negligible after merger, and become large enough to challenge a perturbative treatment in Case II for $F_{22,21}$ (notice the change of scale for the lower right panel).}
-%  \label{fig:TfTA}
-%\end{figure*}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \centering
-%  \includegraphics[width=.48\linewidth]{plots/trig_amp.pdf}
-%  \hspace{0.2cm}
-%  \includegraphics[width=.48\linewidth]{plots/trig_phase.pdf}
   \includegraphics[width=.98\linewidth]{plots/trigopoly_py.pdf}
@@
 The thick and thin vertical lines indicate the frequency of the merger and ringdown (the QNM frequency) respectively.
-The continuous line show the target signal $\tilde{h}(f)$, while the dashed line shows the artificially symmetrized trigonometric polynomial~\eqref{eq:hsymtrigo}.
+The continuous line shows the target signal $\tilde{h}(f)$, while the dashed line shows the artificially symmetrized trigonometric polynomial~\eqref{eq:hsymtrigo}.
 The dots show the discrete samples entering~\eqref{eq:ykDFTprec}.
@@
 The merger and ringdown part of the Fourier-domain signal extends over a wide range of frequencies, while it corresponds to a short interval of times, much shorter than the inspiral part.
-One approach would be to separate the time-domain waveform between the inspiral and merger phase, build a Fourier-domain model for the inspiral part only, to which the formalism of Sec.~\ref{sec:formalism} could be applied, while the merger and ringdown part could be handled by a direct FFT, which would be a cheap operation on a limited range in time.
-Although this approach should be straightforward, here we will use an alternative method, simpler to implement as it allows us to keep the setting of a Fourier-domain precessing-frame waveform combined with a time-domain precession modulation.
+A possible approach would be to separate the time-domain waveform between the inspiral and merger phase, build a Fourier-domain model for the inspiral part only, to which the formalism of Sec.~\ref{sec:formalism} could be applied, while the merger and ringdown part could be handled by a direct FFT, which would be a cheap operation on a limited range in time.
+Although this approach should always be applicable, here we will use an alternative method, simpler to implement as it allows us to keep the setting of a Fourier-domain precessing-frame waveform combined with a time-domain precession modulation.
 
@@
 Taking the point of view of an interpolation problem, Fig~\ref{fig:trigopoly} shows the accuracy of this representation of the high-frequency part of the signal, by comparing the orginal $\tilde{h}(f)$ to its  trigonometric-polynomial representation~\eqref{eq:hsymtrigo}, for the three cases listed in Table~\ref{tab:precexamples}.
-We see that we can achieve a good agreement already for 32 samples (128 counting the symmetrization and 0-padding) \SM{[check this number]}.
+We see that we can achieve a good agreement already for 32 samples (128 counting the symmetrization and 0-padding).
 Errors in the phase $\Psi$ grow towards high frequencies because they are essentially errors in a relative sense, and amplitudes are decaying in this region.
@@
 
-\SM{[update to $10 M_{\odot}$ and $10\mathrm{Hz}$ ?]}
-
-\SM{[the mismatch computation should be averaged over injected phase]}
-
 We now assess the accuracy of the transfer function computation, applying the formalism of Secs.~\ref{subsec:executivesummary} and~\ref{subsec:trigopoly} to the three examples listed in Table~\ref{tab:precexamples}, and comparing the result to a reference numerical computation.
 
-To numerically compute the reference transfer functions, we first have to perform an IFFT of our Fourier-domain P-frame waveform, to obtain $h^{\rm P}_{22}$ as a function of time.
+To obtain the latter, we first have to perform an IFFT of our Fourier-domain P-frame waveform, to obtain $h^{\rm P}_{22}$ as a function of time.
 To mitigate the effect of the necessary tapering of the waveform when computing this numerical inverse Fourier transform, we apply a Planck-window tapering on the range $f\in 16-20 \Hz$ for a total mass of $M=20 \Msol$.
@@
 
-The figures Fig.~\ref{fig:precerrors++}, Fig.~\ref{fig:precerrorspp} and Fig.~\ref{fig:precerrors--} show the result for the cases $++$, $\perp\perp$ and $--$ of Table~\ref{tab:precexamples} at three successive approximations, following the notation of~\eqref{eq:summaryNA}: $\{N:0 | A:0 | \mathrm{No \; Conv.}\}$, which is simply the leading-order transfer function~\eqref{eq:transferlocal}, ignoring all the corrections; $\{N:3 | A:2 | \mathrm{No \; Conv.}\}$, which incorporates both the phase corrections~\eqref{eq:stencilfresnel}, using a stencil size $N=3$, and the amplitude corrections up to second order in~\eqref{eq:summaryNA}; and $\{N:3 | A:2 | \text{conv}\}$, which is the same as the previous setting for the inspiral but uses the convolution formalism of Sec.~\ref{subsec:trigopoly} to cover the high frequency range, with a smooth transition in the shaded range.
-The panels show the Fourier-domain amplitude for each of the modes, both exact and from the reconstruction, the relative errors in amplitude, and the errors in phase.
+The figures Fig.~\ref{fig:precerrors++}, Fig.~\ref{fig:precerrorspp} and Fig.~\ref{fig:precerrors--} show the result for the cases $++$, $\perp\perp$ and $--$ of Table~\ref{tab:precexamples} at three successive approximations, following the notation of~\eqref{eq:summaryNA}: $\{N:0 | A:0 | \mathrm{No \; Conv.}\}$, which is simply the leading-order transfer function~\eqref{eq:transferlocal}, ignoring all the corrections; $\{N:3 | A:2 | \mathrm{No \; Conv.}\}$, which incorporates both the phase corrections~\eqref{eq:stencilfresnel}, using a stencil size $N=3$, and the amplitude corrections up to second order in~\eqref{eq:summaryNA}; and $\{N:3 | A:2 | \text{Conv.}\}$, which is the same as the previous setting for the inspiral but uses the convolution formalism of Sec.~\ref{subsec:trigopoly} to cover the high frequency range, with a smooth transition in the shaded range.
+The panels show the Fourier-domain amplitude for each of the modes, both exact and from the reconstruction, the fractional errors in amplitude, and the errors in phase.
 The errors here are relative to each mode, not to the dominant mode.
@@
 Using the convolution treatment improves the main modes $m=2$, $m=1$, but not the most challenging modes $m=-1$ and $m=-2$, which can be seen to depart from the perturbative treatment before the range covered by the convolution and for which the convolution~\eqref{eq:precconvolution} extends to very high frequencies where our trigonometric polynomial representation of the signal is not accurate (see Fig.~\ref{fig:trigopoly}).
-Note however that, as will be shown below by unfaithfulness computations, these large relative errors for the subdominant modes do not affect much the full waveform.
+Note however that, as will be shown below by unfaithfulness computations, these large fractional errors for the subdominant modes do not affect much the full waveform.
 
@@
 In the norm $(h|h)$, only the dependence on $\delta\Phi$ remains.
-For two signals $h_{1}$ and $h_{2}$, the mismatch between them is then defined as \SM{[check]}
+For two signals $h_{1}$ and $h_{2}$, the mismatch between them is then defined as
 \be\label{eq:defMM}
@@
 Again, we use different approximation levels in~\eqref{eq:summaryNA}, indicated by the color, including or not the treatment of Sec.~\ref{subsec:trigopoly} for the high frequencies as indicated by a continuous or dashed line.
-The total mass ranges from $20\Msol$ (where the inspiral dominates) \SM{[to be updated to $10 \Msol$]} to $400\Msol$ (where the merger-ringdown phase dominates).
-
-As discussed in App.~\ref{app:precpreviousapproaches}, the approximation $\{N:0 | A:0\}$ corresponds to the leading-order formula~\eqref{eq:transferlocal}, which is close to the treatment of PhenomP~\cite{Hannam+13}, with the difference that this model uses effective Fourier-domain Euler angles instead of a time-domain modulation through merger and ringdown.
+The total mass ranges from $20\Msol$ (where the inspiral dominates) to $400\Msol$ (where the merger-ringdown phase dominates).
+
+As discussed in App.~\ref{app:precpreviousapproaches}, the approximation $\{N:0 | A:0 | \text{No Conv.}\}$ corresponds to the leading-order formula~\eqref{eq:transferlocal}, which is close to the treatment of PhenomP~\cite{Hannam+13}, with the difference that this model uses effective Fourier-domain Euler angles instead of a time-domain modulation through merger and ringdown.
 The order of approximation $\{N:3 | A:0\}$ is equivalent to formalism of Ref.~\cite{KCY14} with a stencil order of $N=3$, except that the latter formalism is based on an SPA representation of the P-frame waveform and limited to the inspiral, while our Fourier-domain approach is an attempt at covering the whole frequency band.
@@
 Using all the tools at our hands, we are able to keep the unfaithfulness of our three example waveforms below $2.10^{-3}$ for all masses and inclinations.
+We leave for future work the investigation of the full parameter space of spinning binaries, and of higher harmonics $\ell \neq 2$.
 
@@
   \centering
-%  \includegraphics[width=.32\linewidth]{plots/precerrors_++_00_noconv_col.pdf}
-%  \hspace{0.cm}
-%  \includegraphics[width=.32\linewidth]{plots/precerrors_++_32_noconv_col.pdf}
-%  \hspace{0.cm}
-%  \includegraphics[width=.32\linewidth]{plots/precerrors_++_32_conv_col.pdf}
   \includegraphics[width=.98\linewidth]{plots/precerror_++_py.pdf}
@@
   \centering
-%  \includegraphics[width=.32\linewidth]{plots/precerrors_pp_00_noconv_col.pdf}
-%  \hspace{0.cm}
-%  \includegraphics[width=.32\linewidth]{plots/precerrors_pp_32_noconv_col.pdf}
-%  \hspace{0.cm}
-%  \includegraphics[width=.32\linewidth]{plots/precerrors_pp_32_conv_col.pdf}
   \includegraphics[width=.98\linewidth]{plots/precerror_pp_py.pdf}
@@
   \centering
-%  \includegraphics[width=.32\linewidth]{plots/precerrors_--_00_noconv_col.pdf}
-%  \hspace{0.cm}
-%  \includegraphics[width=.32\linewidth]{plots/precerrors_--_32_noconv_col.pdf}
-%  \hspace{0.cm}
-%  \includegraphics[width=.32\linewidth]{plots/precerrors_--_32_conv_col.pdf}
   \includegraphics[width=.98\linewidth]{plots/precerror_--_py.pdf}
@@
 
-%\begin{figure*}
-%  \centering
-%  \includegraphics[width=.48\linewidth]{plots/precerrorA22caseI.pdf}
-%  \hspace{0.2cm}
-%  \includegraphics[width=.48\linewidth]{plots/precerrorPsi22caseI.pdf}
-%  %
-%  \includegraphics[width=.48\linewidth]{plots/precerrorA21caseI.pdf}
-%  \hspace{0.2cm}
-%  \includegraphics[width=.48\linewidth]{plots/precerrorPsi21caseI.pdf}  
-%  \caption{Errors in the Fourier-domain transfer fonctions $\calT_{22,22}$ and $\calT_{22,21}$ as defined in~\eqref{eq:defTlmlmp}, for Case I and for various approximations.
-The black curve is the exact result provided by the FFT.
-The blue curve gives the leading-order, local approximation.
-The red curve includes quadratic corrections in phase and in amplitude.
-The yellow curve shows the result of the direct conolution approach of Sec.~\ref{subsec:directconvolution}.
-The left panels show the amplitude and the right panels the phase difference with the FFT.
-Note that the vertical scales are not uniform.
-[is it ok to show only 21 and 22 ?
-include the other modes which show problems ?]}
-%  \label{fig:precerrorcaseI}
-%\end{figure*}
-
-%\begin{figure*}
-%  \centering
-%  \includegraphics[width=.48\linewidth]{plots/precerrorA22caseII.pdf}
-%  \hspace{0.2cm}
-%  \includegraphics[width=.48\linewidth]{plots/precerrorPsi22caseII.pdf}
-%  %
-%  \includegraphics[width=.48\linewidth]{plots/precerrorA21caseII.pdf}
-%  \hspace{0.2cm}
-%  \includegraphics[width=.48\linewidth]{plots/precerrorPsi21caseII.pdf}  
-%  \caption{Errors in the Fourier-domain transfer fonctions $\calT_{22,22}$ and $\calT_{22,21}$.
-Conventions are the same as in Fig.~\ref{fig:precerrorcaseI} but for Case II, with a more rapid post-merger frame precession.
-Note that the vertical scales are not uniform.
-[is it ok to show only 21 and 22 ?
-include the other modes which show problems ?]}
-%  \label{fig:precerrorcaseII}
-%\end{figure*}
-
-%\begin{figure*}
-%  \centering
-%  \includegraphics[width=.48\linewidth]{plots/precmmsimple.pdf}
-%  \hspace{0.2cm}
-%  \includegraphics[width=.48\linewidth]{plots/precmmasympt.pdf}
-%  \caption{Mismatches computed between the exact FFT and the various approximations respresented in Figs.~\ref{fig:precerrorcaseI} and~\ref{fig:precerrorcaseII}.
-[log scale for masses ?
-discuss the fact that in case II, leading order is accidentally better than Psi 2]}
-%  \label{fig:precmm}
-%\end{figure*}
-
 \begin{figure*}
   \centering
-%  %
-%  \includegraphics[width=.32\linewidth]{plots/mismatches_++_inc0.pdf}
-%  \hspace{0.1cm}
-%  \includegraphics[width=.32\linewidth]{plots/mismatches_++_incpi3.pdf}
-%  \hspace{0.1cm}
-%  \includegraphics[width=.32\linewidth]{plots/mismatches_++_incpi2.pdf}
-%  %
-%  \includegraphics[width=.32\linewidth]{plots/mismatches_pp_inc0.pdf}
-%  \hspace{0.1cm}
-%  \includegraphics[width=.32\linewidth]{plots/mismatches_pp_incpi3.pdf}
-%  \hspace{0.1cm}
-%  \includegraphics[width=.32\linewidth]{plots/mismatches_pp_incpi2.pdf}
-%  %
-%  \includegraphics[width=.32\linewidth]{plots/mismatches_--_inc0.pdf}
-%  \hspace{0.1cm}
-%  \includegraphics[width=.32\linewidth]{plots/mismatches_--_incpi3.pdf}
-%  \hspace{0.1cm}
-%  \includegraphics[width=.32\linewidth]{plots/mismatches_--_incpi2.pdf}
   \includegraphics[width=.98\linewidth]{plots/precMM_py.pdf}
@@
 The computation is done using~\eqref{eq:defMM} with $f_{\rm min} = 20\mathrm{Hz}$, for masses between 20 and 400 $\Msol$, and with the aLIGO ZDHP noise curve~\cite{LIGOProspects13}.
-The inclination is defined here as the angle between the line of sight and the final $\bm{J}$.
-\SM{[Extend computation to fmin=10 and Mmin=10 ?]} \SM{[re-check definition of inclination: from $J$ or initial $L$ ?]} }
+The inclination is defined here as the angle between the line of sight and the final $\bm{J}$.}
   \label{fig:precunfaithfulness}
@@
 
-This formalism is based on a the separation of timescales between the modulation and the signal.
+This formalism is based on the separation of timescales between the modulation and the signal.
 By a Fourier-domain expansion of the integrand of convolution integrals, corrections beyond the leading order are straightforwardly generated.
@@
 To explore the application of the formalism to precessing binaries, we built a toy model for the precession of a two-spins system including a post-merger precessing-frame trajectory.
-For the high spin, moderate mass ratio examples we explored, we found that subdominant precessing mode contributions can be challenging for the formalism and have large relative errors, although they are suppressed in amplitude in the complete waveform.
+For the high spin, moderate mass ratio examples we explored, we found that subdominant precessing mode contributions can be challenging for the formalism and have large fractional errors, although they are suppressed in amplitude in the complete waveform.
 For the inspiral, the most relevant corrections are the quadratic in phase ones that were already derived in~\cite{KCY14}.
@@
 First, we only considered the leading harmonic of non-spinning waveforms in the LISA response.
-The same approach is also well-suited to treating higher harmonics and aligned spin signal, but precession will leave a definite imprint on the Fourier-domain signal.
+The same approach is also well-suited to treating higher harmonics and aligned spin signals, but precession will leave a definite imprint on the Fourier-domain signal.
 One option in this case would be to handle the precession and response modulation together, as in~\cite{Klein+15}.
@@
 For detectors like the Einstein Telescope, signals are sufficiently long for the rotation of the Earth to matter, and their arms are long enough that the long wavelength approximation is not valid at high frequencies and that the response becomes frequency-dependent~\cite{Essick+17}.
-A preliminary order-of-magnitude analysis shows that the leading-order treatment should be applicable \SM{[check]}, but more investigation is needed.
+A preliminary order-of-magnitude analysis shows that the leading-order treatment should be applicable, but more investigation is needed.
 
@@
 Those signals generally include not only strong precession but also large eccentricity.
-An effective frequency-domain LISA response may be possible by apply our treatment separately to each of the numerous individual harmonics produced (see e.g.~\cite{Chua+17}).
+An effective frequency-domain LISA response may be possible by applying our treatment separately to each of the numerous individual harmonics produced (see e.g.~\cite{Chua+17}).
 
@@
 
-We are indebted to Alejandro Boh\'{e}, Alessandra Buonanno, Scott Field, Philip Graff, Sebastian Khan, Antoine Klein, Serguei Ossokine, Michael P\"{u}rrer for useful discussions and comments.
+We are indebted to Stanislav Babak, Alejandro Boh\'{e}, Alessandra Buonanno, Tito Dal Canton, Scott Field, Philip Graff, Sebastian Khan, Antoine Klein, Serguei Ossokine, Michael P\"{u}rrer for useful discussions and comments.
 We used the software GWFrames~\cite{Boyle13, Boyle+14} in our study.
@@
 \label{app:notation}
-
-\SM{[check part on discrete Fourier coeffs]}
 
@@
 \end{align}
-the range $f<0$ being obtained readily, since $h_{+},h_{\times}$ are real quantities, from $\tilde{h}_{+,\times} (-f) = \tilde{h}_{+,\times}(f)^{*}$.
-
-%In the following, we will drop the mode indices and the sum over modes, and we will denote generically by $h$ an individual mode of the waveform\footnote{This notation for a generic individual mode is not to be confused with the often encountered notation $h\equiv h_{+} - i h_{\times}$.}.
-Note that, if we are considering real signals $s(t)$ (such as $h_{+}$, $h_{\times}$, or any detector output), we can focus  on the positive-frequencies range and reconstruct the negative frequencies as $\tilde{s}(-f) = \tilde{s}(f)^{*}$.
-For positive frequencies $f>0$, we will use the generic notation $h$ \SM{[not using the 'generic notation' anymore]} as
-%\begin{align}\label{eq:notationhsignm}
-%	h(t) &\equiv h_{\ell m}(t) \,, \quad \tilde{h}(f) \equiv \tilde{h}_{\ell m}(f) \text{ for } m>0 \,, \nn\\
-%	h(t) &\equiv h_{\ell m}^{*}(t) \,, \quad \tilde{h}(f) \equiv \tilde{h}_{\ell m}(-f)^{*} \text{ for } m<0 \,.
-%\end{align}
+The range $f<0$ can be obtained readily, since $h_{+},h_{\times}$ are real quantities, from $\tilde{h}_{+,\times} (-f) = \tilde{h}_{+,\times}(f)^{*}$.
+
 Dropping the mode indices $\ell$, $m$, we will decompose a given mode\footnote{Not to be confused with the commonly used notation $h = h_{+} - i h_{\times}$ for the complex strain.} $\tilde{h}_{\ell m} = \tilde{h}$ into a Fourier-domain amplitude $A$ and a phase $\Psi$ according to
@@
 \ee
-%Since we intend to work in the Fourier domain, we will use PhenomD waveforms~\cite{Khan+15}.
-These phenomenological waveforms provide analytic fits for both the Fourier-domain amplitude and phase of the 22 mode, with fit coefficients that were calibrated on EOB/NR hybrids \SM{[describe smoothing across intervals ?]}.
-Fig.~\ref{fig:ampphase} shows an example of amplitude and phase in geometric units for an equal-mass binary.
-\SM{[remove that and point to appendix on interpolation errors]}
-%\begin{figure*}
-%  \centering
-%  \includegraphics[width=.48\linewidth]{plots/Af.pdf}
-%  \hspace{0.2cm}
-%  \includegraphics[width=.48\linewidth]{plots/Psif.pdf}
-%  \caption{Fourier-domain amplitude and phase in geometric units for an equal-mass, non-spinning system.
-The vertical line represents $\omega_{22}^{\rm peak}/2\pi$, the frequency corresponding to the time-domain frequency of the waveform at the peak amplitude.
-The starting frequency $Mf_{min} = 2.6\times 10^{-4}$ is appropriate for a 2-years observation of a $M=10^{6} M_{\odot}$ system by LISA, or for an observation by LIGO/Virgo of a $M=5.4 M_{\odot}$ system entering the instrument band at $10\Hz$.
-[add residuals of interpolation on $~300$ points, to illustrate the compression of Amp/Phase FD waveforms ?
-or show different spins/mass ratios ?] \SM{[Plot perhaps not necessary ?]}}
-%  \label{fig:ampphase}
-%\end{figure*}
-%We recall that, with our sign convention, the effect of a positive shift in time $h' (t) = h(t- \Delta t)$ is
-%\be
-%	\tilde{h}'(f) = e^{2i\pi f \Delta t} \tilde{h}(f) \,.
-%\ee
-%If we also allow for a global shift in phase of the signal, which corresponds to a rotation of the unit vector pointing towards the observer in the frame of the gravitational waves source, a constant term can also be added (note that the effect of such a rotation gives different phase changes $e^{\pm i m \delta \phi}$ for the different modes of the radiation).
-We see that both a constant and a linear term in the Fourier-domain phase can be adjusted by changing extrinsic parameters.
-Thus, as is well known, apart from the relative phases of the modes for waveforms that include higher harmonics, the intrinsic information on the phasing is contained in the second derivative of the Fourier-domain phase $\Psi(f)$.
-\SM{[rewrite this paragraph]}
 
@@
 \ee
-%integrals over the interval of $F$ weighted by the appropriate exponential (as in~\eqref{eq:defcn}).
 In both interpretations, either the truncated approximation of the Fourier series or  the trigonometric interpolation formulation, these coefficients are related to the coefficients of the IFFT.
-If we set $\omega \equiv e^{2i\pi/N}$ and define \SM{[check convention]}
+If we set $\omega \equiv e^{2i\pi/N}$ and define
 \be\label{eq:ykDFT}
@@
 
-Here, we use the dominant eigenvector prescription of~\cite{OShaughnessy+11}, to which we refer (as well as~\cite{Boyle13}) for more details. One defines the matrix
+Here, we use the dominant eigenvector prescription of~\cite{OShaughnessy+11}, to which we refer (as well as~\cite{OOS12, Boyle13}) for more details. One defines the matrix
 \be
@@
 \ee
-%Now, since the Vandermonde system can be seen as a reformulation of the polynomial interpolation problem, we know that $({}^{t}V^{-1})_{ij}$ is the coefficient of $X^{i}$ in the Lagrange interpolation polynomial $\prod_{k\neq j} (X-x_{k})/(x_{j} - x_{k})$.
-If we introduce the symmetric polynomials $\sigma$ such that
-%\be	
-%	\sigma(m, \{x_{1}, \dots, x_{n}\}) = \sum\limits_{1\leq i_{1}<\dots<i_{m}\leq n} x_{i_{1}}\dots x_{i_{m}} \,,
-%\ee
-%we obtain for the inverse of the matrix $V$
-%\be
-%	V^{-1}_{ij} = \frac{1}{\prod_{k\neq i} (i^{2} - k^{2})} (-1)^{N-j} \sigma(N-j, \{0,\dots,N^{2}\}\backslash \{i^{2}\}) \,.
-%\ee
 The expression of the inverse of a Vandermonde matrix in terms of symmetric polynomials then allows us to write the $a_{N,k}^{\epsilon}$, for every finite order $N$, as:
@@
 In the first option, one applies directly the SPA, as described in Sec.~\ref{subsec:SPA}, to the product of the modulation and the signal.
-This is known (see e.g.~\cite{KCY13}) to lead to possible pathologies, since the prefactor $1/\sqrt{\ddot{\phi}_{\rm prec}}$ can blow up due the precessing contributions to the phase of the inertial-frame waveform.
+This is known (see e.g.~\cite{KCY13}) to lead to possible pathologies, since the prefactor $1/\sqrt{\ddot{\phi} + \ddot{\Phi}_{\rm prec}}$ can blow up due the precessing contributions to the phase of the inertial-frame waveform.
 
@@
 In~\cite{LOS13}, only the inspiral phase is modelled, and only single-spin configurations, for which the system undergoes simple precession~\cite{Apostolatos+94, Kidder95}.
-The PhenomP model~\cite{Hannam+13} uses Euler angles computed in the limit of a small opening angle of the precession cone and at the spin-orbit level (see also~\cite{BBF11, MBBB13}).
+The PhenomP model~\cite{Hannam+13} maps the two spins to a single effective spin, and uses Euler angles computed in the limit of a small opening angle of the precession cone and at the spin-orbit level (see also~\cite{BBF11, MBBB13}).
 
@@
 In practice, one finds however that the extended frequency-based expressions~\eqref{eq:precPhenomP} are mildly varying when extrapolating to higher frequencies, and this approach has been validated by comparisons to numerical relativity waveforms~\cite{Hannam+13}.
-
-In the SUA formalism proposed in~\cite{KCY13, KCY14}, one extends the SPA to incorporate the leading-order correction during the inspiral phase.
+Future versions of PhenomP~\cite{PhenomPv3InPrep} will move beyond the single-spin approximation by incorporating analytic solutions for the precession trajectory of double-spin systems~\cite{Chatziioannou+17}.
+
+In the SUA formalism proposed in~\cite{KCY13, KCY14}, one extends the SPA to incorporate the leading-order correction during the inspiral.
 The transfer function is then computed by evaluating the modulation on a stencil of times centered around the time-of-frequency given by the SPA, arriving at the formula~\eqref{eq:stencilfresnel} with the times $t = \tfSPA$ and $T = \TfSPA$ as defined in~\eqref{eq:deftfSPA} and~\eqref{eq:TfSPA}.
@@
 \ee
-where we have used $\ddot{\omega} = 11/3 (\dot{\omega})^{2}/\omega$ following~\eqref{eq:omegaphiN}, the extra Fourier-domain phase contribution~\eqref{eq:additionalPhiprec} cause new contributions in~\eqref{eq:deftf} and~\eqref{eq:defTf} as \SM{[check]}
+where we have used $\ddot{\omega} = 11/3 (\dot{\omega})^{2}/\omega$ following~\eqref{eq:omegaphiN}, the extra Fourier-domain phase contribution~\eqref{eq:additionalPhiprec} cause new contributions in~\eqref{eq:deftf} and~\eqref{eq:defTf} as
 \begin{subequations}
@@
 	-\frac{1}{2\pi} \frac{\ud \Phi_{\rm prec}}{\ud f} &\sim (m' \cos \beta - m) \Tf^{2} \Omega_{\rm prec} \,, \label{eq:additionaltf}\\
-	-\frac{1}{4\pi^{2}} \frac{\ud^{2} \Phi_{\rm prec}}{\ud f^{2}} &\sim (m' \cos \beta - m) \Tf^{2} \left[ \frac{11}{12} \frac{\Omega_{\rm prec}}{\pi f} - \Tf^{2} \dot{\Omega}_{\rm prec} \right] \,, \label{eq:additionalTf}
+	-\frac{1}{4\pi^{2}} \frac{\ud^{2} \Phi_{\rm prec}}{\ud f^{2}} &\sim (m' \cos \beta - m) \Tf^{2} \left[ \frac{11}{6} \frac{\Omega_{\rm prec}}{\pi f} - \Tf^{2} \dot{\Omega}_{\rm prec} \right] \,, \label{eq:additionalTf}
 \end{align}
@@
 
-%\bibliography{ListeRef.bib}
 \bibliography{references.bib}

### 2018-07-23 docs/paper_method.tex

@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/tf_py.pdf}
+  \includegraphics[width=.98\linewidth]{tf_py.pdf}
   \caption{Time-to-frequency correspondence $t_{f}$, as defined directly from the phase of the Fourier-domain signal in~\eqref{eq:deftf}, in geometric units.
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/TfTA_py.pdf}
+  \includegraphics[width=.98\linewidth]{TfTA_py.pdf}
   \caption{Fourier-domain amplitude and phase timescales, as defined in~\eqref{eq:defTf} and~\eqref{eq:defTA}, in geometric units for an equal-mass, non-spinning system.
@@
   \centering
-  \includegraphics[width=.99\linewidth]{plots/lisafom_Psi_McDeltat_py.pdf}
+  \includegraphics[width=.99\linewidth]{lisafom_Psi_McDeltat_py.pdf}
   \caption{Contour levels for the analytical estimate of the error measure $\epsilon_{\Psi 2}$ at the starting frequency, as a function of chirp mass $\Mchirp$ and time to coalescence $\Delta t$.
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/lisafom_py.pdf}
+  \includegraphics[width=.98\linewidth]{lisafom_py.pdf}
   \caption{Error estimates as defined in~\eqref{eq:deffom}, for an equal-mass, non-spinning system and for total masses $M=10^{7} \Msol$, $10^{4} \Msol$ and $10^{2} \Msol$.
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e7dO_py.pdf}
+  \includegraphics[width=.98\linewidth]{LISAtransfererrorM1e7dO_py.pdf}
   \caption{Transfer function and residual for the orbital response~\eqref{eq:defresponse0}, for an equal-mass system with $M=10^{7} \Msol$.
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e2dO_py.pdf}
+  \includegraphics[width=.98\linewidth]{LISAtransfererrorM1e2dO_py.pdf}
   \caption{Transfer function and residual for the orbital response~\eqref{eq:defresponse0}, for an equal-mass system with $M=10^{2} \Msol$.
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e7y12L_py.pdf}
+  \includegraphics[width=.98\linewidth]{LISAtransfererrorM1e7y12L_py.pdf}
   \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:defresponseL}, for an equal-mass system with $M=10^{7} \Msol$.
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorM1e2y12L_py.pdf}
+  \includegraphics[width=.98\linewidth]{LISAtransfererrorM1e2y12L_py.pdf}
   \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:defresponseL}, for an equal-mass system with $M=10^{2} \Msol$.
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/LISACombExtent_py.pdf}
+  \includegraphics[width=.98\linewidth]{LISACombExtent_py.pdf}
   \caption{Extent of the Fourier-domain comb entering the convolution~\eqref{eq:transferdiscreteconvolution}, determined from the criterion~\eqref{eq:criteriontruncationcomb} with truncation levels $\eta=10^{-12}$, $10^{-6}$, $10^{-3}$.
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi1d0corr_py.pdf}
+  \includegraphics[width=.98\linewidth]{LISAtransfererrorSOBHepsPsi1d0corr_py.pdf}
   \caption{Transfer function and residual for the orbital delay $d_{0}$~\eqref{eq:delay0}, for an SOBH with $\epsilon_{\Psi 2}\sim 0.6$ for the orbital response.
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi10d0corr_py.pdf}
+  \includegraphics[width=.98\linewidth]{LISAtransfererrorSOBHepsPsi10d0corr_py.pdf}
   \caption{Same as Fig.~\ref{fig:LISAerrorSOBHepsPsi1d0}, but for an SOBH with $\epsilon_{\Psi 2}\sim 4.5$ for the orbital response.
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi1y12L_py.pdf}
+  \includegraphics[width=.98\linewidth]{LISAtransfererrorSOBHepsPsi1y12L_py.pdf}
   \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:GslrL}, for the same SOBH as in Fig.~\ref{fig:LISAerrorSOBHepsPsi1d0}.
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/LISAtransfererrorSOBHepsPsi10y12L_py.pdf}
+  \includegraphics[width=.98\linewidth]{LISAtransfererrorSOBHepsPsi10y12L_py.pdf}
   \caption{Transfer function and residual for the constellation response $y_{132}$~\eqref{eq:GslrL}, for the same SOBH as in Fig.~\ref{fig:LISAerrorSOBHepsPsi10d0}.
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/eulerSXS0058_py.pdf}
+  \includegraphics[width=.98\linewidth]{eulerSXS0058_py.pdf}
   \caption{Evolution of Euler angles $(\alpha, \gamma)$ (top panel) and $\beta$ (bottom panel) near merger for the example SpEC waveform SXS:BBH:0058~\cite{SpEC, SXScatalog, Mroue+12, Mroue+13}.
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/postmergeromega_py.pdf}
+  \includegraphics[width=.98\linewidth]{postmergeromega_py.pdf}
   \caption{Quasi-normal mode frequencies for the 0th overtone of the modes $22$ and $21$, as a function of the dimensionless spin of the final black hole $\chi_{f}$, as well as their difference $\Omega_{\rm frame}$ (see~\eqref{eq:OmegaframeQNM}).
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/precfom_py.pdf}
+  \includegraphics[width=.98\linewidth]{precfom_py.pdf}
   \caption{Error estimates of the approximation as defined in~\eqref{eq:deffom} for the precession modulations, for the three cases $++$, $\perp\perp$ and $--$ listed in Table~\ref{tab:precexamples}.
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/trigopoly_py.pdf}
+  \includegraphics[width=.98\linewidth]{trigopoly_py.pdf}
   \caption{Amplitude (left panel) and phase (right panel), compared to its trigonometric polynomial representation for the three cases listed in Table~\ref{tab:precexamples}.
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/precerror_++_py.pdf}
+  \includegraphics[width=.98\linewidth]{precerror_++_py.pdf}
   \caption{Fourier-domain amplitudes of the modes $\tilde{h}^{\rm I}_{\ell m}$ (top panels), reconstruction errors in the transfer fonctions $\calT^{2}_{m2}$ for the amplitudes (middle panels) and for the phases (bottom panels),  for the configuration $++$ of Table~\ref{tab:precexamples}, with spins almost aligned.
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/precerror_pp_py.pdf}
+  \includegraphics[width=.98\linewidth]{precerror_pp_py.pdf}
   \caption{Same as Fig.~\ref{fig:precerrors++}, but for the configuration $\perp\perp$ of Table~\ref{tab:precexamples}, with both spins in the orbital plane.}
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/precerror_--_py.pdf}
+  \includegraphics[width=.98\linewidth]{precerror_--_py.pdf}
   \caption{Same as Fig.~\ref{fig:precerrors++}, but for the configuration $--$ of Table~\ref{tab:precexamples}, with almost anti-aligned spins.}
@@
   \centering
-  \includegraphics[width=.98\linewidth]{plots/precMM_py.pdf}
+  \includegraphics[width=.98\linewidth]{precMM_py.pdf}
   \caption{Unfaithfulness at various orders of approximation as compared with an FFT of the full time-domain precessing signal, for the three precessing systems of Table~\ref{tab:precexamples}.

### 2018-09-25 docs/paper_method.tex
Minimal start on referee response actions.

@@
 With the unprecedented recent gravitational-wave detections of coalescencing binary black holes and binary neutron stars, announced by the LIGO-Virgo collaboration~\cite{LIGO-theevent-2016, LIGO-christmasevent-2016, LIGO-O1BBH-2016, LIGO-BNSevent-2017}, gravitational-wave astronomy has entered its observational era.
-As LIGO prepares for even more sensitive observation runs, and with the recent expansion of the ground-based detectors network with Virgo~\cite{Virgo} (and eventually also KAGRA~\cite{KAGRA} and LIGO-India~\cite{INDIGO}), observations of such compact object coalescences are expected at an ever-increasing rate.
-
-Moreover, the European Space Agency has recently selected the Laser Interferometer Space Antenna (LISA)~\cite{LISA17} to realize the ``Gravitational Universe'' science theme~\cite{elisa13} as the 3rd large space mission of its Cosmic Vision program, with a tentative launch around 2034.
+As Advanced LIGO prepares for even more sensitive observation runs, and with the recent expansion of the ground-based detectors network with Advanced Virgo~\cite{Virgo} (and eventually also KAGRA~\cite{KAGRA} and LIGO-India~\cite{INDIGO}), observations of such compact object coalescences are expected at an ever-increasing rate.
+
+Moreover, the European Space Agency has recently selected the Laser Interferometer Space Antenna (LISA)~\cite{LISA17} to realize the ``Gravitational Universe'' science theme~\cite{elisa13} as the third large space mission of its Cosmic Vision program, with a tentative launch around 2034.
 A technology demonstrator, LISA Pathfinder, has tested with great success some of the key mission technologies~\cite{LISAPathfinder2016, LISAPathfinder2018}.
@@
 
-Data analysis for gravitational-wave observations of compact binary coalescences require accurate models (or templates) for the signals, both for ensuring efficient detections of signals that may be buried in instrumental noise, and to extract the physical parameters of the source in a subsequent analysis.
+Data analysis for gravitational-wave observations of compact binary coalescences requires accurate models (or templates) for the signals, both for ensuring efficient detections of signals that may be buried in instrumental noise, and to extract the physical parameters of the source in a subsequent analysis.
 Bayesian analysis for parameter estimation of gravitational-wave signals, as was performed for the LIGO detections~\cite{LIGO-theeventPE-2016,LIGO-O1BBH-2016}, may require millions of evaluations of the likelihood function to sample the posterior probability distribution.
@@
 The GW community is making progress assembling higher-fidelity tools for these coming challenges.
-State-of-the-art IMR templates combine information from the perturbative results of post-Newtonian (PN) theory covering the inspiral (see e.g.~\cite{BlanchetLiving}) and from numerical relativity (NR) simulations covering the end of the inspiral and the merger-ringdown phase (see e.g.~\cite{Pfeiffer12}).
+State-of-the-art inspiral-merger-ringdown (IMR) templates combine information from the perturbative results of post-Newtonian (PN) theory covering the inspiral (see e.g.~\cite{BlanchetLiving}) and from numerical relativity (NR) simulations covering the end of the inspiral and the merger-ringdown phase (see e.g.~\cite{Pfeiffer12}).
 Approaches to template construction include phenomenological templates postulating an analytic ansatz for the Fourier-domain amplitude and phase~\cite{Husa+15, Khan+15,Hannam+13}, and the Effective-One-Body (EOB) approach~\cite{BD99,Taracchini+13, Pan+13, Bohe+16} incorporating PN and NR information.
@@
 A promising approach to modeling the effect of precession on the emitted waveform, as proposed in~\cite{BCV03b, BCPTV05, Schmidt+10, OShaughnessy+11, Boyle+11}, is to decompose precessing waveforms by performing a time-dependent rotation, following the precession of the orbital plane.
-The resulting waveform in the rotated frame can then be modelled by a non-precessing waveform, an approximation which is used both in the construction of the inspiral part of precessing EOB waveforms~\cite{Pan+13} and in the construction of precessing phenomenological waveforms~\cite{Hannam+13}.
+The resulting waveform in the rotating frame can then be modelled by a non-precessing waveform, an approximation which is used both in the construction of the inspiral part of precessing EOB waveforms~\cite{Pan+13} and in the construction of precessing phenomenological waveforms~\cite{Hannam+13}.
 To follow this modelling approach and efficiently create Fourier-domain waveforms, one needs to understand how to translate the time-domain modulations created by the frame rotation into a Fourier-domain transfer function.
@@
 Beyond a fast representation of the incident gravitational wave, observational analyses also require transforming the signals through some instrumental response.
-For short duration mergers such as LIGO  has detected, this can be treated by a simple multiplier and a fixed timeshift between detectors.
+For short duration mergers such as LIGO and Virgo have detected, this can be treated by a simple multiplier and a fixed timeshift between detectors.
 For future instruments though, the instrumental response will be more complicated.
@@
 The PhenomD waveforms have been aligned such that the time-domain amplitude (obtained by an IFFT) peaks at $t=0$.
-The two aligned spins are equal, with components of $0.95$ ($++$, blue), $0$ ($00$, red), and $-0.95$ ($--$, yellow), for mass ratios $q=1$ (full line) and $q=8$ (dashed).
+The two aligned spins are equal, with components of $0.95$ ($++$, blue), $0$ ($00$, red), and $-0.95$ ($--$, yellow), for mass ratios $q=1$ (solid line) and $q=8$ (dashed).
 The vertical lines shows the time-domain instantaneous frequency at the peak $\omega^{22}_{\rm peak}/(2\pi)$, and the fact that the curves $t_{f}$ do not pass exactly by the crossing with the $t=0$ horizontal line reflects the fact that the link between time-domain frequency and Fourier frequency $f$ is only approximate at merger.
@@
 \ee
-which will serve as an estimate for the enveloppe function of the response, devoid of the zero-crossings at high frequencies of the $\sinc$ term in~\eqref{eq:GslrL}.
+which will serve as an estimate for the envelop function of the response, devoid of the zero-crossings at high frequencies of the $\sinc$ term in~\eqref{eq:GslrL}.
 

### 2019-11-08 docs/paper_LDC/paper_LDC.tex
Some notes on symmetry which may be used in an LDC entry paper.

\documentclass[aps,showpacs,twocolumn,prd,superscriptaddress,nofootinbib]{revtex4}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{color}
\usepackage{enumerate}
\usepackage{ulem}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand\ud{{\mathrm{d}}}
\newcommand\uD{{\mathrm{D}}}
\newcommand\calO{{\mathcal{O}}}
\newcommand\bfx{\mathbf{x}}
\newcommand{\ov}[1]{\overline{#1}}
\newcommand{\ph}[1]{\phantom{#1}}
\newcommand{\cte}{\mathrm{cte}}
\newcommand{\nn}{\nonumber}
\newcommand{\hatk}{\hat{k}}
\newcommand{\Hz}{\,\mathrm{Hz}}
\newcommand{\sinc}{\,\mathrm{sinc}}
\newcommand{\Msol}{M_{\odot}}
\newcommand{\bsub}{\begin{subequations}}
\newcommand{\esub}{\end{subequations}}
\newcommand\betaL{{\beta_{L}}}
\newcommand\lambdaL{{\lambda_{L}}}
\newcommand\varphiL{{\varphi_{L}}}
\newcommand\psiL{{\psi_{L}}}

\begin{document}

\title{Methods for black hole merger parameter estimation with LISA.}

\author{John G. Baker}
\affiliation{Gravitational Astrophysics Laboratory, NASA Goddard Space Flight Center, 8800 Greenbelt Rd., Greenbelt, MD 20771, USA}
\author{Sylvain Marsat}

\date{\today}

\begin{abstract}

[Abstract]

\end{abstract}

\pacs{
04.70.Bw, % classical black holes
04.80.Nn, % Gravitational wave detectors and experiments
95.30.Sf, % relativity and gravitation
95.55.Ym, % Gravitational radiation detectors
97.60.Lf  % black holes (astrophysics)
}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}

[Introduction]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Symmetries}

In \cite{MarsatEA2019} we have discussed symmetries which apply to LISA in low-frequency, minimal-motion limit, or more generally any GW instrument in this limit which measures both polarization modes. In this context, we use sky coordinates adapted $\{\beta_{L},\lambda_{a}\}$ adapted to the instrument. Since $\iota \in [0, \pi]$ while $\beta_{L} \in [-\pi/2, \pi/2]$, it will be more convenient to work with $\theta_{L} = \pi/2 - \beta_{L}$ and we further abbreviate notations by using the variables
\be
	t_{\iota} \equiv \tan \frac{\iota}{2} \,, \quad t_{\theta} \equiv \tan \frac{\theta_{L}}{2} \,.
\ee
We write the response signals for a single-mode circularly radiating source via complex combinations of the two polarization channels (having already factored out the intrinsic parameters and having integrated over the Fourier domain),
\be
	\sigma_{\pm} \equiv s_{a}^{22} \pm i s_{e}^{22} \,.
\ee
In this notation, we obtain
\bsub\label{eq:sigmapm}
\begin{align}
  \sigma_{+} &= \rho e^{2i\varphi} \left[ t_{\theta}^{4} e^{-2 i \psiL} + t_{\iota}^{4} e^{2 i \psiL} \right] e^{-2i \lambda_{a}} \,, \\
  \sigma_{-} &= \rho e^{2i\varphi} \left[ e^{-2 i \psiL} + t_{\theta}^{4} t_{\iota}^{4} e^{2 i \psiL} \right] e^{2i \lambda_{a}} \,,
\end{align}
\esub
with a common prefactor 
\be\label{eq:sigmafactorrho}
\rho(d, \iota, \theta_{L}) = \frac{1}{4d\left( 1 + t_{\iota}^{2} \right)^{2} \left(1 + t_{\theta}^{2} \right)^{2}} \,.
\ee

Having measured $\sigma_{+}$ and $\sigma_{-}$ from some observation with assumed intrinsic parameters, we can now count 4 (real) constraints, for the 6 remaining intrinsic parameters.  This leaves us with a degenerate subspace described by disjoint 2-D pieces.

We can regard this as the basic degenerate space for a 2-polarization GW observation, with opportunities for degeneracy breaking with multiple harmonic modes, an accelerating detector, and/or detector with a size comparable to the GW wavelength.  MCMC performance should usually be improved by applying proposals which exploit the symmetries of the degenerate space, which we can understand from examplination of \eqref{eq:sigmapm,eq:sigmafactorrho}.

There are a number of discrete symmetries evident. First, we can rotate the source or instrument through $\pi/4$ with a corresponding rotation of the polarization angle,
\begin{align}
  \phi\rightarrow\phi+\pi/4&\mathrm{and}&\psi\rightarrow\psi+\pi/4\\
  \lambda\rightarrow\lambda+\pi/4&\mathrm{and}&\psi\rightarrow\psi+\pi/4.
\end{align}
The first, for instance, just changes the sign on the $\phi$ exponentials and the $\psi$ exponentials.

Next we note a symmetry associated with reflecting the location of the observation through the instrument plane, 
\be\label{eq:symmetryresponse}
\betaL \rightarrow -\betaL\,, \quad \iota \rightarrow \pi - \iota \,, \quad \psiL \rightarrow \pi - \psiL,
\ee
whereby  $t^4_{\theta}->t^{-4}_{\theta}$ and $t^4_{\iota}->t^{-4}_{\iota}$,
leaving the response exactly invariant.

There is also a symmetry which exchanges the altidual angle of the GW receiver and transmitter, i.e. exchanging $\iota \leftrightarrow \theta_{L}$. While leaving $\sigma_-$ unchanged, this conjugates the factor inside brackets for $\sigma_{+}$, which can be compensated using the phase terms $\varphi \pm \lambda_{a}$. If $\Phi = \mathrm{Arg} \left[ t_{\theta}^{4} e^{-2 i \psiL} + t_{\iota}^{4} e^{2 i \psiL} \right]$, we obtain the symmetry
\begin{align}
	\iota' = \theta_{L} \,, \qquad \theta_{L}' = \iota \,, \nn\\
	\varphi' = \varphi + \frac{1}{2} \Phi \quad \mathrm{mod} \; \pi\,, \nn\\
	\lambda_{a}' = \lambda_{a} - \frac{1}{2} \Phi \quad \mathrm{mod} \; \pi\,.
\end{align}

We can also describe the continuous symmetry describing movement within the 2-D subspace by exchanging distance $d$ with some combination of $\iota$ and $\beta$ in  \eqref{eq:sigmafactorrho}.  This requires also ensuring that changes to $t^4_{\theta}$ and $t^4_{\iota}$ do not alter \eqref{eq:sigmapm}.

First, for an approximate approach, note that in the limit that either $t^4_\theta<<1$ or $t^4_\iota<<1$, then \eqref{eq:sigmapm} become nearly independent of that quantity except through  \eqref{eq:sigmafactorrho}. Then we have an approximate symmetry
\begin{align}
  \iota&\rightarrow\iota'\\
  d&\rightarrow d\frac{(1+t^2_{\iota})^2}{(1+t'^2_{\iota})^2}
\end{align}
A similar symmetry exists when $t^4_\theta<<1$.

It is also possible to explicitly formulate this symmetry in the general case, though here we are still working on a concrete plan.
To approach this, we represent subspace of the parameter space, with coordinates $\{x=\min\{t^4_\theta,t^4_\iota\},y=\max\{t^4_\theta,t^4_\iota\},C=cos(2\psi),d\}$ trivially related to our basic coordinates, but better adapted for this problem.

We will next consider an exact generalization of the distance-inclination symmetry. For this we use coordinates $\{z=xy,\Delta=(|\sigma_-|^2-|\sigma_+|^2)/\rho^2,S=|\sigma_-|^2/\rho^2,\rho\}$.  Of these latter coordinates, the last 3 will be considered fixed within the orbits of the symmetry, though only 2 combinations of these are fully specified by the observed constraints on \eqref{eq:sigmapm}. The remaining coordinate $z$ is free within $S^{1/2}-1\leq z\leq 1-\Delta$. The transformation to these coordinates is
\begin{align}
  z&=xy\\
  \Delta&=(1-x^2)(1-y^2)\\
  S&=1+x^2y^2+2xyC\\
  D&=d(1+x^{1/2})(1+y^{1/2}).
\end{align}
The symmetry is realized by any translation in $z\rightarrow z$, the new coordinates are ten recovered from the inverse transformation is 
\begin{align}
  x&=\sqrt{\Lambda(z,\Delta)-\delta(z,\Delta)}\\
  y&=\sqrt{\Lambda(z,\Delta)+\delta(z,\Delta)}\\
  C&=\frac{S-1-z^2}{2z}\\
  d&=\frac D{(1+x(z,\Delta)^{1/2})(1+y(z,\Delta)^{1/2})},
\end{align}
with $\Lambda=(1+z^2-\Delta^2)/2$ and $\delta=\sqrt{\Lambda^2-z^2}$.
Then to complete the transformation, since the argument of the complex argument of the factors in square brackets in \eqref{eq:sigmapm} will have changed, it is necessary to complete the specification by choosing $\lambda$ and $\phi$ to preserve the argument of the full expressions.

For application as a proposal, the Jacobian of the transformation is needed.  This is straightforward to compute explicitly for the forward transformation above, and its inverse can be evaluated using the same expression with the transformed coordinates.  It is also necessary to include the Jacobian for the steps not made explicit.

We will next consider a full 2-D space of symmstry transformations, this time we use coordinates $\{{z=xy,R=(|\sigma_-/\sigma_+|^2),C,D=|\sigma_-|^{-1}}\}$.  Of these latter coordinates, $R$ and $D$ are fully specified by the observed constraints on \eqref{eq:sigmapm}, but $z$ and $C$ are free within $-1\leq C\leq1$ and $0\leq z\leq T-\sqrt{T^2-1}$ with $T=(1+C)/R-C$. The transformation to this coordinates is
\begin{align}
  z&=xy\\
  R&=\frac{x^2+y^2+2xyC}{1+x^2y^2+2xyC}\\
  C&=C\\
  D&=d\frac{(1+x^{1/2})^2(1+y^{1/2})^2}{1+x^2y^2+2xyC}.
\end{align}
The symmetry is realized by any translation in $z\rightarrow z$, the new coordinates are ten recovered from the inverse transformation is 
\begin{align}
  x&=\sqrt{\Lambda(z,\Delta)-\delta(z,\Delta)}\\
  y&=\sqrt{\Lambda(z,\Delta)+\delta(z,\Delta)}\\
  C&=C\\
  d&=D\frac {1+z^2+2zC}{(1+x(z,\Delta)^{1/2})(1+y(z,\Delta)^{1/2})},
\end{align}
with $\Lambda=(1+z^2-\Delta^2)/2$ and $\delta=\sqrt{\Lambda^2-z^2}$.
The rest of the process follows as above.

An issue with this specification (as with the 1-D version) is that it only encompases transformations which preserve whether $t_\theta$ or $t_\iota$ is larger.  It appears that this can be remedied by starting with a parameter $q=(t_\theta/t_\iota)^{1/4}$ which unlike $z$ is unconstrained. The calculation is similar, using $z=T-\sqrt{T^2-1}$ with $T=((q+q^{-1})/2+C)/R-C$ as intermediates.

\end{document}

### 2019-11-27 docs/paper_LDC/paper_LDC.tex
LDC notes.

@@
 \newcommand\psiL{{\psi_{L}}}
+\newcommand\C{{\cos{4\psi}}}
 
@@
 
-First, for an approximate approach, note that in the limit that either $t^4_\theta<<1$ or $t^4_\iota<<1$, then \eqref{eq:sigmapm} become nearly independent of that quantity except through  \eqref{eq:sigmafactorrho}.
-Then we have an approximate symmetry
+First, for an approximate approach, note that in the limit that either $t^4_\theta<<1$ or $t^4_\iota<<1$, then \eqref{eq:sigmapm} become nearly independent of that quantity except through  $\rho$ in \eqref{eq:sigmafactorrho}.
+Similarly if $t^4_\iota>>1$ then dependence is only through $t^4_\iota\rho$.
+We can approximately represent both possibilities allowing transformations which preserve $d(1+t^2_\iota)^2/(1+t^4_\iota)=2d/(\cos^2(\iota)+1)$,
 \begin{align}
   \iota&\rightarrow\iota'\\
-  d&\rightarrow d\frac{(1+t^2_{\iota})^2}{(1+t'^2_{\iota})^2}
+  d&\rightarrow d\frac{(1+\cos^2{\iota'})}{(1+\cos^2{\iota})}&=d.
 \end{align}
-A similar symmetry exists when $t^4_\theta<<1$.
+Another option at the same level of approximation is preserve $d/\cos(\iota)$, but this may be less appropriately when $\cos(\iota)->0$.
+In some MCMC tests, the latter form acheived better autocorrelations, although the former had slightly higher acceptance rates.
+A corresponding symmetry exists when $t^4_\theta<<1$.
 
@@
 \end{align}
-with $\Lambda=(1+z^2-\Delta^2)/2$ and $\delta=\sqrt{\Lambda^2-z^2}$.
+with $\Lambda=(1+z^2-\Delta^2)/2$ $\Delta=(1-R)*(1+z^2+2*C)$ and $\delta=\sqrt{\Lambda^2-z^2}$.
 The rest of the process follows as above.
@@
 An issue with this specification (as with the 1-D version) is that it only encompases transformations which preserve whether $t_\theta$ or $t_\iota$ is larger.
-It appears that this can be remedied by starting with a parameter $q=(t_\theta/t_\iota)^{1/4}$ which unlike $z$ is unconstrained.
+It appears that this can be remedied by starting with a parameter $q=(t_\theta/t_\iota)^{4}$ which unlike $z$ is unconstrained.
 The calculation is similar, using $z=T-\sqrt{T^2-1}$ with $T=((q+q^{-1})/2+C)/R-C$ as intermediates.
 
+Since we are actually implementing this, we spell it out explicitly.  We use the following notation for convenience:
+\begin{align}
+  x&=t^4_{\theta}=\tan^4{\theta/2}\\
+  y&=t^4_{\iota}=\tan^4{\iota/2}\\
+  z&=xy\\
+  \hat w&=x/y\\
+  R&=|\sigma_-/\sigma_+|^2\\
+  &=\frac{x^2+y^2+2xy\C}{1+x^2y^2+2xy\C}\\
+  &=\frac{\hat w+\hat w^{-1}+2\C}{z+z^{-1}+2\C}\\
+  \Delta&=|\sigma_+|^2-|\sigma_-|^2\\
+  &=\frac{(1-x^2)(1-y^2)}{16d^2(1+x^{1/2})^2(1+y^{1/2})^2}\\
+  &=\frac{(1+z^2-z(\hat w+{\hat w}^{-1})}{16d^2(1+({\hat w}z)^{1/4})^2(1+({\hat w}/z)^{1/4})^2}.
+\end{align}
+Using this shorthand, we transform from variables $\{\psi,x,y,d\}$ to $\{\psi,w,R,D\}$.
+Of these latter coordinates, $R$ and $D$ are fully specified by the observed constraints on \eqref{eq:sigmapm},
+but $\psi$ and $w$ are free. The transformation to these coordinates is
+\begin{align}
+  \psi&=\psi\\
+  w&=\mathrm{sign}(\ln(xy))x/y\\
+  R&=\frac{x^2+y^2+2xy\C}{1+x^2y^2+2xy\C}\\
+  \Delta&=\frac{(1-x^2)(1-y^2)}{16d^2(1+x^{1/2})^2(1+y^{1/2})^2}.
+\end{align}
+Note that we have added a sign to ${\hat w}=|w|$ as a convenient trick to
+store the information about whether $xy$ is greater than $1$ to provide a bijective map over the full domain. 
+The symmetry is realized by any translation in $\{\psi,w\}\rightarrow\{\psi',w'\}$ preserving $R$ and $D$.
+The new coordinates are ten recovered from the inverse transformation, 
+\begin{align}
+  \psi&=\psi\\
+  x&=z\hat w\\
+  y&=z/{\hat w}\\
+  d&=\Delta^{-1/2}\frac {\sqrt{1+z^2-z(\hat w+{\hat w}^{-1})}}{(1+(z \hat w)^{1/4})(1+(z/{\hat w})^{1/4})},
+\end{align}
+with $z=T+\mathrm{sign}(w)\sqrt{T^2-1}$ and $T=(({\hat w}+{\hat w}^{-1})/(2R)+\C(1-R)/R$ as intermediates.
+The Jacobian of the forward coordinate transformation its contribution to the Jacobian of the resulting
+symmetry transformation are conveniently written in mixed notation,
+\begin{align}
+  |J|&=-2\frac{wR\Delta}{xyd}\frac{1-z^2}{1+z^2+2z\C},\\
+  \frac{|J'|}{|J|}&=\frac{w'xyd}{wx'y'd'}\frac{1-z'^2}{1-z^2}\frac{1+z^2+2z\C}{1+z'^2+2z'\cos{4\psi'}}.
+\end{align}
+
+The rest of the calculation proceeds to adjest the angles. Writing $\Psi_{\pm}=\mathrm{Arg}(x^{\pm1}e^{-2i\psi}+ye^{2i\psi})$,
+we have
+\begin{align}
+  \mathrm{Arg}(\frac{\sigma_-}{\sigma_+})&=4\lambda+\Psi_--\Psi_+\\
+  \mathrm{Arg}({\sigma_-}{\sigma_+})&=4\phi+\Psi_-+\Psi_+.
+\end{align}
+Each of these must be preserved through the symmetry transformation, which can be enforced by
+\begin{align}
+  \lambda'&=\lambda+\frac14(\Psi_--\Psi'_--\Psi_++\Psi'_+)\\
+  \phi'&=\phi+\frac14(\Psi_--\Psi'_-+\Psi_+-\Psi'_+).
+\end{align}
+
 \end{document}

### 2020-10-13 docs/paper_LDC/paper_LDC.tex
These were uncommittted changes from last December (LDC process).  Most are the LDC paper start.  There are also some comments added in data_likelihood.c and a flag for extra checking there turned off.

@@
 \newcommand\psiL{{\psi_{L}}}
-\newcommand\C{{\cos{4\psi}}}
+\newcommand\C{{\cos(4\psi)}}
 
@@
   \iota&\rightarrow\iota'\\
-  d&\rightarrow d\frac{(1+\cos^2{\iota'})}{(1+\cos^2{\iota})}&=d.
+  d&\rightarrow d'\frac{(1+\cos^2{\iota'})}{(1+\cos^2{\iota})}.
 \end{align}
@@
 
-It is also possible to explicitly formulate this symmetry in the general case, though here we are still working on a concrete plan.
-To approach this, we represent subspace of the parameter space, with coordinates $\{x=\min\{t^4_\theta,t^4_\iota\},y=\max\{t^4_\theta,t^4_\iota\},C=cos(2\psi),d\}$ trivially related to our basic coordinates, but better adapted for this problem.
+It is also possible to explicitly formulate this symmetry in the general case.
+To approach this, we represent subspace of the parameter space, with coordinates $\{x=\min\{t^4_\theta,t^4_\iota\},y=\max\{t^4_\theta,t^4_\iota\},C=\cos(4\psi),d\}$ trivially related to our basic coordinates, but better adapted for this problem.
 
-We will next consider an exact generalization of the distance-inclination symmetry.
+Consider an exact generalization of the distance-inclination symmetry.
 For this we use coordinates $\{z=xy,\Delta=(|\sigma_-|^2-|\sigma_+|^2)/\rho^2,S=|\sigma_-|^2/\rho^2,\rho\}$.
@@
   S&=1+x^2y^2+2xyC\\
-  D&=d(1+x^{1/2})(1+y^{1/2}).
+  D&=d(1+x^{1/2})^2(1+y^{1/2})^2.
 \end{align}
@@
   C&=\frac{S-1-z^2}{2z}\\
-  d&=\frac D{(1+x(z,\Delta)^{1/2})(1+y(z,\Delta)^{1/2})},
+  d&=\frac D{(1+x(z,\Delta)^{1/2})^2(1+y(z,\Delta)^{1/2})^2},
 \end{align}
@@
 
-We will next consider a full 2-D space of symmstry transformations, this time we use coordinates $\{{z=xy,R=(|\sigma_-/\sigma_+|^2),C,D=|\sigma_-|^{-1}}\}$.
-Of these latter coordinates, $R$ and $D$ are fully specified by the observed constraints on \eqref{eq:sigmapm}, but $z$ and $C$ are free within $-1\leq C\leq1$ and $0\leq z\leq T-\sqrt{T^2-1}$ with $T=(1+C)/R-C$.
-The transformation to this coordinates is
-\begin{align}
-  z&=xy\\
-  R&=\frac{x^2+y^2+2xyC}{1+x^2y^2+2xyC}\\
-  C&=C\\
-  D&=d\frac{(1+x^{1/2})^2(1+y^{1/2})^2}{1+x^2y^2+2xyC}.
-\end{align}
-The symmetry is realized by any translation in $z\rightarrow z$, the new coordinates are ten recovered from the inverse transformation is 
-\begin{align}
-  x&=\sqrt{\Lambda(z,\Delta)-\delta(z,\Delta)}\\
-  y&=\sqrt{\Lambda(z,\Delta)+\delta(z,\Delta)}\\
-  C&=C\\
-  d&=D\frac {1+z^2+2zC}{(1+x(z,\Delta)^{1/2})(1+y(z,\Delta)^{1/2})},
-\end{align}
-with $\Lambda=(1+z^2-\Delta^2)/2$ $\Delta=(1-R)*(1+z^2+2*C)$ and $\delta=\sqrt{\Lambda^2-z^2}$.
-The rest of the process follows as above.
-
-An issue with this specification (as with the 1-D version) is that it only encompases transformations which preserve whether $t_\theta$ or $t_\iota$ is larger.
-It appears that this can be remedied by starting with a parameter $q=(t_\theta/t_\iota)^{4}$ which unlike $z$ is unconstrained.
-The calculation is similar, using $z=T-\sqrt{T^2-1}$ with $T=((q+q^{-1})/2+C)/R-C$ as intermediates.
-
-Since we are actually implementing this, we spell it out explicitly.  We use the following notation for convenience:
+Now consider a full 2-D space of symmstry transformations,
+We use the following notation for convenience:
 \begin{align}
@@
   y&=t^4_{\iota}=\tan^4{\iota/2}\\
-  z&=xy\\
-  \hat w&=x/y\\
+  z&=xy^s\\
+  \hat w&=x/y^s\\
   R&=|\sigma_-/\sigma_+|^2\\
   &=\frac{x^2+y^2+2xy\C}{1+x^2y^2+2xy\C}\\
-  &=\frac{\hat w+\hat w^{-1}+2\C}{z+z^{-1}+2\C}\\
-  \Delta&=|\sigma_+|^2-|\sigma_-|^2\\
-  &=\frac{(1-x^2)(1-y^2)}{16d^2(1+x^{1/2})^2(1+y^{1/2})^2}\\
-  &=\frac{(1+z^2-z(\hat w+{\hat w}^{-1})}{16d^2(1+({\hat w}z)^{1/4})^2(1+({\hat w}/z)^{1/4})^2}.
+  &=\left(\frac{\hat w+\hat w^{-1}+2\C}{z+z^{-1}+2\C}\right)^s\\
+  \Delta&=16(|\sigma_-|^2-|\sigma_+|^2)\\
+  &=\frac{(1-x^2)(1-y^2)}{d^2(1+x^{1/2})^4(1+y^{1/2})^4}\\
+  &=\frac{1-x^2-y^2+x^2y^2}{d^2(1+x^{1/2})^4(1+y^{1/2})^4}\\
+  &=\frac{(xy)^{-1}-(x/y)-(x/y)^{-1}+xy}{d^2(x^{-1/4}+x^{1/4})^4(y^{-1/4}+y^{1/4})^4}\\
+  &=\frac{s}{d^2}\frac{z+z^{-1}-(\hat w+{\hat w}^{-1})}{(z^{1/2}+z^{-1/2}+{\hat w}^{1/2}+{\hat w}^{-1/2})^4}
 \end{align}
+where $s=\pm1$.
 Using this shorthand, we transform from variables $\{\psi,x,y,d\}$ to $\{\psi,w,R,D\}$.
@@
   \psi&=\psi\\
-  w&=\mathrm{sign}(\ln(xy))x/y\\
+  w&=\mathrm{sign}(\ln(xy))x/y^s\\
   R&=\frac{x^2+y^2+2xy\C}{1+x^2y^2+2xy\C}\\
-  \Delta&=\frac{(1-x^2)(1-y^2)}{16d^2(1+x^{1/2})^2(1+y^{1/2})^2}.
+  \Delta&=\frac{(1-x^2)(1-y^2)}{d^2(1+x^{1/2})^4(1+y^{1/2})^4}.
 \end{align}
@@
 store the information about whether $xy$ is greater than $1$ to provide a bijective map over the full domain. 
+Also, swapping the sign $s$, which we have not yet specified, reverses the roles of $w$ and the intermediate quantitity $z$.
 The symmetry is realized by any translation in $\{\psi,w\}\rightarrow\{\psi',w'\}$ preserving $R$ and $D$.
@@
   \psi&=\psi\\
-  x&=z\hat w\\
-  y&=z/{\hat w}\\
-  d&=\Delta^{-1/2}\frac {\sqrt{1+z^2-z(\hat w+{\hat w}^{-1})}}{(1+(z \hat w)^{1/4})(1+(z/{\hat w})^{1/4})},
+  x&=\sqrt{z\hat w}\\
+  y&=\sqrt{z/{\hat w}}\\
+  d&={\Delta^{-1/2}}\frac {\sqrt{1+z^2-z(\hat w+{\hat w}^{-1})}}{(z^{1/2}+z^{-1/2}+{\hat w}^{1/2}+{\hat w}^{-1/2})^2}
 \end{align}
-with $z=T+\mathrm{sign}(w)\sqrt{T^2-1}$ and $T=(({\hat w}+{\hat w}^{-1})/(2R)+\C(1-R)/R$ as intermediates.
+with $z=T+\mathrm{sign}(w)\sqrt{T^2-1}$ and $T=(({\hat w}+{\hat w}^{-1})/(2R^s)+\C(1-R^s)/R^s$ as intermediates.
+Choosing $s=\mathrm{sign}(\ln(R))$, ensures that $0\leq R^s\leq1$, so that $z$ remains real. 
 The Jacobian of the forward coordinate transformation its contribution to the Jacobian of the resulting
@@
 \begin{align}
-  |J|&=-2\frac{wR\Delta}{xyd}\frac{1-z^2}{1+z^2+2z\C},\\
-  \frac{|J'|}{|J|}&=\frac{w'xyd}{wx'y'd'}\frac{1-z'^2}{1-z^2}\frac{1+z^2+2z\C}{1+z'^2+2z'\cos{4\psi'}}.
+  |J|&=4\frac{wR\Delta}{xyd}\frac{z-z^{-1}}{z+z^{-1}+2\C},\\
+  \frac{|J|}{|J'|}&=\frac{wx'y'd'}{w'xyd}\frac{z-z^{-1}}{z'-z'^{-1}}\frac{z'+z'^{-1}+2\cos(4\psi')}{z+z^{-1}+2\cos(4\psi)}.
 \end{align}

### 2020-10-13 tools/data_likelihood.c

@@ -45,7 +45,7 @@
 int IntProdStyle=1;
 int NratioCut = 4;
 int half_edges = 0;
-int in_development=1;
+int in_development=0;
 int use_logarithmic=0;
 
 //Data whitening
@@ -101,7 +101,7 @@ void WhitenData(
 // Functions for computing the integral 
 
 //Perform explicit spline evaluation for inner-product computation of
-//freqseries D with CAmpPhase spline set.  The grid of the D freqseries is summed for a simple integral approximant
+//freqseries D with CAmpPhase spline set.  The grid of the D freqseries is summed for a simple integral approximant.
 double ReIntProdCAmpPhaseSpline(
   CAmpPhaseSpline* splines,                    //input "model"
   ReImUniformFrequencySeries* Dfreqseries,  //input defines ReIm uniform frequency series to be multiplied with spline evals and summed
@@ -153,7 +153,9 @@ double ReIntProdCAmpPhaseSpline(
 };
 
 //Perform explicit spline evaluation for inner-product computation of
-//freqseries D with CAmpPhase spline set.  The grid of the D freqseries is summed for a simple integral approximant
+//freqseries D with CAmpPhase spline set.  
+//Differences of C from plain version (Doesn't include "B" features):
+//  -Switched integration range specification from index-based to freq-based
 double ReIntProdCAmpPhaseSplineC(
   CAmpPhaseSpline* splines,                    //input "model"
   ReImUniformFrequencySeries* Dfreqseries,  //input defines ReIm uniform frequency series to be multiplied with spline evals and summed
@@ -226,7 +228,10 @@ double ReIntProdCAmpPhaseSplineC(
 
 
 //Perform explicit spline evaluation for inner-product computation of
-//freqseries D with CAmpPhase spline set.  The grid of the D freqseries is summed for a simple integral approximant
+//freqseries D with CAmpPhase spline set.
+//Differences in B vs A:
+//  -Added fast computation of exp(i*phase) fac (when sufficient samples in spline interval)
+//  -Also includes diagnostic for estimating the effective time-width of FD spline intervals
 double ReIntProdCAmpPhaseSplineB(
   CAmpPhaseSpline* splines,                    //input "model"
   ReImUniformFrequencySeries* Dfreqseries,  //input defines ReIm uniform frequency series to be multiplied with spline evals and summed
@@ -322,7 +327,10 @@ double ReIntProdCAmpPhaseSplineB(
 };
 
 //Perform explicit spline evaluation for inner-product computation of
[…]

### 2021-06-03 docs/paper_LDC/paper_LDC.tex
Committing stale changes from Oct 2020.
More work.

@@
-\documentclass[aps,showpacs,twocolumn,prd,superscriptaddress,nofootinbib]{revtex4}
+\documentclass[aps,showpacs,12pt,onecolumn,prd,superscriptaddress,nofootinbib]{revtex4}
 
@@
 \newcommand\C{{\cos(4\psi)}}
+\newcommand\mc{{\mathcal{M}}}
+
+\newcommand{\jgb}[1]{{\color{DarkGreen} #1}}
 
@@
 \label{sec:intro}
+This report provides methods and results in a successful effort to characterize a massive black hole binary (MBHB) merger as part fo the first round ``Radler' LISA Data Challenge (LDC) campaign.
+The challenge included simulated data for a LISA observation of a single MBHB system with parameters drawn from a simulated population.
+We conducted a Bayesian treatment of the problem with full-range naive priors for all parameters (except for moderate restriction on the coalescence time.
+This note provides details on the methodology applied to infer MBHB system parameters and more detailed results beyond those included in the broader LDC Radler analysis.
 
-[Introduction]
+\section{Methodology}
+
+The simulated MBHB observation in the LDC1.1 dataset included a signal for a non-precessing MBHB system which merged during the observation period using the PhenomD waveform model.
+The version of LISA assumed in the simulation was ***.
+The LISA response for the frequency domain dataset was implemented using the leading-order approximation of the approach in Ref.~\cite{MarsatEA2018} in a python implementation.
+The dataset corresponds to about 42 million time samples at 10s cadence, adding to the challenge of computing likelihoods fast enough to support a full Bayesian search.
+Sec.~\ref{sec:definition} specifies the problem definition including the parmeterization and generally naive assumed priors.
+
+To enable a suitably fast likelihood computation required several development through the Radler campaign.
+First we developed fast, but python-accessible version of our previous strictly C-based code for the LISA response signal processing defining the core of the \texttt{lisabeta} package, which has also already been applied in several LISA parameter estimation studies.
+Those other studies provided fast likelihood computation by taking we also required a fast-but-accurate method for computing likelihood, but for the LDC we needed a fast way to compute likelihoods with noisy data.
+We discuss out straightforward approach to this in Sec.~\ref{sec:likelihood}.
+
+We applied MCMC sampling to characterize the posterior using the code \texttt{ptmcmc} we had previously developed.
+That code generally takes a relatively naive brute-force approach to MCMC sampling applying parallel tempering to promote full model space exploration an generally relying on a general-purpose proposal distribution components.
+For this problem, we added an interface to enable applying ore problem-specific proposal elements.
+And applied a set of proposals based on approximate symmetries identified in Ref.~\cite{Marsat2020} which can lead to specific parameter degeneracies.
+The details of that treatment are described in ~\ref{sec:MCMC}.
+
+\subsection{ Parameters and Priors }
+\label{sed:definition}
+(So far this is verbatim from our submission.)
+
+%We have attempted to name the parameters consistently with the LDC conventions, using inclination and polarization angles which should be consistent with LDC manual eqs 16/17.
+In the samples, CoalescenceTime is measured at the SSB.
+These are not quite the same as the parameters used in the sampling (see Prior).
+In some cases, for efficiency, we have marginalized over the phi (PhaseAtCoalescence) parameter in the likelihood calculation and though this parameter is present, in should meaninglysly recover the uniform prior.
+
+We generally applied broad uninformative priors across the parameter space. Table~\ref{tab:prior} shows the full list of parameters and their prior distributions.
+
+\begin{table}
+  \label{tab:prior}
+  \centering
+  \begin{tabular}{llll}
+    Name & Symbol & Units & Prior \\
+    \hline
+    chirp mass & $\mc$ & $\Msol$ & $\mathrm{Log}[14650.8,1.46508e+08]$\\
+    mass ratio & $q$ & - &$\mathrm{Log}[0.0333333,1]$\\
+    primary spin & $a1$ & - & $[-0.9,1]$\\
+    secondary spin & $a2$ & - & $[-0.9,1]$\\
+    coalescence time (LISA-frame) & $t_{cL}$ & s & $[2.396e+07,2.596e+07]$\\
+    distance & $d$ & Mpc & $\mathrm{Log}[186.687,1.68018e+07]$\\
+    inclination & $\iota$ & rad & $\mathrm{Polar}[0,3.14159]$\\
+    orbital phase & $\varphi$ & rad & $[0,6.28319]$\\
+    longitude (LISA-frame) & $\lambda_L$ & rad & $[0,6.28319]$\\
+    latitude (LISA-frame) &$\beta_L$ & rad & $\mathrm{CoPolar}[-1.5708,1.5708]$\\
+    polarization angle (LISA-frame) & $\psi_L$ & rad & $[0,3.14159]$\\
+    \hline
+  \end{tabular}
+  \caption{ MBHB parameters and priors.
+Here $\mathrm{Log}[]$ indicates a log-uniform distribution over the range, while $\mathrm{Polar}[]$ ($\mathrm{CoPolar}[]$) correspond to polar (co-polar) angle distributions consistent a uniform distribution on a sphere.}
+\end{table}
+
+Among the intrisic parameters, rather than in terms of individual component masses, we define our problem in terms of chirp mass $\mc$ and mass ratio $q$ applying broad-range log-uniform priors on each parameter, spanning a factor of $10^4$ in $\mc$ and allowing $q \sim [1/30,1]$.
+We applied uniform priors on the spin parameters across almost the full physical range, but excluding relatively extreme anti-aligned spins, for which the waveform model sometimes failed.
+
+The extrinsic parameters which relate to the arrival of the signal at LISA, we use specialized parameters, where the time, sky-location and polarization angle are set with reference to the orientation of LISA at merger time.
+We converted these to the Solar-System Barycentric(SSB) referenced parameters for our LDC entry.
+We assumed full-range circle or sphere-uniform priors for all angle parameters.
+The only parameter for which we did not assume more-or-less the full plausible range was the merger time $t_c$.
+Our method for accelerating the likelihood so far is inaccurate outside of a rough window.
+We use a uniform time window with a width of 2 million seconds.
+
+We also note that we have not attempted to relate our orbital phase $\varphi$ reference with that of the LDC.
+
+\subsection{Likelihood computation}
+\label{sec:likelihood}
+
+\subsection{Bayesian sampling}
+\label{sec:MCMC}
+
+\section{Results}
 
@@
   y&=\sqrt{z/{\hat w}}\\
-  d&={\Delta^{-1/2}}\frac {\sqrt{1+z^2-z(\hat w+{\hat w}^{-1})}}{(z^{1/2}+z^{-1/2}+{\hat w}^{1/2}+{\hat w}^{-1/2})^2}
+  d&={\Delta^{-1/2}}\frac {\sqrt{1+z^2-z(\hat w+{\hat w}^{-1})}}{(z^{1/2}+z^{-1/2}+{\hat w}^{1/2}+{\hat w}^{-1/2})^2}\\
+  d&={\Delta^{-1/2}}\frac {\sqrt{z+z^{-1}-(\hat w+{\hat w}^{-1})}}{(z^{1/2}+z^{-1/2}+{\hat w}^{1/2}+{\hat w}^{-1/2})^2} \mathrm{(Which is right?)}
 \end{align}
